### 2025-05-02 Latex/BIP_predictor.tex
up

% -*- mode: noweb; noweb-default-code-mode: R-mode; -*-
\documentclass[a4paper]{article}

\title{BIP-Predictor}
\author{Marc Wildi\\
Zurich University of Applied Sciences (ZHAW)\\
8000 Zurich, Switzerland\\
marc.wildi@zhaw.ch}

\usepackage{a4wide}
\usepackage{amsmath}
\usepackage{amsfonts} 
\usepackage{amsthm}
\usepackage{ amssymb }
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{float}
\newtheorem{Proposition}{Proposition}
\newtheorem{Corollary}{Corollary}
\newtheorem{Theorem}{Theorem}
\DeclareMathOperator{\sign}{sign}
%- Makes the section title start with Appendix in the appendix environment
\newcommand{\Appendix}
{%\appendix
\def\thesection{Appendix~\Alph{section}}
%\def\thesubsection{\Alph{section}.\arabic{subsection}}
\def\thesubsection{A.\arabic{subsection}}
}

\usepackage{Sweave}
\begin{document}

\maketitle

\begin{abstract}
\noindent

\end{abstract}

~\\
~\\

%\tableofcontents

\section{Introduction}

\section{Data and Dependence}

\subsection{Data}

\begin{itemize}
\item Choice of indicators: BIP, ip, ESI, ifo, spread
\item Quarterly frequency (emphasize $h\geq 2$)
\item Data transformations: Diff-log, standardization (merely for ease of visual inspection), trim Covid ourliers
\item Revisions and publication lags:  BIP lagged by quarter and ip lagged by two months in Jan-2025 data set
\item Real-time data: all series are artificially aligned at end point. BIP and ip are advanced by their lags.
\end{itemize}

The transformed quarterly series are displayed in Fig. \ref{data}, while Fig. \ref{data_lags} offers a close-up view of the financial crisis and the pandemic. The latter figure illustrates that the real-time GDP and industrial production series are synchronized at dips and peaks (coincident), whereas the other indicators tend to lead in relative terms. This characteristic of the additional explanatory variables can be leveraged in a multivariate approach to enhance forecasting performance compared to univariate benchmarks.

\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/Data.pdf}\caption{Sample cross-correlation function (CCF): full data set.\label{data}}\end{center}\end{figure}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/data_lags.pdf}\caption{Data: leads and lags during financial crisis (left) and Pandemic (right).\label{data_lags}}\end{center}\end{figure}

\subsection{Dependence}

We analyze the dependence structure of the data and derive a simple VAR model to establish the multivariate filter. The sample cross-correlation function (CCF) is shown in Figs. \ref{CCF} (entire dataset) and \ref{CCF_wc} (data prior to the pandemic).  
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/CCF.pdf}\caption{Sample cross-correlation function (CCF): full data set.\label{CCF}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/CCF_wc.pdf}\caption{Sample cross-correlation function (CCF): without Pandemic.\label{CCF_wc}}\end{center}\end{figure}The CCF confirms the leads and lags illustrated in Fig. \ref{data_lags}, although the strength of the dependence is influenced by the singular readings during the pandemic\footnote{We discard the corresponding data when finalizing the predictor.}. We can now fit a VARMA model, and according to standard diagnostic tests (not shown here), a simple VAR(1) specification aligns with the data. The resulting impulse responses of the model are presented in Figs. \ref{impulse} (full data set) and \eqref{impulse_wc} (prior the pandemic).
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/ma_inv_multi_ip.pdf}\caption{MA inversion of VAR(1): full data set.\label{impulse}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/ma_inv_multi_ip_wc.pdf}\caption{MA inversion of VAR(1): without Pandemic.\label{impulse_wc}}\end{center}\end{figure}The model summarizes key features of the data, where lagging series (GDP and industrial production, when accounting for publication lags) depend on  indicators leading in relative terms. In particular, the impulse response of the spread (rightmost panel in the figure) suggests the relevance of a univariate model for this specific leading indicator.

\section{Direct Forecasts}\label{direct_forecast}

\subsection{`Classic' Direct Forecast}\label{cdf}

So-called direct forecasts are obtained by regressing the relevant indicators on forward-shifted BIP, accounting for the additional publication lag. For illustration of the concept, we here rely on all indicators and full sample information, excluding the pandemic. In-sample forecasts and the shifted BIP, which serves as the target, are presented in Figs. \ref{direct_wc} (full dataset without the pandemic) and \ref{direct_wc_financial_crisis} (financial crisis) for horizons up to three quarters ahead. 
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/direct_wc_all.pdf}\caption{Direct forecasts (red lines) based on all indicators and full sample information, without the pandemic: BIP (black line) shifted by zero one, two and three quarters.\label{direct_wc}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/direct_wc_financial_crisis.pdf}\caption{Direct forecast  across the financial crisis.\label{direct_wc_financial_crisis}}\end{center}\end{figure}The results indicate that as $h$ increases, the peaks and dips of the direct forecasts become increasingly right-shifted (delayed) relative to the target. Additionally, the F-statistics precipitously decline from $F=3.57$, for $h=2$, to $F=0.46$, for $h=3$, see Table \eqref{f_stat}, suggesting that the predictor becomes statistically insignificant for larger forecast horizons exceeding two quarters (link to Heinisch and Scheufele).

\subsection{Filter: HP(160)}

As shown, the performance of direct forecasts sharply declines for larger forecast horizons. We conjecture that erratic short-term fluctuations -—unpredictable high-frequency noise -— overlay the data, thereby obscuring the effective 'signal' and making a direct regression more susceptible to overfitting. We therefore envisage to highlight the signal by applying a filter to dampen high-frequency noise. 
The so-called Hodrick-Prescott filter is a classic tool used in business cycle analysis: the filter is specified by a single smoothing parameter $\lambda$ and the value $\lambda=1600$ is specifically recommended for quarterly data. However, Phillips and Jin (2021) suggest that the HP(1600) filter removes relevant information due to excessive smoothing.  In our context, this oversmoothing issue would be further exacerbated when considering forecast horizons shorter than a year -—an interval inconsistent with the mean duration of up to several years of business cycles, as highlighted by the HP(1600). Consequently, we here select a more adaptive HP(160) 'target' filter. 
While this choice may appear somewhat arbitrary, an analysis of the filter in the frequency domain confirms a more suitable profile for the corresponding amplitude function, assigning greater weight to yearly components -—relevant in our prediction framework-— than the classic quarterly HP filter. This approach also addresses and controls for undesirable high-frequency noise. See Fig. \ref{hp_160}, which presents both two-sided and (classic) one-sided filters (top panels) along with the associated amplitude functions (bottom panels). We now posit the HP(160) filter as a means to achieve better forecasting performance at larger horizons, noting that a comprehensive technical analysis of the effect of $\lambda$ on the resulting BIP predictor can be conducted using the M-SSA package. 
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/hp_160.pdf}\caption{Two-sided HP-160) (top left) and one-sided (concurrent) HP-C filters (top right) with corresponding amplitude functions (bottom).\label{hp_160}}\end{center}\end{figure}
\subsection{Direct HP Forecast}\label{hpdf}

Extending the `classic' direct forecast design, we here consider filtered indicators, based on the one-sided HP(160), denoted as HP-C: the filtered indicators are explanatory variables for the regressions on forward-shifted BIP\footnote{Katja and Simon: for each forecast horizon $h=0,1,2,3$ I'm using corresponding forecasts of the real-time filter output, assuming the data to be white noise (assuming an ARMA model does not lead to significantly different results). I think we might ignore these technical details in the paper.}. As in the previous section, we rely on all indicators as well as the full sample estimate for illustration, see Figs.\eqref{direct_hp_forecasts} (all observations with exclusion of the pandemic) and \eqref{direct_hp_forecasts_financial_crisis} (financial crisis). A comparison with the classic direct forecasts in Fig.\eqref{direct_wc_financial_crisis} indicates that the new predictors are slightly less retarded (`faster'). 
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/direct_hp_forecasts.pdf}\caption{Direct HP forecasts: entire data set with exclusion of the pandemic.\label{direct_hp_forecasts}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/direct_hp_forecasts_financial_crisis.pdf}\caption{Direct HP forecasts over the curse of the financial crisis.\label{direct_hp_forecasts_financial_crisis}}\end{center}\end{figure}Moreover, the F-statistics are larger than for the direct forecasts, see Table \eqref{f_stat}, but the predictor remains statistically insignificant at forecast horizons exceeding two quarters (even though the evaluation is conducted in-sample and based on all available indicators). 
% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Thu May  1 21:38:41 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & h= 0 & h= 1 & h= 2 & h= 3 \\ 
  \hline
Direct forecast & 7.188 & 5.285 & 3.574 & 0.460 \\ 
  Direct HP forecast & 21.886 & 8.415 & 4.810 & 1.834 \\ 
   \hline
\end{tabular}
\caption{F-statistics of direct forecasts and direct HP forecasts for forecast horizons $h=0,1,2,3$: fully in-sample.  } 
\label{f_stat}
\end{table}We now propose a more refined causal filter design aiming for statistical significance of the resulting predictor at longer forecast horizons, out-of-sample.

\section{Multivariate Causal Filter: M-SSA}

The application of the one-sided HP-C filter to the indicators has demonstrated improvements in forecast performance. However, the univariate nature of this filter limits its ability to leverage information from the cross-section, as supported by leading indicators. Additionally, the noise suppression capability of the one-sided filter is compromised, as illustrated by high-frequency leakage of the amplitude function in Fig. \eqref{hp_160} and further elaborated upon by Wildi (2025). To address these limitations, we propose an extension of the Smooth Sign Accuracy (SSA) framework introduced by Wildi (2025).

\subsection{Optimization criterion}

Let $\mathbf{X}_t$ (of dimension $t\times n$) denote a set of $n$ explanatory series $\mathbf{x}_{1t},...,\mathbf{x}_{nt}$, with observations $\mathbf{x}_{it}=(x_{i1},...,x_{it})'$. Let $\mathbf{z}_t=(z_{1},...,z_t)$  denote a target series, which typically lies outside the linear space spanned by $\mathbf{X}_t$  (and is unknown at time $t$). For simplicity, we assume stationarity of all time series involved. In this framework, $\mathbf{X}_t$ corresponds to a matrix of selected economic indicators with $n=5$, and $z_{it}$   is the output of a two-sided HP(160) filter applied to the $i$-th indicator, $i=1,...,5$. Since $z_{it}$  depends on future observations $x_{it+1},x_{it+2},...$, the prediction task involves `tracking' $z_{it}$  using the estimate $y^i_{t}=\sum_{j=1}^ny^i_{jt}$, where $y^i_{jt}=\mathbf{b}^i_j~'\mathbf{x}_{jt}=\sum_{k=0}^{L} b^i_{jk}x_{jk}$ are the outputs of a multivariate (causal or one-sided) filter $\mathbf{B}^i$, with columns $\mathbf{b}^i_{j}$ of fixed length $L$, $j=1,...,5$, assigning weight to the last $L$ observations of the $j$-th indicator $\mathbf{x}_{jt}$ (the fixed-length assumption is merely used for ease of exposition).
For the sake of clarity, we now omit the upper-case index  $i$ for all variables, assuming $z_t=z^{i_0}_{t}$ for some fixed $i_0$.% Additionally, we assume a fixed filter length $L$, such that $\mathbf{B}^i_t=\mathbf{B}_t=\mathbf{B}$, which has dimension $L\times 5$. 
The task of `tracking'  a target can be formalized in different ways; here, we focus on the target correlation $\rho(z,y,h)$ between $y_t$ and $z_{t+h}$, where $h\geq 0$ denotes the forecast horizon (backcasting with $h<0$ is ignored in this context). To streamline the notation and avoid unnecessary complexity, we assume a fixed horizon, say $h=h_0$, so that we may drop the reference to $h$ in our notation. In the univariate case ($n=1$), Wildi (2025) proposes the Simple Sign Accuracy (SSA) as an optimization criterion for this task:
\begin{eqnarray}\label{critssa}
\left.\begin{array}{c}\rho(z,y)\to\max\\
\rho(y,1)=\rho_1\end{array}\right\}.
\end{eqnarray}
Objective function and constraint are expressed in terms of correlations, namely the target correlation $\rho(z,y)$ between target and predictor and the lag-one autocorrelation $\rho(y,1)$ of the predictor. The parameter $\rho_1$ governs the smoothness of the predictor: higher values of $\rho_1$ generally promote smoother trajectories of $y_t$. Formally, assuming that the process $x_t$ is centered (zero-mean), Wildi (2024) establishes a connection between $\rho(y,1)$ and the expected duration between consecutive zero-crossings (sign changes) of $y_t$, the so-called holding time, denoted by $ht(y)$:
\begin{eqnarray*}\label{ht}
ht(y)=\frac{\pi}{\arccos(\rho(y,1))}.
\end{eqnarray*}
Since the non-linear transformation involved is strictly monotonic, the SSA criterion can be reformulated as the optimization problem:
\begin{eqnarray}\label{critssaht}
\left.\begin{array}{c}\rho(z,y)\to\max\\
\frac{\arccos(\rho(y,1))}{\pi}=1/ht_1\end{array}\right\},
\end{eqnarray}
where the smoothness parameter $1/ht_1$ expresses the rate of zero-crossings of the predictor\footnote{Formally, the exact relationship between the zero-crossing rate and the lag-one autocorrelation of the predictor assumes Gaussian time series, but Wildi (2024) demonstrates robustness to deviations from this assumption.}. \\

We now briefly discuss some implications of criterion \eqref{critssaht}. First, the objective function is indifferent to an affine transformation of the predictor. This ambiguity can be resolved by assuming an arbitrary scale and level for $y_t$ (standardization). Alternatively, a mean-square error norm (MSE) may be substituted for the target correlation, as noted in Wildi (2025). In this case, the classic MSE predictor $y_{t,MSE}$ is obtained as a solution to the SSA criterion by insertion of $1/ht_1:=1/ht_{MSE}$ in the constraint, where $ht_{MSE}$ represents the holding time of $y_{t,MSE}$. %If the MSE predictor is noisy, which is often the case in applications, one can select $ht_1>ht_{MSE}$  in the constraint to address the rate of noisy crossings or false alarms. 
Second, the concept can be extended to a multivariate framework, denoted as M-SSA, by generalizing the correlation functions, as detailed in Wildi (2025b). Lastly, sign changes in trend growth, represented by zero-crossings of the target $z_t$, are indicative of relevant changes in the trajectory of the economy, for example at transitions between expansions and recessions. Unfortunately, classic predictors often generate excessively many false `noisy' crossings, due to high-frequency leakage as illustrated by the amplitude function in Fig. \eqref{hp_160} (bottom right panel) and further elaborated upon by Wildi (2025). The SSA criterion helps control this phenomenon. More precisely, Wildi (2025) derives an equivalent dual formulation of criterion \eqref{critssaht}, which distinguishes the SSA solution as exhibiting the smallest rate of zero-crossings of any (linear) predictor with the same target correlation. We now rely on the M-SSA criterion to derive a `smooth' predictor $y_t$ of $z_t$. %In a second stage, serving as a generalization and refinement of the univariate HP-C utilized in the direct HP forecast.

\subsection{M-MSE and M-SSA Nowcasts}

For illustration, we consider the target $z_t$ generated by the output of the two-sided HP(160) filter applied to BIP, which we denote by HP-BIP. We compare four nowcast methods: the univariate HP-C, the multivariate MSE predictor (denoted by M-MSE), and two M-SSA predictors based on different HT constraints. The first design enforces $ht_1=1.5ht_{MSE}$, aiming to produce approximately $33\%$ fewer zero-crossings than the M-MSE predictor over the long term. The second design sets $ht_1=ht_{HP-C}$, ensuring that its zero-crossing rate matches that of the univariate HP-C. These constraints are intended to systematically compare the effects of different levels of smoothness on prediction performance. The estimation of the underlying VAR(1) model employed for the M-MSE and M-SSA nowcasts is based on data observations up to January 2008. This approach ensures a lengthy out-of-sample period that encompasses significant events such as the financial crisis, the sovereign debt crisis, and the COVID-19 pandemic. Further details and background information are available in Wildi (2025b). All computations are performed utilizing the M-SSA package, as documented in Wildi (2025b).
Figure \ref{bk_gammak} illustrates the coefficients of the multivariate filter designs, where increased smoothing achieved through the M-SSA method manifests as a distinct pattern in the filter weights, notably influencing the rate of decay across increasing lags. The resulting filtered series, presented in Figure \ref{mssa_msse_now}, are standardized to enable direct comparisons. The two M-SSA predictors adhere to the specified constraints: one exhibiting approximately one-third fewer zero-crossings than the M-MSE, and the other matching the zero-crossing rate of the HP-C, confirming the effectiveness of the smoothness control imposed during the design. This is further corroborated by Table \ref{corhtnow}, which reports empirical holding times and target correlations.  In this example, both the univariate design and the M-SSA method, when configured to replicate its holding time, demonstrate the highest level of smoothness, as indicated by the maximal duration between crossings. However, the HP-C also attains the lowest target correlation among the compared approaches. More generally, our example illustrates the prediction dilemma underlying the (M-)SSA criterion, highlighting the inherent trade-off between smoothness (zero-crossing rate) and predictive accuracy (target correlation).
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/bk_gammak.pdf}\caption{Multivariate M-MSE (left) and M-SSA  (middle:smaller HT, right: larger HT) nowcast filters.\label{bk_gammak}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/mssa_msse_now.pdf}\caption{Nowcasting: two-sided HP(160) applied to BIP (target, black), univariate HP-C (violet), M-SSA smaller HT (blue), M-SSA larger HT (orange) and M-MSE (green). The dashed vertical line delimits in- and out-of-sample spans. The first L=31 observations are lost due to initialization and the two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Thu May  1 21:38:41 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & HP-C & M-SSA smaller HT & M-SSA larger HT & M-MSE \\ 
  \hline
Target correlation & 0.55 & 0.63 & 0.59 & 0.68 \\ 
  HT & 8.67 & 6.50 & 8.67 & 4.00 \\ 
   \hline
\end{tabular}
\caption{Target correlations and HTs of HP-C, M-SSA and M-MSE nowcasts.  } 
\label{corhtnow}
\end{table}The results validate that the smoothing constraint incorporated into the multivariate extension of Criterion \eqref{critssa} effectively manages the zero-crossings of the predictor: sample estimates of the holding times (HTs) align with the expected values derived from Equation \eqref{ht}\footnote{Effective convergence of sample estimates to their theoretical expected values —assuming the true data-generating process is known— can be validated using the M-SSA package. This validation is typically performed through analyses of very long samples of simulated (multivariate) data, ensuring the consistency of the estimation procedures under idealized conditions.}. An increase of approximately $50\%$ in the HT relative to the M-MSE predictor has only a modest effect on the target correlation, suggesting that this tradeoff could be beneficial in this application. Conversely, matching the HT of the HP-C predictor through M-SSA results in an increase in the target correlation, demonstrating the impact of zero-crossing control in the multivariate design on predictive performance. Going forward, we now omit the M-MSE, as it is a special case of the more general M-SSA framework.

\subsection{M-SSA Forecasts}

The previously described nowcast example can be systematically extended to generate forecasts of the target variable. This extension utilizes the same empirical framework, incorporating an assumed $50\%$ increase in the HT over the M-MSE, and considers forward shifts of $h=1,...,6$  quarters of HP-BIP (note that M-SSA does not target BIP explicitly). %This approach facilitates the assessment of forecast performance across multiple horizons within the established methodological context. 
The influence of the forecast horizon on the M-SSA design is depicted in Fig. \eqref{bk_h}, which contrasts nowcast and one-year-ahead forecast: as $h$ increases (right panel), the scaling factor diminishes (zero-shrinkage), attributable to increased forecast uncertainty, and there is a relative amplification of the influence of the relevant indicators, leading BIP in relative terms.Additionally, a phase shift associated with the filter applied to BIP is observed (red line, right panel), reflecting the cyclical characteristics inherent to the HP filter. While this phase effect might cause an effective sign change in the BIP filter output as $h$ increases—something to be avoided in this context—the low-pass filters assigned to the additional 'leading' indicators (green and violet lines, right panel) continue to effectively track the low-frequency components of HP-BIP. The combination of the phase effect, applied to BIP, and level-tracking through these additional indicators enables more refined and effective tracking of the target compared to univariate forecasts, which rely solely on the phase-effect of the filter to look ahead of time.\\

M-SSA predictors are compared to the traditional univariate HP-C in Fig. \eqref{multivar_vs_univar} across the periods of the financial crisis (top panels) and the COVID-19 pandemic (bottom panels), for forecast horizons $h=0,...,6$. The nowcast from the previous section is included for reference. All series are standardized to facilitate visual comparison. The primary distinction of the M-SSA approach is the progressive left-shift of the predictor as a function of $h$; this shift becomes more pronounced with increasing $h$ and appears to be consistent across all levels, including the peaks and troughs of the series. Conversely, the positions of peaks and troughs in the HP-C series appear to remain largely unaffected by variations in $h$. The more pronounced and systematic left-shift observed in the M-SSA forecasts can be partly attributed to the increasing weight assigned to the additional indicators leading BIP, which dominate the BIP component at larger forecast horizons. %In comparison, the effect of phase shifts—arising from the cyclical nature of the HP-filter and manifesting in the HP-C forecasts (shown in the left panels)—seems less systematic. 
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/bk_h.pdf}\caption{Multivariate M-SSA: $h=0$ (nowcast, left) and $h=4$ (one year forecast, right).\label{bk_h}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/multivar_vs_univar.pdf}\caption{Now- and forecasts of HP-BIP for $h=0,...,6$: HP-C (left) vs. M-SSA (right) across the financial crisis (top) and the pandemic (bottom).\label{multivar_vs_univar}}\end{center}\end{figure}

\section{M-SSA Component Predictor}

The targets of the multivariate filters are the smoothed outputs obtained from the acausal two-sided HP(160) filter applied to the five indicators. However, our objective is to develop predictors that emphasize the unfiltered BIP. To achieve this, we can utilize the  M-SSA outputs, referred to as M-SSA components.

\subsection{Standardized Predictor: Equal-Weighting}

Assuming that all M-SSA components contribute equally to BIP prediction, an initial step involves averaging the standardized components to form an equally-weighted composite measure. Subsequently, the predictor is derived by regressing this standardized aggregate on the (forward-shifted) BIP series. Fig. \eqref{mssa_predictor} presents the BIP series alongside the equally-weighted M-SSA predictor. All series are standardized to facilitate visual comparison and to highlight the dynamic shifts, particularly the left-shift characteristic of the predictors.\\

To enhance interpretability and assessment of the M-SSA predictor, we propose analyzing its individual components. For illustration, Fig. \eqref{mssa_predictor_interpretability} shows the nowcast at $h=0$. The solid blue line (the nowcast) approaches the zero line near the end of the sample, with a recent upward movement mainly supported by the leading spread component. In contrast, the ifo- and ESI-components remain near the zero line, while the ip- and BIP-components appear to be awaiting additional evidence and confirmation before signaling a definitive movement.  This type of assessment can be useful for evaluating the relevance of recent changes in the forecasted time series dynamics.
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/mssa_predictor.pdf}\caption{BIP and equally-weighted M-SSA predictor: all series standardized.\label{mssa_predictor}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/mssa_predictor_interpretability.pdf}\caption{Equally-weighted M-SSA predictor and M-SSA components: all series standardized.\label{mssa_predictor_interpretability}}\end{center}\end{figure}
%???Eventually add p-values and/rRMSEs???

\subsection{Tracking BIP: Optimal Weighting}

We now utilize the M-SSA components as regressors on forward-shifted BIP, extending the previous (standardized) equally-weighted M-SSA predictor to incorporate optimal weighting of the components. While we keep the M-SSA fixed, as based on data up to January 2008, we implement an expanding window starting from January 2007 for deriving the component weights, which are up-dated on a quarterly basis (this contrasts with Section \eqref{direct_forecast} which was based on the full sample information). \\

There are various combinations of M-SSA components for tracking BIP. For illustration, we here select the single BIP component resulting from the M-SSA filter tracking HP-BIP. Table \eqref{p_val_hpbip_wc} reports HAC-adjusted p-values from regressions of the out-of-sample predictor on HP-BIP, indicating a strong link between the predictor and the target, out-of-sample. 
% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Thu May  1 21:38:45 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.000 & 0.000 & 0.001 & 0.002 & 0.005 & 0.020 & 0.112 \\ 
  Shift=1 & 0.000 & 0.000 & 0.000 & 0.001 & 0.002 & 0.005 & 0.019 \\ 
  Shift=2 & 0.000 & 0.000 & 0.000 & 0.001 & 0.001 & 0.003 & 0.006 \\ 
  Shift=3 & 0.018 & 0.007 & 0.002 & 0.000 & 0.002 & 0.003 & 0.004 \\ 
  Shift=4 & 0.180 & 0.100 & 0.040 & 0.010 & 0.002 & 0.005 & 0.005 \\ 
  Shift=5 & 0.579 & 0.432 & 0.261 & 0.108 & 0.027 & 0.008 & 0.009 \\ 
   \hline
\end{tabular}
\caption{HAC adjusted p-values of regressions of BIP (M-SSA) component optimized for forecast horizons $h=0,...,6$  (the columns) on HP-BIP shifted forward by shift=0,...,5 (the rows). Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{p_val_hpbip_wc}
\end{table}Our choice—focusing on the single M-SSA BIP component—emphasizes simplicity and interpretability, as future BIP and HP-BIP are connected through their shared low-frequency component. Additionally, the M-SSA BIP component is the most important predictor in the regression of the M-SSA outputs on BIP.  Finally, revisions due to quarterly updates of the regression equations are minimized by this predictor, and the estimates are more stable and robust over time\footnote{Other combinations can be experimented with using the M-SSA package. Regression weights of more complex designs, involving multiple components, tend to be more difficult to interpret due to the occurrence of negative weights and potential sign changes of the regressors over time, which may indicate overfitting or non-stationarity.} (see Section \eqref{revisions}). Note that information from the other  indicators is incorporated into the M-SSA BIP component through the multivariate filter design.\\

This setup allows us to evaluate the out-of-sample performance of various predictor approaches, including: (i) the simple mean of BIP, (ii) the direct forecast method described in Section \eqref{cdf}, (iii) the direct HP forecasts from Section \eqref{hpdf}, based on the univariate HP-C, and (iv) the new M-SSA BIP component predictor. %For completeness, we also consider an M-MSE component predictor derived from outputs of the multivariate MSE filter.

\subsection{Out-of-Sample Forecast Performances}\label{oosp}

%Out-of-sample performance metrics include p-values derived from regressions of the out-of-sample predictors on forward-shifted BIP, adjusted for heteroscedasticity and autocorrelation using HAC estimators. Additionally, we report relative root-mean-square forecast errors (rRMSE) benchmarked against various reference models.\\

Tables \eqref{p_val_wc} and \eqref{rRMSE_mSSA_comp_mean_wc} present HAC-adjusted p-values and rRMSEs for the BIP component predictor. The p-values are obtained from a regression of the out-of-sample predictor on forward-shifted BIP (in contrast to Table \eqref{p_val_hpbip_wc}, which emphasized HP-BIP); hence, they do not account for potential bias or mis-scaling of the predictor. In contrast, the rRMSEs, evaluated against the mean benchmark, do incorporate eventual bias and mis-scaling effects. Additional results involving other predictor benchmarks are provided in the appendix, based on data sets with and without the pandemic. The results can be summarized as follows: targeting future HP-BIP (see Table \eqref{p_val_hpbip_wc}) is less challenging than predicting BIP (see Table \eqref{p_val_wc}), due to smoothness (autocorrelation) of the target when compared to BIP. The pandemic period generally obscures the relationships between predictors and BIP. When excluding the pandemic, the direct forecasts outperform the mean forecast up to two quarters ahead. The M-SSA component predictor, optimized for longer forecast horizons, outperforms the mean benchmark up to one year ahead. Additionally, the BIP component predictor also surpasses the direct forecasts for larger forecast horizons ($h>2$ quarters).Statistical significance of the BIP component predictor, optimized for larger forecast horizons, can be established for at least up to one year ahead. Predictors optimized for $h\geq 4$ generally outperform the other designs, irrespective of the forward-shift of BIP.

%???check whether rMSE is decreasing towards sample end (because initial estimate towards 2007 is unreliable)???

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Thu May  1 21:38:45 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.084 & 0.047 & 0.017 & 0.005 & 0.002 & 0.001 & 0.038 \\ 
  Shift=1 & 0.230 & 0.160 & 0.094 & 0.032 & 0.005 & 0.001 & 0.004 \\ 
  Shift=2 & 0.801 & 0.666 & 0.414 & 0.173 & 0.026 & 0.005 & 0.006 \\ 
  Shift=3 & 0.876 & 0.961 & 0.917 & 0.683 & 0.248 & 0.068 & 0.039 \\ 
  Shift=4 & 0.370 & 0.819 & 0.990 & 0.953 & 0.533 & 0.117 & 0.030 \\ 
  Shift=5 & 0.353 & 0.802 & 0.983 & 0.943 & 0.571 & 0.239 & 0.043 \\ 
   \hline
\end{tabular}
\caption{HAC adjusted p-values of regressions of M-SSA component predictors optimized for forecast horizons $h=0,...,6$  (the columns) on BIP shifted forward by shift=0,...,5 (the rows). Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{p_val_wc}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Thu May  1 21:38:45 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.978 & 0.942 & 0.893 & 0.855 & 0.860 & 0.906 & 0.962 \\ 
  Shift=1 & 1.082 & 1.060 & 1.005 & 0.926 & 0.877 & 0.880 & 0.914 \\ 
  Shift=2 & 1.083 & 1.090 & 1.077 & 1.019 & 0.952 & 0.918 & 0.919 \\ 
  Shift=3 & 1.034 & 1.059 & 1.081 & 1.063 & 1.010 & 0.970 & 0.954 \\ 
  Shift=4 & 0.995 & 1.029 & 1.069 & 1.077 & 1.035 & 0.986 & 0.953 \\ 
  Shift=5 & 0.993 & 1.030 & 1.075 & 1.093 & 1.062 & 1.011 & 0.967 \\ 
   \hline
\end{tabular}
\caption{rRMSEs of M-SSA components predictor benchmarked against the expanding mean of BIP. Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{rRMSE_mSSA_comp_mean_wc}
\end{table}

\section{Revisions}\label{revisions}

Revisions refer to changes in the historical values of a time series resulting from updates with new information. Revisions of the M-SSA component predictor can arise from data updates, which we do not analyze further here\footnote{Except for BIP, the other indicators are not or only slightly revised. Moreover, the smoothing effect of filters mitigates the effect of revisions when compared to direct forecasts.}, or from updates to predictor parameters. The latter can be further separated into changes in VAR parameters and regression weights. 
\subsection{Revisions: Regression Parameters}

For illustration, we here focus on a forecast horizon $h=4$ for the BIP component predictor and a forward-shift $shift=4$ for BIP, noting that similar results would be obtained with other combinations of $h$ and $shift$. 
%???Check left-shift of MA-inversion of M-SSA and M-MSE
%??? Check forecast excess for HP-C
%??? Justification HP(160): replace HP(1600) by AR(2) with similar persistency but shorter cycle lengthi, i.e. $a_1=-2"persistency*acos(\theta, a_2=persistency$.
Fig.\eqref{revisions1} compares the final and real-time predictors: as the sample size increases, the real-time predictor converges steadily to the final predictor and the rate of convergence is partially determined by the simplicity of our design, which avoids incorporating additional M-SSA components. 
As a confirmation, Fig.\eqref{revisions2} displays the evolution of the intercept and regression weights over time: with increasing sample size, these estimates seem to converge to fixed points, indicating both the stationarity of the process and the consistency of the estimates (note, however, the slightly elevated values of the regression weight assigned to the M-SSA component during the financial crisis and the pandemic, suggesting a stronger link between the target and the predictor).  In contrast, more complex designs based on multiple M-SSA components can exhibit substantial fluctuations in the regression parameters, sometimes changing signs. This suggests instability and complicates interpretation.
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/revisions1.pdf}\caption{Revisions of the M-SSA BIP component predictor are shown, with the final predictor in blue and the real-time design in red. These revisions result from quarterly updates to the regression equations. The M-SSA filter itself remains fixed, based on data up to January 2008.\label{revisions1}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/revisions2.pdf}\caption{Revisions of intercept and regression weight.\label{revisions2}}\end{center}\end{figure}

\subsection{Revisions: M-SSA}

The second source of revisions in the M-SSA component predictor arises from updates to the VAR model within the M-SSA filter. We compare the original estimates, based on data up to January 2008, with the final estimates obtained using $h=4$ and $shift=4$ (similar findings apply to other values of forecast horizon and forward-shift).  For consistency, we excluded the entire pandemic episode from the data. Fig.\eqref{up_dated_ma_inv_multi_ip} compares the MA-inversions (impulse responses) of the BIP equation derived from the 'old' and the updated VAR models: unlike the former, the latter incorporates ESI as an additional explanatory variable for BIP. Fig.\eqref{bk_2008_all} illustrates the resulting effect on the M-SSA filter targeting HP-BIP at forecast horizon $h=4$. The updated filter (right panel) assigns more weight to the additional 'leading' indicators relative to BIP. Finally, Fig.\eqref{revisions3} compares the M-SSA BIP predictors. To isolate the effect of the regression revision, the figure displays standardized series. The update to the VAR model influences the left-shift of the predictor (blue line), which appears slightly 'faster' than the original predictor (red line). This suggests that the out-of-sample performance of the M-SSA component predictor may be better than reported in Section \eqref{oosp}, which is based on the fixed and increasingly outdated version of M-SSA.
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/up_dated_ma_inv_multi_ip.pdf}\caption{MA-inversion (impulse response) of VAR: data up to Jan-2008 (left) vs. entire data set (right).\label{up_dated_ma_inv_multi_ip}}\end{center}\end{figure}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/bk_2008_all.pdf}\caption{M-SSA filter tracking HP-BIP at forecast horizon $h=4$: data up to Jan-2008 (left) vs. entire data set (right).\label{bk_2008_all}}\end{center}\end{figure}

\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/revisions3.pdf}\caption{Revisions resulting from updates to the VAR model in M-SSA. To isolate the effect of the regression revision, all series are standardized. The up-dated M-SSA (blue) is left-shifted at the zero-crossings and appears slightly faster. For consistency, the entire pandemic episode has been removed.\label{revisions3}}\end{center}\end{figure}

\section{Summary and Conclusion}

Classic direct forecasts of BIP can outperform simple benchmarks, such as the mean, up to two quarters ahead. However, this limited forecast horizon is often insufficient for certain applications. A key challenge with direct forecasts is that the relevant indicators tend to be noisy, which can mask the true signal and lead to overfitting. In particular, the predictor struggles to timely track dips (slowdowns, recessions) or peaks (recoveries, expansions). To address this issue, we apply a HP(160)-filter to smooth the data, with the smoothing parameter $\lambda=160$ reflecting a more adaptive design than the classic quarterly HP, in line with an intended forecast horizon of up to one year ahead. The resulting predictor tends to outperform the classic direct forecasts in terms of statistical significance, mainly due to a slight left-shift noticeable at sharp recession dips. However, despite the filtering, the predictor remains somewhat noisy, potentially producing excessive noisy zero-crossings at the transitions between recessions and expansions. This issue is partly due to noise leakage inherent in the one-sided HP filter. \\

To address these problems, we propose a multivariate extension of SSA based on Wildi (2024). Unlike the univariate HP filter, the M-SSA can leverage cross-sectional information from indicators leading BIP in real time while controlling for the rate of zero-crossings of the predictor. Consequently, the multivariate filter outputs become increasingly left-shifted as the forecast horizon lengthens, allowing dips and peaks of the target series to be tracked more systematically, while reducing the occurrence of noisy zero-crossings. However, since M-SSA does not explicitly target BIP, an additional step is necessary to derive the final BIP predictor. We propose two approaches: a simple equally-weighted aggregate, assuming all M-SSA components are equally important for predicting BIP, and an optimally weighted predictor based on a regression of the M-SSA components on future BIP.\\

For illustration, we consider a simple approach based on regressing the single M-SSA BIP output on future BIP, thereby ignoring the other filter outputs. This predictor is intuitively appealing because the future HP-BIP—i.e., the target of M-SSA—is the low-frequency component of the future BIP target. Consequently, a strong link between M-SSA and future HP-BIP also indicates a connection with BIP, even if the statistical significance is obscured by the noise inherent in BIP. Out-of-sample performance suggests that this predictor is statistically significant at horizons up to one year. Moreover, designs optimized for larger forecast horizons outperform the mean benchmark at all considered forecast horizons, and the predictor exceeds the performance of direct forecasts—based on unfiltered or HP filtered indicators—at horizons longer than two quarters. An analysis of revision errors arising from quarterly updates of the proposed M-SSA BIP component predictor indicates that the regression parameters, the VAR model, and the resulting M-SSA filters stabilize after an initial burn-in period. This stability enhances the explainability and interpretability of the predictor.\\

In conclusion, the new predictor design integrates the traditional direct forecast approach with a novel multivariate filter to target BIP up to one year ahead. This multivariate approach leverages a set of economic indicators leading BIP in real time while controlling the smoothness of the predictor through the zero-crossing rate.

\newpage

\section{Appendix}

\subsection{Performances: Without Pandemic}

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Thu May  1 21:38:46 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.084 & 0.047 & 0.017 & 0.005 & 0.002 & 0.001 & 0.038 \\ 
  Shift=1 & 0.230 & 0.160 & 0.094 & 0.032 & 0.005 & 0.001 & 0.004 \\ 
  Shift=2 & 0.801 & 0.666 & 0.414 & 0.173 & 0.026 & 0.005 & 0.006 \\ 
  Shift=3 & 0.876 & 0.961 & 0.917 & 0.683 & 0.248 & 0.068 & 0.039 \\ 
  Shift=4 & 0.370 & 0.819 & 0.990 & 0.953 & 0.533 & 0.117 & 0.030 \\ 
  Shift=5 & 0.353 & 0.802 & 0.983 & 0.943 & 0.571 & 0.239 & 0.043 \\ 
   \hline
\end{tabular}
\caption{HAC adjusted p-values of regressions of M-SSA component predictors optimized for forecast horizons $h=0,...,6$  (the columns) on BIP shifted forward by shift=0,...,5 (the rows). Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{p_val_wc5}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Thu May  1 21:38:46 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.978 & 0.942 & 0.893 & 0.855 & 0.860 & 0.906 & 0.962 \\ 
  Shift=1 & 1.082 & 1.060 & 1.005 & 0.926 & 0.877 & 0.880 & 0.914 \\ 
  Shift=2 & 1.083 & 1.090 & 1.077 & 1.019 & 0.952 & 0.918 & 0.919 \\ 
  Shift=3 & 1.034 & 1.059 & 1.081 & 1.063 & 1.010 & 0.970 & 0.954 \\ 
  Shift=4 & 0.995 & 1.029 & 1.069 & 1.077 & 1.035 & 0.986 & 0.953 \\ 
  Shift=5 & 0.993 & 1.030 & 1.075 & 1.093 & 1.062 & 1.011 & 0.967 \\ 
   \hline
\end{tabular}
\caption{rRMSEs of M-SSA components predictor benchmarked against the expanding mean of BIP. Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{rRMSE_mSSA_comp_mean6}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Thu May  1 21:38:46 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 1.144 & 1.102 & 1.045 & 0.999 & 1.006 & 1.059 & 1.125 \\ 
  Shift=1 & 1.238 & 1.213 & 1.151 & 1.060 & 1.004 & 1.007 & 1.047 \\ 
  Shift=2 & 1.178 & 1.186 & 1.171 & 1.108 & 1.035 & 0.999 & 0.999 \\ 
  Shift=3 & 1.011 & 1.035 & 1.057 & 1.039 & 0.988 & 0.948 & 0.933 \\ 
  Shift=4 & 1.010 & 1.045 & 1.086 & 1.094 & 1.051 & 1.002 & 0.968 \\ 
  Shift=5 & 0.982 & 1.019 & 1.064 & 1.081 & 1.051 & 1.000 & 0.957 \\ 
   \hline
\end{tabular}
\caption{rRMSEs of M-SSA components predictor benchmarked against the direct forecasts. Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{rRMSE_mSSA_comp_direct_without_covid7}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Thu May  1 21:38:46 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.855 & 0.855 & 0.855 & 0.855 & 0.855 & 0.855 & 0.855 \\ 
  Shift=1 & 0.874 & 0.874 & 0.874 & 0.874 & 0.874 & 0.874 & 0.874 \\ 
  Shift=2 & 0.919 & 0.919 & 0.919 & 0.919 & 0.919 & 0.919 & 0.919 \\ 
  Shift=3 & 1.023 & 1.023 & 1.023 & 1.023 & 1.023 & 1.023 & 1.023 \\ 
  Shift=4 & 0.985 & 0.985 & 0.985 & 0.985 & 0.985 & 0.985 & 0.985 \\ 
  Shift=5 & 1.010 & 1.010 & 1.010 & 1.010 & 1.010 & 1.010 & 1.010 \\ 
   \hline
\end{tabular}
\caption{rRMSEs direct forecasts benchmarked against the expanding mean. Out-of-sample span starting in Jan-2007, without pandemic.} 
\label{rRMSE_mSSA_direct_mean_without_covid8}
\end{table}

\newpage
\subsection{Performances: Including Singular Pandemic Data}

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Thu May  1 21:38:46 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.127 & 0.064 & 0.019 & 0.003 & 0.001 & 0.002 & 0.012 \\ 
  Shift=1 & 0.439 & 0.314 & 0.194 & 0.093 & 0.033 & 0.017 & 0.036 \\ 
  Shift=2 & 0.718 & 0.565 & 0.326 & 0.098 & 0.020 & 0.008 & 0.010 \\ 
  Shift=3 & 0.946 & 0.975 & 0.925 & 0.657 & 0.222 & 0.059 & 0.028 \\ 
  Shift=4 & 0.327 & 0.699 & 0.964 & 0.942 & 0.619 & 0.237 & 0.102 \\ 
  Shift=5 & 0.383 & 0.755 & 0.946 & 0.910 & 0.581 & 0.241 & 0.055 \\ 
   \hline
\end{tabular}
\caption{HAC adjusted p-values of regressions of M-SSA component predictors optimized for forecast horizons $h=0,...,6$  (the columns) on BIP shifted forward by shift=0,...,5 (the rows). Out-of-sample span starting in Jan-2007, including the pandemic.} 
\label{p_val1}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Thu May  1 21:38:46 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 1.008 & 0.990 & 0.964 & 0.939 & 0.930 & 0.942 & 0.967 \\ 
  Shift=1 & 1.064 & 1.057 & 1.036 & 1.001 & 0.976 & 0.968 & 0.973 \\ 
  Shift=2 & 1.033 & 1.034 & 1.026 & 0.999 & 0.969 & 0.955 & 0.957 \\ 
  Shift=3 & 1.025 & 1.034 & 1.040 & 1.028 & 1.001 & 0.979 & 0.967 \\ 
  Shift=4 & 0.995 & 1.011 & 1.030 & 1.037 & 1.022 & 1.003 & 0.990 \\ 
  Shift=5 & 1.001 & 1.018 & 1.039 & 1.048 & 1.034 & 1.009 & 0.986 \\ 
   \hline
\end{tabular}
\caption{rRMSEs of M-SSA components predictor benchmarked against the expanding mean of BIP. Out-of-sample span starting in Jan-2007, including the pandemic.} 
\label{rRMSE_mSSA_comp_mean2}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Thu May  1 21:38:46 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 1.210 & 1.188 & 1.157 & 1.127 & 1.116 & 1.130 & 1.161 \\ 
  Shift=1 & 0.982 & 0.977 & 0.957 & 0.925 & 0.902 & 0.895 & 0.899 \\ 
  Shift=2 & 1.053 & 1.054 & 1.046 & 1.018 & 0.987 & 0.973 & 0.975 \\ 
  Shift=3 & 1.046 & 1.056 & 1.062 & 1.050 & 1.022 & 0.999 & 0.988 \\ 
  Shift=4 & 0.969 & 0.984 & 1.003 & 1.010 & 0.996 & 0.977 & 0.964 \\ 
  Shift=5 & 0.982 & 0.998 & 1.019 & 1.027 & 1.014 & 0.990 & 0.967 \\ 
   \hline
\end{tabular}
\caption{rRMSEs of M-SSA components predictor benchmarked against the direct forecasts. Out-of-sample span starting in Jan-2007, including the pandemic.} 
\label{rRMSE_mSSA_comp_direct3}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Thu May  1 21:38:46 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.833 & 0.833 & 0.833 & 0.833 & 0.833 & 0.833 & 0.833 \\ 
  Shift=1 & 1.083 & 1.083 & 1.083 & 1.083 & 1.083 & 1.083 & 1.083 \\ 
  Shift=2 & 0.981 & 0.981 & 0.981 & 0.981 & 0.981 & 0.981 & 0.981 \\ 
  Shift=3 & 0.979 & 0.979 & 0.979 & 0.979 & 0.979 & 0.979 & 0.979 \\ 
  Shift=4 & 1.027 & 1.027 & 1.027 & 1.027 & 1.027 & 1.027 & 1.027 \\ 
  Shift=5 & 1.020 & 1.020 & 1.020 & 1.020 & 1.020 & 1.020 & 1.020 \\ 
   \hline
\end{tabular}
\caption{rRMSEs direct forecasts benchmarked against the expanding mean. Out-of-sample span starting in Jan-2007, including the pandemic.} 
\label{rRMSE_mSSA_direct_mean4}
\end{table}

\end{document}

### 2025-05-02 R/M_SSA_utility_functions.R

@@ -1,0 +1,952 @@
+
+# Compute target: two-sided HP
+HP_target_sym_T<-function(n,lambda_HP,L)
+{
+  HP_obj<-HP_target_mse_modified_gap(L,lambda_HP)
+  
+  hp_symmetric=HP_obj$target
+  hp_classic_concurrent=HP_obj$hp_trend
+  hp_one_sided<-HP_obj$hp_mse
+  # Target first series  
+  gamma_target<-c(hp_one_sided,rep(0,(n-1)*L))
+  # We now proceed to specifying the targets of the remaining n-1 series
+  for (i in 2:n)
+    gamma_target<-rbind(gamma_target,c(rep(0,(i-1)*L),hp_one_sided,rep(0,(n-i)*L)))
+  # The above target filters are one-sided (right half of two-sided filter)
+  # We now tell M-SSA that it has to mirror the above filters at their center points to obtain two-sided targets
+  symmetric_target<-T
+  return(list(gamma_target=gamma_target,symmetric_target=symmetric_target))
+}  
+
+
+
+# Compute MA-inversion as based on VAR model
+MA_inv_VAR_func<-function(Phi,Theta,L,n,Plot=F)
+{
+  # MA inversion of VAR
+  # MA inversion is used because the M-SSA optimization criterion relies an white noise
+  #   For autocorrelated data, we thus require the MA-inversion of the DGP
+  xi_psi<-PSIwgt(Phi = Phi, Theta = NULL, lag = L, plot = F, output = F)
+  xi_p<-xi_psi$psi.weight
+  # Transform Xi_p into Xi as structured/organized for M-SSA
+  #   First L entries, from left to right, are weights of first explanatory series, next L entries are weights of second WN 
+  xi<-matrix(nrow=n,ncol=n*L)
+  for (i in 1:n)
+  {
+    for (j in 1:L)
+      xi[,(i-1)*L+j]<-xi_p[,i+(j-1)*n]
+  }
+  if (Plot)
+  {
+    # Plot MA inversions  
+    par(mfrow=c(1,n))
+    for (i in 1:n)#i<-1
+    {
+      mplot<-xi[i,1:min(10,L)]
+      
+      for (j in 2:n)
+      {
+        mplot<-cbind(mplot,xi[i,(j-1)*L+1:min(10,L)])
+        
+      }
+      ts.plot(mplot,col=rainbow(ncol(mplot)),main=paste("MA inversion ",colnames(x_mat)[i],sep=""))
+    }
+  }
[…]

### 2025-05-02 SSA Tutorials/Replicate_Wildi_2024_JBCR_paper.R

@@ -1,0 +1,1645 @@
+
+# The present tutorial replicates Figures and Tables in the JBCY-paper
+
+#-----------------------------------------------------------------------
+# Make a clean-sheet, load packages and functions
+rm(list=ls())
+
+library(tis)
+
+library(xts)
+# Load the library mFilter
+# HP and BK filters
+library(mFilter)
+
+library(ggplot2)
+library("gplots")
+
+
+
+# Load all relevant SSA-functions
+source(paste(getwd(),"/R/simple_sign_accuracy.r",sep=""))
+# Load tau-statistic: quantifies time-shift performances (lead/lag)
+source(paste(getwd(),"/R/Tau_statistic.r",sep=""))
+
+# Load signal extraction functions used for JBCY paper (relies on mFilter)
+source(paste(getwd(),"/R/HP_JBCY_functions.r",sep=""))
+
+# Additional functions used in JBCY paper
+# Some of them apply filters to data (can work with xts data!)
+# Others are setting-up specialized plots used in section 4: SSA and HP applied to monthly INDPRO 
+#   (for example plots with shadings of NBER-recessions)
+source(paste(getwd(),"/R/additional_filter_plot_functions_for_JBCY_paper.r",sep=""))
+#-------------------------------------------------------
+###################################################
+### Replicate section figure 1 in 2.2
+###################################################
+# Compute HP and B-K filters and compare concurrent MSE and SSA designs
+# Filter length
+L<-200
+# 1. HP
+lambda_monthly<-14400
+
+HP_obj<-HP_target_mse_modified_gap(L,lambda_monthly)
+# Bi-infinite HP
+hp_target=HP_obj$target
+ts.plot(hp_target)
+# Concurrent gap: as applied to series in levels
+hp_gap=HP_obj$hp_gap
+ts.plot(hp_gap)
+# Concurrent gap: as applied to series in differences
+modified_hp_gap=HP_obj$modified_hp_gap
+ts.plot(modified_hp_gap)
[…]

### 2025-05-02 SSA Tutorials/tutorial_3_SSA_applied_to_Hamilton_filter.R

@@ -1,0 +1,1739 @@
+# In this tutorial we propose an application of SSA to Hamilton's regression  filter (HF)
+# For illustration, we consider quarterly GDP (example 1) and monthly non-farm payroll (examples 2 and 3) data  
+# We analyze long sample periods (starting at WWII) as well as shorter spans (starting at the great moderation)
+
+# Main outcomes: 
+#   1.HF is a lowpass (transformed filter as applied to returns): it does not generate spurious cycles
+#       In contrast to BK, CF or HP-gap bandpass designs, see tutorials 2 (example 7), 4 and 5
+#   2.HF removes the (remaining) weak drift of the returns: it can address arbitrary integration orders by modifying the AR-order p of the regression
+#   3.By its very definition, the HF filter depends on the data-sample (regression model)
+#       -The out-of-sample cycle is non-stationary (regression parameters do not sum to one).
+#       -Therefore, the filter-regression must be continuously up-dated, as new information becomes available.
+#       -Up-dating the regression generates (undesirable) revisions.
+#   4.HF has a rather small holding-time and a rather large phase-lag 
+#       It is subject to noise leakage (`noisy' zero crossings) and retardation
+#       Therefore, an application of SSA is not inopportune
+#   5.We want the SSA-nowcast to be smoother than HF (~30% less crossings in the long run)
+#   6.The SSA forecasts (6 months and 1-year) are smooth, too, and leading 
+
+
+# Note: our intention is not to push a particular BCA-tool. Rather, we strive at illustrating that a particular 
+#   predictor or BCA-filter (any one as long as it's linear in the data) can be replicated and modified by SSA 
+#   in view of addressing 
+# 1. smoothness (noise suppression) and 
+# 2. timeliness (advancement)
+# In this perspective, HF is considered as a basic platform and a vitrine for showcasing SSA
+#   -We offer a number of compelling performance measures, confirming pertinence of a simple novel optimization principle  
+
+#-----------------------------------------------------------------------
+# Make a clean-sheet, load packages and functions
+
+rm(list=ls())
+
+library(xts)
+# This package implements the `never hpfilter', i.e., the Hamilton filter:  
+# It is used here for background information only (generate some plots): we deploy own code for the Hamilton filter 
[…]

### 2025-05-02 SSA Tutorials/tutorial_4_SSA_applied_to_BK_filter.R

@@ -1,0 +1,738 @@
+# In this tutorial we discuss an application of SSA to the Baxter and King (BK) ideal bandpass filter.
+
+# Main outcomes: 
+#   1.The analysis suggests that BK is not well suited for BCA (business cycle analysis) 
+#         because the filter generates a `spurious' cycle (see our analysis of the amplitude functions below). 
+#       -The periodicity of the nearly sinusoidal cycle is determined entirely by the lower bandpass specification (artifact)
+#   2.The BK filter output is very smooth (sinusoidal like) and therefore an application of SSA is not necessary 
+#       or useful in this case. 
+#       -Increasing smoothness even more, by imposing a larger holding-time in SSA, leads to atypical filter-characteristics in this particular setting  
+# Recommendations: 
+#   -Be cautious when applying bandpass designs for the purpose of BCA (HP-gap is subject to similar shortfalls)
+#   -Compute amplitude functions to assess filter characteristics (detect spurious cycles). 
+#   -Verify that SSA does the right job, namely: lessen the number of noisy-crossings attributable to undesirable leakage of high-frequency components by the benchmark  
+# In short: this tutorial is an illustrative counter example with some unexpected (but statistically sound) outcomes 
+#-----------------------------------------------------------------------
+# Make a clean-sheet, load packages and functions
+rm(list=ls())
+
+library(xts)
+# HP and BK filters
+library(mFilter)
+# Load data from FRED
+library(quantmod)
+
+
+# Load all relevant SSA-functions
+source(paste(getwd(),"/R/simple_sign_accuracy.r",sep=""))
+# Load tau-statistic: quantifies time-shift performances (lead/lag)
+source(paste(getwd(),"/R/Tau_statistic.r",sep=""))
+
+# Load signal extraction functions used for JBCY paper (relies on mFilter)
+source(paste(getwd(),"/R/HP_JBCY_functions.r",sep=""))
+
+
+#----------------------------------------------------------------------
+# Source data directly from FRED
+getSymbols('PAYEMS',src='FRED')
+
+# Discard Pandemic: extreme outliers affect regression of Hamilton filter
[…]

### 2025-05-02 SSA Tutorials/tutorial_5_maximal_monotone_nowcast.R

@@ -1,0 +1,333 @@
+# Tutorial on maximal monotone predictors for non-stationary (I(1)) time series
+
+# Previous tutorials 1-4 are based on Wildi, M. (2024) https://doi.org/10.1007/s41549-024-00097-5
+# This (new) tutorial 5 is based on Wildi (2025). The new paper proposes: 
+#   -Technical/formal background to Wildi (2024)
+#   -Introduction of a new SSA predictor for non-stationary time series
+
+# For non-stationary time series, an extension of SSA is proposed which tracks the target optimally in a MSE sense
+#   (predictor and target are cointegrated) and which controls the rate of zero-crossings of (stationary) FIRST DIFFERENCES of the predictor.
+# The predictor tracks the target as closely as possible subject to a holding-time constraint imposed to first differences.
+#   -In its dual form: the predictor generates the least number of zero-crossings in first differences for a given track accuracy (MSE).
+#   -Equivalently: the original predictor (in levels) is maximal monotone
+
+# Application/prediction problem:
+# -We compute a maximal monotone SSA nowcast for the two-sided (bi-infinite) HP(14400) applied to monthly US-INDPRO.
+# -We compute the MSE nowcast of HP based on an ARIMA(1,1,0)-model for INDPRO: benchmark 1
+#   -This MSE nowcast is much noisier than the classic HP concurrent (real-time) filter, widely used in business-cycle applications
+# -Therefore, we consider the classic concurrent one-sided HP (HP-C): benchmark 2
+#   -We replicate the holding-time (smoothness) of HP-C by SSA
+# -By optimization we then expect SSA to equal HP-C in terms of smoothness and to outperform HP-C in terms of MSE (tracking accuracy)
+# -Since SSA is much smoother than the MSE nowcast we are interested in finding out how much SSA looses in terms of MSE performances.
+#   -The SSA criterion minimizes MSE losses for given smoothness (holding-time)
+
[…]

### 2025-05-02 SSA Tutorials/tutorial_7.0_introduction_multivariate_SSA_MSSA.R

@@ -1,0 +1,173 @@
+# We here introduce to the M-SSA. see paper provide link???
+# Work in progress...
+
+# Make a clean-sheet, load packages and functions
+rm(list=ls())
+
+library(xts)
+# Load the library mFilter
+# HP and BK filters
+library(mFilter)
+# Plot for heat map of Trilemma
+library(ggplot2)
+library("gplots")
+
+# Load all relevant SSA-functions
+source(paste(getwd(),"/R/simple_sign_accuracy.r",sep=""))
+# Load tau-statistic: quantifies time-shift performances (lead/lag)
+source(paste(getwd(),"/R/Tau_statistic.r",sep=""))
+
+# Load signal extraction functions used for JBCY paper (relies on mFilter)
+source(paste(getwd(),"/R/HP_JBCY_functions.r",sep=""))
+
+##########################################################################################################
+# Example 1: we replicate SSA by M-SSA
+# We first compute a univariate SSA design
+# For this purpose we rely on example 1 of tutorial 2.1
+# This example relies on targeting a two-sided HP filter
+#--------------------------
+# a. Derivation
+# We use the R-package mFilter for computing HP 
+# Specify filter length: should be an odd number since otherwise the two-sided HP filter could not be adequately centered 
+L<-201
+# Should be an odd number
+if (L/2==as.integer(L/2))
+{
+  print("Filter length should be an odd number")
+  print("If L is even then HP cannot be adequately centered")
+  L<-L+1
+}  
+# Specify lambda: monthly design
+lambda_monthly<-14400
+par(mfrow=c(1,1))
+HP_obj<-HP_target_mse_modified_gap(L,lambda_monthly)
+# Bi-infinite two-sided (symmetric) HP
+hp_target<-HP_obj$target
+ts.plot(hp_target)
+# Concurrent gap: as applied to series in levels: this is a high pass filter
+hp_gap=HP_obj$hp_gap
+ts.plot(hp_gap)
+# Concurrent HP assuming I(2)-process 
+# This is the Classic concurrent or one-sided low pass HP, see e.g. McElroy (2006)
+hp_trend=HP_obj$hp_trend
+ts.plot(hp_trend)
+
[…]

### 2025-05-02 SSA Tutorials/tutorial_7.4_MSSA_component_predictor.R

@@ -1,0 +1,1767 @@
+# Tutorial 7.4
+# Main purposes:
+# 1. Forecasting BIP multiple quarters ahead
+#   -Derive a new M-SSA components predictor which addresses MSE forecast performances when tracking BIP
+#     -The original M-SSA predictor (without the attribute `components') proposed in tutorial 7.3 is 
+#         standardized: neither level nor scale were calibrated to track BIP 
+#     -The original M-SSA predictor was designed to track dynamic changes in the trend growth-rate of BIP (HP-BIP)
+#     -The original M-SSA predictor was based on aggregating equally the standardized M-SSA components 
+#   -We here propose a more refined design, the `M-SSA components predictor', whereby equal weighting 
+#       (of M-SSA components) is replaced by an additional (new) optimal weighting step.
+# 2. Interpretability: 
+#   -Exploit the M-SSA components to gauge forecasts by the proposed M-SSA predictor(s)
+#   -How trustworthy is a change of the predictor outlook?
+# 3. Explainability: 
+#   -Justify the design and construction steps underlying the proposed M-SSA predictor(s)
+#   -Determine which step(s) in the construction of the predictor(s) generate(s) forecast gains?
+
+# Proceeding:
+# -We rely on the design proposed in tutorial 7.3, exercise 1
+# -We then add a new optimal weighting step in the construction of the final predictor
+
+# The tutorial is organized into 6 exercises
+# -Exercise 1
+#   -M-SSA components: derive M-SSA components and replicate the original M-SSA predictor (tutorial 7.3)
+#   -Interpretability: rely on M-SSA components to gauge forecasts (reliability/trustworthiness)
+#   -Forecast BIP: new additional optimization step
+#   -Out-of-sample performance evaluation
+#     -Comparison vs. simple mean, direct forecast and original M-SSA predictor
+# -Exercise 2
+#   -Analysis of revisions of new (real-time out-of-sample) M-SSA components predictor
+# -Exercise 3
+#   -Application of the predictor to German BIP
+# -Exercise 4
+#   -Explainability: why does the M-SSA component predictor outperform specifically at multi-quarters 
+#     ahead forecast horizons?
+# -Exercise 5
[…]

### 2025-05-02 SSA Tutorials/tutorial_3_SSA_applied_to_Hamilton_filter.R
up

@@ -1,1739 +1,0 @@
-# In this tutorial we propose an application of SSA to Hamilton's regression  filter (HF)
-# For illustration, we consider quarterly GDP (example 1) and monthly non-farm payroll (examples 2 and 3) data  
-# We analyze long sample periods (starting at WWII) as well as shorter spans (starting at the great moderation)
-
-# Main outcomes: 
-#   1.HF is a lowpass (transformed filter as applied to returns): it does not generate spurious cycles
-#       In contrast to BK, CF or HP-gap bandpass designs, see tutorials 2 (example 7), 4 and 5
-#   2.HF removes the (remaining) weak drift of the returns: it can address arbitrary integration orders by modifying the AR-order p of the regression
-#   3.By its very definition, the HF filter depends on the data-sample (regression model)
-#       -The out-of-sample cycle is non-stationary (regression parameters do not sum to one).
-#       -Therefore, the filter-regression must be continuously up-dated, as new information becomes available.
-#       -Up-dating the regression generates (undesirable) revisions.
-#   4.HF has a rather small holding-time and a rather large phase-lag 
-#       It is subject to noise leakage (`noisy' zero crossings) and retardation
-#       Therefore, an application of SSA is not inopportune
-#   5.We want the SSA-nowcast to be smoother than HF (~30% less crossings in the long run)
-#   6.The SSA forecasts (6 months and 1-year) are smooth, too, and leading 
-
-
-# Note: our intention is not to push a particular BCA-tool. Rather, we strive at illustrating that a particular 
-#   predictor or BCA-filter (any one as long as it's linear in the data) can be replicated and modified by SSA 
-#   in view of addressing 
-# 1. smoothness (noise suppression) and 
-# 2. timeliness (advancement)
-# In this perspective, HF is considered as a basic platform and a vitrine for showcasing SSA
-#   -We offer a number of compelling performance measures, confirming pertinence of a simple novel optimization principle  
-
-#-----------------------------------------------------------------------
-# Make a clean-sheet, load packages and functions
-
-rm(list=ls())
-
-library(xts)
-# This package implements the `never hpfilter', i.e., the Hamilton filter:  
-# It is used here for background information only (generate some plots): we deploy own code for the Hamilton filter 
[…]

### 2025-05-02 SSA Tutorials/tutorial_4_SSA_applied_to_BK_filter.R

@@ -1,738 +1,0 @@
-# In this tutorial we discuss an application of SSA to the Baxter and King (BK) ideal bandpass filter.
-
-# Main outcomes: 
-#   1.The analysis suggests that BK is not well suited for BCA (business cycle analysis) 
-#         because the filter generates a `spurious' cycle (see our analysis of the amplitude functions below). 
-#       -The periodicity of the nearly sinusoidal cycle is determined entirely by the lower bandpass specification (artifact)
-#   2.The BK filter output is very smooth (sinusoidal like) and therefore an application of SSA is not necessary 
-#       or useful in this case. 
-#       -Increasing smoothness even more, by imposing a larger holding-time in SSA, leads to atypical filter-characteristics in this particular setting  
-# Recommendations: 
-#   -Be cautious when applying bandpass designs for the purpose of BCA (HP-gap is subject to similar shortfalls)
-#   -Compute amplitude functions to assess filter characteristics (detect spurious cycles). 
-#   -Verify that SSA does the right job, namely: lessen the number of noisy-crossings attributable to undesirable leakage of high-frequency components by the benchmark  
-# In short: this tutorial is an illustrative counter example with some unexpected (but statistically sound) outcomes 
-#-----------------------------------------------------------------------
-# Make a clean-sheet, load packages and functions
-rm(list=ls())
-
-library(xts)
-# HP and BK filters
-library(mFilter)
-# Load data from FRED
-library(quantmod)
-
-
-# Load all relevant SSA-functions
-source(paste(getwd(),"/R/simple_sign_accuracy.r",sep=""))
-# Load tau-statistic: quantifies time-shift performances (lead/lag)
-source(paste(getwd(),"/R/Tau_statistic.r",sep=""))
-
-# Load signal extraction functions used for JBCY paper (relies on mFilter)
-source(paste(getwd(),"/R/HP_JBCY_functions.r",sep=""))
-
-
-#----------------------------------------------------------------------
-# Source data directly from FRED
-getSymbols('PAYEMS',src='FRED')
-
-# Discard Pandemic: extreme outliers affect regression of Hamilton filter
[…]

### 2025-05-02 SSA Tutorials/tutorial_5_maximal_monotone_nowcast.R

@@ -1,333 +1,0 @@
-# Tutorial on maximal monotone predictors for non-stationary (I(1)) time series
-
-# Previous tutorials 1-4 are based on Wildi, M. (2024) https://doi.org/10.1007/s41549-024-00097-5
-# This (new) tutorial 5 is based on Wildi (2025). The new paper proposes: 
-#   -Technical/formal background to Wildi (2024)
-#   -Introduction of a new SSA predictor for non-stationary time series
-
-# For non-stationary time series, an extension of SSA is proposed which tracks the target optimally in a MSE sense
-#   (predictor and target are cointegrated) and which controls the rate of zero-crossings of (stationary) FIRST DIFFERENCES of the predictor.
-# The predictor tracks the target as closely as possible subject to a holding-time constraint imposed to first differences.
-#   -In its dual form: the predictor generates the least number of zero-crossings in first differences for a given track accuracy (MSE).
-#   -Equivalently: the original predictor (in levels) is maximal monotone
-
-# Application/prediction problem:
-# -We compute a maximal monotone SSA nowcast for the two-sided (bi-infinite) HP(14400) applied to monthly US-INDPRO.
-# -We compute the MSE nowcast of HP based on an ARIMA(1,1,0)-model for INDPRO: benchmark 1
-#   -This MSE nowcast is much noisier than the classic HP concurrent (real-time) filter, widely used in business-cycle applications
-# -Therefore, we consider the classic concurrent one-sided HP (HP-C): benchmark 2
-#   -We replicate the holding-time (smoothness) of HP-C by SSA
-# -By optimization we then expect SSA to equal HP-C in terms of smoothness and to outperform HP-C in terms of MSE (tracking accuracy)
-# -Since SSA is much smoother than the MSE nowcast we are interested in finding out how much SSA looses in terms of MSE performances.
-#   -The SSA criterion minimizes MSE losses for given smoothness (holding-time)
-
[…]

### 2025-05-02 SSA Tutorials/tutorial_7.0_introduction_multivariate_SSA_MSSA.R

@@ -1,173 +1,0 @@
-# We here introduce to the M-SSA. see paper provide link???
-# Work in progress...
-
-# Make a clean-sheet, load packages and functions
-rm(list=ls())
-
-library(xts)
-# Load the library mFilter
-# HP and BK filters
-library(mFilter)
-# Plot for heat map of Trilemma
-library(ggplot2)
-library("gplots")
-
-# Load all relevant SSA-functions
-source(paste(getwd(),"/R/simple_sign_accuracy.r",sep=""))
-# Load tau-statistic: quantifies time-shift performances (lead/lag)
-source(paste(getwd(),"/R/Tau_statistic.r",sep=""))
-
-# Load signal extraction functions used for JBCY paper (relies on mFilter)
-source(paste(getwd(),"/R/HP_JBCY_functions.r",sep=""))
-
-##########################################################################################################
-# Example 1: we replicate SSA by M-SSA
-# We first compute a univariate SSA design
-# For this purpose we rely on example 1 of tutorial 2.1
-# This example relies on targeting a two-sided HP filter
-#--------------------------
-# a. Derivation
-# We use the R-package mFilter for computing HP 
-# Specify filter length: should be an odd number since otherwise the two-sided HP filter could not be adequately centered 
-L<-201
-# Should be an odd number
-if (L/2==as.integer(L/2))
-{
-  print("Filter length should be an odd number")
-  print("If L is even then HP cannot be adequately centered")
-  L<-L+1
-}  
-# Specify lambda: monthly design
-lambda_monthly<-14400
-par(mfrow=c(1,1))
-HP_obj<-HP_target_mse_modified_gap(L,lambda_monthly)
-# Bi-infinite two-sided (symmetric) HP
-hp_target<-HP_obj$target
-ts.plot(hp_target)
-# Concurrent gap: as applied to series in levels: this is a high pass filter
-hp_gap=HP_obj$hp_gap
-ts.plot(hp_gap)
-# Concurrent HP assuming I(2)-process 
-# This is the Classic concurrent or one-sided low pass HP, see e.g. McElroy (2006)
-hp_trend=HP_obj$hp_trend
-ts.plot(hp_trend)
-
[…]

### 2025-05-02 SSA Tutorials/tutorial_7.4_MSSA_component_predictor.R

@@ -1,1767 +1,0 @@
-# Tutorial 7.4
-# Main purposes:
-# 1. Forecasting BIP multiple quarters ahead
-#   -Derive a new M-SSA components predictor which addresses MSE forecast performances when tracking BIP
-#     -The original M-SSA predictor (without the attribute `components') proposed in tutorial 7.3 is 
-#         standardized: neither level nor scale were calibrated to track BIP 
-#     -The original M-SSA predictor was designed to track dynamic changes in the trend growth-rate of BIP (HP-BIP)
-#     -The original M-SSA predictor was based on aggregating equally the standardized M-SSA components 
-#   -We here propose a more refined design, the `M-SSA components predictor', whereby equal weighting 
-#       (of M-SSA components) is replaced by an additional (new) optimal weighting step.
-# 2. Interpretability: 
-#   -Exploit the M-SSA components to gauge forecasts by the proposed M-SSA predictor(s)
-#   -How trustworthy is a change of the predictor outlook?
-# 3. Explainability: 
-#   -Justify the design and construction steps underlying the proposed M-SSA predictor(s)
-#   -Determine which step(s) in the construction of the predictor(s) generate(s) forecast gains?
-
-# Proceeding:
-# -We rely on the design proposed in tutorial 7.3, exercise 1
-# -We then add a new optimal weighting step in the construction of the final predictor
-
-# The tutorial is organized into 6 exercises
-# -Exercise 1
-#   -M-SSA components: derive M-SSA components and replicate the original M-SSA predictor (tutorial 7.3)
-#   -Interpretability: rely on M-SSA components to gauge forecasts (reliability/trustworthiness)
-#   -Forecast BIP: new additional optimization step
-#   -Out-of-sample performance evaluation
-#     -Comparison vs. simple mean, direct forecast and original M-SSA predictor
-# -Exercise 2
-#   -Analysis of revisions of new (real-time out-of-sample) M-SSA components predictor
-# -Exercise 3
-#   -Application of the predictor to German BIP
-# -Exercise 4
-#   -Explainability: why does the M-SSA component predictor outperform specifically at multi-quarters 
-#     ahead forecast horizons?
-# -Exercise 5
[…]

### 2025-05-03 Sweave/BIP_predictor.Rnw
up

% -*- mode: noweb; noweb-default-code-mode: R-mode; -*-
\documentclass[a4paper]{article}

\title{BIP-Predictor}
\author{Marc Wildi\\
Zurich University of Applied Sciences (ZHAW)\\
8000 Zurich, Switzerland\\
marc.wildi@zhaw.ch}

\SweaveOpts{echo=FALSE}
\usepackage{a4wide}
\usepackage{amsmath}
\usepackage{amsfonts} 
\usepackage{amsthm}
\usepackage{ amssymb }
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{float}
\newtheorem{Proposition}{Proposition}
\newtheorem{Corollary}{Corollary}
\newtheorem{Theorem}{Theorem}
\DeclareMathOperator{\sign}{sign}
%- Makes the section title start with Appendix in the appendix environment
\newcommand{\Appendix}
{%\appendix
\def\thesection{Appendix~\Alph{section}}
%\def\thesubsection{\Alph{section}.\arabic{subsection}}
\def\thesubsection{A.\arabic{subsection}}
}

\begin{document}

\maketitle

\begin{abstract}
\noindent

\end{abstract}

~\\
~\\

%\tableofcontents

\section{Introduction}

\section{Data and Dependence}

<<label=init,results=hide>>=

rm(list=ls())

inst_pack<-rownames(installed.packages())
if (!"mFilter"%in%inst_pack)
  install.packages("mFilter")
if (!"xts"%in%inst_pack)
  install.packages("xts")
if (!"MTS"%in%inst_pack)
  install.packages("MTS")
if (!"sandwich"%in%inst_pack)
  install.packages("sandwich")
if (!"multDM"%in%inst_pack)
  install.packages("multDM")
if (!"fGarch"%in%inst_pack)
  install.packages("fGarch")
if (!"xtable"%in%inst_pack)
  install.packages("xtable")

# Load the required R-libraries
# Standard filter package
library(mFilter)
# Multivariate time series: VARMA model for macro indicators: used here for simulation purposes only
library(MTS)
# HAC estimate of standard deviations in the presence of autocorrelation and heteroscedasticity
library(sandwich)
# Extended time series
library(xts)
# Library for Diebold-Mariano test of equal forecast performance
library(multDM)
# GARCH model: for improving regression estimates
library(fGarch)
# Library for tables
library(xtable)

# Load the relevant M-SSA functionalities
# M-SSA functions
source(paste(getwd(),"/R/functions_MSSA.r",sep=""))
# Load signal extraction functions used for JBCY paper (relies on mFilter)
source(paste(getwd(),"/R/HP_JBCY_functions.r",sep=""))
# Utility functions for M-SSA, see tutorial 
source(paste(getwd(),"/R/M_SSA_utility_functions.r",sep=""))
# Set of performance metrics and tests of unequal predictability
#source(paste(getwd(),"/R/performance_statistics_functions.r",sep=""))

path.main<-getwd()

path.pgm<-paste(path.main,"/R/",sep="")
path.out<-paste(path.main,"/Latex/",sep="")
path.sweave<-paste(path.main,"/Sweave/",sep="")
path.data<-paste(path.main,"/Data/",sep="")
# Savd results from an empirical analysis of S&P500
path.result<-paste(path.main,"/Results/",sep="")
fig_size<-4

#------------------------------

recompute_results<-F

@

<<label=init,results=hide>>=

# Load and transform data
data_file_name<-c("Data_HWI_2025_02.csv","gdp_2025_02.csv")
# Quarterly data: BIP in the first data column
data_quarterly<-read.csv(paste(getwd(),"/Data/",data_file_name[2],sep=""))
BIP_original<-data_quarterly[,"BIP"]

load(file=paste(getwd(),"\\Data\\macro",sep=""))

select_vec_multi<-c("BIP","ip","ifo_c","ESI","spr_10y_3m")
x_mat<-data[,select_vec_multi] 
rownames(x_mat)<-rownames(data)
n<-dim(x_mat)[2]
# Number of observations
len<-dim(x_mat)[1]

date_to_fit<-"2008"
p<-1
q<-0
L<-31
# Publication lag
# -In contrast to M-SSA package (tutorial 7.4) we here assume a lag of one quarter only (instead of two)
# -To replicate the results in package we then have to impose a forecast excess of 5 to BIP (instead of 4)
#   -M-SSA estimation routine shifts HP-BIP by lag[1]+f_excess[1]+h
lag_vec<-c(1,rep(0,ncol(x_mat)-1))
# Forecast excesses: see M-SSA packages
f_excess<-c(5,rep(4,length(select_vec_multi)-1))

# Plot BIP and diff-log BIP
file = "./Figures/data.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
par(mfrow=c(1,1))

# Plot the data
# The real-time BIP (red) is lagging the target (black) by lag_vec[1] quarters (publication lag)
mplot<-x_mat
colo<-c("black",rainbow(ncol(data)-1))
main_title<-"Data: standardized diff-log, Pandemic outliers trimmed"
plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],lwd=1,ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
mtext(colnames(mplot)[1],col=colo[1],line=-1)
for (i in 1:ncol(mplot))
{
  lines(mplot[,i],col=colo[i],lwd=1,lty=1)
  mtext(colnames(mplot)[i],col=colo[i],line=-i)
}
abline(h=0)
axis(1,at=c(1,4*1:(nrow(mplot)/4)),labels=rownames(mplot)[c(1,4*1:(nrow(mplot)/4))])
axis(2)
box()
dev.off()

# Plot BIP and diff-log BIP
file = "./Figures/data_lags.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)

par(mfrow=c(1,2))
# Plot the data
# The real-time BIP (red) is lagging the target (black) by lag_vec[1] quarters (publication lag)
mplot<-x_mat[which(rownames(x_mat)>=2007&rownames(x_mat)<=2011),]
colo<-c("black",rainbow(ncol(data)-1))
main_title<-"Financial Crisis"
plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],lwd=1,ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
mtext(colnames(mplot)[1],col=colo[1],line=-1)
for (i in 1:ncol(mplot))
{
  lines(mplot[,i],col=colo[i],lwd=1,lty=1)
  mtext(colnames(mplot)[i],col=colo[i],line=-i)
}
abline(h=0)
axis(1,at=c(1,4*1:(nrow(mplot)/4)),labels=rownames(mplot)[c(1,4*1:(nrow(mplot)/4))])
axis(2)
box()
mplot<-x_mat[which(rownames(x_mat)>=2019&rownames(x_mat)<=2022),]
colo<-c("black",rainbow(ncol(data)-1))
main_title<-"Pandemic"
plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],lwd=1,ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
mtext(colnames(mplot)[1],col=colo[1],line=-1)
for (i in 1:ncol(mplot))
{
  lines(mplot[,i],col=colo[i],lwd=1,lty=1)
  mtext(colnames(mplot)[i],col=colo[i],line=-i)
}
abline(h=0)
axis(1,at=c(1,4*1:(nrow(mplot)/4)),labels=rownames(mplot)[c(1,4*1:(nrow(mplot)/4))])
axis(2)
box()

dev.off()

acf_mat<-x_mat

# Plot Cross correlation: with Pandemic
file = "./Figures/CCF.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
acf(x_mat)
dev.off()

x_mat_wc<-x_mat[which(rownames(x_mat)<2020|rownames(x_mat)>2021),]

# CCF without Pandemic
file = "./Figures/CCF_wc.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
acf(x_mat[which(rownames(acf_mat)<2020),])
dev.off()

# Without Covid
# Full sample

if (recompute_results)
{
  data_fit<-x_mat

  set.seed(12)
  V_obj<-VARMA(data_fit,p=p,q=q)
  threshold<-1.5
  V_obj<-refVARMA(V_obj, thres = threshold)
  save(V_obj,file=paste(path.result,"/VAR",sep=""))
# Without Pandemic  
  data_fit<-x_mat[which(rownames(x_mat)<date_to_fit),]
  set.seed(12)
  V_obj_wc<-VARMA(data_fit,p=p,q=q)
  threshold<-1.5
  V_obj_wc<-refVARMA(V_obj_wc, thres = threshold)
  save(V_obj_wc,file=paste(path.result,"/VAR_wc",sep=""))

} else
{
    load(file=paste(path.result,"/VAR",sep=""))
    load(file=paste(path.result,"/VAR_wc",sep=""))

}

# Diagnostics are OK
if (F)
  MTSdiag(V_obj)

# MA inversion
# Full data-set
data_fit<-x_mat
# Sigma
Sigma<-V_obj$Sigma
V_obj$Phi
V_obj$Theta
n<-dim(Sigma)[1]
# MA inversion
xi_psi<-PSIwgt(Phi = V_obj$Phi, Theta = V_obj$Theta, lag = L, plot = F, output = F)
xi_p<-xi_psi$psi.weight
# Transform Xi_p into Xi: first L entries, from left to right, are weights of first WN, next L entries are weights of second WN 
xi<-matrix(nrow=n,ncol=n*L)
for (i in 1:n)
{
  for (j in 1:L)
    xi[,(i-1)*L+j]<-xi_p[,i+(j-1)*n]
}

# Plot MA inversions

file = "./Figures/ma_inv_multi_ip.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
colo<-rainbow(ncol(x_mat))
par(mfrow=c(1,n))
for (i in 1:n)#i<-1
{
  select_vec_i<-select_vec_multi
  colo_i<-colo
  mplot<-xi[i,1:min(10,L)]

  for (j in 2:n)
  {
    mplot<-cbind(mplot,xi[i,(j-1)*L+1:min(10,L)])

  }
  ts.plot(mplot,col=colo_i,main=paste("MA inversion ",colnames(data_fit)[i],sep=""),xlab="lag")
  for (k in 1:length(select_vec_i))
    mtext(select_vec_i[k],line=-k,col=colo_i[k])
}

dev.off()

# Without Pandemic
# Sigma
Sigma<-V_obj_wc$Sigma
# MA inversion
xi_psi<-PSIwgt(Phi = V_obj_wc$Phi, Theta = V_obj_wc$Theta, lag = L, plot = F, output = F)
xi_p<-xi_psi$psi.weight
# Transform Xi_p into Xi: first L entries, from left to right, are weights of first WN, next L entries are weights of second WN 
xi<-matrix(nrow=n,ncol=n*L)
for (i in 1:n)
{
  for (j in 1:L)
    xi[,(i-1)*L+j]<-xi_p[,i+(j-1)*n]
}

# Plot MA inversions  
file = "./Figures/ma_inv_multi_ip_wc.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
colo<-rainbow(ncol(x_mat))
par(mfrow=c(1,n))
for (i in 1:n)#i<-1
{
  select_vec_i<-select_vec_multi
  colo_i<-colo
  mplot<-xi[i,1:min(10,L)]

  for (j in 2:n)
  {
    mplot<-cbind(mplot,xi[i,(j-1)*L+1:min(10,L)])

  }
  ts.plot(mplot,col=colo_i,main=paste("MA inversion ",colnames(data_fit)[i],sep=""),xlab="lag")
  for (k in 1:length(select_vec_i))
    mtext(select_vec_i[k],line=-k,col=colo_i[k])
}

dev.off()

@

\subsection{Data}

\begin{itemize}
\item Choice of indicators: BIP, ip, ESI, ifo, spread
\item Quarterly frequency (emphasize $h\geq 2$)
\item Data transformations: Diff-log, standardization (merely for ease of visual inspection), trim Covid ourliers
\item Revisions and publication lags:  BIP lagged by quarter and ip lagged by two months in Jan-2025 data set
\item Real-time data: all series are artificially aligned at end point. BIP and ip are advanced by their lags.
\end{itemize}

The transformed quarterly series are displayed in Fig. \ref{data}, while Fig. \ref{data_lags} offers a close-up view of the financial crisis and the pandemic. The latter figure illustrates that the real-time GDP and industrial production series are synchronized at dips and peaks (coincident), whereas the other indicators tend to lead in relative terms. This characteristic of the additional explanatory variables can be leveraged in a multivariate approach to enhance forecasting performance compared to univariate benchmarks.

<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
file = "./Figures/Data.pdf"
cat("\\begin{figure}[H]")
cat("\\begin{center}")
cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
cat("\\caption{Sample cross-correlation function (CCF): full data set.", sep = "")
cat("\\label{data}}", sep = "")
cat("\\end{center}")
cat("\\end{figure}")
@

<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
file = "./Figures/data_lags.pdf"
cat("\\begin{figure}[H]")
cat("\\begin{center}")
cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
cat("\\caption{Data: leads and lags during financial crisis (left) and Pandemic (right).", sep = "")
cat("\\label{data_lags}}", sep = "")
cat("\\end{center}")
cat("\\end{figure}")
@

\subsection{Dependence}

We analyze the dependence structure of the data and derive a simple VAR model to establish the multivariate filter. The sample cross-correlation function (CCF) is shown in Figs. \ref{CCF} (entire dataset) and \ref{CCF_wc} (data prior to the pandemic).  
<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
file = "./Figures/CCF.pdf"
cat("\\begin{figure}[H]")
cat("\\begin{center}")
cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
cat("\\caption{Sample cross-correlation function (CCF): full data set.", sep = "")
cat("\\label{CCF}}", sep = "")
cat("\\end{center}")
cat("\\end{figure}")
@
<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
file = "./Figures/CCF_wc.pdf"
cat("\\begin{figure}[H]")
cat("\\begin{center}")
cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
cat("\\caption{Sample cross-correlation function (CCF): without Pandemic.", sep = "")
cat("\\label{CCF_wc}}", sep = "")
cat("\\end{center}")
cat("\\end{figure}")
@
The CCF confirms the leads and lags illustrated in Fig. \ref{data_lags}, although the strength of the dependence is influenced by the singular readings during the pandemic\footnote{We discard the corresponding data when finalizing the predictor.}. We can now fit a VARMA model, and according to standard diagnostic tests (not shown here), a simple VAR(1) specification aligns with the data. The resulting impulse responses of the model are presented in Figs. \ref{impulse} (full data set) and \eqref{impulse_wc} (prior the pandemic).
<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
file = "./Figures/ma_inv_multi_ip.pdf"
cat("\\begin{figure}[H]")
cat("\\begin{center}")
cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
cat("\\caption{MA inversion of VAR(1): full data set.", sep = "")
cat("\\label{impulse}}", sep = "")
cat("\\end{center}")
cat("\\end{figure}")
@
<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
file = "./Figures/ma_inv_multi_ip_wc.pdf"
cat("\\begin{figure}[H]")
cat("\\begin{center}")
cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
cat("\\caption{MA inversion of VAR(1): without Pandemic.", sep = "")
cat("\\label{impulse_wc}}", sep = "")
cat("\\end{center}")
cat("\\end{figure}")
@
The model summarizes key features of the data, where lagging series (GDP and industrial production, when accounting for publication lags) depend on  indicators leading in relative terms. In particular, the impulse response of the spread (rightmost panel in the figure) suggests the relevance of a univariate model for this specific leading indicator.

\section{Direct Forecasts}\label{direct_forecast}

\subsection{`Classic' Direct Forecast}\label{cdf}

So-called direct forecasts are obtained by regressing the relevant indicators on forward-shifted BIP, accounting for the additional publication lag. For illustration of the concept, we here rely on all indicators and full sample information, excluding the pandemic. In-sample forecasts and the shifted BIP, which serves as the target, are presented in Figs. \ref{direct_wc} (full dataset without the pandemic) and \ref{direct_wc_financial_crisis} (financial crisis) for horizons up to three quarters ahead. 
<<label=init,results=hide>>=
# Plot of direct forecasts with Pandemic
par(mfrow=c(2,2))
# We can specify the selection of macro-indicators for the direct forecast 
select_direct_indicator<-c("ifo_c","ESI")
# Note: too complex designs (too many indicators) lead to overfitting and thus worse out-of-sample performances
# To illustrate the direct predictor consider the following example of a h-step ahead direct forecast:
for (h in 0:3)
{
  # Shift BIP forward by publication lag+forecast horizon
  forward_shifted_BIP<-c(x_mat[(1+lag_vec[1]+h):nrow(x_mat),"BIP"],rep(NA,h+lag_vec[1]))
  # Regress selected indicators on forward-shifted BIP
  lm_obj<-lm(forward_shifted_BIP~x_mat[,select_direct_indicator])
  # You will probably not find statistically significant regressors for h>2: BIP is a noisy series
  summary(lm_obj)
  # Technical note: 
  # -Residuals are subject to heteroscedasticity (crises) and autocorrelation
  # -Therefore classic OLS tests for statistical significance are biased
  # -We shall rely on HAC-adjusted p-values further down (R-package sandwich)

  # Compute the predictor: one can rely on the generic R-function predict or compute the predictor manually
  direct_forecast<-lm_obj$coef[1]+x_mat[,select_direct_indicator]%*%lm_obj$coef[2:(length(select_direct_indicator)+1)]
  # Note that this is a full-sample predictor (no out-of-sample span)

  # We can now plot target and direct forecast: for h>2 the predictor comes close to a flat line centered at zero

  # We can now plot target and direct forecast: for h>2 the predictor comes close to a flat line centered at zero
  mplot<-cbind(forward_shifted_BIP,direct_forecast)
  colnames(mplot)<-c("BIP","Direct forecast")
  colo<-c("black","red")
  main_title<-paste("BIP shifted by ",h," quarters",sep="")
  plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
  for (i in 1:ncol(mplot))
  {
    lines(mplot[,i],col=colo[i])
    mtext(colnames(mplot)[i],col=colo[i],line=-i)
  }
  abline(h=0)
  axis(1,at=c(1,12*1:(nrow(mplot)/12)),labels=rownames(mplot)[c(1,12*1:(nrow(mplot)/12))])
  axis(2)
  box()
}

# Plot of direct forecasts without Pandemic
file = "./Figures/direct_wc_all.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)

par(mfrow=c(2,2))
# We can specify the selection of macro-indicators for the direct forecast 
select_direct_indicator<-c("ifo_c","ESI")
select_direct_indicator<-select_vec_multi

# Note: too complex designs (too many indicators) lead to overfitting and thus worse out-of-sample performances
# To illustrate the direct predictor consider the following example of a h-step ahead direct forecast:
f_vec_unfiltered<-NULL
for (h in 0:3)
{
  # Shift BIP forward by publication lag+forecast horizon
  forward_shifted_BIP<-c(x_mat_wc[(1+lag_vec[1]+h):nrow(x_mat_wc),"BIP"],rep(NA,h+lag_vec[1]))
  # Regress selected indicators on forward-shifted BIP
  lm_obj<-lm(forward_shifted_BIP~x_mat_wc[,select_direct_indicator])
  # You will probably not find statistically significant regressors for h>2: BIP is a noisy series
  summary(lm_obj)
  sum_obj<-summary(lm_obj)
  f_vec_unfiltered<-c(f_vec_unfiltered,sum_obj$fstatistic["value"])

  # Technical note: 
  # -Residuals are subject to heteroscedasticity (crises) and autocorrelation
  # -Therefore classic OLS tests for statistical significance are biased
  # -We shall rely on HAC-adjusted p-values further down (R-package sandwich)

  # Compute the predictor: one can rely on the generic R-function predict or compute the predictor manually
  direct_forecast<-lm_obj$coef[1]+x_mat_wc[,select_direct_indicator]%*%lm_obj$coef[2:(length(select_direct_indicator)+1)]
  # Note that this is a full-sample predictor (no out-of-sample span)

  # We can now plot target and direct forecast: for h>2 the predictor comes close to a flat line centered at zero
  mplot<-cbind(forward_shifted_BIP,direct_forecast)
  colnames(mplot)<-c("BIP","Direct forecast")
  colo<-c("black","red")
  main_title<-paste("BIP shifted  by ",h," quarters",sep="")
  plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
  for (i in 1:ncol(mplot))
  {
    lines(mplot[,i],col=colo[i])
    mtext(colnames(mplot)[i],col=colo[i],line=-i)
  }
  abline(h=0)
  axis(1,at=c(1,12*1:(nrow(mplot)/12)),labels=rownames(mplot)[c(1,12*1:(nrow(mplot)/12))])
  axis(2)
  box()
}
dev.off()

# Plot of direct forecasts without Pandemic, only financial crisis
file = "./Figures/direct_wc_financial_crisis.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)

par(mfrow=c(2,2))
for (h in 0:3)
{
  # Shift BIP forward by publication lag+forecast horizon
  forward_shifted_BIP<-c(x_mat_wc[(1+lag_vec[1]+h):nrow(x_mat_wc),"BIP"],rep(NA,h+lag_vec[1]))
  # Regress selected indicators on forward-shifted BIP
  lm_obj<-lm(forward_shifted_BIP~x_mat_wc[,select_direct_indicator])
  # You will probably not find statistically significant regressors for h>2: BIP is a noisy series
  sum_obj<-summary(lm_obj)
  sum_obj$fstatistic  
  # Technical note: 
  # -Residuals are subject to heteroscedasticity (crises) and autocorrelation
  # -Therefore classic OLS tests for statistical significance are biased
  # -We shall rely on HAC-adjusted p-values further down (R-package sandwich)

  # Compute the predictor: one can rely on the generic R-function predict or compute the predictor manually
  direct_forecast<-lm_obj$coef[1]+x_mat_wc[,select_direct_indicator]%*%lm_obj$coef[2:(length(select_direct_indicator)+1)]
  # Note that this is a full-sample predictor (no out-of-sample span)

  # We can now plot target and direct forecast: for h>2 the predictor comes close to a flat line centered at zero
  mplot<-cbind(forward_shifted_BIP,direct_forecast)
  mplot<-mplot[which(rownames(mplot)>2008&rownames(mplot)<2011),]
  colnames(mplot)<-c("BIP","Direct forecast")
  colo<-c("black","red")
  main_title<-paste("BIP shifted forward by ",h," quarters",sep="")
  plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
  for (i in 1:ncol(mplot))
  {
    lines(mplot[,i],col=colo[i])
    mtext(colnames(mplot)[i],col=colo[i],line=-i)
  }
  abline(h=0)
  axis(1,at=1:nrow(mplot),labels=rownames(mplot))
  axis(2)
  box()
}
dev.off()
@
<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
file = "./Figures/direct_wc_all.pdf"
cat("\\begin{figure}[H]")
cat("\\begin{center}")
cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
cat("\\caption{Direct forecasts (red lines) based on all indicators and full sample information, without the pandemic: BIP (black line) shifted by zero one, two and three quarters.", sep = "")
cat("\\label{direct_wc}}", sep = "")
cat("\\end{center}")
cat("\\end{figure}")
@
<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
file = "./Figures/direct_wc_financial_crisis.pdf"
cat("\\begin{figure}[H]")
cat("\\begin{center}")
cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
cat("\\caption{Direct forecast  across the financial crisis.", sep = "")
cat("\\label{direct_wc_financial_crisis}}", sep = "")
cat("\\end{center}")
cat("\\end{figure}")
@
The results indicate that as $h$ increases, the peaks and dips of the direct forecasts become increasingly right-shifted (delayed) relative to the target. Additionally, the F-statistics precipitously decline from $F=\Sexpr{round(f_vec_unfiltered[3],2)}$, for $h=2$, to $F=\Sexpr{round(f_vec_unfiltered[4],2)}$, for $h=3$, see Table \eqref{f_stat}, suggesting that the predictor becomes statistically insignificant for larger forecast horizons exceeding two quarters (link to Heinisch and Scheufele).

\subsection{Filter: HP(160)}

As shown, the performance of direct forecasts sharply declines for larger forecast horizons. We conjecture that erratic short-term fluctuations -—unpredictable high-frequency noise -— overlay the data, thereby obscuring the effective 'signal' and making a direct regression more susceptible to overfitting. We therefore envisage to highlight the signal by applying a filter to dampen high-frequency noise. 
The so-called Hodrick-Prescott filter is a classic tool used in business cycle analysis: the filter is specified by a single smoothing parameter $\lambda$ and the value $\lambda=1600$ is specifically recommended for quarterly data. However, Phillips and Jin (2021) suggest that the HP(1600) filter removes relevant information due to excessive smoothing.  In our context, this oversmoothing issue would be further exacerbated when considering forecast horizons shorter than a year -—an interval inconsistent with the mean duration of up to several years of business cycles, as highlighted by the HP(1600). Consequently, we here select a more adaptive HP(160) 'target' filter. 
While this choice may appear somewhat arbitrary, an analysis of the filter in the frequency domain confirms a more suitable profile for the corresponding amplitude function, assigning greater weight to yearly components -—relevant in our prediction framework-— than the classic quarterly HP filter. This approach also addresses and controls for undesirable high-frequency noise. See Fig. \ref{hp_160}, which presents both two-sided and (classic) one-sided filters (top panels) along with the associated amplitude functions (bottom panels). We now posit the HP(160) filter as a means to achieve better forecasting performance at larger horizons, noting that a comprehensive technical analysis of the effect of $\lambda$ on the resulting BIP predictor can be conducted using the M-SSA package. 
<<label=init,results=hide>>=
# Target filter: lambda_HP is the single most important hyperparameter, see tutorial 7.1 for a discussion
lambda_HP<-160
# Filter length: nearly 8 years is fine for the selected lambda_HP (filter weights decay sufficiently fast)
#   The length should be an odd number (see tutorial 7.1)
L<-31
HP_obj<-HP_target_mse_modified_gap(L,lambda_HP)
# Classic concurrent (one-sided) HP filter

file = "./Figures/hp_160.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
par(mfrow=c(2,2))
hp_two<-c(HP_obj$hp_mse[L:2],HP_obj$hp_mse)
ts.plot(hp_two,main=paste("Two-sided HP(",lambda_HP,")",sep=""),xlab="",ylab="")
hp_c<-HP_obj$hp_trend
ts.plot(hp_c,main="One-sided (concurrent) HP-C",xlab="",ylab="")

# Analyze filter in frequency-domain (amplitude function)
# Specify the number of equidistant frequency ordinates in [0,pi]
K<-600
# Compute transfer, amplitude and shift functions (shift=phase divided by frequency)
amp_obj_hp_c<-amp_shift_func(K,hp_two,F)

# Plot amplitude function
mplot<-matrix(amp_obj_hp_c$amp,ncol=1)
colnames(mplot)<-paste("Concurrent HP, lambda=",lambda_HP,sep="")
colo<-c("blue",rainbow(ncol(mplot)))
plot(mplot[,1],type="l",axes=F,xlab="Frequency",ylab="",main=paste("Amplitude HP(",lambda_HP,")",sep=""),ylim=c(min(mplot),max(mplot)),col=colo[1])
if (ncol(mplot)>1)
{
  lines(mplot[,2],col=colo[2])
  abline(v=which(mplot[,1]==max(mplot[,1])),col=colo[1])
  mtext(colnames(mplot)[1],line=-1,col=colo[1])

  for (i in 2:ncol(mplot))
  {
    lines(mplot[,i],col=colo[i])
    mtext(colnames(mplot)[i],col=colo[i],line=-i)
  }
}
axis(1,at=1+0:4*K/4,labels=expression(0, pi/4, 2*pi/4,3*pi/4,pi))
axis(2)
box()
K<-600
# Compute transfer, amplitude and shift functions (shift=phase divided by frequency)
amp_obj_hp_c<-amp_shift_func(K,hp_c,F)

# Plot amplitude function
mplot<-matrix(amp_obj_hp_c$amp,ncol=1)
colnames(mplot)<-paste("Concurrent HP, lambda=",lambda_HP,sep="")
colo<-c("blue",rainbow(ncol(mplot)))
plot(mplot[,1],type="l",axes=F,xlab="Frequency",ylab="",main="Amplitude HP-C",ylim=c(0,max(mplot)),col=colo[1])
if (ncol(mplot)>1)
{
  lines(mplot[,2],col=colo[2])
  abline(v=which(mplot[,1]==max(mplot[,1])),col=colo[1])
  mtext(colnames(mplot)[1],line=-1,col=colo[1])

  for (i in 2:ncol(mplot))
  {
    lines(mplot[,i],col=colo[i])
    mtext(colnames(mplot)[i],col=colo[i],line=-i)
  }
}
abline(h=0)
axis(1,at=1+0:4*K/4,labels=expression(0, pi/4, 2*pi/4,3*pi/4,pi))
axis(2)
box()
dev.off()

@
<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
file = "./Figures/hp_160.pdf"
cat("\\begin{figure}[H]")
cat("\\begin{center}")
cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
cat("\\caption{Two-sided HP-160) (top left) and one-sided (concurrent) HP-C filters (top right) with corresponding amplitude functions (bottom).", sep = "")
cat("\\label{hp_160}}", sep = "")
cat("\\end{center}")
cat("\\end{figure}")
@

\subsection{Direct HP Forecast}\label{hpdf}

Extending the `classic' direct forecast design, we here consider filtered indicators, based on the one-sided HP(160), denoted as HP-C: the filtered indicators are explanatory variables for the regressions on forward-shifted BIP\footnote{Katja and Simon: for each forecast horizon $h=0,1,2,3$ I'm using corresponding forecasts of the real-time filter output, assuming the data to be white noise (assuming an ARMA model does not lead to significantly different results). I think we might ignore these technical details in the paper.}. As in the previous section, we rely on all indicators as well as the full sample estimate for illustration, see Figs.\eqref{direct_hp_forecasts} (all observations with exclusion of the pandemic) and \eqref{direct_hp_forecasts_financial_crisis} (financial crisis). A comparison with the classic direct forecasts in Fig.\eqref{direct_wc_financial_crisis} indicates that the new predictors are slightly less retarded (`faster'). 
<<label=init,results=hide>>=

# Filter the indicators: apply HP-C
hp_c_mat<-NULL
for (i in 1:ncol(x_mat))
{
  hp_c_mat<-cbind(hp_c_mat,filter(x_mat[,i],hp_c,side=1))
}
colnames(hp_c_mat)<-colnames(x_mat)
rownames(hp_c_mat)<-rownames(x_mat)

# Plot
mplot<-hp_c_mat
par(mfrow=c(1,1))
colo<-c(rainbow(ncol(mplot)))
main_title<-paste("Concurrent HP(",lambda_HP,") applied to indicators",sep="")
plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
mtext(colnames(mplot)[1],col=colo[1],line=-1)
for (j in 1:ncol(mplot))
{
  lines(mplot[,j],col=colo[j],lwd=1,lty=1)
  mtext(colnames(mplot)[j],col=colo[j],line=-j)
}
abline(h=0)
axis(1,at=c(1,12*1:(nrow(mplot)/12)),labels=rownames(mplot)[c(1,12*1:(nrow(mplot)/12))])
axis(2)
box()

# Compute filter outputs for forecast horizons 0:6, assuming WN data
# Forecast horizons
h_vec<-0:6
hp_c_array_wc<-array(dim=c(ncol(x_mat),nrow(x_mat_wc),length(h_vec)))
for (j in 1:length(h_vec))
{
  for (i in 1:ncol(x_mat))
  {
# For forecast horizon h_vec[j], the first h_vec[j] filter coefficients are skipped (zeroes are appended at 
#     the end). 
#   -This simple rule is optimal if the data is (close to) WN (white noise).
#   -Log-returns of the indicators are fairly close to WN (one can inspect the ACFs) 
    hp_c_forecast<-c(hp_c[(h_vec[j]+1):L],rep(0,h_vec[j]))
    hp_c_array_wc[i,,j]<-filter(x_mat_wc[,i],hp_c_forecast,side=1)
  }
}
dimnames(hp_c_array_wc)[[1]]<-colnames(x_mat_wc)
dimnames(hp_c_array_wc)[[2]]<-rownames(x_mat_wc)
dimnames(hp_c_array_wc)[[3]]<-paste("h=",h_vec,sep="")

# Same but with pandemic
h_vec<-0:6
hp_c_array<-array(dim=c(ncol(x_mat),nrow(x_mat),length(h_vec)))
for (j in 1:length(h_vec))
{
  for (i in 1:ncol(x_mat))
  {
# For forecast horizon h_vec[j], the first h_vec[j] filter coefficients are skipped (zeroes are appended at 
#     the end). 
#   -This simple rule is optimal if the data is (close to) WN (white noise).
#   -Log-returns of the indicators are fairly close to WN (one can inspect the ACFs) 
    hp_c_forecast<-c(hp_c[(h_vec[j]+1):L],rep(0,h_vec[j]))
    hp_c_array[i,,j]<-filter(x_mat[,i],hp_c_forecast,side=1)
  }
}
dimnames(hp_c_array)[[1]]<-colnames(x_mat)
dimnames(hp_c_array)[[2]]<-rownames(x_mat)
dimnames(hp_c_array)[[3]]<-paste("h=",h_vec,sep="")

# Compute and plot direct HP forecasts: full sample without pandemic
file = "./Figures/direct_hp_forecasts.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)

par(mfrow=c(2,2))
f_vec_hp<-t_mat_hp<-NULL
for (h in 0:3)
{
  # Shift BIP forward by publication lag+forecast horizon
  forward_shifted_BIP<-c(x_mat_wc[(lag_vec[1]+h):nrow(x_mat_wc),"BIP"],rep(NA,h+lag_vec[1]-1))
  # Regress selected indicators on forward-shifted BIP
  length(forward_shifted_BIP)
  lm_obj<-lm(forward_shifted_BIP~t(hp_c_array_wc[,,h+1]))
  # You will probably not find statistically significant regressors for h>2: BIP is a noisy series
  sum_obj<-summary(lm_obj)
  sum_obj<-summary(lm_obj)
  f_vec_hp<-c(f_vec_hp,sum_obj$fstatistic["value"])
  t_mat_hp<-rbind(t_mat_hp,sum_obj$coef[-1,4])
  # Technical note: 
  # -Residuals are subject to heteroscedasticity (crises) and autocorrelation
  # -Therefore classic OLS tests for statistical significance are biased
  # -We shall rely on HAC-adjusted p-values further down (R-package sandwich)

  # Compute the predictor: one can rely on the generic R-function predict or compute the predictor manually
  direct_hp_forecast<-lm_obj$coef[1]+t(hp_c_array_wc[,,h+1])%*%lm_obj$coef[2:(ncol(t(hp_c_array_wc[,,h+1]))+1)]
  # Note that this is a full-sample predictor (no out-of-sample span)

  # We can now plot target and direct forecast: for h>2 the predictor comes close to a flat line centered at zero
  mplot<-cbind(forward_shifted_BIP,direct_hp_forecast)
  colnames(mplot)<-c("BIP","Direct forecast")
  colo<-c("black","red")
  main_title<-paste("BIP shifted forward by ",h," quarters",sep="")
  plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
  for (i in 1:ncol(mplot))
  {
    lines(mplot[,i],col=colo[i])
    mtext(colnames(mplot)[i],col=colo[i],line=-i)
  }
  abline(h=0)
  axis(1,at=1:nrow(mplot),labels=rownames(mplot))
  axis(2)
  box()
}
colnames(t_mat_hp)<-select_vec_multi
rownames(t_mat_hp)<-paste("h=",0:3)
dev.off()

# Compute and plot direct HP forecasts: financial crisis
file = "./Figures/direct_hp_forecasts_financial_crisis.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)

par(mfrow=c(2,2))

for (h in 0:3)
{
  # Shift BIP forward by publication lag+forecast horizon
  forward_shifted_BIP<-c(x_mat_wc[(lag_vec[1]+h):nrow(x_mat_wc),"BIP"],rep(NA,h+lag_vec[1]-1))
  # Regress selected indicators on forward-shifted BIP
  length(forward_shifted_BIP)
  lm_obj<-lm(forward_shifted_BIP~t(hp_c_array_wc[,,h+1]))
  # You will probably not find statistically significant regressors for h>2: BIP is a noisy series
  sum_obj<-summary(lm_obj)
  sum_obj$fstatistic  
  # Technical note: 
  # -Residuals are subject to heteroscedasticity (crises) and autocorrelation
  # -Therefore classic OLS tests for statistical significance are biased
  # -We shall rely on HAC-adjusted p-values further down (R-package sandwich)

  # Compute the predictor: one can rely on the generic R-function predict or compute the predictor manually
  direct_hp_forecast<-lm_obj$coef[1]+t(hp_c_array_wc[,,h+1])%*%lm_obj$coef[2:(ncol(t(hp_c_array_wc[,,h+1]))+1)]
  # Note that this is a full-sample predictor (no out-of-sample span)

  # We can now plot target and direct forecast: for h>2 the predictor comes close to a flat line centered at zero
  mplot<-cbind(forward_shifted_BIP,direct_hp_forecast)
  mplot<-mplot[which(rownames(mplot)>2008&rownames(mplot)<2011),]
  colnames(mplot)<-c("BIP","Direct HP forecast")
  colo<-c("black","red")
  main_title<-paste("BIP shifted forward by ",h," quarters",sep="")
  plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
  for (i in 1:ncol(mplot))
  {
    lines(mplot[,i],col=colo[i])
    mtext(colnames(mplot)[i],col=colo[i],line=-i)
  }
  abline(h=0)
  axis(1,at=1:nrow(mplot),labels=rownames(mplot))
  axis(2)
  box()
}

dev.off()
@
<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
file = "./Figures/direct_hp_forecasts.pdf"
cat("\\begin{figure}[H]")
cat("\\begin{center}")
cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
cat("\\caption{Direct HP forecasts: entire data set with exclusion of the pandemic.", sep = "")
cat("\\label{direct_hp_forecasts}}", sep = "")
cat("\\end{center}")
cat("\\end{figure}")
@
<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
file = "./Figures/direct_hp_forecasts_financial_crisis.pdf"
cat("\\begin{figure}[H]")
cat("\\begin{center}")
cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
cat("\\caption{Direct HP forecasts over the curse of the financial crisis.", sep = "")
cat("\\label{direct_hp_forecasts_financial_crisis}}", sep = "")
cat("\\end{center}")
cat("\\end{figure}")
@
Moreover, the F-statistics are larger than for the direct forecasts, see Table \eqref{f_stat}, but the predictor remains statistically insignificant at forecast horizons exceeding two quarters (even though the evaluation is conducted in-sample and based on all available indicators). 
<<label=ats_mba_2,echo=FALSE,results=tex>>=

mat_re<-rbind(f_vec_unfiltered,f_vec_hp)
rownames(mat_re)<-c("Direct forecast","Direct HP forecast")
colnames(mat_re)<-paste("h=",0:3)
mat1<-round(mat_re,3)
#latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
#center = "centering", file = "", floating = FALSE)
xtable(mat1, dec = 1,digits=rep(3,dim(mat_re)[2]+1),
paste("F-statistics of direct forecasts and direct HP forecasts for forecast horizons $h=0,1,2,3$: fully in-sample.  "),
label=paste("f_stat",sep=""),
center = "centering", file = "", floating = FALSE)
@
We now propose a more refined causal filter design aiming for statistical significance of the resulting predictor at longer forecast horizons, out-of-sample.

\section{Multivariate Causal Filter: M-SSA}

The application of the one-sided HP-C filter to the indicators has demonstrated improvements in forecast performance. However, the univariate nature of this filter limits its ability to leverage information from the cross-section, as supported by leading indicators. Additionally, the noise suppression capability of the one-sided filter is compromised, as illustrated by high-frequency leakage of the amplitude function in Fig. \eqref{hp_160} and further elaborated upon by Wildi (2025). To address these limitations, we propose an extension of the Smooth Sign Accuracy (SSA) framework introduced by Wildi (2025).

\subsection{Optimization criterion}

Let $\mathbf{X}_t$ (of dimension $t\times n$) denote a set of $n$ explanatory series $\mathbf{x}_{1t},...,\mathbf{x}_{nt}$, with observations $\mathbf{x}_{it}=(x_{i1},...,x_{it})'$. Let $\mathbf{z}_t=(z_{1},...,z_t)$  denote a target series, which typically lies outside the linear space spanned by $\mathbf{X}_t$  (and is unknown at time $t$). For simplicity, we assume stationarity of all time series involved. In this framework, $\mathbf{X}_t$ corresponds to a matrix of selected economic indicators with $n=\Sexpr{length(select_vec_multi)}$, and $z_{it}$   is the output of a two-sided HP(160) filter applied to the $i$-th indicator, $i=1,...,\Sexpr{length(select_vec_multi)}$. Since $z_{it}$  depends on future observations $x_{it+1},x_{it+2},...$, the prediction task involves `tracking' $z_{it}$  using the estimate $y^i_{t}=\sum_{j=1}^ny^i_{jt}$, where $y^i_{jt}=\mathbf{b}^i_j~'\mathbf{x}_{jt}=\sum_{k=0}^{L} b^i_{jk}x_{jk}$ are the outputs of a multivariate (causal or one-sided) filter $\mathbf{B}^i$, with columns $\mathbf{b}^i_{j}$ of fixed length $L$, $j=1,...,\Sexpr{length(select_vec_multi)}$, assigning weight to the last $L$ observations of the $j$-th indicator $\mathbf{x}_{jt}$ (the fixed-length assumption is merely used for ease of exposition).
For the sake of clarity, we now omit the upper-case index  $i$ for all variables, assuming $z_t=z^{i_0}_{t}$ for some fixed $i_0$.% Additionally, we assume a fixed filter length $L$, such that $\mathbf{B}^i_t=\mathbf{B}_t=\mathbf{B}$, which has dimension $L\times \Sexpr{length(select_vec_multi)}$. 
The task of `tracking'  a target can be formalized in different ways; here, we focus on the target correlation $\rho(z,y,h)$ between $y_t$ and $z_{t+h}$, where $h\geq 0$ denotes the forecast horizon (backcasting with $h<0$ is ignored in this context). To streamline the notation and avoid unnecessary complexity, we assume a fixed horizon, say $h=h_0$, so that we may drop the reference to $h$ in our notation. In the univariate case ($n=1$), Wildi (2025) proposes the Simple Sign Accuracy (SSA) as an optimization criterion for this task:
\begin{eqnarray}\label{critssa}
\left.\begin{array}{c}\rho(z,y)\to\max\\
\rho(y,1)=\rho_1\end{array}\right\}.
\end{eqnarray}
Objective function and constraint are expressed in terms of correlations, namely the target correlation $\rho(z,y)$ between target and predictor and the lag-one autocorrelation $\rho(y,1)$ of the predictor. The parameter $\rho_1$ governs the smoothness of the predictor: higher values of $\rho_1$ generally promote smoother trajectories of $y_t$. Formally, assuming that the process $x_t$ is centered (zero-mean), Wildi (2024) establishes a connection between $\rho(y,1)$ and the expected duration between consecutive zero-crossings (sign changes) of $y_t$, the so-called holding time, denoted by $ht(y)$:
\begin{eqnarray*}\label{ht}
ht(y)=\frac{\pi}{\arccos(\rho(y,1))}.
\end{eqnarray*}
Since the non-linear transformation involved is strictly monotonic, the SSA criterion can be reformulated as the optimization problem:
\begin{eqnarray}\label{critssaht}
\left.\begin{array}{c}\rho(z,y)\to\max\\
\frac{\arccos(\rho(y,1))}{\pi}=1/ht_1\end{array}\right\},
\end{eqnarray}
where the smoothness parameter $1/ht_1$ expresses the rate of zero-crossings of the predictor\footnote{Formally, the exact relationship between the zero-crossing rate and the lag-one autocorrelation of the predictor assumes Gaussian time series, but Wildi (2024) demonstrates robustness to deviations from this assumption.}. \\

We now briefly discuss some implications of criterion \eqref{critssaht}. First, the objective function is indifferent to an affine transformation of the predictor. This ambiguity can be resolved by assuming an arbitrary scale and level for $y_t$ (standardization). Alternatively, a mean-square error norm (MSE) may be substituted for the target correlation, as noted in Wildi (2025). In this case, the classic MSE predictor $y_{t,MSE}$ is obtained as a solution to the SSA criterion by insertion of $1/ht_1:=1/ht_{MSE}$ in the constraint, where $ht_{MSE}$ represents the holding time of $y_{t,MSE}$. %If the MSE predictor is noisy, which is often the case in applications, one can select $ht_1>ht_{MSE}$  in the constraint to address the rate of noisy crossings or false alarms. 
Second, the concept can be extended to a multivariate framework, denoted as M-SSA, by generalizing the correlation functions, as detailed in Wildi (2025b). Lastly, sign changes in trend growth, represented by zero-crossings of the target $z_t$, are indicative of relevant changes in the trajectory of the economy, for example at transitions between expansions and recessions. Unfortunately, classic predictors often generate excessively many false `noisy' crossings, due to high-frequency leakage as illustrated by the amplitude function in Fig. \eqref{hp_160} (bottom right panel) and further elaborated upon by Wildi (2025). The SSA criterion helps control this phenomenon. More precisely, Wildi (2025) derives an equivalent dual formulation of criterion \eqref{critssaht}, which distinguishes the SSA solution as exhibiting the smallest rate of zero-crossings of any (linear) predictor with the same target correlation. We now rely on the M-SSA criterion to derive a `smooth' predictor $y_t$ of $z_t$. %In a second stage, serving as a generalization and refinement of the univariate HP-C utilized in the direct HP forecast.

\subsection{M-MSE and M-SSA Nowcasts}

For illustration, we consider the target $z_t$ generated by the output of the two-sided HP(160) filter applied to BIP, which we denote by HP-BIP. We compare four nowcast methods: the univariate HP-C, the multivariate MSE predictor (denoted by M-MSE), and two M-SSA predictors based on different HT constraints. The first design enforces $ht_1=1.5ht_{MSE}$, aiming to produce approximately $33\%$ fewer zero-crossings than the M-MSE predictor over the long term. The second design sets $ht_1=ht_{HP-C}$, ensuring that its zero-crossing rate matches that of the univariate HP-C. These constraints are intended to systematically compare the effects of different levels of smoothness on prediction performance. The estimation of the underlying VAR(1) model employed for the M-MSE and M-SSA nowcasts is based on data observations up to January 2008. This approach ensures a lengthy out-of-sample period that encompasses significant events such as the financial crisis, the sovereign debt crisis, and the COVID-19 pandemic. Further details and background information are available in Wildi (2025b). All computations are performed utilizing the M-SSA package, as documented in Wildi (2025b).
<<label=init,results=hide>>=
# Target filter: lambda_HP is the single most important hyperparameter, see tutorial 7.1 for a discussion
lambda_HP<-160
# Filter length: nearly 8 years is fine for the selected lambda_HP (filter weights decay sufficiently fast)
#   The length should be an odd number (see tutorial 7.1)
L<-31
# In-sample span for VAR, i.e., M-SSA (the proposed design is quite insensitive to this specification because the VAR is parsimoniously parameterized
date_to_fit<-date_to_fit
# VARMA model orders: keep the model simple in particular for short/tight in-sample spans
p<-1
q<-0
# Holding-times (HT): controls smoothness of M-SSA (the following numbers are pasted from the original predictor)
#   -Increasing these numbers leads to predictors with less zero-crossings (smoother), see tutorial 7.1
ht_mssa_vec<-c(6.380160,  6.738270,   7.232453,   7.225927,   7.033768)
names(ht_mssa_vec)<-colnames(x_mat)
# Forecast horizons: M-SSA is optimized for each forecast horizon in h_vec 
h_vec<-0

if (recompute_results)
{
# Run the wrapper, see tutorial 7.2
#   -The function computes M-SSA for each forecast horizon h in h_vec
  mssa_indicator_obj<-compute_mssa_BIP_predictors_func(x_mat,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi)

# Replicate HT of HP-C
# Theoretical HT  
  ht_hp_c<-compute_holding_time_func(hp_c)$ht
# Sample HT: want to replicate sample HT in nowcast example and compare target correlations
  ht_hp_c<-8

  ht_mssa_vec<-c(ht_hp_c,  6.738270,   7.232453,   7.225927,   7.033768)

  mssa_indicator_HP_HT_obj<-compute_mssa_BIP_predictors_func(x_mat,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi)

  save(mssa_indicator_obj,file=paste(path.result,"/nowcast_obj",sep=""))
  save(mssa_indicator_HP_HT_obj,file=paste(path.result,"/nowcast_HT_HP_obj",sep=""))
} else
{
  load(file=paste(path.result,"/nowcast_obj",sep=""))
  load(file=paste(path.result,"/nowcast_HT_HP_obj",sep=""))
}

select<-"BIP"

dim(mssa_indicator_obj$bk_x_array)
# Dimensions of bk_x_array:
# 1. Which series do we target: depending on the series various settings may be used (forecast excess)
# 2. The corresponding M-SSA components
# 3. L*length(select_vec): filter weights attributed to the indicators
# 4. Forecast horizon h
bk<-matrix(mssa_indicator_obj$bk_x_array[select,select,,"h=0"],nrow=L)
gammak<-matrix(mssa_indicator_obj$gammak_x_mse[select,select,,"h=0"],nrow=L)
colnames(bk)<-colnames(gammak)<-select_vec_multi

# Target series: output of two-sided HP applied to BIP: 
#   -This is the target for which the original M-SSA predictor (tutorial 7.3) has been designed
#   -The target is forward-shifted by the forecast horizon (plus publication lag)
target_shifted_mat<-mssa_indicator_obj$target_shifted_mat
# Original M-SSA predictor, see tutorial 7.3
#   -One predictor available for each forecast horizon in h_vec
predictor_mssa_mat<-mssa_indicator_obj$predictor_mssa_mat
# M-SSA components, see tutorial 7.2
#   -This is a three dimensional array
#   -For each forecast horizon and for each indicator we obtain the M-SSA predictor when targeting 
#     the two-sided HP applied to this indicator, see exercise 1.1 below
mssa_array<-mssa_indicator_obj$mssa_array
# M-MSE components
# -Same as mssa_array but without HT imposed, i.e., classic multivariate mean-square error signal extraction
# -Forecast performances, see exercise 5 below 
mmse_array<-mssa_indicator_obj$mmse_array

# Select M-SSA output of design replicating HT of HP-C
mssa_array_HT_HP_C<-mssa_indicator_HP_HT_obj$mssa_array
bk_HP_HP_C<-matrix(mssa_indicator_HP_HT_obj$bk_x_array[select,select,,"h=0"],nrow=L)
colnames(bk_HP_HP_C)<-select_vec_multi

file = "./Figures/bk_gammak.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)

par(mfrow=c(1,3))
mplot<-gammak
colo<-rainbow(ncol(mplot))
ts.plot(mplot,col=colo,main="M-MSE")
for (i in 1:ncol(mplot))
  mtext(colnames(mplot)[i],col=colo[i],line=-i)
mplot<-bk
colo<-rainbow(ncol(mplot))
ts.plot(mplot,col=colo,main="M-SSA smaller HT")
for (i in 1:ncol(mplot))
  mtext(colnames(mplot)[i],col=colo[i],line=-i)
mplot<-bk_HP_HP_C
colo<-rainbow(ncol(mplot))
ts.plot(mplot,col=colo,main="M-SSA larger HT")
for (i in 1:ncol(mplot))
  mtext(colnames(mplot)[i],col=colo[i],line=-i)

dev.off()

dim(mmse_array)

mssa<-t(mssa_array[,,1])
mmse<-t(mmse_array[,,1])
mssa_HT_HP_C<-t(mssa_array_HT_HP_C[,,1])

file = "./Figures/mssa_msse_now.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)

hp_c_out<-filter(x_mat[,select],hp_c,side=1)
par(mfrow=c(1,1))
mplot<-scale(cbind(target_shifted_mat,hp_c_out,mssa[,select],mssa_HT_HP_C[,select],mmse[,select]))
dim(mplot)
colnames(mplot)<-c("Target: HP-BIP","HP-C","M-SSA smaller HT","M-SSA larger HT","M-MSE")
rownames(mplot)<-rownames(x_mat)
colo<-c("black","violet","blue","orange","green")
main_title<-"Nowcasting HP-BIP"
plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],lwd=1,ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
mtext(colnames(mplot)[1],col=colo[1],line=-1)
for (i in 1:ncol(mplot))
{
  lines(mplot[,i],col=colo[i],lwd=1,lty=1)
  mtext(colnames(mplot)[i],col=colo[i],line=-i)
}
abline(h=0)
abline(v=which(rownames(mplot)>date_to_fit)[1],lty=2,lwd=2)
axis(1,at=c(1,4*1:(nrow(mplot)/4)),labels=rownames(mplot)[c(1,4*1:(nrow(mplot)/4))])
axis(2)
box()

dev.off()

mat_cor_ht<-rbind(c(cor(na.exclude(mplot[,c("Target: HP-BIP","HP-C")]))[1,2],
cor(na.exclude(mplot[,c("Target: HP-BIP","M-SSA smaller HT")]))[1,2],
cor(na.exclude(mplot[,c("Target: HP-BIP","M-SSA larger HT")]))[1,2],
cor(na.exclude(mplot[,c("Target: HP-BIP","M-MSE")]))[1,2]),
c(compute_empirical_ht_func(na.exclude(hp_c_array[select,,"h=0"]))$empirical_ht,
compute_empirical_ht_func(na.exclude(mssa[,select]))$empirical_ht,
compute_empirical_ht_func(na.exclude(mssa_HT_HP_C[,select]))$empirical_ht,
compute_empirical_ht_func(na.exclude(mmse[,select]))$empirical_ht))

colnames(mat_cor_ht)<-c("HP-C","M-SSA smaller HT","M-SSA larger HT","M-MSE")
rownames(mat_cor_ht)<-c("Target correlation","HT")

mat_cor_ht

@
Figure \ref{bk_gammak} illustrates the coefficients of the multivariate filter designs, where increased smoothing achieved through the M-SSA method manifests as a distinct pattern in the filter weights, notably influencing the rate of decay across increasing lags. The resulting filtered series, presented in Figure \ref{mssa_msse_now}, are standardized to enable direct comparisons. The two M-SSA predictors adhere to the specified constraints: one exhibiting approximately one-third fewer zero-crossings than the M-MSE, and the other matching the zero-crossing rate of the HP-C, confirming the effectiveness of the smoothness control imposed during the design. This is further corroborated by Table \ref{corhtnow}, which reports empirical holding times and target correlations.  In this example, both the univariate design and the M-SSA method, when configured to replicate its holding time, demonstrate the highest level of smoothness, as indicated by the maximal duration between crossings. However, the HP-C also attains the lowest target correlation among the compared approaches. More generally, our example illustrates the prediction dilemma underlying the (M-)SSA criterion, highlighting the inherent trade-off between smoothness (zero-crossing rate) and predictive accuracy (target correlation).
<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
file = "./Figures/bk_gammak.pdf"
cat("\\begin{figure}[H]")
cat("\\begin{center}")
cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
cat("\\caption{Multivariate M-MSE (left) and M-SSA  (middle:smaller HT, right: larger HT) nowcast filters.", sep = "")
cat("\\label{bk_gammak}}", sep = "")
cat("\\end{center}")
cat("\\end{figure}")
@
<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
file = "./Figures/mssa_msse_now.pdf"
cat("\\begin{figure}[H]")
cat("\\begin{center}")
cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
cat("\\caption{Nowcasting: two-sided HP(160) applied to BIP (target, black), univariate HP-C (violet), M-SSA smaller HT (blue), M-SSA larger HT (orange) and M-MSE (green). The dashed vertical line delimits in- and out-of-sample spans. The first L=31 observations are lost due to initialization and the two-sided filter does not extend to the sample end.", sep = "")
cat("\\label{mssa_msse_now}}", sep = "")
cat("\\end{center}")
cat("\\end{figure}")
@
<<label=ats_mba_2,echo=FALSE,results=tex>>=
mat1<-round(mat_cor_ht,3)
#latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
#center = "centering", file = "", floating = FALSE)
xtable(mat1, dec = 1,digits=rep(2,dim(mat1)[2]+1),
paste("Target correlations and HTs of HP-C, M-SSA and M-MSE nowcasts.  "),
label=paste("corhtnow",sep=""),
center = "centering", file = "", floating = FALSE)
@
The results validate that the smoothing constraint incorporated into the multivariate extension of Criterion \eqref{critssa} effectively manages the zero-crossings of the predictor: sample estimates of the holding times (HTs) align with the expected values derived from Equation \eqref{ht}\footnote{Effective convergence of sample estimates to their theoretical expected values —assuming the true data-generating process is known— can be validated using the M-SSA package. This validation is typically performed through analyses of very long samples of simulated (multivariate) data, ensuring the consistency of the estimation procedures under idealized conditions.}. An increase of approximately $50\%$ in the HT relative to the M-MSE predictor has only a modest effect on the target correlation, suggesting that this tradeoff could be beneficial in this application. Conversely, matching the HT of the HP-C predictor through M-SSA results in an increase in the target correlation, demonstrating the impact of zero-crossing control in the multivariate design on predictive performance. Going forward, we now omit the M-MSE, as it is a special case of the more general M-SSA framework.

\subsection{M-SSA Forecasts}

The previously described nowcast example can be systematically extended to generate forecasts of the target variable. This extension utilizes the same empirical framework, incorporating an assumed $50\%$ increase in the HT over the M-MSE, and considers forward shifts of $h=1,...,6$  quarters of HP-BIP (note that M-SSA does not target BIP explicitly). %This approach facilitates the assessment of forecast performance across multiple horizons within the established methodological context. 
The influence of the forecast horizon on the M-SSA design is depicted in Fig. \eqref{bk_h}, which contrasts nowcast and one-year-ahead forecast: as $h$ increases (right panel), the scaling factor diminishes (zero-shrinkage), attributable to increased forecast uncertainty, and there is a relative amplification of the influence of the relevant indicators, leading BIP in relative terms.Additionally, a phase shift associated with the filter applied to BIP is observed (red line, right panel), reflecting the cyclical characteristics inherent to the HP filter. While this phase effect might cause an effective sign change in the BIP filter output as $h$ increases—something to be avoided in this context—the low-pass filters assigned to the additional 'leading' indicators (green and violet lines, right panel) continue to effectively track the low-frequency components of HP-BIP. The combination of the phase effect, applied to BIP, and level-tracking through these additional indicators enables more refined and effective tracking of the target compared to univariate forecasts, which rely solely on the phase-effect of the filter to look ahead of time.\\

M-SSA predictors are compared to the traditional univariate HP-C in Fig. \eqref{multivar_vs_univar} across the periods of the financial crisis (top panels) and the COVID-19 pandemic (bottom panels), for forecast horizons $h=0,...,6$. The nowcast from the previous section is included for reference. All series are standardized to facilitate visual comparison. The primary distinction of the M-SSA approach is the progressive left-shift of the predictor as a function of $h$; this shift becomes more pronounced with increasing $h$ and appears to be consistent across all levels, including the peaks and troughs of the series. Conversely, the positions of peaks and troughs in the HP-C series appear to remain largely unaffected by variations in $h$. The more pronounced and systematic left-shift observed in the M-SSA forecasts can be partly attributed to the increasing weight assigned to the additional indicators leading BIP, which dominate the BIP component at larger forecast horizons. %In comparison, the effect of phase shifts—arising from the cyclical nature of the HP-filter and manifesting in the HP-C forecasts (shown in the left panels)—seems less systematic. 
<<label=init,results=hide>>=
# Target filter: lambda_HP is the single most important hyperparameter, see tutorial 7.1 for a discussion
lambda_HP<-160
# Filter length: nearly 8 years is fine for the selected lambda_HP (filter weights decay sufficiently fast)
#   The length should be an odd number (see tutorial 7.1)
L<-31
# In-sample span for VAR, i.e., M-SSA (the proposed design is quite insensitive to this specification because the VAR is parsimoniously parameterized
date_to_fit<-date_to_fit
# VARMA model orders: keep the model simple in particular for short/tight in-sample spans
p<-1
q<-0
# Holding-times (HT): controls smoothness of M-SSA (the following numbers are pasted from the original predictor)
#   -Increasing these numbers leads to predictors with less zero-crossings (smoother), see tutorial 7.1
ht_mssa_vec<-c(6.380160,  6.738270,   7.232453,   7.225927,   7.033768)
names(ht_mssa_vec)<-colnames(x_mat)
# Forecast horizons: M-SSA is optimized for each forecast horizon in h_vec 
h_vec<-0:6

if (recompute_results)
{
# Run the wrapper, see tutorial 7.2
#   -The function computes M-SSA for each forecast horizon h in h_vec
  mssa_indicator_obj<-compute_mssa_BIP_predictors_func(x_mat,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi)
  save(mssa_indicator_obj,file=paste(path.result,"/mssa_indicator_obj",sep=""))
} else
{
  load(file=paste(path.result,"/mssa_indicator_obj",sep=""))
}
# Target series: output of two-sided HP applied to BIP: 
#   -This is the target for which the original M-SSA predictor (tutorial 7.3) has been designed
#   -The target is forward-shifted by the forecast horizon (plus publication lag)
target_shifted_mat<-mssa_indicator_obj$target_shifted_mat
# Original M-SSA predictor, see tutorial 7.3
#   -One predictor available for each forecast horizon in h_vec
predictor_mssa_mat<-mssa_indicator_obj$predictor_mssa_mat
# M-SSA components, see tutorial 7.2
#   -This is a three dimensional array
#   -For each forecast horizon and for each indicator we obtain the M-SSA predictor when targeting 
#     the two-sided HP applied to this indicator, see exercise 1.1 below
mssa_array<-mssa_indicator_obj$mssa_array

# M-SSA outputs targeting HP-BIP at forecasts horizons h=0,...,6
mssa_array["BIP",,]

dim(mssa_indicator_obj$bk_x_array)
# Dimensions of bk_x_array:
# 1. Which series do we target: depending on the series various settings may be used (forecast excess)
# 2. The corresponding M-SSA components
# 3. L*length(select_vec): filter weights attributed to the indicators for each target
# 4. Forecast horizon h
bk<-matrix(mssa_indicator_obj$bk_x_array[select,select,,"h=0"],nrow=L)
gammak<-matrix(mssa_indicator_obj$gammak_x_mse[select,select,,"h=0"],nrow=L)
colnames(bk)<-colnames(gammak)<-select_vec_multi

file = "./Figures/bk_h.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)

par(mfrow=c(1,2))
mplot<-bk
colo<-rainbow(ncol(mplot))
ts.plot(mplot,col=colo,main="M-SSA: h=0")
for (i in 1:ncol(mplot))
  mtext(colnames(mplot)[i],col=colo[i],line=-i)
bk<-matrix(mssa_indicator_obj$bk_x_array[select,select,,"h=4"],nrow=L)
colnames(bk)<-colnames(gammak)<-select_vec_multi

mplot<-bk
colo<-rainbow(ncol(mplot))
ts.plot(mplot,col=colo,main="M-SSA: h=4")
for (i in 1:ncol(mplot))
  mtext(colnames(mplot)[i],col=colo[i],line=-i)

dev.off()

file = "./Figures/multivar_vs_univar.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)

# Select BIP
i<-1
# HP-C Financial crisis
mplot<-scale(hp_c_array[i,,])
colnames(mplot)<-paste("HP-",colnames(x_mat)[i],": h=",h_vec,sep="")
par(mfrow=c(2,2))
colo<-c(rainbow(ncol(mplot)))
mplot<-mplot[which(rownames(mplot)>2007&rownames(mplot)<2011),]
main_title<-paste("HP-C targeting HP-",colnames(x_mat)[i],sep="")
main_title<-"HP-C: financial crisis"
plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
mtext(colnames(mplot)[1],col=colo[1],line=-1)
for (j in 1:ncol(mplot))
{
  lines(mplot[,j],col=colo[j],lwd=1,lty=1)
  mtext(colnames(mplot)[j],col=colo[j],line=-j)
}
abline(h=0)
axis(1,at=1:nrow(mplot),labels=rownames(mplot))
axis(2)
box()

# HP-C Pandemic
mplot<-scale(hp_c_array[i,,])
colnames(mplot)<-paste("HP-",colnames(x_mat)[i],": h=",h_vec,sep="")
colo<-c(rainbow(ncol(mplot)))
mplot<-mplot[which(rownames(mplot)>2017&rownames(mplot)<2022),]
main_title<-paste("HP-C targeting HP-",colnames(x_mat)[i],sep="")
main_title<-"HP-C: pandemic"
plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
mtext(colnames(mplot)[1],col=colo[1],line=-1)
for (j in 1:ncol(mplot))
{
  lines(mplot[,j],col=colo[j],lwd=1,lty=1)
  mtext(colnames(mplot)[j],col=colo[j],line=-j)
}
abline(h=0)
axis(1,at=1:nrow(mplot),labels=rownames(mplot))
axis(2)
box()

# M-SSA financial crisis 
mplot<-scale(mssa_array[i,,])
colnames(mplot)<-paste("HP-",colnames(x_mat)[i],": h=",h_vec,sep="")
mplot<-mplot[which(rownames(mplot)>2007&rownames(mplot)<2011),]

colo<-c(rainbow(ncol(mplot)))
main_title<-paste("M-SSA targeting HP-",colnames(x_mat)[i],sep="")
main_title<-"MSSA: financial crisis"
plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
mtext(colnames(mplot)[1],col=colo[1],line=-1)
for (j in 1:ncol(mplot))
{
  lines(mplot[,j],col=colo[j],lwd=1,lty=1)
  mtext(colnames(mplot)[j],col=colo[j],line=-j)
}
abline(h=0)
axis(1,at=1:nrow(mplot),labels=rownames(mplot))
axis(2)
box()

# M-SSA: pandemic 
mplot<-scale(mssa_array[i,,])
colnames(mplot)<-paste("HP-",colnames(x_mat)[i],": h=",h_vec,sep="")
mplot<-mplot[which(rownames(mplot)>2017&rownames(mplot)<2022),]

colo<-c(rainbow(ncol(mplot)))
main_title<-paste("M-SSA targeting HP-",colnames(x_mat)[i],sep="")
main_title<-"MSSA: pandemic"
plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
mtext(colnames(mplot)[1],col=colo[1],line=-1)
for (j in 1:ncol(mplot))
{
  lines(mplot[,j],col=colo[j],lwd=1,lty=1)
  mtext(colnames(mplot)[j],col=colo[j],line=-j)
}
abline(h=0)
axis(1,at=1:nrow(mplot),labels=rownames(mplot))
axis(2)
box()

dev.off()

file = "./Figures/mssa_predictor.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)

par(mfrow=c(1,1))
mplot<-cbind(x_mat[,"BIP"],predictor_mssa_mat)
rownames(mplot)<-rownames(x_mat)
colo<-c("black",rainbow(ncol(mplot)-1))
colnames(mplot)[1]<-"BIP"
plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
mtext(colnames(mplot)[1],col=colo[1],line=-1)
for (j in 1:ncol(mplot))
{
  lines(mplot[,j],col=colo[j],lwd=1,lty=1)
  mtext(colnames(mplot)[j],col=colo[j],line=-j)
}
abline(h=0)
axis(1,at=1:nrow(mplot),labels=rownames(mplot))
axis(2)
box()

dev.off()

file = "./Figures/mssa_predictor_interpretability.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
j_now<-1
# Plot M-SSA nowcast and components
par(mfrow=c(1,1))
# Scale the data 
mplot<-scale(cbind(predictor_mssa_mat[,j_now],scale(t(mssa_array[,,j_now]))))
rownames(mplot)<-rownames(x_mat)
colnames(mplot)<-c(paste("M-SSA predictor optimized for h=",h_vec[j_now],sep=""),
                   paste("Subseries ",select_vec_multi,sep=""))
colo<-c("blue",rainbow(length(select_vec_multi)))
main_title<-c(paste("M-SSA predictor for h=",h_vec[j_now]," (solid blue) and sub-series (dashed lines)",sep=""),"In-sample span up to black vertical (dashed) line")
plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],lwd=2,ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
mtext(colnames(mplot)[1],col=colo[1],line=-1)
for (i in 1:ncol(mplot))
{
  lines(mplot[,i],col=colo[i],lwd=1,lty=2)
  mtext(colnames(mplot)[i],col=colo[i],line=-i)
}
abline(h=0)
abline(v=which(rownames(mplot)<=date_to_fit)[length(which(rownames(mplot)<=date_to_fit))],lwd=2,lty=2)
axis(1,at=c(1,4*1:(nrow(mplot)/4)),labels=rownames(mplot)[c(1,4*1:(nrow(mplot)/4))])
axis(2)
box()
dev.off()

# Compute performances of original M-SSA predictor: these will be used as a benchmark when evaluating 
#   the new design
# Select start of out-of-sample span (entire financial crisis is out-of-sample)
in_out_separator<-"2007"
# We can specify the selection of macro-indicators for the direct forecast, see tutorial 7.3
# Note: these results will not be used in this tutorial (but we need to specify a selection anyway)
select_direct_indicator<-c("ifo_c","ESI")
perf_obj<-compute_perf_func(x_mat,target_shifted_mat,predictor_mssa_mat,predictor_mmse_mat,in_out_separator,select_direct_indicator,h_vec)

# The above function generates a rich output, see tutorial 7.3
#   -But we need only one (important/representative) performance number for our comparisons further down
# HAC adjusted p-values of regression of original M-SSA predictor on forward-shifted BIP, see tutorial 7.3 (exercise 1.2.2)
p_value_HAC_BIP_oos=perf_obj$p_value_HAC_BIP_oos
# Regression on HP-BIP, without pandemic
p_value_HAC_BIP_oos_wc<-perf_obj$p_value_HAC_HP_BIP_oos_wc

p_value_HAC_BIP_oos_wc
# The p-values are empirical significance levels of the regressions of the (out-of-sample) M-SSA predictors
#   (optimized for forecast horizon h in h_vec: the columns of the matrix) on BIP shifted forward by shift 
#   plus publication lag (the rows of the matrix)

@
<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
file = "./Figures/bk_h.pdf"
cat("\\begin{figure}[H]")
cat("\\begin{center}")
cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
cat("\\caption{Multivariate M-SSA: $h=0$ (nowcast, left) and $h=4$ (one year forecast, right).", sep = "")
cat("\\label{bk_h}}", sep = "")
cat("\\end{center}")
cat("\\end{figure}")
@
<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
file = "./Figures/multivar_vs_univar.pdf"
cat("\\begin{figure}[H]")
cat("\\begin{center}")
cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
cat("\\caption{Now- and forecasts of HP-BIP for $h=0,...,6$: HP-C (left) vs. M-SSA (right) across the financial crisis (top) and the pandemic (bottom).", sep = "")
cat("\\label{multivar_vs_univar}}", sep = "")
cat("\\end{center}")
cat("\\end{figure}")
@

\section{M-SSA Component Predictor}

The targets of the multivariate filters are the smoothed outputs obtained from the acausal two-sided HP(160) filter applied to the five indicators. However, our objective is to develop predictors that emphasize the unfiltered BIP. To achieve this, we can utilize the  M-SSA outputs, referred to as M-SSA components.

\subsection{Standardized Predictor: Equal-Weighting}

Assuming that all M-SSA components contribute equally to BIP prediction, an initial step involves averaging the standardized components to form an equally-weighted composite measure. Subsequently, the predictor is derived by regressing this standardized aggregate on the (forward-shifted) BIP series. Fig. \eqref{mssa_predictor} presents the BIP series alongside the equally-weighted M-SSA predictor. All series are standardized to facilitate visual comparison and to highlight the dynamic shifts, particularly the left-shift characteristic of the predictors.\\

To enhance interpretability and assessment of the M-SSA predictor, we propose analyzing its individual components. For illustration, Fig. \eqref{mssa_predictor_interpretability} shows the nowcast at $h=0$. The solid blue line (the nowcast) approaches the zero line near the end of the sample, with a recent upward movement mainly supported by the leading spread component. In contrast, the ifo- and ESI-components remain near the zero line, while the ip- and BIP-components appear to be awaiting additional evidence and confirmation before signaling a definitive movement.  This type of assessment can be useful for evaluating the relevance of recent changes in the forecasted time series dynamics.
<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
file = "./Figures/mssa_predictor.pdf"
cat("\\begin{figure}[H]")
cat("\\begin{center}")
cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
cat("\\caption{BIP and equally-weighted M-SSA predictor: all series standardized.", sep = "")
cat("\\label{mssa_predictor}}", sep = "")
cat("\\end{center}")
cat("\\end{figure}")
@
<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
file = "./Figures/mssa_predictor_interpretability.pdf"
cat("\\begin{figure}[H]")
cat("\\begin{center}")
cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
cat("\\caption{Equally-weighted M-SSA predictor and M-SSA components: all series standardized.", sep = "")
cat("\\label{mssa_predictor_interpretability}}", sep = "")
cat("\\end{center}")
cat("\\end{figure}")
@

%???Eventually add p-values and/rRMSEs???

\subsection{Tracking BIP: Optimal Weighting}

We now utilize the M-SSA components as regressors on forward-shifted BIP, extending the previous (standardized) equally-weighted M-SSA predictor to incorporate optimal weighting of the components. While we keep the M-SSA fixed, as based on data up to January 2008, we implement an expanding window starting from January 2007 for deriving the component weights, which are up-dated on a quarterly basis (this contrasts with Section \eqref{direct_forecast} which was based on the full sample information). \\

There are various combinations of M-SSA components for tracking BIP. For illustration, we here select the single BIP component resulting from the M-SSA filter tracking HP-BIP. Table \eqref{p_val_hpbip_wc} reports HAC-adjusted p-values from regressions of the out-of-sample predictor on HP-BIP, indicating a strong link between the predictor and the target, out-of-sample. 
<<label=ats_mba_2,echo=FALSE,results=tex>>=

shift_vec<-0:5

mat1<-round(p_value_HAC_BIP_oos_wc,3)
rownames(mat1)<-paste("Shift=",shift_vec,sep="")
colnames(mat1)<-paste("h=",h_vec,sep="")
#latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
#center = "centering", file = "", floating = FALSE)
xtable(mat1, dec = 1,digits=rep(3,dim(mat1)[2]+1),
paste("HAC adjusted p-values of regressions of BIP (M-SSA) component optimized for forecast horizons $h=0,...,6$  (the columns) on HP-BIP shifted forward by shift=0,...,5 (the rows). Out-of-sample span starting in Jan-2007, without pandemic."),
label=paste("p_val_hpbip_wc",sep=""),
center = "centering", file = "", floating = FALSE)
@
Our choice—focusing on the single M-SSA BIP component—emphasizes simplicity and interpretability, as future BIP and HP-BIP are connected through their shared low-frequency component. Additionally, the M-SSA BIP component is the most important predictor in the regression of the M-SSA outputs on BIP.  Finally, revisions due to quarterly updates of the regression equations are minimized by this predictor, and the estimates are more stable and robust over time\footnote{Other combinations can be experimented with using the M-SSA package. Regression weights of more complex designs, involving multiple components, tend to be more difficult to interpret due to the occurrence of negative weights and potential sign changes of the regressors over time, which may indicate overfitting or non-stationarity.} (see Section \eqref{revisions}). Note that information from the other  indicators is incorporated into the M-SSA BIP component through the multivariate filter design.\\

This setup allows us to evaluate the out-of-sample performance of various predictor approaches, including: (i) the simple mean of BIP, (ii) the direct forecast method described in Section \eqref{cdf}, (iii) the direct HP forecasts from Section \eqref{hpdf}, based on the univariate HP-C, and (iv) the new M-SSA BIP component predictor. %For completeness, we also consider an M-MSE component predictor derived from outputs of the multivariate MSE filter.

\subsection{Out-of-Sample Forecast Performances}\label{oosp}

%Out-of-sample performance metrics include p-values derived from regressions of the out-of-sample predictors on forward-shifted BIP, adjusted for heteroscedasticity and autocorrelation using HAC estimators. Additionally, we report relative root-mean-square forecast errors (rRMSE) benchmarked against various reference models.\\

Tables \eqref{p_val_wc} and \eqref{rRMSE_mSSA_comp_mean_wc} present HAC-adjusted p-values and rRMSEs for the BIP component predictor. The p-values are obtained from a regression of the out-of-sample predictor on forward-shifted BIP (in contrast to Table \eqref{p_val_hpbip_wc}, which emphasized HP-BIP); hence, they do not account for potential bias or mis-scaling of the predictor. In contrast, the rRMSEs, evaluated against the mean benchmark, do incorporate eventual bias and mis-scaling effects. Additional results involving other predictor benchmarks are provided in the appendix, based on data sets with and without the pandemic. The results can be summarized as follows: targeting future HP-BIP (see Table \eqref{p_val_hpbip_wc}) is less challenging than predicting BIP (see Table \eqref{p_val_wc}), due to smoothness (autocorrelation) of the target when compared to BIP. The pandemic period generally obscures the relationships between predictors and BIP. When excluding the pandemic, the direct forecasts outperform the mean forecast up to two quarters ahead. The M-SSA component predictor, optimized for longer forecast horizons, outperforms the mean benchmark up to one year ahead. Additionally, the BIP component predictor also surpasses the direct forecasts for larger forecast horizons ($h>2$ quarters).Statistical significance of the BIP component predictor, optimized for larger forecast horizons, can be established for at least up to one year ahead. Predictors optimized for $h\geq 4$ generally outperform the other designs, irrespective of the forward-shift of BIP.

%???check whether rMSE is decreasing towards sample end (because initial estimate towards 2007 is unreliable)???

<<label=init,results=hide>>=
recompute_results<-F
sel_vec_pred<-c("BIP")

shift_vec<-0:5
if (recompute_results)
{
# Initialize performance matrices
  MSE_oos_mssa_comp_without_covid_mat<-MSE_oos_mssa_comp_mat<-p_mat_mssa<-p_mat_mssa_components<-p_mat_mssa_components_without_covid<-p_mat_direct<-rRMSE_mSSA_comp_direct<-rRMSE_mSSA_comp_mean<-rRMSE_mSSA_comp_direct_without_covid<-rRMSE_mSSA_comp_mean_without_covid<-rRMSE_mSSA_direct_mean_without_covid<-rRMSE_mSSA_direct_mean<-p_mat_direct_without_covid<-matrix(ncol=length(h_vec),nrow=length(h_vec)-1)
  final_components_preditor_array<-oos_components_preditor_array<-array(dim=c(length(shift_vec),length(h_vec),nrow(x_mat)))
  dimnames(final_components_preditor_array)<-dimnames(oos_components_preditor_array)<-list(paste("shift=",shift_vec,sep=""),
                                                                                           paste("h=",h_vec,sep=""),rownames(x_mat))
# Initialize arrays collecting final predictors, real-time predictors and regression weights
#   -These will be used when analyzing revisions
  track_weights_array<-array(dim=c(length(shift_vec),length(h_vec),nrow(x_mat),length(sel_vec_pred)+1))
  dimnames(track_weights_array)<-list(paste("shift=",shift_vec,sep=""),
                                      paste("h=",h_vec,sep=""),rownames(x_mat),c("Intercept",sel_vec_pred))

# Use OLS
  use_garch<-F
# Set-up progress bar: indicates progress in R-console
  pb <- txtProgressBar(min=min(h_vec),max=max(h_vec)-1,style=3)

# The following double loop computes all combinations of forward-shifts (of BIP) and forecast horizons (of M-SSA)
  for (shift in shift_vec)#shift<-5
  {
# Progress bar: see R-console
    setTxtProgressBar(pb, shift)
    for (j in h_vec)#j<-6
    {
# Horizon j corresponds to k=j+1-th entry of array    
      k<-j+1
# A. M-SSA component predictor
# Specify data matrix for WLS regression
      if (length(sel_vec_pred)>1)
      {
        dat<-cbind(c(x_mat[(shift+lag_vec[1]+1):nrow(x_mat),1],rep(NA,shift+lag_vec[1])),t(mssa_array[sel_vec_pred,,k]))
      } else
      {
        dat<-cbind(c(x_mat[(shift+lag_vec[1]+1):nrow(x_mat),1],rep(NA,shift+lag_vec[1])),(mssa_array[sel_vec_pred,,k]))
      }
      rownames(dat)<-rownames(x_mat)
      colnames(dat)<-c(colnames(x_mat)[1],sel_vec_pred)
# Apply the previous function: compute GARCH, WLS regression, out-of-sample MSEs and p-values    
      perf_obj<-optimal_weight_predictor_func(dat,in_out_separator,use_garch,shift,lag_vec)
# Retrieve out-of-sample performances 
# a. p-values with/without Pandemic    
      p_mat_mssa_components[shift+1,k]<-perf_obj$p_value
      p_mat_mssa_components_without_covid[shift+1,k]<-perf_obj$p_value_without_covid
# b. MSE forecast error out-of-sample
#   -M-SSA components with/without Pandemic    
      MSE_oos_mssa_comp_mat[shift+1,k]<-MSE_oos_mssa_comp<-perf_obj$MSE_oos
      MSE_oos_mssa_comp_without_covid_mat[shift+1,k]<-MSE_oos_mssa_comp_without_covid<-perf_obj$MSE_oos_without_covid
#   -mean-benchmark with/without Pandemic    
      MSE_mean_oos<-perf_obj$MSE_mean_oos
      MSE_mean_oos_without_covid<-perf_obj$MSE_mean_oos_without_covid
# Here we retrieve the final in-sample predictor (based on the full-sample WLS regression) as well as the 
#    real-time out-of-sample predictor (re-adjusted to new data at each time point)
# We can plot both predictors to illustrate revisions (due to WLS estimation at each time point), see below
      final_in_sample_preditor<-perf_obj$final_in_sample_preditor
      final_components_preditor_array[shift+1,j+1,(nrow(x_mat)-length(final_in_sample_preditor)+1):nrow(x_mat)]<-final_in_sample_preditor
      cal_oos_pred<-perf_obj$cal_oos_pred
      oos_components_preditor_array[shift+1,j+1,(nrow(x_mat)-length(cal_oos_pred)+1):nrow(x_mat)]<-cal_oos_pred
# We can also obtain the regression weights to track changes (systematic vs. noisy revisions) over time
# Note: the variable will be overwritten, i.e., we keep only the last run through the double loop, 
#   corresponding to maximal shift and maximal forecast horizon, see exercise 2.2 below 
      track_weights<-perf_obj$track_weights
      track_weights_array[shift+1,j+1,(nrow(x_mat)-nrow(track_weights)+1):nrow(x_mat),]<-track_weights

# B. Direct forecasts
# -The main difference to M-SSA above is the specification of the explanatory variables in the data 
#     matrix dat: we here use x_mat instead of mssa_array. 
#   -We select all indicators (one could easily change this setting but results are only marginally effected as long as ifo and ESi are included)
#   -Note that the data matrix here does not depend on j, in contrast  to the M-SSA components above    
      dat<-cbind(c(x_mat[(shift+lag_vec[1]+1):nrow(x_mat),1],rep(NA,shift+lag_vec[1])),x_mat)
      rownames(dat)<-rownames(x_mat)
      dat<-dat

      perf_obj<-optimal_weight_predictor_func(dat,in_out_separator,use_garch,shift,lag_vec)
# Retrieve out-of-sample performances: p-values and forecast MSE, with/without Pandemic 
      p_mat_direct[shift+1,k]<-perf_obj$p_value 
      p_mat_direct_without_covid[shift+1,k]<-perf_obj$p_value_without_covid 
      MSE_oos_direct<-perf_obj$MSE_oos
      MSE_oos_direct_without_covid<-perf_obj$MSE_oos_without_covid

# Compute rRMSEs
# a. M-SSA Components vs. direct forecast    
      rRMSE_mSSA_comp_direct[shift+1,k]<-sqrt(MSE_oos_mssa_comp/MSE_oos_direct)
# b. M-SSA Components vs. mean benchmark    
      rRMSE_mSSA_comp_mean[shift+1,k]<-sqrt(MSE_oos_mssa_comp/MSE_mean_oos)
# c. Direct forecast vs. mean benchmark    
      rRMSE_mSSA_direct_mean[shift+1,k]<-sqrt(MSE_oos_direct/MSE_mean_oos)
# Same as a, b, c but without Pandemic
      rRMSE_mSSA_comp_direct_without_covid[shift+1,k]<-sqrt(MSE_oos_mssa_comp_without_covid/MSE_oos_direct_without_covid)
      rRMSE_mSSA_comp_mean_without_covid[shift+1,k]<-sqrt(MSE_oos_mssa_comp_without_covid/MSE_mean_oos_without_covid)
      rRMSE_mSSA_direct_mean_without_covid[shift+1,k]<-sqrt(MSE_oos_direct_without_covid/MSE_mean_oos_without_covid)
    }
  }
# Close progress bar
  close(pb)
# Note: possible warnings issued by the GARCH estimation routine during computations can be ignored

# Assign column and rownames
  colnames(p_mat_mssa_components)<-colnames(p_mat_direct)<-colnames(p_mat_mssa_components_without_covid)<-
    colnames(rRMSE_mSSA_comp_direct)<-colnames(rRMSE_mSSA_comp_mean)<-
    colnames(rRMSE_mSSA_comp_direct_without_covid)<-colnames(rRMSE_mSSA_comp_mean_without_covid)<-
    colnames(rRMSE_mSSA_direct_mean)<-colnames(rRMSE_mSSA_direct_mean_without_covid)<-
    colnames(p_mat_direct_without_covid)<-colnames(MSE_oos_mssa_comp_mat)<-
    colnames(MSE_oos_mssa_comp_without_covid_mat)<-paste("h=",h_vec,sep="")
  rownames(p_mat_mssa_components)<-rownames(p_mat_direct)<-rownames(p_mat_mssa_components_without_covid)<-
    rownames(rRMSE_mSSA_comp_direct)<-rownames(rRMSE_mSSA_comp_mean)<-
    rownames(rRMSE_mSSA_comp_direct_without_covid)<-rownames(rRMSE_mSSA_comp_mean_without_covid)<-
    rownames(rRMSE_mSSA_direct_mean)<-rownames(rRMSE_mSSA_direct_mean_without_covid)<-
    rownames(p_mat_direct_without_covid)<-rownames(MSE_oos_mssa_comp_mat)<-
    rownames(MSE_oos_mssa_comp_without_covid_mat)<-paste("Shift=",shift_vec,sep="")
# Define list for saving all matrices  
    list_perf<-list(p_mat_mssa_components=p_mat_mssa_components,p_mat_direct=p_mat_direct,
    p_mat_mssa_components_without_covid=p_mat_mssa_components_without_covid,rRMSE_mSSA_comp_direct=rRMSE_mSSA_comp_direct,
    rRMSE_mSSA_comp_mean=rRMSE_mSSA_comp_mean, rRMSE_mSSA_comp_direct_without_covid=rRMSE_mSSA_comp_direct_without_covid,
    rRMSE_mSSA_comp_mean_without_covid=rRMSE_mSSA_comp_mean_without_covid,rRMSE_mSSA_direct_mean=rRMSE_mSSA_direct_mean,
    rRMSE_mSSA_direct_mean_without_covid=rRMSE_mSSA_direct_mean_without_covid,p_mat_direct_without_covid=p_mat_direct_without_covid,
    final_components_preditor_array=final_components_preditor_array,oos_components_preditor_array=oos_components_preditor_array,track_weights_array=track_weights_array,MSE_oos_mssa_comp_mat=MSE_oos_mssa_comp_mat,
    MSE_oos_mssa_comp_without_covid_mat=MSE_oos_mssa_comp_without_covid_mat)
# The results can be saved (overwritten)    
    if (F)
    {
      save(list_perf,file=paste(path.result,"list_perf",sep=""))
    }
} else
{
# Load all results  
  load(file=paste(path.result,"list_perf",sep=""))
  p_mat_mssa_components=list_perf$p_mat_mssa_components
  p_mat_direct=list_perf$p_mat_direct
  p_mat_mssa_components_without_covid=list_perf$p_mat_mssa_components_without_covid
  rRMSE_mSSA_comp_direct=list_perf$rRMSE_mSSA_comp_direct
  rRMSE_mSSA_comp_mean=list_perf$rRMSE_mSSA_comp_mean
  rRMSE_mSSA_comp_direct_without_covid=list_perf$rRMSE_mSSA_comp_direct_without_covid
  rRMSE_mSSA_comp_mean_without_covid=list_perf$rRMSE_mSSA_comp_mean_without_covid
  rRMSE_mSSA_direct_mean=list_perf$rRMSE_mSSA_direct_mean
  rRMSE_mSSA_direct_mean_without_covid=list_perf$rRMSE_mSSA_direct_mean_without_covid
  p_mat_direct_without_covid=list_perf$p_mat_direct_without_covid  
  final_components_preditor_array=list_perf$final_components_preditor_array
  oos_components_preditor_array=list_perf$oos_components_preditor_array
  track_weights_array=list_perf$track_weights_array
  MSE_oos_mssa_comp_mat=list_perf$MSE_oos_mssa_comp_mat
  MSE_oos_mssa_comp_without_covid_mat=list_perf$MSE_oos_mssa_comp_without_covid_mat
}

# HAC-adjusted p-values of out-of-sample (M-SSA) components predictor when targeting forward-shifted BIP
#   -Evaluation based on out-of-sample span starting at in_out_separator and ending on Jan-2025
p_mat_mssa_components
p_mat_direct
# Same but without singular Pandemic
p_mat_mssa_components_without_covid
p_mat_direct_without_covid
# The link between the new predictor and future BIP is statistically significant up to multiple quarters ahead 
#   -Designs optimized for larger forecast horizons (columns with h>=4) seem to perform significantly up to 
#     one year ahead

# -The above p-values are based on regressions (of out-of-sample predictors on BIP) thereby ignoring 
#   `static' level and scale parameters (which are automatically adjusted by the regressions)
# -The following rRMSEs account (also) for scale and level of the new predictor
#   -All results out-of-sample (starting just before the financial crisis)
# a. rRMSE of M-SSA components when benchmarked against mean, out-of-sample
rRMSE_mSSA_comp_mean
# b. rRMSE of M-SSA components when benchmarked against direct forecast, out-of-sample
rRMSE_mSSA_comp_direct
# c. rRMSE of direct forecasts when benchmarked against mean benchmark, out-of-sample
#   Note: the columns are all the same because for a given shift, the explanatory variables of the 
#     direct forecast do not depend on j in the above loop  
rRMSE_mSSA_direct_mean

# Same but without Pandemic
rRMSE_mSSA_comp_mean_without_covid
rRMSE_mSSA_comp_direct_without_covid
rRMSE_mSSA_direct_mean_without_covid

@
<<label=ats_mba_2,echo=FALSE,results=tex>>=

mat1<-round(p_mat_mssa_components_without_covid,3)
#latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
#center = "centering", file = "", floating = FALSE)
xtable(mat1, dec = 1,digits=rep(3,dim(mat1)[2]+1),
paste("HAC adjusted p-values of regressions of M-SSA component predictors optimized for forecast horizons $h=0,...,6$  (the columns) on BIP shifted forward by shift=0,...,5 (the rows). Out-of-sample span starting in Jan-2007, without pandemic."),
label=paste("p_val_wc",sep=""),
center = "centering", file = "", floating = FALSE)

mat1<-round(rRMSE_mSSA_comp_mean_without_covid,3)
#latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
#center = "centering", file = "", floating = FALSE)
xtable(mat1, dec = 1,digits=rep(3,dim(mat1)[2]+1),
paste("rRMSEs of M-SSA components predictor benchmarked against the expanding mean of BIP. Out-of-sample span starting in Jan-2007, without pandemic."),
label=paste("rRMSE_mSSA_comp_mean_wc",sep=""),
center = "centering", file = "", floating = FALSE)

@

\section{Revisions}\label{revisions}

<<label=init,results=hide>>=

# Select h and shift
h<-4
shift<-4

file = "./Figures/revisions1.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)

par(mfrow=c(1,1))
mplot<-cbind(final_components_preditor_array[shift,h,],oos_components_preditor_array[shift,h,])
colnames(mplot)<-c("Final predictor","Real-time out-of-sample predictor")
colo<-c("blue",rainbow(length(select_vec_multi)))
main_title<-paste("Final vs. real-time predictor: h=",h,", shift=",shift,sep=""
)
plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],lwd=1,ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
mtext(colnames(mplot)[1],col=colo[1],line=-1)
for (i in 1:ncol(mplot))
{
  lines(mplot[,i],col=colo[i],lwd=1,lty=1)
  mtext(colnames(mplot)[i],col=colo[i],line=-i)
}
abline(h=0)
#abline(v=which(rownames(mplot)<=date_to_fit)[length(which(rownames(mplot)<=date_to_fit))],lwd=2,lty=2)
axis(1,at=c(1,4*1:(nrow(mplot)/4)),labels=rownames(mplot)[c(1,4*1:(nrow(mplot)/4))])
axis(2)
box()

dev.off()
# Notes: 
# 1. The above plot compares real-time and final predictors for maximal shift and maximal forecast horizon
#     -Last run in the double-loop of exercise 1.3.5 above 
#     -Similar plots could be obtained for all combinations of forward-shift and forecast horizon (with no systematic change/difference)
# 2. The last data point in the above plot does not correspond to Jan-2025 (because the target is forward-shifted and NAs are removed)
tail(mplot)
#     -The purpose of the above plot is to illustrate revisions 
#     -It does not show the current forecast at the series end (Jan-2025)

# Comments (revisions): 
# -To the left of the plot, the real-time predictor is volatile because the sample is short (revisions are large)
# -With increasing sample size (from left to right), the real-time predictor approaches the final estimate 
# -The vertical black line in the plot indicates the start of the evaluation period, relevant for 
#     out-of-sample MSEs and p-value statistics, see exercise 1.3.5 above
# -As time progresses, we expect better out-of-sample forecast performances because the part imputable to the 
#   above revision error will decrease

# 2.2 We now examine the effect of the revisions on the regression weights

file = "./Figures/revisions2.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)

par(mfrow=c(1,1))
mplot<-track_weights_array[shift,h,,]
#colnames(mplot)[2:ncol(track_weights)]<-paste("Weight of M-SSA component ",colnames(track_weights)[2:ncol(track_weights)])
colo<-c("black","blue","red")
main_title<-paste("Intercept and regression weight over time: h=",h,", shift=",shift,sep="")
plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],lwd=1,ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
mtext(colnames(mplot)[1],col=colo[1],line=-1)
for (i in 1:ncol(mplot))
{
  lines(mplot[,i],col=colo[i],lwd=1,lty=1)
  mtext(colnames(mplot)[i],col=colo[i],line=-i)
}
abline(h=0)
#abline(v=which(rownames(mplot)<=date_to_fit)[length(which(rownames(mplot)<=date_to_fit))],lwd=2,lty=2)
axis(1,at=c(1,4*1:(nrow(mplot)/4)),labels=rownames(mplot)[c(1,4*1:(nrow(mplot)/4))])
axis(2)
box()
dev.off()
@
Revisions refer to changes in the historical values of a time series resulting from updates with new information. Revisions of the M-SSA component predictor can arise from data updates, which we do not analyze further here\footnote{Except for BIP, the other indicators are not or only slightly revised. Moreover, the smoothing effect of filters mitigates the effect of revisions when compared to direct forecasts.}, or from updates to predictor parameters. The latter can be further separated into changes in VAR parameters and regression weights. 
\subsection{Revisions: Regression Parameters}

For illustration, we here focus on a forecast horizon $h=\Sexpr{h}$ for the BIP component predictor and a forward-shift $shift=\Sexpr{shift}$ for BIP, noting that similar results would be obtained with other combinations of $h$ and $shift$. 
%???Check left-shift of MA-inversion of M-SSA and M-MSE
%??? Check forecast excess for HP-C
%??? Justification HP(160): replace HP(1600) by AR(2) with similar persistency but shorter cycle lengthi, i.e. $a_1=-2"persistency*acos(\theta, a_2=persistency$.
Fig.\eqref{revisions1} compares the final and real-time predictors: as the sample size increases, the real-time predictor converges steadily to the final predictor and the rate of convergence is partially determined by the simplicity of our design, which avoids incorporating additional M-SSA components. 
As a confirmation, Fig.\eqref{revisions2} displays the evolution of the intercept and regression weights over time: with increasing sample size, these estimates seem to converge to fixed points, indicating both the stationarity of the process and the consistency of the estimates (note, however, the slightly elevated values of the regression weight assigned to the M-SSA component during the financial crisis and the pandemic, suggesting a stronger link between the target and the predictor).  In contrast, more complex designs based on multiple M-SSA components can exhibit substantial fluctuations in the regression parameters, sometimes changing signs. This suggests instability and complicates interpretation.
<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
file = "./Figures/revisions1.pdf"
cat("\\begin{figure}[H]")
cat("\\begin{center}")
cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
cat("\\caption{Revisions of the M-SSA BIP component predictor are shown, with the final predictor in blue and the real-time design in red. These revisions result from quarterly updates to the regression equations. The M-SSA filter itself remains fixed, based on data up to January 2008.", sep = "")
cat("\\label{revisions1}}", sep = "")
cat("\\end{center}")
cat("\\end{figure}")
@
<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
file = "./Figures/revisions2.pdf"
cat("\\begin{figure}[H]")
cat("\\begin{center}")
cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
cat("\\caption{Revisions of intercept and regression weight.", sep = "")
cat("\\label{revisions2}}", sep = "")
cat("\\end{center}")
cat("\\end{figure}")
@

\subsection{Revisions: M-SSA}

The second source of revisions in the M-SSA component predictor arises from updates to the VAR model within the M-SSA filter. We compare the original estimates, based on data up to January 2008, with the final estimates obtained using $h=\Sexpr{h}$ and $shift=\Sexpr{shift}$ (similar findings apply to other values of forecast horizon and forward-shift).  For consistency, we excluded the entire pandemic episode from the data. Fig.\eqref{up_dated_ma_inv_multi_ip} compares the MA-inversions (impulse responses) of the BIP equation derived from the 'old' and the updated VAR models: unlike the former, the latter incorporates ESI as an additional explanatory variable for BIP. Fig.\eqref{bk_2008_all} illustrates the resulting effect on the M-SSA filter targeting HP-BIP at forecast horizon $h=\Sexpr{h}$. The updated filter (right panel) assigns more weight to the additional 'leading' indicators relative to BIP. Finally, Fig.\eqref{revisions3} compares the M-SSA BIP predictors. To isolate the effect of the regression revision, the figure displays standardized series. The update to the VAR model influences the left-shift of the predictor (blue line), which appears slightly 'faster' than the original predictor (red line). This suggests that the out-of-sample performance of the M-SSA component predictor may be better than reported in Section \eqref{oosp}, which is based on the fixed and increasingly outdated version of M-SSA.
<<label=init,results=hide>>=

# Up-date M-SSA and M-MSE
# In-sample span for VAR, use all data for VAR
date_to_fit_all<-"3000"
# Remove Pandemic
x_mat_wc<-x_mat[c(which(rownames(x_mat)<2020),which(rownames(x_mat)>2021)),]
par(mfrow=c(1,1))
ts.plot(x_mat_wc)

# Run the M-SSA wrapper, see tutorial 7.2
#   -The function computes M-SSA and M-MSE for each forecast horizon h in h_vec
if (recompute_results)
{
  final_mssa_indicator_obj<-compute_mssa_BIP_predictors_func(x_mat_wc,lambda_HP,L,date_to_fit_all,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi)
  save(final_mssa_indicator_obj,file=paste(path.result,"/final_mssa_indicator_obj",sep=""))
} else
{
  load(file=paste(path.result,"/final_mssa_indicator_obj",sep=""))
}
# Final M-SSA components
final_mssa_array<-final_mssa_indicator_obj$mssa_array

#----------------------
# Compute up-dated M-SSA component predictors

if (length(sel_vec_pred)>1)
{
  dat<-cbind(c(x_mat_wc[(shift+lag_vec[1]+1):nrow(x_mat_wc),1],rep(NA,shift+lag_vec[1])),t(final_mssa_array[sel_vec_pred,,h+1]))
} else
{
  dat<-cbind(c(x_mat_wc[(shift+lag_vec[1]+1):nrow(x_mat_wc),1],rep(NA,shift+lag_vec[1])),(final_mssa_array[sel_vec_pred,,h+1]))
}
lm_obj<-lm(dat[,1]~dat[,2:ncol(dat)])
summary(lm_obj)

optimal_weights<-lm_obj$coef
# Compute predictor for each forward-shift  
if (length(sel_vec_pred)>1)
{
  final_mssa_predictor<-optimal_weights[1]+dat[,2:ncol(dat)]%*%optimal_weights[2:length(optimal_weights)]
} else
{
  final_mssa_predictor<-optimal_weights[1]+dat[,2:ncol(dat)]*optimal_weights[2:length(optimal_weights)]
}

# Compare and plot old and final M-SSA BIP component predictors
#   Standardize predictors to isolate VAR-revision from total revision (including regression revission)
mssa_final_vs_prem_wc<-cbind(final_mssa_predictor,final_components_preditor_array[shift,h,c(which(rownames(x_mat)<2020),which(rownames(x_mat)>2021))])
rownames(mssa_final_vs_prem_wc)<-rownames(x_mat_wc)
colnames(mssa_final_vs_prem_wc)<-c("Predictor based on final M-SSA","Predictor based on 2007-M-SSA")

file = "./Figures/revisions3.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
par(mfrow=c(1,1))

colo<-c("blue","red")
mplot<-scale(mssa_final_vs_prem_wc)
colnames(mplot)<-colnames(mssa_final_vs_prem_wc)
main_title<-paste("BIP component predictors: final and 2007-MSSA (h=",h,", shift=",shift,")",sep="")
plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
mtext(colnames(mplot)[1],col=colo[1],line=-1)
for (jj in 1:ncol(mplot))
{
  lines(mplot[,jj],col=colo[jj],lwd=1,lty=1)
  mtext(colnames(mplot)[jj],col=colo[jj],line=-jj)
}
abline(h=0)
axis(1,at=c(1,12*1:(nrow(mplot)/12)),labels=rownames(mplot)[c(1,12*1:(nrow(mplot)/12))])
axis(2)
box()
dev.off()

#---------------------------------------------
# Compute impulse responses of old and of up-dated VAR
if (recompute_results)
{
# VAR based on full data set without pandemic  
  data_fit<-x_mat_wc

  set.seed(12)
  V_obj_full<-VARMA(data_fit,p=p,q=q)
  threshold<-1.5
  V_obj_full<-refVARMA(V_obj_full, thres = threshold)
  save(V_obj_full,file=paste(path.result,"/VAR_full",sep=""))

# VAR based on data pre 2008  
  data_fit<-x_mat_wc[which(rownames(x_mat_wc)<date_to_fit),]

  set.seed(12)
  V_obj_2008<-VARMA(data_fit,p=p,q=q)
  threshold<-1.5
  V_obj_2008<-refVARMA(V_obj_2008, thres = threshold)
  save(V_obj_2008,file=paste(path.result,"/VAR_2008",sep=""))

} else
{
    load(file=paste(path.result,"/VAR_full",sep=""))
    load(file=paste(path.result,"/VAR_2008",sep=""))

}

# Diagnostics are OK
if (F)
{
  MTSdiag(V_obj_full)
  MTSdiag(V_obj_2008)
}

# Compute MA-inversion
compute_xi_func<-function(V_obj,L)
{
  Sigma<-V_obj$Sigma
  V_obj$Phi
  V_obj$Theta
  n<-dim(Sigma)[1]
  # MA inversion
  xi_psi<-PSIwgt(Phi = V_obj$Phi, Theta = V_obj$Theta, lag = L, plot = F, output = F)
  xi_p<-xi_psi$psi.weight
  # Transform Xi_p into Xi: first L entries, from left to right, are weights of first WN, next L entries are weights of second WN 
  xi<-matrix(nrow=n,ncol=n*L)
  for (i in 1:n)
  {
    for (j in 1:L)
      xi[,(i-1)*L+j]<-xi_p[,i+(j-1)*n]
  }
  return(list(xi=xi))
}

xi_full<-compute_xi_func(V_obj_full,L)$xi
xi_2008<-compute_xi_func(V_obj_2008,L)$xi

# Plot MA inversions

file = "./Figures/up_dated_ma_inv_multi_ip.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
colo<-rainbow(ncol(x_mat))
# Select BIP component
i<-1
par(mfrow=c(1,2))
# 2008
xi<-xi_2008
mplot<-xi[i,1:min(10,L)]
for (j in 2:n)
{
  mplot<-cbind(mplot,xi[i,(j-1)*L+1:min(10,L)])
}
ts.plot(mplot,col=colo_i,main=paste("2008: MA inversion ",colnames(data_fit)[i],sep=""),xlab="lag")
for (k in 1:length(select_vec_i))
  mtext(select_vec_i[k],line=-k,col=colo_i[k])

# full data set
xi<-xi_full
mplot<-xi[i,1:min(10,L)]
for (j in 2:n)
{
  mplot<-cbind(mplot,xi[i,(j-1)*L+1:min(10,L)])
}
ts.plot(mplot,col=colo_i,main=paste("Full data: MA inversion ",colnames(data_fit)[i],sep=""),xlab="lag")
for (k in 1:length(select_vec_i))
  mtext(select_vec_i[k],line=-k,col=colo_i[k])

dev.off()

#---------------------------------------------
# Plot M-SSA filter weights of old and of up-dated VAR

bk_all<-matrix(final_mssa_indicator_obj$bk_x_array[select,select,,"h=4"],nrow=L)
bk_2008<-matrix(mssa_indicator_obj$bk_x_array[select,select,,"h=4"],nrow=L)
colnames(bk_2008)<-colnames(bk_all)<-select_vec_multi

file = "./Figures/bk_2008_all.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)

par(mfrow=c(1,2))
mplot<-bk_2008
colo<-rainbow(ncol(mplot))
ts.plot(mplot,col=colo,main=paste("M-SSA 2008: h=",h,sep=""))
for (i in 1:ncol(mplot))
  mtext(colnames(mplot)[i],col=colo[i],line=-i)
mplot<-bk_all
colo<-rainbow(ncol(mplot))
ts.plot(mplot,col=colo,main=paste("M-SSA full data: h=",h,sep=""))
for (i in 1:ncol(mplot))
  mtext(colnames(mplot)[i],col=colo[i],line=-i)

dev.off()

@
<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
file = "./Figures/up_dated_ma_inv_multi_ip.pdf"
cat("\\begin{figure}[H]")
cat("\\begin{center}")
cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
cat("\\caption{MA-inversion (impulse response) of VAR: data up to Jan-2008 (left) vs. entire data set (right).", sep = "")
cat("\\label{up_dated_ma_inv_multi_ip}}", sep = "")
cat("\\end{center}")
cat("\\end{figure}")
@

<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
file = "./Figures/bk_2008_all.pdf"
cat("\\begin{figure}[H]")
cat("\\begin{center}")
cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
cat("\\caption{M-SSA filter tracking HP-BIP at forecast horizon $h=4$: data up to Jan-2008 (left) vs. entire data set (right).", sep = "")
cat("\\label{bk_2008_all}}", sep = "")
cat("\\end{center}")
cat("\\end{figure}")
@

<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
file = "./Figures/revisions3.pdf"
cat("\\begin{figure}[H]")
cat("\\begin{center}")
cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
cat("\\caption{Revisions resulting from updates to the VAR model in M-SSA. To isolate the effect of the regression revision, all series are standardized. The up-dated M-SSA (blue) is left-shifted at the zero-crossings and appears slightly faster. For consistency, the entire pandemic episode has been removed.", sep = "")
cat("\\label{revisions3}}", sep = "")
cat("\\end{center}")
cat("\\end{figure}")
@

\section{Summary and Conclusion}

Classic direct forecasts of BIP can outperform simple benchmarks, such as the mean, up to two quarters ahead. However, this limited forecast horizon is often insufficient for certain applications. A key challenge with direct forecasts is that the relevant indicators tend to be noisy, which can mask the true signal and lead to overfitting. In particular, the predictor struggles to timely track dips (slowdowns, recessions) or peaks (recoveries, expansions). To address this issue, we apply a HP(160)-filter to smooth the data, with the smoothing parameter $\lambda=160$ reflecting a more adaptive design than the classic quarterly HP, in line with an intended forecast horizon of up to one year ahead. The resulting predictor tends to outperform the classic direct forecasts in terms of statistical significance, mainly due to a slight left-shift noticeable at sharp recession dips. However, despite the filtering, the predictor remains somewhat noisy, potentially producing excessive noisy zero-crossings at the transitions between recessions and expansions. This issue is partly due to noise leakage inherent in the one-sided HP filter. \\

To address these problems, we propose a multivariate extension of SSA based on Wildi (2024). Unlike the univariate HP filter, the M-SSA can leverage cross-sectional information from indicators leading BIP in real time while controlling for the rate of zero-crossings of the predictor. Consequently, the multivariate filter outputs become increasingly left-shifted as the forecast horizon lengthens, allowing dips and peaks of the target series to be tracked more systematically, while reducing the occurrence of noisy zero-crossings. However, since M-SSA does not explicitly target BIP, an additional step is necessary to derive the final BIP predictor. We propose two approaches: a simple equally-weighted aggregate, assuming all M-SSA components are equally important for predicting BIP, and an optimally weighted predictor based on a regression of the M-SSA components on future BIP.\\

For illustration, we consider a simple approach based on regressing the single M-SSA BIP output on future BIP, thereby ignoring the other filter outputs. This predictor is intuitively appealing because the future HP-BIP—i.e., the target of M-SSA—is the low-frequency component of the future BIP target. Consequently, a strong link between M-SSA and future HP-BIP also indicates a connection with BIP, even if the statistical significance is obscured by the noise inherent in BIP. Out-of-sample performance suggests that this predictor is statistically significant at horizons up to one year. Moreover, designs optimized for larger forecast horizons outperform the mean benchmark at all considered forecast horizons, and the predictor exceeds the performance of direct forecasts—based on unfiltered or HP filtered indicators—at horizons longer than two quarters. An analysis of revision errors arising from quarterly updates of the proposed M-SSA BIP component predictor indicates that the regression parameters, the VAR model, and the resulting M-SSA filters stabilize after an initial burn-in period. This stability enhances the explainability and interpretability of the predictor.\\

In conclusion, the new predictor design integrates the traditional direct forecast approach with a novel multivariate filter to target BIP up to one year ahead. This multivariate approach leverages a set of economic indicators leading BIP in real time while controlling the smoothness of the predictor through the zero-crossing rate.

\newpage

\section{Appendix}

\subsection{Performances: Without Pandemic}

<<label=ats_mba_2,echo=FALSE,results=tex>>=

mat1<-round(p_mat_mssa_components_without_covid,3)
#latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
#center = "centering", file = "", floating = FALSE)
xtable(mat1, dec = 1,digits=rep(3,dim(mat1)[2]+1),
paste("HAC adjusted p-values of regressions of M-SSA component predictors optimized for forecast horizons $h=0,...,6$  (the columns) on BIP shifted forward by shift=0,...,5 (the rows). Out-of-sample span starting in Jan-2007, without pandemic."),
label=paste("p_val_wc5",sep=""),
center = "centering", file = "", floating = FALSE)

mat1<-round(rRMSE_mSSA_comp_mean_without_covid,3)
#latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
#center = "centering", file = "", floating = FALSE)
xtable(mat1, dec = 1,digits=rep(3,dim(mat1)[2]+1),
paste("rRMSEs of M-SSA components predictor benchmarked against the expanding mean of BIP. Out-of-sample span starting in Jan-2007, without pandemic."),
label=paste("rRMSE_mSSA_comp_mean6",sep=""),
center = "centering", file = "", floating = FALSE)

mat1<-round(rRMSE_mSSA_comp_direct_without_covid,3)
#latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
#center = "centering", file = "", floating = FALSE)
xtable(mat1, dec = 1,digits=rep(3,dim(mat1)[2]+1),
paste("rRMSEs of M-SSA components predictor benchmarked against the direct forecasts. Out-of-sample span starting in Jan-2007, without pandemic."),
label=paste("rRMSE_mSSA_comp_direct_without_covid7",sep=""),
center = "centering", file = "", floating = FALSE)

mat1<-round(rRMSE_mSSA_direct_mean_without_covid,3)
#latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
#center = "centering", file = "", floating = FALSE)
xtable(mat1, dec = 1,digits=rep(3,dim(mat1)[2]+1),
paste("rRMSEs direct forecasts benchmarked against the expanding mean. Out-of-sample span starting in Jan-2007, without pandemic."),
label=paste("rRMSE_mSSA_direct_mean_without_covid8",sep=""),
center = "centering", file = "", floating = FALSE)

@

\newpage
\subsection{Performances: Including Singular Pandemic Data}

<<label=ats_mba_2,echo=FALSE,results=tex>>=

mat1<-round(p_mat_mssa_components,3)
#latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
#center = "centering", file = "", floating = FALSE)
xtable(mat1, dec = 1,digits=rep(3,dim(mat1)[2]+1),
paste("HAC adjusted p-values of regressions of M-SSA component predictors optimized for forecast horizons $h=0,...,6$  (the columns) on BIP shifted forward by shift=0,...,5 (the rows). Out-of-sample span starting in Jan-2007, including the pandemic."),
label=paste("p_val1",sep=""),
center = "centering", file = "", floating = FALSE)

mat1<-round(rRMSE_mSSA_comp_mean,3)
#latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
#center = "centering", file = "", floating = FALSE)
xtable(mat1, dec = 1,digits=rep(3,dim(mat1)[2]+1),
paste("rRMSEs of M-SSA components predictor benchmarked against the expanding mean of BIP. Out-of-sample span starting in Jan-2007, including the pandemic."),
label=paste("rRMSE_mSSA_comp_mean2",sep=""),
center = "centering", file = "", floating = FALSE)

mat1<-round(rRMSE_mSSA_comp_direct,3)
#latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
#center = "centering", file = "", floating = FALSE)
xtable(mat1, dec = 1,digits=rep(3,dim(mat1)[2]+1),
paste("rRMSEs of M-SSA components predictor benchmarked against the direct forecasts. Out-of-sample span starting in Jan-2007, including the pandemic."),
label=paste("rRMSE_mSSA_comp_direct3",sep=""),
center = "centering", file = "", floating = FALSE)

mat1<-round(rRMSE_mSSA_direct_mean,3)
#latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
#center = "centering", file = "", floating = FALSE)
xtable(mat1, dec = 1,digits=rep(3,dim(mat1)[2]+1),
paste("rRMSEs direct forecasts benchmarked against the expanding mean. Out-of-sample span starting in Jan-2007, including the pandemic."),
label=paste("rRMSE_mSSA_direct_mean4",sep=""),
center = "centering", file = "", floating = FALSE)

@

\end{document}

### 2025-05-03 Sweave/BIP_predictor.Rnw
up

@@
 
-@
-
-<<label=init,results=hide>>=
-
-# Load and transform data
+# Load data
 data_file_name<-c("Data_HWI_2025_02.csv","gdp_2025_02.csv")
-# Quarterly data: BIP in the first data column
+# Original (un-transformed) indicators
 data_quarterly<-read.csv(paste(getwd(),"/Data/",data_file_name[2],sep=""))
 BIP_original<-data_quarterly[,"BIP"]
-
+# Transformed indicators: differences, trimmed, standardized
 load(file=paste(getwd(),"\\Data\\macro",sep=""))
@@
 
+# Settings for M-SSA
+# In-smaple span for VAR
 date_to_fit<-"2008"
+# Model orders
 p<-1
 q<-0
+# Filter length
 L<-31
-# Publication lag
-# -In contrast to M-SSA package (tutorial 7.4) we here assume a lag of one quarter only (instead of two)
-# -To replicate the results in package we then have to impose a forecast excess of 5 to BIP (instead of 4)
-#   -M-SSA estimation routine shifts HP-BIP by lag[1]+f_excess[1]+h
+# Publication lag: one quarter for BIP
 lag_vec<-c(1,rep(0,ncol(x_mat)-1))
@@
 f_excess<-c(5,rep(4,length(select_vec_multi)-1))
+
+#------------------------------------------------------
+# Plots section data dependence
 
@@
 select_direct_indicator<-c("ifo_c","ESI")
+select_direct_indicator<-select_vec_multi
 # Note: too complex designs (too many indicators) lead to overfitting and thus worse out-of-sample performances
@@
 
-# Filter the indicators: apply HP-C
-hp_c_mat<-NULL
-for (i in 1:ncol(x_mat))
-{
-  hp_c_mat<-cbind(hp_c_mat,filter(x_mat[,i],hp_c,side=1))
-}
-colnames(hp_c_mat)<-colnames(x_mat)
-rownames(hp_c_mat)<-rownames(x_mat)
-
-# Plot
-mplot<-hp_c_mat
-par(mfrow=c(1,1))
-colo<-c(rainbow(ncol(mplot)))
-main_title<-paste("Concurrent HP(",lambda_HP,") applied to indicators",sep="")
-plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
-mtext(colnames(mplot)[1],col=colo[1],line=-1)
-for (j in 1:ncol(mplot))
-{
-  lines(mplot[,j],col=colo[j],lwd=1,lty=1)
-  mtext(colnames(mplot)[j],col=colo[j],line=-j)
-}
-abline(h=0)
-axis(1,at=c(1,12*1:(nrow(mplot)/12)),labels=rownames(mplot)[c(1,12*1:(nrow(mplot)/12))])
-axis(2)
-box()
-
 # Compute filter outputs for forecast horizons 0:6, assuming WN data
@@
 
-# Same but with pandemic
+# Apply HP-C to selected indicators
 h_vec<-0:6
@@
 <<label=init,results=hide>>=
-# Target filter: lambda_HP is the single most important hyperparameter, see tutorial 7.1 for a discussion
-lambda_HP<-160
-# Filter length: nearly 8 years is fine for the selected lambda_HP (filter weights decay sufficiently fast)
-#   The length should be an odd number (see tutorial 7.1)
-L<-31
-# In-sample span for VAR, i.e., M-SSA (the proposed design is quite insensitive to this specification because the VAR is parsimoniously parameterized
-date_to_fit<-date_to_fit
-# VARMA model orders: keep the model simple in particular for short/tight in-sample spans
-p<-1
-q<-0
-# Holding-times (HT): controls smoothness of M-SSA (the following numbers are pasted from the original predictor)
-#   -Increasing these numbers leads to predictors with less zero-crossings (smoother), see tutorial 7.1
-ht_mssa_vec<-c(6.380160,  6.738270,   7.232453,   7.225927,   7.033768)
-names(ht_mssa_vec)<-colnames(x_mat)
-# Forecast horizons: M-SSA is optimized for each forecast horizon in h_vec 
+
+# Compute and compare M-SSA and M-MSE nowcasts
 h_vec<-0
@@
 
+# Select M-SSA output for BIP: tracks HP-BIP
 select<-"BIP"
@@
 # Target series: output of two-sided HP applied to BIP: 
-#   -This is the target for which the original M-SSA predictor (tutorial 7.3) has been designed
 #   -The target is forward-shifted by the forecast horizon (plus publication lag)
 target_shifted_mat<-mssa_indicator_obj$target_shifted_mat
-# Original M-SSA predictor, see tutorial 7.3
+# Original M-SSA predictor, see tutorial 7.3 (equal-weighting of M-SSA outputs)
 #   -One predictor available for each forecast horizon in h_vec
@@
 <<label=init,results=hide>>=
-# Target filter: lambda_HP is the single most important hyperparameter, see tutorial 7.1 for a discussion
-lambda_HP<-160
-# Filter length: nearly 8 years is fine for the selected lambda_HP (filter weights decay sufficiently fast)
-#   The length should be an odd number (see tutorial 7.1)
-L<-31
-# In-sample span for VAR, i.e., M-SSA (the proposed design is quite insensitive to this specification because the VAR is parsimoniously parameterized
-date_to_fit<-date_to_fit
-# VARMA model orders: keep the model simple in particular for short/tight in-sample spans
-p<-1
-q<-0
-# Holding-times (HT): controls smoothness of M-SSA (the following numbers are pasted from the original predictor)
-#   -Increasing these numbers leads to predictors with less zero-crossings (smoother), see tutorial 7.1
-ht_mssa_vec<-c(6.380160,  6.738270,   7.232453,   7.225927,   7.033768)
-names(ht_mssa_vec)<-colnames(x_mat)
 # Forecast horizons: M-SSA is optimized for each forecast horizon in h_vec 
@@
 
+# M-SSA predictor: for each forecast horizon h=0,...,6 this is the equally-weighted average of all M-SSA outputs
 file = "./Figures/mssa_predictor.pdf"
@@
 colnames(mplot)[1]<-"BIP"
+main_title<-"Equally-weighted average of M-SSA outputs"
 plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
@@
 
-# Compute performances of original M-SSA predictor: these will be used as a benchmark when evaluating 
-#   the new design
+# Compute performances of original M-SSA predictor
 # Select start of out-of-sample span (entire financial crisis is out-of-sample)
@@
 
-# The above function generates a rich output, see tutorial 7.3
-#   -But we need only one (important/representative) performance number for our comparisons further down
-# HAC adjusted p-values of regression of original M-SSA predictor on forward-shifted BIP, see tutorial 7.3 (exercise 1.2.2)
-p_value_HAC_BIP_oos=perf_obj$p_value_HAC_BIP_oos
 # Regression on HP-BIP, without pandemic
-p_value_HAC_BIP_oos_wc<-perf_obj$p_value_HAC_HP_BIP_oos_wc
-
-p_value_HAC_BIP_oos_wc
+p_value_HAC_HP_BIP_oos_wc<-perf_obj$p_value_HAC_HP_BIP_oos_wc
+
+p_value_HAC_HP_BIP_oos_wc
 # The p-values are empirical significance levels of the regressions of the (out-of-sample) M-SSA predictors
@@
 
-mat1<-round(p_value_HAC_BIP_oos_wc,3)
+mat1<-round(p_value_HAC_HP_BIP_oos_wc,3)
 rownames(mat1)<-paste("Shift=",shift_vec,sep="")
@@
 #   -Evaluation based on out-of-sample span starting at in_out_separator and ending on Jan-2025
-p_mat_mssa_components
-p_mat_direct
-# Same but without singular Pandemic
+#   -Without singular Pandemic
 p_mat_mssa_components_without_covid
-p_mat_direct_without_covid
 # The link between the new predictor and future BIP is statistically significant up to multiple quarters ahead 
@@
 
-# -The above p-values are based on regressions (of out-of-sample predictors on BIP) thereby ignoring 
-#   `static' level and scale parameters (which are automatically adjusted by the regressions)
-# -The following rRMSEs account (also) for scale and level of the new predictor
-#   -All results out-of-sample (starting just before the financial crisis)
-# a. rRMSE of M-SSA components when benchmarked against mean, out-of-sample
-rRMSE_mSSA_comp_mean
-# b. rRMSE of M-SSA components when benchmarked against direct forecast, out-of-sample
-rRMSE_mSSA_comp_direct
-# c. rRMSE of direct forecasts when benchmarked against mean benchmark, out-of-sample
-#   Note: the columns are all the same because for a given shift, the explanatory variables of the 
-#     direct forecast do not depend on j in the above loop  
-rRMSE_mSSA_direct_mean
-
-# Same but without Pandemic
+# Same but for direct forecast
+p_mat_direct_without_covid
+
+# -Out of sample rRMSEs
+# -Without pandemic
+# a. M-SSA against mean: numbers smaller one signify outperformance of M-SSA
 rRMSE_mSSA_comp_mean_without_covid
+# b. M-SSA against direct forecast
 rRMSE_mSSA_comp_direct_without_covid
-rRMSE_mSSA_direct_mean_without_covid
+# c. Direct forecast against mean
+rRMSE_mSSA_direct_mean_without_covid[,1]
 

### 2025-05-03 Sweave/BIP_predictor.Rnw
up

@@
 # Same but for direct forecast
-p_mat_direct_without_covid
+p_mat_direct_without_covid[,1]
 
@@
 lm_obj<-lm(dat[,1]~dat[,2:ncol(dat)])
-summary(lm_obj)
 

### 2025-05-06 Sweave/BIP_predictor_new.Rnw
up

--- 2025-05-03 Sweave/BIP_predictor.Rnw
@@
 # Settings for M-SSA
+ht_mssa_vec<-c(6.380160,  6.738270,   7.232453,   7.225927,   7.033768)
+f_excess<-c(5,rep(4,length(select_vec_multi)-1))
 # In-smaple span for VAR
@@
 lag_vec<-c(1,rep(0,ncol(x_mat)-1))
-# Forecast excesses: see M-SSA packages
-f_excess<-c(5,rep(4,length(select_vec_multi)-1))
 
@@
 
-  # We can now plot target and direct forecast: for h>2 the predictor comes close to a flat line centered at zero
-
-  # We can now plot target and direct forecast: for h>2 the predictor comes close to a flat line centered at zero
-  mplot<-cbind(forward_shifted_BIP,direct_forecast)
-  colnames(mplot)<-c("BIP","Direct forecast")
-  colo<-c("black","red")
-  main_title<-paste("BIP shifted by ",h," quarters",sep="")
-  plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
-  for (i in 1:ncol(mplot))
-  {
-    lines(mplot[,i],col=colo[i])
-    mtext(colnames(mplot)[i],col=colo[i],line=-i)
-  }
-  abline(h=0)
-  axis(1,at=c(1,12*1:(nrow(mplot)/12)),labels=rownames(mplot)[c(1,12*1:(nrow(mplot)/12))])
-  axis(2)
-  box()
 }
@@
 # Forecast horizons
+# Without Pandemic
 h_vec<-0:6
@@
 
-# Apply HP-C to selected indicators
+# Same but data with Pandemic
 h_vec<-0:6
@@
 
-# Compute and plot direct HP forecasts: financial crisis
+# Financial crisis
 file = "./Figures/direct_hp_forecasts_financial_crisis.pdf"
@@
 
+# We rely on data with Pandemic
+#   -M-SSA is not affected since the in-sample span ends in 2008
 if (recompute_results)
@@
 <<label=init,results=hide>>=
-recompute_results<-F
 sel_vec_pred<-c("BIP")
@@
 # The following double loop computes all combinations of forward-shifts (of BIP) and forecast horizons (of M-SSA)
-  for (shift in shift_vec)#shift<-5
+  for (shift in shift_vec)#shift<-1
   {
@@
     setTxtProgressBar(pb, shift)
-    for (j in h_vec)#j<-6
+    for (j in h_vec)#j<-0
     {
@@
 
+# P-values
 mat1<-round(p_mat_mssa_components_without_covid,3)
@@
 
+# rRMSEs: M-SSA vs. mean
 mat1<-round(rRMSE_mSSA_comp_mean_without_covid,3)
@@
 <<label=init,results=hide>>=
+# Revisions
 
@@
 
+# Plot final predictor (based on full-sample regression of M-SSA on future BIP)
 file = "./Figures/revisions1.pdf"
@@
 dev.off()
-# Notes: 
-# 1. The above plot compares real-time and final predictors for maximal shift and maximal forecast horizon
-#     -Last run in the double-loop of exercise 1.3.5 above 
-#     -Similar plots could be obtained for all combinations of forward-shift and forecast horizon (with no systematic change/difference)
-# 2. The last data point in the above plot does not correspond to Jan-2025 (because the target is forward-shifted and NAs are removed)
-tail(mplot)
-#     -The purpose of the above plot is to illustrate revisions 
-#     -It does not show the current forecast at the series end (Jan-2025)
-
-# Comments (revisions): 
-# -To the left of the plot, the real-time predictor is volatile because the sample is short (revisions are large)
-# -With increasing sample size (from left to right), the real-time predictor approaches the final estimate 
-# -The vertical black line in the plot indicates the start of the evaluation period, relevant for 
-#     out-of-sample MSEs and p-value statistics, see exercise 1.3.5 above
-# -As time progresses, we expect better out-of-sample forecast performances because the part imputable to the 
-#   above revision error will decrease
-
-# 2.2 We now examine the effect of the revisions on the regression weights
-
+
+# Plot the regression weights 
 file = "./Figures/revisions2.pdf"

### 2025-05-06 Sweave/BIP_predictor.Rnw
up

@@
 
-recompute_results<-F
+recompute_results<-T
 
@@
 
-  ht_mssa_vec<-c(ht_hp_c,  6.738270,   7.232453,   7.225927,   7.033768)
-
-  mssa_indicator_HP_HT_obj<-compute_mssa_BIP_predictors_func(x_mat,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi)
+  ht_mssa_vec_HP_C<-c(ht_hp_c,  6.738270,   7.232453,   7.225927,   7.033768)
+
+  mssa_indicator_HP_HT_obj<-compute_mssa_BIP_predictors_func(x_mat,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec_HP_C,h_vec,f_excess,lag_vec,select_vec_multi)
 

### 2025-05-06 Sweave/BIP_predictor.Rnw
up

@@
 
-dim(mmse_array)
-
 mssa<-t(mssa_array[,,1])
@@
 mssa_array<-mssa_indicator_obj$mssa_array
+# Same but M-MSE
+mmse_array<-mssa_indicator_obj$mmse_array
 

### 2025-05-06 Sweave/generate_sweave.R
up

@@ -3,50 +3,15 @@
 rm(list=ls())
 
 
-# Load the required R-libraries
-# Standard filter package
-library(mFilter)
-# Multivariate time series: VARMA model for macro indicators: used here for simulation purposes only
-library(MTS)
-# HAC estimate of standard deviations in the presence of autocorrelation and heteroscedasticity
-library(sandwich)
-# Extended time series
-library(xts)
-# Library for Diebold-Mariano test of equal forecast performance
-library(multDM)
-# GARCH model: for improving regression estimates
-library(fGarch)
-
-
-
-# Load the relevant M-SSA functionalities
-# M-SSA functions
-source(paste(getwd(),"/R/functions_MSSA.r",sep=""))
-# Load signal extraction functions used for JBCY paper (relies on mFilter)
-source(paste(getwd(),"/R/HP_JBCY_functions.r",sep=""))
-# Utility functions for M-SSA, see tutorial 
-source(paste(getwd(),"/R/M_SSA_utility_functions.r",sep=""))
-# Set of performance metrics and tests of unequal predictability
-#source(paste(getwd(),"/R/performance_statistics_functions.r",sep=""))
-
-
-
-
 
  
 path.main<-getwd()
 
-path.pgm<-paste(path.main,"/R/",sep="")
 path.out<-paste(path.main,"/Latex/",sep="")
 path.sweave<-paste(path.main,"/Sweave/",sep="")
-path.data<-paste(path.main,"/Data/",sep="")
-# Savd results from an empirical analysis of S&P500
-path.result<-paste(path.main,"/Results/",sep="")
-fig_size<-4
-
 #------------------------------
 
-recompute_results<-T
+recompute_results<-
 
 paper<-"BIP_predictor"
 

### 2025-05-06 Sweave/BIP_predictor.Rnw
up

@@
 
-recompute_results<-T
+recompute_results<-F
 
@@
 #------------------------------------------------------
-# Plots section data dependence
+# Section 1: data
+# Plots section data
 
@@
 
+#--------------------------------------------------------
+# Section 2: dependence
+
 acf_mat<-x_mat
@@
 
+# Data without pandemic
 x_mat_wc<-x_mat[which(rownames(x_mat)<2020|rownames(x_mat)>2021),]
@@
 <<label=init,results=hide>>=
-# Plot of direct forecasts with Pandemic
-par(mfrow=c(2,2))
-# We can specify the selection of macro-indicators for the direct forecast 
-select_direct_indicator<-c("ifo_c","ESI")
+# Section 3.1: classic direct forecasts
+
+# Select indicators: we select all indicators and regression is entirely in-sample
+#   -Ideal case: performances too good (overfitting)
+#   -in applications the regression is out-of-sample (ifo and ESI work best out-of-sample) 
 select_direct_indicator<-select_vec_multi
-# Note: too complex designs (too many indicators) lead to overfitting and thus worse out-of-sample performances
-# To illustrate the direct predictor consider the following example of a h-step ahead direct forecast:
-for (h in 0:3)
-{
-  # Shift BIP forward by publication lag+forecast horizon
-  forward_shifted_BIP<-c(x_mat[(1+lag_vec[1]+h):nrow(x_mat),"BIP"],rep(NA,h+lag_vec[1]))
-  # Regress selected indicators on forward-shifted BIP
-  lm_obj<-lm(forward_shifted_BIP~x_mat[,select_direct_indicator])
-  # You will probably not find statistically significant regressors for h>2: BIP is a noisy series
-  summary(lm_obj)
-  # Technical note: 
-  # -Residuals are subject to heteroscedasticity (crises) and autocorrelation
-  # -Therefore classic OLS tests for statistical significance are biased
-  # -We shall rely on HAC-adjusted p-values further down (R-package sandwich)
-
-  # Compute the predictor: one can rely on the generic R-function predict or compute the predictor manually
-  direct_forecast<-lm_obj$coef[1]+x_mat[,select_direct_indicator]%*%lm_obj$coef[2:(length(select_direct_indicator)+1)]
-  # Note that this is a full-sample predictor (no out-of-sample span)
-
-}
 
@@
 par(mfrow=c(2,2))
-# We can specify the selection of macro-indicators for the direct forecast 
-select_direct_indicator<-c("ifo_c","ESI")
-select_direct_indicator<-select_vec_multi
 
@@
   mplot<-cbind(forward_shifted_BIP,direct_forecast)
+  rownames(mplot)<-rownames(x_mat_wc)
   colnames(mplot)<-c("BIP","Direct forecast")
@@
   mplot<-cbind(forward_shifted_BIP,direct_forecast)
+  rownames(mplot)<-rownames(x_mat_wc)
   mplot<-mplot[which(rownames(mplot)>2008&rownames(mplot)<2011),]
@@
 <<label=init,results=hide>>=
+# Section 3.2: HP filter
+
 # Target filter: lambda_HP is the single most important hyperparameter, see tutorial 7.1 for a discussion
@@
 <<label=init,results=hide>>=
+# Section 3.3: HP-filtered data used in regression on future BIP
 
@@
   mplot<-cbind(forward_shifted_BIP,direct_hp_forecast)
+  rownames(mplot)<-rownames(x_mat_wc)
+
   colnames(mplot)<-c("BIP","Direct forecast")
@@
   mplot<-cbind(forward_shifted_BIP,direct_hp_forecast)
+    rownames(mplot)<-rownames(x_mat_wc)
+
   mplot<-mplot[which(rownames(mplot)>2008&rownames(mplot)<2011),]
   colnames(mplot)<-c("BIP","Direct HP forecast")
+
   colo<-c("black","red")
@@
 rownames(mat_re)<-c("Direct forecast","Direct HP forecast")
-colnames(mat_re)<-paste("h=",0:3)
+colnames(mat_re)<-paste("h=",0:3,sep="")
 mat1<-round(mat_re,3)
@@
 <<label=init,results=hide>>=
+# Section 4.2: Nowcast M-SSA and M-MSE
 
@@
 <<label=init,results=hide>>=
+# Section 4.3: M-SSA forecasts
 # Forecast horizons: M-SSA is optimized for each forecast horizon in h_vec 
@@
 par(mfrow=c(1,2))
+# h=0, see above
 mplot<-bk
@@
   mtext(colnames(mplot)[i],col=colo[i],line=-i)
+
+# h=4
 bk<-matrix(mssa_indicator_obj$bk_x_array[select,select,,"h=4"],nrow=L)
 colnames(bk)<-colnames(gammak)<-select_vec_multi
-
 mplot<-bk
@@
 
+#----------------------------------------------------------------------------
+# Section 5.1
 # M-SSA predictor: for each forecast horizon h=0,...,6 this is the equally-weighted average of all M-SSA outputs
@@
 
+#------------------------------------------------------------
+# Section 5.2
+
 # Compute performances of original M-SSA predictor
+#   -Predictor: equal weighting of all M-SSA components
+#   -Target: two-sided HP applied to BIP (easier to forecast than BIP)
+
 # Select start of out-of-sample span (entire financial crisis is out-of-sample)
@@
 # We can specify the selection of macro-indicators for the direct forecast, see tutorial 7.3
-# Note: these results will not be used in this tutorial (but we need to specify a selection anyway)
+#   Note: these results will not be used here (but we need to specify a selection anyway)
 select_direct_indicator<-c("ifo_c","ESI")
@@
 p_value_HAC_HP_BIP_oos_wc
-# The p-values are empirical significance levels of the regressions of the (out-of-sample) M-SSA predictors
-#   (optimized for forecast horizon h in h_vec: the columns of the matrix) on BIP shifted forward by shift 
-#   plus publication lag (the rows of the matrix)
+# Strong significance
+# Systematic pattern: minimal p-Values on diagonal of matrix
+# Can predict low-frequency part of BIP
 
@@
 <<label=init,results=hide>>=
+# Section 5.3
+
 sel_vec_pred<-c("BIP")
@@
 <<label=init,results=hide>>=
-# Revisions
+# Section 6.1: revisions (regression)
 
@@
 The latter can be further separated into changes in VAR parameters and regression weights.
-\subsection{Revisions: Regression Parameters}
+
+\subsection{Revisions: Regression Equation Only}
 
@@
 
-\subsection{Revisions: M-SSA}
+\subsection{Revisions: M-SSA Only}
 
@@
 <<label=init,results=hide>>=
+# Section 6.2: revisions due to VAR-updating
 

### 2025-05-07 Latex/BIP_predictor.tex
up

@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  1 21:38:41 2025
+% Wed May  7 07:36:27 2025
 \begin{table}[ht]
@@
   \hline
- & h= 0 & h= 1 & h= 2 & h= 3 \\ 
+ & h=0 & h=1 & h=2 & h=3 \\ 
   \hline
@@
 The first L=31 observations are lost due to initialization and the two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  1 21:38:41 2025
+% Wed May  7 07:36:27 2025
 \begin{table}[ht]
@@
   \hline
-Target correlation & 0.55 & 0.63 & 0.59 & 0.68 \\ 
-  HT & 8.67 & 6.50 & 8.67 & 4.00 \\ 
+Target correlation & 0.55 & 0.68 & 0.64 & 0.69 \\ 
+  HT & 8.67 & 6.50 & 7.43 & 3.71 \\ 
    \hline
@@
 
-M-SSA predictors are compared to the traditional univariate HP-C in Fig.
-\eqref{multivar_vs_univar} across the periods of the financial crisis (top panels) and the COVID-19 pandemic (bottom panels), for forecast horizons $h=0,...,6$.
+M-SSA predictors are compared to the traditional univariate HP-C in Figs.
+\eqref{multivar_vs_univar} and \eqref{mssa_hpc_financial_crisis}.
+In the first figure, we compare both filters across the periods of the financial crisis (top panels) and the COVID-19 pandemic (bottom panels), for forecast horizons $h=0,...,6$.
 The nowcast from the previous section is included for reference.
+In the second figure we compare both filters to the forward-shifted BIP for shifts larger than two quarters.
 All series are standardized to facilitate visual comparison.
@@
 %In comparison, the effect of phase shifts—arising from the cyclical nature of the HP-filter and manifesting in the HP-C forecasts (shown in the left panels)—seems less systematic.
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/bk_h.pdf}\caption{Multivariate M-SSA: $h=0$ (nowcast, left) and $h=4$ (one year forecast, right).\label{bk_h}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/multivar_vs_univar.pdf}\caption{Now- and forecasts of HP-BIP for $h=0,...,6$: HP-C (left) vs.
-M-SSA (right) across the financial crisis (top) and the pandemic (bottom).\label{multivar_vs_univar}}\end{center}\end{figure}
+\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/bk_h.pdf}\caption{Multivariate M-SSA: $h=0$ (nowcast, left) and $h=4$ (one year forecast, right).\label{bk_h}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/multivar_vs_univar.pdf}\caption{Now- and forecasts of HP-BIP for $h=0,...,6$: HP-C (top) vs.
+M-SSA (bottom) across the financial crisis (left) and the pandemic (right).\label{multivar_vs_univar}}\end{center}\end{figure}
+\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/mssa_hpc_financial_crisis.pdf}\caption{Forward-shifted BIP (black), M-SSA (blue) and HP-C (red) for $h=3,4,5,6$ across the financial crisis.\label{mssa_hpc_financial_crisis}}\end{center}\end{figure}
 
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  1 21:38:45 2025
+% Wed May  7 07:36:30 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  1 21:38:45 2025
+% Wed May  7 07:36:30 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  1 21:38:46 2025
+% Wed May  7 07:36:32 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  1 21:38:46 2025
+% Wed May  7 07:36:32 2025
 \begin{table}[ht]

### 2025-05-07 Sweave/BIP_predictor.Rnw
up

@@
   mssa_indicator_obj<-compute_mssa_BIP_predictors_func(x_mat,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi)
+# Without pandemic  
+  mssa_indicator_obj_wc<-compute_mssa_BIP_predictors_func(x_mat_wc,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi)
+
   save(mssa_indicator_obj,file=paste(path.result,"/mssa_indicator_obj",sep=""))
+  save(mssa_indicator_obj_wc,file=paste(path.result,"/mssa_indicator_obj_wc",sep=""))
 } else
@@
   load(file=paste(path.result,"/mssa_indicator_obj",sep=""))
+  load(file=paste(path.result,"/mssa_indicator_obj_wc",sep=""))
 }
@@
 
+# Plot of M-SSA and HP-C, only financial crisis
+file = "./Figures/mssa_hpc_financial_crisis.pdf"
+pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
+
+par(mfrow=c(2,2))
+for (h in 3:6)#h<-1
+{
+  # Shift BIP forward by publication lag+forecast horizon
+  forward_shifted_BIP<-c(x_mat[(1+lag_vec[1]+h):nrow(x_mat),"BIP"],rep(NA,h+lag_vec[1]))
+
+  # We can now plot target and direct forecast: for h>2 the predictor comes close to a flat line centered at zero
+#  mplot<-scale(cbind(forward_shifted_BIP,hp_c_array_wc[select,,h+1],mssa_array_wc[select,,h+3]))
+#  mplot<-scale(cbind(forward_shifted_BIP,hp_c_array_wc[select,,h+1],predictor_mssa_mat_wc[,h+3]))
+  mplot<-scale(cbind(forward_shifted_BIP,mssa_array["BIP",,h+1],hp_c_array["BIP",,h+1]))
+  rownames(mplot)<-rownames(x_mat)
+
+  mplot<-mplot[which(rownames(mplot)>2006&rownames(mplot)<2011),]
+  colnames(mplot)<-c("BIP","M-SSA","HP-C")
+  colo<-c("black","blue","red")
+  main_title<-paste("BIP shifted forward by ",h," quarters",sep="")
+  plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
+  for (i in 1:ncol(mplot))
+  {
+    lines(mplot[,i],col=colo[i])
+    mtext(colnames(mplot)[i],col=colo[i],line=-i)
+  }
+  abline(h=0)
+  axis(1,at=1:nrow(mplot),labels=rownames(mplot))
+  axis(2)
+  box()
+}
+dev.off()
+
 #----------------------------------------------------------------------------
@@
 cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
-cat("\\caption{Now- and forecasts of HP-BIP for $h=0,...,6$: HP-C (left) vs. M-SSA (right) across the financial crisis (top) and the pandemic (bottom).", sep = "")
+cat("\\caption{Now- and forecasts of HP-BIP for $h=0,...,6$: HP-C (top) vs. M-SSA (bottom) across the financial crisis (left) and the pandemic (right).", sep = "")
 cat("\\label{multivar_vs_univar}}", sep = "")
+cat("\\end{center}")
+cat("\\end{figure}")
+@
+
+<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
+file = "./Figures/mssa_hpc_financial_crisis.pdf"
+cat("\\begin{figure}[H]")
+cat("\\begin{center}")
+cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
+cat("\\caption{Forward-shifted BIP (black), M-SSA (blue) and HP-C (red) for $h=3,4,5,6$ across the financial crisis.", sep = "")
+cat("\\label{mssa_hpc_financial_crisis}}", sep = "")
 cat("\\end{center}")

### 2025-05-07 Latex/BIP_predictor.tex
up

@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 07:36:27 2025
+% Wed May  7 08:19:43 2025
 \begin{table}[ht]
@@
 
-For illustration, we consider the target $z_t$ generated by the output of the two-sided HP(160) filter applied to BIP, which we denote by HP-BIP.
-We compare four nowcast methods: the univariate HP-C, the multivariate MSE predictor (denoted by M-MSE), and two M-SSA predictors based on different HT constraints.
-The first design enforces $ht_1=1.5ht_{MSE}$, aiming to produce approximately $33\%$ fewer zero-crossings than the M-MSE predictor over the long term.
-The second design sets $ht_1=ht_{HP-C}$, ensuring that its zero-crossing rate matches that of the univariate HP-C.
-These constraints are intended to systematically compare the effects of different levels of smoothness on prediction performance.
-The estimation of the underlying VAR(1) model employed for the M-MSE and M-SSA nowcasts is based on data observations up to January 2008.
+For illustration, we consider a nowcast of the target $z_t$ generated by the output of the two-sided HP(160) filter applied to BIP, which we denote by HP-BIP.
+We compare the classic multivariate mean squared error (MSE) predictor, denoted by M-MSE, with the M-SSA predictor, where we set $ht_1=1.5ht_{MSE}$ in the holding time constraint.
+This configuration ensures that the M-SSA predictor produces approximately 33$\%$ fewer zero-crossings than M-MSE over the long term.
+The purpose of this constraint is to compare the effect of a different level of smoothness on the prediction performance (target correlation).
+The estimation of the underlying VAR(1) model employed for the M-MSE and M-SSA is based on data observations up to January 2008.
 This approach ensures a lengthy out-of-sample period that encompasses significant events such as the financial crisis, the sovereign debt crisis, and the COVID-19 pandemic.
 Further details and background information are available in Wildi (2025b).
-All computations are performed utilizing the M-SSA package, as documented in Wildi (2025b).
-Figure \ref{bk_gammak} illustrates the coefficients of the multivariate filter designs, where increased smoothing achieved through the M-SSA method manifests as a distinct pattern in the filter weights, notably influencing the rate of decay across increasing lags.
+All computations are performed utilizing the M-SSA package, as documented in Wildi (2025b).\\
+
+Figure \ref{bk_gammak} displays the coefficients of the multivariate nowcasts, where increased smoothing achieved through the M-SSA method manifests as a distinct pattern in the filter weights, notably influencing the rate of decay across increasing lags.
 The resulting filtered series, presented in Figure \ref{mssa_msse_now}, are standardized to enable direct comparisons.
-The two M-SSA predictors adhere to the specified constraints: one exhibiting approximately one-third fewer zero-crossings than the M-MSE, and the other matching the zero-crossing rate of the HP-C, confirming the effectiveness of the smoothness control imposed during the design.
-This is further corroborated by Table \ref{corhtnow}, which reports empirical holding times and target correlations.
-In this example, both the univariate design and the M-SSA method, when configured to replicate its holding time, demonstrate the highest level of smoothness, as indicated by the maximal duration between crossings.
-However, the HP-C also attains the lowest target correlation among the compared approaches.
-More generally, our example illustrates the prediction dilemma underlying the (M-)SSA criterion, highlighting the inherent trade-off between smoothness (zero-crossing rate) and predictive accuracy (target correlation).
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/bk_gammak.pdf}\caption{Multivariate M-MSE (left) and M-SSA  (middle:smaller HT, right: larger HT) nowcast filters.\label{bk_gammak}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/mssa_msse_now.pdf}\caption{Nowcasting: two-sided HP(160) applied to BIP (target, black), univariate HP-C (violet), M-SSA smaller HT (blue), M-SSA larger HT (orange) and M-MSE (green).
+The M-SSA nowcast adheres to the specified constraint, exhibiting approximately one-third fewer zero-crossings than the M-MSE, confirming the effectiveness of the smoothness control imposed during the design.
+This is further corroborated by Table \ref{corhtnow}, which reports empirical holding times and target correlations: increased smoothness is obtained by a slight degradation of tracking accuracy, highlighting the inherent trade-off (prediction dilemma) between smoothness (zero-crossing rate) and predictive accuracy (target correlation) in the M-SSA criterion.
+\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/bk_gammak.pdf}\caption{Multivariate M-MSE (left) and M-SSA  (right) nowcast filters.\label{bk_gammak}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/mssa_msse_now.pdf}\caption{Nowcasting: two-sided HP(160) applied to BIP (target, black),  M-SSA (blue) and and M-MSE (green).
 The dashed vertical line delimits in- and out-of-sample spans.
 The first L=31 observations are lost due to initialization and the two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 07:36:27 2025
+% Wed May  7 08:19:43 2025
 \begin{table}[ht]
 \centering
-\begin{tabular}{rrrrr}
+\begin{tabular}{rrr}
   \hline
- & HP-C & M-SSA smaller HT & M-SSA larger HT & M-MSE \\ 
+ & M-SSA & M-MSE \\ 
   \hline
-Target correlation & 0.55 & 0.68 & 0.64 & 0.69 \\ 
-  HT & 8.67 & 6.50 & 7.43 & 3.71 \\ 
+Target correlation & 0.68 & 0.69 \\ 
+  HT & 6.50 & 3.71 \\ 
    \hline
 \end{tabular}
-\caption{Target correlations and HTs of HP-C, M-SSA and M-MSE nowcasts.  } 
+\caption{Target correlations and HTs of M-SSA and M-MSE nowcasts.  } 
 \label{corhtnow}
@@
 This validation is typically performed through analyses of very long samples of simulated (multivariate) data, ensuring the consistency of the estimation procedures under idealized conditions.}.
-An increase of approximately $50\%$ in the HT relative to the M-MSE predictor has only a modest effect on the target correlation, suggesting that this tradeoff could be beneficial in this application.
-Conversely, matching the HT of the HP-C predictor through M-SSA results in an increase in the target correlation, demonstrating the impact of zero-crossing control in the multivariate design on predictive performance.
+An increase of $50\%$ in the HT relative to the M-MSE predictor has only a modest effect on the target correlation, suggesting that this tradeoff could be beneficial in our application.
 Going forward, we now omit the M-MSE, as it is a special case of the more general M-SSA framework.
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 07:36:30 2025
+% Wed May  7 08:19:47 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 07:36:30 2025
+% Wed May  7 08:19:47 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 07:36:32 2025
+% Wed May  7 08:19:48 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 07:36:32 2025
+% Wed May  7 08:19:48 2025
 \begin{table}[ht]

### 2025-05-07 Latex/BIP_predictor.tex
up

@@
 \ref{direct_wc} (full dataset without the pandemic) and \ref{direct_wc_financial_crisis} (financial crisis) for horizons up to three quarters ahead.
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/direct_wc_all.pdf}\caption{Direct forecasts (red lines) based on all indicators and full sample information, without the pandemic: BIP (black line) shifted by zero one, two and three quarters.\label{direct_wc}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/direct_wc_financial_crisis.pdf}\caption{Direct forecast  across the financial crisis.\label{direct_wc_financial_crisis}}\end{center}\end{figure}The results indicate that as $h$ increases, the peaks and dips of the direct forecasts become increasingly right-shifted (delayed) relative to the target.
+\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/direct_wc_all.pdf}\caption{Direct forecasts (red lines) based on all indicators and full sample information, without the pandemic: BIP (black line) shifted by zero one, two and three quarters.\label{direct_wc}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/direct_wc_financial_crisis.pdf}\caption{Direct forecast  across the financial crisis:  regression based on all indicators and full sample information.\label{direct_wc_financial_crisis}}\end{center}\end{figure}The results indicate that as $h$ increases, the peaks and dips of the direct forecasts become increasingly right-shifted (delayed) relative to the target.
 Additionally, the F-statistics precipitously decline from $F=3.57$, for $h=2$, to $F=0.46$, for $h=3$, see Table \eqref{f_stat}, suggesting that the predictor becomes statistically insignificant for larger forecast horizons exceeding two quarters (link to Heinisch and Scheufele).
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:19:43 2025
+% Wed May  7 08:45:24 2025
 \begin{table}[ht]
@@
 The first L=31 observations are lost due to initialization and the two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:19:43 2025
+% Wed May  7 08:45:24 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:19:47 2025
+% Wed May  7 08:45:30 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:19:47 2025
+% Wed May  7 08:45:30 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:19:48 2025
+% Wed May  7 08:45:31 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:19:48 2025
+% Wed May  7 08:45:31 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:19:48 2025
+% Wed May  7 08:45:32 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:19:48 2025
+% Wed May  7 08:45:32 2025
 \begin{table}[ht]

### 2025-05-07 Sweave/BIP_predictor.Rnw
up

@@
 cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
-cat("\\caption{Direct forecast  across the financial crisis.", sep = "")
+cat("\\caption{Direct forecast  across the financial crisis:  regression based on all indicators and full sample information.", sep = "")
 cat("\\label{direct_wc_financial_crisis}}", sep = "")
@@
 
-For illustration, we consider the target $z_t$ generated by the output of the two-sided HP(160) filter applied to BIP, which we denote by HP-BIP.
-We compare four nowcast methods: the univariate HP-C, the multivariate MSE predictor (denoted by M-MSE), and two M-SSA predictors based on different HT constraints.
-The first design enforces $ht_1=1.5ht_{MSE}$, aiming to produce approximately $33\%$ fewer zero-crossings than the M-MSE predictor over the long term.
-The second design sets $ht_1=ht_{HP-C}$, ensuring that its zero-crossing rate matches that of the univariate HP-C.
-These constraints are intended to systematically compare the effects of different levels of smoothness on prediction performance.
-The estimation of the underlying VAR(1) model employed for the M-MSE and M-SSA nowcasts is based on data observations up to January 2008.
+For illustration, we consider a nowcast of the target $z_t$ generated by the output of the two-sided HP(160) filter applied to BIP, which we denote by HP-BIP.
+We compare the classic multivariate mean squared error (MSE) predictor, denoted by M-MSE, with the M-SSA predictor, where we set $ht_1=1.5ht_{MSE}$ in the holding time constraint.
+This configuration ensures that the M-SSA predictor produces approximately 33$\%$ fewer zero-crossings than M-MSE over the long term.
+The purpose of this constraint is to compare the effect of a different level of smoothness on the prediction performance (target correlation).
+The estimation of the underlying VAR(1) model employed for the M-MSE and M-SSA is based on data observations up to January 2008.
 This approach ensures a lengthy out-of-sample period that encompasses significant events such as the financial crisis, the sovereign debt crisis, and the COVID-19 pandemic.
 Further details and background information are available in Wildi (2025b).
-All computations are performed utilizing the M-SSA package, as documented in Wildi (2025b).
+All computations are performed utilizing the M-SSA package, as documented in Wildi (2025b).\\
 <<label=init,results=hide>>=
@@
 
-par(mfrow=c(1,3))
+par(mfrow=c(1,2))
 mplot<-gammak
 colo<-rainbow(ncol(mplot))
-ts.plot(mplot,col=colo,main="M-MSE")
+ts.plot(mplot,col=colo,main="M-MSE",xlab="")
 for (i in 1:ncol(mplot))
@@
 colo<-rainbow(ncol(mplot))
-ts.plot(mplot,col=colo,main="M-SSA smaller HT")
+ts.plot(mplot,col=colo,main="M-SSA",xlab="")
 for (i in 1:ncol(mplot))
   mtext(colnames(mplot)[i],col=colo[i],line=-i)
-mplot<-bk_HP_HP_C
-colo<-rainbow(ncol(mplot))
-ts.plot(mplot,col=colo,main="M-SSA larger HT")
-for (i in 1:ncol(mplot))
-  mtext(colnames(mplot)[i],col=colo[i],line=-i)
+#mplot<-bk_HP_HP_C
+#colo<-rainbow(ncol(mplot))
+#ts.plot(mplot,col=colo,main="M-SSA larger HT")
+#for (i in 1:ncol(mplot))
+#  mtext(colnames(mplot)[i],col=colo[i],line=-i)
 
@@
 par(mfrow=c(1,1))
-mplot<-scale(cbind(target_shifted_mat,hp_c_out,mssa[,select],mssa_HT_HP_C[,select],mmse[,select]))
+#mplot<-scale(cbind(target_shifted_mat,hp_c_out,mssa[,select],mssa_HT_HP_C[,select],mmse[,select]))
+#dim(mplot)
+#colnames(mplot)<-c("Target: HP-BIP","HP-C","M-SSA smaller HT","M-SSA larger HT","M-MSE")
+mplot<-scale(cbind(target_shifted_mat,mssa[,select],mmse[,select]))
 dim(mplot)
-colnames(mplot)<-c("Target: HP-BIP","HP-C","M-SSA smaller HT","M-SSA larger HT","M-MSE")
+colnames(mplot)<-c("Target: HP-BIP","M-SSA","M-MSE")
+
 rownames(mplot)<-rownames(x_mat)
 colo<-c("black","violet","blue","orange","green")
+colo<-c("black","blue","green")
 main_title<-"Nowcasting HP-BIP"
@@
 
-mat_cor_ht<-rbind(c(cor(na.exclude(mplot[,c("Target: HP-BIP","HP-C")]))[1,2],
-cor(na.exclude(mplot[,c("Target: HP-BIP","M-SSA smaller HT")]))[1,2],
-cor(na.exclude(mplot[,c("Target: HP-BIP","M-SSA larger HT")]))[1,2],
+mat_cor_ht<-rbind(c(cor(na.exclude(mplot[,c("Target: HP-BIP","M-SSA")]))[1,2],
 cor(na.exclude(mplot[,c("Target: HP-BIP","M-MSE")]))[1,2]),
-c(compute_empirical_ht_func(na.exclude(hp_c_array[select,,"h=0"]))$empirical_ht,
-compute_empirical_ht_func(na.exclude(mssa[,select]))$empirical_ht,
-compute_empirical_ht_func(na.exclude(mssa_HT_HP_C[,select]))$empirical_ht,
+c(compute_empirical_ht_func(na.exclude(mssa[,select]))$empirical_ht,
 compute_empirical_ht_func(na.exclude(mmse[,select]))$empirical_ht))
 
-colnames(mat_cor_ht)<-c("HP-C","M-SSA smaller HT","M-SSA larger HT","M-MSE")
+colnames(mat_cor_ht)<-c("M-SSA","M-MSE")
 rownames(mat_cor_ht)<-c("Target correlation","HT")
-
-mat_cor_ht
-
-@
-Figure \ref{bk_gammak} illustrates the coefficients of the multivariate filter designs, where increased smoothing achieved through the M-SSA method manifests as a distinct pattern in the filter weights, notably influencing the rate of decay across increasing lags.
+# Keep M-MSE and M-SSA only
+
+@
+
+Figure \ref{bk_gammak} displays the coefficients of the multivariate nowcasts, where increased smoothing achieved through the M-SSA method manifests as a distinct pattern in the filter weights, notably influencing the rate of decay across increasing lags.
 The resulting filtered series, presented in Figure \ref{mssa_msse_now}, are standardized to enable direct comparisons.
-The two M-SSA predictors adhere to the specified constraints: one exhibiting approximately one-third fewer zero-crossings than the M-MSE, and the other matching the zero-crossing rate of the HP-C, confirming the effectiveness of the smoothness control imposed during the design.
-This is further corroborated by Table \ref{corhtnow}, which reports empirical holding times and target correlations.
-In this example, both the univariate design and the M-SSA method, when configured to replicate its holding time, demonstrate the highest level of smoothness, as indicated by the maximal duration between crossings.
-However, the HP-C also attains the lowest target correlation among the compared approaches.
-More generally, our example illustrates the prediction dilemma underlying the (M-)SSA criterion, highlighting the inherent trade-off between smoothness (zero-crossing rate) and predictive accuracy (target correlation).
+The M-SSA nowcast adheres to the specified constraint, exhibiting approximately one-third fewer zero-crossings than the M-MSE, confirming the effectiveness of the smoothness control imposed during the design.
+This is further corroborated by Table \ref{corhtnow}, which reports empirical holding times and target correlations: increased smoothness is obtained by a slight degradation of tracking accuracy, highlighting the inherent trade-off (prediction dilemma) between smoothness (zero-crossing rate) and predictive accuracy (target correlation) in the M-SSA criterion.
 <<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
@@
 cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
-cat("\\caption{Multivariate M-MSE (left) and M-SSA  (middle:smaller HT, right: larger HT) nowcast filters.", sep = "")
+cat("\\caption{Multivariate M-MSE (left) and M-SSA  (right) nowcast filters.", sep = "")
 cat("\\label{bk_gammak}}", sep = "")
@@
 cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
-cat("\\caption{Nowcasting: two-sided HP(160) applied to BIP (target, black), univariate HP-C (violet), M-SSA smaller HT (blue), M-SSA larger HT (orange) and M-MSE (green).
+cat("\\caption{Nowcasting: two-sided HP(160) applied to BIP (target, black),  M-SSA (blue) and and M-MSE (green).
 The dashed vertical line delimits in- and out-of-sample spans.
@@
 xtable(mat1, dec = 1,digits=rep(2,dim(mat1)[2]+1),
-paste("Target correlations and HTs of HP-C, M-SSA and M-MSE nowcasts.  "),
+paste("Target correlations and HTs of M-SSA and M-MSE nowcasts.  "),
 label=paste("corhtnow",sep=""),

### 2025-05-07 Latex/BIP_predictor.tex
up

@@
 A comparison with the classic direct forecasts in Fig.\eqref{direct_wc_financial_crisis} indicates that the new predictors are slightly less retarded (`faster').
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/direct_hp_forecasts.pdf}\caption{Direct HP forecasts: entire data set with exclusion of the pandemic.\label{direct_hp_forecasts}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/direct_hp_forecasts_financial_crisis.pdf}\caption{Direct HP forecasts over the curse of the financial crisis.\label{direct_hp_forecasts_financial_crisis}}\end{center}\end{figure}Moreover, the F-statistics are larger than for the direct forecasts, see Table \eqref{f_stat}, but the predictor remains statistically insignificant at forecast horizons exceeding two quarters (even though the evaluation is conducted in-sample and based on all available indicators).
+\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/direct_hp_forecasts.pdf}\caption{Direct HP forecasts: entire data set with exclusion of the pandemic.\label{direct_hp_forecasts}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/direct_hp_forecasts_financial_crisis.pdf}\caption{Direct HP forecasts over the curse of the financial crisis:  regression based on all indicators and full sample information.\label{direct_hp_forecasts_financial_crisis}}\end{center}\end{figure}Moreover, the F-statistics are larger than for the direct forecasts, see Table \eqref{f_stat}, but the predictor remains statistically insignificant at forecast horizons exceeding two quarters (even though the evaluation is conducted in-sample and based on all available indicators).
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:45:24 2025
+% Wed May  7 08:46:33 2025
 \begin{table}[ht]
@@
 The first L=31 observations are lost due to initialization and the two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:45:24 2025
+% Wed May  7 08:46:33 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:45:30 2025
+% Wed May  7 08:46:38 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:45:30 2025
+% Wed May  7 08:46:38 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:45:31 2025
+% Wed May  7 08:46:39 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:45:31 2025
+% Wed May  7 08:46:39 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:45:32 2025
+% Wed May  7 08:46:39 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:45:32 2025
+% Wed May  7 08:46:39 2025
 \begin{table}[ht]

### 2025-05-07 Sweave/BIP_predictor.Rnw
up

@@
 cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
-cat("\\caption{Direct HP forecasts over the curse of the financial crisis.", sep = "")
+cat("\\caption{Direct HP forecasts over the curse of the financial crisis:  regression based on all indicators and full sample information.", sep = "")
 cat("\\label{direct_hp_forecasts_financial_crisis}}", sep = "")

### 2025-05-07 Latex/BIP_predictor.tex
up

@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:46:33 2025
+% Wed May  7 08:51:37 2025
 \begin{table}[ht]
@@
 The M-SSA nowcast adheres to the specified constraint, exhibiting approximately one-third fewer zero-crossings than the M-MSE, confirming the effectiveness of the smoothness control imposed during the design.
-This is further corroborated by Table \ref{corhtnow}, which reports empirical holding times and target correlations: increased smoothness is obtained by a slight degradation of tracking accuracy, highlighting the inherent trade-off (prediction dilemma) between smoothness (zero-crossing rate) and predictive accuracy (target correlation) in the M-SSA criterion.
+This is further corroborated by Table \ref{corhtnow}, which reports empirical holding times and target correlations.
+%: increased smoothness is obtained by a slight degradation of tracking accuracy, highlighting the inherent trade-off (prediction dilemma) between smoothness (zero-crossing rate) and predictive accuracy (target correlation) in the M-SSA criterion.
 \begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/bk_gammak.pdf}\caption{Multivariate M-MSE (left) and M-SSA  (right) nowcast filters.\label{bk_gammak}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/mssa_msse_now.pdf}\caption{Nowcasting: two-sided HP(160) applied to BIP (target, black),  M-SSA (blue) and and M-MSE (green).
@@
 The first L=31 observations are lost due to initialization and the two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:46:33 2025
+% Wed May  7 08:51:38 2025
 \begin{table}[ht]
@@
 \end{tabular}
-\caption{Target correlations and HTs of M-SSA and M-MSE nowcasts.  } 
+\caption{Sample target correlations and HTs of M-SSA and M-MSE nowcasts.  } 
 \label{corhtnow}
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:46:38 2025
+% Wed May  7 08:51:42 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:46:38 2025
+% Wed May  7 08:51:43 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:46:38 2025
+% Wed May  7 08:51:43 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:46:39 2025
+% Wed May  7 08:51:44 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:46:39 2025
+% Wed May  7 08:51:44 2025
 \begin{table}[ht]

### 2025-05-07 Sweave/BIP_predictor.Rnw
up

@@
 The M-SSA nowcast adheres to the specified constraint, exhibiting approximately one-third fewer zero-crossings than the M-MSE, confirming the effectiveness of the smoothness control imposed during the design.
-This is further corroborated by Table \ref{corhtnow}, which reports empirical holding times and target correlations: increased smoothness is obtained by a slight degradation of tracking accuracy, highlighting the inherent trade-off (prediction dilemma) between smoothness (zero-crossing rate) and predictive accuracy (target correlation) in the M-SSA criterion.
+This is further corroborated by Table \ref{corhtnow}, which reports empirical holding times and target correlations.
+%: increased smoothness is obtained by a slight degradation of tracking accuracy, highlighting the inherent trade-off (prediction dilemma) between smoothness (zero-crossing rate) and predictive accuracy (target correlation) in the M-SSA criterion.
 <<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
@@
 xtable(mat1, dec = 1,digits=rep(2,dim(mat1)[2]+1),
-paste("Target correlations and HTs of M-SSA and M-MSE nowcasts.  "),
+paste("Sample target correlations and HTs of M-SSA and M-MSE nowcasts.  "),
 label=paste("corhtnow",sep=""),

### 2025-05-07 Latex/BIP_predictor.tex
up

@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:51:37 2025
+% Wed May  7 08:57:13 2025
 \begin{table}[ht]
@@
 %: increased smoothness is obtained by a slight degradation of tracking accuracy, highlighting the inherent trade-off (prediction dilemma) between smoothness (zero-crossing rate) and predictive accuracy (target correlation) in the M-SSA criterion.
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/bk_gammak.pdf}\caption{Multivariate M-MSE (left) and M-SSA  (right) nowcast filters.\label{bk_gammak}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/mssa_msse_now.pdf}\caption{Nowcasting: two-sided HP(160) applied to BIP (target, black),  M-SSA (blue) and and M-MSE (green).
+\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/bk_gammak.pdf}\caption{Multivariate M-MSE (left) and M-SSA  (right) nowcast filters.\label{bk_gammak}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/mssa_msse_now.pdf}\caption{Two-sided HP(160) applied to BIP (target, black) and nowcasts:  M-SSA (blue) and  M-MSE (green).
 The dashed vertical line delimits in- and out-of-sample spans.
 The first L=31 observations are lost due to initialization and the two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:51:38 2025
+% Wed May  7 08:57:14 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:51:42 2025
+% Wed May  7 08:57:19 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:51:43 2025
+% Wed May  7 08:57:19 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:51:43 2025
+% Wed May  7 08:57:19 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:51:44 2025
+% Wed May  7 08:57:20 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:51:44 2025
+% Wed May  7 08:57:20 2025
 \begin{table}[ht]

### 2025-05-07 Sweave/BIP_predictor.Rnw
up

@@
 cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
-cat("\\caption{Nowcasting: two-sided HP(160) applied to BIP (target, black),  M-SSA (blue) and and M-MSE (green).
+cat("\\caption{Two-sided HP(160) applied to BIP (target, black) and nowcasts:  M-SSA (blue) and  M-MSE (green).
 The dashed vertical line delimits in- and out-of-sample spans.

### 2025-05-07 Latex/BIP_predictor.tex
up

@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:57:13 2025
+% Wed May  7 11:13:49 2025
 \begin{table}[ht]
@@
 This is further corroborated by Table \ref{corhtnow}, which reports empirical holding times and target correlations.
-%: increased smoothness is obtained by a slight degradation of tracking accuracy, highlighting the inherent trade-off (prediction dilemma) between smoothness (zero-crossing rate) and predictive accuracy (target correlation) in the M-SSA criterion.
+Increased smoothness achieved by the M-SSA comes at the cost of a slight reduction in tracking accuracy, as indicated by a marginally lower target correlation.
+This highlights the inherent trade-off—often referred to as the prediction dilemma—between smoothness (zero-crossing rate) and predictive accuracy (target correlation) within the M-SSA criterion.
 \begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/bk_gammak.pdf}\caption{Multivariate M-MSE (left) and M-SSA  (right) nowcast filters.\label{bk_gammak}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/mssa_msse_now.pdf}\caption{Two-sided HP(160) applied to BIP (target, black) and nowcasts:  M-SSA (blue) and  M-MSE (green).
@@
 The first L=31 observations are lost due to initialization and the two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:57:14 2025
+% Wed May  7 11:13:50 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:57:19 2025
+% Wed May  7 11:13:55 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:57:19 2025
+% Wed May  7 11:13:56 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:57:19 2025
+% Wed May  7 11:13:56 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:57:20 2025
+% Wed May  7 11:13:57 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 08:57:20 2025
+% Wed May  7 11:13:57 2025
 \begin{table}[ht]

### 2025-05-07 Sweave/BIP_predictor.Rnw
up

@@
 #------------------------------
-
+# Section 2.1: data
 recompute_results<-F
@@
 #--------------------------------------------------------
-# Section 2: dependence
+# Section 2.2: dependence
 
@@
 This is further corroborated by Table \ref{corhtnow}, which reports empirical holding times and target correlations.
-%: increased smoothness is obtained by a slight degradation of tracking accuracy, highlighting the inherent trade-off (prediction dilemma) between smoothness (zero-crossing rate) and predictive accuracy (target correlation) in the M-SSA criterion.
+Increased smoothness achieved by the M-SSA comes at the cost of a slight reduction in tracking accuracy, as indicated by a marginally lower target correlation.
+This highlights the inherent trade-off—often referred to as the prediction dilemma—between smoothness (zero-crossing rate) and predictive accuracy (target correlation) within the M-SSA criterion.
 <<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=

### 2025-05-07 Latex/BIP_predictor.tex
up

@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 11:13:49 2025
+% Wed May  7 11:21:15 2025
 \begin{table}[ht]
@@
 The first L=31 observations are lost due to initialization and the two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 11:13:50 2025
+% Wed May  7 11:21:16 2025
 \begin{table}[ht]
@@
 M-SSA (bottom) across the financial crisis (left) and the pandemic (right).\label{multivar_vs_univar}}\end{center}\end{figure}
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/mssa_hpc_financial_crisis.pdf}\caption{Forward-shifted BIP (black), M-SSA (blue) and HP-C (red) for $h=3,4,5,6$ across the financial crisis.\label{mssa_hpc_financial_crisis}}\end{center}\end{figure}
+\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/mssa_hpc_financial_crisis.pdf}\caption{Forward-shifted BIP (black), M-SSA (blue) and HP-C (red) for $h=3,4,5,6$ across the financial crisis.
+All series are standardized for better visual inspection.\label{mssa_hpc_financial_crisis}}\end{center}\end{figure}
 
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 11:13:55 2025
+% Wed May  7 11:21:19 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 11:13:56 2025
+% Wed May  7 11:21:19 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 11:13:56 2025
+% Wed May  7 11:21:19 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 11:13:57 2025
+% Wed May  7 11:21:20 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 11:13:57 2025
+% Wed May  7 11:21:20 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 11:13:57 2025
+% Wed May  7 11:21:21 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 11:13:57 2025
+% Wed May  7 11:21:21 2025
 \begin{table}[ht]

### 2025-05-07 Sweave/BIP_predictor.Rnw
up

@@
 cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
-cat("\\caption{Forward-shifted BIP (black), M-SSA (blue) and HP-C (red) for $h=3,4,5,6$ across the financial crisis.", sep = "")
+cat("\\caption{Forward-shifted BIP (black), M-SSA (blue) and HP-C (red) for $h=3,4,5,6$ across the financial crisis. All series are standardized for better visual inspection.", sep = "")
 cat("\\label{mssa_hpc_financial_crisis}}", sep = "")

### 2025-05-07 Latex/BIP_predictor.tex
up

@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 11:21:15 2025
+% Wed May  7 12:49:46 2025
 \begin{table}[ht]
@@
 The first L=31 observations are lost due to initialization and the two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 11:21:16 2025
+% Wed May  7 12:49:47 2025
 \begin{table}[ht]
@@
 The previously described nowcast example can be systematically extended to generate forecasts of the target variable.
-This extension utilizes the same empirical framework, incorporating an assumed $50\%$ increase in the HT over the M-MSE, and considers forward shifts of $h=1,...,6$  quarters of HP-BIP (note that M-SSA does not target BIP explicitly).
+This extension utilizes the same empirical framework, incorporating an assumed $50\%$ increase in the HT over the M-MSE, and considers forward shifts of $h=0,...,6$  quarters of HP-BIP (note that M-SSA does not target BIP explicitly).
 %This approach facilitates the assessment of forecast performance across multiple horizons within the established methodological context.
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 11:21:19 2025
+% Wed May  7 12:49:50 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 11:21:19 2025
+% Wed May  7 12:49:50 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 11:21:20 2025
+% Wed May  7 12:49:51 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 11:21:20 2025
+% Wed May  7 12:49:51 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 11:21:21 2025
+% Wed May  7 12:49:51 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 11:21:21 2025
+% Wed May  7 12:49:51 2025
 \begin{table}[ht]

### 2025-05-07 Sweave/BIP_predictor.Rnw
up

@@
 # Set of performance metrics and tests of unequal predictability
-#source(paste(getwd(),"/R/performance_statistics_functions.r",sep=""))
 
@@
 # 4. Forecast horizon h
+select<-"BIP"
 bk<-matrix(mssa_indicator_obj$bk_x_array[select,select,,"h=0"],nrow=L)

### 2025-05-08 Sweave/BIP_predictor.Rnw
up

@@
 source(paste(getwd(),"/R/M_SSA_utility_functions.r",sep=""))
-# Set of performance metrics and tests of unequal predictability
+source(paste(getwd(),"/R/M_SSA_paper_functions.r",sep=""))
 
@@
 len<-dim(x_mat)[1]
+
+# Read original data 
+h0<-0
+trim_threshold<-3
+na_rw_lin<-T
+data_file_name<-c("Data_HWI_2025_02.csv","gdp_2025_02.csv")
+select_vec<-select_vec_multi
+diff_lag<-3
+
+read_obj<-read_data_func_monthly(path.data,select_vec,h0,trim_threshold,na_rw_lin,data_file_name,diff_lag)
+data_table<-read_obj$data_select_unshifted_xts
+colnames(data_table)[1]<-"GDP"
+
+tail(data_table)
 
@@
 \end{itemize}
+
+<<label=ats_mba_2,echo=FALSE,results=tex>>=
+
+mat1<-as.matrix(tail(data_table))
+#latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
+#center = "centering", file = "", floating = FALSE)
+xtable(mat1, dec = 1,digits=rep(3,dim(mat1)[2]+1),
+paste("Ragged end.  "),
+label=paste("f_stat",sep=""),
+center = "centering", file = "", floating = FALSE)
+
+math<-as.matrix(tail(data_table,10))
+mat2<-cbind(math[2:7,1],math[3:8,2],math[5:10,3:5])
+colnames(mat2)<-colnames(mat1)
+#latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
+#center = "centering", file = "", floating = FALSE)
+xtable(mat2, dec = 1,digits=rep(3,dim(mat1)[2]+1),
+paste("Data aligned at sample end."),
+label=paste("f_stat",sep=""),
+center = "centering", file = "", floating = FALSE)
+
+@
 

### 2025-05-08 R/M_SSA_paper_functions.R

@@ -1,0 +1,2152 @@
+
+
+
+read_data_func<-function(path.data,select_vec,h,trim_threshold,na_rw_lin,data_file_name,quarterly_TF,diff_lag)
+{
+  if (quarterly_TF)
+  {
+    read_obj<-read_data_func_quarterly(path.data,select_vec,h,trim_threshold,na_rw_lin,data_file_name,diff_lag)
+      
+  } else
+  {
+    read_obj<-read_data_func_monthly(path.data,select_vec,h,trim_threshold,na_rw_lin,data_file_name,diff_lag)
+    
+  }
+  
+  # Main data set for computing predictors
+  data=read_obj$data
+  # Same as above but xts object: can play with dates and shift data
+  data_xts=read_obj$data_xts
+  # Same as above but without transformation: data in levels, no trimming, but real-time with target 
+  data_select_xts_filled=read_obj$data_select_xts_filled 
+  # Same as above but data not shifted, no interpolation of BIP (if selected)
+  data_select_unshifted_xts=read_obj$data_select_unshifted_xts
+  # Original data set, i.e., all columns, xts format
+  # Same as above but selected columns only
+  data_select_xts=read_obj$data_select_xts
+  # Same as data above but without standardization: used for calibration of predictor
+  returns_unscaled=read_obj$returns_unscaled
+  # Vector of publication lags
+  lag_vec=read_obj$lag_vec
+# Lags of series in months (in original monthly/quarterly data sets): typically, BIP and ip are lagged   
+  lag_data=read_obj$lag_data
+  
+  ts.plot(data)
+  
+
+  return(list(data=data,data_xts=data_xts,data_select_xts_filled=data_select_xts_filled,data_select_unshifted_xts=data_select_unshifted_xts,data_select_xts=data_select_xts,lag_vec=lag_vec,returns_unscaled=returns_unscaled,lag_data=lag_data))
+}  
+  
+
+# Monthly data sample
+read_data_func_monthly<-function(path.data,select_vec,h,trim_threshold,na_rw_lin,data_file_name,diff_lag)
+{
+  # 1 Data
+  # Read data
+  data_all<-read.csv(paste(path.data,data_file_name[1],sep=""))
+  tail(data_all)
+  tail(data_all)[,"ifo_exp"]
+  
+  
[…]

### 2025-05-08 Latex/BIP_predictor.tex
up

@@
 
+% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Thu May  8 07:34:02 2025
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrrr}
+  \hline
+ & GDP & ip & ifo\_c & ESI & spr\_10y\_3m \\ 
+  \hline
+2024-08-01 &  & 93.200 & 84.900 & 90.900 & -1.400 \\ 
+  2024-09-01 &  & 91.200 & 83.700 & 89.800 & -1.400 \\ 
+  2024-10-01 & 900.761 & 90.800 & 84.100 & 90.700 & -1.200 \\ 
+  2024-11-01 &  & 92.200 & 83.800 & 89.300 & -0.900 \\ 
+  2024-12-01 &  &  & 82.800 & 86.900 & -0.800 \\ 
+  2025-01-01 &  &  & 82.500 & 88.100 & -0.200 \\ 
+   \hline
+\end{tabular}
+\caption{Ragged end.  } 
+\label{f_stat}
+\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Thu May  8 07:34:02 2025
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrrr}
+  \hline
+ & GDP & ip & ifo\_c & ESI & spr\_10y\_3m \\ 
+  \hline
+2024-05-01 &  & 93.400 & 84.900 & 90.900 & -1.400 \\ 
+  2024-06-01 &  & 90.700 & 83.700 & 89.800 & -1.400 \\ 
+  2024-07-01 & 902.571 & 93.200 & 84.100 & 90.700 & -1.200 \\ 
+  2024-08-01 &  & 91.200 & 83.800 & 89.300 & -0.900 \\ 
+  2024-09-01 &  & 90.800 & 82.800 & 86.900 & -0.800 \\ 
+  2024-10-01 & 900.761 & 92.200 & 82.500 & 88.100 & -0.200 \\ 
+   \hline
+\end{tabular}
+\caption{Data aligned at sample end.} 
+\label{f_stat}
+\end{table}
+
 The transformed quarterly series are displayed in Fig.
@@
 \ref{CCF} (entire dataset) and \ref{CCF_wc} (data prior to the pandemic).
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/CCF.pdf}\caption{Sample cross-correlation function (CCF): full data set.\label{CCF}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/CCF_wc.pdf}\caption{Sample cross-correlation function (CCF): without Pandemic.\label{CCF_wc}}\end{center}\end{figure}The CCF confirms the leads and lags illustrated in Fig.
+\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/CCF.pdf}\caption{Sample cross-correlation function: GDP and ifo (left), GDP and ESI (right): full data set, including the pandemic.
+Correlations at positive lags mean that real-time GDP is lagging in relative terms.\label{CCF}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/CCF_wc.pdf}\caption{Sample cross-correlation function: GDP and ifo (left), GDP and ESI (right): without Pandemic.
+Correlations at positive lags mean that real-time GDP is lagging in relative terms.\label{CCF_wc}}\end{center}\end{figure}The CCF confirms the leads and lags illustrated in Fig.
 \ref{data_lags}, although the strength of the dependence is influenced by the singular readings during the pandemic\footnote{We discard the corresponding data when finalizing the predictor.}.
@@
 \ref{impulse} (full data set) and \eqref{impulse_wc} (prior the pandemic).
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/ma_inv_multi_ip.pdf}\caption{MA inversion of VAR(1): full data set.\label{impulse}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/ma_inv_multi_ip_wc.pdf}\caption{MA inversion of VAR(1): without Pandemic.\label{impulse_wc}}\end{center}\end{figure}The model summarizes key features of the data, where lagging series (GDP and industrial production, when accounting for publication lags) depend on  indicators leading in relative terms.
+\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/ma_inv_multi_ip.pdf}\caption{MA inversion of VAR(1): full data set.\label{impulse}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/ma_inv_multi_ip_wc.pdf}\caption{MA inversion of VAR(1): without Pandemic.\label{impulse_wc}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=6in]{./Figures/ma_inv_BIP.pdf}\caption{MA inversion of VAR(1) for GDP: with pandemic (left) and without pandemic (right).\label{ma_inv_BIP}}\end{center}\end{figure}
+
+The model summarizes key features of the data, where lagging series (GDP and industrial production, when accounting for publication lags) depend on  indicators leading in relative terms.
 In particular, the impulse response of the spread (rightmost panel in the figure) suggests the relevance of a univariate model for this specific leading indicator.
@@
 \ref{direct_wc} (full dataset without the pandemic) and \ref{direct_wc_financial_crisis} (financial crisis) for horizons up to three quarters ahead.
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/direct_wc_all.pdf}\caption{Direct forecasts (red lines) based on all indicators and full sample information, without the pandemic: BIP (black line) shifted by zero one, two and three quarters.\label{direct_wc}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/direct_wc_financial_crisis.pdf}\caption{Direct forecast  across the financial crisis:  regression based on all indicators and full sample information.\label{direct_wc_financial_crisis}}\end{center}\end{figure}The results indicate that as $h$ increases, the peaks and dips of the direct forecasts become increasingly right-shifted (delayed) relative to the target.
+\begin{figure}[H]\begin{center}\includegraphics[height=4in, width=5in]{./Figures/direct_wc_all.pdf}\caption{Direct forecasts (red lines) based on all indicators and full sample information, without the pandemic: BIP (black line) shifted by zero one, two and three quarters.\label{direct_wc}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=4in, width=5in]{./Figures/direct_wc_financial_crisis.pdf}\caption{Direct forecast  across the financial crisis:  regression based on all indicators and full sample information.\label{direct_wc_financial_crisis}}\end{center}\end{figure}The results indicate that as $h$ increases, the peaks and dips of the direct forecasts become increasingly right-shifted (delayed) relative to the target.
 Additionally, the F-statistics precipitously decline from $F=3.57$, for $h=2$, to $F=0.46$, for $h=3$, see Table \eqref{f_stat}, suggesting that the predictor becomes statistically insignificant for larger forecast horizons exceeding two quarters (link to Heinisch and Scheufele).
@@
 A comparison with the classic direct forecasts in Fig.\eqref{direct_wc_financial_crisis} indicates that the new predictors are slightly less retarded (`faster').
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/direct_hp_forecasts.pdf}\caption{Direct HP forecasts: entire data set with exclusion of the pandemic.\label{direct_hp_forecasts}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/direct_hp_forecasts_financial_crisis.pdf}\caption{Direct HP forecasts over the curse of the financial crisis:  regression based on all indicators and full sample information.\label{direct_hp_forecasts_financial_crisis}}\end{center}\end{figure}Moreover, the F-statistics are larger than for the direct forecasts, see Table \eqref{f_stat}, but the predictor remains statistically insignificant at forecast horizons exceeding two quarters (even though the evaluation is conducted in-sample and based on all available indicators).
+\begin{figure}[H]\begin{center}\includegraphics[height=4in, width=5in]{./Figures/direct_hp_forecasts.pdf}\caption{Direct HP forecasts: entire data set with exclusion of the pandemic.\label{direct_hp_forecasts}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=5in, width=6in]{./Figures/direct_hp_forecasts_financial_crisis.pdf}\caption{Shifted GDP (black), direct HP forecast (blue) and classic direct forecast (red) over the curse of the financial crisis:  regression based on all indicators and full sample information (without pandemic).\label{direct_hp_forecasts_financial_crisis}}\end{center}\end{figure}Moreover, the F-statistics are larger than for the direct forecasts, see Table \eqref{f_stat}, but the predictor remains statistically insignificant at forecast horizons exceeding two quarters (even though the evaluation is conducted in-sample and based on all available indicators).
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 12:49:46 2025
+% Thu May  8 07:34:04 2025
 \begin{table}[ht]
@@
 The first L=31 observations are lost due to initialization and the two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 12:49:47 2025
+% Thu May  8 07:34:04 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 12:49:50 2025
+% Thu May  8 07:34:09 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 12:49:50 2025
+% Thu May  8 07:34:09 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 12:49:51 2025
+% Thu May  8 07:34:10 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 12:49:51 2025
+% Thu May  8 07:34:10 2025
 \begin{table}[ht]

### 2025-05-08 Sweave/BIP_predictor.Rnw
up

@@
 pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
-acf(x_mat)
+#acf(x_mat)
+acf_result <- acf(x_mat, plot = FALSE)
+
+# Extract the ACF values and lags
+acf_values <- acf_result$acf
+lags <- acf_result$lag
+
+par(mfrow=c(1,2))
+k<-3
+plot(lags[,1,length(select_vec_multi)],acf_values[,1,k], type = "h", main = paste("GDP vs. ",select_vec_multi[k],sep=""), xlab = "Lag", ylab = "ACF")
+abline(h = 0, col = "red")
+k<-4
+plot(lags[,1,length(select_vec_multi)],acf_values[,1,k], type = "h", main = paste("GDP vs. ",select_vec_multi[k],sep=""), xlab = "Lag", ylab = "ACF")
+abline(h = 0, col = "red")
+
 dev.off()
@@
 pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
-acf(x_mat[which(rownames(acf_mat)<2020),])
+#acf(x_mat[which(rownames(acf_mat)<2020),])
+acf_result <- acf(x_mat_wc, plot = FALSE)
+
+# Extract the ACF values and lags
+acf_values <- acf_result$acf
+lags <- acf_result$lag
+
+par(mfrow=c(1,2))
+k<-3
+plot(lags[,1,length(select_vec_multi)],acf_values[,1,k], type = "h", main = paste("GDP vs. ",select_vec_multi[k],sep=""), xlab = "Lag", ylab = "ACF")
+abline(h = 0, col = "red")
+k<-4
+plot(lags[,1,length(select_vec_multi)],acf_values[,1,k], type = "h", main = paste("GDP vs. ",select_vec_multi[k],sep=""), xlab = "Lag", ylab = "ACF")
+abline(h = 0, col = "red")
+
 dev.off()
@@
 # Without Pandemic  
-  data_fit<-x_mat[which(rownames(x_mat)<date_to_fit),]
+  data_fit<-x_mat_wc
   set.seed(12)
@@
 colo<-rainbow(ncol(x_mat))
-par(mfrow=c(1,n))
-for (i in 1:n)#i<-1
+k<-1
+par(mfrow=c(1,k))
+for (i in 1:k)#i<-1
 {
@@
   }
+  mplot_with_pandemic<-mplot
   ts.plot(mplot,col=colo_i,main=paste("MA inversion ",colnames(data_fit)[i],sep=""),xlab="lag")
@@
   }
+    mplot_without_pandemic<-mplot
+
   ts.plot(mplot,col=colo_i,main=paste("MA inversion ",colnames(data_fit)[i],sep=""),xlab="lag")
@@
 }
+
+dev.off()
+
+file = "./Figures/ma_inv_BIP.pdf"
+pdf(file = paste(path.out,file,sep=""), paper = "special", width = 9, height = 6)
+par(mfrow=c(1,2))
+  ts.plot(mplot_with_pandemic,col=colo_i,main=paste("MA inversion GDP: with pandemic",sep=""),xlab="lag")
+  for (k in 1:length(select_vec_i))
+    mtext(select_vec_i[k],line=-k,col=colo_i[k])
+    ts.plot(mplot_without_pandemic,col=colo_i,main="Without pandemic",xlab="lag")
+#  for (k in 1:length(select_vec_i))
+#    mtext(select_vec_i[k],line=-k,col=colo_i[k])
 
@@
 cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
-cat("\\caption{Sample cross-correlation function (CCF): full data set.", sep = "")
+cat("\\caption{Sample cross-correlation function: GDP and ifo (left), GDP and ESI (right): full data set, including the pandemic.
+Correlations at positive lags mean that real-time GDP is lagging in relative terms.", sep = "")
 cat("\\label{CCF}}", sep = "")
@@
 cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
-cat("\\caption{Sample cross-correlation function (CCF): without Pandemic.", sep = "")
+cat("\\caption{Sample cross-correlation function: GDP and ifo (left), GDP and ESI (right): without Pandemic.
+Correlations at positive lags mean that real-time GDP is lagging in relative terms.", sep = "")
 cat("\\label{CCF_wc}}", sep = "")
@@
 @
+<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
+file = "./Figures/ma_inv_BIP.pdf"
+cat("\\begin{figure}[H]")
+cat("\\begin{center}")
+cat("\\includegraphics[height=3in, width=6in]{", file, "}\n",sep = "")
+cat("\\caption{MA inversion of VAR(1) for GDP: with pandemic (left) and without pandemic (right).", sep = "")
+cat("\\label{ma_inv_BIP}}", sep = "")
+cat("\\end{center}")
+cat("\\end{figure}")
+@
+
 The model summarizes key features of the data, where lagging series (GDP and industrial production, when accounting for publication lags) depend on  indicators leading in relative terms.
@@
 file = "./Figures/direct_wc_all.pdf"
-pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
+pdf(file = paste(path.out,file,sep=""), paper = "special", width = 7.5, height = 6)
 
@@
 file = "./Figures/direct_wc_financial_crisis.pdf"
-pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
+pdf(file = paste(path.out,file,sep=""), paper = "special", width = 7.5, height = 6)
 
@@
 cat("\\begin{center}")
-cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
+cat("\\includegraphics[height=4in, width=5in]{", file, "}\n",sep = "")
 cat("\\caption{Direct forecasts (red lines) based on all indicators and full sample information, without the pandemic: BIP (black line) shifted by zero one, two and three quarters.", sep = "")
@@
 cat("\\begin{center}")
-cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
+cat("\\includegraphics[height=4in, width=5in]{", file, "}\n",sep = "")
 cat("\\caption{Direct forecast  across the financial crisis:  regression based on all indicators and full sample information.", sep = "")
@@
 file = "./Figures/direct_hp_forecasts.pdf"
-pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
+pdf(file = paste(path.out,file,sep=""), paper = "special", width = 7.5, height = 6)
 
@@
 file = "./Figures/direct_hp_forecasts_financial_crisis.pdf"
-pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
+pdf(file = paste(path.out,file,sep=""), paper = "special", width = 7.5, height = 6)
 
@@
   # Shift BIP forward by publication lag+forecast horizon
-  forward_shifted_BIP<-c(x_mat_wc[(lag_vec[1]+h):nrow(x_mat_wc),"BIP"],rep(NA,h+lag_vec[1]-1))
+  forward_shifted_BIP<-c(x_mat_wc[(1+lag_vec[1]+h):nrow(x_mat_wc),"BIP"],rep(NA,h+lag_vec[1]))
   # Regress selected indicators on forward-shifted BIP
@@
 
+    # Shift BIP forward by publication lag+forecast horizon
+  forward_shifted_BIP<-c(x_mat_wc[(1+lag_vec[1]+h):nrow(x_mat_wc),"BIP"],rep(NA,h+lag_vec[1]))
+  # Regress selected indicators on forward-shifted BIP
+  lm_obj<-lm(forward_shifted_BIP~x_mat_wc[,select_direct_indicator])
+  # You will probably not find statistically significant regressors for h>2: BIP is a noisy series
+  sum_obj<-summary(lm_obj)
+  sum_obj$fstatistic  
+  # Technical note: 
+  # -Residuals are subject to heteroscedasticity (crises) and autocorrelation
+  # -Therefore classic OLS tests for statistical significance are biased
+  # -We shall rely on HAC-adjusted p-values further down (R-package sandwich)
+
+  # Compute the predictor: one can rely on the generic R-function predict or compute the predictor manually
+  direct_forecast<-lm_obj$coef[1]+x_mat_wc[,select_direct_indicator]%*%lm_obj$coef[2:(length(select_direct_indicator)+1)]
+
   # We can now plot target and direct forecast: for h>2 the predictor comes close to a flat line centered at zero
-  mplot<-cbind(forward_shifted_BIP,direct_hp_forecast)
+  mplot<-cbind(forward_shifted_BIP,direct_hp_forecast,direct_forecast)
     rownames(mplot)<-rownames(x_mat_wc)
 
-  mplot<-mplot[which(rownames(mplot)>2008&rownames(mplot)<2011),]
-  colnames(mplot)<-c("BIP","Direct HP forecast")
-
-  colo<-c("black","red")
-  main_title<-paste("BIP shifted forward by ",h," quarters",sep="")
+  mplot<-mplot[which(rownames(mplot)>2007&rownames(mplot)<2010),]
+  colnames(mplot)<-c("GDP","Direct HP forecast","Direct forecast")
+
+  colo<-c("black","red","blue")
+  main_title<-paste("GDP shifted forward by ",h," quarters",sep="")
   plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
@@
     lines(mplot[,i],col=colo[i])
+    if (h==0)
     mtext(colnames(mplot)[i],col=colo[i],line=-i)
@@
 cat("\\begin{center}")
-cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
+cat("\\includegraphics[height=4in, width=5in]{", file, "}\n",sep = "")
 cat("\\caption{Direct HP forecasts: entire data set with exclusion of the pandemic.", sep = "")
@@
 cat("\\begin{center}")
-cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
-cat("\\caption{Direct HP forecasts over the curse of the financial crisis:  regression based on all indicators and full sample information.", sep = "")
+cat("\\includegraphics[height=5in, width=6in]{", file, "}\n",sep = "")
+cat("\\caption{Shifted GDP (black), direct HP forecast (red) and classic direct forecast (blue) over the curse of the financial crisis:  regression based on all indicators and full sample information (without pandemic).", sep = "")
 cat("\\label{direct_hp_forecasts_financial_crisis}}", sep = "")

### 2025-05-08 Latex/BIP_predictor.tex
up

@@
 
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 07:34:02 2025
+% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Thu May  8 14:00:18 2025
 \begin{table}[ht]
@@
 \label{f_stat}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 07:34:02 2025
+\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Thu May  8 14:00:18 2025
 \begin{table}[ht]
@@
 \ref{direct_wc} (full dataset without the pandemic) and \ref{direct_wc_financial_crisis} (financial crisis) for horizons up to three quarters ahead.
-\begin{figure}[H]\begin{center}\includegraphics[height=4in, width=5in]{./Figures/direct_wc_all.pdf}\caption{Direct forecasts (red lines) based on all indicators and full sample information, without the pandemic: BIP (black line) shifted by zero one, two and three quarters.\label{direct_wc}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=4in, width=5in]{./Figures/direct_wc_financial_crisis.pdf}\caption{Direct forecast  across the financial crisis:  regression based on all indicators and full sample information.\label{direct_wc_financial_crisis}}\end{center}\end{figure}The results indicate that as $h$ increases, the peaks and dips of the direct forecasts become increasingly right-shifted (delayed) relative to the target.
+\begin{figure}[H]\begin{center}\includegraphics[height=5in, width=6in]{./Figures/direct_wc_all.pdf}\caption{Direct forecasts (red lines) based on all indicators and full sample information, without the pandemic: GDP (black line) shifted by zero one, two and three quarters.\label{direct_wc}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=5in, width=6in]{./Figures/direct_wc_financial_crisis.pdf}\caption{Direct forecast  across the financial crisis:  regression based on all indicators and full sample information.\label{direct_wc_financial_crisis}}\end{center}\end{figure}The results indicate that as $h$ increases, the peaks and dips of the direct forecasts become increasingly right-shifted (delayed) relative to the target.
 Additionally, the F-statistics precipitously decline from $F=3.57$, for $h=2$, to $F=0.46$, for $h=3$, see Table \eqref{f_stat}, suggesting that the predictor becomes statistically insignificant for larger forecast horizons exceeding two quarters (link to Heinisch and Scheufele).
@@
 We now posit the HP(160) filter as a means to achieve better forecasting performance at larger horizons, noting that a comprehensive technical analysis of the effect of $\lambda$ on the resulting BIP predictor can be conducted using the M-SSA package.
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/hp_160.pdf}\caption{Two-sided HP-160) (top left) and one-sided (concurrent) HP-C filters (top right) with corresponding amplitude functions (bottom).\label{hp_160}}\end{center}\end{figure}
+\begin{figure}[H]\begin{center}\includegraphics[height=5in, width=6in]{./Figures/hp_160.pdf}\caption{Two-sided HP-160) (top left) and one-sided (concurrent) HP-C filters (top right) with corresponding amplitude functions (bottom).\label{hp_160}}\end{center}\end{figure}
 \subsection{Direct HP Forecast}\label{hpdf}
@@
 A comparison with the classic direct forecasts in Fig.\eqref{direct_wc_financial_crisis} indicates that the new predictors are slightly less retarded (`faster').
-\begin{figure}[H]\begin{center}\includegraphics[height=4in, width=5in]{./Figures/direct_hp_forecasts.pdf}\caption{Direct HP forecasts: entire data set with exclusion of the pandemic.\label{direct_hp_forecasts}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=5in, width=6in]{./Figures/direct_hp_forecasts_financial_crisis.pdf}\caption{Shifted GDP (black), direct HP forecast (blue) and classic direct forecast (red) over the curse of the financial crisis:  regression based on all indicators and full sample information (without pandemic).\label{direct_hp_forecasts_financial_crisis}}\end{center}\end{figure}Moreover, the F-statistics are larger than for the direct forecasts, see Table \eqref{f_stat}, but the predictor remains statistically insignificant at forecast horizons exceeding two quarters (even though the evaluation is conducted in-sample and based on all available indicators).
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 07:34:04 2025
+\begin{figure}[H]\begin{center}\includegraphics[height=5in, width=6in]{./Figures/direct_hp_forecasts.pdf}\caption{Direct HP forecasts: entire data set with exclusion of the pandemic.\label{direct_hp_forecasts}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=5in, width=6in]{./Figures/direct_hp_forecasts_financial_crisis.pdf}\caption{Shifted GDP (black), direct HP forecast (blue) and classic direct forecast (red) over the curse of the financial crisis:  regression based on all indicators and full sample information (without pandemic).\label{direct_hp_forecasts_financial_crisis}}\end{center}\end{figure}Moreover, the F-statistics are larger than for the direct forecasts, see Table \eqref{f_stat}, but the predictor remains statistically insignificant at forecast horizons exceeding two quarters (even though the evaluation is conducted in-sample and based on all available indicators).
+% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Thu May  8 14:00:18 2025
 \begin{table}[ht]
@@
 This highlights the inherent trade-off—often referred to as the prediction dilemma—between smoothness (zero-crossing rate) and predictive accuracy (target correlation) within the M-SSA criterion.
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/bk_gammak.pdf}\caption{Multivariate M-MSE (left) and M-SSA  (right) nowcast filters.\label{bk_gammak}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/mssa_msse_now.pdf}\caption{Two-sided HP(160) applied to BIP (target, black) and nowcasts:  M-SSA (blue) and  M-MSE (green).
+\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=5in]{./Figures/bk_gammak.pdf}\caption{Multivariate M-MSE (left) and M-SSA  (right) nowcast filters.\label{bk_gammak}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=4in, width=5.5in]{./Figures/mssa_msse_now.pdf}\caption{Two-sided HP(160) applied to GDP (target, black) and nowcasts:  M-SSA (blue) and  M-MSE (green).
 The dashed vertical line delimits in- and out-of-sample spans.
-The first L=31 observations are lost due to initialization and the two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 07:34:04 2025
+The two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Thu May  8 14:00:18 2025
 \begin{table}[ht]
@@
 %In comparison, the effect of phase shifts—arising from the cyclical nature of the HP-filter and manifesting in the HP-C forecasts (shown in the left panels)—seems less systematic.
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/bk_h.pdf}\caption{Multivariate M-SSA: $h=0$ (nowcast, left) and $h=4$ (one year forecast, right).\label{bk_h}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/multivar_vs_univar.pdf}\caption{Now- and forecasts of HP-BIP for $h=0,...,6$: HP-C (top) vs.
+\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=5.5in]{./Figures/bk_h.pdf}\caption{Multivariate M-SSA: $h=0$ (nowcast, left) and $h=4$ (one year forecast, right).\label{bk_h}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=4in, width=5.5in]{./Figures/multivar_vs_univar.pdf}\caption{Now- and forecasts of HP-GDP for $h=0,...,6$: HP-C (top) vs.
 M-SSA (bottom) across the financial crisis (left) and the pandemic (right).\label{multivar_vs_univar}}\end{center}\end{figure}
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/mssa_hpc_financial_crisis.pdf}\caption{Forward-shifted BIP (black), M-SSA (blue) and HP-C (red) for $h=3,4,5,6$ across the financial crisis.
+\begin{figure}[H]\begin{center}\includegraphics[height=4in, width=5.5in]{./Figures/mssa_hpc_financial_crisis.pdf}\caption{Forward-shifted GDP (black), M-SSA (blue) and HP-C (red) for $h=3,4,5,6$ across the financial crisis.
 All series are standardized for better visual inspection.\label{mssa_hpc_financial_crisis}}\end{center}\end{figure}
@@
 Table \eqref{p_val_hpbip_wc} reports HAC-adjusted p-values from regressions of the out-of-sample predictor on HP-BIP, indicating a strong link between the predictor and the target, out-of-sample.
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 07:34:09 2025
+% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Thu May  8 14:00:20 2025
 \begin{table}[ht]
@@
 
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 07:34:09 2025
+% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Thu May  8 14:00:20 2025
 \begin{table}[ht]
@@
 \label{p_val_wc}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 07:34:09 2025
+\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Thu May  8 14:00:20 2025
 \begin{table}[ht]
@@
 This suggests instability and complicates interpretation.
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/revisions1.pdf}\caption{Revisions of the M-SSA BIP component predictor are shown, with the final predictor in blue and the real-time design in red.
+\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/revisions1.pdf}\caption{Revisions of the M-SSA GDP component predictor are shown, with the final predictor in blue and the real-time design in red.
 These revisions result from quarterly updates to the regression equations.
@@
 This suggests that the out-of-sample performance of the M-SSA component predictor may be better than reported in Section \eqref{oosp}, which is based on the fixed and increasingly outdated version of M-SSA.
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/up_dated_ma_inv_multi_ip.pdf}\caption{MA-inversion (impulse response) of VAR: data up to Jan-2008 (left) vs.
+\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=5in]{./Figures/up_dated_ma_inv_multi_ip.pdf}\caption{MA-inversion (impulse response) of VAR: data up to Jan-2008 (left) vs.
 entire data set (right).\label{up_dated_ma_inv_multi_ip}}\end{center}\end{figure}
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/bk_2008_all.pdf}\caption{M-SSA filter tracking HP-BIP at forecast horizon $h=4$: data up to Jan-2008 (left) vs.
+\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=5in]{./Figures/bk_2008_all.pdf}\caption{M-SSA filter tracking HP-GDP at forecast horizon $h=4$: data up to Jan-2008 (left) vs.
 entire data set (right).\label{bk_2008_all}}\end{center}\end{figure}
 
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/revisions3.pdf}\caption{Revisions resulting from updates to the VAR model in M-SSA.
+\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=5in]{./Figures/revisions3.pdf}\caption{Revisions resulting from updates to the VAR model in M-SSA optimized for $h=shift=4$.
 To isolate the effect of the regression revision, all series are standardized.
@@
 
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 07:34:10 2025
+% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Thu May  8 14:00:20 2025
 \begin{table}[ht]
@@
 \label{p_val_wc5}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 07:34:10 2025
+\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Thu May  8 14:00:20 2025
 \begin{table}[ht]
@@
 \label{rRMSE_mSSA_comp_mean6}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 07:34:10 2025
+\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Thu May  8 14:00:20 2025
 \begin{table}[ht]
@@
 \label{rRMSE_mSSA_comp_direct_without_covid7}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 07:34:10 2025
+\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Thu May  8 14:00:20 2025
 \begin{table}[ht]
@@
 \label{p_val1}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 07:34:10 2025
+\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Thu May  8 14:00:20 2025
 \begin{table}[ht]
@@
 \label{rRMSE_mSSA_comp_mean2}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 07:34:10 2025
+\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Thu May  8 14:00:20 2025
 \begin{table}[ht]
@@
 \label{rRMSE_mSSA_comp_direct3}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 07:34:10 2025
+\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Thu May  8 14:00:20 2025
 \begin{table}[ht]

### 2025-05-08 Sweave/BIP_predictor.Rnw
up

@@
   rownames(mplot)<-rownames(x_mat_wc)
-  colnames(mplot)<-c("BIP","Direct forecast")
+  colnames(mplot)<-c("GDP","Direct forecast")
   colo<-c("black","red")
-  main_title<-paste("BIP shifted  by ",h," quarters",sep="")
+  main_title<-paste("GDP shifted  by ",h," quarters",sep="")
   plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
@@
   mplot<-mplot[which(rownames(mplot)>2008&rownames(mplot)<2011),]
-  colnames(mplot)<-c("BIP","Direct forecast")
+  colnames(mplot)<-c("GDP","Direct forecast")
   colo<-c("black","red")
-  main_title<-paste("BIP shifted forward by ",h," quarters",sep="")
+  main_title<-paste("GDP shifted forward by ",h," quarters",sep="")
   plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
@@
 cat("\\begin{center}")
-cat("\\includegraphics[height=4in, width=5in]{", file, "}\n",sep = "")
-cat("\\caption{Direct forecasts (red lines) based on all indicators and full sample information, without the pandemic: BIP (black line) shifted by zero one, two and three quarters.", sep = "")
+cat("\\includegraphics[height=5in, width=6in]{", file, "}\n",sep = "")
+cat("\\caption{Direct forecasts (red lines) based on all indicators and full sample information, without the pandemic: GDP (black line) shifted by zero one, two and three quarters.", sep = "")
 cat("\\label{direct_wc}}", sep = "")
@@
 cat("\\begin{center}")
-cat("\\includegraphics[height=4in, width=5in]{", file, "}\n",sep = "")
+cat("\\includegraphics[height=5in, width=6in]{", file, "}\n",sep = "")
 cat("\\caption{Direct forecast  across the financial crisis:  regression based on all indicators and full sample information.", sep = "")
@@
 file = "./Figures/hp_160.pdf"
-pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
+pdf(file = paste(path.out,file,sep=""), paper = "special", width = 7.5, height = 6)
 par(mfrow=c(2,2))
@@
 cat("\\begin{center}")
-cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
+cat("\\includegraphics[height=5in, width=6in]{", file, "}\n",sep = "")
 cat("\\caption{Two-sided HP-160) (top left) and one-sided (concurrent) HP-C filters (top right) with corresponding amplitude functions (bottom).", sep = "")
@@
 
-  colnames(mplot)<-c("BIP","Direct forecast")
-  colo<-c("black","red")
-  main_title<-paste("BIP shifted forward by ",h," quarters",sep="")
+  colnames(mplot)<-c("GDP","Direct forecast")
+  colo<-c("black","blue")
+  main_title<-paste("GDP shifted forward by ",h," quarters",sep="")
   plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
@@
 
-  colo<-c("black","red","blue")
+  colo<-c("black","blue","red")
   main_title<-paste("GDP shifted forward by ",h," quarters",sep="")
@@
 cat("\\begin{center}")
-cat("\\includegraphics[height=4in, width=5in]{", file, "}\n",sep = "")
+cat("\\includegraphics[height=5in, width=6in]{", file, "}\n",sep = "")
 cat("\\caption{Direct HP forecasts: entire data set with exclusion of the pandemic.", sep = "")
@@
 cat("\\includegraphics[height=5in, width=6in]{", file, "}\n",sep = "")
-cat("\\caption{Shifted GDP (black), direct HP forecast (red) and classic direct forecast (blue) over the curse of the financial crisis:  regression based on all indicators and full sample information (without pandemic).", sep = "")
+cat("\\caption{Shifted GDP (black), direct HP forecast (blue) and classic direct forecast (red) over the curse of the financial crisis:  regression based on all indicators and full sample information (without pandemic).", sep = "")
 cat("\\label{direct_hp_forecasts_financial_crisis}}", sep = "")
@@
 file = "./Figures/bk_gammak.pdf"
-pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
+pdf(file = paste(path.out,file,sep=""), paper = "special", width = 7, height = 7)
 
@@
 mplot<-gammak
+colnames(mplot)[1]<-"GDP"
 colo<-rainbow(ncol(mplot))
@@
 mplot<-bk
+colnames(mplot)[1]<-"GDP"
+
 colo<-rainbow(ncol(mplot))
 ts.plot(mplot,col=colo,main="M-SSA",xlab="")
-for (i in 1:ncol(mplot))
-  mtext(colnames(mplot)[i],col=colo[i],line=-i)
+#for (i in 1:ncol(mplot))
+#  mtext(colnames(mplot)[i],col=colo[i],line=-i)
 #mplot<-bk_HP_HP_C
@@
 file = "./Figures/mssa_msse_now.pdf"
-pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
+pdf(file = paste(path.out,file,sep=""), paper = "special", width = 7, height = 7)
 
@@
 dim(mplot)
-colnames(mplot)<-c("Target: HP-BIP","M-SSA","M-MSE")
+colnames(mplot)<-c("Target: HP-GDP","M-SSA","M-MSE")
 
 rownames(mplot)<-rownames(x_mat)
+mplot<-mplot[L:nrow(mplot),]
 colo<-c("black","violet","blue","orange","green")
 colo<-c("black","blue","green")
-main_title<-"Nowcasting HP-BIP"
+main_title<-"Nowcasting HP-GDP"
 plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],lwd=1,ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
@@
 
-mat_cor_ht<-rbind(c(cor(na.exclude(mplot[,c("Target: HP-BIP","M-SSA")]))[1,2],
-cor(na.exclude(mplot[,c("Target: HP-BIP","M-MSE")]))[1,2]),
+mat_cor_ht<-rbind(c(cor(na.exclude(mplot[,c("Target: HP-GDP","M-SSA")]))[1,2],
+cor(na.exclude(mplot[,c("Target: HP-GDP","M-MSE")]))[1,2]),
 c(compute_empirical_ht_func(na.exclude(mssa[,select]))$empirical_ht,
@@
 cat("\\begin{center}")
-cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
+cat("\\includegraphics[height=3in, width=5in]{", file, "}\n",sep = "")
 cat("\\caption{Multivariate M-MSE (left) and M-SSA  (right) nowcast filters.", sep = "")
@@
 cat("\\begin{center}")
-cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
-cat("\\caption{Two-sided HP(160) applied to BIP (target, black) and nowcasts:  M-SSA (blue) and  M-MSE (green).
+cat("\\includegraphics[height=4in, width=5.5in]{", file, "}\n",sep = "")
+cat("\\caption{Two-sided HP(160) applied to GDP (target, black) and nowcasts:  M-SSA (blue) and  M-MSE (green).
 The dashed vertical line delimits in- and out-of-sample spans.
-The first L=31 observations are lost due to initialization and the two-sided filter does not extend to the sample end.", sep = "")
+The two-sided filter does not extend to the sample end.", sep = "")
 cat("\\label{mssa_msse_now}}", sep = "")
@@
 file = "./Figures/bk_h.pdf"
-pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
+pdf(file = paste(path.out,file,sep=""), paper = "special", width = 8, height = 6)
 
@@
 colo<-rainbow(ncol(mplot))
-ts.plot(mplot,col=colo,main="M-SSA: h=0")
+ts.plot(mplot,col=colo,main="M-SSA: h=0",xlab="")
 for (i in 1:ncol(mplot))
@@
 colo<-rainbow(ncol(mplot))
-ts.plot(mplot,col=colo,main="M-SSA: h=4")
-for (i in 1:ncol(mplot))
-  mtext(colnames(mplot)[i],col=colo[i],line=-i)
+ts.plot(mplot,col=colo,main="M-SSA: h=4",xlab="")
+#for (i in 1:ncol(mplot))
+#  mtext(colnames(mplot)[i],col=colo[i],line=-i)
 
@@
 file = "./Figures/multivar_vs_univar.pdf"
-pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
+pdf(file = paste(path.out,file,sep=""), paper = "special", width = 8, height = 8)
 
@@
 mplot<-scale(hp_c_array[i,,])
-colnames(mplot)<-paste("HP-",colnames(x_mat)[i],": h=",h_vec,sep="")
+colnames(mplot)<-paste("h=",h_vec,sep="")
 par(mfrow=c(2,2))
@@
 plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
-mtext(colnames(mplot)[1],col=colo[1],line=-1)
+#mtext(colnames(mplot)[1],col=colo[1],line=-1)
 for (j in 1:ncol(mplot))
@@
   lines(mplot[,j],col=colo[j],lwd=1,lty=1)
-  mtext(colnames(mplot)[j],col=colo[j],line=-j)
+#  mtext(colnames(mplot)[j],col=colo[j],line=-j)
 }
@@
 file = "./Figures/mssa_hpc_financial_crisis.pdf"
-pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
+pdf(file = paste(path.out,file,sep=""), paper = "special", width = 8, height = 8)
 
@@
   colo<-c("black","blue","red")
-  main_title<-paste("BIP shifted forward by ",h," quarters",sep="")
+  main_title<-paste("GDP shifted forward by ",h," quarters",sep="")
   plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
@@
     lines(mplot[,i],col=colo[i])
-    mtext(colnames(mplot)[i],col=colo[i],line=-i)
+#    mtext(colnames(mplot)[i],col=colo[i],line=-i)
   }
@@
 cat("\\begin{center}")
-cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
+cat("\\includegraphics[height=3in, width=5.5in]{", file, "}\n",sep = "")
 cat("\\caption{Multivariate M-SSA: $h=0$ (nowcast, left) and $h=4$ (one year forecast, right).", sep = "")
@@
 cat("\\begin{center}")
-cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
-cat("\\caption{Now- and forecasts of HP-BIP for $h=0,...,6$: HP-C (top) vs. M-SSA (bottom) across the financial crisis (left) and the pandemic (right).", sep = "")
+cat("\\includegraphics[height=4in, width=5.5in]{", file, "}\n",sep = "")
+cat("\\caption{Now- and forecasts of HP-GDP for $h=0,...,6$: HP-C (top) vs. M-SSA (bottom) across the financial crisis (left) and the pandemic (right).", sep = "")
 cat("\\label{multivar_vs_univar}}", sep = "")
@@
 cat("\\begin{center}")
-cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
-cat("\\caption{Forward-shifted BIP (black), M-SSA (blue) and HP-C (red) for $h=3,4,5,6$ across the financial crisis. All series are standardized for better visual inspection.", sep = "")
+cat("\\includegraphics[height=4in, width=5.5in]{", file, "}\n",sep = "")
+cat("\\caption{Forward-shifted GDP (black), M-SSA (blue) and HP-C (red) for $h=3,4,5,6$ across the financial crisis. All series are standardized for better visual inspection.", sep = "")
 cat("\\label{mssa_hpc_financial_crisis}}", sep = "")
@@
 mplot<-cbind(final_components_preditor_array[shift,h,],oos_components_preditor_array[shift,h,])
+mplot<-mplot[L:nrow(mplot),]
 colnames(mplot)<-c("Final predictor","Real-time out-of-sample predictor")
@@
 mplot<-track_weights_array[shift,h,,]
+mplot<-mplot[(L:nrow(mplot)),]
+#mplot<-t(t(mplot)/mplot[nrow(mplot),])
 #colnames(mplot)[2:ncol(track_weights)]<-paste("Weight of M-SSA component ",colnames(track_weights)[2:ncol(track_weights)])
@@
 cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
-cat("\\caption{Revisions of the M-SSA BIP component predictor are shown, with the final predictor in blue and the real-time design in red.
+cat("\\caption{Revisions of the M-SSA GDP component predictor are shown, with the final predictor in blue and the real-time design in red.
 These revisions result from quarterly updates to the regression equations.
@@
 file = "./Figures/revisions3.pdf"
-pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
+pdf(file = paste(path.out,file,sep=""), paper = "special", width = 7, height = 7)
 par(mfrow=c(1,1))
@@
 mplot<-scale(mssa_final_vs_prem_wc)
+mplot<-mplot[L:nrow(mplot),]
 colnames(mplot)<-colnames(mssa_final_vs_prem_wc)
-main_title<-paste("BIP component predictors: final and 2007-MSSA (h=",h,", shift=",shift,")",sep="")
+main_title<-paste("GDP component predictors: final and 2007-MSSA (h=",h,", shift=",shift,")",sep="")
 plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
@@
 file = "./Figures/up_dated_ma_inv_multi_ip.pdf"
-pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
+pdf(file = paste(path.out,file,sep=""), paper = "special", width = 7, height = 7)
 colo<-rainbow(ncol(x_mat))
@@
 }
-ts.plot(mplot,col=colo_i,main=paste("2008: MA inversion ",colnames(data_fit)[i],sep=""),xlab="lag")
+ts.plot(mplot,col=colo_i,main="MA inversion GDP: 2008",xlab="")
 for (k in 1:length(select_vec_i))
@@
 }
-ts.plot(mplot,col=colo_i,main=paste("Full data: MA inversion ",colnames(data_fit)[i],sep=""),xlab="lag")
-for (k in 1:length(select_vec_i))
-  mtext(select_vec_i[k],line=-k,col=colo_i[k])
+ts.plot(mplot,col=colo_i,main="Full data set",xlab="")
+#for (k in 1:length(select_vec_i))
+#  mtext(select_vec_i[k],line=-k,col=colo_i[k])
 
@@
 file = "./Figures/bk_2008_all.pdf"
-pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
+pdf(file = paste(path.out,file,sep=""), paper = "special", width = 7, height = 7)
 
@@
 colo<-rainbow(ncol(mplot))
-ts.plot(mplot,col=colo,main=paste("M-SSA 2008: h=",h,sep=""))
+ts.plot(mplot,col=colo,main=paste("M-SSA data up to 2008: h=",h,sep=""),xlab="")
 for (i in 1:ncol(mplot))
@@
 colo<-rainbow(ncol(mplot))
-ts.plot(mplot,col=colo,main=paste("M-SSA full data: h=",h,sep=""))
-for (i in 1:ncol(mplot))
-  mtext(colnames(mplot)[i],col=colo[i],line=-i)
+ts.plot(mplot,col=colo,main="Full data set",xlab="")
+#for (i in 1:ncol(mplot))
+#  mtext(colnames(mplot)[i],col=colo[i],line=-i)
 
@@
 cat("\\begin{center}")
-cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
+cat("\\includegraphics[height=3in, width=5in]{", file, "}\n",sep = "")
 cat("\\caption{MA-inversion (impulse response) of VAR: data up to Jan-2008 (left) vs. entire data set (right).", sep = "")
@@
 cat("\\begin{center}")
-cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
-cat("\\caption{M-SSA filter tracking HP-BIP at forecast horizon $h=4$: data up to Jan-2008 (left) vs. entire data set (right).", sep = "")
+cat("\\includegraphics[height=3in, width=5in]{", file, "}\n",sep = "")
+cat("\\caption{M-SSA filter tracking HP-GDP at forecast horizon $h=4$: data up to Jan-2008 (left) vs. entire data set (right).", sep = "")
 cat("\\label{bk_2008_all}}", sep = "")
@@
 cat("\\begin{center}")
-cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
-cat("\\caption{Revisions resulting from updates to the VAR model in M-SSA.
+cat("\\includegraphics[height=3in, width=5in]{", file, "}\n",sep = "")
+cat("\\caption{Revisions resulting from updates to the VAR model in M-SSA optimized for $h=shift=4$.
 To isolate the effect of the regression revision, all series are standardized.

### 2025-05-08 Latex/BIP_predictor.tex
up

@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu May  8 14:00:18 2025
+% Thu May  8 14:11:36 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu May  8 14:00:18 2025
+% Thu May  8 14:11:36 2025
 \begin{table}[ht]
@@
 
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/Data.pdf}\caption{Sample cross-correlation function (CCF): full data set.\label{data}}\end{center}\end{figure}
+\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/Data.pdf}\caption{Data: standardized log-differences: pandemic trimmed to $\pm 3$.\label{data}}\end{center}\end{figure}
 \begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/data_lags.pdf}\caption{Data: leads and lags during financial crisis (left) and Pandemic (right).\label{data_lags}}\end{center}\end{figure}
@@
 \begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/CCF.pdf}\caption{Sample cross-correlation function: GDP and ifo (left), GDP and ESI (right): full data set, including the pandemic.
-Correlations at positive lags mean that real-time GDP is lagging in relative terms.\label{CCF}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/CCF_wc.pdf}\caption{Sample cross-correlation function: GDP and ifo (left), GDP and ESI (right): without Pandemic.
-Correlations at positive lags mean that real-time GDP is lagging in relative terms.\label{CCF_wc}}\end{center}\end{figure}The CCF confirms the leads and lags illustrated in Fig.
+Correlations at positive lags signify that (real-time) GDP is lagging when referenced against the two (real-time) indicators.\label{CCF}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/CCF_wc.pdf}\caption{Sample cross-correlation function: GDP and ifo (left), GDP and ESI (right): without pandemic.
+Correlations at positive lags signify that (real-time) GDP is lagging when referenced against the two (real-time) indicators.\label{CCF_wc}}\end{center}\end{figure}The CCF confirms the leads and lags illustrated in Fig.
 \ref{data_lags}, although the strength of the dependence is influenced by the singular readings during the pandemic\footnote{We discard the corresponding data when finalizing the predictor.}.
@@
 The two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu May  8 14:00:18 2025
+% Thu May  8 14:11:36 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu May  8 14:00:20 2025
+% Thu May  8 14:11:38 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu May  8 14:00:20 2025
+% Thu May  8 14:11:38 2025
 \begin{table}[ht]

### 2025-05-08 Sweave/BIP_predictor.Rnw
up

@@
 mplot<-x_mat
+colnames(mplot)[1]<-"GDP"
 colo<-c("black",rainbow(ncol(data)-1))
@@
 mplot<-x_mat[which(rownames(x_mat)>=2007&rownames(x_mat)<=2011),]
+colnames(mplot)[1]<-"GDP"
 colo<-c("black",rainbow(ncol(data)-1))
@@
 mplot<-x_mat[which(rownames(x_mat)>=2019&rownames(x_mat)<=2022),]
+colnames(mplot)[1]<-"GDP"
 colo<-c("black",rainbow(ncol(data)-1))
@@
 k<-3
-plot(lags[,1,length(select_vec_multi)],acf_values[,1,k], type = "h", main = paste("GDP vs. ",select_vec_multi[k],sep=""), xlab = "Lag", ylab = "ACF")
+plot(lags[,1,length(select_vec_multi)],acf_values[,1,k], type = "h", main = paste("GDP vs. ",select_vec_multi[k],sep=""), xlab = "Lag", ylab = "CCF")
 abline(h = 0, col = "red")
 k<-4
-plot(lags[,1,length(select_vec_multi)],acf_values[,1,k], type = "h", main = paste("GDP vs. ",select_vec_multi[k],sep=""), xlab = "Lag", ylab = "ACF")
+plot(lags[,1,length(select_vec_multi)],acf_values[,1,k], type = "h", main = paste("GDP vs. ",select_vec_multi[k],sep=""), xlab = "Lag", ylab = "")
 abline(h = 0, col = "red")
@@
 cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
-cat("\\caption{Sample cross-correlation function (CCF): full data set.", sep = "")
+cat("\\caption{Data: standardized log-differences: pandemic trimmed to $\\pm 3$.", sep = "")
 cat("\\label{data}}", sep = "")
@@
 cat("\\caption{Sample cross-correlation function: GDP and ifo (left), GDP and ESI (right): full data set, including the pandemic.
-Correlations at positive lags mean that real-time GDP is lagging in relative terms.", sep = "")
+Correlations at positive lags signify that (real-time) GDP is lagging when referenced against the two (real-time) indicators.", sep = "")
 cat("\\label{CCF}}", sep = "")
@@
 cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
-cat("\\caption{Sample cross-correlation function: GDP and ifo (left), GDP and ESI (right): without Pandemic.
-Correlations at positive lags mean that real-time GDP is lagging in relative terms.", sep = "")
+cat("\\caption{Sample cross-correlation function: GDP and ifo (left), GDP and ESI (right): without pandemic.
+Correlations at positive lags signify that (real-time) GDP is lagging when referenced against the two (real-time) indicators.", sep = "")
 cat("\\label{CCF_wc}}", sep = "")

### 2025-05-08 Latex/BIP_predictor.tex
up

@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu May  8 14:11:36 2025
+% Thu May  8 14:21:03 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu May  8 14:11:36 2025
+% Thu May  8 14:21:03 2025
 \begin{table}[ht]
@@
 
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/Data.pdf}\caption{Data: standardized log-differences: pandemic trimmed to $\pm 3$.\label{data}}\end{center}\end{figure}
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/data_lags.pdf}\caption{Data: leads and lags during financial crisis (left) and Pandemic (right).\label{data_lags}}\end{center}\end{figure}
+\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/Data.pdf}\caption{Data: standardized log-differences.
+Data trimmed to $\pm 3$ for ease of visual inspection.\label{data}}\end{center}\end{figure}
+\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/data_lags.pdf}\caption{Data: leads and lags during financial crisis (left) and Pandemic (right).
+Data trimmed to $\pm 3$ for ease of visual inspection.\label{data_lags}}\end{center}\end{figure}
 
@@
 \ref{direct_wc} (full dataset without the pandemic) and \ref{direct_wc_financial_crisis} (financial crisis) for horizons up to three quarters ahead.
-\begin{figure}[H]\begin{center}\includegraphics[height=5in, width=6in]{./Figures/direct_wc_all.pdf}\caption{Direct forecasts (red lines) based on all indicators and full sample information, without the pandemic: GDP (black line) shifted by zero one, two and three quarters.\label{direct_wc}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=5in, width=6in]{./Figures/direct_wc_financial_crisis.pdf}\caption{Direct forecast  across the financial crisis:  regression based on all indicators and full sample information.\label{direct_wc_financial_crisis}}\end{center}\end{figure}The results indicate that as $h$ increases, the peaks and dips of the direct forecasts become increasingly right-shifted (delayed) relative to the target.
+\begin{figure}[H]\begin{center}\includegraphics[height=5in, width=6in]{./Figures/direct_wc_all.pdf}\caption{Direct forecasts (red lines) based on all indicators and full sample information, without the pandemic: GDP (black line) shifted by zero one, two and three quarters.\label{direct_wc}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=5in, width=6in]{./Figures/direct_wc_financial_crisis.pdf}\caption{Direct forecast  across the financial crisis:  regression based on all indicators and full sample information, without the pandemic.\label{direct_wc_financial_crisis}}\end{center}\end{figure}The results indicate that as $h$ increases, the peaks and dips of the direct forecasts become increasingly right-shifted (delayed) relative to the target.
 Additionally, the F-statistics precipitously decline from $F=3.57$, for $h=2$, to $F=0.46$, for $h=3$, see Table \eqref{f_stat}, suggesting that the predictor becomes statistically insignificant for larger forecast horizons exceeding two quarters (link to Heinisch and Scheufele).
@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu May  8 14:11:36 2025
+% Thu May  8 14:21:04 2025
 \begin{table}[ht]
@@
 The two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu May  8 14:11:36 2025
+% Thu May  8 14:21:04 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu May  8 14:11:38 2025
+% Thu May  8 14:21:05 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu May  8 14:11:38 2025
+% Thu May  8 14:21:05 2025
 \begin{table}[ht]

### 2025-05-08 Sweave/BIP_predictor.Rnw
up

@@
 par(mfrow=c(1,2))
-  ts.plot(mplot_with_pandemic,col=colo_i,main=paste("MA inversion GDP: with pandemic",sep=""),xlab="lag")
+  ts.plot(mplot_with_pandemic,col=colo_i,main=paste("MA inversion GDP: with pandemic",sep=""),xlab="")
   for (k in 1:length(select_vec_i))
     mtext(select_vec_i[k],line=-k,col=colo_i[k])
-    ts.plot(mplot_without_pandemic,col=colo_i,main="Without pandemic",xlab="lag")
+    ts.plot(mplot_without_pandemic,col=colo_i,main="Without pandemic",xlab="")
 #  for (k in 1:length(select_vec_i))
@@
 cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
-cat("\\caption{Data: standardized log-differences: pandemic trimmed to $\\pm 3$.", sep = "")
+cat("\\caption{Data: standardized log-differences. Data trimmed to $\\pm 3$ for ease of visual inspection.", sep = "")
 cat("\\label{data}}", sep = "")
@@
 cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
-cat("\\caption{Data: leads and lags during financial crisis (left) and Pandemic (right).", sep = "")
+cat("\\caption{Data: leads and lags during financial crisis (left) and Pandemic (right). Data trimmed to $\\pm 3$ for ease of visual inspection.", sep = "")
 cat("\\label{data_lags}}", sep = "")
@@
 cat("\\includegraphics[height=5in, width=6in]{", file, "}\n",sep = "")
-cat("\\caption{Direct forecast  across the financial crisis:  regression based on all indicators and full sample information.", sep = "")
+cat("\\caption{Direct forecast  across the financial crisis:  regression based on all indicators and full sample information, without the pandemic.", sep = "")
 cat("\\label{direct_wc_financial_crisis}}", sep = "")

### 2025-05-08 Latex/BIP_predictor.tex
up

@@
 
-% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu May  8 14:21:03 2025
+% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Thu May  8 14:43:48 2025
 \begin{table}[ht]
@@
   \hline
+2024-07-01 & 902.571 & 90.700 & 85.800 & 92.500 & -1.300 \\ 
   2024-08-01 &  & 93.200 & 84.900 & 90.900 & -1.400 \\ 
@@
 \label{f_stat}
-\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu May  8 14:21:03 2025
+\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Thu May  8 14:43:48 2025
 \begin{table}[ht]
@@
 \begin{figure}[H]\begin{center}\includegraphics[height=5in, width=6in]{./Figures/direct_hp_forecasts.pdf}\caption{Direct HP forecasts: entire data set with exclusion of the pandemic.\label{direct_hp_forecasts}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=5in, width=6in]{./Figures/direct_hp_forecasts_financial_crisis.pdf}\caption{Shifted GDP (black), direct HP forecast (blue) and classic direct forecast (red) over the curse of the financial crisis:  regression based on all indicators and full sample information (without pandemic).\label{direct_hp_forecasts_financial_crisis}}\end{center}\end{figure}Moreover, the F-statistics are larger than for the direct forecasts, see Table \eqref{f_stat}, but the predictor remains statistically insignificant at forecast horizons exceeding two quarters (even though the evaluation is conducted in-sample and based on all available indicators).
-% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu May  8 14:21:04 2025
+% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Thu May  8 14:43:51 2025
 \begin{table}[ht]
@@
 The dashed vertical line delimits in- and out-of-sample spans.
-The two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu May  8 14:21:04 2025
+The two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Thu May  8 14:43:51 2025
 \begin{table}[ht]
@@
 Table \eqref{p_val_hpbip_wc} reports HAC-adjusted p-values from regressions of the out-of-sample predictor on HP-BIP, indicating a strong link between the predictor and the target, out-of-sample.
-% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu May  8 14:21:05 2025
+% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Thu May  8 14:43:56 2025
 \begin{table}[ht]
@@
 
-% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu May  8 14:21:05 2025
+% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Thu May  8 14:43:56 2025
 \begin{table}[ht]
@@
 \label{p_val_wc}
-\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu May  8 14:21:05 2025
+\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Thu May  8 14:43:56 2025
 \begin{table}[ht]
@@
 
-% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu May  8 14:21:05 2025
+% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Thu May  8 14:43:57 2025
 \begin{table}[ht]
@@
 \label{p_val_wc5}
-\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu May  8 14:21:05 2025
+\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Thu May  8 14:43:57 2025
 \begin{table}[ht]
@@
 \label{rRMSE_mSSA_comp_mean6}
-\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu May  8 14:21:05 2025
+\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Thu May  8 14:43:57 2025
 \begin{table}[ht]
@@
 \label{rRMSE_mSSA_comp_direct_without_covid7}
-\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu May  8 14:21:05 2025
+\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Thu May  8 14:43:57 2025
 \begin{table}[ht]
@@
 \label{p_val1}
-\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu May  8 14:21:05 2025
+\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Thu May  8 14:43:57 2025
 \begin{table}[ht]
@@
 \label{rRMSE_mSSA_comp_mean2}
-\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu May  8 14:21:05 2025
+\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Thu May  8 14:43:57 2025
 \begin{table}[ht]
@@
 \label{rRMSE_mSSA_comp_direct3}
-\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Thu May  8 14:21:05 2025
+\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Thu May  8 14:43:57 2025
 \begin{table}[ht]

### 2025-05-08 Sweave/BIP_predictor.Rnw
up

@@
 
-mat1<-as.matrix(tail(data_table))
+mat1<-as.matrix(tail(data_table,7))
 #latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
@@
 mplot<-mplot[(L:nrow(mplot)),]
+colnames(mplot)[2]<-"GDP"
 #mplot<-t(t(mplot)/mplot[nrow(mplot),])
@@
 rownames(mssa_final_vs_prem_wc)<-rownames(x_mat_wc)
-colnames(mssa_final_vs_prem_wc)<-c("Predictor based on final M-SSA","Predictor based on 2007-M-SSA")
+colnames(mssa_final_vs_prem_wc)<-c("Predictor based on final M-SSA","Predictor based on 2008-M-SSA")
 
@@
 colnames(mplot)<-colnames(mssa_final_vs_prem_wc)
-main_title<-paste("GDP component predictors: final and 2007-MSSA (h=",h,", shift=",shift,")",sep="")
+main_title<-paste("GDP component predictors: final and 2008-MSSA (h=",h,", shift=",shift,")",sep="")
 plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))

### 2025-05-09 Sweave/BIP_predictor.Rnw
up

@@
   for (k in 1:length(select_vec_i))
+  { 
+    if (k==1)
+    {
+      mtext("GDP",line=-k,col=colo_i[k])
+    } else
+    {
       mtext(select_vec_i[k],line=-k,col=colo_i[k])
+    }
+  }
 }
@@
   for (k in 1:length(select_vec_i))
+  { 
+    if (k==1)
+    {
+      mtext("GDP",line=-k,col=colo_i[k])
+    } else
+    {
       mtext(select_vec_i[k],line=-k,col=colo_i[k])
+    }
+  }
 ts.plot(mplot_without_pandemic,col=colo_i,main="Without pandemic",xlab="")

### 2025-05-09 Sweave/BIP_predictor.Rnw
up

@@
   forward_shifted_BIP<-c(x_mat_wc[(1+lag_vec[1]+h):nrow(x_mat_wc),"BIP"],rep(NA,h+lag_vec[1]))
-  # Regress selected indicators on forward-shifted BIP
+  # Regress selected filtered indicators on forward-shifted BIP
   length(forward_shifted_BIP)
-  lm_obj<-lm(forward_shifted_BIP~t(hp_c_array_wc[,,h+1]))
+  lm_obj<-lm(forward_shifted_BIP~t(hp_c_array_wc[,,1]))
   # You will probably not find statistically significant regressors for h>2: BIP is a noisy series

### 2025-05-09 Sweave/BIP_predictor.Rnw
up

--- 2025-05-09 Sweave/BIP_predictor.Rnw
@@
 # To illustrate the direct predictor consider the following example of a h-step ahead direct forecast:
-f_vec_unfiltered<-NULL
+f_vec_unfiltered<-p_val_unfiltered<-NULL
 for (h in 0:3)
@@
   f_vec_unfiltered<-c(f_vec_unfiltered,sum_obj$fstatistic["value"])  
+  p_val_unfiltered<-c(p_val_unfiltered,pf(sum_obj$fstatistic["value"], sum_obj$df[1],sum_obj$df[2], lower.tail = FALSE))
 
@@
 par(mfrow=c(2,2))
-f_vec_hp<-t_mat_hp<-NULL
+f_vec_hp<-t_mat_hp<-p_val_hp<-NULL
 for (h in 0:3)
@@
   sum_obj<-summary(lm_obj)
+
+  p_val_hp<-c(p_val_hp,pf(sum_obj$fstatistic["value"], sum_obj$df[1],sum_obj$df[2], lower.tail = FALSE))
   f_vec_hp<-c(f_vec_hp,sum_obj$fstatistic["value"])
@@
 mat_re<-rbind(f_vec_unfiltered,f_vec_hp)
+mat_re<-rbind(p_val_unfiltered,p_val_hp)
 rownames(mat_re)<-c("Direct forecast","Direct HP forecast")
 colnames(mat_re)<-paste("h=",0:3,sep="")
-mat1<-round(mat_re,3)
+mat1<-round(mat_re,5)
 #latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
@@
 xtable(mat1, dec = 1,digits=rep(3,dim(mat_re)[2]+1),
-paste("F-statistics of direct forecasts and direct HP forecasts for forecast horizons $h=0,1,2,3$: fully in-sample.  "),
+paste("P-values of of F-statistics of direct forecasts and direct HP forecasts for forecast horizons $h=0,1,2,3$: fully in-sample.  "),
 label=paste("f_stat",sep=""),

### 2025-05-09 Latex/BIP_predictor.tex
up

@@
 
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 14:43:48 2025
+% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Fri May  9 17:38:47 2025
 \begin{table}[ht]
@@
 \label{f_stat}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 14:43:48 2025
+\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Fri May  9 17:38:47 2025
 \begin{table}[ht]
@@
 \begin{figure}[H]\begin{center}\includegraphics[height=5in, width=6in]{./Figures/direct_hp_forecasts.pdf}\caption{Direct HP forecasts: entire data set with exclusion of the pandemic.\label{direct_hp_forecasts}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=5in, width=6in]{./Figures/direct_hp_forecasts_financial_crisis.pdf}\caption{Shifted GDP (black), direct HP forecast (blue) and classic direct forecast (red) over the curse of the financial crisis:  regression based on all indicators and full sample information (without pandemic).\label{direct_hp_forecasts_financial_crisis}}\end{center}\end{figure}Moreover, the F-statistics are larger than for the direct forecasts, see Table \eqref{f_stat}, but the predictor remains statistically insignificant at forecast horizons exceeding two quarters (even though the evaluation is conducted in-sample and based on all available indicators).
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 14:43:51 2025
+% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Fri May  9 17:38:48 2025
 \begin{table}[ht]
@@
   \hline
-Direct forecast & 7.188 & 5.285 & 3.574 & 0.460 \\ 
-  Direct HP forecast & 21.886 & 8.415 & 4.810 & 1.834 \\ 
+Direct forecast & 0.000 & 0.000 & 0.003 & 0.837 \\ 
+  Direct HP forecast & 0.000 & 0.000 & 0.047 & 0.113 \\ 
    \hline
 \end{tabular}
-\caption{F-statistics of direct forecasts and direct HP forecasts for forecast horizons $h=0,1,2,3$: fully in-sample.  } 
+\caption{P-values of of F-statistics of direct forecasts and direct HP forecasts for forecast horizons $h=0,1,2,3$: fully in-sample.  } 
 \label{f_stat}
@@
 The dashed vertical line delimits in- and out-of-sample spans.
-The two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 14:43:51 2025
+The two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Fri May  9 17:38:48 2025
 \begin{table}[ht]
@@
 Table \eqref{p_val_hpbip_wc} reports HAC-adjusted p-values from regressions of the out-of-sample predictor on HP-BIP, indicating a strong link between the predictor and the target, out-of-sample.
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 14:43:56 2025
+% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Fri May  9 17:38:49 2025
 \begin{table}[ht]
@@
 
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 14:43:56 2025
+% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Fri May  9 17:38:49 2025
 \begin{table}[ht]
@@
 \label{p_val_wc}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 14:43:56 2025
+\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Fri May  9 17:38:49 2025
 \begin{table}[ht]
@@
 
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 14:43:57 2025
+% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Fri May  9 17:38:49 2025
 \begin{table}[ht]
@@
 \label{p_val_wc5}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 14:43:57 2025
+\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Fri May  9 17:38:49 2025
 \begin{table}[ht]
@@
 \label{rRMSE_mSSA_comp_mean6}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 14:43:57 2025
+\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Fri May  9 17:38:49 2025
 \begin{table}[ht]
@@
 \label{rRMSE_mSSA_comp_direct_without_covid7}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 14:43:57 2025
+\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Fri May  9 17:38:49 2025
 \begin{table}[ht]
@@
 \label{p_val1}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 14:43:57 2025
+\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Fri May  9 17:38:49 2025
 \begin{table}[ht]
@@
 \label{rRMSE_mSSA_comp_mean2}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 14:43:57 2025
+\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Fri May  9 17:38:49 2025
 \begin{table}[ht]
@@
 \label{rRMSE_mSSA_comp_direct3}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 14:43:57 2025
+\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Fri May  9 17:38:49 2025
 \begin{table}[ht]

### 2025-05-09 Sweave/BIP_predictor.Rnw
up

@@
 # To illustrate the direct predictor consider the following example of a h-step ahead direct forecast:
-f_vec_unfiltered<-p_val_unfiltered<-NULL
-for (h in 0:3)
+f_vec_unfiltered<-p_val_unfiltered<-MSE_unfiltered_vec<-NULL
+for (h in 0:6)
 {
@@
   forward_shifted_BIP<-c(x_mat_wc[(1+lag_vec[1]+h):nrow(x_mat_wc),"BIP"],rep(NA,h+lag_vec[1]))
-  # Regress selected indicators on forward-shifted BIP
-  lm_obj<-lm(forward_shifted_BIP~x_mat_wc[,select_direct_indicator])
+  # Regress selected indicators on forward-shifted BIP: start at t=L (same sample as HP-C)
+  lm_obj<-lm(forward_shifted_BIP[L:length(forward_shifted_BIP)]~x_mat_wc[L:nrow(x_mat_wc),select_direct_indicator])
   # You will probably not find statistically significant regressors for h>2: BIP is a noisy series
   summary(lm_obj)
-  sum_obj<-summary(lm_obj)
+  sum_obj<-sum_obj_unfiltered<-summary(lm_obj)
+  MSE_unfiltered_vec<-c(MSE_unfiltered_vec,sum_obj$sigma^2)
+
   f_vec_unfiltered<-c(f_vec_unfiltered,sum_obj$fstatistic["value"])  
@@
   forward_shifted_BIP<-c(x_mat_wc[(1+lag_vec[1]+h):nrow(x_mat_wc),"BIP"],rep(NA,h+lag_vec[1]))
-  # Regress selected indicators on forward-shifted BIP
-  lm_obj<-lm(forward_shifted_BIP~x_mat_wc[,select_direct_indicator])
+  # Regress selected indicators on forward-shifted BIP: start at t=L (same sample as HP-C)
+  lm_obj<-lm(forward_shifted_BIP[L:length(forward_shifted_BIP)]~x_mat_wc[L:nrow(x_mat_wc),select_direct_indicator])
   # You will probably not find statistically significant regressors for h>2: BIP is a noisy series
@@
 dev.off()
+
 @
@@
 par(mfrow=c(2,2))
-f_vec_hp<-t_mat_hp<-p_val_hp<-NULL
-for (h in 0:3)
+f_vec_hp<-t_mat_hp<-p_val_hp<-MSE_hp_vec<-NULL
+for (h in 0:6)
 {
@@
   # You will probably not find statistically significant regressors for h>2: BIP is a noisy series
+  sum_obj_hp<-summary(lm_obj)
   sum_obj<-summary(lm_obj)
-  sum_obj<-summary(lm_obj)
-
+  MSE_hp_vec<-c(MSE_hp_vec,sum_obj$sigma^2)
   p_val_hp<-c(p_val_hp,pf(sum_obj$fstatistic["value"], sum_obj$df[1],sum_obj$df[2], lower.tail = FALSE))
@@
 dev.off()
+
+# Compare MSEs and p-values
+sqrt(MSE_hp_vec/MSE_unfiltered_vec)
+
+p_val_unfiltered
+p_val_hp
 

### 2025-05-10 Latex/BIP_predictor.tex
up

--- 2025-05-08 Latex/BIP_predictor.tex
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 14:43:48 2025
+% Sat May 10 08:52:45 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 14:43:48 2025
+% Sat May 10 08:52:45 2025
 \begin{table}[ht]
@@
 \begin{figure}[H]\begin{center}\includegraphics[height=5in, width=6in]{./Figures/direct_wc_all.pdf}\caption{Direct forecasts (red lines) based on all indicators and full sample information, without the pandemic: GDP (black line) shifted by zero one, two and three quarters.\label{direct_wc}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=5in, width=6in]{./Figures/direct_wc_financial_crisis.pdf}\caption{Direct forecast  across the financial crisis:  regression based on all indicators and full sample information, without the pandemic.\label{direct_wc_financial_crisis}}\end{center}\end{figure}The results indicate that as $h$ increases, the peaks and dips of the direct forecasts become increasingly right-shifted (delayed) relative to the target.
-Additionally, the F-statistics precipitously decline from $F=3.57$, for $h=2$, to $F=0.46$, for $h=3$, see Table \eqref{f_stat}, suggesting that the predictor becomes statistically insignificant for larger forecast horizons exceeding two quarters (link to Heinisch and Scheufele).
+Additionally, the F-statistics precipitously decline from $F=3.21$, for $h=2$, to $F=0.28$, for $h=3$, see Table \eqref{f_stat}, suggesting that the predictor becomes statistically insignificant for larger forecast horizons exceeding two quarters (link to Heinisch and Scheufele).
 
@@
 \begin{figure}[H]\begin{center}\includegraphics[height=5in, width=6in]{./Figures/direct_hp_forecasts.pdf}\caption{Direct HP forecasts: entire data set with exclusion of the pandemic.\label{direct_hp_forecasts}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=5in, width=6in]{./Figures/direct_hp_forecasts_financial_crisis.pdf}\caption{Shifted GDP (black), direct HP forecast (blue) and classic direct forecast (red) over the curse of the financial crisis:  regression based on all indicators and full sample information (without pandemic).\label{direct_hp_forecasts_financial_crisis}}\end{center}\end{figure}Moreover, the F-statistics are larger than for the direct forecasts, see Table \eqref{f_stat}, but the predictor remains statistically insignificant at forecast horizons exceeding two quarters (even though the evaluation is conducted in-sample and based on all available indicators).
+
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 14:43:51 2025
+% Sat May 10 08:52:46 2025
 \begin{table}[ht]
@@
   \hline
- & h=0 & h=1 & h=2 & h=3 \\ 
+ & h=3 & h=4 & h=5 & h=6 \\ 
   \hline
-Direct forecast & 7.188 & 5.285 & 3.574 & 0.460 \\ 
-  Direct HP forecast & 21.886 & 8.415 & 4.810 & 1.834 \\ 
+rRMSE & 0.962 & 0.967 & 0.990 & 0.967 \\ 
    \hline
 \end{tabular}
-\caption{F-statistics of direct forecasts and direct HP forecasts for forecast horizons $h=0,1,2,3$: fully in-sample.  } 
+\caption{Relative root mean-square error of HP-C against classic direct forecast: numbers smaller one signify an outperformance by the direct HP-C forecast.  } 
 \label{f_stat}
-\end{table}We now propose a more refined causal filter design aiming for statistical significance of the resulting predictor at longer forecast horizons, out-of-sample.
+\end{table}
+
+% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Sat May 10 08:52:46 2025
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrrrrr}
+  \hline
+ & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
+  \hline
+P-values & 0.000 & 0.000 & 0.007 & 0.946 & 0.603 & 0.096 & 0.181 \\ 
+   \hline
+\end{tabular}
+\caption{Statistical significance of direct forecasts.  } 
+\label{f_stat}
+\end{table}
+We now propose a more refined causal filter design aiming for statistical significance of the resulting predictor at longer forecast horizons, out-of-sample.
 
@@
 The two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 14:43:51 2025
+% Sat May 10 08:52:47 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 14:43:56 2025
+% Sat May 10 08:52:49 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 14:43:56 2025
+% Sat May 10 08:52:49 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 14:43:57 2025
+% Sat May 10 08:52:50 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 14:43:57 2025
+% Sat May 10 08:52:50 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 14:43:57 2025
+% Sat May 10 08:52:50 2025
 \begin{table}[ht]
 \centering
-\begin{tabular}{rrrrrrrr}
+\begin{tabular}{rrrrrrr}
   \hline
- & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
+ & Shift=0 & Shift=1 & Shift=2 & Shift=3 & Shift=4 & Shift=5 \\ 
   \hline
-Shift=0 & 0.855 & 0.855 & 0.855 & 0.855 & 0.855 & 0.855 & 0.855 \\ 
-  Shift=1 & 0.874 & 0.874 & 0.874 & 0.874 & 0.874 & 0.874 & 0.874 \\ 
-  Shift=2 & 0.919 & 0.919 & 0.919 & 0.919 & 0.919 & 0.919 & 0.919 \\ 
-  Shift=3 & 1.023 & 1.023 & 1.023 & 1.023 & 1.023 & 1.023 & 1.023 \\ 
-  Shift=4 & 0.985 & 0.985 & 0.985 & 0.985 & 0.985 & 0.985 & 0.985 \\ 
-  Shift=5 & 1.010 & 1.010 & 1.010 & 1.010 & 1.010 & 1.010 & 1.010 \\ 
+RRMSE & 0.855 & 0.874 & 0.919 & 1.023 & 0.985 & 1.010 \\ 
    \hline
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May  8 14:43:57 2025
+% Sat May 10 08:52:50 2025
 \begin{table}[ht]
 \centering
-\begin{tabular}{rrrrrrrr}
+\begin{tabular}{rrrrrrr}
   \hline
- & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
+ & Shift=0 & Shift=1 & Shift=2 & Shift=3 & Shift=4 & Shift=5 \\ 
   \hline
-Shift=0 & 0.833 & 0.833 & 0.833 & 0.833 & 0.833 & 0.833 & 0.833 \\ 
-  Shift=1 & 1.083 & 1.083 & 1.083 & 1.083 & 1.083 & 1.083 & 1.083 \\ 
-  Shift=2 & 0.981 & 0.981 & 0.981 & 0.981 & 0.981 & 0.981 & 0.981 \\ 
-  Shift=3 & 0.979 & 0.979 & 0.979 & 0.979 & 0.979 & 0.979 & 0.979 \\ 
-  Shift=4 & 1.027 & 1.027 & 1.027 & 1.027 & 1.027 & 1.027 & 1.027 \\ 
-  Shift=5 & 1.020 & 1.020 & 1.020 & 1.020 & 1.020 & 1.020 & 1.020 \\ 
+RRMSE & 0.833 & 1.083 & 0.981 & 0.979 & 1.027 & 1.020 \\ 
    \hline

### 2025-05-10 Sweave/BIP_predictor.Rnw
up

@@
 #   -in applications the regression is out-of-sample (ifo and ESI work best out-of-sample)
-select_direct_indicator<-select_vec_multi
 
@@
 par(mfrow=c(2,2))
+select_direct_indicator<-select_vec_multi
 
@@
 f_vec_hp<-t_mat_hp<-p_val_hp<-MSE_hp_vec<-NULL
-for (h in 0:6)
+for (h in 0:3)
 {
@@
 
+if (F)
+{
+# rRMSEs are computed for h=0:6 (just increase h in the above lopp to recompute)
 # Compare MSEs and p-values
-sqrt(MSE_hp_vec/MSE_unfiltered_vec)
-
-p_val_unfiltered
-p_val_hp
+  rrmse_hp_direct<-sqrt(MSE_hp_vec/MSE_unfiltered_vec)
+  names(rrmse_hp_direct)<-names(p_val_hp)<-names(p_val_unfiltered)<-paste("h=",0:6,sep="")
+
+  save(rrmse_hp_direct,file=paste(path.result,"/rrmse_hp_direct",sep=""))
+  save(p_val_unfiltered,file=paste(path.result,"/p_val_unfiltered",sep=""))
+  save(p_val_hp,file=paste(path.result,"/p_val_hp",sep=""))
+} else
+{
+  load(file=paste(path.result,"/rrmse_hp_direct",sep=""))
+  load(file=paste(path.result,"/p_val_unfiltered",sep=""))
+  load(file=paste(path.result,"/p_val_hp",sep=""))
+
+}
+rrmse_hp_direct
 
@@
 Moreover, the F-statistics are larger than for the direct forecasts, see Table \eqref{f_stat}, but the predictor remains statistically insignificant at forecast horizons exceeding two quarters (even though the evaluation is conducted in-sample and based on all available indicators).
+
 <<label=ats_mba_2,echo=FALSE,results=tex>>=
 
-mat_re<-rbind(f_vec_unfiltered,f_vec_hp)
-mat_re<-rbind(p_val_unfiltered,p_val_hp)
-rownames(mat_re)<-c("Direct forecast","Direct HP forecast")
-colnames(mat_re)<-paste("h=",0:3,sep="")
-mat1<-round(mat_re,5)
+mat1<-matrix(round(rrmse_hp_direct[4:7],5),nrow=1)
+colnames(mat1)<-names(rrmse_hp_direct)[4:7]
+rownames(mat1)<-"rRMSE"
 #latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
 #center = "centering", file = "", floating = FALSE)
-xtable(mat1, dec = 1,digits=rep(3,dim(mat_re)[2]+1),
-paste("P-values of of F-statistics of direct forecasts and direct HP forecasts for forecast horizons $h=0,1,2,3$: fully in-sample.  "),
+xtable(mat1, dec = 1,digits=rep(3,dim(mat1)[2]+1),
+paste("Relative root mean-square error of HP-C against classic direct forecast: numbers smaller one signify an outperformance by the direct HP-C forecast.  "),
 label=paste("f_stat",sep=""),
@@
 @
+
+<<label=ats_mba_2,echo=FALSE,results=tex>>=
+
+mat1<-matrix(round(p_val_unfiltered,5),nrow=1)
+colnames(mat1)<-names(p_val_unfiltered)
+rownames(mat1)<-"P-values"
+#latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
+#center = "centering", file = "", floating = FALSE)
+xtable(mat1, dec = 1,digits=rep(3,dim(mat1)[2]+1),
+paste("Statistical significance of direct forecasts.  "),
+label=paste("f_stat",sep=""),
+center = "centering", file = "", floating = FALSE)
+@
+
 We now propose a more refined causal filter design aiming for statistical significance of the resulting predictor at longer forecast horizons, out-of-sample.
@@
 
-mat1<-round(rRMSE_mSSA_direct_mean_without_covid,3)
+mat1<-matrix(round(rRMSE_mSSA_direct_mean_without_covid[,1],3),nrow=1)
+colnames(mat1)<-rownames(rRMSE_mSSA_direct_mean_without_covid)
+rownames(mat1)<-"RRMSE"
 #latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
@@
 
-mat1<-round(rRMSE_mSSA_direct_mean,3)
+mat1<-matrix(round(rRMSE_mSSA_direct_mean[,1],3),nrow=1)
+colnames(mat1)<-rownames(rRMSE_mSSA_direct_mean)
+rownames(mat1)<-"RRMSE"
 #latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",

### 2025-05-10 Latex/BIP_predictor.tex
up

@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sat May 10 08:52:45 2025
+% Sat May 10 15:26:21 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sat May 10 08:52:45 2025
+% Sat May 10 15:26:21 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sat May 10 08:52:46 2025
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrr}
-  \hline
- & h=3 & h=4 & h=5 & h=6 \\ 
-  \hline
-rRMSE & 0.962 & 0.967 & 0.990 & 0.967 \\ 
-   \hline
-\end{tabular}
-\caption{Relative root mean-square error of HP-C against classic direct forecast: numbers smaller one signify an outperformance by the direct HP-C forecast.  } 
-\label{f_stat}
-\end{table}
-
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sat May 10 08:52:46 2025
+% Sat May 10 15:26:24 2025
 \begin{table}[ht]
@@
   \hline
- & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
+ & h= 0 & h= 1 & h= 2 & h= 3 & h= 4 & h= 5 & h= 6 \\ 
   \hline
-P-values & 0.000 & 0.000 & 0.007 & 0.946 & 0.603 & 0.096 & 0.181 \\ 
+HP-C filtered & 0.000 & 0.001 & 0.036 & 0.046 & 0.049 & 0.012 & 0.022 \\ 
+  Unfiltered & 0.028 & 0.000 & 0.001 & 0.910 & 0.509 & 0.091 & 0.178 \\ 
    \hline
 \end{tabular}
-\caption{Statistical significance of direct forecasts.  } 
+\caption{Statistical significance of unfiltered and HP-C filtered forecasts based on HAC-adjusted Wald test: all indicators and full sample (without pandemic).  } 
 \label{f_stat}
@@
 The two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sat May 10 08:52:47 2025
+% Sat May 10 15:26:25 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sat May 10 08:52:49 2025
+% Sat May 10 15:26:30 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sat May 10 08:52:49 2025
+% Sat May 10 15:26:30 2025
 \begin{table}[ht]
@@
 \begin{figure}[H]\begin{center}\includegraphics[height=3in, width=5in]{./Figures/up_dated_ma_inv_multi_ip.pdf}\caption{MA-inversion (impulse response) of VAR: data up to Jan-2008 (left) vs.
-entire data set (right).\label{up_dated_ma_inv_multi_ip}}\end{center}\end{figure}
+entire data set  without pandemic (right).\label{up_dated_ma_inv_multi_ip}}\end{center}\end{figure}
 \begin{figure}[H]\begin{center}\includegraphics[height=3in, width=5in]{./Figures/bk_2008_all.pdf}\caption{M-SSA filter tracking HP-GDP at forecast horizon $h=4$: data up to Jan-2008 (left) vs.
-entire data set (right).\label{bk_2008_all}}\end{center}\end{figure}
+entire data set without pandemic (right).\label{bk_2008_all}}\end{center}\end{figure}
 
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sat May 10 08:52:50 2025
+% Sat May 10 15:26:31 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sat May 10 08:52:50 2025
+% Sat May 10 15:26:31 2025
 \begin{table}[ht]

### 2025-05-10 Sweave/BIP_predictor.Rnw
up

@@
 par(mfrow=c(1,1))
-
-# Plot the data
-# The real-time BIP (red) is lagging the target (black) by lag_vec[1] quarters (publication lag)
+colo<-c("black", "red", "green", "blue", "orange", "dimgray")
 mplot<-x_mat
 colnames(mplot)[1]<-"GDP"
-colo<-c("black",rainbow(ncol(data)-1))
 main_title<-"Data: standardized diff-log, Pandemic outliers trimmed"
 plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],lwd=1,ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
-mtext(colnames(mplot)[1],col=colo[1],line=-1)
 for (i in 1:ncol(mplot))
@@
   lines(mplot[,i],col=colo[i],lwd=1,lty=1)
-  mtext(colnames(mplot)[i],col=colo[i],line=-i)
 }
@@
 box()
+legend("bottomleft", legend = colnames(mplot), col = colo, lty = 1, lwd = 1, bty = "n")
+
 dev.off()
@@
 colnames(mplot)[1]<-"GDP"
-colo<-c("black",rainbow(ncol(data)-1))
-main_title<-"Financial Crisis"
-plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],lwd=1,ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
-mtext(colnames(mplot)[1],col=colo[1],line=-1)
+plot(mplot[,1],main="Financial crisis",axes=F,type="l",xlab="",ylab="",col=colo[1],lwd=1,ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
 for (i in 1:ncol(mplot))
@@
 box()
+#legend("bottomleft", legend = colnames(mplot), col = colo, lty = 1, lwd = 1, bty = "n")
 mplot<-x_mat[which(rownames(x_mat)>=2019&rownames(x_mat)<=2022),]
 colnames(mplot)[1]<-"GDP"
-colo<-c("black",rainbow(ncol(data)-1))
-main_title<-"Pandemic"
-plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],lwd=1,ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
-mtext(colnames(mplot)[1],col=colo[1],line=-1)
+plot(mplot[,1],main="Pandemic",axes=F,type="l",xlab="",ylab="",col=colo[1],lwd=1,ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
 for (i in 1:ncol(mplot))
@@
 box()
+#legend("bottomleft", legend = colnames(mplot), col = colo, lty = 1, lwd = 1, bty = "n")
 
@@
   mplot_with_pandemic<-mplot
+  colnames(mplot_with_pandemic)<-c("GDP",select_vec_i[2:length(select_vec_i)])
   ts.plot(mplot,col=colo_i,main=paste("MA inversion ",colnames(data_fit)[i],sep=""),xlab="lag")
@@
     mplot_without_pandemic<-mplot
+  colnames(mplot_without_pandemic)<-c("GDP",select_vec_i[2:length(select_vec_i)])
 
@@
 pdf(file = paste(path.out,file,sep=""), paper = "special", width = 9, height = 6)
+colo<-c("black", "red", "green", "blue", "orange", "dimgray")
 par(mfrow=c(1,2))
-ts.plot(mplot_with_pandemic,col=colo_i,main=paste("MA inversion GDP: with pandemic",sep=""),xlab="")
-  for (k in 1:length(select_vec_i))
-  { 
-    if (k==1)
-    {
-      mtext("GDP",line=-k,col=colo_i[k])
-    } else
-    {
-      mtext(select_vec_i[k],line=-k,col=colo_i[k])
-    }
-  }
-ts.plot(mplot_without_pandemic,col=colo_i,main="Without pandemic",xlab="")
-#  for (k in 1:length(select_vec_i))
-#    mtext(select_vec_i[k],line=-k,col=colo_i[k])
+plot(mplot_with_pandemic[,1],main=paste("MA inversion GDP: with pandemic",sep=""),axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot_with_pandemic)),max(na.exclude(mplot_with_pandemic))))
+
+for (i in 1:ncol(mplot_with_pandemic))
+{
+  lines(mplot_with_pandemic[,i],col=colo[i],lwd=1,lty=1)
+}
+abline(h=0)
+axis(1,at=1:nrow(mplot),labels=1:nrow(mplot)-1)
+axis(2)
+legend("topright", legend = colnames(mplot_with_pandemic), col = colo, lty = 1, lwd = 1, bty = "n")
+box()
+
+plot(mplot_without_pandemic[,1],main=paste("Without pandemic",sep=""),axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot_without_pandemic)),max(na.exclude(mplot_without_pandemic))))
+
+for (i in 1:ncol(mplot_without_pandemic))
+{
+  lines(mplot_without_pandemic[,i],col=colo[i],lwd=1,lty=1)
+}
+abline(h=0)
+axis(1,at=1:nrow(mplot),labels=1:nrow(mplot)-1)
+axis(2)
+box()
 
@@
 f_vec_unfiltered<-p_val_unfiltered<-MSE_unfiltered_vec<-NULL
-for (h in 0:6)
+for (h in 0:3)
 {
@@
     lines(mplot[,i],col=colo[i])
+  }
+  abline(h=0)
+  axis(1,at=c(1,12*1:(nrow(mplot)/12)),labels=rownames(mplot)[c(1,12*1:(nrow(mplot)/12))])
+  axis(2)
       if (h==0)
-    mtext(colnames(mplot)[i],col=colo[i],line=-i)
-  }
-  abline(h=0)
-  axis(1,at=c(1,12*1:(nrow(mplot)/12)),labels=rownames(mplot)[c(1,12*1:(nrow(mplot)/12))])
-  axis(2)
+    legend("bottomleft", legend = colnames(mplot), col = colo, lty = 1, lwd = 1, bty = "n")
+
   box()
@@
     lines(mplot[,i],col=colo[i])
+#    if (h==0)
+#    mtext(colnames(mplot)[i],col=colo[i],line=-i)
+  }
+  abline(h=0)
+  axis(1,at=1:nrow(mplot),labels=rownames(mplot))
+  axis(2)
     if (h==0)
-    mtext(colnames(mplot)[i],col=colo[i],line=-i)
-  }
-  abline(h=0)
-  axis(1,at=1:nrow(mplot),labels=rownames(mplot))
-  axis(2)
+    legend("bottomright", legend = colnames(mplot), col = colo, lty = 1, lwd = 1, bty = "n")
+
   box()
@@
   MSE_hp_vec<-c(MSE_hp_vec,sum_obj$sigma^2)
-  p_val_hp<-c(p_val_hp,pf(sum_obj$fstatistic["value"], sum_obj$df[1],sum_obj$df[2], lower.tail = FALSE))
-  f_vec_hp<-c(f_vec_hp,sum_obj$fstatistic["value"])
-  t_mat_hp<-rbind(t_mat_hp,sum_obj$coef[-1,4])
   # Technical note: 
@@
   # Compute the predictor: one can rely on the generic R-function predict or compute the predictor manually
-  direct_hp_forecast<-lm_obj$coef[1]+t(hp_c_array_wc[,,h+1])%*%lm_obj$coef[2:(ncol(t(hp_c_array_wc[,,h+1]))+1)]
+  direct_hp_forecast<-lm_obj$coef[1]+t(hp_c_array_wc[,,1])%*%lm_obj$coef[2:(ncol(t(hp_c_array_wc[,,1]))+1)]
   # Note that this is a full-sample predictor (no out-of-sample span)
@@
     lines(mplot[,i],col=colo[i])
+#    if (h==0)
+#    mtext(colnames(mplot)[i],col=colo[i],line=-i)
+  }
+  abline(h=0)
+  axis(1,at=1:nrow(mplot),labels=rownames(mplot))
+  axis(2)
   if (h==0)
-    mtext(colnames(mplot)[i],col=colo[i],line=-i)
-  }
-  abline(h=0)
-  axis(1,at=1:nrow(mplot),labels=rownames(mplot))
-  axis(2)
-  box()
-}
-colnames(t_mat_hp)<-select_vec_multi
-rownames(t_mat_hp)<-paste("h=",0:3)
+    legend("bottomleft", legend = colnames(mplot), col = colo, lty = 1, lwd = 1, bty = "n")
+
+  box()
+}
+#colnames(t_mat_hp)<-select_vec_multi
+#rownames(t_mat_hp)<-paste("h=",0:3)
 dev.off()
 
-if (F)
-{
-# rRMSEs are computed for h=0:6 (just increase h in the above lopp to recompute)
-# Compare MSEs and p-values
-  rrmse_hp_direct<-sqrt(MSE_hp_vec/MSE_unfiltered_vec)
-  names(rrmse_hp_direct)<-names(p_val_hp)<-names(p_val_unfiltered)<-paste("h=",0:6,sep="")
-
-  save(rrmse_hp_direct,file=paste(path.result,"/rrmse_hp_direct",sep=""))
-  save(p_val_unfiltered,file=paste(path.result,"/p_val_unfiltered",sep=""))
-  save(p_val_hp,file=paste(path.result,"/p_val_hp",sep=""))
-} else
-{
-  load(file=paste(path.result,"/rrmse_hp_direct",sep=""))
-  load(file=paste(path.result,"/p_val_unfiltered",sep=""))
-  load(file=paste(path.result,"/p_val_hp",sep=""))
-
-}
-rrmse_hp_direct
-
-# Financial crisis
+# Same but financial crisis only
 file = "./Figures/direct_hp_forecasts_financial_crisis.pdf"
@@
   length(forward_shifted_BIP)
-  lm_obj<-lm(forward_shifted_BIP~t(hp_c_array_wc[,,h+1]))
+  lm_obj<-lm(forward_shifted_BIP~t(hp_c_array_wc[,,1]))
   # You will probably not find statistically significant regressors for h>2: BIP is a noisy series
@@
     lines(mplot[,i],col=colo[i])
+#    if (h==0)
+#    mtext(colnames(mplot)[i],col=colo[i],line=-i)
+  }
+  abline(h=0)
+  axis(1,at=1:nrow(mplot),labels=rownames(mplot))
+  axis(2)
   if (h==0)
-    mtext(colnames(mplot)[i],col=colo[i],line=-i)
-  }
-  abline(h=0)
-  axis(1,at=1:nrow(mplot),labels=rownames(mplot))
-  axis(2)
+legend("bottomleft", legend = colnames(mplot), col = colo, lty = 1, lwd = 1, bty = "n")
   box()
@@
 dev.off()
+
+#-------------------
+# Compute p-values of direct and HP-c forecasts: rely on HAC adjusted Wald-statistic
+# -There is no HAC-adjustment for F-statistic
+# -Use Wald-test with HAC-adjusted variance of estimators
+
+ws_vec_unfiltered<-f_vec_unfiltered<-p_val_unfiltered<-MSE_unfiltered_vec<-NULL
+for (h in 0:6)#h<-0
+{
+  # Shift BIP forward by publication lag+forecast horizon
+  forward_shifted_BIP<-c(x_mat_wc[(1+lag_vec[1]+h):nrow(x_mat_wc),"BIP"],rep(NA,h+lag_vec[1]))
+  # Regress selected indicators on forward-shifted BIP: start at t=L (same sample as HP-C)
+  lm_obj<-lm(forward_shifted_BIP[L:length(forward_shifted_BIP)]~x_mat_wc[L:nrow(x_mat_wc),])
+# HAC adjustment
+  var_HAC<-vcovHAC(lm_obj)
+  coef<-lm_obj$coef
+# Wald statistic  
+  chisq<-t(coef)%*%solve(var_HAC)%*%coef
+# P-Value  
+  ws_vec_unfiltered<-c(ws_vec_unfiltered,pchisq(chisq,ncol(x_mat_wc)+1,lower.tail = F))
+}
+
+ws_vec_hp<-f_vec_hp<-t_mat_hp<-p_val_hp<-MSE_hp_vec<-NULL
+for (h in 0:6)#h<-3
+{
+  # Shift BIP forward by publication lag+forecast horizon
+  forward_shifted_BIP<-c(x_mat_wc[(1+lag_vec[1]+h):nrow(x_mat_wc),"BIP"],rep(NA,h+lag_vec[1]))
+  # Regress selected filtered indicators on forward-shifted BIP
+  length(forward_shifted_BIP)
+  lm_obj<-lm(forward_shifted_BIP~t(hp_c_array_wc[,,1]))
+# HAC adjustment  
+  var_HAC<-vcovHAC(lm_obj)
+  coef<-lm_obj$coef
+# Wald statistic  
+  chisq<-t(coef)%*%solve(var_HAC)%*%coef
+# P-value  
+  ws_vec_hp<-c(ws_vec_hp,pchisq(chisq,ncol(x_mat_wc)+1,lower.tail = F))
+
+}
+
 @
@@
 
-mat1<-matrix(round(rrmse_hp_direct[4:7],5),nrow=1)
-colnames(mat1)<-names(rrmse_hp_direct)[4:7]
-rownames(mat1)<-"rRMSE"
+mat1<-rbind(round(ws_vec_hp,5),round(ws_vec_unfiltered,5))
+colnames(mat1)<-paste("h=",0:(length(ws_vec_unfiltered)-1))
+rownames(mat1)<-c("HP-C filtered","Unfiltered")
 #latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
@@
 xtable(mat1, dec = 1,digits=rep(3,dim(mat1)[2]+1),
-paste("Relative root mean-square error of HP-C against classic direct forecast: numbers smaller one signify an outperformance by the direct HP-C forecast.  "),
-label=paste("f_stat",sep=""),
-center = "centering", file = "", floating = FALSE)
-@
-
-<<label=ats_mba_2,echo=FALSE,results=tex>>=
-
-mat1<-matrix(round(p_val_unfiltered,5),nrow=1)
-colnames(mat1)<-names(p_val_unfiltered)
-rownames(mat1)<-"P-values"
-#latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
-#center = "centering", file = "", floating = FALSE)
-xtable(mat1, dec = 1,digits=rep(3,dim(mat1)[2]+1),
-paste("Statistical significance of direct forecasts.  "),
+paste("Statistical significance of unfiltered and HP-C filtered forecasts based on HAC-adjusted Wald test: all indicators and full sample (without pandemic).  "),
 label=paste("f_stat",sep=""),
@@
 par(mfrow=c(1,2))
+colo<-c("black", "red", "green", "blue", "orange", "dimgray")
 mplot<-gammak
-colnames(mplot)[1]<-"GDP"
-colo<-rainbow(ncol(mplot))
-ts.plot(mplot,col=colo,main="M-MSE",xlab="")
+plot(mplot[,1],main="M-MSE",axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
+
 for (i in 1:ncol(mplot))
-  mtext(colnames(mplot)[i],col=colo[i],line=-i)
+{
+  lines(mplot[,i],col=colo[i],lwd=1,lty=1)
+}
+abline(h=0)
+axis(1,at=c(1,1+5*1:(nrow(mplot)/5)),labels=c(0,5*1:(nrow(mplot)/5)))
+axis(2)
+legend("topright", legend = colnames(mplot), col = colo, lty = 1, lwd = 1, bty = "n")
+box()
+
 mplot<-bk
-colnames(mplot)[1]<-"GDP"
-
-colo<-rainbow(ncol(mplot))
-ts.plot(mplot,col=colo,main="M-SSA",xlab="")
-#for (i in 1:ncol(mplot))
-#  mtext(colnames(mplot)[i],col=colo[i],line=-i)
-#mplot<-bk_HP_HP_C
-#colo<-rainbow(ncol(mplot))
-#ts.plot(mplot,col=colo,main="M-SSA larger HT")
-#for (i in 1:ncol(mplot))
-#  mtext(colnames(mplot)[i],col=colo[i],line=-i)
-
+plot(mplot[,1],main="M-SSA",axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
+
+for (i in 1:ncol(mplot))
+{
+  lines(mplot[,i],col=colo[i],lwd=1,lty=1)
+}
+abline(h=0)
+axis(1,at=c(1,1+5*1:(nrow(mplot)/5)),labels=c(0,5*1:(nrow(mplot)/5)))
+axis(2)
+box()
 dev.off()
@@
 par(mfrow=c(1,2))
-# h=0, see above
+colo<-c("black", "red", "green", "blue", "orange", "dimgray")
 mplot<-bk
-colo<-rainbow(ncol(mplot))
-ts.plot(mplot,col=colo,main="M-SSA: h=0",xlab="")
+plot(mplot[,1],main="M-SSA: h=0",axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
+
 for (i in 1:ncol(mplot))
-  mtext(colnames(mplot)[i],col=colo[i],line=-i)
-
-# h=4
+{
+  lines(mplot[,i],col=colo[i],lwd=1,lty=1)
+}
+abline(h=0)
+axis(1,at=c(1,1+5*1:(nrow(mplot)/5)),labels=c(0,5*1:(nrow(mplot)/5)))
+axis(2)
+legend("topright", legend = colnames(mplot), col = colo, lty = 1, lwd = 1, bty = "n")
+box()
+
 bk<-matrix(mssa_indicator_obj$bk_x_array[select,select,,"h=4"],nrow=L)
@@
 mplot<-bk
-colo<-rainbow(ncol(mplot))
-ts.plot(mplot,col=colo,main="M-SSA: h=4",xlab="")
-#for (i in 1:ncol(mplot))
-#  mtext(colnames(mplot)[i],col=colo[i],line=-i)
+plot(mplot[,1],main="M-SSA: h=4",axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
+
+for (i in 1:ncol(mplot))
+{
+  lines(mplot[,i],col=colo[i],lwd=1,lty=1)
+}
+abline(h=0)
+axis(1,at=c(1,1+5*1:(nrow(mplot)/5)),labels=c(0,5*1:(nrow(mplot)/5)))
+axis(2)
+box()
 
@@
 axis(2)
+legend("bottomleft", legend = colnames(mplot), col = colo, lty = 1, lwd = 1, bty = "n")
 box()
@@
   axis(2)
+  if (h==3)
+    legend("bottomleft", legend = colnames(mplot), col = colo, lty = 1, lwd = 1, bty = "n")
+
   box()
@@
 rownames(mplot)<-rownames(x_mat)
+mplot<-mplot[L:nrow(mplot),]
 colo<-c("black",rainbow(ncol(mplot)-1))
@@
 axis(2)
+legend("bottomleft", legend = colnames(mplot), col = colo, lty = 1, lwd = 1, bty = "n")
+
 box()
@@
 pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
+
 j_now<-1
@@
                    paste("Subseries ",select_vec_multi,sep=""))
-colo<-c("blue",rainbow(length(select_vec_multi)))
+mplot<-mplot[L:nrow(mplot),]
+colo<-c("blue","black", "red", "green", "blue", "orange", "dimgray")
+
 main_title<-c(paste("M-SSA predictor for h=",h_vec[j_now]," (solid blue) and sub-series (dashed lines)",sep=""),"In-sample span up to black vertical (dashed) line")
 plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],lwd=2,ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
-mtext(colnames(mplot)[1],col=colo[1],line=-1)
+#mtext(colnames(mplot)[1],col=colo[1],line=-1)
 for (i in 1:ncol(mplot))
@@
   lines(mplot[,i],col=colo[i],lwd=1,lty=2)
-  mtext(colnames(mplot)[i],col=colo[i],line=-i)
+#  mtext(colnames(mplot)[i],col=colo[i],line=-i)
 }
@@
 plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],lwd=1,ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
-mtext(colnames(mplot)[1],col=colo[1],line=-1)
+#mtext(colnames(mplot)[1],col=colo[1],line=-1)
 for (i in 1:ncol(mplot))
@@
   lines(mplot[,i],col=colo[i],lwd=1,lty=1)
-  mtext(colnames(mplot)[i],col=colo[i],line=-i)
+#  mtext(colnames(mplot)[i],col=colo[i],line=-i)
 }
@@
 axis(2)
-box()
+legend("bottomleft", legend = colnames(mplot), col = colo, lty = 1, lwd = 1, bty = "n")
+box()
+
 dev.off()
@@
 plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
-mtext(colnames(mplot)[1],col=colo[1],line=-1)
+#mtext(colnames(mplot)[1],col=colo[1],line=-1)
 for (jj in 1:ncol(mplot))
@@
   lines(mplot[,jj],col=colo[jj],lwd=1,lty=1)
-  mtext(colnames(mplot)[jj],col=colo[jj],line=-jj)
+#  mtext(colnames(mplot)[jj],col=colo[jj],line=-jj)
 }
@@
 # Plot MA inversions
-
 file = "./Figures/up_dated_ma_inv_multi_ip.pdf"
 pdf(file = paste(path.out,file,sep=""), paper = "special", width = 7, height = 7)
-colo<-rainbow(ncol(x_mat))
+colo<-c("black", "red", "green", "blue", "orange", "dimgray")
 # Select BIP component
-i<-1
+ih<-1
 par(mfrow=c(1,2))
 # 2008
-xi<-xi_2008
-mplot<-xi[i,1:min(10,L)]
+xi<-compute_xi_func(V_obj_2008,L)$xi
+mplot<-xi[ih,1:min(10,L)]
 for (j in 2:n)
 {
-  mplot<-cbind(mplot,xi[i,(j-1)*L+1:min(10,L)])
-}
-ts.plot(mplot,col=colo_i,main="MA inversion GDP: 2008",xlab="")
-for (k in 1:length(select_vec_i))
-  mtext(select_vec_i[k],line=-k,col=colo_i[k])
+  mplot<-cbind(mplot,xi[ih,(j-1)*L+1:min(10,L)])
+}
+colnames(mplot)<-select_vec_multi
+plot(mplot[,1],main=paste("MA inversion GDP: 2008",sep=""),axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
+
+for (i in 1:ncol(mplot))
+{
+  lines(mplot[,i],col=colo[i],lwd=1,lty=1)
+}
+abline(h=0)
+axis(1,at=1:nrow(mplot),labels=1:nrow(mplot)-1)
+axis(2)
+legend("topright", legend = colnames(mplot), col = colo, lty = 1, lwd = 1, bty = "n")
+box()
 
 # full data set
-xi<-xi_full
-mplot<-xi[i,1:min(10,L)]
+xi<-compute_xi_func(V_obj_full,L)$xi
+
+mplot<-xi[ih,1:min(10,L)]
 for (j in 2:n)
 {
-  mplot<-cbind(mplot,xi[i,(j-1)*L+1:min(10,L)])
-}
-ts.plot(mplot,col=colo_i,main="Full data set",xlab="")
-#for (k in 1:length(select_vec_i))
-#  mtext(select_vec_i[k],line=-k,col=colo_i[k])
+  mplot<-cbind(mplot,xi[ih,(j-1)*L+1:min(10,L)])
+}
+
+plot(mplot[,1],main=paste("Full data set",sep=""),axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
+
+for (i in 1:ncol(mplot))
+{
+  lines(mplot[,i],col=colo[i],lwd=1,lty=1)
+}
+abline(h=0)
+axis(1,at=1:nrow(mplot),labels=1:nrow(mplot)-1)
+axis(2)
+box()
 
@@
 pdf(file = paste(path.out,file,sep=""), paper = "special", width = 7, height = 7)
+colo<-c("black", "red", "green", "blue", "orange", "dimgray")
 
@@
 mplot<-bk_2008
-colo<-rainbow(ncol(mplot))
 ts.plot(mplot,col=colo,main=paste("M-SSA data up to 2008: h=",h,sep=""),xlab="")
-for (i in 1:ncol(mplot))
-  mtext(colnames(mplot)[i],col=colo[i],line=-i)
+#for (i in 1:ncol(mplot))
+#  mtext(colnames(mplot)[i],col=colo[i],line=-i)
+legend("topright", legend = colnames(mplot), col = colo, lty = 1, lwd = 1, bty = "n")
 mplot<-bk_all
-colo<-rainbow(ncol(mplot))
 ts.plot(mplot,col=colo,main="Full data set",xlab="")
@@
 cat("\\includegraphics[height=3in, width=5in]{", file, "}\n",sep = "")
-cat("\\caption{MA-inversion (impulse response) of VAR: data up to Jan-2008 (left) vs. entire data set (right).", sep = "")
+cat("\\caption{MA-inversion (impulse response) of VAR: data up to Jan-2008 (left) vs. entire data set  without pandemic (right).", sep = "")
 cat("\\label{up_dated_ma_inv_multi_ip}}", sep = "")
@@
 cat("\\includegraphics[height=3in, width=5in]{", file, "}\n",sep = "")
-cat("\\caption{M-SSA filter tracking HP-GDP at forecast horizon $h=4$: data up to Jan-2008 (left) vs. entire data set (right).", sep = "")
+cat("\\caption{M-SSA filter tracking HP-GDP at forecast horizon $h=4$: data up to Jan-2008 (left) vs. entire data set without pandemic (right).", sep = "")
 cat("\\label{bk_2008_all}}", sep = "")

### 2025-05-10 Sweave/BIP_predictor.Rnw
up

@@
 mplot<-gammak
+colnames(mplot)[1]<-"GDP"
 plot(mplot[,1],main="M-MSE",axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
@@
 mplot<-bk
+colnames(mplot)[1]<-"GDP"
 plot(mplot[,1],main="M-SSA: h=0",axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
@@
   mplot<-mplot[which(rownames(mplot)>2006&rownames(mplot)<2011),]
-  colnames(mplot)<-c("BIP","M-SSA","HP-C")
+  colnames(mplot)<-c("GDP","M-SSA","HP-C")
   colo<-c("black","blue","red")
@@
 colo<-c("black",rainbow(ncol(mplot)-1))
-colnames(mplot)[1]<-"BIP"
+colnames(mplot)[1]<-"GDP"
 main_title<-"Equally-weighted average of M-SSA outputs"
@@
 colo<-c("blue","black", "red", "green", "blue", "orange", "dimgray")
+colnames(mplot)[2]<-"Subseries GDP"
 
@@
 colnames(mplot)<-select_vec_multi
+colnames(mplot)[1]<-"GDP"
 plot(mplot[,1],main=paste("MA inversion GDP: 2008",sep=""),axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
@@
 mplot<-bk_2008
+colnames(mplot)[1]<-"GDP"
 ts.plot(mplot,col=colo,main=paste("M-SSA data up to 2008: h=",h,sep=""),xlab="")

### 2025-05-12 Latex/BIP_predictor.tex
up

@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sat May 10 15:26:21 2025
+% Sun May 11 12:36:10 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sat May 10 15:26:21 2025
+% Sun May 11 12:36:10 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sat May 10 15:26:24 2025
+% Sun May 11 12:36:11 2025
 \begin{table}[ht]
@@
 The two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sat May 10 15:26:25 2025
+% Sun May 11 12:36:12 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sat May 10 15:26:30 2025
+% Sun May 11 12:36:15 2025
 \begin{table}[ht]
@@
 
-%Out-of-sample performance metrics include p-values derived from regressions of the out-of-sample predictors on forward-shifted BIP, adjusted for heteroscedasticity and autocorrelation using HAC estimators.
-Additionally, we report relative root-mean-square forecast errors (rRMSE) benchmarked against various reference models.\\
+Tables \ref{p_val_wc} and \ref{rRMSE_mSSA_comp_mean_wc} present HAC-adjusted p-values and rRMSEs for the GDP component predictor. They are obtained from the simple regressions
 
-Tables \eqref{p_val_wc} and \eqref{rRMSE_mSSA_comp_mean_wc} present HAC-adjusted p-values and rRMSEs for the BIP component predictor.
-The p-values are obtained from a regression of the out-of-sample predictor on forward-shifted BIP (in contrast to Table \eqref{p_val_hpbip_wc}, which emphasized HP-BIP); hence, they do not account for potential bias or mis-scaling of the predictor.
-In contrast, the rRMSEs, evaluated against the mean benchmark, do incorporate eventual bias and mis-scaling effects.
+\begin{equation}
+    \textrm{GDP}_{t+shift}=\beta_0(shift,h)+\beta_1(shift,h) \cdot \textrm{M-SSA}^{\textrm{GDP}}_t(h) + \epsilon_t^{shift,h}
+    \label{eq:GDP_M-SSA}
+\end{equation}
+
+where $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$ is the out-of-sample M-SSA predictor for GDP (not HP-GDP.)
+%(in contrast to Table \ref{p_val_hpbip_wc}, which emphasized HP-GDP.) 
+The p-values test $H_0 : 0 = \beta_1(shift,h)$ and do not account for potential bias or mis-scaling of the GDP predictor.
+In contrast, the rRMSEs, evaluated against the mean benchmark, incorporate eventual bias and mis-scaling effects.
 Additional results involving other predictor benchmarks are provided in the appendix, based on data sets with and without the pandemic.
-The results can be summarized as follows: targeting future HP-BIP (see Table \eqref{p_val_hpbip_wc}) is less challenging than predicting BIP (see Table \eqref{p_val_wc}), due to smoothness (autocorrelation) of the target when compared to BIP.
-The pandemic period generally obscures the relationships between predictors and BIP.
-When excluding the pandemic, the direct forecasts outperform the mean forecast up to two quarters ahead.
-The M-SSA component predictor, optimized for longer forecast horizons, outperforms the mean benchmark up to one year ahead.
-Additionally, the BIP component predictor also surpasses the direct forecasts for larger forecast horizons ($h>2$ quarters).Statistical significance of the BIP component predictor, optimized for larger forecast horizons, can be established for at least up to one year ahead.
-Predictors optimized for $h\geq 4$ generally outperform the other designs, irrespective of the forward-shift of BIP.
+\\
+
+The purposes of Equation \ref{eq:GDP_M-SSA} are twofold: first, the regression parameters enable affine transformations of the left-hand side variable, so that the standardized target $\textrm{GDP}_{t+shift}$ can be replaced by the original log-returns without affecting the p-values or rRMSEs reported in Tables \ref{p_val_wc} and \ref{rRMSE_mSSA_comp_mean_wc} (noting that the benchmark predictors, i.e., the direct forecasts and the arithmetic mean, similarly adjust to affine transformations); second, the static scale adjustment by $\beta_1(shift,h)$ compensates for the increasing zero-shrinkage of $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$, reflecting greater uncertainty  as the forecast horizon $h$ increases.
+This simple transformation allows $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$ to be used as a predictor for $\textrm{GDP}_{t+shift}$ even when $h\neq shift$.
+Such flexibility enables assessment of the dynamic left-shift of $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$, as illustrated in Figure \ref{{multivar_vs_univar}}, at various forward-shifts of $\textrm{GDP}_{t+shift}$.\\
+
+When excluding the pandemic, the direct forecasts outperform the mean forecast up to two quarters ahead, see Table \ref{rRMSE_mSSA_direct_mean_without_covid8}.\footnote{However, the pandemic period generally obscures the relationships between predictors and GDP.} However, the M-SSA component predictor, optimized for longer forecast horizons, outperforms the mean benchmark up to 6Q ahead.
+Additionally, the GDP component predictor optimized for large forecast horizons $h\geq 4$ surpass the direct forecasts for $shift>2$, see Table \ref{rRMSE_mSSA_comp_direct_without_covid7}.
+Statistical significance of the GDP component predictor, optimized for larger forecast horizons, can be established for at least up to one year ahead.
 
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sat May 10 15:26:30 2025
+% Sun May 11 12:36:15 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sat May 10 15:26:31 2025
+% Sun May 11 12:36:16 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sat May 10 15:26:31 2025
+% Sun May 11 12:36:16 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 1.144 & 1.102 & 1.045 & 0.999 & 1.006 & 1.059 & 1.125 \\ 
-  Shift=1 & 1.238 & 1.213 & 1.151 & 1.060 & 1.004 & 1.007 & 1.047 \\ 
-  Shift=2 & 1.178 & 1.186 & 1.171 & 1.108 & 1.035 & 0.999 & 0.999 \\ 
-  Shift=3 & 1.011 & 1.035 & 1.057 & 1.039 & 0.988 & 0.948 & 0.933 \\ 
-  Shift=4 & 1.010 & 1.045 & 1.086 & 1.094 & 1.051 & 1.002 & 0.968 \\ 
-  Shift=5 & 0.982 & 1.019 & 1.064 & 1.081 & 1.051 & 1.000 & 0.957 \\ 
+Shift=0 & 1.091 & 1.051 & 0.996 & 0.953 & 0.959 & 1.010 & 1.073 \\ 
+  Shift=1 & 1.215 & 1.190 & 1.129 & 1.040 & 0.985 & 0.988 & 1.027 \\ 
+  Shift=2 & 1.121 & 1.128 & 1.114 & 1.055 & 0.985 & 0.950 & 0.951 \\ 
+  Shift=3 & 0.959 & 0.982 & 1.003 & 0.986 & 0.937 & 0.900 & 0.885 \\ 
+  Shift=4 & 0.950 & 0.982 & 1.021 & 1.028 & 0.988 & 0.942 & 0.910 \\ 
+  Shift=5 & 0.935 & 0.970 & 1.013 & 1.029 & 1.000 & 0.952 & 0.911 \\ 
    \hline
@@
   \hline
-RRMSE & 0.855 & 0.874 & 0.919 & 1.023 & 0.985 & 1.010 \\ 
+RRMSE & 0.897 & 0.890 & 0.966 & 1.078 & 1.047 & 1.062 \\ 
    \hline
@@
   \hline
-Shift=0 & 1.210 & 1.188 & 1.157 & 1.127 & 1.116 & 1.130 & 1.161 \\ 
-  Shift=1 & 0.982 & 0.977 & 0.957 & 0.925 & 0.902 & 0.895 & 0.899 \\ 
-  Shift=2 & 1.053 & 1.054 & 1.046 & 1.018 & 0.987 & 0.973 & 0.975 \\ 
-  Shift=3 & 1.046 & 1.056 & 1.062 & 1.050 & 1.022 & 0.999 & 0.988 \\ 
-  Shift=4 & 0.969 & 0.984 & 1.003 & 1.010 & 0.996 & 0.977 & 0.964 \\ 
-  Shift=5 & 0.982 & 0.998 & 1.019 & 1.027 & 1.014 & 0.990 & 0.967 \\ 
+Shift=0 & 1.165 & 1.144 & 1.114 & 1.085 & 1.075 & 1.089 & 1.118 \\ 
+  Shift=1 & 0.942 & 0.936 & 0.917 & 0.887 & 0.864 & 0.857 & 0.862 \\ 
+  Shift=2 & 1.041 & 1.042 & 1.034 & 1.006 & 0.976 & 0.962 & 0.964 \\ 
+  Shift=3 & 1.023 & 1.032 & 1.039 & 1.027 & 0.999 & 0.977 & 0.966 \\ 
+  Shift=4 & 0.946 & 0.961 & 0.980 & 0.986 & 0.972 & 0.954 & 0.942 \\ 
+  Shift=5 & 0.943 & 0.959 & 0.979 & 0.987 & 0.974 & 0.951 & 0.929 \\ 
    \hline
@@
   \hline
-RRMSE & 0.833 & 1.083 & 0.981 & 0.979 & 1.027 & 1.020 \\ 
+RRMSE & 0.865 & 1.130 & 0.993 & 1.001 & 1.051 & 1.061 \\ 
    \hline

### 2025-05-12 Sweave/BIP_predictor.Rnw
up

@@
 
-%Out-of-sample performance metrics include p-values derived from regressions of the out-of-sample predictors on forward-shifted BIP, adjusted for heteroscedasticity and autocorrelation using HAC estimators.
-Additionally, we report relative root-mean-square forecast errors (rRMSE) benchmarked against various reference models.\\
-
-Tables \eqref{p_val_wc} and \eqref{rRMSE_mSSA_comp_mean_wc} present HAC-adjusted p-values and rRMSEs for the BIP component predictor.
-The p-values are obtained from a regression of the out-of-sample predictor on forward-shifted BIP (in contrast to Table \eqref{p_val_hpbip_wc}, which emphasized HP-BIP); hence, they do not account for potential bias or mis-scaling of the predictor.
-In contrast, the rRMSEs, evaluated against the mean benchmark, do incorporate eventual bias and mis-scaling effects.
+Tables \ref{p_val_wc} and \ref{rRMSE_mSSA_comp_mean_wc} present HAC-adjusted p-values and rRMSEs for the GDP component predictor. They are obtained from the simple regressions
+
+\begin{equation}
+    \textrm{GDP}_{t+shift}=\beta_0(shift,h)+\beta_1(shift,h) \cdot \textrm{M-SSA}^{\textrm{GDP}}_t(h) + \epsilon_t^{shift,h}
+    \label{eq:GDP_M-SSA}
+\end{equation}
+
+where $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$ is the out-of-sample M-SSA predictor for GDP (not HP-GDP.)
+%(in contrast to Table \ref{p_val_hpbip_wc}, which emphasized HP-GDP.) 
+The p-values test $H_0 : 0 = \beta_1(shift,h)$ and do not account for potential bias or mis-scaling of the GDP predictor.
+In contrast, the rRMSEs, evaluated against the mean benchmark, incorporate eventual bias and mis-scaling effects.
 Additional results involving other predictor benchmarks are provided in the appendix, based on data sets with and without the pandemic.
-The results can be summarized as follows: targeting future HP-BIP (see Table \eqref{p_val_hpbip_wc}) is less challenging than predicting BIP (see Table \eqref{p_val_wc}), due to smoothness (autocorrelation) of the target when compared to BIP.
-The pandemic period generally obscures the relationships between predictors and BIP.
-When excluding the pandemic, the direct forecasts outperform the mean forecast up to two quarters ahead.
-The M-SSA component predictor, optimized for longer forecast horizons, outperforms the mean benchmark up to one year ahead.
-Additionally, the BIP component predictor also surpasses the direct forecasts for larger forecast horizons ($h>2$ quarters).Statistical significance of the BIP component predictor, optimized for larger forecast horizons, can be established for at least up to one year ahead.
-Predictors optimized for $h\geq 4$ generally outperform the other designs, irrespective of the forward-shift of BIP.
+\\
+
+The purposes of Equation \ref{eq:GDP_M-SSA} are twofold: first, the regression parameters enable affine transformations of the left-hand side variable, so that the standardized target $\textrm{GDP}_{t+shift}$ can be replaced by the original log-returns without affecting the p-values or rRMSEs reported in Tables \ref{p_val_wc} and \ref{rRMSE_mSSA_comp_mean_wc} (noting that the benchmark predictors, i.e., the direct forecasts and the arithmetic mean, similarly adjust to affine transformations); second, the static scale adjustment by $\beta_1(shift,h)$ compensates for the increasing zero-shrinkage of $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$, reflecting greater uncertainty  as the forecast horizon $h$ increases.
+This simple transformation allows $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$ to be used as a predictor for $\textrm{GDP}_{t+shift}$ even when $h\neq shift$.
+Such flexibility enables assessment of the dynamic left-shift of $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$, as illustrated in Figure \ref{{multivar_vs_univar}}, at various forward-shifts of $\textrm{GDP}_{t+shift}$.\\
+
+When excluding the pandemic, the direct forecasts outperform the mean forecast up to two quarters ahead, see Table \ref{rRMSE_mSSA_direct_mean_without_covid8}.\footnote{However, the pandemic period generally obscures the relationships between predictors and GDP.} However, the M-SSA component predictor, optimized for longer forecast horizons, outperforms the mean benchmark up to 6Q ahead.
+Additionally, the GDP component predictor optimized for large forecast horizons $h\geq 4$ surpass the direct forecasts for $shift>2$, see Table \ref{rRMSE_mSSA_comp_direct_without_covid7}.
+Statistical significance of the GDP component predictor, optimized for larger forecast horizons, can be established for at least up to one year ahead.
 
@@
     setTxtProgressBar(pb, shift)
-    for (j in h_vec)#j<-0
+    for (j in h_vec)#j<-1
     {
@@
       colnames(dat)<-c(colnames(x_mat)[1],sel_vec_pred)
+      nrow(na.exclude(dat))
+
 # Apply the previous function: compute GARCH, WLS regression, out-of-sample MSEs and p-values    
@@
       rownames(dat)<-rownames(x_mat)
-      dat<-dat
+# Two variants for data set:  
+      if (F)
+      {
+# 1. Same sample as M-SSA (due to filter initialization the first L observations are lost) 
+        dat<-dat[L:nrow(dat),]
+# Same length as M-SSA      
+        nrow(na.exclude(dat))
+      }
+# 2. Full data, including the first L observations (makes sense since the data is available) 
+      dat<-dat[1:nrow(dat),]
 

### 2025-05-12 R/M_SSA_utility_functions.R

@@ -598,7 +598,7 @@ optimal_weight_predictor_func<-function(dat,in_out_separator,use_garch,shift,lag
   cal_oos_pred<-cal_oos_mean_pred<-rep(NA,len)
   track_weights<-NULL
   start_non_na<-which(!is.na(dat[,2]))[1]
-  for (i in (n+start_non_na):(len-shift-lag_vec[1])) #i<-n+2
+  for (i in (n+start_non_na):(len-shift-lag_vec[1])) #i<-n+start_non_na+2
   {
 # If use_garch==T then the regression relies on weighted least-squares, whereby the weights are based 
 #   on volatility obtained from a GARCH(1,1) model fitted to target (first column of dat)
@@ -630,6 +630,10 @@ optimal_weight_predictor_func<-function(dat,in_out_separator,use_garch,shift,lag
       cal_oos_pred[i+shift+lag_vec[1]]<-(lm_obj$coef[1]+lm_obj$coef[2:(n+1)]%*%dat[i+shift+lag_vec[1],2:(n+1)]) 
     }
 # 2. Use mean as predictor (simplest benchmark)
+# Two variants: 
+# 1. Same data span as filter (due to initialization the filter losts L values at start)    
+    cal_oos_mean_pred[i+shift+lag_vec[1]]<-mean(dat[start_non_na:i,1])
+# 2. Full sample, including first L observations (makes sense since the data is available)    
     cal_oos_mean_pred[i+shift+lag_vec[1]]<-mean(dat[1:i,1])
   }
   colnames(track_weights)<-c("intercept",colnames(dat)[2:ncol(dat)])

### 2025-05-13 Latex/BIP_predictor.tex
up

@@
 
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sun May 11 12:36:10 2025
+% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Mon May 12 12:05:30 2025
 \begin{table}[ht]
@@
 \label{f_stat}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sun May 11 12:36:10 2025
+\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Mon May 12 12:05:30 2025
 \begin{table}[ht]
@@
 
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sun May 11 12:36:11 2025
+% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Mon May 12 12:05:31 2025
 \begin{table}[ht]
@@
 The dashed vertical line delimits in- and out-of-sample spans.
-The two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sun May 11 12:36:12 2025
+The two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Mon May 12 12:05:31 2025
 \begin{table}[ht]
@@
 Table \eqref{p_val_hpbip_wc} reports HAC-adjusted p-values from regressions of the out-of-sample predictor on HP-BIP, indicating a strong link between the predictor and the target, out-of-sample.
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sun May 11 12:36:15 2025
+% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Mon May 12 12:05:32 2025
 \begin{table}[ht]
@@
 
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sun May 11 12:36:15 2025
+% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Mon May 12 12:05:32 2025
 \begin{table}[ht]
@@
 \label{p_val_wc}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sun May 11 12:36:15 2025
+\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Mon May 12 12:05:32 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 0.978 & 0.942 & 0.893 & 0.855 & 0.860 & 0.906 & 0.962 \\ 
-  Shift=1 & 1.082 & 1.060 & 1.005 & 0.926 & 0.877 & 0.880 & 0.914 \\ 
-  Shift=2 & 1.083 & 1.090 & 1.077 & 1.019 & 0.952 & 0.918 & 0.919 \\ 
-  Shift=3 & 1.034 & 1.059 & 1.081 & 1.063 & 1.010 & 0.970 & 0.954 \\ 
-  Shift=4 & 0.995 & 1.029 & 1.069 & 1.077 & 1.035 & 0.986 & 0.953 \\ 
-  Shift=5 & 0.993 & 1.030 & 1.075 & 1.093 & 1.062 & 1.011 & 0.967 \\ 
+Shift=0 & 0.973 & 0.937 & 0.888 & 0.850 & 0.855 & 0.900 & 0.956 \\ 
+  Shift=1 & 1.073 & 1.051 & 0.997 & 0.918 & 0.870 & 0.873 & 0.906 \\ 
+  Shift=2 & 1.069 & 1.077 & 1.063 & 1.006 & 0.940 & 0.907 & 0.907 \\ 
+  Shift=3 & 1.020 & 1.045 & 1.066 & 1.048 & 0.997 & 0.957 & 0.941 \\ 
+  Shift=4 & 0.983 & 1.016 & 1.056 & 1.064 & 1.023 & 0.974 & 0.942 \\ 
+  Shift=5 & 0.988 & 1.026 & 1.071 & 1.088 & 1.057 & 1.007 & 0.963 \\ 
    \hline
@@
 
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sun May 11 12:36:16 2025
+% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Mon May 12 12:05:32 2025
 \begin{table}[ht]
@@
 \label{p_val_wc5}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sun May 11 12:36:16 2025
+\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Mon May 12 12:05:32 2025
 \begin{table}[ht]
@@
 \label{rRMSE_mSSA_comp_mean6}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sun May 11 12:36:16 2025
+\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Mon May 12 12:05:32 2025
 \begin{table}[ht]
@@
 \label{rRMSE_mSSA_comp_direct_without_covid7}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sun May 11 12:36:16 2025
+\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Mon May 12 12:05:32 2025
 \begin{table}[ht]
@@
   \hline
-RRMSE & 0.897 & 0.890 & 0.966 & 1.078 & 1.047 & 1.062 \\ 
+RRMSE & 0.892 & 0.883 & 0.954 & 1.063 & 1.035 & 1.058 \\ 
    \hline
@@
 \label{p_val1}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sun May 11 12:36:16 2025
+\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Mon May 12 12:05:32 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 1.008 & 0.990 & 0.964 & 0.939 & 0.930 & 0.942 & 0.967 \\ 
-  Shift=1 & 1.064 & 1.057 & 1.036 & 1.001 & 0.976 & 0.968 & 0.973 \\ 
-  Shift=2 & 1.033 & 1.034 & 1.026 & 0.999 & 0.969 & 0.955 & 0.957 \\ 
-  Shift=3 & 1.025 & 1.034 & 1.040 & 1.028 & 1.001 & 0.979 & 0.967 \\ 
-  Shift=4 & 0.995 & 1.011 & 1.030 & 1.037 & 1.022 & 1.003 & 0.990 \\ 
-  Shift=5 & 1.001 & 1.018 & 1.039 & 1.048 & 1.034 & 1.009 & 0.986 \\ 
+Shift=0 & 1.005 & 0.986 & 0.960 & 0.935 & 0.927 & 0.939 & 0.964 \\ 
+  Shift=1 & 1.060 & 1.054 & 1.032 & 0.998 & 0.973 & 0.965 & 0.970 \\ 
+  Shift=2 & 1.028 & 1.029 & 1.020 & 0.993 & 0.963 & 0.950 & 0.952 \\ 
+  Shift=3 & 1.018 & 1.027 & 1.034 & 1.022 & 0.994 & 0.972 & 0.961 \\ 
+  Shift=4 & 0.990 & 1.006 & 1.025 & 1.032 & 1.017 & 0.998 & 0.985 \\ 
+  Shift=5 & 1.000 & 1.016 & 1.037 & 1.046 & 1.032 & 1.008 & 0.985 \\ 
    \hline
@@
 \label{rRMSE_mSSA_comp_mean2}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sun May 11 12:36:16 2025
+\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Mon May 12 12:05:32 2025
 \begin{table}[ht]
@@
 \label{rRMSE_mSSA_comp_direct3}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Sun May 11 12:36:16 2025
+\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
+% Mon May 12 12:05:32 2025
 \begin{table}[ht]
@@
   \hline
-RRMSE & 0.865 & 1.130 & 0.993 & 1.001 & 1.051 & 1.061 \\ 
+RRMSE & 0.862 & 1.126 & 0.987 & 0.995 & 1.046 & 1.060 \\ 
    \hline

### 2025-05-13 Latex/BIP_predictor.tex
up

@@
 
-% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Mon May 12 12:05:30 2025
+% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Tue May 13 07:02:03 2025
 \begin{table}[ht]
@@
 \label{f_stat}
-\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Mon May 12 12:05:30 2025
+\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Tue May 13 07:02:03 2025
 \begin{table}[ht]
@@
 
-% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Mon May 12 12:05:31 2025
+% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Tue May 13 07:02:05 2025
 \begin{table}[ht]
@@
 The dashed vertical line delimits in- and out-of-sample spans.
-The two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Mon May 12 12:05:31 2025
+The two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Tue May 13 07:02:05 2025
 \begin{table}[ht]
@@
 Table \eqref{p_val_hpbip_wc} reports HAC-adjusted p-values from regressions of the out-of-sample predictor on HP-BIP, indicating a strong link between the predictor and the target, out-of-sample.
-% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Mon May 12 12:05:32 2025
+% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Tue May 13 07:02:08 2025
 \begin{table}[ht]
@@
 
-% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Mon May 12 12:05:32 2025
+% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Tue May 13 07:02:08 2025
 \begin{table}[ht]
@@
 \label{p_val_wc}
-\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Mon May 12 12:05:32 2025
+\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Tue May 13 07:02:08 2025
 \begin{table}[ht]
@@
 
-% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Mon May 12 12:05:32 2025
+% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Tue May 13 07:02:09 2025
 \begin{table}[ht]
@@
 \label{p_val_wc5}
-\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Mon May 12 12:05:32 2025
+\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Tue May 13 07:02:09 2025
 \begin{table}[ht]
@@
 \label{rRMSE_mSSA_comp_mean6}
-\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Mon May 12 12:05:32 2025
+\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Tue May 13 07:02:09 2025
 \begin{table}[ht]
@@
 \label{rRMSE_mSSA_comp_direct_without_covid7}
-\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Mon May 12 12:05:32 2025
+\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Tue May 13 07:02:09 2025
 \begin{table}[ht]
@@
 \label{p_val1}
-\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Mon May 12 12:05:32 2025
+\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Tue May 13 07:02:09 2025
 \begin{table}[ht]
@@
 \label{rRMSE_mSSA_comp_mean2}
-\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Mon May 12 12:05:32 2025
+\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Tue May 13 07:02:09 2025
 \begin{table}[ht]
@@
 \label{rRMSE_mSSA_comp_direct3}
-\end{table}% latex table generated in R 4.4.1 by xtable 1.8-4 package
-% Mon May 12 12:05:32 2025
+\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Tue May 13 07:02:09 2025
 \begin{table}[ht]

### 2025-05-13 Sweave/BIP_predictor.Rnw
up

@@
 lag_vec<-c(1,rep(0,ncol(x_mat)-1))
+
+# Use OLS or WLS in regression of shifted BIP on M-SSA component predictor 
+use_garch<-F
 
@@
 
-sel_vec_pred<-c("BIP")
-
+# Select M-SSA components: use M-SSA output tracking BIP
+sel_vec_mssa_comp<-select_vec_multi
+sel_vec_mssa_comp<-c("BIP")
+# Select indicators for direct forecast: all indicators lead to overfitting (worse out-of-sample performances)
+# Best design out-of-sample is based on ifo_c and ESI
+sel_vec_direct_forecast<-select_vec_multi
+sel_vec_direct_forecast<-c("ifo_c","ESI")
+# M-SSA looses L observations at start when compared to direct forecast or mean benchmarks: 
+#   -therefore the samples are slightly different
+# If align_sample==T then the first L observations of mean and direct forecasts are removed to align with M-SSA
+#   -Avoid random differences due to unequal samples.
+align_sample<-T
+# Shift BIP forward by shift_vec + publication lag
 shift_vec<-0:5
@@
 {
+# Select indicators for direct forecast: use all indicators
+  sel_indicator_out_sample<-select_vec_multi
 # Initialize performance matrices
-  MSE_oos_mssa_comp_without_covid_mat<-MSE_oos_mssa_comp_mat<-p_mat_mssa<-p_mat_mssa_components<-p_mat_mssa_components_without_covid<-p_mat_direct<-rRMSE_mSSA_comp_direct<-rRMSE_mSSA_comp_mean<-rRMSE_mSSA_comp_direct_without_covid<-rRMSE_mSSA_comp_mean_without_covid<-rRMSE_mSSA_direct_mean_without_covid<-rRMSE_mSSA_direct_mean<-p_mat_direct_without_covid<-matrix(ncol=length(h_vec),nrow=length(h_vec)-1)
+  MSE_oos_mssa_comp_without_covid_mat<-MSE_oos_mssa_comp_mat<-p_mat_mssa<-p_mat_mssa_components<-p_mat_mssa_components_without_covid<-p_mat_direct<-rRMSE_mSSA_comp_direct<-rRMSE_mSSA_comp_mean<-rRMSE_mSSA_comp_direct_without_covid<-rRMSE_mSSA_comp_mean_without_covid<-rRMSE_mSSA_direct_mean_without_covid<-rRMSE_mSSA_direct_mean<-p_mat_direct_without_covid<-matrix(nrow=length(shift_vec),ncol=length(h_vec))
   final_components_preditor_array<-oos_components_preditor_array<-array(dim=c(length(shift_vec),length(h_vec),nrow(x_mat)))
-  dimnames(final_components_preditor_array)<-dimnames(oos_components_preditor_array)<-list(paste("shift=",shift_vec,sep=""),
-                                                                                           paste("h=",h_vec,sep=""),rownames(x_mat))
+  dimnames(final_components_preditor_array)<-dimnames(oos_components_preditor_array)<-list(paste("shift=",shift_vec,sep=""), paste("h=",h_vec,sep=""),rownames(x_mat))
+
+  dim(p_mat_mssa_components)
 # Initialize arrays collecting final predictors, real-time predictors and regression weights
 #   -These will be used when analyzing revisions
-  track_weights_array<-array(dim=c(length(shift_vec),length(h_vec),nrow(x_mat),length(sel_vec_pred)+1))
+  track_weights_array<-array(dim=c(length(shift_vec),length(h_vec),nrow(x_mat),length(sel_vec_mssa_comp)+1))
   dimnames(track_weights_array)<-list(paste("shift=",shift_vec,sep=""),
-                                      paste("h=",h_vec,sep=""),rownames(x_mat),c("Intercept",sel_vec_pred))
+                                      paste("h=",h_vec,sep=""),rownames(x_mat),c("Intercept",sel_vec_mssa_comp))
 
 # Use OLS
-  use_garch<-F
 # Set-up progress bar: indicates progress in R-console
@@
 # Specify data matrix for WLS regression
-      if (length(sel_vec_pred)>1)
-      {
-        dat<-cbind(c(x_mat[(shift+lag_vec[1]+1):nrow(x_mat),1],rep(NA,shift+lag_vec[1])),t(mssa_array[sel_vec_pred,,k]))
+      if (length(sel_vec_mssa_comp)>1)
+      {
+        dat<-cbind(c(x_mat[(shift+lag_vec[1]+1):nrow(x_mat),1],rep(NA,shift+lag_vec[1])),t(mssa_array[sel_vec_mssa_comp,,k]))
       } else
       {
-        dat<-cbind(c(x_mat[(shift+lag_vec[1]+1):nrow(x_mat),1],rep(NA,shift+lag_vec[1])),(mssa_array[sel_vec_pred,,k]))
+        dat<-cbind(c(x_mat[(shift+lag_vec[1]+1):nrow(x_mat),1],rep(NA,shift+lag_vec[1])),(mssa_array[sel_vec_mssa_comp,,k]))
       }
       rownames(dat)<-rownames(x_mat)
-      colnames(dat)<-c(colnames(x_mat)[1],sel_vec_pred)
+      colnames(dat)<-c(colnames(x_mat)[1],sel_vec_mssa_comp)
       nrow(na.exclude(dat))
@@
 # Apply the previous function: compute GARCH, WLS regression, out-of-sample MSEs and p-values    
-      perf_obj<-optimal_weight_predictor_func(dat,in_out_separator,use_garch,shift,lag_vec)
+      perf_obj<-optimal_weight_predictor_func(dat,in_out_separator,use_garch,shift,lag_vec,align_sample)
 # Retrieve out-of-sample performances 
@@
 
+#----------------      
 # B. Direct forecasts
@@
 #   -Note that the data matrix here does not depend on j, in contrast  to the M-SSA components above    
-      dat<-cbind(c(x_mat[(shift+lag_vec[1]+1):nrow(x_mat),1],rep(NA,shift+lag_vec[1])),x_mat)
+      dat<-cbind(c(x_mat[(shift+lag_vec[1]+1):nrow(x_mat),1],rep(NA,shift+lag_vec[1])),x_mat[,sel_vec_direct_forecast])
       rownames(dat)<-rownames(x_mat)
 # Two variants for data set:  
-      if (F)
+      if (T)
       {
@@
         nrow(na.exclude(dat))
-      }
+      } else
+      {
 # 2. Full data, including the first L observations (makes sense since the data is available) 
         dat<-dat[1:nrow(dat),]
-
-      perf_obj<-optimal_weight_predictor_func(dat,in_out_separator,use_garch,shift,lag_vec)
+      }
+      perf_obj<-optimal_weight_predictor_func(dat,in_out_separator,use_garch,shift,lag_vec,align_sample)
 # Retrieve out-of-sample performances: p-values and forecast MSE, with/without Pandemic 
@@
 
-if (length(sel_vec_pred)>1)
-{
-  dat<-cbind(c(x_mat_wc[(shift+lag_vec[1]+1):nrow(x_mat_wc),1],rep(NA,shift+lag_vec[1])),t(final_mssa_array[sel_vec_pred,,h+1]))
+if (length(sel_vec_mssa_comp)>1)
+{
+  dat<-cbind(c(x_mat_wc[(shift+lag_vec[1]+1):nrow(x_mat_wc),1],rep(NA,shift+lag_vec[1])),t(final_mssa_array[sel_vec_mssa_comp,,h+1]))
 } else
 {
-  dat<-cbind(c(x_mat_wc[(shift+lag_vec[1]+1):nrow(x_mat_wc),1],rep(NA,shift+lag_vec[1])),(final_mssa_array[sel_vec_pred,,h+1]))
+  dat<-cbind(c(x_mat_wc[(shift+lag_vec[1]+1):nrow(x_mat_wc),1],rep(NA,shift+lag_vec[1])),(final_mssa_array[sel_vec_mssa_comp,,h+1]))
 }
@@
 # Compute predictor for each forward-shift  
-if (length(sel_vec_pred)>1)
+if (length(sel_vec_mssa_comp)>1)
 {

### 2025-05-13 R/M_SSA_utility_functions.R

@@ -588,7 +588,7 @@ HAC_ajusted_p_value_func<-function(da)
 # MSE_mean_oos_without_covid: same but without Pandemic
 # MSE_oos_without_covid: without Pandemic
 # p_value_without_covid: without Pandemic
-optimal_weight_predictor_func<-function(dat,in_out_separator,use_garch,shift,lag_vec)
+optimal_weight_predictor_func<-function(dat,in_out_separator,use_garch,shift,lag_vec,align_sample)
 {
   
   len<-dim(dat)[1]
@@ -631,7 +631,7 @@ optimal_weight_predictor_func<-function(dat,in_out_separator,use_garch,shift,lag
     }
 # 2. Use mean as predictor (simplest benchmark)
 # Two variants: 
-    if (T)
+    if (align_sample)
     {
 # 1. Same data span as filter (due to initialization the filter losts L values at start)    
       cal_oos_mean_pred[i+shift+lag_vec[1]]<-mean(dat[start_non_na:i,1])

### 2025-05-13 Sweave/BIP_predictor.Rnw
up

@@
 align_sample<-T
+reg_type<-"OLS"
+reg_type<-"LASSO"
+reg_type<-"Ridge"
+lambda_reg<-10
 # Shift BIP forward by shift_vec + publication lag
@@
 # Apply the previous function: compute GARCH, WLS regression, out-of-sample MSEs and p-values    
-      perf_obj<-optimal_weight_predictor_func(dat,in_out_separator,use_garch,shift,lag_vec,align_sample)
+      perf_obj<-optimal_weight_predictor_func(dat,in_out_separator,use_garch,shift,lag_vec,align_sample,reg_type,lambda_reg)
 # Retrieve out-of-sample performances 
@@
       }
-      perf_obj<-optimal_weight_predictor_func(dat,in_out_separator,use_garch,shift,lag_vec,align_sample)
+      perf_obj<-optimal_weight_predictor_func(dat,in_out_separator,use_garch,shift,lag_vec,align_sample,reg_type,lambda_reg)
 # Retrieve out-of-sample performances: p-values and forecast MSE, with/without Pandemic 

### 2025-05-13 Latex/BIP_predictor.tex
up

@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Tue May 13 07:02:03 2025
+% Tue May 13 15:04:25 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Tue May 13 07:02:03 2025
+% Tue May 13 15:04:25 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Tue May 13 07:02:05 2025
+% Tue May 13 15:04:27 2025
 \begin{table}[ht]
@@
 The two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Tue May 13 07:02:05 2025
+% Tue May 13 15:04:28 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Tue May 13 07:02:08 2025
+% Tue May 13 15:04:31 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 0.084 & 0.047 & 0.017 & 0.005 & 0.002 & 0.001 & 0.038 \\ 
-  Shift=1 & 0.230 & 0.160 & 0.094 & 0.032 & 0.005 & 0.001 & 0.004 \\ 
-  Shift=2 & 0.801 & 0.666 & 0.414 & 0.173 & 0.026 & 0.005 & 0.006 \\ 
-  Shift=3 & 0.876 & 0.961 & 0.917 & 0.683 & 0.248 & 0.068 & 0.039 \\ 
-  Shift=4 & 0.370 & 0.819 & 0.990 & 0.953 & 0.533 & 0.117 & 0.030 \\ 
-  Shift=5 & 0.353 & 0.802 & 0.983 & 0.943 & 0.571 & 0.239 & 0.043 \\ 
+Shift=0 & 0.086 & 0.049 & 0.018 & 0.005 & 0.002 & 0.002 & 0.042 \\ 
+  Shift=1 & 0.238 & 0.165 & 0.097 & 0.033 & 0.005 & 0.001 & 0.005 \\ 
+  Shift=2 & 0.815 & 0.680 & 0.428 & 0.180 & 0.029 & 0.007 & 0.008 \\ 
+  Shift=3 & 0.866 & 0.963 & 0.922 & 0.699 & 0.269 & 0.079 & 0.047 \\ 
+  Shift=4 & 0.361 & 0.808 & 0.989 & 0.954 & 0.546 & 0.127 & 0.034 \\ 
+  Shift=5 & 0.351 & 0.786 & 0.982 & 0.943 & 0.579 & 0.245 & 0.042 \\ 
    \hline
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Tue May 13 07:02:08 2025
+% Tue May 13 15:04:31 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 0.973 & 0.937 & 0.888 & 0.850 & 0.855 & 0.900 & 0.956 \\ 
-  Shift=1 & 1.073 & 1.051 & 0.997 & 0.918 & 0.870 & 0.873 & 0.906 \\ 
-  Shift=2 & 1.069 & 1.077 & 1.063 & 1.006 & 0.940 & 0.907 & 0.907 \\ 
-  Shift=3 & 1.020 & 1.045 & 1.066 & 1.048 & 0.997 & 0.957 & 0.941 \\ 
-  Shift=4 & 0.983 & 1.016 & 1.056 & 1.064 & 1.023 & 0.974 & 0.942 \\ 
-  Shift=5 & 0.988 & 1.026 & 1.071 & 1.088 & 1.057 & 1.007 & 0.963 \\ 
+Shift=0 & 0.950 & 0.917 & 0.877 & 0.851 & 0.865 & 0.910 & 0.960 \\ 
+  Shift=1 & 1.028 & 1.007 & 0.960 & 0.899 & 0.867 & 0.878 & 0.912 \\ 
+  Shift=2 & 1.040 & 1.043 & 1.030 & 0.984 & 0.933 & 0.910 & 0.913 \\ 
+  Shift=3 & 1.006 & 1.024 & 1.040 & 1.025 & 0.983 & 0.952 & 0.941 \\ 
+  Shift=4 & 0.977 & 1.003 & 1.034 & 1.040 & 1.005 & 0.964 & 0.937 \\ 
+  Shift=5 & 0.984 & 1.011 & 1.044 & 1.057 & 1.031 & 0.986 & 0.949 \\ 
    \hline
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Tue May 13 07:02:09 2025
+% Tue May 13 15:04:31 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Tue May 13 07:02:09 2025
+% Tue May 13 15:04:31 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 1.091 & 1.051 & 0.996 & 0.953 & 0.959 & 1.010 & 1.073 \\ 
-  Shift=1 & 1.215 & 1.190 & 1.129 & 1.040 & 0.985 & 0.988 & 1.027 \\ 
-  Shift=2 & 1.121 & 1.128 & 1.114 & 1.055 & 0.985 & 0.950 & 0.951 \\ 
-  Shift=3 & 0.959 & 0.982 & 1.003 & 0.986 & 0.937 & 0.900 & 0.885 \\ 
-  Shift=4 & 0.950 & 0.982 & 1.021 & 1.028 & 0.988 & 0.942 & 0.910 \\ 
-  Shift=5 & 0.935 & 0.970 & 1.013 & 1.029 & 1.000 & 0.952 & 0.911 \\ 
+Shift=0 & 1.104 & 1.066 & 1.019 & 0.989 & 1.006 & 1.058 & 1.116 \\ 
+  Shift=1 & 1.232 & 1.206 & 1.150 & 1.076 & 1.038 & 1.051 & 1.092 \\ 
+  Shift=2 & 1.146 & 1.150 & 1.136 & 1.085 & 1.028 & 1.002 & 1.006 \\ 
+  Shift=3 & 1.008 & 1.027 & 1.042 & 1.028 & 0.986 & 0.954 & 0.943 \\ 
+  Shift=4 & 0.988 & 1.014 & 1.045 & 1.051 & 1.016 & 0.975 & 0.948 \\ 
+  Shift=5 & 0.967 & 0.995 & 1.027 & 1.039 & 1.013 & 0.970 & 0.933 \\ 
    \hline
@@
   \hline
-RRMSE & 0.892 & 0.883 & 0.954 & 1.063 & 1.035 & 1.058 \\ 
+RRMSE & 0.860 & 0.835 & 0.907 & 0.998 & 0.989 & 1.017 \\ 
    \hline
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Tue May 13 07:02:09 2025
+% Tue May 13 15:04:32 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 0.127 & 0.064 & 0.019 & 0.003 & 0.001 & 0.002 & 0.012 \\ 
-  Shift=1 & 0.439 & 0.314 & 0.194 & 0.093 & 0.033 & 0.017 & 0.036 \\ 
-  Shift=2 & 0.718 & 0.565 & 0.326 & 0.098 & 0.020 & 0.008 & 0.010 \\ 
-  Shift=3 & 0.946 & 0.975 & 0.925 & 0.657 & 0.222 & 0.059 & 0.028 \\ 
-  Shift=4 & 0.327 & 0.699 & 0.964 & 0.942 & 0.619 & 0.237 & 0.102 \\ 
-  Shift=5 & 0.383 & 0.755 & 0.946 & 0.910 & 0.581 & 0.241 & 0.055 \\ 
+Shift=0 & 0.142 & 0.074 & 0.023 & 0.003 & 0.001 & 0.002 & 0.013 \\ 
+  Shift=1 & 0.475 & 0.344 & 0.217 & 0.111 & 0.044 & 0.023 & 0.044 \\ 
+  Shift=2 & 0.724 & 0.564 & 0.325 & 0.101 & 0.023 & 0.010 & 0.012 \\ 
+  Shift=3 & 0.943 & 0.979 & 0.933 & 0.670 & 0.234 & 0.066 & 0.031 \\ 
+  Shift=4 & 0.322 & 0.666 & 0.960 & 0.946 & 0.644 & 0.263 & 0.119 \\ 
+  Shift=5 & 0.386 & 0.721 & 0.940 & 0.910 & 0.591 & 0.248 & 0.060 \\ 
    \hline
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Tue May 13 07:02:09 2025
+% Tue May 13 15:04:32 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 1.005 & 0.986 & 0.960 & 0.935 & 0.927 & 0.939 & 0.964 \\ 
-  Shift=1 & 1.060 & 1.054 & 1.032 & 0.998 & 0.973 & 0.965 & 0.970 \\ 
-  Shift=2 & 1.028 & 1.029 & 1.020 & 0.993 & 0.963 & 0.950 & 0.952 \\ 
-  Shift=3 & 1.018 & 1.027 & 1.034 & 1.022 & 0.994 & 0.972 & 0.961 \\ 
-  Shift=4 & 0.990 & 1.006 & 1.025 & 1.032 & 1.017 & 0.998 & 0.985 \\ 
-  Shift=5 & 1.000 & 1.016 & 1.037 & 1.046 & 1.032 & 1.008 & 0.985 \\ 
+Shift=0 & 0.991 & 0.974 & 0.952 & 0.933 & 0.929 & 0.942 & 0.965 \\ 
+  Shift=1 & 1.037 & 1.030 & 1.012 & 0.985 & 0.966 & 0.963 & 0.969 \\ 
+  Shift=2 & 1.014 & 1.014 & 1.005 & 0.983 & 0.959 & 0.950 & 0.953 \\ 
+  Shift=3 & 1.010 & 1.017 & 1.021 & 1.011 & 0.988 & 0.969 & 0.961 \\ 
+  Shift=4 & 0.987 & 0.999 & 1.015 & 1.020 & 1.008 & 0.992 & 0.981 \\ 
+  Shift=5 & 0.996 & 1.008 & 1.023 & 1.030 & 1.018 & 0.997 & 0.976 \\ 
    \hline
@@
   \hline
-Shift=0 & 1.165 & 1.144 & 1.114 & 1.085 & 1.075 & 1.089 & 1.118 \\ 
-  Shift=1 & 0.942 & 0.936 & 0.917 & 0.887 & 0.864 & 0.857 & 0.862 \\ 
-  Shift=2 & 1.041 & 1.042 & 1.034 & 1.006 & 0.976 & 0.962 & 0.964 \\ 
-  Shift=3 & 1.023 & 1.032 & 1.039 & 1.027 & 0.999 & 0.977 & 0.966 \\ 
-  Shift=4 & 0.946 & 0.961 & 0.980 & 0.986 & 0.972 & 0.954 & 0.942 \\ 
-  Shift=5 & 0.943 & 0.959 & 0.979 & 0.987 & 0.974 & 0.951 & 0.929 \\ 
+Shift=0 & 1.228 & 1.207 & 1.180 & 1.156 & 1.151 & 1.167 & 1.196 \\ 
+  Shift=1 & 0.978 & 0.972 & 0.955 & 0.929 & 0.912 & 0.908 & 0.915 \\ 
+  Shift=2 & 1.030 & 1.029 & 1.021 & 0.998 & 0.974 & 0.964 & 0.968 \\ 
+  Shift=3 & 1.037 & 1.044 & 1.048 & 1.038 & 1.014 & 0.995 & 0.986 \\ 
+  Shift=4 & 0.973 & 0.985 & 1.000 & 1.006 & 0.994 & 0.978 & 0.966 \\ 
+  Shift=5 & 0.981 & 0.993 & 1.008 & 1.015 & 1.003 & 0.982 & 0.962 \\ 
    \hline
@@
   \hline
-RRMSE & 0.862 & 1.126 & 0.987 & 0.995 & 1.046 & 1.060 \\ 
+RRMSE & 0.807 & 1.060 & 0.985 & 0.974 & 1.015 & 1.015 \\ 
    \hline

### 2025-05-13 Sweave/BIP_predictor.Rnw
up

@@
   install.packages("xtable")
+if (!"MASS"%in%inst_pack)
+  install.packages("MASS")
+if (!"glmnet"%in%inst_pack)
+  install.packages("glmnet")
 
@@
 library(sandwich)
-# Extended time series
-library(xts)
 # Library for Diebold-Mariano test of equal forecast performance
@@
 library(xtable)
+library(MASS)
+library(glmnet) 
+# Extended time series
+library(xts)
 
@@
 align_sample<-T
+# Can define type of regression
 reg_type<-"OLS"
@@
 reg_type<-"Ridge"
+# Penalty weight in regularization
 lambda_reg<-10

### 2025-05-13 R/M_SSA_utility_functions.R

@@ -588,101 +588,122 @@ HAC_ajusted_p_value_func<-function(da)
 # MSE_mean_oos_without_covid: same but without Pandemic
 # MSE_oos_without_covid: without Pandemic
 # p_value_without_covid: without Pandemic
-optimal_weight_predictor_func<-function(dat,in_out_separator,use_garch,shift,lag_vec,align_sample)
+optimal_weight_predictor_func<-function(dat,in_out_separator,use_garch,shift,lag_vec,align_sample,reg_type,lambda_reg=NULL)
 {
-  
   len<-dim(dat)[1]
-# First column is target, i.e. dimension is dim(dat)[2]-1
+  # First column is target, i.e. dimension is dim(dat)[2]-1
   n<-dim(dat)[2]-1
-# Compute calibrated out-of-sample predictor, based on expanding window
+  # Compute calibrated out-of-sample predictor, based on expanding window
   cal_oos_pred<-cal_oos_mean_pred<-rep(NA,len)
   track_weights<-NULL
   start_non_na<-which(!is.na(dat[,2]))[1]
-  for (i in (n+start_non_na):(len-shift-lag_vec[1])) #i<-n+start_non_na+2
+  for (i in (n+start_non_na):(len-shift-lag_vec[1])) #i<-n+start_non_na+2  i<-100
   {
-# If use_garch==T then the regression relies on weighted least-squares, whereby the weights are based 
-#   on volatility obtained from a GARCH(1,1) model fitted to target (first column of dat)
+    # If use_garch==T then the regression relies on weighted least-squares, whereby the weights are based 
+    #   on volatility obtained from a GARCH(1,1) model fitted to target (first column of dat)
     if (use_garch)
     {
-      weight<-garch_vola_func(dat,shift,lag_vec)$weight[1:i]
+      weights<-garch_vola_func(dat,shift,lag_vec)$weight[start_non_na:i]
     } else
     {
-# Fixed weight      
-      weight<-rep(1,i)
+      # Fixed weight      
+      weights<-rep(1,i-start_non_na+1)
     }
-#    ts.plot(cbind(dat[1:i,1],1/weight),col=c("red","blue"))
-# 1. Use predictors in columns 2,..., of dat    
-# Fit model with data up to time point i; weighted least-squares relying on weight as defined above  
[…]

### 2025-05-14 Sweave/BIP_predictor.Rnw
up

@@
 library(xtable)
+# Packages for Ridge and LASSO
 library(MASS)

### 2025-05-14 Sweave/BIP_predictor.Rnw
Update BIP_predictor.Rnw

@@
 # Two variants for data set:  
-      if (T)
+      if (align_sample)
       {

### 2025-05-20 Sweave/BIP_predictor.Rnw
up

@@
   install.packages("glmnet")
+if (!"pROC"%in%inst_pack)
+  install.packages("pROC")
 
@@
 library(xts)
+# Extended time series
+library(pROC)
 

### 2025-05-20 R/ROCplots.R

@@ -1,0 +1,79 @@
+# ROCplots : a function for comparing AUCs and ROC plots
+#   by Simon van Norden, HEC Montreal
+#   19-05-2025
+#
+# SYNTAX
+#   
+#   auc <- ROCplots(df, showROC = TRUE, smoothROC = FALSE, showLegend = TRUE)
+#
+# where
+#   df  is a data.frame of series to analyse. See below for format.
+#   showROC  controls creation of the plot comparing ROC curves
+#   smooth   controls whether ROC curves should be smoothed
+#   showLegend controls where a legend is added to the plot of ROC curves
+#
+# VALUE
+#
+#   auc   a 1-column data.frame containing the AUC for each predictor in df.
+#
+# NOTES
+#
+# The data.frame df should contain
+# - the target series in column 1, consisting only of 0s and 1s
+# - an arbitrary number of additional numeric series in the remaining columns
+# All series should be the same length and contain no NAs
+
+inst_pack<-rownames(installed.packages())
+if (!"pROC"%in%inst_pack)
+install.packages("pROC")
+
+library(pROC)
+
+ROCplots <- function(df, showROC = T, smoothROC = FALSE, showLegend = TRUE) {
+  df_names <- names(df) # series names
+  k <- length(df_names) # 1 + number of predictors to compare
+  print(paste("Target variable is", df_names[1]))
+  # Create table to hold AUCs
+  aurocs <- as.data.frame(matrix(NA, nrow = k-1))
+  row.names(aurocs) <- df_names[-1]
+  names(aurocs) <- "AUC"
+  
+  # make the first plot and store the first AUC
+  ROC_OBJ <- roc_(df, response = df_names[1], 
+                  predictor = df_names[2],
+                  quiet = T, plot = showROC,
+                  smooth = smoothROC)
+  aurocs[1,1] <- ROC_OBJ$auc
+  
+  if (k>2) {
+  # Calculate and plot for the rest of the series
+    for (j in 3:k) {
+      ROC_OBJ <- roc_(df, response = df_names[1], 
+                      predictor = df_names[j], 
+                      quiet = T, plot = F,
+                      smooth = smoothROC)
+      aurocs[j-1,1] <- ROC_OBJ$auc
+      # Plot the ROC?
+      if (showROC) lines(ROC_OBJ$specificities, 
+                         ROC_OBJ$sensitivities, 
+                         col = j-2, lwd = 2)
+    }
+  }
+  
+  if (showROC){
+    if (k > 5) {
[…]

### 2025-05-20 R/compute_ROC_results.R

@@ -1,0 +1,143 @@
+# ROC curve sample code with artificial data
+
+sd <- seq(from = 1, to = 0.1, by = -0.1) # Std Dev's of Errors
+k <- length(sd) # No. of Series
+n <- 100 # No. of Observations
+
+# Generate target and k series of length n
+set.seed(20250519)
+mytarget <- matrix(rnorm(n), nrow = n)
+mypreds  <- matrix(rnorm(n*k, mean = 0, sd = sd), 
+                   nrow = n, byrow = T)
+# Predictors = Target + noise
+mypreds <- matrix(mytarget, ncol = k, nrow = n) + mypreds
+
+plot(mytarget,mypreds[,1], main = "Demo Data",
+     xlab = "Target", ylab = "Predictors") 
+for (j in 2:10) points(mytarget, mypreds[,j], col = j)
+
+Mydata <- data.frame(cbind(mytarget,mypreds))
+names(Mydata) <- c("Target", paste0("P_",1:10))
+# convert Target >0 to 1, rest is 0
+Mydata$Target <- ifelse(Mydata$Target > 0, 1, 0 )
+
+
+
+ROCplots(Mydata, smoothROC = T)
+
+########################################################################################################################
+# Alternative BVAR(4) model specification
+
+# Try a BVAR(4)
+V0_BVAR <- diag(rep(1,n))  # prior for Sigma_a matrix - K x K
+
+nlags <- 24
+lambda_BVAR <- 10  # scaling factor for precision matrix
+p_BVAR <- 4
+C_BVAR <- lambda_BVAR*diag(rep(1,k*p_BVAR+1))
+BVAR4_obj <- BVAR(x_mat_wc,
+                  p = p_BVAR, 
+                  C = C_BVAR,
+                  V0 = V0_BVAR)
+
+BVAR4_MA <- VARpsi(Phi = BVAR4_obj$Phi, lag = nlags)
+IRF_BIP_BVAR4 <- as.data.frame(t(matrix(BVAR4_MA$psi[1,], ncol = 1+nlags)))
+names(IRF_BIP_BVAR4) <- labels(x_mat)[[2]]   # labels source of shocks
+rownames(IRF_BIP_BVAR4) <- seq.int(from = 0, to = nlags) # labels lags
+
+MyColours <- c("blue", "darkgreen", "red3", "gray80", "lightgreen", "orange")
+# plot the results
+for (j in names(IRF_BIP_VAR1)) {
[…]

### 2025-05-20 R/compute_ROC_results.R
Update compute_ROC_results.R

@@ -138,6 +138,71 @@ for (h in 0:max(h_vec))
 }
 colnames(forward_shifted_HP_GDP_mat)<-paste("shift",h_vec,sep="")
 
+#-------------------------
+# Apply ROC
+ROCplots <- function(df, showROC = T, smoothROC = FALSE, showLegend = TRUE) {
+  df_names <- names(df) # series names
+  k <- length(df_names) # 1 + number of predictors to compare
+  print(paste("Target variable is", df_names[1]))
+  # Create table to hold AUCs
+  aurocs <- as.data.frame(matrix(NA, nrow = k-1))
+  row.names(aurocs) <- df_names[-1]
+  names(aurocs) <- "AUC"
+  
+  # make the first plot and store the first AUC
+  ROC_OBJ <- roc_(df, response = df_names[1], 
+                  predictor = df_names[2],
+                  quiet = T, plot = showROC,
+                  smooth = smoothROC)
+  aurocs[1,1] <- ROC_OBJ$auc
+  
+  if (k>2) {
+    # Calculate and plot for the rest of the series
+    for (j in 3:k) {
+      ROC_OBJ <- roc_(df, response = df_names[1], 
+                      predictor = df_names[j], 
+                      quiet = T, plot = F,
+                      smooth = smoothROC)
+      aurocs[j-1,1] <- ROC_OBJ$auc
+      # Plot the ROC?
+      if (showROC) lines(ROC_OBJ$specificities, 
+                         ROC_OBJ$sensitivities, 
+                         col = j-2, lwd = 2)
+    }
+  }
+  
+  if (showROC){
+    if (k > 5) {
+      roc_cex = 0.75
+      roc_ncol = 2
+    } else {
+      roc_cex = 1
+      roc_ncol = 1
+    }
+    if (showLegend) legend("bottomright", 
+                           legend = df_names[-1], 
+                           col = 1:(k-1), 
+                           lwd = 3,
+                           cex = roc_cex,
+                           ncol = roc_ncol)
+  }
+  return(aurocs)
+}
+
+ROCplots(Mydata, smoothROC = T)
+
+# Select shift and h
+shift<-4
+h<-6
+# Select target
+target<-as.integer(forward_shifted_HP_GDP_mat[,shift+1]>0)
+
[…]

### 2025-05-20 Sweave/BIP_predictor.Rnw
up

@@
 source(paste(getwd(),"/R/M_SSA_paper_functions.r",sep=""))
+# ROC curve: plots, AUC
+source(paste(getwd(),"/R/ROCplots.r",sep=""))
 

### 2025-05-20 R/ROCplots.R

@@ -23,11 +23,6 @@
 # - an arbitrary number of additional numeric series in the remaining columns
 # All series should be the same length and contain no NAs
 
-inst_pack<-rownames(installed.packages())
-if (!"pROC"%in%inst_pack)
-install.packages("pROC")
-
-library(pROC)
 
 ROCplots <- function(df, showROC = T, smoothROC = FALSE, showLegend = TRUE) {
   df_names <- names(df) # series names
@@ -56,7 +51,7 @@ ROCplots <- function(df, showROC = T, smoothROC = FALSE, showLegend = TRUE) {
       # Plot the ROC?
       if (showROC) lines(ROC_OBJ$specificities, 
                          ROC_OBJ$sensitivities, 
-                         col = j-2, lwd = 2)
+                         col = j-1, lwd = 2)
     }
   }
   
@@ -77,3 +72,7 @@ ROCplots <- function(df, showROC = T, smoothROC = FALSE, showLegend = TRUE) {
   }
   return(aurocs)
 }
+
+
+
+

### 2025-05-20 R/compute_ROC_results.R
up

@@ -80,6 +80,7 @@ for (j in names(IRF_BIP_VAR1)) {
 # 1.1. Compute direct forecasts for h in h_vec
 
 select_direct_indicator<-select_vec_multi
+select_direct_indicator<-c("ifo_c","ESI")
 # Select data matrix
 data_roc<-x_mat_wc
 # Note: too complex designs (too many indicators) lead to overfitting and thus worse out-of-sample performances
@@ -115,7 +116,7 @@ for (h in 0:max(h_vec))
   colnames(hp_mat)<-select_vec_multi
   lm_obj<-lm(forward_shifted_GDP~hp_mat[,select_direct_indicator])
   # Compute the predictor: one can rely on the generic R-function predict or compute the predictor manually
-  direct_hp_forecast_mat<-cbind(direct_hp_forecast_mat,lm_obj$coef[1]+t(hp_c_array_wc[,,1])%*%lm_obj$coef[2:(ncol(t(hp_c_array_wc[,,1]))+1)])
+  direct_hp_forecast_mat<-cbind(direct_hp_forecast_mat,lm_obj$coef[1]+hp_mat[,select_direct_indicator]%*%lm_obj$coef[2:(ncol(hp_mat[,select_direct_indicator])+1)])
 }
 colnames(direct_hp_forecast_mat)<-colnames(forward_shifted_GDP_mat)<-paste("h=",h_vec,sep="")
 
@@ -140,69 +141,29 @@ colnames(forward_shifted_HP_GDP_mat)<-paste("shift",h_vec,sep="")
 
 #-------------------------
 # Apply ROC
-ROCplots <- function(df, showROC = T, smoothROC = FALSE, showLegend = TRUE) {
-  df_names <- names(df) # series names
-  k <- length(df_names) # 1 + number of predictors to compare
-  print(paste("Target variable is", df_names[1]))
-  # Create table to hold AUCs
-  aurocs <- as.data.frame(matrix(NA, nrow = k-1))
-  row.names(aurocs) <- df_names[-1]
-  names(aurocs) <- "AUC"
-  
-  # make the first plot and store the first AUC
-  ROC_OBJ <- roc_(df, response = df_names[1], 
-                  predictor = df_names[2],
-                  quiet = T, plot = showROC,
-                  smooth = smoothROC)
-  aurocs[1,1] <- ROC_OBJ$auc
-  
-  if (k>2) {
-    # Calculate and plot for the rest of the series
[…]

### 2025-05-20 R/compute_ROC_results.R
Update compute_ROC_results.R

@@ -116,7 +116,7 @@ for (h in 0:max(h_vec))
   colnames(hp_mat)<-select_vec_multi
   lm_obj<-lm(forward_shifted_GDP~hp_mat[,select_direct_indicator])
   # Compute the predictor: one can rely on the generic R-function predict or compute the predictor manually
-  direct_hp_forecast_mat<-cbind(direct_hp_forecast_mat,lm_obj$coef[1]+hp_mat[,select_direct_indicator]%*%lm_obj$coef[2:(ncol(hp_mat[,select_direct_indicator])+1)])
+  direct_hp_forecast_mat<-cbind(direct_hp_forecast_mat,lm_obj$coef[1]+hp_mat[,select_direct_indicator]%*%lm_obj$coef[2:(length(select_direct_indicator)+1)])
 }
 colnames(direct_hp_forecast_mat)<-colnames(forward_shifted_GDP_mat)<-paste("h=",h_vec,sep="")
 
@@ -126,6 +126,21 @@ final_mssa_array<-final_mssa_indicator_obj$mssa_array
 mssa<-final_mssa_array["BIP",,]
 final_mmse_array<-final_mssa_indicator_obj$mmse_array
 mmse<-final_mmse_array["BIP",,]
+
+mssa_mat<-mmse_mat<-NULL
+for (h in 0:max(h_vec))
+{
+  # Shift GDP forward by publication lag+forecast horizon
+  forward_shifted_GDP<-c(data_roc[(1+lag_vec[1]+h):nrow(data_roc),"BIP"],rep(NA,h+lag_vec[1]))
+  forward_shifted_GDP_mat<-cbind(forward_shifted_GDP_mat,forward_shifted_GDP)
+  lm_obj<-lm(forward_shifted_GDP~t(final_mssa_array[select_direct_indicator,,h+1]))
+# Compute the predictor: one can rely on the generic R-function predict or compute the predictor manually
+  mssa_mat<-cbind(mssa_mat,lm_obj$coef[1]+t(final_mssa_array[select_direct_indicator,,h+1])%*%lm_obj$coef[2:(length(select_direct_indicator)+1)])
+  lm_obj<-lm(forward_shifted_GDP~t(final_mmse_array[select_direct_indicator,,h+1]))
[…]

### 2025-05-20 R/compute_ROC_results.R
up

@@ -157,8 +157,8 @@ colnames(forward_shifted_HP_GDP_mat)<-paste("shift",h_vec,sep="")
 #-------------------------
 # Apply ROC
 
-shift_vec<-0:5
-hh_vec<-0:6
+shift_vec<-3:5
+hh_vec<-3:6
 AUC_array<-array(dim=c(length(shift_vec),length(hh_vec),4))
 dimnames(AUC_array)<-list(paste("shift=",shift_vec,sep=""),paste("h=",hh_vec,sep=""),c("Direct forecast","Direct HP forecast","M-MSE","M-SSA"))
 par(mfrow=c(length(shift_vec),length(hh_vec)))
@@ -171,14 +171,17 @@ for (i in 1:length(shift_vec))
     
 # Select target
     target<-as.integer(forward_shifted_HP_GDP_mat[,shift+1]>0)
-    target<-as.integer(forward_shifted_GDP_mat[,shift+1]>0)
+#    target<-as.integer(forward_shifted_GDP_mat[,shift+1]>0)
     
     ROC_data<-cbind(target,direct_forecast_mat[,h+1],direct_hp_forecast_mat[,h+1],mmse_mat[,h+1],mssa_mat[,h+1])
     rownames(ROC_data)<-rownames(data_roc)
     colnames(ROC_data)<-c("Target","Direct forecast","Direct HP forecast","M-MSE","M-SSA")
     ROC_data<-as.data.frame(na.exclude(ROC_data))
-    AUC<-ROCplots(ROC_data, smoothROC = T)
-    AUC_array[i,j,]<-unlist(AUC)
+    showLegend<-ifelse(i==1&j==1,T,F)
+    ROCplots(ROC_data, smoothROC = T,showLegend=showLegend)
+#    AUC_array[i,j,]<-unlist(AUC)
+#    AUC<-ROCplots(ROC_data, smoothROC = T)
+#    AUC_array[i,j,]<-unlist(AUC)
   }
 }
 

### 2025-05-20 R/compute_ROC_results.R
Update compute_ROC_results.R

@@ -79,8 +79,8 @@ for (j in names(IRF_BIP_VAR1)) {
 
 # 1.1. Compute direct forecasts for h in h_vec
 
-select_direct_indicator<-select_vec_multi
 select_direct_indicator<-c("ifo_c","ESI")
+select_direct_indicator<-select_vec_multi
 # Select data matrix
 data_roc<-x_mat_wc
 # Note: too complex designs (too many indicators) lead to overfitting and thus worse out-of-sample performances
@@ -123,9 +123,7 @@ colnames(direct_hp_forecast_mat)<-colnames(forward_shifted_GDP_mat)<-paste("h=",
 #--------------
 # 1.3. M-SSA and M-MSE
 final_mssa_array<-final_mssa_indicator_obj$mssa_array
-mssa<-final_mssa_array["BIP",,]
 final_mmse_array<-final_mssa_indicator_obj$mmse_array
-mmse<-final_mmse_array["BIP",,]
 
 mssa_mat<-mmse_mat<-NULL
 for (h in 0:max(h_vec))
@@ -155,33 +153,41 @@ for (h in 0:max(h_vec))
 colnames(forward_shifted_HP_GDP_mat)<-paste("shift",h_vec,sep="")
 
 #-------------------------
-# Apply ROC
+# Apply ROC: generate curves and compute AUCs
 
-shift_vec<-3:5
-hh_vec<-3:6
+shift_vec<-0:5
+hh_vec<-0:6
 AUC_array<-array(dim=c(length(shift_vec),length(hh_vec),4))
 dimnames(AUC_array)<-list(paste("shift=",shift_vec,sep=""),paste("h=",hh_vec,sep=""),c("Direct forecast","Direct HP forecast","M-MSE","M-SSA"))
-par(mfrow=c(length(shift_vec),length(hh_vec)))
-for (i in 1:length(shift_vec))
+par(mfrow=c(length(which(shift_vec>2)),length(which(hh_vec>2))))
+showl<-T
+for (i in 1:length(shift_vec))#i<-1
 { 
-  for (j in 1:length(hh_vec))
+  for (j in 1:length(hh_vec))#j<-1
   {
     shift<-shift_vec[i]
     h<-hh_vec[j]
     
 # Select target
     target<-as.integer(forward_shifted_HP_GDP_mat[,shift+1]>0)
[…]

### 2025-05-21 R/compute_ROC_results.R
up

@@ -1,5 +1,99 @@
 # ROC curve sample code with artificial data
 
+
+ROCplots <- function(df, showROC = TRUE, main = NULL, lbls = "roc",
+                     smoothROC = FALSE, colours = NULL, lwd = 2,
+                     showLegend = TRUE, lg_cex = 1, lg_ncol = 1) {
+  df_names <- names(df) # series names
+  k <- length(df_names) # 1 + number of predictors to compare
+  print(paste("Target variable is", df_names[1]))
+  # Create table to hold AUCs
+  aurocs <- as.data.frame(matrix(NA, nrow = k-1))
+  row.names(aurocs) <- df_names[-1]
+  names(aurocs) <- "AUC"
+  
+  # make the first plot and store the first AUC
+  # ROC_OBJ <- roc_(df, response = df_names[1],
+  #                 predictor = df_names[2],
+  #                 quiet = T, plot = showROC,
+  #                 smooth = smoothROC)
+  # aurocs[1,1] <- ROC_OBJ$auc
+  
+  ROC_OBJ <- roc_(df, response = df_names[1],
+                  predictor = df_names[2],
+                  quiet = T, plot = F,
+                  smooth = smoothROC)
+  # Store the AUC
+  aurocs[1,1] <- ROC_OBJ$auc
+  
+  if (showROC) {
+    # preliminaries for all plots
+    xlim = c(1,0)
+    a_line = 1
+    b_line = -1
+    xlbl = "Specificity"
+    ylbl = "Sensitivity"
+    # User-defined colours
+    if (is.null(colours)) colours = 1:(k-1)
+    
+    # reset axes and their labels if needed
+    if (lbls == 'Hit') {
+      xlbl <- 'False Alarm Rate'
+      ylbl <- "Hit Rate"
+      xlim = c(0,1)
+      a_line = 0
+      b_line = 1
+    } else if (lbls == ""){
+      xlbl <- ""
+      ylbl <- ""
+    }
+    
+    # plot the 1st ROC
+    # plotting specificity or False Alarms?
+    specifty <- ROC_OBJ$specificities
+    if(lbls == "Hit") specifty <- 1 - specifty
+    
+    plot(specifty, 
+         ROC_OBJ$sensitivities,
+         xlim = xlim, ylim = c(0,1), 
+         type = "l", lwd = lwd, 
+         col = colours[1],
[…]

### 2025-05-21 Latex/BIP_predictor.tex
up

@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Tue May 13 15:04:25 2025
+% Wed May 21 09:45:42 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Tue May 13 15:04:25 2025
+% Wed May 21 09:45:42 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Tue May 13 15:04:27 2025
+% Wed May 21 09:45:44 2025
 \begin{table}[ht]
@@
 The two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Tue May 13 15:04:28 2025
+% Wed May 21 09:45:44 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Tue May 13 15:04:31 2025
+% Wed May 21 09:45:47 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Tue May 13 15:04:31 2025
+% Wed May 21 09:45:47 2025
 \begin{table}[ht]
@@
 
+\section{Hit Rate vs. False Alarm Rate}
+
+%As shown in Wildi (2025), the SSA criterion's objective function—the target correlation $\rho(y,z,h)$—can be linked to the sign accuracy of the predictor $SA(y_t):=\text{P}\left(\text{sign}(z_{t+h}) = \text{sign}(y_t)\right)$, i.e., the probability that the signs of the predictor $y_t$ and the target $z_{t+h}$ agree:
+%\begin{eqnarray}\label{sig_acc}
+%SA(y_t)=0.5+\frac{\arcsin(\rho(y,z,h))}{\pi}.
+%\end{eqnarray}
+%Due to the strict monotonicity of the arcsin function, maximizing the target correlation also maximizes the sign accuracy of the predictor.\footnote{Equation \ref{sig_acc} holds exactly for Gaussian processes.
+Due to the central limit theorem, the relationship holds approximately for low-pass filtered non-Gaussian series, see Wildi (2025).}
+\begin{itemize}
+\item Alternative to MSE, target correlation, holding-time: hit-rate, false alarm. 
+\item We can link objective function (target correlation) with sign accuracy, see Wildi (2025). BUT: risk that reviewer asks for comparison with logit or whatever sign-forecast tool.
+\item Context ROC-curve
+\end{itemize}
+For the ROC curves, we consider predictors of GDP based on both original (unfiltered) and filtered indicators. For the direct forecast, we use the regression
+\[
+GDP_{t+shift}=\beta_0^{shift}+\sum_{i=1}^5\beta_i^{shift}x_{ti}+\epsilon_{t}^{shift,direct}=\hat{GDP}_{t+shift}^{direct}+\epsilon_{t}^{shift,direct},
+\]
+where $x_{ti}$ designates the original indicators. 
+For the filtered indicators, we consider similar regression equations:
+\[GDP_{t+shift}=\beta_0^{shift,h,filt}+\sum_{i=1}^5\beta_i^{shift,h,filt}y_{ti}^{h,filt}+\epsilon_{t}^{h,shift,filt}=\hat{GDP}_{t+shift}^{filt}+\epsilon_{t}^{shift,filt},\]
+where $y_{ti}^{h,filt}$ represents the $i$-th filtered indicator, based on either HP-C, M-MSE, or M-SSA.
+For simplicity of exposition, we present results based on the full sample, excluding the pandemic period.
+Fig.
+
+\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=5in]{./Figures/ROC_GDP_shift_1_4.pdf}\caption{Hit rate vs.
+false alarm rate of the predictors at small (one quarter ahead:left panel) and yearly forecast horizons (right panel).\label{ROC_GDP_shift_1_4}}\end{center}\end{figure}
+% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Wed May 21 09:45:50 2025
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrr}
+  \hline
+ & Direct forecast & Direct HP forecast & M-MSE & M-SSA \\ 
+  \hline
+Forecast horizon 1 & 0.752 & 0.715 & 0.746 & 0.747 \\ 
+  Forecast horizon 2 & 0.697 & 0.663 & 0.729 & 0.720 \\ 
+  Forecast horizon 3 & 0.554 & 0.622 & 0.684 & 0.709 \\ 
+  Forecast horizon 4 & 0.486 & 0.624 & 0.680 & 0.708 \\ 
+  Forecast horizon 5 & 0.593 & 0.636 & 0.643 & 0.701 \\ 
+   \hline
+\end{tabular}
+\caption{AUCs of different predictors against GDP at various forecast horizons..} 
+\label{p_val_wc5}
+\end{table}
+
 \section{Summary and Conclusion}
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Tue May 13 15:04:31 2025
+% Wed May 21 09:45:50 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Tue May 13 15:04:31 2025
+% Wed May 21 09:45:50 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Tue May 13 15:04:32 2025
+% Wed May 21 09:45:50 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Tue May 13 15:04:32 2025
+% Wed May 21 09:45:50 2025
 \begin{table}[ht]

### 2025-05-21 Sweave/BIP_predictor.Rnw
up

@@
 
+# We can select type of VAR model
+VAR_type="BVAR"
+VAR_type="VAR"
+if (VAR_type=="BVAR")
+{
+# Simon's settings for BVAR(4)  
+  lambda_BVAR <- 10 
+  p_full<-4
+  q_full<-0
+}
+if (VAR_type=="VAR")
+{
+# Marc's settings for VAR  
+  p_full<-p
+  q_full<-q
+}
+
 # Run the M-SSA wrapper, see tutorial 7.2
@@
 {
-  final_mssa_indicator_obj<-compute_mssa_BIP_predictors_func(x_mat_wc,lambda_HP,L,date_to_fit_all,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi)
+  final_mssa_indicator_obj<-compute_mssa_BIP_predictors_func(x_mat_wc,lambda_HP,L,date_to_fit_all,p_full,q_full,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi,VAR_type,lambda_BVAR)
   save(final_mssa_indicator_obj,file=paste(path.result,"/final_mssa_indicator_obj",sep=""))
@@
 legend("bottomleft", legend = colnames(mplot), col = colo, lty = 1, lwd = 1, bty = "n")
-
 box()
-dev.off()
+invisible(dev.off())
 
@@
 
-dev.off()
+invisible(dev.off())
 
@@
 
+\section{Hit Rate vs. False Alarm Rate}
+
+%As shown in Wildi (2025), the SSA criterion's objective function—the target correlation $\rho(y,z,h)$—can be linked to the sign accuracy of the predictor $SA(y_t):=\text{P}\left(\text{sign}(z_{t+h}) = \text{sign}(y_t)\right)$, i.e., the probability that the signs of the predictor $y_t$ and the target $z_{t+h}$ agree:
+%\begin{eqnarray}\label{sig_acc}
+%SA(y_t)=0.5+\frac{\arcsin(\rho(y,z,h))}{\pi}.
+%\end{eqnarray}
+%Due to the strict monotonicity of the arcsin function, maximizing the target correlation also maximizes the sign accuracy of the predictor.\footnote{Equation \ref{sig_acc} holds exactly for Gaussian processes.
+Due to the central limit theorem, the relationship holds approximately for low-pass filtered non-Gaussian series, see Wildi (2025).}
+\begin{itemize}
+\item Alternative to MSE, target correlation, holding-time: hit-rate, false alarm. 
+\item We can link objective function (target correlation) with sign accuracy, see Wildi (2025). BUT: risk that reviewer asks for comparison with logit or whatever sign-forecast tool.
+\item Context ROC-curve
+\end{itemize}
+For the ROC curves, we consider predictors of GDP based on both original (unfiltered) and filtered indicators. For the direct forecast, we use the regression
+\[
+GDP_{t+shift}=\beta_0^{shift}+\sum_{i=1}^5\beta_i^{shift}x_{ti}+\epsilon_{t}^{shift,direct}=\hat{GDP}_{t+shift}^{direct}+\epsilon_{t}^{shift,direct},
+\]
+where $x_{ti}$ designates the original indicators. 
+For the filtered indicators, we consider similar regression equations:
+\[GDP_{t+shift}=\beta_0^{shift,h,filt}+\sum_{i=1}^5\beta_i^{shift,h,filt}y_{ti}^{h,filt}+\epsilon_{t}^{h,shift,filt}=\hat{GDP}_{t+shift}^{filt}+\epsilon_{t}^{shift,filt},\]
+where $y_{ti}^{h,filt}$ represents the $i$-th filtered indicator, based on either HP-C, M-MSE, or M-SSA.
+For simplicity of exposition, we present results based on the full sample, excluding the pandemic period.
+Fig.
+<<label=init,results=hide>>=
+
+#########################################################################################################
+# Apply ROC to direct forecast, direct HP forecast, M-MSE and M-SSA
+# Data: 
+#   -In-sample vs. out-of-sample
+#   -Same sample (account for different NAs of various designs)
+# Target: 
+#   -GDP vs. HP-GDP
+#   -forward-shift against h
+
+# 1. Full sample without pandemic; all explanatories
+
+# 1.1. Compute direct forecasts for h in h_vec
+# Select all indicators
+select_direct_indicator<-select_vec_multi
+# Select data matrix: full sample without pandemic
+data_roc<-x_mat_wc
+direct_forecast_mat<-NULL
+for (h in 0:max(h_vec))#h<-0
+{
+  # Shift GDP forward by publication lag+forecast horizon
+  forward_shifted_GDP<-c(data_roc[(1+lag_vec[1]+h):nrow(data_roc),"BIP"],rep(NA,h+lag_vec[1]))
+  # Regress selected indicators on forward-shifted GDP: start at t=L (same sample as HP-C)
+  lm_obj<-lm(forward_shifted_GDP[L:length(forward_shifted_GDP)]~data_roc[L:nrow(data_roc),select_direct_indicator])
+  summary(lm_obj)
+  # Compute the predictor: one can rely on the generic R-function predict or compute the predictor manually
+  direct_forecast<-lm_obj$coef[1]+data_roc[,select_direct_indicator]%*%lm_obj$coef[2:(length(select_direct_indicator)+1)]
+  # Note that this is a full-sample predictor (no out-of-sample span)
+
+  # We can now plot target and direct forecast: for h>2 the predictor comes close to a flat line centered at zero
+  direct_forecast_mat<-cbind(direct_forecast_mat,direct_forecast)
+}
+colnames(direct_forecast_mat)<-paste("h=",h_vec,sep="")
+
+#-----------
+# 1.2. Direct HP forecast
+direct_hp_forecast_mat<-forward_shifted_GDP_mat<-NULL
+for (h in 0:max(h_vec))
+{
+  # Shift GDP forward by publication lag+forecast horizon
+  forward_shifted_GDP<-c(data_roc[(1+lag_vec[1]+h):nrow(data_roc),"BIP"],rep(NA,h+lag_vec[1]))
+  forward_shifted_GDP_mat<-cbind(forward_shifted_GDP_mat,forward_shifted_GDP)
+  hp_mat<-NULL
+  for (i in 1:ncol(data_roc))
+   hp_mat<-cbind(hp_mat,filter(data_roc[,i],hp_c,side=1))
+  colnames(hp_mat)<-select_vec_multi
+  lm_obj<-lm(forward_shifted_GDP~hp_mat[,select_direct_indicator])
+  # Compute the predictor: one can rely on the generic R-function predict or compute the predictor manually
+  direct_hp_forecast_mat<-cbind(direct_hp_forecast_mat,lm_obj$coef[1]+hp_mat[,select_direct_indicator]%*%lm_obj$coef[2:(length(select_direct_indicator)+1)])
+}
+colnames(direct_hp_forecast_mat)<-colnames(forward_shifted_GDP_mat)<-paste("h=",h_vec,sep="")
+
+#--------------
+# 1.3. M-SSA and M-MSE
+final_mssa_array<-final_mssa_indicator_obj$mssa_array
+final_mmse_array<-final_mssa_indicator_obj$mmse_array
+
+mssa_mat<-mmse_mat<-NULL
+for (h in 0:max(h_vec))
+{
+  # Shift GDP forward by publication lag+forecast horizon
+  forward_shifted_GDP<-c(data_roc[(1+lag_vec[1]+h):nrow(data_roc),"BIP"],rep(NA,h+lag_vec[1]))
+  forward_shifted_GDP_mat<-cbind(forward_shifted_GDP_mat,forward_shifted_GDP)
+  lm_obj<-lm(forward_shifted_GDP~t(final_mssa_array[select_direct_indicator,,h+1]))
+# Compute the predictor: one can rely on the generic R-function predict or compute the predictor manually
+  mssa_mat<-cbind(mssa_mat,lm_obj$coef[1]+t(final_mssa_array[select_direct_indicator,,h+1])%*%lm_obj$coef[2:(length(select_direct_indicator)+1)])
+  lm_obj<-lm(forward_shifted_GDP~t(final_mmse_array[select_direct_indicator,,h+1]))
+  mmse_mat<-cbind(mmse_mat,lm_obj$coef[1]+t(final_mmse_array[select_direct_indicator,,h+1])%*%lm_obj$coef[2:(length(select_direct_indicator)+1)])
+}
+colnames(mssa_mat)<-colnames(mmse_mat)<-paste("h=",h_vec,sep="")
+
+#-------------
+# 1.4. Targets: forward-shifted GDP and HP-GDP
+forward_shifted_GDP_mat<-forward_shifted_GDP_mat
+forward_shifted_HP_GDP_mat<-NULL
+for (h in 0:max(h_vec))
+{
+  # Shift GDP forward by publication lag+forecast horizon
+  forward_shifted_GDP<-c(data_roc[(1+lag_vec[1]+h):nrow(data_roc),"BIP"],rep(NA,h+lag_vec[1]))
+  forward_shifted_HP_GDP_mat<-cbind(forward_shifted_HP_GDP_mat,filter(forward_shifted_GDP,hp_two,side=2))
+}
+colnames(forward_shifted_HP_GDP_mat)<-paste("shift",h_vec,sep="")
+
+#-------------------------
+# Apply ROC: generate curves and compute AUCs
+
+# Combination of forward-shift (of target) and forecast horizon (underlying optimization)
+shift_vec<-1:5
+hh_vec<-1:6
+AUC_array_GDP<-AUC_array_HP_GDP<-array(dim=c(length(shift_vec),length(hh_vec),4))
+dimnames(AUC_array_GDP)<-dimnames(AUC_array_HP_GDP)<-list(paste("shift=",shift_vec,sep=""),paste("h=",hh_vec,sep=""),c("Direct forecast","Direct HP forecast","M-MSE","M-SSA"))
+# Select shifts>k (otherwise too many plots in single window)
+k<-2
+# Show plot
+showl<-T
+# Labels: hit rate vs. false alarm rate
+lbls = "Hit"
+
+#----------------------
+# Target GDP
+select_target<-"GDP"
+
+#file = "./Figures/ROC_GDP_many.pdf"
+#pdf(file = paste(path.out,file,sep=""), paper = "special", width = 7, height = 7)
+par(mfrow=c(length(which(shift_vec>k)),length(which(hh_vec>k))))
+
+for (i in 1:length(shift_vec))#i<-1
+{ 
+  for (j in 1:length(hh_vec))#j<-1
+  {
+    shift<-shift_vec[i]
+    h<-hh_vec[j]
+
+# Select target: shifted GDP or HP-GDP
+    if (select_target=="GDP")
+    {
+      target<-as.integer(forward_shifted_GDP_mat[,shift+1]>0)
+    } else
+
+    {
+      target<-as.integer(forward_shifted_HP_GDP_mat[,shift+1]>0)
+    }
+# Set up data matrix for ROC calculation    
+    ROC_data<-cbind(target,direct_forecast_mat[,h+1],direct_hp_forecast_mat[,h+1],mmse_mat[,h+1],mssa_mat[,h+1])
+    rownames(ROC_data)<-rownames(data_roc)
+    colnames(ROC_data)<-c("Target","Direct forecast","Direct HP forecast","M-MSE","M-SSA")
+    ROC_data<-as.data.frame(na.exclude(ROC_data))
+# Select plots that will be shown    
+    showROC = ifelse(i>k&j>k,T,F)
+    if (showROC&showl)
+    {
+      showLegend<-T
+      showl<-F
+    } else
+    {
+      showLegend<-F
+    }
+# Apply smoothing    
+    smoothROC<-T
+
+    AUC<-ROCplots(ROC_data, showROC , main = paste("shift=",shift,", h=",h,sep=""), lbls = lbls,
+                  smoothROC , colours =NULL, lwd = 2,
+                  showLegend , lg_cex = 1, lg_ncol = 1)
+    AUC_array_GDP[i,j,]<-unlist(AUC)
+  }
+}
+#invisible(dev.off())
+
+#--------------------------
+# Select  HP-GDP
+select_target<-"HP-GDP"
+
+#file = "./Figures/ROC_HP_GDP_many.pdf"
+#pdf(file = paste(path.out,file,sep=""), paper = "special", width = 7, height = 7)
+par(mfrow=c(length(which(shift_vec>k)),length(which(hh_vec>k))))
+
+for (i in 1:length(shift_vec))#i<-1
+{ 
+  for (j in 1:length(hh_vec))#j<-1
+  {
+    shift<-shift_vec[i]
+    h<-hh_vec[j]
+
+# Select target: shifted GDP or HP-GDP
+    if (select_target=="GDP")
+    {
+      target<-as.integer(forward_shifted_GDP_mat[,shift+1]>0)
+    } else
+
+    {
+      target<-as.integer(forward_shifted_HP_GDP_mat[,shift+1]>0)
+    }
+# Set up data matrix for ROC calculation    
+    ROC_data<-cbind(target,direct_forecast_mat[,h+1],direct_hp_forecast_mat[,h+1],mmse_mat[,h+1],mssa_mat[,h+1])
+    rownames(ROC_data)<-rownames(data_roc)
+    colnames(ROC_data)<-c("Target","Direct forecast","Direct HP forecast","M-MSE","M-SSA")
+    ROC_data<-as.data.frame(na.exclude(ROC_data))
+# Select plots that will be shown    
+    showROC = ifelse(i>k&j>k,T,F)
+    if (showROC&showl)
+    {
+      showLegend<-T
+      showl<-F
+    } else
+    {
+      showLegend<-F
+    }
+# Apply smoothing    
+    smoothROC<-T
+
+    AUC<-ROCplots(ROC_data, showROC , main = paste("shift=",shift,", h=",h,sep=""), lbls = lbls,
+                  smoothROC , colours =NULL, lwd = 2,
+                  showLegend , lg_cex = 1, lg_ncol = 1)
+    AUC_array_HP_GDP[i,j,]<-unlist(AUC)
+  }
+}
+#invisible(dev.off())
+
+#-----------------------------
+# Shifts 1 and 4 only
+
+select_target<-"GDP"
+
+file = "./Figures/ROC_GDP_shift_1_4.pdf"
+pdf(file = paste(path.out,file,sep=""), paper = "special", width = 7, height = 7)
+par(mfrow=c(1,2))
+
+shift<-1
+h<-1
+
+# Select target: shifted GDP or HP-GDP
+if (select_target=="GDP")
+{
+  target<-as.integer(forward_shifted_GDP_mat[,shift+1]>0)
+} else
+{
+  target<-as.integer(forward_shifted_HP_GDP_mat[,shift+1]>0)
+}
+# Set up data matrix for ROC calculation    
+ROC_data<-cbind(target,direct_forecast_mat[,h+1],direct_hp_forecast_mat[,h+1],mmse_mat[,h+1],mssa_mat[,h+1])
+rownames(ROC_data)<-rownames(data_roc)
+colnames(ROC_data)<-c("Target","Direct forecast","Direct HP forecast","M-MSE","M-SSA")
+ROC_data<-as.data.frame(na.exclude(ROC_data))
+# Select plots that will be shown    
+showLegend<-T
+AUC<-ROCplots(ROC_data, showROC , main = paste("shift=",shift,sep=""), lbls = lbls,
+                  smoothROC , colours =NULL, lwd = 2,
+                  showLegend , lg_cex = 1, lg_ncol = 1)
+shift<-4
+h<-shift
+
+# Select target: shifted GDP or HP-GDP
+if (select_target=="GDP")
+{
+  target<-as.integer(forward_shifted_GDP_mat[,shift+1]>0)
+} else
+{
+  target<-as.integer(forward_shifted_HP_GDP_mat[,shift+1]>0)
+}
+# Set up data matrix for ROC calculation    
+ROC_data<-cbind(target,direct_forecast_mat[,h+1],direct_hp_forecast_mat[,h+1],mmse_mat[,h+1],mssa_mat[,h+1])
+rownames(ROC_data)<-rownames(data_roc)
+colnames(ROC_data)<-c("Target","Direct forecast","Direct HP forecast","M-MSE","M-SSA")
+ROC_data<-as.data.frame(na.exclude(ROC_data))
+# Select plots that will be shown    
+showLegend<-F
+AUC<-ROCplots(ROC_data, showROC , main = paste("shift=",shift,sep=""), lbls = lbls,
+                  smoothROC , colours =NULL, lwd = 2,
+                  showLegend , lg_cex = 1, lg_ncol = 1)
+
+invisible(dev.off())
+
+# Shifts 1 and 4 only
+
+select_target<-"HP-GDP"
+
+file = "./Figures/ROC_HP_GDP_shift_1_4.pdf"
+pdf(file = paste(path.out,file,sep=""), paper = "special", width = 7, height = 7)
+par(mfrow=c(1,2))
+
+shift<-1
+h<-1
+
+# Select target: shifted GDP or HP-GDP
+if (select_target=="GDP")
+{
+  target<-as.integer(forward_shifted_GDP_mat[,shift+1]>0)
+} else
+{
+  target<-as.integer(forward_shifted_HP_GDP_mat[,shift+1]>0)
+}
+# Set up data matrix for ROC calculation    
+ROC_data<-cbind(target,direct_forecast_mat[,h+1],direct_hp_forecast_mat[,h+1],mmse_mat[,h+1],mssa_mat[,h+1])
+rownames(ROC_data)<-rownames(data_roc)
+colnames(ROC_data)<-c("Target","Direct forecast","Direct HP forecast","M-MSE","M-SSA")
+ROC_data<-as.data.frame(na.exclude(ROC_data))
+# Select plots that will be shown    
+showLegend<-T
+AUC<-ROCplots(ROC_data, showROC , main = paste("HP-GDP: shift=",shift,sep=""), lbls = lbls,
+                  smoothROC , colours =NULL, lwd = 2,
+                  showLegend , lg_cex = 1, lg_ncol = 1)
+shift<-4
+h<-shift
+
+# Select target: shifted GDP or HP-GDP
+if (select_target=="GDP")
+{
+  target<-as.integer(forward_shifted_GDP_mat[,shift+1]>0)
+} else
+{
+  target<-as.integer(forward_shifted_HP_GDP_mat[,shift+1]>0)
+}
+# Set up data matrix for ROC calculation    
+ROC_data<-cbind(target,direct_forecast_mat[,h+1],direct_hp_forecast_mat[,h+1],mmse_mat[,h+1],mssa_mat[,h+1])
+rownames(ROC_data)<-rownames(data_roc)
+colnames(ROC_data)<-c("Target","Direct forecast","Direct HP forecast","M-MSE","M-SSA")
+ROC_data<-as.data.frame(na.exclude(ROC_data))
+# Select plots that will be shown    
+showLegend<-F
+AUC<-ROCplots(ROC_data, showROC , main = paste("HP-GDP: shift=",shift,sep=""), lbls = lbls,
+                  smoothROC , colours =NULL, lwd = 2,
+                  showLegend , lg_cex = 1, lg_ncol = 1)
+
+invisible(dev.off())
+
+#----------------------------
+# LOOK at AUCs
+
+AUC_array_GDP["shift=1",,]
+AUC_array_GDP["shift=2",,]
+AUC_array_GDP["shift=3",,]
+AUC_array_GDP["shift=4",,]
+AUC_array_GDP["shift=5",,]
+
+# Compute table with diagonal (shift=h)
+auc_GDP_mat<-auc_HP_GDP_mat<-NULL
+for (i in 1:length(shift_vec))
+{
+  auc_GDP_mat<-rbind(auc_GDP_mat,AUC_array_GDP[i,i,])
+  auc_HP_GDP_mat<-rbind(auc_HP_GDP_mat,AUC_array_HP_GDP[i,i,])
+}
+rownames(auc_GDP_mat)<-rownames(auc_HP_GDP_mat)<-paste("Forecast horizon ",shift_vec,sep="")
+auc_GDP_mat
+auc_HP_GDP_mat
+
+@
+
+<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
+file = "./Figures/ROC_GDP_shift_1_4.pdf"
+cat("\\begin{figure}[H]")
+cat("\\begin{center}")
+cat("\\includegraphics[height=3in, width=5in]{", file, "}\n",sep = "")
+cat("\\caption{Hit rate vs. false alarm rate of the predictors at small (one quarter ahead:left panel) and yearly forecast horizons (right panel).", sep = "")
+cat("\\label{ROC_GDP_shift_1_4}}", sep = "")
+cat("\\end{center}")
+cat("\\end{figure}")
+@
+
+<<label=ats_mba_2,echo=FALSE,results=tex>>=
+
+mat1<-round(auc_GDP_mat,3)
+#latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
+#center = "centering", file = "", floating = FALSE)
+xtable(mat1, dec = 1,digits=rep(3,dim(mat1)[2]+1),
+paste("AUCs of different predictors against GDP at various forecast horizons.."),
+label=paste("p_val_wc5",sep=""),
+center = "centering", file = "", floating = FALSE)
+@
+
 \section{Summary and Conclusion}

### 2025-05-21 Latex/BIP_predictor.tex
up

@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 09:45:42 2025
+% Wed May 21 10:24:26 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 09:45:42 2025
+% Wed May 21 10:24:26 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 09:45:44 2025
+% Wed May 21 10:24:28 2025
 \begin{table}[ht]
@@
 The two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 09:45:44 2025
+% Wed May 21 10:24:28 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 09:45:47 2025
+% Wed May 21 10:24:32 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 09:45:47 2025
+% Wed May 21 10:24:33 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 09:45:47 2025
+% Wed May 21 10:24:33 2025
 \begin{table}[ht]
@@
 Due to the central limit theorem, the relationship holds approximately for low-pass filtered non-Gaussian series, see Wildi (2025).}
+\textbf{Remarks:}
 \begin{itemize}
@@
 For the filtered indicators, we consider similar regression equations:
-\[GDP_{t+shift}=\beta_0^{shift,h,filt}+\sum_{i=1}^5\beta_i^{shift,h,filt}y_{ti}^{h,filt}+\epsilon_{t}^{h,shift,filt}=\hat{GDP}_{t+shift}^{filt}+\epsilon_{t}^{shift,filt},\]
-where $y_{ti}^{h,filt}$ represents the $i$-th filtered indicator, based on either HP-C, M-MSE, or M-SSA.
+\[GDP_{t+shift}=\beta_0^{shift,h,filt}+\sum_{i=1}^5\beta_i^{shift,h,filt}y_{ti}^{h,filt}+\epsilon_{t}^{h,shift,filt}=\hat{GDP}_{t+shift}^{h,filt}+\epsilon_{t}^{h,shift,filt},\]
+where $y_{ti}^{h,filt}$ represents the $i$-th filtered indicator optimized at forecast horizon $h$, based on either HP-C, M-MSE, or M-SSA. \\
+\textbf{Remarks:}
+\begin{itemize}
+\item I'm computing all combinations of shift and h. 
+\item But we could emphasize the diagonal only (h=shift): simplifies notation/explanation.
+\item The table of AUCs (below) is based on diagonal only
+\end{itemize}
 For simplicity of exposition, we present results based on the full sample, excluding the pandemic period.
-Fig.
+Fig.\ref{ROC_GDP_shift_1_4} compares ROCs of the predictors for small and mid-term forecast horizons and Table \ref{auc} presents AUCs (area under the curve) for forecast horizons 1-5 quarters ahead.
 
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 09:45:50 2025
+% Wed May 21 10:24:36 2025
 \begin{table}[ht]
@@
 \end{tabular}
-\caption{AUCs of different predictors against GDP at various forecast horizons..} 
+\caption{AUCs of predictors against GDP at various forecast horizons.} 
 \label{p_val_wc5}
-\end{table}
+\end{table}At forecast horizons less than or equal to two quarters, the direct forecast challenges the filters.
+However, for horizons $h>2$, the filters outperform the direct approach: the univariate HP-C design is surpassed by the classic multivariate MSE filter, which is in turn dominated by the M-SSA.
+These results confirm that controlling the rate of sign changes of the predictor, while maximizing its target correlation through M-SSA can enhance predictor performance in terms of sign accuracy and false alarm rate.
 
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 09:45:50 2025
+% Wed May 21 10:24:36 2025
 \begin{table}[ht]

### 2025-05-21 archive/overleaf_latex.tex
up

\documentclass[11pt,a4paper]{article}

\usepackage{amsmath}
\usepackage[a4paper,includehead]{geometry}

\usepackage{graphicx} % Required for inserting images
\usepackage{tabularx}
\usepackage{natbib}
\usepackage[spaces,hyphens]{url}
\usepackage[hidelinks]{hyperref}
\usepackage{setspace}
\usepackage{booktabs}
\usepackage{hyphenat}
\usepackage{color}
%\usepackage{subfig}
% \usepackage[justification=left,hang]{caption}
\usepackage[justification=raggedright,hang]{caption}
\usepackage{subcaption}
	%\captionsetup[subfigure]{labelfont=normal,textfont=normalfont,singlelinecheck=off,justification=centering}
\usepackage[table]{xcolor} 
\usepackage{colortbl}
\usepackage[hang]{footmisc}
\usepackage{appendix}

\geometry{left=2.25cm,right=2.25cm,top=1.5cm,bottom=2cm}
%\hyphenation{EMU} \hyphenation{CEECs}
%\setlength{\parindent}{0pt} \setlength{\parskip}{10pt} %\setlength{\abovecaptionskip}{-5pt}
%\setlength{\belowcaptionskip}{-5pt} \setlength{\floatsep}{10pt}
\clubpenalty = 10000 
\widowpenalty = 10000
\displaywidowpenalty = 10000

% \makeatletter
% \setlength{\@fptop}{5pt} %setzt den abstand der floats vom oberen Rand auf Null- 5 entspricht der gleichen Höhe wie die wo der Text anfängt
% 		%\setlength\@fptop{0\p@ \@plus 1fil}
% \setlength\@fpsep{10pt plus 2pt minus 2pt} %Abstand zwischen den Floats
% \makeatother

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\setlength{\abovedisplayskip}{2cm}
\setlength{\abovedisplayshortskip}{2cm}
\setlength{\belowdisplayskip}{0cm}
\setlength{\belowdisplayshortskip}{2cm} %Abstand zwischen Text und Formel

\title{Smooth and Persistent Forecasts of German GDP}
\author{Katja Heinisch\thanks{Halle Institute for Economic Research (IWH).} \hspace{0.2cm}
Simon van Norden\thanks{HEC Montréal and CIREQ.} \hspace{0.2cm}
Marc Wildi\thanks{Zurich University of Applied Sciences (ZHAW).}}
\date{May 2025 \\
\vspace{2cm} \textit{Work in progress, not for circulation}}

\maketitle
%%%%%%%%%%%%%%%%%%%%%%%
\maketitle

\begin{abstract} %max.
This paper addresses the trade-off between informational efficiency and forecast smoothness in economic forecasting. Efficient forecasts are often volatile, limiting their practical usefulness. Building on the Smooth Sign Accuracy (SSA) framework, we propose a novel method that integrates Multivariate Smooth Sign Accuracy (M-SSA) to extract smoothed components from leading indicators. This approach enhances the signal-to-noise ratio and allows for systematic control over forecast volatility and timing. Applied to German GDP, our method yields smoothed forecasts that match or exceed the accuracy of traditional models, particularly over medium-term horizons. While smoother forecasts may lag slightly, this can be mitigated by adjusting the forecast horizon. The findings offer practical tools for forecasters and policymakers operating under adjustment constraints.

\end{abstract}

\bigskip \vspace{10pt}
\textbf{Keywords:} Forecast Smoothing, Smooth Sign Accuracy 
\newline
\textbf{JEL Classification:} C53, E37, E66
%C53 – Forecasting and Prediction Methods; Simulation Methods
%E37 – Forecasting and Simulation: Models and Applications
%E66 – General Outlook and Conditions: Forecasts and Simulations
\thispagestyle{empty}

\onehalfspacing
%\doublespacing
\renewcommand{\thepage}{\arabic{page}} \setcounter{page}{1}
\newpage
\section{Motivation}\label{sec:motivation}

\begin{quote}
     \quad As we parse the incoming information, we are focused on separating the signal from the noise as the outlook evolves.\\
    \textit{Speech by Federal Reserve Board Chair Jerome H. Powell, March 07, 2025, \\
    University of Chicago Booth School of Business}
\end{quote}

Economic forecasters face a dilemma: While efficiency suggests that their forecasts should incorporate all available information, the resulting forecasts are often volatile, which may reflect greater information content or contamination by excessive ``noise''. Further, forecast users (including public and private decision makers) may be constrained by how rapidly they can adjust to changing forecasts.\footnote{For example, private or public spending decisions may be constrained by annual budgets. Similarly, central banks increasingly attempt to smooth interest rate changes to allow them to signal the future direction of policy to markets, while private firms attempt to smooth changes in dividends.}

%%% Further recent literature
% \cite{lehmann2021predicting} show the performance of ifo busines cycles indicator for the nowcast (h=0) and forecast (h=1)

\cite{Wildi2024,Wildi2025} and \cite{McElroy2019,McElroy2020} propose novel methods that optimally smooth forecasts by controlling the expected frequency of sign changes. The resulting Smooth Sign Accuracy (SSA) framework encompasses efficient mean-square error (MSE) as a special case and produces increasingly smooth forecasts as changes in forecast sign are increasingly penalized.

% First, direct AR-forecast (benchmarks) are fairly noisy. In contrast,
% trend forecast are generally (substantially) smoother
% Second, zero-crossings are possibly indicative of relevant changes in
% the economy (trend growth cycle) and the rate of
% zero-crossings can be controlled by SSA.
% Third, trend forecasts and direct AR forecasts are similar with
% respect to mean-square error (forecast) performances
% Fourth, lead/lag and smoothness can be controlled (simultaneously
% up to some point) by SSA:
% Finally, multivariate designs can provide added value.

This study proposes a methodological innovation that extends the classical direct forecasting framework by integrating dimension-reduced information extracted through Multivariate Smooth Sign Accuracy (M-SSA). The core idea is to replace raw
%, unprocessed 
leading indicators with smoothed and filtered components derived from the M-SSA procedure, thereby improving the signal-to-noise ratio in the predictor variables used to forecast GDP. We apply this framework to well-established German economic indicators to show its ability to improve forecasts of GDP and GDP trends, especially over mid-term horizons ranging from two to four quarters ahead, using a variety of sample periods and performance metrics.

Surprisingly, we find that smoothing indicator-based forecasts often provides useful leading information for GDP. We also show that while smoother GDP forecasts lag efficient MSE forecasts slightly, this can be mitigated by adjustments in the forecast horizon.\footnote{The lag reflects the trilemma highlighted by \cite{McElroy2019} in which forecasters choose between forecast accuracy, timeliness and volatility.}

%Previous research often highlights the utility of direct forecasts, particularly for short-term predictions such as nowcasting and forecasts one quarter ahead. However, this paper advances the literature by showing that more sophisticated predictors—specifically, filtered components obtained via M-SSA yield superior performance, especially over mid-term horizons ranging from two to four quarters ahead.

Section \ref{sec:data} describe our data series, the sample period used, and documents their dynamic behaviour. Section \ref{sec:direct_forecast} provides benchmark forecasts using simple direct projections of GDP based on our indicators. It then demonstrates how in-sample forecast performance may improve when we ``de-noise'' the indicators prior to projection with simple smoothing filters. Section \ref{sec:mSSA} defines the M-SSA method and shows how its forecast performance compares to that of direct projections as well as how it varies with forecast horizon. Section \ref{sec:mSSA_component} considers the ability of our indicators to forecast German GDP growth using the M-SSA framework, and Section \ref{sec:revisions} investigates the temporal stability of the estimated forecasting equations. Section \ref{sec:conclusions} summarizes the results and concludes. The Appendix provides some additional results.

\section{Data and Dependence}\label{sec:data}
\subsection{Data}
The analysis employs a set of well-established leading indicators to forecast quarterly German GDP \citep{Drechsel2012financial,Heinisch2018bottom}, including industrial production (IP), the ifo Business Climate Index (ifo\_c), the Economic Sentiment Indicator (ESI), and the term spread between 10-year bond yields and the 3-month EURIBOR rate (spre\_10y\_3m) covering the period 1995 to 2024. All series are used at a quarterly frequency, with a particular emphasis on forecast horizons of 2 or more quarters, reflecting medium-term forecasting objectives. 
To ensure comparability and mitigate scale effects, all indicators (except the spread, which is differenced only) are log-differenced and standardized, the latter applied solely for ease of visual inspection.
Additionally, extreme outliers associated with the COVID-19 pandemic are trimmed to prevent distortion of the underlying signal.
We take into account the publication lag of the indicators and the GDP series.\footnote{Data revisions are not considered, given that previous studies have shown that their overall effect on forecast performance is minor.} 
While GDP is available with a one-quarter delay, and IP is released with a two-month lag, business surveys and financial data are published without any lag. All series are therefore re-aligned to reflect the forecasters' available information, with GDP and IP shifted forward by their respective lags. Tables \ref{tab:data_stucture} and \ref{tab:data_stucture_aligned} show the data structure, as published by mid-February 2025.

%\begin{table}[ht]
%\centering
%\begin{tabular}{rrrrrr}
%  \hline
% & GDP & ip & ifo\_c & ESI & spr\_10y\_3m \\ 
%  \hline
%2024-07-01 & & 90.700 & 85.800 & 92.500 & -1.300 \\ 
%  2024-08-01 &  & 93.200 & 84.900 & 90.900 & -1.400 \\ 
%  2024-09-01 &902.571   & 91.200 & 83.700 & 89.800 & -1.400 \\ 
%  2024-10-01 &  & 90.800 & 84.100 & 90.700 & -1.200 \\ 
%  2024-11-01 &  & 92.200 & 83.800 & 89.300 & -0.900 \\ 
%  2024-12-01 & 900.761 &  & 82.800 & 86.900 & -0.800 \\ 
%  2025-01-01 &  &  & 82.500 & 88.100 & -0.200 \\ 
%   \hline
%\end{tabular}
%\caption{Ragged end.  } 
%\label{tab:data_stucture}
%\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
  \hline
 & GDP & ip & ifo\_c & ESI & spr\_10y\_3m \\ 
  \hline
2024-07-01 & 902.571 & 90.700 & 85.800 & 92.500 & -1.300 \\ 
  2024-08-01 &  & 93.200 & 84.900 & 90.900 & -1.400 \\ 
  2024-09-01 &  & 91.200 & 83.700 & 89.800 & -1.400 \\ 
  2024-10-01 & 900.761 & 90.800 & 84.100 & 90.700 & -1.200 \\ 
  2024-11-01 &  & 92.200 & 83.800 & 89.300 & -0.900 \\ 
  2024-12-01 &  &  & 82.800 & 86.900 & -0.800 \\ 
  2025-01-01 &  &  & 82.500 & 88.100 & -0.200 \\ 
   \hline
\end{tabular}
\caption{Ragged end.  } 
\label{tab:data_stucture}
\end{table}
% Thu May  8 05:39:20 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
  \hline
 & GDP & ip & ifo\_c & ESI & spr\_10y\_3m \\ 
  \hline
2024-05-01 &  & 93.400 & 84.900 & 90.900 & -1.400 \\ 
  2024-08-01 &  & 90.700 & 83.700 & 89.800 & -1.400 \\ 
  2024-09-01 & 902.571 & 93.200 & 84.100 & 90.700 & -1.200 \\ 
  2024-10-01 &  & 91.200 & 83.800 & 89.300 & -0.900 \\ 
  2024-11-01 &  & 90.800 & 82.800 & 86.900 & -0.800 \\ 
  2024-12-01 & 900.761 & 92.200 & 82.500 & 88.100 & -0.200 \\ 
   \hline
\end{tabular}
\caption{Data aligned at sample end.} 
\label{tab:data_stucture_aligned}
\end{table}

 Monthly series are transformed to quarterly series and are displayed in Figure \ref{fig:data}, while Figure \ref{fig:data_lags} offers a close-up view of the financial crisis and the pandemic.\footnote{Monthly series are averaged to quarterly values.} The latter figure illustrates that the real-time GDP and industrial production series are synchronized at dips and peaks (coincident), whereas the other indicators tend to lead somewhat. The latter can be leveraged in a multivariate approach to enhance forecasting performance compared to univariate benchmarks.

% \begin{figure}[htbp]
% \begin{center}
% \includegraphics[height=3in, width=4.5in]{./Figures/Data.pdf}
% \caption{Data: standardized log-differences, trimmed to $\pm 3$.
% \label{data}}
% \end{center}
% \end{figure}

% \begin{figure}[htbp]
% \begin{center}
% \includegraphics[height=3in, width=4.5in]{./Figures/data_lags.pdf}
% \end{center}
% \caption{Data: leads and lags during financial crisis (left) and Pandemic (right).
% \label{data_lags}}
% \end{figure}

\begin{figure}
  \begin{subfigure}[t]{0.49\textwidth}
    \centering
    \includegraphics[width=\linewidth]{./Figures/Data.pdf}
    \subcaption{Full Sample.}
    \label{fig:data}
\end{subfigure}
\hfill
\begin{subfigure}[t]{0.49\textwidth}
    \centering
    \includegraphics[width=\linewidth]{./Figures/data_lags.pdf}
    \subcaption{Leads \& lags during financial crisis \& Pandemic.}
    \label{fig:data_lags}
\end{subfigure}
\caption{GDP \& Indicators: standardized log-differences, trimmed to $\pm 3$ standard deviations.}
\end{figure}

%Hence, we are targeting a series that comprises the publication lag and the forecast horizon.

%[Simon and Katja: not sure if we need the following sentences:]
%Benchmark for analysis are the historical mean and the AR-predictions

%%%%%%%%%%%%%%%%%

%\footnote{Code is provided at \url{ }.}
%https://github.com/wiaidp/R-package-SSA-Predictor/blob/main/SSA%20Tutorials/tutorial_7.4_MSSA_component_predictor.R}}
%Among our results to date, we find that industrial production and economic outlook surveys provide much more valuable information than the yield curve. We also find that extreme events, such as the Great Financial Crisis or the COVID-19 pandemic, have an important impact on the forecasts. This reflects a common tradeoff facing forecasters: the choice between high accuracy in over the whole period and avoiding rare but large and costly forecast errors in extraordinary situations.

%%%%%%%%%%%%%%%%%%% here comes the GIT-file

\subsection{Dependence}
To examine the dynamic interdependencies among the variables, we estimate their cross-correlations as well as their vector moving-average (VMA) representations. The latter is a key input to the multivariate filters we develop in Sections \ref{sec:mSSA} and \ref{sec:mSSA_component}.

Figures \ref{CCF} (full sample) and \ref{CCF_wc} (data prior to the pandemic) show the sample cross-correlation functions (CCF's) between current and lagged GDP and the two survey indicators. Both show strong correlations that peak at leads relative to GDP. Results for IP showed no lead, while correlations for the term spread were weaker. The Figures also show that the precise strengths of the correlations are influenced by the outliers associated with the COVID-19 pandemic. In the interest of model stability, we exclude data from Q4-2019 to Q4-2020 (pandemic) when modelling the dynamics for the multivariate filters developed below.

\begin{figure}[ht]
    \begin{subfigure}[htpb]{\textwidth}
        \begin{center}
            \includegraphics[height=3in, width=4.5in]{./Figures/CCF.pdf}
            \caption{Full data set from Q2-1991 to Q4-2024.\label{CCF}}
        \end{center}
    \end{subfigure}
    \\
    \begin{subfigure}[htpb]{\textwidth}
        \begin{center}
            \includegraphics[height=3in, width=4.5in]{./Figures/CCF_wc.pdf}
            \caption{Without Pandemic (omitting Q4-2019 to Q4-2020).\label{CCF_wc}}
        \end{center}
    \end{subfigure}
    \caption{Sample cross-correlation functions (CCF) showing correlations with lags of GDP growth.}
\end{figure}

We then estimated VARMA models to capture these dynamics, selecting a parsimonious VAR(1) specification.\footnote{We used the \texttt{refVARMA()} procedure in the \cite{MTSpackage} \texttt{MTS} package, which sets VARMA coefficients with t-statistics $< 1.5$ to zero. See \cite{tsay2013multivariate}. We intend to explore richer dynamic structures in subsequent work.} Figure \ref{ma_inv_BIP} shows the implied responses of GDP to shocks to each variable.\footnote{Shown are the MA coefficients for GDP from the vector moving-average (VMA) model obtained by inversion of the VAR(1) representation (i.e. the impluse-response functions, or IRFs, of GDP to the correlated reduced-form shocks.)}

\begin{figure}[h]
    \begin{center}
        \includegraphics[height=3in, width=6in]{./Figures/ma_inv_BIP.pdf}
        \caption{VMA Coefficients for GDP Implied by VAR(1)\\Estimation with pandemic (left) and without pandemic (right).
        \label{ma_inv_BIP}}\end{center}
\end{figure}

The model summarizes key features of the data after we account for publication lags. GDP responds most strongly to shocks to the two survey indicators (ifo\_c and ESI) after a modest lag. There also appears to be a more modest response to IP shocks after a slightly longer lag, but this vanishes once data after the start of the pandemic is included.\footnote{There is also weaker evidence of a response to the term spread (spr\_10y\_3m) at longer lags. This should not be confused with the implied response to a \textit{structural} shock monetary policy shock.} The delayed response of GDP to movements in these various indicators suggests that multivariate models may provide better GDP forecasts than univariate models, a question we pursue in the next section.
%, where lagging series (GDP and IP, after accounting for publication lags) depend on indicators leading in relative terms. In particular, the impulse response of the spread suggests the relevance of a univariate model for this specific leading indicator.

\section{Forecasts with Pre-Filtering}\label{sec:direct_forecast}

In this section, we demonstrate how pre-filtering indicator variables may improve the performance of simple forecasts by removing noise. For simplicity, all the models in this section are direct projections of GDP on our indicators and take the form

\begin{equation}
    u_{t+h} = \beta \cdot \mathbf{X}_t + \varepsilon_t
    \label{eq:direct_forecast}
\end{equation}

where
\begin{itemize}
    \item $u_{t+h}$ is the growth rate of GDP, $h = \{0, 1, 2, ... \}$
    \item $\mathbf{X}_t \equiv [1,u_{t-1}, ip_t, ifo\_c_t, ESI_t, spr\_10y\_3m_t]'$ or their transformed counterparts and $u_{t-1}$ accounts for the publication lag of GDP.\footnote{Including lagged GDP enables the nesting of the AR forecast within the broader modeling framework.}
    \item $\hat{\beta}$ is estimated by least-squares, omitting observations affected by the pandemic.
\end{itemize}

We then demonstrate how in-sample forecast performance improves when we ``de-noise'' the indicators prior to projection with simple smoothing filters.\footnote{The smoothing filters are one-sided (concurrent) filters of length 31, therefore, estimation of both forecasting models omits the first 30 observations.}

\subsection{Unfiltered Indicators}\label{cdf}

%So-called direct forecasts are obtained by regressing the relevant indicators on forward-shifted GDP, accounting for the additional publication lag. To illustrate the concept, we rely on all indicators and full sample information, excluding observations affected by the pandemic. In-sample forecasts and the forward-shifted GDP, which serves as the target, are presented in Figures \ref{direct_wc} (full dataset without the pandemic) and \ref{direct_wc_financial_crisis} (financial crisis) for forecast horizons up to three quarters ahead.

\begin{figure}[h]
    \begin{center}
        \includegraphics[width=0.9\textwidth]{./Figures/direct_wc_all.pdf}
        %\includegraphics[height=3in, width=4.5in]{./Figures/direct_wc_all.pdf}
        \caption{Direct forecasts (red lines) based on all indicators and full sample information (without pandemic): GDP (black line) shifted by zero, one, two, and three quarters.
        \label{direct_wc}}
    \end{center}
\end{figure}

Figures \ref{direct_wc} and \ref{direct_wc_financial_crisis} compare actual and fitted values for GDP at forecast horizons $h = 0, \ldots, 3$, using the four indicators and lagged GDP, without applying any filtering. Figure \ref{direct_wc} shows results over the sample without the pandemic, while \ref{direct_wc_financial_crisis} provides a closer look at the fit during the financial crisis. Table \ref{tab:pvaluedhp} shows p-values for tests of the joint null hypothesis that all coefficients on the indicators are zero (HAC-adjusted Wald test).

\begin{figure}[h]
    \begin{center}
        \includegraphics[width=0.9\textwidth]{./Figures/direct_wc_financial_crisis.pdf}
        \caption{Direct forecast across the financial crisis:  regression based on all indicators and full sample information, without pandemic.
        \label{direct_wc_financial_crisis}}
    \end{center}
\end{figure}

The results indicate that as the forecast horizon $h$ increases, the turning points (i.e., peaks and troughs) of the direct forecasts become increasingly right-shifted (lagged) relative to the target GDP. Additionally, the p-values of the Hac-adjusted Wald test precipitously increase from $h=2$ to $h=3$, see Table \ref{tab:pvaluedhp}, implying that the indicators become statistically insignificant for forecast horizons of more than two quarters.\footnote{ 
These findings are confirmed by the relative root mean-square errors comparing direct forecasts against the mean benchmark, see Table \ref{rRMSE_mSSA_direct_mean_without_covid8} in the appendix.}

%\begin{table}[ht]
%\centering
%\begin{tabular}{rrrrrrrr}
%  \hline
% & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
%  \hline
%P-values & 0.000 & 0.000 & 0.007 & 0.946 & 0.603 & 0.096 & 0.181 \\ 
%   \hline
%\end{tabular}
%\caption{Statistical significance of direct forecasts.  } 
%\label{tab:f_stats}
%\end{table}

\subsection{Direct Forecasts Based on Filtered Indicators}\label{hpdf}
%\subsection{Filter: HP(160)}

%As shown, the performance of direct forecasts sharply declines for longer forecast horizons. 
We conjecture that economic indicators may be contaminated by unpredictable high-frequency noise, thereby obscuring the effective `signal' and making a direct regression more susceptible to overfitting. If so, it may be possible to highlight the signal by filtering our indicators to dampen high-frequency noise.

The Hodrick-Prescott (HP) filter is a classic tool used in business cycle analysis: the filter is specified by a single smoothing parameter $\lambda$ and the value $\lambda=1600$ is specifically recommended for quarterly data. However, \cite{Phillips_Jin_2021} noted that the HP(1600) filter removes relevant information due to excessive smoothing. In our context, this oversmoothing issue would be further exacerbated when considering forecast horizons shorter than a year---an interval inconsistent with the mean duration of up to several years of business cycles, as highlighted by the HP(1600). To address this concern, we suggest a more adaptive HP(160) `target' filter.

Figure \ref{hp_160} presents both classic two-sided and one-sided HP(160) (denoted HP-C for \textit{Concurrent}) filters (top panels) along with their associated amplitude functions in the frequency domain (bottom panels). It confirms that the HP-C filter assigns greater weight to yearly components—relevant in a mid-term forecast context— than the classic quarterly HP filter, while also reducing undesirable high-frequency noise. Therefore, while the choice of the HP(160) filter in this example
is somewhat arbitrary, it should serve the purpose of reducing high-frequency noise while preserving longer-duration movements.\footnote{A comprehensive technical analysis of the effect of $\lambda$ on the resulting GDP predictor can be conducted using the M-SSA package (\url{https://github.com/wiaidp/R-package-SSA-Predictor}).}

\begin{figure}[h]
    \begin{center}
        \includegraphics[height=4in, width=5.5in]{./Figures/hp_160.pdf}
        \caption{Two-sided HP(160) (top left) and one-sided (concurrent) HP-C filters (top right) with corresponding amplitude functions (bottom).
        \label{hp_160}}
    \end{center}
\end{figure}

%\subsection{Direct HP Forecast}\label{hpdf}

%Extending the `classic' direct forecast design, we consider filtered indicators, based on the one-sided HP(160), denoted as the HP-C filter: the filtered indicators are explanatory variables for the regressions on forward-shifted GDP. As in the previous section, we rely on all indicators as well as the full sample estimate for illustration,

Figures \ref{direct_hp_forecasts} and \ref{direct_hp_forecasts_financial_crisis} provide results comparable to those of Figures \ref{direct_wc} and \ref{direct_wc_financial_crisis} but now using indicators smoothed with the HP-C filter. As before, the first Figure covers the full sample without the pandemic while the latter provides a closer look at performance during the financial crisis.  We see that the fitted values lag GDP slightly less and track turning points somewhat more quickly. Table \ref{tab:pvaluedhp} shows p-values for tests of the joint null hypothesis that all coefficients $\hat{\beta} = 0$. We see that the use of filtered indicators improves their statistical significance at almost all forecast horizons, particularly for $h > 2$ quarters. 
%Figures \ref{direct_hp_forecasts} (all observations with exclusion of the pandemic) and \ref{direct_hp_forecasts_financial_crisis} (financial crisis). A comparison with the classic (unfiltered) direct forecasts in the latter figure %in Fig.\eqref{direct_wc_financial_crisis} 
% indicates that the new predictors are slightly less retarded and track the turning points more promptly (`faster'). 
% The filtered approach also exhibits superior statistical properties. As shown in Table \ref{tab:f_stats}, 
%The p-values of the F-statistics for the HP-based forecasts are smaller than for the direct forecasts specifically at $h=3$, but the predictor remains statistically insignificant at forecast horizons exceeding two quarters (even though the evaluation is conducted in-sample and based on all available indicators).

\begin{figure}[h]
    \begin{center}
        \includegraphics[width=0.85\textwidth]{./Figures/direct_hp_forecasts.pdf}
        \caption{Direct HP forecasts: entire data set with exclusion of the pandemic.
        \label{direct_hp_forecasts}}
    \end{center}
\end{figure}

\begin{figure}[h]
    \begin{center}
        \includegraphics[height=5in, width=6in]{./Figures/direct_hp_forecasts_financial_crisis.pdf}
        \caption{Shifted GDP (black), direct HP forecast (red) and classic direct forecast (blue) over the curse of the financial crisis: regression based on all indicators and full sample information (without pandemic).
        \label{direct_hp_forecasts_financial_crisis}}
    \end{center}
\end{figure}

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May  7 12:49:46 2025

%\begin{table}[ht]
%\centering
%\begin{tabular}{rrrrr}
%  \hline
% & h=3 & h=4 & h=5 & h=6 \\ 
%  \hline
%rRMSE & 0.962 & 0.967 & 0.990 & 0.967 \\ 
%   \hline
%\end{tabular}
%\caption{Relative root mean-square error of HP-C against classic direct forecast for $h\geq %3$.\\Values $<1$ imply  HP-C has smaller RMSE.  } 
%\label{tab:rRMSE}
%\end{table}
\begin{table}[h]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h= 0 & h= 1 & h= 2 & h= 3 & h= 4 & h= 5 & h= 6 \\ 
  \hline
  Unfiltered & 0.028 & 0.000 & 0.001 & 0.910 & 0.509 & 0.091 & 0.178 \\ 
HP-C filtered & 0.000 & 0.001 & 0.036 & 0.046 & 0.049 & 0.012 & 0.022 \\ 
   \hline
\end{tabular}
\caption{p-values for $H_0: \hat{\beta} = 0$, HAC-adjusted Wald test.\\Estimation over full sample (without pandemic.)} 
\label{tab:pvaluedhp}
\end{table}

%We now propose a more refined causal filter design aiming for statistical significance of the resulting predictor at longer forecast horizons, out-of-sample.

\section{Multivariate Causal Filter: M-SSA}\label{sec:mSSA}

The application of the one-sided HP-C filter to the indicators yields some measurable improvements in forecast performance. However, its univariate nature excludes additional information such as that provided by leading indicators. The noise suppression capability of the one-sided filter is compromised, as explained by \cite{Wildi2025} and illustrated by the high-frequency leakage of the amplitude function in Figure \ref{hp_160}. To address these limitations, we propose an extension of the Smooth Sign Accuracy (SSA) framework introduced by \cite{Wildi2025}.

\subsection{Optimization criterion}

Let $\mathbf{X}_t$ (of dimension $t\times n$) denote a set of $n$ explanatory series $\mathbf{x}_{1t},...,\mathbf{x}_{nt}$, with observations $\mathbf{x}_{it}=(x_{i1},...,x_{it})'$. Let $\mathbf{z}_t=(z_{1},...,z_t)$  denote a target series, which typically lies outside the linear space spanned by $\mathbf{X}_t$  (and is unknown at time $t$). For simplicity, we assume stationarity of all time series involved. In this framework, $\mathbf{X}_t$ corresponds to a matrix of selected economic indicators with $n=5$, and $z_{it}$, i=1,...,5,   is the output of a two-sided HP(160) filter applied to the four indicators as well as to GDP. Since $z_{it}$  depends on future observations $x_{it+1},x_{it+2},...$, the prediction task involves `tracking' $z_{it}$  using the estimate $y^i_{t}=\sum_{j=1}^ny^i_{jt}$, where $y^i_{jt}=\mathbf{b}^i_j~'\mathbf{x}_{jt}=\sum_{k=0}^{L} b^i_{jk}x_{j,t-k}$ are the outputs of a multivariate (causal or one-sided) filter $\mathbf{B}^i$, with columns $\mathbf{b}^i_{j}$ of fixed length $L$, $j=1,...,5$, assigning weights to the last $L$ observations of the $j$-th indicator $\mathbf{x}_{jt}$ (the fixed-length assumption is merely used for ease of exposition).
For the sake of clarity, we now omit the superscript index $i$ for all variables, assuming $z_t=z^{i_0}_{t}$ for some fixed $i_0$.% Additionally, we assume a fixed filter length $L$, such that $\mathbf{B}^i_t=\mathbf{B}_t=\mathbf{B}$, which has dimension $L\times 5$.

The task of `tracking' a target can be formalized in different ways; here, we focus on the correlation $\rho(z,y,h)$ between $y_t$ and target $z_{t+h}$, where $h\geq 0$ denotes the forecast horizon (backcasting with $h<0$ is ignored in this context). For simplicity, assume a fixed horizon, say $h=h_0$, so that we may drop the reference to $h$ in our notation. In the univariate case ($n=1$), \cite{Wildi2025} proposes the Smooth Sign Accuracy (SSA) as an optimization criterion:
\begin{eqnarray}\label{critssa}
\left.\begin{array}{c}\rho(z,y,h)\to\max\\
s.t. \quad \rho(y,1)=\rho_1\end{array}\right\}
\end{eqnarray}
where $\rho(y,1)$ is the first-order autocorrelation of the predictor $y_t$. The parameter $\rho_1$ controls the smoothness of the predictor: higher values of $\rho_1$ favour smoother trajectories of $y_t$. Assuming that the process $x_t$ has mean zero, \cite{Wildi2024} establishes a connection between $\rho(y,1)$ and the expected duration between consecutive `zero-crossings' (sign changes) of $y_t$, the ``holding time'', denoted by $ht(y)$:
\begin{eqnarray}\label{ht}
ht(y)=\frac{\pi}{\arccos(\rho(y,1))}.
\label{eq:ht_arccos}
\end{eqnarray}
Since \ref{eq:ht_arccos} is strictly monotonic, the SSA criterion can be reformulated as the optimization problem:
\begin{eqnarray}\label{critssaht}
\left.\begin{array}{c}\rho(z,y)\to\max\\
\frac{\arccos(\rho(y,1))}{\pi}=1/ht_1\end{array}\right\},
\end{eqnarray}
where the smoothness parameter $1/ht_1$ expresses the expected rate of sign changes of the predictor.\footnote{Formally, the exact relationship between the rate of sign changes and the lag-one autocorrelation of the predictor assumes Gaussian time series, but \cite{Wildi2024} demonstrates robustness to deviations from this assumption.} \\

\eqref{critssaht} has several important implications. First, the objective function is indifferent to an affine transformation of the predictor. This ambiguity can be resolved by assuming an arbitrary scale and level for $y_t$ (standardization). Alternatively, a mean-square error norm (MSE) may be substituted for the target correlation, as noted in \cite{Wildi2025}. In this case, the classic MSE predictor $y_{t,MSE}$ is obtained as a solution to the SSA criterion by insertion of $1/ht_1:=1/ht_{MSE}$ in the constraint, where $ht_{MSE}$ represents the holding time of $y_{t,MSE}$. %If the MSE predictor is noisy, which is often the case in applications, one can select $ht_1>ht_{MSE}$  in the constraint to address the rate of noisy crossings or false alarms. 
Second, the concept can be extended to a multivariate framework, denoted M-SSA, by generalizing the correlation functions as detailed in \cite{Wildi2025}. 
%Lastly, sign changes in trend growth, represented by zero-crossings of the target $z_t$, are indicative of relevant changes in the trajectory of the economy, for example, at transitions between expansions and recessions. 
Lastly, when $z_t$ is GDP growth, sign changes are indicative of expansions and recessions.
Unfortunately, classic predictors often generate excessively many false (`noisy') sign changes due to high-frequency leakage as discussed above.
%as illustrated by the amplitude function in Figure \eqref{hp_160} (bottom right panel) and further elaborated upon by \cite{Wildi2025}. 
 \cite{Wildi2025} shows that the SSA criterion helps control this phenomenon. Using the equivalent dual formulation of criterion \eqref{critssaht}, the author demonstrates that the SSA solution exhibits the smallest rate of zero-crossings of any (linear) predictor with the same target correlation. Below, we therefore use the M-SSA framework to construct a `smooth' predictor $y_t$ of $z_t$. %In a second stage, serving as a generalization and refinement of the univariate HP-C utilized in the direct HP forecast.

\subsection{Example: Out of Sample M-MSE vs M-SSA Nowcasts}

For illustration, consider the problem of nowcasting HP-filtered GDP, denoted by HP-GDP, where the target $z_t$ is generated using the two-sided HP(160) filter examined previously. We compare the classic multivariate mean squared error (MSE) predictor (denoted by M-MSE) with the M-SSA predictor where we set $ht_1=1.5ht_{MSE}$. This constraint ensures that the M-SSA predictor produces roughly 33$\%$ fewer sign changes than M-MSE in large samples. Comparisons with M-MSE allow us to see the cost of this constraint in terms of prediction performance (target correlation). 
Both predictors use the same underlying VAR(1) model introduced in Section \ref{sec:direct_forecast} and estimated on data up to January 2008. This ensures a lengthy out-of-sample period that encompasses significant events such as the financial crisis, the sovereign debt crisis, and the COVID-19 pandemic.\footnote{Further details and background information are available in \cite{Wildi2025}. All computations are performed utilizing the M-SSA package, as documented in \cite{Wildi2025}. \url{https://github.com/wiaidp/R-package-SSA-Predictor}}

\begin{figure}[htpb]
    \begin{center}
        \includegraphics[height=3in, width=4.5in]{./Figures/bk_gammak.pdf}
        \caption{Multivariate M-MSE (left) and M-SSA (right) nowcast filter weights.
        \label{bk_gammak}}
    \end{center}
\end{figure}

\begin{figure}[htpb]
    \begin{center}
        \includegraphics[height=3in, width=4.5in]{./Figures/mssa_msse_now.pdf}
        \caption{Two-sided HP(160) applied to GDP (black)\\ Nowcasts: M-SSA (blue) and M-MSE (green). \\The dashed vertical line delimits in- and out-of sample periods.  \\The two-sided filter does not extend to the sample end.
        \label{mssa_msse_now}}
    \end{center}
\end{figure}

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May  7 12:49:47 2025
\begin{table}[ht]
\centering
\begin{tabular}{lrr}
  \hline
 & M-SSA & M-MSE \\ 
  \hline
Target correlation & 0.68 & 0.69 \\ 
  HT & 6.50 & 3.71 \\ 
   \hline
\end{tabular}
\caption{Sample target correlations and holding time (HT) of M-SSA and M-MSE nowcasts.} 
\label{corhtnow}
\end{table}

Figure \ref{bk_gammak} displays the coefficients of the multivariate nowcasts, where M-SSA's increased smoothing reduces the maximum amplitudes of the filter weights and slows their rate of decay at longer lags. The resulting filtered series, presented in Figure \ref{mssa_msse_now}, are standardized to enable direct comparisons. The M-SSA nowcast is slightly smoother than that of the M-MSE, exhibiting approximately one-third fewer zero-crossings. Table \ref{corhtnow} compares empirical holding times and target correlations for both nowcasts. Increased smoothness achieved by the M-SSA comes at the cost of a slight reduction (0.68 vs 0.69) in target correlation. This highlights the inherent trade-off—often referred to as the prediction dilemma—between smoothness (zero-crossing rate) and predictive accuracy (target correlation) within the M-SSA criterion.

The results validate that the smoothing constraint incorporated into the multivariate extension of Criterion \eqref{critssa} effectively manages the zero-crossings of the predictor: sample estimates of the holding times (HTs) align with the expected values derived from Equation \eqref{ht}.\footnote{Effective convergence of sample estimates to their theoretical expected values —assuming the true data-generating process is known— can be validated using the M-SSA package. This validation is typically performed through analyses of very long samples of simulated (multivariate) data, ensuring the consistency of the estimation procedures under idealized conditions.} An increase of $50\%$ in the HT relative to the M-MSE predictor has only a modest effect on the target correlation, suggesting that this trade-off could be beneficial in our application. %Going forward, we now omit the M-MSE, as it is a special case of the more general M-SSA framework.

\subsection{Example: M-SSA Forecasts}

Next, we extend the previous example to consider forecasts of HP-GDP. The only change this requires is to change $h$ to $h=0,...,6$ quarters. %The previously described nowcast example can be systematically extended to generate forecasts of the target variable. This extension utilizes the same empirical framework, incorporating an assumed $50\%$ increase in the HT over the M-MSE, and considers forward shifts of   (note that M-SSA does not target GDP explicitly). %This approach facilitates the assessment of forecast performance across multiple horizons within the established methodological context. 
The influence of the forecast horizon on the M-SSA design is depicted in Figure \ref{bk_h}, which contrasts M-SSA filter weights for the nowcast and the one-year-ahead forecast. Noting the difference in vertical scales, we see that as $h$ increases (right panel), the filter weights diminish, reflecting increased forecast uncertainty. We also see increased weight on the leading indicators relative to that on GDP.\footnote{Weights on IP and ESI are indistinguishable from zero.} Additionally, a phase shift associated with the filter applied to GDP is observed (black line, right panel), reflecting the cyclical characteristics inherent to the HP filter. While this phase effect might cause an effective sign change in the GDP filter output as $h$ increases—something to be avoided in this context—the low-pass filters assigned to the additional 'leading' indicators (green and violet lines, right panel) continue to effectively track the low-frequency components of HP-GDP. The combination of the phase effect, applied to GDP, and level-tracking through these additional indicators enables more refined and effective tracking of the target compared to univariate forecasts, which rely solely on the phase-effect of the filter to look ahead.

\begin{figure}[h]
    \begin{center}
        \includegraphics[width=0.85\textwidth]{./Figures/bk_h.pdf}
        \caption{Multivariate M-SSA: $h=0$ (nowcast, left) and $h=4$ (one year forecast, right).
        \label{bk_h}}
    \end{center}
\end{figure}

\begin{figure}[h]
    \begin{center}
        \includegraphics[width=0.85\textwidth]{./Figures/multivar_vs_univar.pdf}
        \caption{Now- and forecasts of HP-GDP for $h=0,...,6$: HP-C (top) vs. M-SSA (bottom) across the financial crisis (left) and the pandemic (right).
        \label{multivar_vs_univar}}
    \end{center}
\end{figure}

\begin{figure}[h]
    \begin{center}
        \includegraphics[width=0.85\textwidth]{./Figures/mssa_hpc_financial_crisis.pdf}
        \caption{Forward-shifted GDP (black), M-SSA (blue) and HP-C (red) for $h=3,4,5,6$ across the financial crisis. All series are standardized for ease of comparison.
        \label{mssa_hpc_financial_crisis}}
    \end{center}
\end{figure}

M-SSA predictors are compared to the traditional univariate HP-C in Figures  \ref{multivar_vs_univar} and \ref{mssa_hpc_financial_crisis}. In the first figure, we compare both filters across the periods of the financial crisis (top panels) and the COVID-19 pandemic (bottom panels), for forecast horizons $h=0,...,6$. 
%The nowcast from the previous section is included for reference. 
In the second figure, we compare both filters to the forward-shifted GDP for shifts larger than two quarters. All series are standardized for ease of comparison. Forecasts using the M-SSA approach consistently shift leftwards $h$ increases; this shift appears to be consistent across all levels, including the peaks and troughs of the series. Conversely, the positions of peaks and troughs in the HP-C series are largely unaffected by increases in $h$. The leftward-shift of  the M-SSA forecasts can be partly attributed to the increasing weight assigned to the additional leading indicators, which dominate the GDP component at larger forecast horizons. %In comparison, the effect of phase shifts—arising from the cyclical nature of the HP-filter and manifesting in the HP-C forecasts (shown in the left panels)—seems less systematic.

\subsection{Hit Rate vs. False Alarm Rate}

%As shown in Wildi (2025), the SSA criterion's objective function—the target correlation $\rho(y,z,h)$—can be linked to the sign accuracy of the predictor $SA(y_t):=\text{P}\left(\text{sign}(z_{t+h}) = \text{sign}(y_t)\right)$, i.e., the probability that the signs of the predictor $y_t$ and the target $z_{t+h}$ agree:
%\begin{eqnarray}\label{sig_acc}
%SA(y_t)=0.5+\frac{\arcsin(\rho(y,z,h))}{\pi}.
%\end{eqnarray}
%Due to the strict monotonicity of the arcsin function, maximizing the target correlation also maximizes the sign accuracy of the predictor.\footnote{Equation \ref{sig_acc} holds exactly for Gaussian processes. Due to the central limit theorem, the relationship holds approximately for low-pass filtered non-Gaussian series, see Wildi (2025).} 
\textbf{Remarks:}
\begin{itemize}
\item Alternative to MSE, target correlation, holding-time: hit-rate, false alarm. 
\item We can link objective function (target correlation) with sign accuracy, see Wildi (2025). BUT: risk that reviewer asks for comparison with logit or whatever sign-forecast tool.
\item Context ROC/AUC  is missing.
\end{itemize}
For the ROC curves, we consider predictors of GDP based on both original (unfiltered) and filtered indicators. For the direct forecast, we use the regression
\[
GDP_{t+shift}=\beta_0^{shift}+\sum_{i=1}^5\beta_i^{shift}x_{ti}+\epsilon_{t}^{shift,direct}=\hat{GDP}_{t+shift}^{direct}+\epsilon_{t}^{shift,direct},
\]
where $x_{ti}$ designates the original indicators. 
For the filtered indicators, we consider similar regression equations:
\[GDP_{t+shift}=\beta_0^{shift,h,filt}+\sum_{i=1}^5\beta_i^{shift,h,filt}y_{ti}^{h,filt}+\epsilon_{t}^{h,shift,filt}=\hat{GDP}_{t+shift}^{h,filt}+\epsilon_{t}^{h,shift,filt},\]
where $y_{ti}^{h,filt}$ represents the $i$-th filtered indicator optimized at forecast horizon $h$, based on either HP-C, M-MSE, or M-SSA. \\
\textbf{Remarks:}
\begin{itemize}
\item I'm computing all combinations of shift and h. 
\item But we could emphasize the diagonal only (h=shift): simplifies notation/explanation.
\item The table of AUCs (below) is based on diagonal only
\end{itemize}
 ROCs are based on the comparison of signs of the target $GDP_{t+shift}$ and of predictors $\hat{GDP}_{t+shift}^{h,filt}$ (remark: we are using standardized GDP but the regressions account automatically for scales or level shifts, i.e., the results apply directly to original GDP). 
For simplicity of exposition, we present results based on the entire sample, excluding the pandemic period. Fig.\ref{ROC_GDP_shift_1_4} compares ROCs of the predictors for small and mid-term forecast horizons and Table \ref{auc} presents AUC (area under the curve) for forecast horizons 1-5 quarters ahead.
\begin{figure}[h]\begin{center}\includegraphics[height=3in, width=5in]{./Figures/ROC_GDP_shift_1_4.pdf}\caption{Hit rate vs. false alarm rate of the predictors at small (one quarter ahead:left panel) and yearly forecast horizons (right panel).\label{ROC_GDP_shift_1_4}}\end{center}\end{figure}
% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May 21 10:06:12 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Direct forecast & Direct HP forecast & M-MSE & M-SSA \\ 
  \hline
Forecast horizon 1 & 0.752 & 0.715 & 0.746 & 0.747 \\ 
  Forecast horizon 2 & 0.697 & 0.663 & 0.729 & 0.720 \\ 
  Forecast horizon 3 & 0.554 & 0.622 & 0.684 & 0.709 \\ 
  Forecast horizon 4 & 0.486 & 0.624 & 0.680 & 0.708 \\ 
  Forecast horizon 5 & 0.593 & 0.636 & 0.643 & 0.701 \\ 
   \hline
\end{tabular}
\caption{AUCs of predictors against GDP at various forecast horizons.} 
\label{auc}
\end{table}At forecast horizons less than or equal to two quarters, the direct forecast challenges the filters. However, for horizons $h>2$, the filters outperform the direct approach: the univariate HP-C design is surpassed by the classic multivariate MSE filter, which is in turn dominated by the M-SSA. These results confirm that controlling the rate of sign changes of the predictor, while maximizing its target correlation through M-SSA can enhance predictor performance in terms of sign accuracy and false alarm rate.

\newpage

\section{M-SSA Component Predictor}\label{sec:mSSA_component}

The targets of the multivariate filters above were the smoothed outputs obtained from the acausal two-sided HP(160) filter applied to the five indicators. However, our objective is to develop predictors for unfiltered GDP. To achieve this, we use the above M-SSA outputs $y_t^i, \: i = 1, \ldots, 5$ (referred to as M-SSA components) as predictors for forward-shifted GDP.
While we keep the M-SSA fixed, as based on data up to January 2008, we implement an expanding window starting from January 2008 for deriving the component weights, which are up-dated on a quarterly basis. 
%(this contrasts with Section \ref{sec:direct_forecast}, which was based on the full sample information). 
All the results that we present here exclude the pandemic; results including the pandemic are in the appendix.\\

\newpage

% \subsection{Standardized Predictor: Equal-Weighting}

% We begin with a simplifying assumption (relaxed in the next section) that all M-SSA components contribute equally to GDP prediction. This suggests averaging the standardized components to form an equally-weighted composite measure. Subsequently, the predictor is derived by regressing the (forward-shifted) GDP series on these standardized components. Figure \eqref{mssa_predictor} presents the GDP series alongside the equally-weighted M-SSA predictor. All series are standardized to facilitate visual comparison and to highlight the dynamic shifts, particularly the left-shift characteristic of the predictors.

% To enhance interpretability and assessment of the M-SSA predictor, we propose analyzing its individual components. For illustration, Figure \ref{mssa_predictor_interpretability} shows the nowcast at $h=0$. The solid blue line (the nowcast) approaches the zero line near the end of the sample, with a recent upward movement mainly supported by the leading spread component. In contrast, the ifo- and ESI-components remain near the zero line, while the ip- and GDP-components appear to be awaiting additional evidence and confirmation before signaling a definitive movement. This type of assessment can be useful for evaluating the relevance of recent changes in the forecasted time series dynamics.
% \begin{figure}[htpb]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/mssa_predictor.pdf}\caption{GDP and equally-weighted M-SSA predictor: all series standardized.\label{mssa_predictor}}\end{center}\end{figure}\begin{figure}[htpb]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/mssa_predictor_interpretability.pdf}\caption{Equally-weighted M-SSA predictor and M-SSA components: all series standardized.\label{mssa_predictor_interpretability}}\end{center}\end{figure}

% \subsection{Tracking GDP: Optimal Weighting}

%We now utilize the M-SSA components as predictors for forward-shifted GDP.
%, extending the previous (standardized) equally-weighted M-SSA predictor to incorporate optimal weighting of the components.

% There are various combinations of M-SSA components $y_t^i$ for tracking GDP. For illustration, we here select the single GDP component resulting from the M-SSA filter tracking HP-GDP
% \begin{equation}
%     \textrm{HP-GDP}_{t+shift}=\beta_0(shift,h)+\beta_1(shift,h)\cdot \textrm{M-SSA}^{\textrm{GDP}}_t(h) + \epsilon_t^{shift,h}
% \end{equation}

% In this expression, $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$ is optimized for forecast horizon $h$, based on data up to January 2008. Table \eqref{p_val_hpbip_wc} reports HAC-adjusted p-values from regressions of the out-of-sample predictor on HP-GDP, indicating a strong link between the predictor and the target, out-of-sample. 
% % latex table generated in R 4.2.2 by xtable 1.8-4 package
% % Wed May  7 12:49:50 2025
% \begin{table}[ht]
% \centering
% \begin{tabular}{rrrrrrrr}
%   \hline
%  & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
%   \hline
% Shift=0 & 0.000 & 0.000 & 0.001 & 0.002 & 0.005 & 0.020 & 0.112 \\ 
%   Shift=1 & 0.000 & 0.000 & 0.000 & 0.001 & 0.002 & 0.005 & 0.019 \\ 
%   Shift=2 & 0.000 & 0.000 & 0.000 & 0.001 & 0.001 & 0.003 & 0.006 \\ 
%   Shift=3 & 0.018 & 0.007 & 0.002 & 0.000 & 0.002 & 0.003 & 0.004 \\ 
%   Shift=4 & 0.180 & 0.100 & 0.040 & 0.010 & 0.002 & 0.005 & 0.005 \\ 
%   Shift=5 & 0.579 & 0.432 & 0.261 & 0.108 & 0.027 & 0.008 & 0.009 \\ 
%    \hline
% \end{tabular}
% \caption{Regression of shifted HP-GDP on M-SSA($h$) GDP component\\p-values for $H_0: \hat{\beta} = 0$, HAC-adjusted Wald test\\Out-of-sample evaluation starting in Jan-2007, without pandemic.} 
% \label{p_val_hpbip_wc}
% \end{table}Our choice—focusing on the single M-SSA GDP component—emphasizes simplicity and interpretability, as future GDP and HP-GDP are connected through their shared low-frequency component. Additionally, the M-SSA GDP component is the most important predictor in the regression of the M-SSA outputs on GDP.  Finally, revisions due to quarterly updates of the regression equations are minimized by this predictor, and the estimates are more stable and robust over time\footnote{Other combinations can be experimented with using the M-SSA package. Regression weights of more complex designs, involving multiple components, tend to be more difficult to interpret due to the occurrence of negative weights and potential sign changes of the regressors over time, which may indicate overfitting or non-stationarity.} (see Section \eqref{sec:revisions}). Note that information from the other indicators is incorporated into the M-SSA GDP component through the multivariate filter design.\\

% This setup allows us to evaluate the out-of-sample performance of various predictor approaches, including: (i) the simple mean of GDP, (ii) the direct forecast method described in Section \eqref{cdf}, (iii) the direct HP forecasts from Section \eqref{hpdf}, based on the univariate HP-C, and (iv) the new M-SSA GDP component predictor. %For completeness, we also consider an M-MSE component predictor derived from outputs of the multivariate MSE filter.

%Out-of-sample performance metrics include p-values derived from regressions of the out-of-sample predictors on forward-shifted GDP, adjusted for heteroscedasticity and autocorrelation using HAC estimators. Additionally, we report relative root-mean-square forecast errors (rRMSE) benchmarked against various reference models.\\

\subsection{Tracking GDP: Optimal Weighting}

Let $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$ denote the M-SSA predictor of HP-GDP at forecast horizon $h$. A predictor of GDP can be obtained from the simple regression equation
\begin{equation}
\textrm{GDP}_{t+shift}=\beta_0(shift,h)+\beta_1(shift,h) \cdot \textrm{M-SSA}^{\textrm{GDP}}_t(h) + \epsilon_t^{shift,h}
\label{eq:GDP_M-SSA}
\end{equation}
The purposes of the regression are twofold: first, the parameters enable affine transformations of the left-hand side variable, so that the standardized target $\textrm{GDP}_{t+shift}$ can be replaced by the original log-returns of GDP; second, the scale adjustment by $\beta_1(shift,h)$ compensates for the deflation (zero-shrinkage) of $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$, reflecting greater uncertainty  as the forecast horizon $h$ extends. This simple transformation allows $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$ to serve as a predictor for $\textrm{GDP}_{t+shift}$ even when $h\neq shift$. This flexibility enables leveraging the dynamic left-shift of $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$, illustrated in Figure \ref{multivar_vs_univar}, when $shift<h$.\\

Define the (out-of-sample) GDP component predictor $\hat{\textrm{GDP}}_{t+shift}(h)$:  
\begin{equation}\label{oos_pred}
\hat{\textrm{GDP}}_{t+shift}(h):=\hat{\beta}_0(shift,h,t)+\hat{\beta}_1(shift,h,t) \cdot \textrm{M-SSA}^{\textrm{GDP}}_t(h),
\end{equation}
where the regression parameters $\hat{\beta}_0(shift,h,t),\hat{\beta}_1(shift,h,t)$ are estimated using data up to time point $t-1$ (expanding window). Since M-SSA is based on data up to January 2008, $\hat{\textrm{GDP}}_{t+shift}(h)$ constitutes an out-of-sample predictor of $\textrm{GDP}_{t+shift}$ for all positive $shift\geq 0$ and all $t$ larger than January 2008. Define the out-of-sample forecast error
\begin{equation}\label{oosfe}
\hat{\epsilon}_t^{shift,h}:={\textrm{GDP}}_{t+shift}-\hat{\textrm{GDP}}_{t+shift}(h).
\end{equation}
We now assume that the regression parameters $\hat{\beta}_0(shift,h,t)$ and$,\hat{\beta}_1(shift,h,t)$ are up-dated quarterly. Note that Equations \ref{eq:GDP_M-SSA} and \ref{oos_pred} could be extended to incorporate additional M-SSA components, but for illustration purposes, we focus on the simplest (and most robust) design.

\subsection{Out-of-Sample Forecast Performances}\label{oosp}

The statistical significance of the (out of sample) GDP component predictor in Equation \ref{oos_pred}  can be evaluated by regressing 
$\textrm{GDP}_{t+shift}$ on $\hat{\textrm{GDP}}_{t+shift}(h)$:
\begin{equation}\label{eq_gdp_pred}
\textrm{GDP}_{t+shift}=\alpha_0+\alpha_1\hat{\textrm{GDP}}_{t+shift}(h)+\nu_{t}^{shift,h},
\end{equation}
for observations after January 2008: the corresponding HAC-adjusted p-values are reported in Table \ref{p_val_wc}.\footnote{Statistical significance is influenced by the relatively short out-of-sample span, of length 17 years.} The MSE forecast performance of $\hat{\textrm{GDP}}_{t+shift}(h)$ can be benchmarked against the mean predictor (see Table \ref{rRMSE_mSSA_comp_mean_wc}) as well as against the direct forecast (see Table \ref{rRMSE_mSSA_comp_direct_without_covid7}). For consistency, both benchmark predictors are also quarterly updated out-of-sample designs.\\

% Obviously, predicting HP-GDP (see Table \ref{p_val_hpbip_wc}) is less challenging than predicting GDP (see Table \eqref{p_val_wc}), due to smoothness (autocorrelation) of the target when compared to GDP.

Our results can be summarized as follows. When excluding the pandemic, the direct forecasts outperform the mean forecast up to two quarters ahead, see Table \ref{rRMSE_mSSA_direct_mean_without_covid8}\footnote{However, the pandemic period generally obscures the relationships between predictors and GDP.}. However, the GDP component predictors $\hat{\textrm{GDP}}_{t+shift}(h)$, optimized for longer forecast horizons $h\geq 5$, outperform the mean benchmark up to 5Q ahead, see Table \ref{rRMSE_mSSA_comp_mean_wc}. Additionally,  $\hat{\textrm{GDP}}_{t+shift}(h)$, optimized for large forecast horizons $h\geq 4$, surpass the direct forecasts for $shift\geq 3$, see Table \ref{rRMSE_mSSA_comp_direct_without_covid7}. Statistical significance of the GDP component predictor, optimized for larger forecast horizons, can be established for at least up to one year ahead, see Table \ref{p_val_wc}. Finally, the strong performance of predictors optimized for larger forecast horizons ($h\geq 4$), at smaller shifts ($shifts< 4$) corroborates the importance of the left-shift of the corresponding predictors in relation to forecast performance.

%???check whether rMSE is decreasing towards sample end (because initial estimate towards 2007 is unreliable)???

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May  7 12:49:50 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.084 & 0.047 & 0.017 & 0.005 & 0.002 & 0.001 & 0.038 \\ 
  Shift=1 & 0.230 & 0.160 & 0.094 & 0.032 & 0.005 & 0.001 & 0.004 \\ 
  Shift=2 & 0.801 & 0.666 & 0.414 & 0.173 & 0.026 & 0.005 & 0.006 \\ 
  Shift=3 & 0.876 & 0.961 & 0.917 & 0.683 & 0.248 & 0.068 & 0.039 \\ 
  Shift=4 & 0.370 & 0.819 & 0.990 & 0.953 & 0.533 & 0.117 & 0.030 \\ 
  Shift=5 & 0.353 & 0.802 & 0.983 & 0.943 & 0.571 & 0.239 & 0.043 \\ 
   \hline
\end{tabular}
\caption{p-values for $H_0: \hat{\beta_1} = 0$, HAC-adjusted Wald test, Equation \ref{eq_gdp_pred}.\\Sample excludes pandemic.} 
\label{p_val_wc}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May  7 12:49:50 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.978 & 0.942 & 0.893 & 0.855 & 0.860 & 0.906 & 0.962 \\ 
  Shift=1 & 1.082 & 1.060 & 1.005 & 0.926 & 0.877 & 0.880 & 0.914 \\ 
  Shift=2 & 1.083 & 1.090 & 1.077 & 1.019 & 0.952 & 0.918 & 0.919 \\ 
  Shift=3 & 1.034 & 1.059 & 1.081 & 1.063 & 1.010 & 0.970 & 0.954 \\ 
  Shift=4 & 0.995 & 1.029 & 1.069 & 1.077 & 1.035 & 0.986 & 0.953 \\ 
  Shift=5 & 0.993 & 1.030 & 1.075 & 1.093 & 1.062 & 1.011 & 0.967 \\ 
   \hline
\end{tabular}
\caption{rRMSEs: out-of-sample forecast errors Equation \ref{oosfe} vs expanding mean of GDP\\Out-of-sample evaluation starting in Jan-2008, without pandemic.\\Values $< 1$ indicate superior predictions from Equation \ref{oos_pred}.} 
\label{rRMSE_mSSA_comp_mean_wc}
\end{table}

\section{Revisions}\label{sec:revisions}

Revisions refer to changes in the historical values of a time series resulting from updates with new information. Revisions of the M-SSA component predictor can arise from data updates, which we do not analyze further here\footnote{Except for GDP, the other indicators are not or only slightly revised. Moreover, the smoothing effect of filters mitigates the effect of revisions when compared to direct forecasts.}, or from updates to predictor parameters. The latter can be further separated into changes in VAR parameters and regression weights.

\subsection{Revisions: Regression Equation Only}

\begin{figure}[h]
    \begin{center}
        \includegraphics[height=3in, width=4.5in]{./Figures/revisions1.pdf}
        \caption{Revisions of the M-SSA GDP component predictor are shown, with the final predictor in blue and the real-time design in red. These revisions result from quarterly updates to the regression coefficients of equation \ref{oos_pred}. The M-SSA filter    itself remains fixed, based on data up to January 2008.
        \label{revisions1}}
    \end{center}
    \end{figure}

\begin{figure}[h!]
    \begin{center}
        \includegraphics[height=3in, width=4.5in]{./Figures/revisions2.pdf}
        \caption{Revisions of intercept and regression weight of Equation \ref{oos_pred}.
        \label{revisions2}}
    \end{center}
\end{figure}

For illustration, we focus on a forecast horizon $h=4$ for the GDP component predictor and a forward-shift $shift=4$ for GDP, noting that similar results would be obtained with other combinations of $h$ and $shift$. 
%???Check left-shift of MA-inversion of M-SSA and M-MSE
%??? Check forecast excess for HP-C
%??? Justification HP(160): replace HP(1600) by AR(2) with similar persistency but shorter cycle lengthi, i.e. $a_1=-2"persistency*acos(\theta, a_2=persistency$.
Figure \ref{revisions1} compares the final and real-time predictors: as the sample size increases, the real-time predictor converges steadily to the final predictor.\footnote{The rate of convergence is partially determined by the simplicity of our design, which avoids incorporating additional M-SSA components.} 
As a confirmation, Figure \ref{revisions2} displays the evolution of the intercept and regression weights over time: with increasing sample size, these estimates seem to converge to fixed points, indicating both the stationarity of the process and the consistency of the estimates. (Note, however, the slightly elevated values of the regression weight assigned to the M-SSA component during the financial crisis and the pandemic, suggesting a stronger link between the target and the predictor).  In contrast, more complex designs based on multiple M-SSA components can exhibit substantial fluctuations in the regression parameters, sometimes changing signs. This suggests instability and complicates interpretation.

\subsection{Revisions: M-SSA Only}

\begin{figure}[h!]
    \begin{center}
        \includegraphics[height=3in, width=4.5in]{./Figures/up_dated_ma_inv_multi_ip.pdf}
        \caption{MA-inversion (impulse response) of VAR: data up to Jan-2008 (left) vs. entire data set without pandemic (right).
        \label{up_dated_ma_inv_multi_ip}}
    \end{center}
\end{figure}

\begin{figure}[h!]
    \begin{center}
        \includegraphics[height=3in, width=4.5in]{./Figures/bk_2008_all.pdf}
        \caption{M-SSA filter tracking HP-GDP at forecast horizon $h=4$: data up to Jan-2008 (left) vs. entire data set without pandemic (right).
        \label{bk_2008_all}}
    \end{center}
\end{figure}

\begin{figure}[h!]
    \begin{center}
        \includegraphics[height=3in, width=4.5in]{./Figures/revisions3.pdf}
        \caption{Revisions resulting from updates to the VAR model in M-SSA optimized for $h=shift=4$. To isolate the effect of the regression revision, all series are standardized. The up-dated M-SSA (blue) is left-shifted at the zero-crossings and appears slightly faster. For consistency, the entire pandemic episode has been removed.
        \label{revisions3}}
    \end{center}
\end{figure}

The second source of revisions in the M-SSA component predictor arises from updates to the VAR model within the M-SSA filter. We compare the original estimates, based on data up to January 2008, with the final estimates obtained using $h=4$ and $shift=4$ (similar findings apply to other values of forecast horizon and forward-shift).  For consistency, we excluded the entire pandemic episode from the data. Figure \ref{up_dated_ma_inv_multi_ip} compares the MA-inversions (impulse responses) of the GDP equation derived from the 'old' and the updated VAR models: unlike the former, the latter incorporates ESI as an additional explanatory variable for GDP. Figure \ref{bk_2008_all} illustrates the resulting effect on the M-SSA filter targeting HP-GDP at forecast horizon $h=4$. The updated filter (right panel) assigns more weight to the additional 'leading' indicators relative to GDP. Finally, Figure \ref{revisions3} compares the M-SSA GDP predictors. To isolate the effect of the regression revision, the figure displays standardized series. The update to the VAR model influences the left-shift of the predictor (blue line), which appears slightly 'faster' than the original predictor (red line). This suggests that the out-of-sample performance of the M-SSA component predictor may be better than reported in Section \ref{oosp}, which is based on the fixed and increasingly outdated version of M-SSA.

\section{Summary and Conclusion}\label{sec:conclusions}

Classic direct forecasts of GDP can outperform simple benchmarks, such as the mean, up to two quarters ahead. However, this limited forecast horizon is often insufficient for certain applications. A key challenge with direct forecasts is that the relevant indicators tend to be noisy, which can mask the true signal and lead to overfitting, see \cite{Hastie2009}. In particular, the predictor struggles to timely track dips (slowdowns, recessions) or peaks (recoveries, expansions). To address this issue, we apply a HP(160)-filter to smooth the data, with the smoothing parameter $\lambda=160$ reflecting a more adaptive design than the classic quarterly HP, in line with an intended forecast horizon of up to one year ahead. The resulting predictor tends to outperform the classic direct forecasts in terms of statistical significance, mainly due to a slight left-shift noticeable at sharp recessions such as the Great Financial Crisis or Covid19-Pandemic. However, despite the filtering, the predictor remains somewhat noisy, potentially producing excessive noisy sign changes at the transitions between recessions and expansions. This issue is partly due to noise leakage inherent in the one-sided HP filter. \\

To address these problems, we propose a multivariate extension of SSA based on \cite{Wildi2024}. Unlike the univariate HP filter, the M-SSA can leverage cross-sectional information from indicators leading GDP in real time while controlling for the rate of zero-crossings of the predictor. Consequently, the multivariate filter outputs become increasingly left-shifted as the forecast horizon lengthens, allowing dips and peaks of the target series to be tracked more systematically, while reducing the occurrence of noisy zero-crossings. However, since M-SSA does not explicitly target GDP, an additional step is necessary to derive the final GDP predictor.\\% We propose two approaches: a simple equally-weighted aggregate, assuming all M-SSA components are equally important for predicting GDP, and an optimally weighted predictor based on a regression of the M-SSA components on future GDP.\\

For illustration, we consider a simple approach based on regressing the single M-SSA GDP output on future GDP, thereby ignoring the other filter outputs. This predictor is intuitively appealing because the future HP-GDP, i.e., the target of M-SSA—is the low-frequency component of the future GDP target. Consequently, a strong link between M-SSA and future HP-GDP also indicates a connection with GDP, even if the statistical significance is obscured by the noise inherent in GDP. Out-of-sample performance suggests that this predictor is statistically significant at horizons up to one year. Moreover, designs optimized for larger forecast horizons outperform the mean benchmark at all considered forecast horizons, and the predictor exceeds the performance of direct forecasts---based on unfiltered or HP filtered indicators---at horizons longer than two quarters. An analysis of revision errors arising from quarterly updates of the proposed M-SSA GDP component predictor indicates that the regression parameters, the VAR model, and the resulting M-SSA filters stabilize after an initial burn-in period. This stability enhances the explainability and interpretability of the predictor.\\

In conclusion, the new predictor design integrates the traditional direct forecast approach with a novel multivariate filter to target quarterly GDP growth up to one year ahead. This multivariate approach leverages a set of economic indicators leading GDP in real time while controlling the smoothness of the predictor through the zero-crossing rate.

%%%%%%%%%%%%%%%%%%%
%\newpage
\bibliographystyle{apalike}
\bibliography{references}
%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\section{Appendix}\label{sec:Appendix}

\subsection{Additional Performances Results: Without Pandemic}

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May  7 12:49:51 2025
% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May  7 12:49:51 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 1.144 & 1.102 & 1.045 & 0.999 & 1.006 & 1.059 & 1.125 \\ 
  Shift=1 & 1.238 & 1.213 & 1.151 & 1.060 & 1.004 & 1.007 & 1.047 \\ 
  Shift=2 & 1.178 & 1.186 & 1.171 & 1.108 & 1.035 & 0.999 & 0.999 \\ 
  Shift=3 & 1.011 & 1.035 & 1.057 & 1.039 & 0.988 & 0.948 & 0.933 \\ 
  Shift=4 & 1.010 & 1.045 & 1.086 & 1.094 & 1.051 & 1.002 & 0.968 \\ 
  Shift=5 & 0.982 & 1.019 & 1.064 & 1.081 & 1.051 & 1.000 & 0.957 \\ 
   \hline
\end{tabular}
\caption{Regression of shifted GDP on M-SSA($h$) GDP component.\\rRMSEs benchmarked against the direct forecast.\\Out-of-sample evaluation starting in Jan-2008, without pandemic.} 
\label{rRMSE_mSSA_comp_direct_without_covid7}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May  7 12:49:51 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrr}
  \hline
 & Shift=0 & Shift=1 & Shift=2 & Shift=3 & Shift=4 & Shift=5 \\ 
  \hline
RRMSE & 0.855 & 0.874 & 0.919 & 1.023 & 0.985 & 1.010 \\ 
   \hline
\end{tabular}
\caption{Direct forecasts.\\
rRMSEs benchmarked against the expanding mean.\\Out-of-sample evaluation starting in Jan-2008, without pandemic.} 
\label{rRMSE_mSSA_direct_mean_without_covid8}
\end{table}

\newpage
\subsection{Performances: Including Singular Pandemic Data}

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May  7 12:49:51 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 0.127 & 0.064 & 0.019 & 0.003 & 0.001 & 0.002 & 0.012 \\ 
  Shift=1 & 0.439 & 0.314 & 0.194 & 0.093 & 0.033 & 0.017 & 0.036 \\ 
  Shift=2 & 0.718 & 0.565 & 0.326 & 0.098 & 0.020 & 0.008 & 0.010 \\ 
  Shift=3 & 0.946 & 0.975 & 0.925 & 0.657 & 0.222 & 0.059 & 0.028 \\ 
  Shift=4 & 0.327 & 0.699 & 0.964 & 0.942 & 0.619 & 0.237 & 0.102 \\ 
  Shift=5 & 0.383 & 0.755 & 0.946 & 0.910 & 0.581 & 0.241 & 0.055 \\ 
   \hline
\end{tabular}
\caption{p-values for $H_0: \hat{\beta_1} = 0$, HAC-adjusted Wald test, Equation \ref{eq_gdp_pred}.\\ including the pandemic.} 
\label{p_val1}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May  7 12:49:51 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 1.008 & 0.990 & 0.964 & 0.939 & 0.930 & 0.942 & 0.967 \\ 
  Shift=1 & 1.064 & 1.057 & 1.036 & 1.001 & 0.976 & 0.968 & 0.973 \\ 
  Shift=2 & 1.033 & 1.034 & 1.026 & 0.999 & 0.969 & 0.955 & 0.957 \\ 
  Shift=3 & 1.025 & 1.034 & 1.040 & 1.028 & 1.001 & 0.979 & 0.967 \\ 
  Shift=4 & 0.995 & 1.011 & 1.030 & 1.037 & 1.022 & 1.003 & 0.990 \\ 
  Shift=5 & 1.001 & 1.018 & 1.039 & 1.048 & 1.034 & 1.009 & 0.986 \\ 
   \hline
\end{tabular}
\caption{rRMSEs: out-of-sample forecast errors Equation \ref{oosfe} vs expanding mean of GDP\\Out-of-sample evaluation starting in Jan-2008, including the pandemic.\\Values $< 1$ indicate superior predictions from Equation \ref{oos_pred}.} 
\label{rRMSE_mSSA_comp_mean2}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May  7 12:49:51 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
  \hline
Shift=0 & 1.210 & 1.188 & 1.157 & 1.127 & 1.116 & 1.130 & 1.161 \\ 
  Shift=1 & 0.982 & 0.977 & 0.957 & 0.925 & 0.902 & 0.895 & 0.899 \\ 
  Shift=2 & 1.053 & 1.054 & 1.046 & 1.018 & 0.987 & 0.973 & 0.975 \\ 
  Shift=3 & 1.046 & 1.056 & 1.062 & 1.050 & 1.022 & 0.999 & 0.988 \\ 
  Shift=4 & 0.969 & 0.984 & 1.003 & 1.010 & 0.996 & 0.977 & 0.964 \\ 
  Shift=5 & 0.982 & 0.998 & 1.019 & 1.027 & 1.014 & 0.990 & 0.967 \\ 
   \hline
\end{tabular}
\caption{Regression of shifted GDP on M-SSA($h$) GDP component.\\rRMSEs benchmarked against the direct forecast.\\Out-of-sample evaluation starting in Jan-2008, including the pandemic.} 
\label{rRMSE_mSSA_comp_direct3}
\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Wed May  7 12:49:51 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrr}
  \hline
 & Shift=0 & Shift=1 & Shift=2 & Shift=3 & Shift=4 & Shift=5 \\ 
  \hline
RRMSE & 0.833 & 1.083 & 0.981 & 0.979 & 1.027 & 1.020 \\ 
   \hline
\end{tabular}
\caption{Direct forecasts.\\
rRMSEs benchmarked against the expanding mean.\\Out-of-sample evaluation starting in Jan-2008, including the pandemic.} 
\label{rRMSE_mSSA_direct_mean4}
\end{table}

\end{document}

### 2025-05-21 R/compute_ROC_results.R
up

@@ -249,13 +249,21 @@ colnames(forward_shifted_HP_GDP_mat)<-paste("shift",h_vec,sep="")
 #-------------------------
 # Apply ROC: generate curves and compute AUCs
 
+# Combination of forward-shift (of target) and forecast horizon (underlying optimization)
 shift_vec<-1:5
-hh_vec<-1:5
+hh_vec<-1:6
 AUC_array<-array(dim=c(length(shift_vec),length(hh_vec),4))
 dimnames(AUC_array)<-list(paste("shift=",shift_vec,sep=""),paste("h=",hh_vec,sep=""),c("Direct forecast","Direct HP forecast","M-MSE","M-SSA"))
+# Select shifts>k (otherwise too many plots in single window)
 k<-2
 par(mfrow=c(length(which(shift_vec>k)),length(which(hh_vec>k))))
+# Show plot
 showl<-T
+# Labels: hit rate vs. false alarm rate
+lbls = "Hit"
+# Select target: GDP or HP-GDP
+select_target<-"HP-GDP"
+select_target<-"GDP"
 for (i in 1:length(shift_vec))#i<-1
 { 
   for (j in 1:length(hh_vec))#j<-1
@@ -264,13 +272,20 @@ for (i in 1:length(shift_vec))#i<-1
     h<-hh_vec[j]
     
 # Select target: shifted GDP or HP-GDP
-    target<-as.integer(forward_shifted_HP_GDP_mat[,shift+1]>0)
-    target<-as.integer(forward_shifted_GDP_mat[,shift+1]>0)
+    if (select_target=="GDP")
+    {
+      target<-as.integer(forward_shifted_GDP_mat[,shift+1]>0)
+    } else
     
+    {
+      target<-as.integer(forward_shifted_HP_GDP_mat[,shift+1]>0)
+    }
+# Set up data matrix for ROC calculation    
     ROC_data<-cbind(target,direct_forecast_mat[,h+1],direct_hp_forecast_mat[,h+1],mmse_mat[,h+1],mssa_mat[,h+1])
     rownames(ROC_data)<-rownames(data_roc)
     colnames(ROC_data)<-c("Target","Direct forecast","Direct HP forecast","M-MSE","M-SSA")
     ROC_data<-as.data.frame(na.exclude(ROC_data))
+# Select plots that will be shown    
     showROC = ifelse(i>k&j>k,T,F)
     if (showROC&showl)
[…]

### 2025-05-21 Sweave/BIP_predictor.Rnw
up

@@
 For the filtered indicators, we consider similar regression equations:
-\[GDP_{t+shift}=\beta_0^{shift,h,filt}+\sum_{i=1}^5\beta_i^{shift,h,filt}y_{ti}^{h,filt}+\epsilon_{t}^{h,shift,filt}=\hat{GDP}_{t+shift}^{filt}+\epsilon_{t}^{shift,filt},\]
-where $y_{ti}^{h,filt}$ represents the $i$-th filtered indicator, based on either HP-C, M-MSE, or M-SSA.
+\[GDP_{t+shift}=\beta_0^{shift,h,filt}+\sum_{i=1}^5\beta_i^{shift,h,filt}y_{ti}^{h,filt}+\epsilon_{t}^{h,shift,filt}=\hat{GDP}_{t+shift}^{h,filt}+\epsilon_{t}^{h,shift,filt},\]
+where $y_{ti}^{h,filt}$ represents the $i$-th filtered indicator optimized at forecast horizon $h$, based on either HP-C, M-MSE, or M-SSA. \\
+\textbf{Remarks:}
+\begin{itemize}
+\item I'm computing all combinations of shift and h. 
+\item But we could emphasize the diagonal only (h=shift): simplifies notation/explanation.
+\item The table of AUCs (below) is based on diagonal only
+\end{itemize}
 For simplicity of exposition, we present results based on the full sample, excluding the pandemic period.
-Fig.
+Fig.\ref{ROC_GDP_shift_1_4} compares ROCs of the predictors for small and mid-term forecast horizons and Table \ref{auc} presents AUCs (area under the curve) for forecast horizons 1-5 quarters ahead.
 <<label=init,results=hide>>=
@@
 k<-2
-# Show plot
-showl<-T
 # Labels: hit rate vs. false alarm rate
 lbls = "Hit"
+# Size of legend box in plot
+lg_cex<-0.5
 
@@
 par(mfrow=c(length(which(shift_vec>k)),length(which(hh_vec>k))))
+# Show legend
+showl<-T
 
@@
                   smoothROC , colours =NULL, lwd = 2,
-                  showLegend , lg_cex = 1, lg_ncol = 1)
+                  showLegend , lg_cex = lg_cex, lg_ncol = 1)
     AUC_array_GDP[i,j,]<-unlist(AUC)
@@
 par(mfrow=c(length(which(shift_vec>k)),length(which(hh_vec>k))))
-
+showl<-T
 for (i in 1:length(shift_vec))#i<-1
@@
                   smoothROC , colours =NULL, lwd = 2,
-                  showLegend , lg_cex = 1, lg_ncol = 1)
+                  showLegend , lg_cex = lg_cex, lg_ncol = 1)
     AUC_array_HP_GDP[i,j,]<-unlist(AUC)
@@
 showLegend<-T
-AUC<-ROCplots(ROC_data, showROC , main = paste("shift=",shift,sep=""), lbls = lbls,
+AUC<-ROCplots(ROC_data, showROC , main = "One quarter ahead", lbls = lbls,
                   smoothROC , colours =NULL, lwd = 2,
@@
 showLegend<-F
-AUC<-ROCplots(ROC_data, showROC , main = paste("shift=",shift,sep=""), lbls = lbls,
+AUC<-ROCplots(ROC_data, showROC , main = "One year ahead", lbls = lbls,
                   smoothROC , colours =NULL, lwd = 2,
@@
 xtable(mat1, dec = 1,digits=rep(3,dim(mat1)[2]+1),
-paste("AUCs of different predictors against GDP at various forecast horizons.."),
+paste("AUCs of predictors against GDP at various forecast horizons."),
 label=paste("p_val_wc5",sep=""),
@@
 @
+At forecast horizons less than or equal to two quarters, the direct forecast challenges the filters.
+However, for horizons $h>2$, the filters outperform the direct approach: the univariate HP-C design is surpassed by the classic multivariate MSE filter, which is in turn dominated by the M-SSA.
+These results confirm that controlling the rate of sign changes of the predictor, while maximizing its target correlation through M-SSA can enhance predictor performance in terms of sign accuracy and false alarm rate.
 

### 2025-05-22 Sweave/BIP_predictor.Rnw
up

@@
 date_to_fit<-"2008"
-# Model orders
+# We can select type of VAR model
+VAR_type="BVAR"
+VAR_type="VAR"
+if (VAR_type=="BVAR")
+{
+# Simon's settings for BVAR(4)  
+  lambda_BVAR <- 0.001
+  p<-3
+  q<-0
+}
+if (VAR_type=="VAR")
+{
+# Marc's settings for VAR  
   p<-1
   q<-0
+}
+
 # Filter length
@@
 # Use OLS or WLS in regression of shifted BIP on M-SSA component predictor
+# Default (better): use_garch<-F
 use_garch<-F
@@
 
-# Without Covid
-# Full sample
-
+# Full sample estimation with/without pandemic
 if (recompute_results)
 {
+# With pandemic  
   data_fit<-x_mat
+  if (VAR_type=="VAR")
+  {
 
@@
     V_obj<-VARMA(data_fit,p=p,q=q)
+# Apply regularization: see vignette MTS package
     threshold<-1.5
@@
     save(V_obj,file=paste(path.result,"/VAR",sep=""))
+  }
+  if (VAR_type=="BVAR")
+  {
+    V0_BVAR <- diag(rep(1,n))  # prior for Sigma_a matrix - K x K
+    # precision matrix of size (kp+1) x (kp+1)  if there's a constant
+    #   low values <=> low precision
+
+    C_BVAR <- lambda_BVAR*diag(rep(1,n*p+1))
+    V_obj <- BVAR(data_fit,p = p,C = C_BVAR,V0 = V0_BVAR)
+    save(V_obj,file=paste(path.result,"/BVAR",sep=""))
+  }
 # Without Pandemic  
   data_fit<-x_mat_wc
+  if (VAR_type=="VAR")
+  {
+
     set.seed(12)
-  V_obj_wc<-VARMA(data_fit,p=p,q=q)
+    V_obj<-VARMA(data_fit,p=p,q=q)
+# Apply regularization: see vignette MTS package
     threshold<-1.5
-  V_obj_wc<-refVARMA(V_obj_wc, thres = threshold)
+    V_obj<-refVARMA(V_obj, thres = threshold)
     save(V_obj_wc,file=paste(path.result,"/VAR_wc",sep=""))  
+  }
+  if (VAR_type=="BVAR")
+  {
+    V0_BVAR <- diag(rep(1,n))  # prior for Sigma_a matrix - K x K
+    # precision matrix of size (kp+1) x (kp+1)  if there's a constant
+    #   low values <=> low precision
+    C_BVAR <- lambda_BVAR*diag(rep(1,n*p+1))
+    V_obj_wc <- BVAR(data_fit,p = p,C = C_BVAR,V0 = V0_BVAR)
+    save(V_obj_wc,file=paste(path.result,"/BVAR_wc",sep=""))
+  }
 
 } else
+{
+  if (VAR_type=="VAR")
   {
@@
     load(file=paste(path.result,"/VAR_wc",sep=""))
-
+  } else
+  {
+    load(file=paste(path.result,"/BVAR",sep=""))
+    load(file=paste(path.result,"/BVAR_wc",sep=""))
+  }
 }
@@
 #   -The function computes M-SSA for each forecast horizon h in h_vec
-  mssa_indicator_obj<-compute_mssa_BIP_predictors_func(x_mat,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi)
+  mssa_indicator_obj<-compute_mssa_BIP_predictors_func(x_mat,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi,VAR_type,lambda_BVAR)
 
@@
 
-  mssa_indicator_HP_HT_obj<-compute_mssa_BIP_predictors_func(x_mat,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec_HP_C,h_vec,f_excess,lag_vec,select_vec_multi)
-
+  mssa_indicator_HP_HT_obj<-compute_mssa_BIP_predictors_func(x_mat,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec_HP_C,h_vec,f_excess,lag_vec,select_vec_multi,VAR_type,lambda_BVAR)
+  if (VAR_type=="VAR")
+  {
     save(mssa_indicator_obj,file=paste(path.result,"/nowcast_obj",sep=""))
     save(mssa_indicator_HP_HT_obj,file=paste(path.result,"/nowcast_HT_HP_obj",sep=""))
+  }
+  if (VAR_type=="BVAR")
+  {
+    save(mssa_indicator_obj,file=paste(path.result,"/BV_nowcast_obj",sep=""))
+    save(mssa_indicator_HP_HT_obj,file=paste(path.result,"/BV_nowcast_HT_HP_obj",sep=""))
+  }
 } else
+{
+  if (VAR_type=="VAR")
   {
@@
     load(file=paste(path.result,"/nowcast_HT_HP_obj",sep=""))
+  }
+  if (VAR_type=="BVAR")
+  {
+    load(file=paste(path.result,"/BV_nowcast_obj",sep=""))
+    load(file=paste(path.result,"/BV_nowcast_HT_HP_obj",sep=""))
+  }
+
 }
@@
 #   -The function computes M-SSA for each forecast horizon h in h_vec
-  mssa_indicator_obj<-compute_mssa_BIP_predictors_func(x_mat,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi)
+  mssa_indicator_obj<-compute_mssa_BIP_predictors_func(x_mat,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi,VAR_type,lambda_BVAR)
 # Without pandemic  
-  mssa_indicator_obj_wc<-compute_mssa_BIP_predictors_func(x_mat_wc,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi)
-
+  mssa_indicator_obj_wc<-compute_mssa_BIP_predictors_func(x_mat_wc,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi,VAR_type,lambda_BVAR)
+  if (VAR_type=="VAR")
+  {
     save(mssa_indicator_obj,file=paste(path.result,"/mssa_indicator_obj",sep=""))
     save(mssa_indicator_obj_wc,file=paste(path.result,"/mssa_indicator_obj_wc",sep=""))
+  }
+  if (VAR_type=="BVAR")
+  {
+    save(mssa_indicator_obj,file=paste(path.result,"/BV_mssa_indicator_obj",sep=""))
+    save(mssa_indicator_obj_wc,file=paste(path.result,"/BV_mssa_indicator_obj_wc",sep=""))
+  }
+
 } else
+{
+  if (VAR_type=="VAR")
   {
@@
     load(file=paste(path.result,"/mssa_indicator_obj_wc",sep=""))
+  }
+  if (VAR_type=="BVAR")
+  {
+    load(file=paste(path.result,"/BV_mssa_indicator_obj",sep=""))
+    load(file=paste(path.result,"/BV_mssa_indicator_obj_wc",sep=""))
+  }
 }
@@
 # The results can be saved (overwritten)    
-    if (F)
+  if (VAR_type=="VAR")
   {
@@
   }
+  if (VAR_type=="BVAR")
+  {
+    save(list_perf,file=paste(path.result,"BV_list_perf",sep=""))
+  }
 } else
@@
 # Load all results
+  if (VAR_type=="VAR")
+  {
     load(file=paste(path.result,"list_perf",sep=""))
+  }
+  if (VAR_type=="BVAR")
+  {
+    load(file=paste(path.result,"BV_list_perf",sep=""))
+  }
   p_mat_mssa_components=list_perf$p_mat_mssa_components
@@
 
-# We can select type of VAR model
-VAR_type="BVAR"
-VAR_type="VAR"
-if (VAR_type=="BVAR")
-{
-# Simon's settings for BVAR(4)  
-  lambda_BVAR <- 10 
-  p_full<-4
-  q_full<-0
-}
-if (VAR_type=="VAR")
-{
-# Marc's settings for VAR  
+# Model orders
 p_full<-p
 q_full<-q
-}
 
@@
   final_mssa_indicator_obj<-compute_mssa_BIP_predictors_func(x_mat_wc,lambda_HP,L,date_to_fit_all,p_full,q_full,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi,VAR_type,lambda_BVAR)
+
+  if (VAR_type=="VAR")
+  {
     save(final_mssa_indicator_obj,file=paste(path.result,"/final_mssa_indicator_obj",sep=""))
+  }
+  if (VAR_type=="BVAR")
+  {
+    save(final_mssa_indicator_obj,file=paste(path.result,"/BV_final_mssa_indicator_obj",sep=""))
+  }
 } else
 {
+  if (VAR_type=="VAR")
+  {
     load(file=paste(path.result,"/final_mssa_indicator_obj",sep=""))
+  }
+  if (VAR_type=="BVAR")
+  {
+    load(file=paste(path.result,"/BV_final_mssa_indicator_obj",sep=""))
+  }
 }
@@
 
+  if (VAR_type=="VAR")
+  {
     set.seed(12)
@@
     save(V_obj_full,file=paste(path.result,"/VAR_full",sep=""))
+  }
+  if (VAR_type=="BVAR")
+  {
+    V0_BVAR <- diag(rep(1,n))  # prior for Sigma_a matrix - K x K
+    # precision matrix of size (kp+1) x (kp+1)  if there's a constant
+    #   low values <=> low precision
+    C_BVAR <- lambda_BVAR*diag(rep(1,n*p+1))
+    V_obj_full <- BVAR(data_fit,p = p,C = C_BVAR,V0 = V0_BVAR)
+    save(V_obj_full,file=paste(path.result,"/BVAR_full",sep=""))
+  }
 
@@
   data_fit<-x_mat_wc[which(rownames(x_mat_wc)<date_to_fit),]
-
+  if (VAR_type=="VAR")
+  {
     set.seed(12)
@@
     save(V_obj_2008,file=paste(path.result,"/VAR_2008",sep=""))
+  }
+  if (VAR_type=="BVAR")
+  {
+    V0_BVAR <- diag(rep(1,n))  # prior for Sigma_a matrix - K x K
+    # precision matrix of size (kp+1) x (kp+1)  if there's a constant
+    #   low values <=> low precision
+    C_BVAR <- lambda_BVAR*diag(rep(1,n*p+1))
+    V_obj_2008 <- BVAR(data_fit,p = p,C = C_BVAR,V0 = V0_BVAR)
+    save(V_obj_2008,file=paste(path.result,"/BVAR_2008",sep=""))
+  }
 
 } else
+{
+  if (VAR_type=="VAR")
   {
@@
     load(file=paste(path.result,"/VAR_2008",sep=""))
-
+  }
+  if (VAR_type=="BVAR")
+  {
+    load(file=paste(path.result,"/BVAR_full",sep=""))
+    load(file=paste(path.result,"/BVAR_2008",sep=""))
+  }
 }

### 2025-05-23 Latex/BIP_predictor.tex
up

@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 10:24:26 2025
+% Fri May 23 07:41:20 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 10:24:26 2025
+% Fri May 23 07:41:20 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 10:24:28 2025
+% Fri May 23 07:41:23 2025
 \begin{table}[ht]
@@
 The two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 10:24:28 2025
+% Fri May 23 07:41:35 2025
 \begin{table}[ht]
@@
   \hline
-Target correlation & 0.68 & 0.69 \\ 
-  HT & 6.50 & 3.71 \\ 
+Target correlation & 0.77 & 0.70 \\ 
+  HT & 5.47 & 2.89 \\ 
    \hline
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 10:24:32 2025
+% Fri May 23 07:42:27 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 0.000 & 0.000 & 0.001 & 0.002 & 0.005 & 0.020 & 0.112 \\ 
-  Shift=1 & 0.000 & 0.000 & 0.000 & 0.001 & 0.002 & 0.005 & 0.019 \\ 
-  Shift=2 & 0.000 & 0.000 & 0.000 & 0.001 & 0.001 & 0.003 & 0.006 \\ 
-  Shift=3 & 0.018 & 0.007 & 0.002 & 0.000 & 0.002 & 0.003 & 0.004 \\ 
-  Shift=4 & 0.180 & 0.100 & 0.040 & 0.010 & 0.002 & 0.005 & 0.005 \\ 
-  Shift=5 & 0.579 & 0.432 & 0.261 & 0.108 & 0.027 & 0.008 & 0.009 \\ 
+Shift=0 & 0.000 & 0.000 & 0.002 & 0.006 & 0.023 & 0.092 & 0.356 \\ 
+  Shift=1 & 0.000 & 0.000 & 0.000 & 0.001 & 0.003 & 0.010 & 0.037 \\ 
+  Shift=2 & 0.000 & 0.000 & 0.000 & 0.000 & 0.001 & 0.003 & 0.007 \\ 
+  Shift=3 & 0.001 & 0.000 & 0.000 & 0.000 & 0.001 & 0.002 & 0.002 \\ 
+  Shift=4 & 0.033 & 0.006 & 0.000 & 0.000 & 0.001 & 0.003 & 0.002 \\ 
+  Shift=5 & 0.269 & 0.097 & 0.018 & 0.002 & 0.003 & 0.006 & 0.006 \\ 
    \hline
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 10:24:33 2025
+% Fri May 23 07:42:39 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 0.086 & 0.049 & 0.018 & 0.005 & 0.002 & 0.002 & 0.042 \\ 
-  Shift=1 & 0.238 & 0.165 & 0.097 & 0.033 & 0.005 & 0.001 & 0.005 \\ 
-  Shift=2 & 0.815 & 0.680 & 0.428 & 0.180 & 0.029 & 0.007 & 0.008 \\ 
-  Shift=3 & 0.866 & 0.963 & 0.922 & 0.699 & 0.269 & 0.079 & 0.047 \\ 
-  Shift=4 & 0.361 & 0.808 & 0.989 & 0.954 & 0.546 & 0.127 & 0.034 \\ 
-  Shift=5 & 0.351 & 0.786 & 0.982 & 0.943 & 0.579 & 0.245 & 0.042 \\ 
+Shift=0 & 0.021 & 0.009 & 0.008 & 0.005 & 0.003 & 0.043 & 0.448 \\ 
+  Shift=1 & 0.113 & 0.053 & 0.018 & 0.007 & 0.003 & 0.007 & 0.041 \\ 
+  Shift=2 & 0.402 & 0.209 & 0.089 & 0.023 & 0.002 & 0.003 & 0.010 \\ 
+  Shift=3 & 0.962 & 0.703 & 0.249 & 0.081 & 0.013 & 0.004 & 0.007 \\ 
+  Shift=4 & 0.934 & 0.996 & 0.738 & 0.262 & 0.054 & 0.008 & 0.005 \\ 
+  Shift=5 & 0.751 & 0.987 & 0.953 & 0.510 & 0.195 & 0.039 & 0.013 \\ 
    \hline
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 10:24:33 2025
+% Fri May 23 07:42:39 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 0.950 & 0.917 & 0.877 & 0.851 & 0.865 & 0.910 & 0.960 \\ 
-  Shift=1 & 1.028 & 1.007 & 0.960 & 0.899 & 0.867 & 0.878 & 0.912 \\ 
-  Shift=2 & 1.040 & 1.043 & 1.030 & 0.984 & 0.933 & 0.910 & 0.913 \\ 
-  Shift=3 & 1.006 & 1.024 & 1.040 & 1.025 & 0.983 & 0.952 & 0.941 \\ 
-  Shift=4 & 0.977 & 1.003 & 1.034 & 1.040 & 1.005 & 0.964 & 0.937 \\ 
-  Shift=5 & 0.984 & 1.011 & 1.044 & 1.057 & 1.031 & 0.986 & 0.949 \\ 
+Shift=0 & 0.875 & 0.847 & 0.846 & 0.878 & 0.924 & 0.962 & 0.990 \\ 
+  Shift=1 & 0.947 & 0.896 & 0.853 & 0.852 & 0.884 & 0.919 & 0.949 \\ 
+  Shift=2 & 1.002 & 0.977 & 0.932 & 0.891 & 0.882 & 0.897 & 0.918 \\ 
+  Shift=3 & 1.002 & 0.997 & 0.972 & 0.933 & 0.905 & 0.899 & 0.907 \\ 
+  Shift=4 & 1.000 & 1.009 & 0.999 & 0.969 & 0.935 & 0.911 & 0.902 \\ 
+  Shift=5 & 1.003 & 1.017 & 1.018 & 0.999 & 0.971 & 0.945 & 0.928 \\ 
    \hline
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 10:24:36 2025
+% Fri May 23 07:43:07 2025
 \begin{table}[ht]
@@
   \hline
-Forecast horizon 1 & 0.752 & 0.715 & 0.746 & 0.747 \\ 
-  Forecast horizon 2 & 0.697 & 0.663 & 0.729 & 0.720 \\ 
-  Forecast horizon 3 & 0.554 & 0.622 & 0.684 & 0.709 \\ 
-  Forecast horizon 4 & 0.486 & 0.624 & 0.680 & 0.708 \\ 
-  Forecast horizon 5 & 0.593 & 0.636 & 0.643 & 0.701 \\ 
+Forecast horizon 1 & 0.752 & 0.715 & 0.748 & 0.764 \\ 
+  Forecast horizon 2 & 0.697 & 0.663 & 0.741 & 0.754 \\ 
+  Forecast horizon 3 & 0.554 & 0.622 & 0.708 & 0.709 \\ 
+  Forecast horizon 4 & 0.486 & 0.624 & 0.683 & 0.700 \\ 
+  Forecast horizon 5 & 0.593 & 0.636 & 0.656 & 0.690 \\ 
    \hline
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 10:24:36 2025
+% Fri May 23 07:43:07 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 1.104 & 1.066 & 1.019 & 0.989 & 1.006 & 1.058 & 1.116 \\ 
-  Shift=1 & 1.232 & 1.206 & 1.150 & 1.076 & 1.038 & 1.051 & 1.092 \\ 
-  Shift=2 & 1.146 & 1.150 & 1.136 & 1.085 & 1.028 & 1.002 & 1.006 \\ 
-  Shift=3 & 1.008 & 1.027 & 1.042 & 1.028 & 0.986 & 0.954 & 0.943 \\ 
-  Shift=4 & 0.988 & 1.014 & 1.045 & 1.051 & 1.016 & 0.975 & 0.948 \\ 
-  Shift=5 & 0.967 & 0.995 & 1.027 & 1.039 & 1.013 & 0.970 & 0.933 \\ 
+Shift=0 & 1.017 & 0.985 & 0.984 & 1.021 & 1.074 & 1.119 & 1.151 \\ 
+  Shift=1 & 1.134 & 1.073 & 1.022 & 1.020 & 1.058 & 1.101 & 1.136 \\ 
+  Shift=2 & 1.104 & 1.076 & 1.027 & 0.982 & 0.972 & 0.989 & 1.012 \\ 
+  Shift=3 & 1.004 & 0.999 & 0.974 & 0.935 & 0.907 & 0.902 & 0.909 \\ 
+  Shift=4 & 1.011 & 1.020 & 1.010 & 0.980 & 0.945 & 0.921 & 0.912 \\ 
+  Shift=5 & 0.986 & 1.000 & 1.001 & 0.982 & 0.955 & 0.930 & 0.913 \\ 
    \hline
@@
   \hline
-Shift=0 & 0.142 & 0.074 & 0.023 & 0.003 & 0.001 & 0.002 & 0.013 \\ 
-  Shift=1 & 0.475 & 0.344 & 0.217 & 0.111 & 0.044 & 0.023 & 0.044 \\ 
-  Shift=2 & 0.724 & 0.564 & 0.325 & 0.101 & 0.023 & 0.010 & 0.012 \\ 
-  Shift=3 & 0.943 & 0.979 & 0.933 & 0.670 & 0.234 & 0.066 & 0.031 \\ 
-  Shift=4 & 0.322 & 0.666 & 0.960 & 0.946 & 0.644 & 0.263 & 0.119 \\ 
-  Shift=5 & 0.386 & 0.721 & 0.940 & 0.910 & 0.591 & 0.248 & 0.060 \\ 
+Shift=0 & 0.122 & 0.101 & 0.085 & 0.068 & 0.082 & 0.194 & 0.562 \\ 
+  Shift=1 & 0.460 & 0.393 & 0.324 & 0.248 & 0.159 & 0.093 & 0.099 \\ 
+  Shift=2 & 0.269 & 0.131 & 0.057 & 0.047 & 0.036 & 0.021 & 0.029 \\ 
+  Shift=3 & 0.977 & 0.754 & 0.397 & 0.175 & 0.083 & 0.052 & 0.045 \\ 
+  Shift=4 & 0.708 & 0.962 & 0.891 & 0.545 & 0.257 & 0.117 & 0.062 \\ 
+  Shift=5 & 0.792 & 0.975 & 0.926 & 0.548 & 0.258 & 0.105 & 0.051 \\ 
    \hline
@@
   \hline
-Shift=0 & 0.991 & 0.974 & 0.952 & 0.933 & 0.929 & 0.942 & 0.965 \\ 
-  Shift=1 & 1.037 & 1.030 & 1.012 & 0.985 & 0.966 & 0.963 & 0.969 \\ 
-  Shift=2 & 1.014 & 1.014 & 1.005 & 0.983 & 0.959 & 0.950 & 0.953 \\ 
-  Shift=3 & 1.010 & 1.017 & 1.021 & 1.011 & 0.988 & 0.969 & 0.961 \\ 
-  Shift=4 & 0.987 & 0.999 & 1.015 & 1.020 & 1.008 & 0.992 & 0.981 \\ 
-  Shift=5 & 0.996 & 1.008 & 1.023 & 1.030 & 1.018 & 0.997 & 0.976 \\ 
+Shift=0 & 0.987 & 0.983 & 0.981 & 0.981 & 0.983 & 0.989 & 0.998 \\ 
+  Shift=1 & 1.043 & 1.036 & 1.022 & 1.007 & 0.993 & 0.983 & 0.980 \\ 
+  Shift=2 & 0.997 & 0.992 & 0.985 & 0.977 & 0.969 & 0.964 & 0.965 \\ 
+  Shift=3 & 1.004 & 1.004 & 1.001 & 0.994 & 0.984 & 0.974 & 0.968 \\ 
+  Shift=4 & 0.997 & 1.005 & 1.009 & 1.008 & 1.000 & 0.987 & 0.975 \\ 
+  Shift=5 & 1.009 & 1.013 & 1.013 & 1.010 & 1.001 & 0.988 & 0.976 \\ 
    \hline
@@
   \hline
-Shift=0 & 1.228 & 1.207 & 1.180 & 1.156 & 1.151 & 1.167 & 1.196 \\ 
-  Shift=1 & 0.978 & 0.972 & 0.955 & 0.929 & 0.912 & 0.908 & 0.915 \\ 
-  Shift=2 & 1.030 & 1.029 & 1.021 & 0.998 & 0.974 & 0.964 & 0.968 \\ 
-  Shift=3 & 1.037 & 1.044 & 1.048 & 1.038 & 1.014 & 0.995 & 0.986 \\ 
-  Shift=4 & 0.973 & 0.985 & 1.000 & 1.006 & 0.994 & 0.978 & 0.966 \\ 
-  Shift=5 & 0.981 & 0.993 & 1.008 & 1.015 & 1.003 & 0.982 & 0.962 \\ 
+Shift=0 & 1.223 & 1.218 & 1.216 & 1.216 & 1.219 & 1.226 & 1.237 \\ 
+  Shift=1 & 0.984 & 0.977 & 0.965 & 0.950 & 0.937 & 0.927 & 0.924 \\ 
+  Shift=2 & 1.012 & 1.007 & 1.000 & 0.992 & 0.983 & 0.979 & 0.980 \\ 
+  Shift=3 & 1.031 & 1.031 & 1.028 & 1.021 & 1.010 & 1.000 & 0.994 \\ 
+  Shift=4 & 0.982 & 0.991 & 0.995 & 0.993 & 0.985 & 0.973 & 0.961 \\ 
+  Shift=5 & 0.994 & 0.998 & 0.998 & 0.995 & 0.986 & 0.974 & 0.961 \\ 
    \hline

### 2025-05-23 Sweave/BIP_predictor.Rnw
up

@@
 # Section 2.1: data
-recompute_results<-F
+recompute_results<-T
 
@@
 VAR_type="BVAR"
-VAR_type="VAR"
+#VAR_type="VAR"
 if (VAR_type=="BVAR")
@@
 # Simon's settings for BVAR(4)  
-  lambda_BVAR <- 0.001
+  lambda_BVAR <- 1.001
   p<-3
@@
     threshold<-1.5
-    V_obj<-refVARMA(V_obj, thres = threshold)
+    V_obj_wc<-refVARMA(V_obj, thres = threshold)
     save(V_obj_wc,file=paste(path.result,"/VAR_wc",sep=""))  
@@
 # The following double loop computes all combinations of forward-shifts (of BIP) and forecast horizons (of M-SSA)
-  for (shift in shift_vec)#shift<-1
+  for (shift in shift_vec)#shift<-4
   {
@@
     setTxtProgressBar(pb, shift)
-    for (j in h_vec)#j<-1
+    for (j in h_vec)#j<-4
     {
@@
 }
+if (reg_type=="OLS")
+{
   lm_obj<-lm(dat[,1]~dat[,2:ncol(dat)])
-
+}  
+if (reg_type=="Ridge")
+{
+  lm_obj<-lm.ridge(dat[,1]~matrix(cbind(rep(1,ncol(dat)-1),dat[,2:ncol(dat)]),ncol=ncol(dat))-1, lambda = lambda_reg)
+}
+if (reg_type=="Lasso")
+{
+  lm_obj<-glmnet(x=dat[,2:ncol(dat)],y=dat[,1],alpha=0,lambda = lambda_reg)
+}
 optimal_weights<-lm_obj$coef

### 2025-05-23 Sweave/generate_sweave.R

@@ -11,7 +11,6 @@ path.out<-paste(path.main,"/Latex/",sep="")
 path.sweave<-paste(path.main,"/Sweave/",sep="")
 #------------------------------
 
-recompute_results<-
 
 paper<-"BIP_predictor"
 

### 2025-05-26 Latex/BIP_predictor.tex
up

--- 2025-05-21 Latex/BIP_predictor.tex
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 10:24:26 2025
+% Mon May 26 07:26:22 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 10:24:26 2025
+% Mon May 26 07:26:22 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 10:24:28 2025
+% Mon May 26 07:26:23 2025
 \begin{table}[ht]
@@
 The two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 10:24:28 2025
+% Mon May 26 07:26:35 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 10:24:32 2025
+% Mon May 26 07:27:26 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 10:24:33 2025
+% Mon May 26 07:27:36 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 0.086 & 0.049 & 0.018 & 0.005 & 0.002 & 0.002 & 0.042 \\ 
-  Shift=1 & 0.238 & 0.165 & 0.097 & 0.033 & 0.005 & 0.001 & 0.005 \\ 
-  Shift=2 & 0.815 & 0.680 & 0.428 & 0.180 & 0.029 & 0.007 & 0.008 \\ 
-  Shift=3 & 0.866 & 0.963 & 0.922 & 0.699 & 0.269 & 0.079 & 0.047 \\ 
-  Shift=4 & 0.361 & 0.808 & 0.989 & 0.954 & 0.546 & 0.127 & 0.034 \\ 
-  Shift=5 & 0.351 & 0.786 & 0.982 & 0.943 & 0.579 & 0.245 & 0.042 \\ 
+Shift=0 & 0.084 & 0.047 & 0.017 & 0.005 & 0.002 & 0.001 & 0.038 \\ 
+  Shift=1 & 0.230 & 0.160 & 0.094 & 0.032 & 0.005 & 0.001 & 0.004 \\ 
+  Shift=2 & 0.801 & 0.666 & 0.414 & 0.173 & 0.026 & 0.005 & 0.006 \\ 
+  Shift=3 & 0.876 & 0.961 & 0.917 & 0.683 & 0.248 & 0.068 & 0.039 \\ 
+  Shift=4 & 0.370 & 0.819 & 0.990 & 0.953 & 0.533 & 0.117 & 0.030 \\ 
+  Shift=5 & 0.353 & 0.802 & 0.983 & 0.943 & 0.571 & 0.239 & 0.043 \\ 
    \hline
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 10:24:33 2025
+% Mon May 26 07:27:36 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 0.950 & 0.917 & 0.877 & 0.851 & 0.865 & 0.910 & 0.960 \\ 
-  Shift=1 & 1.028 & 1.007 & 0.960 & 0.899 & 0.867 & 0.878 & 0.912 \\ 
-  Shift=2 & 1.040 & 1.043 & 1.030 & 0.984 & 0.933 & 0.910 & 0.913 \\ 
-  Shift=3 & 1.006 & 1.024 & 1.040 & 1.025 & 0.983 & 0.952 & 0.941 \\ 
-  Shift=4 & 0.977 & 1.003 & 1.034 & 1.040 & 1.005 & 0.964 & 0.937 \\ 
-  Shift=5 & 0.984 & 1.011 & 1.044 & 1.057 & 1.031 & 0.986 & 0.949 \\ 
+Shift=0 & 0.973 & 0.937 & 0.888 & 0.850 & 0.855 & 0.900 & 0.956 \\ 
+  Shift=1 & 1.073 & 1.051 & 0.997 & 0.918 & 0.870 & 0.873 & 0.906 \\ 
+  Shift=2 & 1.069 & 1.077 & 1.063 & 1.006 & 0.940 & 0.907 & 0.907 \\ 
+  Shift=3 & 1.020 & 1.045 & 1.066 & 1.048 & 0.997 & 0.957 & 0.941 \\ 
+  Shift=4 & 0.983 & 1.016 & 1.056 & 1.064 & 1.023 & 0.974 & 0.942 \\ 
+  Shift=5 & 0.988 & 1.026 & 1.071 & 1.088 & 1.057 & 1.007 & 0.963 \\ 
    \hline
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 10:24:36 2025
+% Mon May 26 07:28:11 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 10:24:36 2025
+% Mon May 26 07:28:11 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 1.104 & 1.066 & 1.019 & 0.989 & 1.006 & 1.058 & 1.116 \\ 
-  Shift=1 & 1.232 & 1.206 & 1.150 & 1.076 & 1.038 & 1.051 & 1.092 \\ 
-  Shift=2 & 1.146 & 1.150 & 1.136 & 1.085 & 1.028 & 1.002 & 1.006 \\ 
-  Shift=3 & 1.008 & 1.027 & 1.042 & 1.028 & 0.986 & 0.954 & 0.943 \\ 
-  Shift=4 & 0.988 & 1.014 & 1.045 & 1.051 & 1.016 & 0.975 & 0.948 \\ 
-  Shift=5 & 0.967 & 0.995 & 1.027 & 1.039 & 1.013 & 0.970 & 0.933 \\ 
+Shift=0 & 1.128 & 1.086 & 1.030 & 0.985 & 0.992 & 1.044 & 1.109 \\ 
+  Shift=1 & 1.280 & 1.254 & 1.189 & 1.096 & 1.037 & 1.041 & 1.081 \\ 
+  Shift=2 & 1.173 & 1.181 & 1.166 & 1.104 & 1.031 & 0.995 & 0.995 \\ 
+  Shift=3 & 1.012 & 1.037 & 1.058 & 1.040 & 0.989 & 0.949 & 0.934 \\ 
+  Shift=4 & 0.982 & 1.016 & 1.056 & 1.063 & 1.022 & 0.974 & 0.941 \\ 
+  Shift=5 & 0.961 & 0.997 & 1.041 & 1.058 & 1.028 & 0.979 & 0.936 \\ 
    \hline
@@
   \hline
-RRMSE & 0.860 & 0.835 & 0.907 & 0.998 & 0.989 & 1.017 \\ 
+RRMSE & 0.862 & 0.838 & 0.912 & 1.008 & 1.001 & 1.028 \\ 
    \hline
@@
   \hline
-Shift=0 & 0.142 & 0.074 & 0.023 & 0.003 & 0.001 & 0.002 & 0.013 \\ 
-  Shift=1 & 0.475 & 0.344 & 0.217 & 0.111 & 0.044 & 0.023 & 0.044 \\ 
-  Shift=2 & 0.724 & 0.564 & 0.325 & 0.101 & 0.023 & 0.010 & 0.012 \\ 
-  Shift=3 & 0.943 & 0.979 & 0.933 & 0.670 & 0.234 & 0.066 & 0.031 \\ 
-  Shift=4 & 0.322 & 0.666 & 0.960 & 0.946 & 0.644 & 0.263 & 0.119 \\ 
-  Shift=5 & 0.386 & 0.721 & 0.940 & 0.910 & 0.591 & 0.248 & 0.060 \\ 
+Shift=0 & 0.127 & 0.064 & 0.019 & 0.003 & 0.001 & 0.002 & 0.012 \\ 
+  Shift=1 & 0.439 & 0.314 & 0.194 & 0.093 & 0.033 & 0.017 & 0.036 \\ 
+  Shift=2 & 0.718 & 0.565 & 0.326 & 0.098 & 0.020 & 0.008 & 0.010 \\ 
+  Shift=3 & 0.946 & 0.975 & 0.925 & 0.657 & 0.222 & 0.059 & 0.028 \\ 
+  Shift=4 & 0.327 & 0.699 & 0.964 & 0.942 & 0.619 & 0.237 & 0.102 \\ 
+  Shift=5 & 0.383 & 0.755 & 0.946 & 0.910 & 0.581 & 0.241 & 0.055 \\ 
    \hline
@@
   \hline
-Shift=0 & 0.991 & 0.974 & 0.952 & 0.933 & 0.929 & 0.942 & 0.965 \\ 
-  Shift=1 & 1.037 & 1.030 & 1.012 & 0.985 & 0.966 & 0.963 & 0.969 \\ 
-  Shift=2 & 1.014 & 1.014 & 1.005 & 0.983 & 0.959 & 0.950 & 0.953 \\ 
-  Shift=3 & 1.010 & 1.017 & 1.021 & 1.011 & 0.988 & 0.969 & 0.961 \\ 
-  Shift=4 & 0.987 & 0.999 & 1.015 & 1.020 & 1.008 & 0.992 & 0.981 \\ 
-  Shift=5 & 0.996 & 1.008 & 1.023 & 1.030 & 1.018 & 0.997 & 0.976 \\ 
+Shift=0 & 1.005 & 0.986 & 0.960 & 0.935 & 0.927 & 0.939 & 0.964 \\ 
+  Shift=1 & 1.060 & 1.054 & 1.032 & 0.998 & 0.973 & 0.965 & 0.970 \\ 
+  Shift=2 & 1.028 & 1.029 & 1.020 & 0.993 & 0.963 & 0.950 & 0.952 \\ 
+  Shift=3 & 1.018 & 1.027 & 1.034 & 1.022 & 0.994 & 0.972 & 0.961 \\ 
+  Shift=4 & 0.990 & 1.006 & 1.025 & 1.032 & 1.017 & 0.998 & 0.985 \\ 
+  Shift=5 & 1.000 & 1.016 & 1.037 & 1.046 & 1.032 & 1.008 & 0.985 \\ 
    \hline
@@
   \hline
-Shift=0 & 1.228 & 1.207 & 1.180 & 1.156 & 1.151 & 1.167 & 1.196 \\ 
-  Shift=1 & 0.978 & 0.972 & 0.955 & 0.929 & 0.912 & 0.908 & 0.915 \\ 
-  Shift=2 & 1.030 & 1.029 & 1.021 & 0.998 & 0.974 & 0.964 & 0.968 \\ 
-  Shift=3 & 1.037 & 1.044 & 1.048 & 1.038 & 1.014 & 0.995 & 0.986 \\ 
-  Shift=4 & 0.973 & 0.985 & 1.000 & 1.006 & 0.994 & 0.978 & 0.966 \\ 
-  Shift=5 & 0.981 & 0.993 & 1.008 & 1.015 & 1.003 & 0.982 & 0.962 \\ 
+Shift=0 & 1.237 & 1.215 & 1.183 & 1.152 & 1.141 & 1.156 & 1.187 \\ 
+  Shift=1 & 0.986 & 0.980 & 0.960 & 0.929 & 0.905 & 0.898 & 0.903 \\ 
+  Shift=2 & 1.036 & 1.037 & 1.029 & 1.001 & 0.971 & 0.957 & 0.960 \\ 
+  Shift=3 & 1.042 & 1.051 & 1.058 & 1.045 & 1.018 & 0.995 & 0.983 \\ 
+  Shift=4 & 0.970 & 0.985 & 1.004 & 1.010 & 0.996 & 0.978 & 0.965 \\ 
+  Shift=5 & 0.970 & 0.986 & 1.007 & 1.015 & 1.002 & 0.978 & 0.956 \\ 
    \hline
@@
   \hline
-RRMSE & 0.807 & 1.060 & 0.985 & 0.974 & 1.015 & 1.015 \\ 
+RRMSE & 0.812 & 1.075 & 0.992 & 0.977 & 1.021 & 1.031 \\ 
    \hline

### 2025-05-26 Latex/BIP_predictor.tex
up

@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 07:26:22 2025
+% Mon May 26 11:43:54 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 07:26:22 2025
+% Mon May 26 11:43:54 2025
 \begin{table}[ht]
@@
   \hline
-2024-05-01 &  & 93.400 & 84.900 & 90.900 & -1.400 \\ 
-  2024-06-01 &  & 90.700 & 83.700 & 89.800 & -1.400 \\ 
-  2024-07-01 & 902.571 & 93.200 & 84.100 & 90.700 & -1.200 \\ 
-  2024-08-01 &  & 91.200 & 83.800 & 89.300 & -0.900 \\ 
-  2024-09-01 &  & 90.800 & 82.800 & 86.900 & -0.800 \\ 
-  2024-10-01 & 900.761 & 92.200 & 82.500 & 88.100 & -0.200 \\ 
+2024-08-01 &  & 93.400 & 84.900 & 90.900 & -1.400 \\ 
+  2024-09-01 &  & 90.700 & 83.700 & 89.800 & -1.400 \\ 
+  2024-10-01 & 902.571 & 93.200 & 84.100 & 90.700 & -1.200 \\ 
+  2024-11-01 &  & 91.200 & 83.800 & 89.300 & -0.900 \\ 
+  2024-12-01 &  & 90.800 & 82.800 & 86.900 & -0.800 \\ 
+  2025-01-01 & 900.761 & 92.200 & 82.500 & 88.100 & -0.200 \\ 
    \hline
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 07:26:23 2025
+% Mon May 26 11:43:56 2025
 \begin{table}[ht]
@@
 The dashed vertical line delimits in- and out-of-sample spans.
-The two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 07:26:35 2025
+The two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=4in, width=5.5in]{./Figures/mssa_msse_zc.pdf}\caption{Sign changes of the M-MSE (green) and M-SSA (blue) nowcasts are examined in relation to the financial crisis, with sign changes of the M-MSE marked by green vertical lines.
+The sample holding time (HT) of each filter corresponds to the average duration between its consecutive zero-crossings.
+For each filter, its sample HT converges to the respective expected HT asymptotically, given sufficiently long samples.
+In this analysis, we impose a $50\%$ larger expected HT for M-SSA $HT_{MSSA}:=1.5×HT_{MSE}$.
+Consequently, in the long run, we expect the M-MSE to exhibit approximately $50\%$ more sign changes than the M-SSA.\label{mssa_msse_zc}}\end{center}\end{figure}
+
+% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Mon May 26 11:43:56 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 07:27:26 2025
+% Mon May 26 11:43:59 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 07:27:36 2025
+% Mon May 26 11:43:59 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 07:27:36 2025
+% Mon May 26 11:43:59 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 07:28:11 2025
+% Mon May 26 11:44:02 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 07:28:11 2025
+% Mon May 26 11:44:02 2025
 \begin{table}[ht]

### 2025-06-02 Latex/BIP_predictor.tex
up

@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 11:43:54 2025
+% Thu May 29 23:26:52 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 11:43:54 2025
+% Thu May 29 23:26:52 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 11:43:56 2025
+% Thu May 29 23:26:53 2025
 \begin{table}[ht]
@@
 For each filter, its sample HT converges to the respective expected HT asymptotically, given sufficiently long samples.
-In this analysis, we impose a $50\%$ larger expected HT for M-SSA $HT_{MSSA}:=1.5×HT_{MSE}$.
-Consequently, in the long run, we expect the M-MSE to exhibit approximately $50\%$ more sign changes than the M-SSA.\label{mssa_msse_zc}}\end{center}\end{figure}
+Here, we impose  $HT_{MSSA}=1.5\times HT_{MSE}$: in the long run, we expect the M-MSE to exhibit approximately $50\%$ more sign changes than the M-SSA.\label{mssa_msse_zc}}\end{center}\end{figure}
 
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 11:43:56 2025
+% Thu May 29 23:26:54 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 11:43:59 2025
+% Thu May 29 23:26:56 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 11:43:59 2025
+% Thu May 29 23:26:56 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 11:44:02 2025
+% Thu May 29 23:26:59 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 11:44:02 2025
+% Thu May 29 23:26:59 2025
 \begin{table}[ht]

### 2025-06-02 Sweave/BIP_predictor.Rnw
up

@@
 # Section 2.1: data
-recompute_results<-T
+recompute_results<-F
 
@@
 x_mat<-data[,select_vec_multi]
+tail(x_mat)
 rownames(x_mat)<-rownames(data)
@@
 colnames(data_table)[1]<-"GDP"
-
-tail(data_table)
+tail(data_table,20)
 
@@
 VAR_type="BVAR"
-#VAR_type="VAR"
+VAR_type="VAR"
 if (VAR_type=="BVAR")
@@
 # Use OLS or WLS in regression of shifted BIP on M-SSA component predictor
-# Default (better): use_garch<-F
+# Default (better): use_garch<-F (otherwise important crises are underweighted)
 use_garch<-F
@@
 acf_mat<-x_mat
+tail(x_mat)
 
@@
 colnames(mat2)<-colnames(mat1)
+rownames(mat2)<-as.character(index(data_table))[(nrow(data_table)-nrow(mat2)+1):nrow(data_table)]
 #latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
@@
 f_vec_unfiltered<-p_val_unfiltered<-MSE_unfiltered_vec<-NULL
-for (h in 0:3)
+for (h in 0:3)#h<-0
 {
@@
 hp_two<-c(HP_obj$hp_mse[L:2],HP_obj$hp_mse)
-ts.plot(hp_two,main=paste("Two-sided HP(",lambda_HP,")",sep=""),xlab="",ylab="")
+plot(hp_two,main=paste("Two-sided HP(",lambda_HP,")",sep=""),axes=F,type="l",xlab="Lags",ylab="")
+abline(h=0)
+axis(1,at=c(1,11,21,L,L-1+10,L-1+20,length(hp_two)),labels=c(1,11,21,L,L+10,L+20,length(hp_two))-L)
+axis(2)
+box()
+
 hp_c<-HP_obj$hp_trend
-ts.plot(hp_c,main="One-sided (concurrent) HP-C",xlab="",ylab="")
+plot(hp_c,main=paste("Two-sided HP(",lambda_HP,")",sep=""),axes=F,type="l",xlab="Lags",ylab="")
+abline(h=0)
+axis(1,at=c(1,11,21,L),labels=-1+c(1,11,21,L))
+axis(2)
+box()
 
@@
 abline(v=which(rownames(mplot)>date_to_fit)[1],lty=2,lwd=2)
+#abline(v=1+which(sign(mplot[2:nrow(mplot),3])!=sign(mplot[1:(nrow(mplot)-1),3])),col=colo[3])
+#abline(v=1+which(sign(mplot[2:nrow(mplot),2])!=sign(mplot[1:(nrow(mplot)-1),2])),col=colo[2])
 axis(1,at=c(1,4*1:(nrow(mplot)/4)),labels=rownames(mplot)[c(1,4*1:(nrow(mplot)/4))])
@@
 
+# Table with empirical HT and target correlations
 mat_cor_ht<-rbind(c(cor(na.exclude(mplot[,c("Target: HP-GDP","M-SSA")]))[1,2],
@@
 # Keep M-MSE and M-SSA only
+
+# Plot emphasizing zero-crossings towaards financial crisis
+file = "./Figures/mssa_msse_zc.pdf"
+pdf(file = paste(path.out,file,sep=""), paper = "special", width = 7, height = 7)
+
+hp_c_out<-filter(x_mat[,select],hp_c,side=1)
+par(mfrow=c(1,1))
+#mplot<-scale(cbind(target_shifted_mat,hp_c_out,mssa[,select],mssa_HT_HP_C[,select],mmse[,select]))
+#dim(mplot)
+#colnames(mplot)<-c("Target: HP-BIP","HP-C","M-SSA smaller HT","M-SSA larger HT","M-MSE")
+mplot<-scale(cbind(mssa[,select],mmse[,select]))
+dim(mplot)
+colnames(mplot)<-c("M-SSA","M-MSE")
+
+rownames(mplot)<-rownames(x_mat)
+
+mplot<-mplot[L:nrow(mplot),]
+mplot<-mplot[which(rownames(mplot)>2006&rownames(mplot)<2013),]
+
+colo<-c("black","violet","blue","orange","green")
+colo<-c("blue","green")
+main_title<-"Zero-Crossings of M-SSA and M-MSE towards the financial crisis"
+plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],lwd=1,ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
+mtext(colnames(mplot)[1],col=colo[1],line=-1)
+for (i in 1:ncol(mplot))
+{
+  lines(mplot[,i],col=colo[i],lwd=1,lty=1)
+  mtext(colnames(mplot)[i],col=colo[i],line=-i)
+}
+abline(h=0)
+abline(v=1+which(sign(mplot[2:nrow(mplot),1])!=sign(mplot[1:(nrow(mplot)-1),1])),col=colo[1],lty=2)
+abline(v=1+which(sign(mplot[2:nrow(mplot),2])!=sign(mplot[1:(nrow(mplot)-1),2])),col=colo[2],lty=2)
+axis(1,at=c(1,4*1:(nrow(mplot)/4)),labels=rownames(mplot)[c(1,4*1:(nrow(mplot)/4))])
+axis(2)
+box()
+
+dev.off()
 
@@
 @
+<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
+file = "./Figures/mssa_msse_zc.pdf"
+cat("\\begin{figure}[H]")
+cat("\\begin{center}")
+cat("\\includegraphics[height=4in, width=5.5in]{", file, "}\n",sep = "")
+cat("\\caption{Sign changes of the M-MSE (green) and M-SSA (blue) nowcasts are examined in relation to the financial crisis, with sign changes of the M-MSE marked by green vertical lines.
+The sample holding time (HT) of each filter corresponds to the average duration between its consecutive zero-crossings.
+For each filter, its sample HT converges to the respective expected HT asymptotically, given sufficiently long samples.
+Here, we impose  $HT_{MSSA}=1.5\\times HT_{MSE}$: in the long run, we expect the M-MSE to exhibit approximately $50\\%$ more sign changes than the M-SSA.", sep = "")
+cat("\\label{mssa_msse_zc}}", sep = "")
+cat("\\end{center}")
+cat("\\end{figure}")
+@
+
 <<label=ats_mba_2,echo=FALSE,results=tex>>=
@@
 # Can define type of regression
-reg_type<-"OLS"
 reg_type<-"LASSO"
 reg_type<-"Ridge"
+reg_type<-"OLS"
 # Penalty weight in regularization
@@
 #colnames(mplot)[2:ncol(track_weights)]<-paste("Weight of M-SSA component ",colnames(track_weights)[2:ncol(track_weights)])
-colo<-c("black","blue","red")
+colo<-c("orange","blue","red")
 main_title<-paste("Intercept and regression weight over time: h=",h,", shift=",shift,sep="")

### 2025-06-04 Latex/BIP_predictor.tex
up

--- 2025-05-26 Latex/BIP_predictor.tex
@@
 \section{Data and Dependence}
-
-\subsection{Data}
-
-\begin{itemize}
-\item Choice of indicators: BIP, ip, ESI, ifo, spread
-\item Quarterly frequency (emphasize $h\geq 2$)
-\item Data transformations: Diff-log, standardization (merely for ease of visual inspection), trim Covid ourliers
-\item Revisions and publication lags:  BIP lagged by quarter and ip lagged by two months in Jan-2025 data set
-\item Real-time data: all series are artificially aligned at end point. BIP and ip are advanced by their lags.
-\end{itemize}
-
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 07:26:22 2025
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrr}
-  \hline
- & GDP & ip & ifo\_c & ESI & spr\_10y\_3m \\ 
-  \hline
-2024-07-01 & 902.571 & 90.700 & 85.800 & 92.500 & -1.300 \\ 
-  2024-08-01 &  & 93.200 & 84.900 & 90.900 & -1.400 \\ 
-  2024-09-01 &  & 91.200 & 83.700 & 89.800 & -1.400 \\ 
-  2024-10-01 & 900.761 & 90.800 & 84.100 & 90.700 & -1.200 \\ 
-  2024-11-01 &  & 92.200 & 83.800 & 89.300 & -0.900 \\ 
-  2024-12-01 &  &  & 82.800 & 86.900 & -0.800 \\ 
-  2025-01-01 &  &  & 82.500 & 88.100 & -0.200 \\ 
-   \hline
-\end{tabular}
-\caption{Ragged end.  } 
-\label{f_stat}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 07:26:22 2025
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrr}
-  \hline
- & GDP & ip & ifo\_c & ESI & spr\_10y\_3m \\ 
-  \hline
-2024-05-01 &  & 93.400 & 84.900 & 90.900 & -1.400 \\ 
-  2024-06-01 &  & 90.700 & 83.700 & 89.800 & -1.400 \\ 
-  2024-07-01 & 902.571 & 93.200 & 84.100 & 90.700 & -1.200 \\ 
-  2024-08-01 &  & 91.200 & 83.800 & 89.300 & -0.900 \\ 
-  2024-09-01 &  & 90.800 & 82.800 & 86.900 & -0.800 \\ 
-  2024-10-01 & 900.761 & 92.200 & 82.500 & 88.100 & -0.200 \\ 
-   \hline
-\end{tabular}
-\caption{Data aligned at sample end.} 
-\label{f_stat}
-\end{table}
-
-The transformed quarterly series are displayed in Fig.
-\ref{data}, while Fig.
-\ref{data_lags} offers a close-up view of the financial crisis and the pandemic.
-The latter figure illustrates that the real-time GDP and industrial production series are synchronized at dips and peaks (coincident), whereas the other indicators tend to lead in relative terms.
-This characteristic of the additional explanatory variables can be leveraged in a multivariate approach to enhance forecasting performance compared to univariate benchmarks.
-
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/Data.pdf}\caption{Data: standardized log-differences.
-Data trimmed to $\pm 3$ for ease of visual inspection.\label{data}}\end{center}\end{figure}
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/data_lags.pdf}\caption{Data: leads and lags during financial crisis (left) and Pandemic (right).
-Data trimmed to $\pm 3$ for ease of visual inspection.\label{data_lags}}\end{center}\end{figure}
-
-\subsection{Dependence}
-
-We analyze the dependence structure of the data and derive a simple VAR model to establish the multivariate filter.
-The sample cross-correlation function (CCF) is shown in Figs.
-\ref{CCF} (entire dataset) and \ref{CCF_wc} (data prior to the pandemic).
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/CCF.pdf}\caption{Sample cross-correlation function: GDP and ifo (left), GDP and ESI (right): full data set, including the pandemic.
-Correlations at positive lags signify that (real-time) GDP is lagging when referenced against the two (real-time) indicators.\label{CCF}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/CCF_wc.pdf}\caption{Sample cross-correlation function: GDP and ifo (left), GDP and ESI (right): without pandemic.
-Correlations at positive lags signify that (real-time) GDP is lagging when referenced against the two (real-time) indicators.\label{CCF_wc}}\end{center}\end{figure}The CCF confirms the leads and lags illustrated in Fig.
-\ref{data_lags}, although the strength of the dependence is influenced by the singular readings during the pandemic\footnote{We discard the corresponding data when finalizing the predictor.}.
-We can now fit a VARMA model, and according to standard diagnostic tests (not shown here), a simple VAR(1) specification aligns with the data.
-The resulting impulse responses of the model are presented in Figs.
-\ref{impulse} (full data set) and \eqref{impulse_wc} (prior the pandemic).
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/ma_inv_multi_ip.pdf}\caption{MA inversion of VAR(1): full data set.\label{impulse}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/ma_inv_multi_ip_wc.pdf}\caption{MA inversion of VAR(1): without Pandemic.\label{impulse_wc}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=6in]{./Figures/ma_inv_BIP.pdf}\caption{MA inversion of VAR(1) for GDP: with pandemic (left) and without pandemic (right).\label{ma_inv_BIP}}\end{center}\end{figure}
-
-The model summarizes key features of the data, where lagging series (GDP and industrial production, when accounting for publication lags) depend on  indicators leading in relative terms.
-In particular, the impulse response of the spread (rightmost panel in the figure) suggests the relevance of a univariate model for this specific leading indicator.
-
-\section{Direct Forecasts}\label{direct_forecast}
-
-\subsection{`Classic' Direct Forecast}\label{cdf}
-
-So-called direct forecasts are obtained by regressing the relevant indicators on forward-shifted BIP, accounting for the additional publication lag.
-For illustration of the concept, we here rely on all indicators and full sample information, excluding the pandemic.
-In-sample forecasts and the shifted BIP, which serves as the target, are presented in Figs.
-\ref{direct_wc} (full dataset without the pandemic) and \ref{direct_wc_financial_crisis} (financial crisis) for horizons up to three quarters ahead.
-\begin{figure}[H]\begin{center}\includegraphics[height=5in, width=6in]{./Figures/direct_wc_all.pdf}\caption{Direct forecasts (red lines) based on all indicators and full sample information, without the pandemic: GDP (black line) shifted by zero one, two and three quarters.\label{direct_wc}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=5in, width=6in]{./Figures/direct_wc_financial_crisis.pdf}\caption{Direct forecast  across the financial crisis:  regression based on all indicators and full sample information, without the pandemic.\label{direct_wc_financial_crisis}}\end{center}\end{figure}The results indicate that as $h$ increases, the peaks and dips of the direct forecasts become increasingly right-shifted (delayed) relative to the target.
-Additionally, the F-statistics precipitously decline from $F=3.21$, for $h=2$, to $F=0.28$, for $h=3$, see Table \eqref{f_stat}, suggesting that the predictor becomes statistically insignificant for larger forecast horizons exceeding two quarters (link to Heinisch and Scheufele).
-
-\subsection{Filter: HP(160)}
-
-As shown, the performance of direct forecasts sharply declines for larger forecast horizons.
-We conjecture that erratic short-term fluctuations -—unpredictable high-frequency noise -— overlay the data, thereby obscuring the effective 'signal' and making a direct regression more susceptible to overfitting.
-We therefore envisage to highlight the signal by applying a filter to dampen high-frequency noise.
-The so-called Hodrick-Prescott filter is a classic tool used in business cycle analysis: the filter is specified by a single smoothing parameter $\lambda$ and the value $\lambda=1600$ is specifically recommended for quarterly data.
-However, Phillips and Jin (2021) suggest that the HP(1600) filter removes relevant information due to excessive smoothing.
-In our context, this oversmoothing issue would be further exacerbated when considering forecast horizons shorter than a year -—an interval inconsistent with the mean duration of up to several years of business cycles, as highlighted by the HP(1600).
-Consequently, we here select a more adaptive HP(160) 'target' filter.
-While this choice may appear somewhat arbitrary, an analysis of the filter in the frequency domain confirms a more suitable profile for the corresponding amplitude function, assigning greater weight to yearly components -—relevant in our prediction framework-— than the classic quarterly HP filter.
-This approach also addresses and controls for undesirable high-frequency noise.
-See Fig.
-\ref{hp_160}, which presents both two-sided and (classic) one-sided filters (top panels) along with the associated amplitude functions (bottom panels).
-We now posit the HP(160) filter as a means to achieve better forecasting performance at larger horizons, noting that a comprehensive technical analysis of the effect of $\lambda$ on the resulting BIP predictor can be conducted using the M-SSA package.
-\begin{figure}[H]\begin{center}\includegraphics[height=5in, width=6in]{./Figures/hp_160.pdf}\caption{Two-sided HP-160) (top left) and one-sided (concurrent) HP-C filters (top right) with corresponding amplitude functions (bottom).\label{hp_160}}\end{center}\end{figure}
-\subsection{Direct HP Forecast}\label{hpdf}
-
-Extending the `classic' direct forecast design, we here consider filtered indicators, based on the one-sided HP(160), denoted as HP-C: the filtered indicators are explanatory variables for the regressions on forward-shifted BIP\footnote{Katja and Simon: for each forecast horizon $h=0,1,2,3$ I'm using corresponding forecasts of the real-time filter output, assuming the data to be white noise (assuming an ARMA model does not lead to significantly different results).
-I think we might ignore these technical details in the paper.}.
-As in the previous section, we rely on all indicators as well as the full sample estimate for illustration, see Figs.\eqref{direct_hp_forecasts} (all observations with exclusion of the pandemic) and \eqref{direct_hp_forecasts_financial_crisis} (financial crisis).
-A comparison with the classic direct forecasts in Fig.\eqref{direct_wc_financial_crisis} indicates that the new predictors are slightly less retarded (`faster').
-\begin{figure}[H]\begin{center}\includegraphics[height=5in, width=6in]{./Figures/direct_hp_forecasts.pdf}\caption{Direct HP forecasts: entire data set with exclusion of the pandemic.\label{direct_hp_forecasts}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=5in, width=6in]{./Figures/direct_hp_forecasts_financial_crisis.pdf}\caption{Shifted GDP (black), direct HP forecast (blue) and classic direct forecast (red) over the curse of the financial crisis:  regression based on all indicators and full sample information (without pandemic).\label{direct_hp_forecasts_financial_crisis}}\end{center}\end{figure}Moreover, the F-statistics are larger than for the direct forecasts, see Table \eqref{f_stat}, but the predictor remains statistically insignificant at forecast horizons exceeding two quarters (even though the evaluation is conducted in-sample and based on all available indicators).
-
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 07:26:23 2025
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrrrr}
-  \hline
- & h= 0 & h= 1 & h= 2 & h= 3 & h= 4 & h= 5 & h= 6 \\ 
-  \hline
-HP-C filtered & 0.000 & 0.001 & 0.036 & 0.046 & 0.049 & 0.012 & 0.022 \\ 
-  Unfiltered & 0.028 & 0.000 & 0.001 & 0.910 & 0.509 & 0.091 & 0.178 \\ 
-   \hline
-\end{tabular}
-\caption{Statistical significance of unfiltered and HP-C filtered forecasts based on HAC-adjusted Wald test: all indicators and full sample (without pandemic).  } 
-\label{f_stat}
-\end{table}
-We now propose a more refined causal filter design aiming for statistical significance of the resulting predictor at longer forecast horizons, out-of-sample.
-
-\section{Multivariate Causal Filter: M-SSA}
-
-The application of the one-sided HP-C filter to the indicators has demonstrated improvements in forecast performance.
-However, the univariate nature of this filter limits its ability to leverage information from the cross-section, as supported by leading indicators.
-Additionally, the noise suppression capability of the one-sided filter is compromised, as illustrated by high-frequency leakage of the amplitude function in Fig.
-\eqref{hp_160} and further elaborated upon by Wildi (2025).
-To address these limitations, we propose an extension of the Smooth Sign Accuracy (SSA) framework introduced by Wildi (2025).
-
-\subsection{Optimization criterion}
-
-Let $\mathbf{X}_t$ (of dimension $t\times n$) denote a set of $n$ explanatory series $\mathbf{x}_{1t},...,\mathbf{x}_{nt}$, with observations $\mathbf{x}_{it}=(x_{i1},...,x_{it})'$.
-Let $\mathbf{z}_t=(z_{1},...,z_t)$  denote a target series, which typically lies outside the linear space spanned by $\mathbf{X}_t$  (and is unknown at time $t$).
-For simplicity, we assume stationarity of all time series involved.
-In this framework, $\mathbf{X}_t$ corresponds to a matrix of selected economic indicators with $n=5$, and $z_{it}$   is the output of a two-sided HP(160) filter applied to the $i$-th indicator, $i=1,...,5$.
-Since $z_{it}$  depends on future observations $x_{it+1},x_{it+2},...$, the prediction task involves `tracking' $z_{it}$  using the estimate $y^i_{t}=\sum_{j=1}^ny^i_{jt}$, where $y^i_{jt}=\mathbf{b}^i_j~'\mathbf{x}_{jt}=\sum_{k=0}^{L} b^i_{jk}x_{jk}$ are the outputs of a multivariate (causal or one-sided) filter $\mathbf{B}^i$, with columns $\mathbf{b}^i_{j}$ of fixed length $L$, $j=1,...,5$, assigning weight to the last $L$ observations of the $j$-th indicator $\mathbf{x}_{jt}$ (the fixed-length assumption is merely used for ease of exposition).
-For the sake of clarity, we now omit the upper-case index  $i$ for all variables, assuming $z_t=z^{i_0}_{t}$ for some fixed $i_0$.% Additionally, we assume a fixed filter length $L$, such that $\mathbf{B}^i_t=\mathbf{B}_t=\mathbf{B}$, which has dimension $L\times 5$.
-The task of `tracking'  a target can be formalized in different ways; here, we focus on the target correlation $\rho(z,y,h)$ between $y_t$ and $z_{t+h}$, where $h\geq 0$ denotes the forecast horizon (backcasting with $h<0$ is ignored in this context).
-To streamline the notation and avoid unnecessary complexity, we assume a fixed horizon, say $h=h_0$, so that we may drop the reference to $h$ in our notation.
-In the univariate case ($n=1$), Wildi (2025) proposes the Simple Sign Accuracy (SSA) as an optimization criterion for this task:
-\begin{eqnarray}\label{critssa}
-\left.\begin{array}{c}\rho(z,y)\to\max\\
-\rho(y,1)=\rho_1\end{array}\right\}.
-\end{eqnarray}
-Objective function and constraint are expressed in terms of correlations, namely the target correlation $\rho(z,y)$ between target and predictor and the lag-one autocorrelation $\rho(y,1)$ of the predictor.
-The parameter $\rho_1$ governs the smoothness of the predictor: higher values of $\rho_1$ generally promote smoother trajectories of $y_t$.
-Formally, assuming that the process $x_t$ is centered (zero-mean), Wildi (2024) establishes a connection between $\rho(y,1)$ and the expected duration between consecutive zero-crossings (sign changes) of $y_t$, the so-called holding time, denoted by $ht(y)$:
-\begin{eqnarray*}\label{ht}
-ht(y)=\frac{\pi}{\arccos(\rho(y,1))}.
-\end{eqnarray*}
-Since the non-linear transformation involved is strictly monotonic, the SSA criterion can be reformulated as the optimization problem:
-\begin{eqnarray}\label{critssaht}
-\left.\begin{array}{c}\rho(z,y)\to\max\\
-\frac{\arccos(\rho(y,1))}{\pi}=1/ht_1\end{array}\right\},
-\end{eqnarray}
-where the smoothness parameter $1/ht_1$ expresses the rate of zero-crossings of the predictor\footnote{Formally, the exact relationship between the zero-crossing rate and the lag-one autocorrelation of the predictor assumes Gaussian time series, but Wildi (2024) demonstrates robustness to deviations from this assumption.}.
-\\
-
-We now briefly discuss some implications of criterion \eqref{critssaht}.
-First, the objective function is indifferent to an affine transformation of the predictor.
-This ambiguity can be resolved by assuming an arbitrary scale and level for $y_t$ (standardization).
-Alternatively, a mean-square error norm (MSE) may be substituted for the target correlation, as noted in Wildi (2025).
-In this case, the classic MSE predictor $y_{t,MSE}$ is obtained as a solution to the SSA criterion by insertion of $1/ht_1:=1/ht_{MSE}$ in the constraint, where $ht_{MSE}$ represents the holding time of $y_{t,MSE}$.
-%If the MSE predictor is noisy, which is often the case in applications, one can select $ht_1>ht_{MSE}$  in the constraint to address the rate of noisy crossings or false alarms.
-Second, the concept can be extended to a multivariate framework, denoted as M-SSA, by generalizing the correlation functions, as detailed in Wildi (2025b).
-Lastly, sign changes in trend growth, represented by zero-crossings of the target $z_t$, are indicative of relevant changes in the trajectory of the economy, for example at transitions between expansions and recessions.
-Unfortunately, classic predictors often generate excessively many false `noisy' crossings, due to high-frequency leakage as illustrated by the amplitude function in Fig.
-\eqref{hp_160} (bottom right panel) and further elaborated upon by Wildi (2025).
-The SSA criterion helps control this phenomenon.
-More precisely, Wildi (2025) derives an equivalent dual formulation of criterion \eqref{critssaht}, which distinguishes the SSA solution as exhibiting the smallest rate of zero-crossings of any (linear) predictor with the same target correlation.
-We now rely on the M-SSA criterion to derive a `smooth' predictor $y_t$ of $z_t$.
-%In a second stage, serving as a generalization and refinement of the univariate HP-C utilized in the direct HP forecast.
-
-\subsection{M-MSE and M-SSA Nowcasts}
-
-For illustration, we consider a nowcast of the target $z_t$ generated by the output of the two-sided HP(160) filter applied to BIP, which we denote by HP-BIP.
-We compare the classic multivariate mean squared error (MSE) predictor, denoted by M-MSE, with the M-SSA predictor, where we set $ht_1=1.5ht_{MSE}$ in the holding time constraint.
-This configuration ensures that the M-SSA predictor produces approximately 33$\%$ fewer zero-crossings than M-MSE over the long term.
-The purpose of this constraint is to compare the effect of a different level of smoothness on the prediction performance (target correlation).
-The estimation of the underlying VAR(1) model employed for the M-MSE and M-SSA is based on data observations up to January 2008.
-This approach ensures a lengthy out-of-sample period that encompasses significant events such as the financial crisis, the sovereign debt crisis, and the COVID-19 pandemic.
-Further details and background information are available in Wildi (2025b).
-All computations are performed utilizing the M-SSA package, as documented in Wildi (2025b).\\
-
-Figure \ref{bk_gammak} displays the coefficients of the multivariate nowcasts, where increased smoothing achieved through the M-SSA method manifests as a distinct pattern in the filter weights, notably influencing the rate of decay across increasing lags.
-The resulting filtered series, presented in Figure \ref{mssa_msse_now}, are standardized to enable direct comparisons.
-The M-SSA nowcast adheres to the specified constraint, exhibiting approximately one-third fewer zero-crossings than the M-MSE, confirming the effectiveness of the smoothness control imposed during the design.
-This is further corroborated by Table \ref{corhtnow}, which reports empirical holding times and target correlations.
-Increased smoothness achieved by the M-SSA comes at the cost of a slight reduction in tracking accuracy, as indicated by a marginally lower target correlation.
-This highlights the inherent trade-off—often referred to as the prediction dilemma—between smoothness (zero-crossing rate) and predictive accuracy (target correlation) within the M-SSA criterion.
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=5in]{./Figures/bk_gammak.pdf}\caption{Multivariate M-MSE (left) and M-SSA  (right) nowcast filters.\label{bk_gammak}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=4in, width=5.5in]{./Figures/mssa_msse_now.pdf}\caption{Two-sided HP(160) applied to GDP (target, black) and nowcasts:  M-SSA (blue) and  M-MSE (green).
-The dashed vertical line delimits in- and out-of-sample spans.
-The two-sided filter does not extend to the sample end.\label{mssa_msse_now}}\end{center}\end{figure}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 07:26:35 2025
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrr}
-  \hline
- & M-SSA & M-MSE \\ 
-  \hline
-Target correlation & 0.68 & 0.69 \\ 
-  HT & 6.50 & 3.71 \\ 
-   \hline
-\end{tabular}
-\caption{Sample target correlations and HTs of M-SSA and M-MSE nowcasts.  } 
-\label{corhtnow}
-\end{table}The results validate that the smoothing constraint incorporated into the multivariate extension of Criterion \eqref{critssa} effectively manages the zero-crossings of the predictor: sample estimates of the holding times (HTs) align with the expected values derived from Equation \eqref{ht}\footnote{Effective convergence of sample estimates to their theoretical expected values —assuming the true data-generating process is known— can be validated using the M-SSA package.
-This validation is typically performed through analyses of very long samples of simulated (multivariate) data, ensuring the consistency of the estimation procedures under idealized conditions.}.
-An increase of $50\%$ in the HT relative to the M-MSE predictor has only a modest effect on the target correlation, suggesting that this tradeoff could be beneficial in our application.
-Going forward, we now omit the M-MSE, as it is a special case of the more general M-SSA framework.
-
-\subsection{M-SSA Forecasts}
-
-The previously described nowcast example can be systematically extended to generate forecasts of the target variable.
-This extension utilizes the same empirical framework, incorporating an assumed $50\%$ increase in the HT over the M-MSE, and considers forward shifts of $h=0,...,6$  quarters of HP-BIP (note that M-SSA does not target BIP explicitly).
-%This approach facilitates the assessment of forecast performance across multiple horizons within the established methodological context.
-The influence of the forecast horizon on the M-SSA design is depicted in Fig.
-\eqref{bk_h}, which contrasts nowcast and one-year-ahead forecast: as $h$ increases (right panel), the scaling factor diminishes (zero-shrinkage), attributable to increased forecast uncertainty, and there is a relative amplification of the influence of the relevant indicators, leading BIP in relative terms.Additionally, a phase shift associated with the filter applied to BIP is observed (red line, right panel), reflecting the cyclical characteristics inherent to the HP filter.
-While this phase effect might cause an effective sign change in the BIP filter output as $h$ increases—something to be avoided in this context—the low-pass filters assigned to the additional 'leading' indicators (green and violet lines, right panel) continue to effectively track the low-frequency components of HP-BIP.
-The combination of the phase effect, applied to BIP, and level-tracking through these additional indicators enables more refined and effective tracking of the target compared to univariate forecasts, which rely solely on the phase-effect of the filter to look ahead of time.\\
-
-M-SSA predictors are compared to the traditional univariate HP-C in Figs.
-\eqref{multivar_vs_univar} and \eqref{mssa_hpc_financial_crisis}.
-In the first figure, we compare both filters across the periods of the financial crisis (top panels) and the COVID-19 pandemic (bottom panels), for forecast horizons $h=0,...,6$.
-The nowcast from the previous section is included for reference.
-In the second figure we compare both filters to the forward-shifted BIP for shifts larger than two quarters.
-All series are standardized to facilitate visual comparison.
-The primary distinction of the M-SSA approach is the progressive left-shift of the predictor as a function of $h$; this shift becomes more pronounced with increasing $h$ and appears to be consistent across all levels, including the peaks and troughs of the series.
-Conversely, the positions of peaks and troughs in the HP-C series appear to remain largely unaffected by variations in $h$.
-The more pronounced and systematic left-shift observed in the M-SSA forecasts can be partly attributed to the increasing weight assigned to the additional indicators leading BIP, which dominate the BIP component at larger forecast horizons.
-%In comparison, the effect of phase shifts—arising from the cyclical nature of the HP-filter and manifesting in the HP-C forecasts (shown in the left panels)—seems less systematic.
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=5.5in]{./Figures/bk_h.pdf}\caption{Multivariate M-SSA: $h=0$ (nowcast, left) and $h=4$ (one year forecast, right).\label{bk_h}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=4in, width=5.5in]{./Figures/multivar_vs_univar.pdf}\caption{Now- and forecasts of HP-GDP for $h=0,...,6$: HP-C (top) vs.
-M-SSA (bottom) across the financial crisis (left) and the pandemic (right).\label{multivar_vs_univar}}\end{center}\end{figure}
-\begin{figure}[H]\begin{center}\includegraphics[height=4in, width=5.5in]{./Figures/mssa_hpc_financial_crisis.pdf}\caption{Forward-shifted GDP (black), M-SSA (blue) and HP-C (red) for $h=3,4,5,6$ across the financial crisis.
-All series are standardized for better visual inspection.\label{mssa_hpc_financial_crisis}}\end{center}\end{figure}
-
-\section{M-SSA Component Predictor}
-
-The targets of the multivariate filters are the smoothed outputs obtained from the acausal two-sided HP(160) filter applied to the five indicators.
-However, our objective is to develop predictors that emphasize the unfiltered BIP.
-To achieve this, we can utilize the  M-SSA outputs, referred to as M-SSA components.
-
-\subsection{Standardized Predictor: Equal-Weighting}
-
-Assuming that all M-SSA components contribute equally to BIP prediction, an initial step involves averaging the standardized components to form an equally-weighted composite measure.
-Subsequently, the predictor is derived by regressing this standardized aggregate on the (forward-shifted) BIP series.
-Fig.
-\eqref{mssa_predictor} presents the BIP series alongside the equally-weighted M-SSA predictor.
-All series are standardized to facilitate visual comparison and to highlight the dynamic shifts, particularly the left-shift characteristic of the predictors.\\
-
-To enhance interpretability and assessment of the M-SSA predictor, we propose analyzing its individual components.
-For illustration, Fig.
-\eqref{mssa_predictor_interpretability} shows the nowcast at $h=0$.
-The solid blue line (the nowcast) approaches the zero line near the end of the sample, with a recent upward movement mainly supported by the leading spread component.
-In contrast, the ifo- and ESI-components remain near the zero line, while the ip- and BIP-components appear to be awaiting additional evidence and confirmation before signaling a definitive movement.
-This type of assessment can be useful for evaluating the relevance of recent changes in the forecasted time series dynamics.
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/mssa_predictor.pdf}\caption{BIP and equally-weighted M-SSA predictor: all series standardized.\label{mssa_predictor}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/mssa_predictor_interpretability.pdf}\caption{Equally-weighted M-SSA predictor and M-SSA components: all series standardized.\label{mssa_predictor_interpretability}}\end{center}\end{figure}
-%???Eventually add p-values and/rRMSEs???
-
-\subsection{Tracking BIP: Optimal Weighting}
-
-We now utilize the M-SSA components as regressors on forward-shifted BIP, extending the previous (standardized) equally-weighted M-SSA predictor to incorporate optimal weighting of the components.
-While we keep the M-SSA fixed, as based on data up to January 2008, we implement an expanding window starting from January 2007 for deriving the component weights, which are up-dated on a quarterly basis (this contrasts with Section \eqref{direct_forecast} which was based on the full sample information).
-\\
-
-There are various combinations of M-SSA components for tracking BIP.
-For illustration, we here select the single BIP component resulting from the M-SSA filter tracking HP-BIP.
-Table \eqref{p_val_hpbip_wc} reports HAC-adjusted p-values from regressions of the out-of-sample predictor on HP-BIP, indicating a strong link between the predictor and the target, out-of-sample.
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 07:27:26 2025
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrrrr}
-  \hline
- & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
-  \hline
-Shift=0 & 0.000 & 0.000 & 0.001 & 0.002 & 0.005 & 0.020 & 0.112 \\ 
-  Shift=1 & 0.000 & 0.000 & 0.000 & 0.001 & 0.002 & 0.005 & 0.019 \\ 
-  Shift=2 & 0.000 & 0.000 & 0.000 & 0.001 & 0.001 & 0.003 & 0.006 \\ 
-  Shift=3 & 0.018 & 0.007 & 0.002 & 0.000 & 0.002 & 0.003 & 0.004 \\ 
-  Shift=4 & 0.180 & 0.100 & 0.040 & 0.010 & 0.002 & 0.005 & 0.005 \\ 
-  Shift=5 & 0.579 & 0.432 & 0.261 & 0.108 & 0.027 & 0.008 & 0.009 \\ 
-   \hline
-\end{tabular}
-\caption{HAC adjusted p-values of regressions of BIP (M-SSA) component optimized for forecast horizons $h=0,...,6$  (the columns) on HP-BIP shifted forward by shift=0,...,5 (the rows).
-Out-of-sample span starting in Jan-2007, without pandemic.}
-\label{p_val_hpbip_wc}
-\end{table}Our choice—focusing on the single M-SSA BIP component—emphasizes simplicity and interpretability, as future BIP and HP-BIP are connected through their shared low-frequency component.
-Additionally, the M-SSA BIP component is the most important predictor in the regression of the M-SSA outputs on BIP.
-Finally, revisions due to quarterly updates of the regression equations are minimized by this predictor, and the estimates are more stable and robust over time\footnote{Other combinations can be experimented with using the M-SSA package.
-Regression weights of more complex designs, involving multiple components, tend to be more difficult to interpret due to the occurrence of negative weights and potential sign changes of the regressors over time, which may indicate overfitting or non-stationarity.} (see Section \eqref{revisions}).
-Note that information from the other  indicators is incorporated into the M-SSA BIP component through the multivariate filter design.\\
-
-This setup allows us to evaluate the out-of-sample performance of various predictor approaches, including: (i) the simple mean of BIP, (ii) the direct forecast method described in Section \eqref{cdf}, (iii) the direct HP forecasts from Section \eqref{hpdf}, based on the univariate HP-C, and (iv) the new M-SSA BIP component predictor.
-%For completeness, we also consider an M-MSE component predictor derived from outputs of the multivariate MSE filter.
-
-\subsection{Out-of-Sample Forecast Performances}\label{oosp}
-
-Tables \ref{p_val_wc} and \ref{rRMSE_mSSA_comp_mean_wc} present HAC-adjusted p-values and rRMSEs for the GDP component predictor. They are obtained from the simple regressions
-
-\begin{equation}
-    \textrm{GDP}_{t+shift}=\beta_0(shift,h)+\beta_1(shift,h) \cdot \textrm{M-SSA}^{\textrm{GDP}}_t(h) + \epsilon_t^{shift,h}
-    \label{eq:GDP_M-SSA}
-\end{equation}
-
-where $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$ is the out-of-sample M-SSA predictor for GDP (not HP-GDP.)
-%(in contrast to Table \ref{p_val_hpbip_wc}, which emphasized HP-GDP.) 
-The p-values test $H_0 : 0 = \beta_1(shift,h)$ and do not account for potential bias or mis-scaling of the GDP predictor.
-In contrast, the rRMSEs, evaluated against the mean benchmark, incorporate eventual bias and mis-scaling effects.
-Additional results involving other predictor benchmarks are provided in the appendix, based on data sets with and without the pandemic.
-\\
-
-The purposes of Equation \ref{eq:GDP_M-SSA} are twofold: first, the regression parameters enable affine transformations of the left-hand side variable, so that the standardized target $\textrm{GDP}_{t+shift}$ can be replaced by the original log-returns without affecting the p-values or rRMSEs reported in Tables \ref{p_val_wc} and \ref{rRMSE_mSSA_comp_mean_wc} (noting that the benchmark predictors, i.e., the direct forecasts and the arithmetic mean, similarly adjust to affine transformations); second, the static scale adjustment by $\beta_1(shift,h)$ compensates for the increasing zero-shrinkage of $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$, reflecting greater uncertainty  as the forecast horizon $h$ increases.
-This simple transformation allows $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$ to be used as a predictor for $\textrm{GDP}_{t+shift}$ even when $h\neq shift$.
-Such flexibility enables assessment of the dynamic left-shift of $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$, as illustrated in Figure \ref{{multivar_vs_univar}}, at various forward-shifts of $\textrm{GDP}_{t+shift}$.\\
-
-When excluding the pandemic, the direct forecasts outperform the mean forecast up to two quarters ahead, see Table \ref{rRMSE_mSSA_direct_mean_without_covid8}.\footnote{However, the pandemic period generally obscures the relationships between predictors and GDP.} However, the M-SSA component predictor, optimized for longer forecast horizons, outperforms the mean benchmark up to 6Q ahead.
-Additionally, the GDP component predictor optimized for large forecast horizons $h\geq 4$ surpass the direct forecasts for $shift>2$, see Table \ref{rRMSE_mSSA_comp_direct_without_covid7}.
-Statistical significance of the GDP component predictor, optimized for larger forecast horizons, can be established for at least up to one year ahead.
-
-%???check whether rMSE is decreasing towards sample end (because initial estimate towards 2007 is unreliable)???
-
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 07:27:36 2025
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrrrr}
-  \hline
- & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
-  \hline
-Shift=0 & 0.084 & 0.047 & 0.017 & 0.005 & 0.002 & 0.001 & 0.038 \\ 
-  Shift=1 & 0.230 & 0.160 & 0.094 & 0.032 & 0.005 & 0.001 & 0.004 \\ 
-  Shift=2 & 0.801 & 0.666 & 0.414 & 0.173 & 0.026 & 0.005 & 0.006 \\ 
-  Shift=3 & 0.876 & 0.961 & 0.917 & 0.683 & 0.248 & 0.068 & 0.039 \\ 
-  Shift=4 & 0.370 & 0.819 & 0.990 & 0.953 & 0.533 & 0.117 & 0.030 \\ 
-  Shift=5 & 0.353 & 0.802 & 0.983 & 0.943 & 0.571 & 0.239 & 0.043 \\ 
-   \hline
-\end{tabular}
-\caption{HAC adjusted p-values of regressions of M-SSA component predictors optimized for forecast horizons $h=0,...,6$  (the columns) on BIP shifted forward by shift=0,...,5 (the rows).
-Out-of-sample span starting in Jan-2007, without pandemic.}
-\label{p_val_wc}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 07:27:36 2025
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrrrr}
-  \hline
- & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
-  \hline
-Shift=0 & 0.973 & 0.937 & 0.888 & 0.850 & 0.855 & 0.900 & 0.956 \\ 
-  Shift=1 & 1.073 & 1.051 & 0.997 & 0.918 & 0.870 & 0.873 & 0.906 \\ 
-  Shift=2 & 1.069 & 1.077 & 1.063 & 1.006 & 0.940 & 0.907 & 0.907 \\ 
-  Shift=3 & 1.020 & 1.045 & 1.066 & 1.048 & 0.997 & 0.957 & 0.941 \\ 
-  Shift=4 & 0.983 & 1.016 & 1.056 & 1.064 & 1.023 & 0.974 & 0.942 \\ 
-  Shift=5 & 0.988 & 1.026 & 1.071 & 1.088 & 1.057 & 1.007 & 0.963 \\ 
-   \hline
-\end{tabular}
-\caption{rRMSEs of M-SSA components predictor benchmarked against the expanding mean of BIP. Out-of-sample span starting in Jan-2007, without pandemic.} 
-\label{rRMSE_mSSA_comp_mean_wc}
-\end{table}
-
-\section{Revisions}\label{revisions}
-
-Revisions refer to changes in the historical values of a time series resulting from updates with new information.
-Revisions of the M-SSA component predictor can arise from data updates, which we do not analyze further here\footnote{Except for BIP, the other indicators are not or only slightly revised.
-Moreover, the smoothing effect of filters mitigates the effect of revisions when compared to direct forecasts.}, or from updates to predictor parameters.
-The latter can be further separated into changes in VAR parameters and regression weights.
-
-\subsection{Revisions: Regression Equation Only}
-
-For illustration, we here focus on a forecast horizon $h=4$ for the BIP component predictor and a forward-shift $shift=4$ for BIP, noting that similar results would be obtained with other combinations of $h$ and $shift$.
-%???Check left-shift of MA-inversion of M-SSA and M-MSE
-%??? Check forecast excess for HP-C
-%??? Justification HP(160): replace HP(1600) by AR(2) with similar persistency but shorter cycle lengthi, i.e. $a_1=-2"persistency*acos(\theta, a_2=persistency$.
-Fig.\eqref{revisions1} compares the final and real-time predictors: as the sample size increases, the real-time predictor converges steadily to the final predictor and the rate of convergence is partially determined by the simplicity of our design, which avoids incorporating additional M-SSA components.
-As a confirmation, Fig.\eqref{revisions2} displays the evolution of the intercept and regression weights over time: with increasing sample size, these estimates seem to converge to fixed points, indicating both the stationarity of the process and the consistency of the estimates (note, however, the slightly elevated values of the regression weight assigned to the M-SSA component during the financial crisis and the pandemic, suggesting a stronger link between the target and the predictor).
-In contrast, more complex designs based on multiple M-SSA components can exhibit substantial fluctuations in the regression parameters, sometimes changing signs.
-This suggests instability and complicates interpretation.
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/revisions1.pdf}\caption{Revisions of the M-SSA GDP component predictor are shown, with the final predictor in blue and the real-time design in red.
-These revisions result from quarterly updates to the regression equations.
-The M-SSA filter itself remains fixed, based on data up to January 2008.\label{revisions1}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/revisions2.pdf}\caption{Revisions of intercept and regression weight.\label{revisions2}}\end{center}\end{figure}
-
-\subsection{Revisions: M-SSA Only}
-
-The second source of revisions in the M-SSA component predictor arises from updates to the VAR model within the M-SSA filter.
-We compare the original estimates, based on data up to January 2008, with the final estimates obtained using $h=4$ and $shift=4$ (similar findings apply to other values of forecast horizon and forward-shift).
-For consistency, we excluded the entire pandemic episode from the data.
-Fig.\eqref{up_dated_ma_inv_multi_ip} compares the MA-inversions (impulse responses) of the BIP equation derived from the 'old' and the updated VAR models: unlike the former, the latter incorporates ESI as an additional explanatory variable for BIP.
-Fig.\eqref{bk_2008_all} illustrates the resulting effect on the M-SSA filter targeting HP-BIP at forecast horizon $h=4$.
-The updated filter (right panel) assigns more weight to the additional 'leading' indicators relative to BIP.
-Finally, Fig.\eqref{revisions3} compares the M-SSA BIP predictors.
-To isolate the effect of the regression revision, the figure displays standardized series.
-The update to the VAR model influences the left-shift of the predictor (blue line), which appears slightly 'faster' than the original predictor (red line).
-This suggests that the out-of-sample performance of the M-SSA component predictor may be better than reported in Section \eqref{oosp}, which is based on the fixed and increasingly outdated version of M-SSA.
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=5in]{./Figures/up_dated_ma_inv_multi_ip.pdf}\caption{MA-inversion (impulse response) of VAR: data up to Jan-2008 (left) vs.
-entire data set  without pandemic (right).\label{up_dated_ma_inv_multi_ip}}\end{center}\end{figure}
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=5in]{./Figures/bk_2008_all.pdf}\caption{M-SSA filter tracking HP-GDP at forecast horizon $h=4$: data up to Jan-2008 (left) vs.
-entire data set without pandemic (right).\label{bk_2008_all}}\end{center}\end{figure}
-
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=5in]{./Figures/revisions3.pdf}\caption{Revisions resulting from updates to the VAR model in M-SSA optimized for $h=shift=4$.
-To isolate the effect of the regression revision, all series are standardized.
-The up-dated M-SSA (blue) is left-shifted at the zero-crossings and appears slightly faster.
-For consistency, the entire pandemic episode has been removed.\label{revisions3}}\end{center}\end{figure}
-
-\section{Hit Rate vs. False Alarm Rate}
-
-%As shown in Wildi (2025), the SSA criterion's objective function—the target correlation $\rho(y,z,h)$—can be linked to the sign accuracy of the predictor $SA(y_t):=\text{P}\left(\text{sign}(z_{t+h}) = \text{sign}(y_t)\right)$, i.e., the probability that the signs of the predictor $y_t$ and the target $z_{t+h}$ agree:
-%\begin{eqnarray}\label{sig_acc}
-%SA(y_t)=0.5+\frac{\arcsin(\rho(y,z,h))}{\pi}.
-%\end{eqnarray}
-%Due to the strict monotonicity of the arcsin function, maximizing the target correlation also maximizes the sign accuracy of the predictor.\footnote{Equation \ref{sig_acc} holds exactly for Gaussian processes.
-Due to the central limit theorem, the relationship holds approximately for low-pass filtered non-Gaussian series, see Wildi (2025).}
-\textbf{Remarks:}
-\begin{itemize}
-\item Alternative to MSE, target correlation, holding-time: hit-rate, false alarm. 
-\item We can link objective function (target correlation) with sign accuracy, see Wildi (2025). BUT: risk that reviewer asks for comparison with logit or whatever sign-forecast tool.
-\item Context ROC-curve
-\end{itemize}
-For the ROC curves, we consider predictors of GDP based on both original (unfiltered) and filtered indicators. For the direct forecast, we use the regression
-\[
-GDP_{t+shift}=\beta_0^{shift}+\sum_{i=1}^5\beta_i^{shift}x_{ti}+\epsilon_{t}^{shift,direct}=\hat{GDP}_{t+shift}^{direct}+\epsilon_{t}^{shift,direct},
-\]
-where $x_{ti}$ designates the original indicators. 
-For the filtered indicators, we consider similar regression equations:
-\[GDP_{t+shift}=\beta_0^{shift,h,filt}+\sum_{i=1}^5\beta_i^{shift,h,filt}y_{ti}^{h,filt}+\epsilon_{t}^{h,shift,filt}=\hat{GDP}_{t+shift}^{h,filt}+\epsilon_{t}^{h,shift,filt},\]
-where $y_{ti}^{h,filt}$ represents the $i$-th filtered indicator optimized at forecast horizon $h$, based on either HP-C, M-MSE, or M-SSA. \\
-\textbf{Remarks:}
-\begin{itemize}
-\item I'm computing all combinations of shift and h. 
-\item But we could emphasize the diagonal only (h=shift): simplifies notation/explanation.
-\item The table of AUCs (below) is based on diagonal only
-\end{itemize}
-For simplicity of exposition, we present results based on the full sample, excluding the pandemic period.
-Fig.\ref{ROC_GDP_shift_1_4} compares ROCs of the predictors for small and mid-term forecast horizons and Table \ref{auc} presents AUCs (area under the curve) for forecast horizons 1-5 quarters ahead.
-
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=5in]{./Figures/ROC_GDP_shift_1_4.pdf}\caption{Hit rate vs.
-false alarm rate of the predictors at small (one quarter ahead:left panel) and yearly forecast horizons (right panel).\label{ROC_GDP_shift_1_4}}\end{center}\end{figure}
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 07:28:11 2025
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrr}
-  \hline
- & Direct forecast & Direct HP forecast & M-MSE & M-SSA \\ 
-  \hline
-Forecast horizon 1 & 0.752 & 0.715 & 0.746 & 0.747 \\ 
-  Forecast horizon 2 & 0.697 & 0.663 & 0.729 & 0.720 \\ 
-  Forecast horizon 3 & 0.554 & 0.622 & 0.684 & 0.709 \\ 
-  Forecast horizon 4 & 0.486 & 0.624 & 0.680 & 0.708 \\ 
-  Forecast horizon 5 & 0.593 & 0.636 & 0.643 & 0.701 \\ 
-   \hline
-\end{tabular}
-\caption{AUCs of predictors against GDP at various forecast horizons.} 
-\label{p_val_wc5}
-\end{table}At forecast horizons less than or equal to two quarters, the direct forecast challenges the filters.
-However, for horizons $h>2$, the filters outperform the direct approach: the univariate HP-C design is surpassed by the classic multivariate MSE filter, which is in turn dominated by the M-SSA.
-These results confirm that controlling the rate of sign changes of the predictor, while maximizing its target correlation through M-SSA can enhance predictor performance in terms of sign accuracy and false alarm rate.
-
-\section{Summary and Conclusion}
-
-Classic direct forecasts of BIP can outperform simple benchmarks, such as the mean, up to two quarters ahead.
-However, this limited forecast horizon is often insufficient for certain applications.
-A key challenge with direct forecasts is that the relevant indicators tend to be noisy, which can mask the true signal and lead to overfitting.
-In particular, the predictor struggles to timely track dips (slowdowns, recessions) or peaks (recoveries, expansions).
-To address this issue, we apply a HP(160)-filter to smooth the data, with the smoothing parameter $\lambda=160$ reflecting a more adaptive design than the classic quarterly HP, in line with an intended forecast horizon of up to one year ahead.
-The resulting predictor tends to outperform the classic direct forecasts in terms of statistical significance, mainly due to a slight left-shift noticeable at sharp recession dips.
-However, despite the filtering, the predictor remains somewhat noisy, potentially producing excessive noisy zero-crossings at the transitions between recessions and expansions.
-This issue is partly due to noise leakage inherent in the one-sided HP filter.
-\\
-
-To address these problems, we propose a multivariate extension of SSA based on Wildi (2024).
-Unlike the univariate HP filter, the M-SSA can leverage cross-sectional information from indicators leading BIP in real time while controlling for the rate of zero-crossings of the predictor.
-Consequently, the multivariate filter outputs become increasingly left-shifted as the forecast horizon lengthens, allowing dips and peaks of the target series to be tracked more systematically, while reducing the occurrence of noisy zero-crossings.
-However, since M-SSA does not explicitly target BIP, an additional step is necessary to derive the final BIP predictor.
-We propose two approaches: a simple equally-weighted aggregate, assuming all M-SSA components are equally important for predicting BIP, and an optimally weighted predictor based on a regression of the M-SSA components on future BIP.\\
-
-For illustration, we consider a simple approach based on regressing the single M-SSA BIP output on future BIP, thereby ignoring the other filter outputs.
-This predictor is intuitively appealing because the future HP-BIP—i.e., the target of M-SSA—is the low-frequency component of the future BIP target.
-Consequently, a strong link between M-SSA and future HP-BIP also indicates a connection with BIP, even if the statistical significance is obscured by the noise inherent in BIP.
-Out-of-sample performance suggests that this predictor is statistically significant at horizons up to one year.
-Moreover, designs optimized for larger forecast horizons outperform the mean benchmark at all considered forecast horizons, and the predictor exceeds the performance of direct forecasts—based on unfiltered or HP filtered indicators—at horizons longer than two quarters.
-An analysis of revision errors arising from quarterly updates of the proposed M-SSA BIP component predictor indicates that the regression parameters, the VAR model, and the resulting M-SSA filters stabilize after an initial burn-in period.
-This stability enhances the explainability and interpretability of the predictor.\\
-
-In conclusion, the new predictor design integrates the traditional direct forecast approach with a novel multivariate filter to target BIP up to one year ahead.
-This multivariate approach leverages a set of economic indicators leading BIP in real time while controlling the smoothness of the predictor through the zero-crossing rate.
-
-\newpage
-
-\section{Appendix}
-
-\subsection{Performances: Without Pandemic}
-
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 07:28:11 2025
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrrrr}
-  \hline
- & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
-  \hline
-Shift=0 & 0.084 & 0.047 & 0.017 & 0.005 & 0.002 & 0.001 & 0.038 \\ 
-  Shift=1 & 0.230 & 0.160 & 0.094 & 0.032 & 0.005 & 0.001 & 0.004 \\ 
-  Shift=2 & 0.801 & 0.666 & 0.414 & 0.173 & 0.026 & 0.005 & 0.006 \\ 
-  Shift=3 & 0.876 & 0.961 & 0.917 & 0.683 & 0.248 & 0.068 & 0.039 \\ 
-  Shift=4 & 0.370 & 0.819 & 0.990 & 0.953 & 0.533 & 0.117 & 0.030 \\ 
-  Shift=5 & 0.353 & 0.802 & 0.983 & 0.943 & 0.571 & 0.239 & 0.043 \\ 
-   \hline
-\end{tabular}
-\caption{HAC adjusted p-values of regressions of M-SSA component predictors optimized for forecast horizons $h=0,...,6$  (the columns) on BIP shifted forward by shift=0,...,5 (the rows).
-Out-of-sample span starting in Jan-2007, without pandemic.}
-\label{p_val_wc5}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 07:28:11 2025
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrrrr}
-  \hline
- & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
-  \hline
-Shift=0 & 0.973 & 0.937 & 0.888 & 0.850 & 0.855 & 0.900 & 0.956 \\ 
-  Shift=1 & 1.073 & 1.051 & 0.997 & 0.918 & 0.870 & 0.873 & 0.906 \\ 
-  Shift=2 & 1.069 & 1.077 & 1.063 & 1.006 & 0.940 & 0.907 & 0.907 \\ 
-  Shift=3 & 1.020 & 1.045 & 1.066 & 1.048 & 0.997 & 0.957 & 0.941 \\ 
-  Shift=4 & 0.983 & 1.016 & 1.056 & 1.064 & 1.023 & 0.974 & 0.942 \\ 
-  Shift=5 & 0.988 & 1.026 & 1.071 & 1.088 & 1.057 & 1.007 & 0.963 \\ 
-   \hline
-\end{tabular}
-\caption{rRMSEs of M-SSA components predictor benchmarked against the expanding mean of BIP. Out-of-sample span starting in Jan-2007, without pandemic.} 
-\label{rRMSE_mSSA_comp_mean6}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 07:28:11 2025
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrrrr}
-  \hline
- & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
-  \hline
-Shift=0 & 1.128 & 1.086 & 1.030 & 0.985 & 0.992 & 1.044 & 1.109 \\ 
-  Shift=1 & 1.280 & 1.254 & 1.189 & 1.096 & 1.037 & 1.041 & 1.081 \\ 
-  Shift=2 & 1.173 & 1.181 & 1.166 & 1.104 & 1.031 & 0.995 & 0.995 \\ 
-  Shift=3 & 1.012 & 1.037 & 1.058 & 1.040 & 0.989 & 0.949 & 0.934 \\ 
-  Shift=4 & 0.982 & 1.016 & 1.056 & 1.063 & 1.022 & 0.974 & 0.941 \\ 
-  Shift=5 & 0.961 & 0.997 & 1.041 & 1.058 & 1.028 & 0.979 & 0.936 \\ 
-   \hline
-\end{tabular}
-\caption{rRMSEs of M-SSA components predictor benchmarked against the direct forecasts. Out-of-sample span starting in Jan-2007, without pandemic.} 
-\label{rRMSE_mSSA_comp_direct_without_covid7}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 07:28:11 2025
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrrr}
-  \hline
- & Shift=0 & Shift=1 & Shift=2 & Shift=3 & Shift=4 & Shift=5 \\ 
-  \hline
-RRMSE & 0.862 & 0.838 & 0.912 & 1.008 & 1.001 & 1.028 \\ 
-   \hline
-\end{tabular}
-\caption{rRMSEs direct forecasts benchmarked against the expanding mean. Out-of-sample span starting in Jan-2007, without pandemic.} 
-\label{rRMSE_mSSA_direct_mean_without_covid8}
-\end{table}
-
-\newpage
-\subsection{Performances: Including Singular Pandemic Data}
-
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 07:28:11 2025
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrrrr}
-  \hline
- & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
-  \hline
-Shift=0 & 0.127 & 0.064 & 0.019 & 0.003 & 0.001 & 0.002 & 0.012 \\ 
-  Shift=1 & 0.439 & 0.314 & 0.194 & 0.093 & 0.033 & 0.017 & 0.036 \\ 
-  Shift=2 & 0.718 & 0.565 & 0.326 & 0.098 & 0.020 & 0.008 & 0.010 \\ 
-  Shift=3 & 0.946 & 0.975 & 0.925 & 0.657 & 0.222 & 0.059 & 0.028 \\ 
-  Shift=4 & 0.327 & 0.699 & 0.964 & 0.942 & 0.619 & 0.237 & 0.102 \\ 
-  Shift=5 & 0.383 & 0.755 & 0.946 & 0.910 & 0.581 & 0.241 & 0.055 \\ 
-   \hline
-\end{tabular}
-\caption{HAC adjusted p-values of regressions of M-SSA component predictors optimized for forecast horizons $h=0,...,6$  (the columns) on BIP shifted forward by shift=0,...,5 (the rows).
-Out-of-sample span starting in Jan-2007, including the pandemic.}
-\label{p_val1}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 07:28:11 2025
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrrrr}
-  \hline
- & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
-  \hline
-Shift=0 & 1.005 & 0.986 & 0.960 & 0.935 & 0.927 & 0.939 & 0.964 \\ 
-  Shift=1 & 1.060 & 1.054 & 1.032 & 0.998 & 0.973 & 0.965 & 0.970 \\ 
-  Shift=2 & 1.028 & 1.029 & 1.020 & 0.993 & 0.963 & 0.950 & 0.952 \\ 
-  Shift=3 & 1.018 & 1.027 & 1.034 & 1.022 & 0.994 & 0.972 & 0.961 \\ 
-  Shift=4 & 0.990 & 1.006 & 1.025 & 1.032 & 1.017 & 0.998 & 0.985 \\ 
-  Shift=5 & 1.000 & 1.016 & 1.037 & 1.046 & 1.032 & 1.008 & 0.985 \\ 
-   \hline
-\end{tabular}
-\caption{rRMSEs of M-SSA components predictor benchmarked against the expanding mean of BIP. Out-of-sample span starting in Jan-2007, including the pandemic.} 
-\label{rRMSE_mSSA_comp_mean2}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 07:28:11 2025
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrrrr}
-  \hline
- & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
-  \hline
-Shift=0 & 1.237 & 1.215 & 1.183 & 1.152 & 1.141 & 1.156 & 1.187 \\ 
-  Shift=1 & 0.986 & 0.980 & 0.960 & 0.929 & 0.905 & 0.898 & 0.903 \\ 
-  Shift=2 & 1.036 & 1.037 & 1.029 & 1.001 & 0.971 & 0.957 & 0.960 \\ 
-  Shift=3 & 1.042 & 1.051 & 1.058 & 1.045 & 1.018 & 0.995 & 0.983 \\ 
-  Shift=4 & 0.970 & 0.985 & 1.004 & 1.010 & 0.996 & 0.978 & 0.965 \\ 
-  Shift=5 & 0.970 & 0.986 & 1.007 & 1.015 & 1.002 & 0.978 & 0.956 \\ 
-   \hline
-\end{tabular}
-\caption{rRMSEs of M-SSA components predictor benchmarked against the direct forecasts. Out-of-sample span starting in Jan-2007, including the pandemic.} 
-\label{rRMSE_mSSA_comp_direct3}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Mon May 26 07:28:11 2025
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrrr}
-  \hline
- & Shift=0 & Shift=1 & Shift=2 & Shift=3 & Shift=4 & Shift=5 \\ 
-  \hline
-RRMSE & 0.812 & 1.075 & 0.992 & 0.977 & 1.021 & 1.031 \\ 
-   \hline
-\end{tabular}
-\caption{rRMSEs direct forecasts benchmarked against the expanding mean. Out-of-sample span starting in Jan-2007, including the pandemic.} 
-\label{rRMSE_mSSA_direct_mean4}
-\end{table}
-
-\end{document}

### 2025-06-04 Sweave/BIP_predictor.Rnw
up

@@
 library(pROC)
+# Elastic net for estimating VAR
+library(elasticnet)
 
@@
 source(paste(getwd(),"/R/ROCplots.r",sep=""))
+# Elastic net functions for VAR estimation
+source(paste(getwd(),"/R/elastic_net_functions.r",sep=""))
 
@@
 date_to_fit<-"2008"
+# Filter length
+L<-31
+
 # We can select type of VAR model
@@
 VAR_type="VAR"
+VAR_type="elastic_net"
 if (VAR_type=="BVAR")
@@
 }
-
-# Filter length
-L<-31
+if (VAR_type=="elastic_net")
+{
+# Marc's settings for VAR  
+  p<-2
+  q<-0
+# Length of MA inversion (impulse response function)  
+  nirf<-L
+# Number of non-zero coefficients in VAR for each dimension  
+  nsteps<-5
+}
+
 # Publication lag: one quarter for BIP
@@
   }
+  if (VAR_type=="elastic_net")
+  {
+    V_obj<-enetVAR(data_fit, p, nIRF, nsteps  )
+
+    save(V_obj,file=paste(path.result,"/en_VAR",sep=""))
+  }
+
 # Without Pandemic  
@@
   } 
+  if (VAR_type=="elastic_net")
+  {
+    V_obj_wc<-enetVAR(data_fit, p, nIRF, nsteps  )
+
+    save(V_obj_wc,file=paste(path.result,"/en_VAR_wc",sep=""))
+  }
 
@@
     load(file=paste(path.result,"/VAR_wc",sep=""))
-  } else
+  } 
+  if (VAR_type=="BVAR")
   {
@@
     load(file=paste(path.result,"/BVAR_wc",sep=""))
+  }
+  if (VAR_type=="elastic_net")
+  {
+    load(file=paste(path.result,"/en_VAR",sep=""))
+    load(file=paste(path.result,"/en_VAR_wc",sep=""))
   }
@@
 #   -The function computes M-SSA for each forecast horizon h in h_vec
-  mssa_indicator_obj<-compute_mssa_BIP_predictors_func(x_mat,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi,VAR_type,lambda_BVAR)
+  mssa_indicator_obj<-compute_mssa_BIP_predictors_func(x_mat,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi,nsteps,VAR_type,lambda_BVAR)
 
@@
 
-  mssa_indicator_HP_HT_obj<-compute_mssa_BIP_predictors_func(x_mat,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec_HP_C,h_vec,f_excess,lag_vec,select_vec_multi,VAR_type,lambda_BVAR)
+  mssa_indicator_HP_HT_obj<-compute_mssa_BIP_predictors_func(x_mat,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec_HP_C,h_vec,f_excess,lag_vec,select_vec_multi,nsteps,VAR_type,lambda_BVAR)
   if (VAR_type=="VAR")
@@
 #   -The function computes M-SSA for each forecast horizon h in h_vec
-  mssa_indicator_obj<-compute_mssa_BIP_predictors_func(x_mat,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi,VAR_type,lambda_BVAR)
+  mssa_indicator_obj<-compute_mssa_BIP_predictors_func(x_mat,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi,nsteps,VAR_type,lambda_BVAR)
 # Without pandemic  
-  mssa_indicator_obj_wc<-compute_mssa_BIP_predictors_func(x_mat_wc,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi,VAR_type,lambda_BVAR)
+  mssa_indicator_obj_wc<-compute_mssa_BIP_predictors_func(x_mat_wc,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi,nsteps,VAR_type,lambda_BVAR)
   if (VAR_type=="VAR")
@@
 {
-  final_mssa_indicator_obj<-compute_mssa_BIP_predictors_func(x_mat_wc,lambda_HP,L,date_to_fit_all,p_full,q_full,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi,VAR_type,lambda_BVAR)
+  final_mssa_indicator_obj<-compute_mssa_BIP_predictors_func(x_mat_wc,lambda_HP,L,date_to_fit_all,p_full,q_full,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi,nsteps,VAR_type,lambda_BVAR)
 

### 2025-06-04 R/M_SSA_utility_functions.R

@@ -185,7 +185,7 @@ filter_func<-function(x_mat,bk_x_mat,gammak_x_mse,gamma_target,symmetric_target,
 # This function operationalizes the M-SSA concept for predicting quarterly (German) GDP
 # It relies on hyperparameters specifying the design: lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess
 # It returns M-SSA and M-MSE predictors as well as forward-shifted HP-BIP (two-sided HP applied to BIP)
-compute_mssa_BIP_predictors_func<-function(x_mat,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi,VAR_type="VAR",lambda_BVAR=NULL)
+compute_mssa_BIP_predictors_func<-function(x_mat,lambda_HP,L,date_to_fit,p,q,ht_mssa_vec,h_vec,f_excess,lag_vec,select_vec_multi,nsteps,VAR_type="VAR",lambda_BVAR=NULL)
 {
 # 1. Compute target
   n<-dim(x_mat)[2] 
@@ -228,6 +228,15 @@ compute_mssa_BIP_predictors_func<-function(x_mat,lambda_HP,L,date_to_fit,p,q,ht_
     
     
   }
+  if (VAR_type=="elastic_net")
+  {
+    V_obj<-enetVAR(data_fit, p,L, nsteps  )
+    Sigma<-V_obj$Sigma
+    Phi<-V_obj$Phi
+    Theta<-NULL
+    
+  }
+    
 #---------------------------------------
 # 3. MA inversion: M-SSA relies on MA-inversion of VAR
   

### 2025-06-04 R/elastic_net_functions.R

@@ -1,29 +1,4 @@
 
-# Created 25-05-2025 by Simon van Norden
-# This code just loads the BIP data set.
-# It is based on snippets from BIP_predictor.Rnw Section 1 written by Marc Wildi
-#
-GetBIPdata <- function(main_path = NULL, DataFrame = F) {
-  # if main_path omitted, it defaults to ~/GitHub/BIP-predictor
-  path.data <- ifelse( is.null(main_path),
-                       "C:/Users/svn/Documents/GitHub/BIP-predictor/Data/",
-                       main_path)
-  
-  #------------------------------
-  
-  # Load data
-  # Transformed indicators: differences, trimmed, standardized
-  load(file=paste0(path.data,"macro"))
-  
-  select_vec_multi<-c("BIP","ip","ifo_c","ESI","spr_10y_3m")
-  x_mat<-data[,select_vec_multi] 
-  rownames(x_mat)<-rownames(data)
-  n<-dim(x_mat)[2]
-  # Number of observations
-  len<-dim(x_mat)[1]
-  
-  return(x_mat)
-}
 
 
 mat_lag <- function(X, p) {
@@ -137,92 +112,3 @@ enetVAR <- function(X, p = 1, nIRF = 20, nsteps = NULL, mode = 'step',
 }
 
 
-library(elasticnet)
-
-
-main_path<-paste(getwd(),"/Data/",sep="")
-MyData <-  GetBIPdata(main_path)
-tail(MyData)
-
-
-p <- 4  # number of lags to include
-MyX <- mat_lag(MyData[,-1], p)  # Include 1:p lags of all variables but the 1st
-MyY <- MyData[-(1:p),1]         # Drop the first p lags of the 1st variable
-
-# Estimate the enet for ALL POSSIBLE values
-# Includes intercept, all variables normalized before estimate
-enet_BIP <- enet(MyX, MyY)
-# plot the results
-plot(enet_BIP, use.color = T, lwd=3)
-
-print(enet_BIP)
-
-# show the estimated coefficients at each step in the path.
-K <- ncol(MyX)
-predict(enet_BIP, type = "coefficients", s = 2:(K+1))
-
-
-# What's the order of the VAR we're modeling?
-p<-3
-p <- 2
-# What's the depth of the ENet we want to use?
-nsteps <- 6
-nsteps <- 5
-# How many variables in our VAR?
[…]

### 2025-06-04 Latex/BIP_predictor.tex
up

--- 2025-06-02 Latex/BIP_predictor.tex
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May 29 23:26:52 2025
+% Wed Jun  4 12:00:56 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May 29 23:26:52 2025
+% Wed Jun  4 12:00:56 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May 29 23:26:53 2025
+% Wed Jun  4 12:00:58 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May 29 23:26:54 2025
+% Wed Jun  4 12:00:59 2025
 \begin{table}[ht]
@@
   \hline
-Target correlation & 0.68 & 0.69 \\ 
-  HT & 6.50 & 3.71 \\ 
+Target correlation & 0.69 & 0.73 \\ 
+  HT & 8.67 & 4.73 \\ 
    \hline
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May 29 23:26:56 2025
+% Wed Jun  4 12:01:02 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 0.000 & 0.000 & 0.001 & 0.002 & 0.005 & 0.020 & 0.112 \\ 
-  Shift=1 & 0.000 & 0.000 & 0.000 & 0.001 & 0.002 & 0.005 & 0.019 \\ 
-  Shift=2 & 0.000 & 0.000 & 0.000 & 0.001 & 0.001 & 0.003 & 0.006 \\ 
-  Shift=3 & 0.018 & 0.007 & 0.002 & 0.000 & 0.002 & 0.003 & 0.004 \\ 
-  Shift=4 & 0.180 & 0.100 & 0.040 & 0.010 & 0.002 & 0.005 & 0.005 \\ 
-  Shift=5 & 0.579 & 0.432 & 0.261 & 0.108 & 0.027 & 0.008 & 0.009 \\ 
+Shift=0 & 0.000 & 0.000 & 0.000 & 0.000 & 0.001 & 0.017 & 0.244 \\ 
+  Shift=1 & 0.000 & 0.000 & 0.000 & 0.000 & 0.000 & 0.003 & 0.026 \\ 
+  Shift=2 & 0.001 & 0.000 & 0.000 & 0.000 & 0.000 & 0.001 & 0.007 \\ 
+  Shift=3 & 0.052 & 0.023 & 0.005 & 0.001 & 0.002 & 0.001 & 0.002 \\ 
+  Shift=4 & 0.370 & 0.240 & 0.100 & 0.015 & 0.006 & 0.004 & 0.002 \\ 
+  Shift=5 & 0.792 & 0.673 & 0.450 & 0.157 & 0.016 & 0.013 & 0.007 \\ 
    \hline
@@
   \hline
-Shift=0 & 0.084 & 0.047 & 0.017 & 0.005 & 0.002 & 0.001 & 0.038 \\ 
-  Shift=1 & 0.230 & 0.160 & 0.094 & 0.032 & 0.005 & 0.001 & 0.004 \\ 
-  Shift=2 & 0.801 & 0.666 & 0.414 & 0.173 & 0.026 & 0.005 & 0.006 \\ 
-  Shift=3 & 0.876 & 0.961 & 0.917 & 0.683 & 0.248 & 0.068 & 0.039 \\ 
-  Shift=4 & 0.370 & 0.819 & 0.990 & 0.953 & 0.533 & 0.117 & 0.030 \\ 
-  Shift=5 & 0.353 & 0.802 & 0.983 & 0.943 & 0.571 & 0.239 & 0.043 \\ 
+Shift=0 & 0.090 & 0.060 & 0.025 & 0.006 & 0.002 & 0.001 & 0.031 \\ 
+  Shift=1 & 0.271 & 0.202 & 0.126 & 0.053 & 0.007 & 0.001 & 0.006 \\ 
+  Shift=2 & 0.820 & 0.723 & 0.502 & 0.225 & 0.050 & 0.005 & 0.005 \\ 
+  Shift=3 & 0.878 & 0.952 & 0.906 & 0.678 & 0.251 & 0.049 & 0.026 \\ 
+  Shift=4 & 0.641 & 0.916 & 0.986 & 0.938 & 0.549 & 0.147 & 0.020 \\ 
+  Shift=5 & 0.494 & 0.842 & 0.994 & 0.973 & 0.675 & 0.298 & 0.058 \\ 
    \hline
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May 29 23:26:56 2025
+% Wed Jun  4 12:01:02 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 0.973 & 0.937 & 0.888 & 0.850 & 0.855 & 0.900 & 0.956 \\ 
-  Shift=1 & 1.073 & 1.051 & 0.997 & 0.918 & 0.870 & 0.873 & 0.906 \\ 
-  Shift=2 & 1.069 & 1.077 & 1.063 & 1.006 & 0.940 & 0.907 & 0.907 \\ 
-  Shift=3 & 1.020 & 1.045 & 1.066 & 1.048 & 0.997 & 0.957 & 0.941 \\ 
-  Shift=4 & 0.983 & 1.016 & 1.056 & 1.064 & 1.023 & 0.974 & 0.942 \\ 
-  Shift=5 & 0.988 & 1.026 & 1.071 & 1.088 & 1.057 & 1.007 & 0.963 \\ 
+Shift=0 & 0.997 & 0.966 & 0.911 & 0.856 & 0.847 & 0.890 & 0.953 \\ 
+  Shift=1 & 1.099 & 1.088 & 1.038 & 0.951 & 0.884 & 0.875 & 0.912 \\ 
+  Shift=2 & 1.082 & 1.099 & 1.092 & 1.032 & 0.952 & 0.906 & 0.904 \\ 
+  Shift=3 & 1.033 & 1.068 & 1.097 & 1.076 & 1.009 & 0.954 & 0.934 \\ 
+  Shift=4 & 1.005 & 1.051 & 1.101 & 1.108 & 1.054 & 0.982 & 0.935 \\ 
+  Shift=5 & 1.006 & 1.054 & 1.104 & 1.121 & 1.092 & 1.031 & 0.969 \\ 
    \hline
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May 29 23:26:59 2025
+% Wed Jun  4 12:01:06 2025
 \begin{table}[ht]
@@
   \hline
-Forecast horizon 1 & 0.752 & 0.715 & 0.746 & 0.747 \\ 
-  Forecast horizon 2 & 0.697 & 0.663 & 0.729 & 0.720 \\ 
-  Forecast horizon 3 & 0.554 & 0.622 & 0.684 & 0.709 \\ 
-  Forecast horizon 4 & 0.486 & 0.624 & 0.680 & 0.708 \\ 
-  Forecast horizon 5 & 0.593 & 0.636 & 0.643 & 0.701 \\ 
+Forecast horizon 1 & 0.752 & 0.715 & 0.749 & 0.752 \\ 
+  Forecast horizon 2 & 0.697 & 0.663 & 0.715 & 0.740 \\ 
+  Forecast horizon 3 & 0.554 & 0.622 & 0.669 & 0.706 \\ 
+  Forecast horizon 4 & 0.486 & 0.624 & 0.706 & 0.737 \\ 
+  Forecast horizon 5 & 0.593 & 0.636 & 0.660 & 0.714 \\ 
    \hline
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Thu May 29 23:26:59 2025
+% Wed Jun  4 12:01:06 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 1.128 & 1.086 & 1.030 & 0.985 & 0.992 & 1.044 & 1.109 \\ 
-  Shift=1 & 1.280 & 1.254 & 1.189 & 1.096 & 1.037 & 1.041 & 1.081 \\ 
-  Shift=2 & 1.173 & 1.181 & 1.166 & 1.104 & 1.031 & 0.995 & 0.995 \\ 
-  Shift=3 & 1.012 & 1.037 & 1.058 & 1.040 & 0.989 & 0.949 & 0.934 \\ 
-  Shift=4 & 0.982 & 1.016 & 1.056 & 1.063 & 1.022 & 0.974 & 0.941 \\ 
-  Shift=5 & 0.961 & 0.997 & 1.041 & 1.058 & 1.028 & 0.979 & 0.936 \\ 
+Shift=0 & 1.157 & 1.120 & 1.056 & 0.992 & 0.982 & 1.032 & 1.105 \\ 
+  Shift=1 & 1.312 & 1.298 & 1.238 & 1.135 & 1.055 & 1.044 & 1.089 \\ 
+  Shift=2 & 1.187 & 1.205 & 1.198 & 1.132 & 1.045 & 0.994 & 0.991 \\ 
+  Shift=3 & 1.025 & 1.060 & 1.089 & 1.067 & 1.001 & 0.946 & 0.927 \\ 
+  Shift=4 & 1.004 & 1.050 & 1.100 & 1.108 & 1.053 & 0.982 & 0.935 \\ 
+  Shift=5 & 0.978 & 1.025 & 1.073 & 1.090 & 1.062 & 1.003 & 0.942 \\ 
    \hline
@@
   \hline
-Shift=0 & 0.127 & 0.064 & 0.019 & 0.003 & 0.001 & 0.002 & 0.012 \\ 
-  Shift=1 & 0.439 & 0.314 & 0.194 & 0.093 & 0.033 & 0.017 & 0.036 \\ 
-  Shift=2 & 0.718 & 0.565 & 0.326 & 0.098 & 0.020 & 0.008 & 0.010 \\ 
-  Shift=3 & 0.946 & 0.975 & 0.925 & 0.657 & 0.222 & 0.059 & 0.028 \\ 
-  Shift=4 & 0.327 & 0.699 & 0.964 & 0.942 & 0.619 & 0.237 & 0.102 \\ 
-  Shift=5 & 0.383 & 0.755 & 0.946 & 0.910 & 0.581 & 0.241 & 0.055 \\ 
+Shift=0 & 0.112 & 0.062 & 0.019 & 0.002 & 0.000 & 0.001 & 0.006 \\ 
+  Shift=1 & 0.451 & 0.327 & 0.206 & 0.108 & 0.042 & 0.031 & 0.073 \\ 
+  Shift=2 & 0.755 & 0.646 & 0.425 & 0.150 & 0.025 & 0.008 & 0.014 \\ 
+  Shift=3 & 0.936 & 0.952 & 0.901 & 0.664 & 0.229 & 0.046 & 0.017 \\ 
+  Shift=4 & 0.502 & 0.838 & 0.956 & 0.918 & 0.613 & 0.230 & 0.092 \\ 
+  Shift=5 & 0.414 & 0.820 & 0.967 & 0.944 & 0.684 & 0.303 & 0.057 \\ 
    \hline
@@
   \hline
-Shift=0 & 1.005 & 0.986 & 0.960 & 0.935 & 0.927 & 0.939 & 0.964 \\ 
-  Shift=1 & 1.060 & 1.054 & 1.032 & 0.998 & 0.973 & 0.965 & 0.970 \\ 
-  Shift=2 & 1.028 & 1.029 & 1.020 & 0.993 & 0.963 & 0.950 & 0.952 \\ 
-  Shift=3 & 1.018 & 1.027 & 1.034 & 1.022 & 0.994 & 0.972 & 0.961 \\ 
-  Shift=4 & 0.990 & 1.006 & 1.025 & 1.032 & 1.017 & 0.998 & 0.985 \\ 
-  Shift=5 & 1.000 & 1.016 & 1.037 & 1.046 & 1.032 & 1.008 & 0.985 \\ 
+Shift=0 & 1.010 & 0.993 & 0.964 & 0.934 & 0.920 & 0.925 & 0.949 \\ 
+  Shift=1 & 1.061 & 1.057 & 1.037 & 1.003 & 0.979 & 0.975 & 0.981 \\ 
+  Shift=2 & 1.034 & 1.040 & 1.035 & 1.005 & 0.969 & 0.952 & 0.958 \\ 
+  Shift=3 & 1.021 & 1.035 & 1.046 & 1.034 & 1.000 & 0.971 & 0.956 \\ 
+  Shift=4 & 0.999 & 1.020 & 1.044 & 1.050 & 1.030 & 1.003 & 0.986 \\ 
+  Shift=5 & 1.005 & 1.027 & 1.052 & 1.063 & 1.050 & 1.021 & 0.989 \\ 
    \hline
@@
   \hline
-Shift=0 & 1.237 & 1.215 & 1.183 & 1.152 & 1.141 & 1.156 & 1.187 \\ 
-  Shift=1 & 0.986 & 0.980 & 0.960 & 0.929 & 0.905 & 0.898 & 0.903 \\ 
-  Shift=2 & 1.036 & 1.037 & 1.029 & 1.001 & 0.971 & 0.957 & 0.960 \\ 
-  Shift=3 & 1.042 & 1.051 & 1.058 & 1.045 & 1.018 & 0.995 & 0.983 \\ 
-  Shift=4 & 0.970 & 0.985 & 1.004 & 1.010 & 0.996 & 0.978 & 0.965 \\ 
-  Shift=5 & 0.970 & 0.986 & 1.007 & 1.015 & 1.002 & 0.978 & 0.956 \\ 
+Shift=0 & 1.244 & 1.223 & 1.187 & 1.150 & 1.133 & 1.140 & 1.169 \\ 
+  Shift=1 & 0.987 & 0.984 & 0.965 & 0.933 & 0.911 & 0.907 & 0.912 \\ 
+  Shift=2 & 1.043 & 1.049 & 1.043 & 1.013 & 0.977 & 0.960 & 0.966 \\ 
+  Shift=3 & 1.044 & 1.059 & 1.071 & 1.058 & 1.024 & 0.993 & 0.978 \\ 
+  Shift=4 & 0.978 & 0.999 & 1.023 & 1.029 & 1.009 & 0.983 & 0.966 \\ 
+  Shift=5 & 0.975 & 0.997 & 1.021 & 1.031 & 1.019 & 0.991 & 0.960 \\ 
    \hline

### 2025-06-04 Sweave/BIP_predictor.Rnw
up

@@
   q<-0
-# Length of MA inversion (impulse response function)  
-  nirf<-L
 # Number of non-zero coefficients in VAR for each dimension  
@@
   {
-    V_obj<-enetVAR(data_fit, p, nIRF, nsteps  )
+    V_obj<-enetVAR(data_fit, p, L, nsteps  )
 
@@
   {
-    V_obj_wc<-enetVAR(data_fit, p, nIRF, nsteps  )
+    V_obj_wc<-enetVAR(data_fit, p, L, nsteps  )
 
@@
   }
+  if (VAR_type=="elastic_net")
+  {
+    save(mssa_indicator_obj,file=paste(path.result,"/en_nowcast_obj",sep=""))
+    save(mssa_indicator_HP_HT_obj,file=paste(path.result,"/en_nowcast_HT_HP_obj",sep=""))
+  }
+
 } else
@@
     load(file=paste(path.result,"/BV_nowcast_HT_HP_obj",sep=""))
+  }
+  if (VAR_type=="elastic_net")
+  {
+    load(file=paste(path.result,"/en_nowcast_obj",sep=""))
+    load(file=paste(path.result,"/en_nowcast_HT_HP_obj",sep=""))
   }
@@
   }
+  if (VAR_type=="elastic_net")
+  {
+    save(mssa_indicator_obj,file=paste(path.result,"/en_mssa_indicator_obj",sep=""))
+    save(mssa_indicator_obj_wc,file=paste(path.result,"/en_mssa_indicator_obj_wc",sep=""))
+  }
 
@@
     load(file=paste(path.result,"/BV_mssa_indicator_obj_wc",sep=""))
+  }
+  if (VAR_type=="elastic_net")
+  {
+    load(file=paste(path.result,"/en_mssa_indicator_obj",sep=""))
+    load(file=paste(path.result,"/en_mssa_indicator_obj_wc",sep=""))
   }
@@
   }
+  if (VAR_type=="elastic_net")
+  {
+    save(list_perf,file=paste(path.result,"en_list_perf",sep=""))
+  }
 } else
@@
     load(file=paste(path.result,"BV_list_perf",sep=""))
+  }
+  if (VAR_type=="elastic_net")
+  {
+    load(file=paste(path.result,"en_list_perf",sep=""))
   }
@@
   }
+  if (VAR_type=="elastic_net")
+  {
+    save(final_mssa_indicator_obj,file=paste(path.result,"/en_final_mssa_indicator_obj",sep=""))
+  }
 } else
@@
   }
+  if (VAR_type=="elastic_net")
+  {
+    load(file=paste(path.result,"/en_final_mssa_indicator_obj",sep=""))
+  }
+
 }
@@
   }
+  if (VAR_type=="elastic_net")
+  {
+    V_obj_full<-enetVAR(data_fit, p, L, nsteps  )
+
+    save(V_obj_full,file=paste(path.result,"/enVAR_full",sep=""))
+  }
 
@@
   }
+  if (VAR_type=="elastic_net")
+  {
+    V_obj_2008<-enetVAR(data_fit, p, L, nsteps  )
+
+    save(V_obj_2008,file=paste(path.result,"/enVAR_2008",sep=""))
+  }
 
@@
   }
+  if (VAR_type=="elastic_net")
+  {
+    load(file=paste(path.result,"/enVAR_full",sep=""))
+    load(file=paste(path.result,"/enVAR_2008",sep=""))
+  }
+
 }

### 2025-06-04 Latex/BIP_predictor.tex
up

@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:00:56 2025
+% Wed Jun  4 14:24:40 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:00:56 2025
+% Wed Jun  4 14:24:41 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:00:58 2025
+% Wed Jun  4 14:24:43 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:00:59 2025
+% Wed Jun  4 14:24:53 2025
 \begin{table}[ht]
@@
   \hline
-Target correlation & 0.69 & 0.73 \\ 
-  HT & 8.67 & 4.73 \\ 
+Target correlation & 0.71 & 0.73 \\ 
+  HT & 7.43 & 5.78 \\ 
    \hline
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:01:02 2025
+% Wed Jun  4 14:25:53 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 0.000 & 0.000 & 0.000 & 0.000 & 0.001 & 0.017 & 0.244 \\ 
-  Shift=1 & 0.000 & 0.000 & 0.000 & 0.000 & 0.000 & 0.003 & 0.026 \\ 
-  Shift=2 & 0.001 & 0.000 & 0.000 & 0.000 & 0.000 & 0.001 & 0.007 \\ 
-  Shift=3 & 0.052 & 0.023 & 0.005 & 0.001 & 0.002 & 0.001 & 0.002 \\ 
-  Shift=4 & 0.370 & 0.240 & 0.100 & 0.015 & 0.006 & 0.004 & 0.002 \\ 
-  Shift=5 & 0.792 & 0.673 & 0.450 & 0.157 & 0.016 & 0.013 & 0.007 \\ 
+Shift=0 & 0.000 & 0.000 & 0.000 & 0.000 & 0.002 & 0.031 & 0.469 \\ 
+  Shift=1 & 0.000 & 0.000 & 0.000 & 0.000 & 0.000 & 0.005 & 0.051 \\ 
+  Shift=2 & 0.001 & 0.000 & 0.000 & 0.000 & 0.000 & 0.001 & 0.010 \\ 
+  Shift=3 & 0.058 & 0.023 & 0.003 & 0.002 & 0.001 & 0.001 & 0.004 \\ 
+  Shift=4 & 0.398 & 0.244 & 0.079 & 0.006 & 0.006 & 0.002 & 0.003 \\ 
+  Shift=5 & 0.815 & 0.681 & 0.405 & 0.092 & 0.019 & 0.011 & 0.006 \\ 
    \hline
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:01:02 2025
+% Wed Jun  4 14:26:07 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 0.090 & 0.060 & 0.025 & 0.006 & 0.002 & 0.001 & 0.031 \\ 
-  Shift=1 & 0.271 & 0.202 & 0.126 & 0.053 & 0.007 & 0.001 & 0.006 \\ 
-  Shift=2 & 0.820 & 0.723 & 0.502 & 0.225 & 0.050 & 0.005 & 0.005 \\ 
-  Shift=3 & 0.878 & 0.952 & 0.906 & 0.678 & 0.251 & 0.049 & 0.026 \\ 
-  Shift=4 & 0.641 & 0.916 & 0.986 & 0.938 & 0.549 & 0.147 & 0.020 \\ 
-  Shift=5 & 0.494 & 0.842 & 0.994 & 0.973 & 0.675 & 0.298 & 0.058 \\ 
+Shift=0 & 0.071 & 0.045 & 0.017 & 0.004 & 0.002 & 0.001 & 0.068 \\ 
+  Shift=1 & 0.235 & 0.177 & 0.107 & 0.039 & 0.004 & 0.001 & 0.008 \\ 
+  Shift=2 & 0.793 & 0.679 & 0.432 & 0.184 & 0.029 & 0.004 & 0.006 \\ 
+  Shift=3 & 0.918 & 0.949 & 0.877 & 0.584 & 0.187 & 0.032 & 0.022 \\ 
+  Shift=4 & 0.742 & 0.949 & 0.984 & 0.898 & 0.466 & 0.080 & 0.013 \\ 
+  Shift=5 & 0.609 & 0.893 & 0.995 & 0.960 & 0.595 & 0.235 & 0.035 \\ 
    \hline
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:01:02 2025
+% Wed Jun  4 14:26:07 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 0.997 & 0.966 & 0.911 & 0.856 & 0.847 & 0.890 & 0.953 \\ 
-  Shift=1 & 1.099 & 1.088 & 1.038 & 0.951 & 0.884 & 0.875 & 0.912 \\ 
-  Shift=2 & 1.082 & 1.099 & 1.092 & 1.032 & 0.952 & 0.906 & 0.904 \\ 
-  Shift=3 & 1.033 & 1.068 & 1.097 & 1.076 & 1.009 & 0.954 & 0.934 \\ 
-  Shift=4 & 1.005 & 1.051 & 1.101 & 1.108 & 1.054 & 0.982 & 0.935 \\ 
-  Shift=5 & 1.006 & 1.054 & 1.104 & 1.121 & 1.092 & 1.031 & 0.969 \\ 
+Shift=0 & 0.975 & 0.941 & 0.886 & 0.839 & 0.843 & 0.897 & 0.965 \\ 
+  Shift=1 & 1.085 & 1.068 & 1.011 & 0.927 & 0.871 & 0.875 & 0.919 \\ 
+  Shift=2 & 1.076 & 1.089 & 1.074 & 1.011 & 0.938 & 0.902 & 0.909 \\ 
+  Shift=3 & 1.037 & 1.069 & 1.090 & 1.059 & 0.992 & 0.943 & 0.930 \\ 
+  Shift=4 & 1.013 & 1.058 & 1.101 & 1.097 & 1.037 & 0.967 & 0.925 \\ 
+  Shift=5 & 1.016 & 1.062 & 1.106 & 1.115 & 1.078 & 1.015 & 0.955 \\ 
    \hline
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:01:06 2025
+% Wed Jun  4 14:26:39 2025
 \begin{table}[ht]
@@
   \hline
-Forecast horizon 1 & 0.752 & 0.715 & 0.749 & 0.752 \\ 
-  Forecast horizon 2 & 0.697 & 0.663 & 0.715 & 0.740 \\ 
-  Forecast horizon 3 & 0.554 & 0.622 & 0.669 & 0.706 \\ 
-  Forecast horizon 4 & 0.486 & 0.624 & 0.706 & 0.737 \\ 
-  Forecast horizon 5 & 0.593 & 0.636 & 0.660 & 0.714 \\ 
+Forecast horizon 1 & 0.752 & 0.715 & 0.750 & 0.756 \\ 
+  Forecast horizon 2 & 0.697 & 0.663 & 0.714 & 0.744 \\ 
+  Forecast horizon 3 & 0.554 & 0.622 & 0.666 & 0.716 \\ 
+  Forecast horizon 4 & 0.486 & 0.624 & 0.707 & 0.732 \\ 
+  Forecast horizon 5 & 0.593 & 0.636 & 0.660 & 0.713 \\ 
    \hline
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:01:06 2025
+% Wed Jun  4 14:26:39 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 1.157 & 1.120 & 1.056 & 0.992 & 0.982 & 1.032 & 1.105 \\ 
-  Shift=1 & 1.312 & 1.298 & 1.238 & 1.135 & 1.055 & 1.044 & 1.089 \\ 
-  Shift=2 & 1.187 & 1.205 & 1.198 & 1.132 & 1.045 & 0.994 & 0.991 \\ 
-  Shift=3 & 1.025 & 1.060 & 1.089 & 1.067 & 1.001 & 0.946 & 0.927 \\ 
-  Shift=4 & 1.004 & 1.050 & 1.100 & 1.108 & 1.053 & 0.982 & 0.935 \\ 
-  Shift=5 & 0.978 & 1.025 & 1.073 & 1.090 & 1.062 & 1.003 & 0.942 \\ 
+Shift=0 & 1.131 & 1.092 & 1.028 & 0.973 & 0.977 & 1.040 & 1.119 \\ 
+  Shift=1 & 1.295 & 1.274 & 1.206 & 1.106 & 1.039 & 1.044 & 1.097 \\ 
+  Shift=2 & 1.180 & 1.194 & 1.178 & 1.109 & 1.029 & 0.990 & 0.997 \\ 
+  Shift=3 & 1.029 & 1.061 & 1.082 & 1.051 & 0.984 & 0.936 & 0.923 \\ 
+  Shift=4 & 1.012 & 1.057 & 1.100 & 1.096 & 1.036 & 0.967 & 0.924 \\ 
+  Shift=5 & 0.988 & 1.033 & 1.076 & 1.084 & 1.048 & 0.987 & 0.929 \\ 
    \hline
@@
   \hline
-Shift=0 & 0.112 & 0.062 & 0.019 & 0.002 & 0.000 & 0.001 & 0.006 \\ 
-  Shift=1 & 0.451 & 0.327 & 0.206 & 0.108 & 0.042 & 0.031 & 0.073 \\ 
-  Shift=2 & 0.755 & 0.646 & 0.425 & 0.150 & 0.025 & 0.008 & 0.014 \\ 
-  Shift=3 & 0.936 & 0.952 & 0.901 & 0.664 & 0.229 & 0.046 & 0.017 \\ 
-  Shift=4 & 0.502 & 0.838 & 0.956 & 0.918 & 0.613 & 0.230 & 0.092 \\ 
-  Shift=5 & 0.414 & 0.820 & 0.967 & 0.944 & 0.684 & 0.303 & 0.057 \\ 
+Shift=0 & 0.080 & 0.042 & 0.011 & 0.001 & 0.000 & 0.002 & 0.011 \\ 
+  Shift=1 & 0.403 & 0.297 & 0.188 & 0.095 & 0.042 & 0.039 & 0.103 \\ 
+  Shift=2 & 0.718 & 0.594 & 0.359 & 0.114 & 0.019 & 0.009 & 0.022 \\ 
+  Shift=3 & 0.955 & 0.950 & 0.871 & 0.569 & 0.158 & 0.031 & 0.013 \\ 
+  Shift=4 & 0.593 & 0.879 & 0.957 & 0.893 & 0.541 & 0.183 & 0.080 \\ 
+  Shift=5 & 0.531 & 0.875 & 0.971 & 0.929 & 0.605 & 0.239 & 0.041 \\ 
    \hline
@@
   \hline
-Shift=0 & 1.010 & 0.993 & 0.964 & 0.934 & 0.920 & 0.925 & 0.949 \\ 
-  Shift=1 & 1.061 & 1.057 & 1.037 & 1.003 & 0.979 & 0.975 & 0.981 \\ 
-  Shift=2 & 1.034 & 1.040 & 1.035 & 1.005 & 0.969 & 0.952 & 0.958 \\ 
-  Shift=3 & 1.021 & 1.035 & 1.046 & 1.034 & 1.000 & 0.971 & 0.956 \\ 
-  Shift=4 & 0.999 & 1.020 & 1.044 & 1.050 & 1.030 & 1.003 & 0.986 \\ 
-  Shift=5 & 1.005 & 1.027 & 1.052 & 1.063 & 1.050 & 1.021 & 0.989 \\ 
+Shift=0 & 0.999 & 0.980 & 0.951 & 0.925 & 0.915 & 0.926 & 0.955 \\ 
+  Shift=1 & 1.057 & 1.051 & 1.028 & 0.996 & 0.977 & 0.978 & 0.985 \\ 
+  Shift=2 & 1.031 & 1.035 & 1.026 & 0.996 & 0.964 & 0.952 & 0.964 \\ 
+  Shift=3 & 1.022 & 1.035 & 1.042 & 1.026 & 0.992 & 0.965 & 0.953 \\ 
+  Shift=4 & 1.002 & 1.023 & 1.044 & 1.046 & 1.024 & 0.999 & 0.984 \\ 
+  Shift=5 & 1.010 & 1.032 & 1.054 & 1.060 & 1.043 & 1.013 & 0.983 \\ 
    \hline
@@
   \hline
-Shift=0 & 1.244 & 1.223 & 1.187 & 1.150 & 1.133 & 1.140 & 1.169 \\ 
-  Shift=1 & 0.987 & 0.984 & 0.965 & 0.933 & 0.911 & 0.907 & 0.912 \\ 
-  Shift=2 & 1.043 & 1.049 & 1.043 & 1.013 & 0.977 & 0.960 & 0.966 \\ 
-  Shift=3 & 1.044 & 1.059 & 1.071 & 1.058 & 1.024 & 0.993 & 0.978 \\ 
-  Shift=4 & 0.978 & 0.999 & 1.023 & 1.029 & 1.009 & 0.983 & 0.966 \\ 
-  Shift=5 & 0.975 & 0.997 & 1.021 & 1.031 & 1.019 & 0.991 & 0.960 \\ 
+Shift=0 & 1.230 & 1.207 & 1.172 & 1.139 & 1.127 & 1.140 & 1.177 \\ 
+  Shift=1 & 0.984 & 0.977 & 0.956 & 0.927 & 0.909 & 0.910 & 0.916 \\ 
+  Shift=2 & 1.039 & 1.044 & 1.035 & 1.004 & 0.972 & 0.960 & 0.972 \\ 
+  Shift=3 & 1.046 & 1.059 & 1.067 & 1.050 & 1.015 & 0.987 & 0.975 \\ 
+  Shift=4 & 0.982 & 1.002 & 1.023 & 1.024 & 1.003 & 0.978 & 0.964 \\ 
+  Shift=5 & 0.979 & 1.001 & 1.022 & 1.028 & 1.012 & 0.983 & 0.953 \\ 
    \hline

### 2025-06-05 Latex/BIP_predictor.tex
up

--- 2025-06-04 Latex/BIP_predictor.tex
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:00:56 2025
+% Thu Jun  5 09:11:25 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:00:56 2025
+% Thu Jun  5 09:11:26 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:00:58 2025
+% Thu Jun  5 09:11:27 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:00:59 2025
+% Thu Jun  5 09:11:34 2025
 \begin{table}[ht]
@@
   \hline
-Target correlation & 0.69 & 0.73 \\ 
-  HT & 8.67 & 4.73 \\ 
+Target correlation & 0.70 & 0.73 \\ 
+  HT & 7.43 & 5.78 \\ 
    \hline
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:01:02 2025
+% Thu Jun  5 09:12:18 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 0.000 & 0.000 & 0.000 & 0.000 & 0.001 & 0.017 & 0.244 \\ 
-  Shift=1 & 0.000 & 0.000 & 0.000 & 0.000 & 0.000 & 0.003 & 0.026 \\ 
-  Shift=2 & 0.001 & 0.000 & 0.000 & 0.000 & 0.000 & 0.001 & 0.007 \\ 
-  Shift=3 & 0.052 & 0.023 & 0.005 & 0.001 & 0.002 & 0.001 & 0.002 \\ 
-  Shift=4 & 0.370 & 0.240 & 0.100 & 0.015 & 0.006 & 0.004 & 0.002 \\ 
-  Shift=5 & 0.792 & 0.673 & 0.450 & 0.157 & 0.016 & 0.013 & 0.007 \\ 
+Shift=0 & 0.000 & 0.000 & 0.000 & 0.000 & 0.002 & 0.036 & 0.488 \\ 
+  Shift=1 & 0.000 & 0.000 & 0.000 & 0.000 & 0.000 & 0.006 & 0.058 \\ 
+  Shift=2 & 0.001 & 0.000 & 0.000 & 0.000 & 0.000 & 0.002 & 0.012 \\ 
+  Shift=3 & 0.067 & 0.027 & 0.004 & 0.002 & 0.001 & 0.001 & 0.005 \\ 
+  Shift=4 & 0.429 & 0.271 & 0.087 & 0.007 & 0.006 & 0.003 & 0.004 \\ 
+  Shift=5 & 0.835 & 0.710 & 0.428 & 0.097 & 0.018 & 0.011 & 0.007 \\ 
    \hline
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:01:02 2025
+% Thu Jun  5 09:12:28 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 0.090 & 0.060 & 0.025 & 0.006 & 0.002 & 0.001 & 0.031 \\ 
-  Shift=1 & 0.271 & 0.202 & 0.126 & 0.053 & 0.007 & 0.001 & 0.006 \\ 
-  Shift=2 & 0.820 & 0.723 & 0.502 & 0.225 & 0.050 & 0.005 & 0.005 \\ 
-  Shift=3 & 0.878 & 0.952 & 0.906 & 0.678 & 0.251 & 0.049 & 0.026 \\ 
-  Shift=4 & 0.641 & 0.916 & 0.986 & 0.938 & 0.549 & 0.147 & 0.020 \\ 
-  Shift=5 & 0.494 & 0.842 & 0.994 & 0.973 & 0.675 & 0.298 & 0.058 \\ 
+Shift=0 & 0.077 & 0.049 & 0.017 & 0.004 & 0.002 & 0.001 & 0.079 \\ 
+  Shift=1 & 0.249 & 0.186 & 0.108 & 0.039 & 0.004 & 0.001 & 0.008 \\ 
+  Shift=2 & 0.805 & 0.694 & 0.436 & 0.181 & 0.029 & 0.003 & 0.005 \\ 
+  Shift=3 & 0.905 & 0.949 & 0.878 & 0.585 & 0.187 & 0.034 & 0.023 \\ 
+  Shift=4 & 0.722 & 0.944 & 0.984 & 0.900 & 0.471 & 0.075 & 0.015 \\ 
+  Shift=5 & 0.591 & 0.892 & 0.995 & 0.958 & 0.588 & 0.232 & 0.031 \\ 
    \hline
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:01:02 2025
+% Thu Jun  5 09:12:28 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 0.997 & 0.966 & 0.911 & 0.856 & 0.847 & 0.890 & 0.953 \\ 
-  Shift=1 & 1.099 & 1.088 & 1.038 & 0.951 & 0.884 & 0.875 & 0.912 \\ 
-  Shift=2 & 1.082 & 1.099 & 1.092 & 1.032 & 0.952 & 0.906 & 0.904 \\ 
-  Shift=3 & 1.033 & 1.068 & 1.097 & 1.076 & 1.009 & 0.954 & 0.934 \\ 
-  Shift=4 & 1.005 & 1.051 & 1.101 & 1.108 & 1.054 & 0.982 & 0.935 \\ 
-  Shift=5 & 1.006 & 1.054 & 1.104 & 1.121 & 1.092 & 1.031 & 0.969 \\ 
+Shift=0 & 0.981 & 0.947 & 0.888 & 0.837 & 0.841 & 0.897 & 0.967 \\ 
+  Shift=1 & 1.087 & 1.071 & 1.012 & 0.926 & 0.866 & 0.870 & 0.919 \\ 
+  Shift=2 & 1.075 & 1.089 & 1.074 & 1.010 & 0.936 & 0.897 & 0.904 \\ 
+  Shift=3 & 1.036 & 1.069 & 1.090 & 1.059 & 0.991 & 0.943 & 0.931 \\ 
+  Shift=4 & 1.012 & 1.058 & 1.101 & 1.096 & 1.036 & 0.966 & 0.927 \\ 
+  Shift=5 & 1.015 & 1.064 & 1.107 & 1.114 & 1.078 & 1.015 & 0.951 \\ 
    \hline
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:01:06 2025
+% Thu Jun  5 09:12:55 2025
 \begin{table}[ht]
@@
   \hline
-Forecast horizon 1 & 0.752 & 0.715 & 0.749 & 0.752 \\ 
-  Forecast horizon 2 & 0.697 & 0.663 & 0.715 & 0.740 \\ 
-  Forecast horizon 3 & 0.554 & 0.622 & 0.669 & 0.706 \\ 
-  Forecast horizon 4 & 0.486 & 0.624 & 0.706 & 0.737 \\ 
-  Forecast horizon 5 & 0.593 & 0.636 & 0.660 & 0.714 \\ 
+Forecast horizon 1 & 0.752 & 0.715 & 0.750 & 0.750 \\ 
+  Forecast horizon 2 & 0.697 & 0.663 & 0.714 & 0.740 \\ 
+  Forecast horizon 3 & 0.554 & 0.622 & 0.666 & 0.710 \\ 
+  Forecast horizon 4 & 0.486 & 0.624 & 0.707 & 0.731 \\ 
+  Forecast horizon 5 & 0.593 & 0.636 & 0.660 & 0.710 \\ 
    \hline
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:01:06 2025
+% Thu Jun  5 09:12:55 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 1.157 & 1.120 & 1.056 & 0.992 & 0.982 & 1.032 & 1.105 \\ 
-  Shift=1 & 1.312 & 1.298 & 1.238 & 1.135 & 1.055 & 1.044 & 1.089 \\ 
-  Shift=2 & 1.187 & 1.205 & 1.198 & 1.132 & 1.045 & 0.994 & 0.991 \\ 
-  Shift=3 & 1.025 & 1.060 & 1.089 & 1.067 & 1.001 & 0.946 & 0.927 \\ 
-  Shift=4 & 1.004 & 1.050 & 1.100 & 1.108 & 1.053 & 0.982 & 0.935 \\ 
-  Shift=5 & 0.978 & 1.025 & 1.073 & 1.090 & 1.062 & 1.003 & 0.942 \\ 
+Shift=0 & 1.138 & 1.098 & 1.029 & 0.971 & 0.975 & 1.040 & 1.121 \\ 
+  Shift=1 & 1.297 & 1.277 & 1.207 & 1.104 & 1.034 & 1.038 & 1.096 \\ 
+  Shift=2 & 1.180 & 1.195 & 1.179 & 1.108 & 1.027 & 0.984 & 0.992 \\ 
+  Shift=3 & 1.028 & 1.061 & 1.082 & 1.051 & 0.984 & 0.936 & 0.924 \\ 
+  Shift=4 & 1.011 & 1.058 & 1.101 & 1.095 & 1.035 & 0.965 & 0.926 \\ 
+  Shift=5 & 0.987 & 1.034 & 1.076 & 1.084 & 1.048 & 0.987 & 0.925 \\ 
    \hline
@@
   \hline
-Shift=0 & 0.112 & 0.062 & 0.019 & 0.002 & 0.000 & 0.001 & 0.006 \\ 
-  Shift=1 & 0.451 & 0.327 & 0.206 & 0.108 & 0.042 & 0.031 & 0.073 \\ 
-  Shift=2 & 0.755 & 0.646 & 0.425 & 0.150 & 0.025 & 0.008 & 0.014 \\ 
-  Shift=3 & 0.936 & 0.952 & 0.901 & 0.664 & 0.229 & 0.046 & 0.017 \\ 
-  Shift=4 & 0.502 & 0.838 & 0.956 & 0.918 & 0.613 & 0.230 & 0.092 \\ 
-  Shift=5 & 0.414 & 0.820 & 0.967 & 0.944 & 0.684 & 0.303 & 0.057 \\ 
+Shift=0 & 0.087 & 0.046 & 0.011 & 0.001 & 0.000 & 0.002 & 0.011 \\ 
+  Shift=1 & 0.414 & 0.303 & 0.188 & 0.098 & 0.055 & 0.063 & 0.125 \\ 
+  Shift=2 & 0.736 & 0.614 & 0.364 & 0.111 & 0.017 & 0.008 & 0.028 \\ 
+  Shift=3 & 0.947 & 0.949 & 0.872 & 0.568 & 0.154 & 0.026 & 0.009 \\ 
+  Shift=4 & 0.572 & 0.874 & 0.957 & 0.894 & 0.555 & 0.210 & 0.108 \\ 
+  Shift=5 & 0.506 & 0.869 & 0.971 & 0.927 & 0.597 & 0.232 & 0.037 \\ 
    \hline
@@
   \hline
-Shift=0 & 1.010 & 0.993 & 0.964 & 0.934 & 0.920 & 0.925 & 0.949 \\ 
-  Shift=1 & 1.061 & 1.057 & 1.037 & 1.003 & 0.979 & 0.975 & 0.981 \\ 
-  Shift=2 & 1.034 & 1.040 & 1.035 & 1.005 & 0.969 & 0.952 & 0.958 \\ 
-  Shift=3 & 1.021 & 1.035 & 1.046 & 1.034 & 1.000 & 0.971 & 0.956 \\ 
-  Shift=4 & 0.999 & 1.020 & 1.044 & 1.050 & 1.030 & 1.003 & 0.986 \\ 
-  Shift=5 & 1.005 & 1.027 & 1.052 & 1.063 & 1.050 & 1.021 & 0.989 \\ 
+Shift=0 & 1.001 & 0.982 & 0.952 & 0.924 & 0.912 & 0.918 & 0.948 \\ 
+  Shift=1 & 1.057 & 1.050 & 1.028 & 0.997 & 0.980 & 0.984 & 0.989 \\ 
+  Shift=2 & 1.031 & 1.036 & 1.027 & 0.995 & 0.962 & 0.951 & 0.968 \\ 
+  Shift=3 & 1.022 & 1.035 & 1.043 & 1.025 & 0.991 & 0.962 & 0.948 \\ 
+  Shift=4 & 1.002 & 1.023 & 1.045 & 1.046 & 1.024 & 1.001 & 0.989 \\ 
+  Shift=5 & 1.009 & 1.032 & 1.054 & 1.060 & 1.043 & 1.013 & 0.981 \\ 
    \hline
@@
   \hline
-Shift=0 & 1.244 & 1.223 & 1.187 & 1.150 & 1.133 & 1.140 & 1.169 \\ 
-  Shift=1 & 0.987 & 0.984 & 0.965 & 0.933 & 0.911 & 0.907 & 0.912 \\ 
-  Shift=2 & 1.043 & 1.049 & 1.043 & 1.013 & 0.977 & 0.960 & 0.966 \\ 
-  Shift=3 & 1.044 & 1.059 & 1.071 & 1.058 & 1.024 & 0.993 & 0.978 \\ 
-  Shift=4 & 0.978 & 0.999 & 1.023 & 1.029 & 1.009 & 0.983 & 0.966 \\ 
-  Shift=5 & 0.975 & 0.997 & 1.021 & 1.031 & 1.019 & 0.991 & 0.960 \\ 
+Shift=0 & 1.233 & 1.210 & 1.173 & 1.138 & 1.123 & 1.131 & 1.167 \\ 
+  Shift=1 & 0.983 & 0.977 & 0.956 & 0.928 & 0.912 & 0.916 & 0.920 \\ 
+  Shift=2 & 1.040 & 1.044 & 1.035 & 1.004 & 0.970 & 0.959 & 0.976 \\ 
+  Shift=3 & 1.045 & 1.059 & 1.067 & 1.049 & 1.014 & 0.984 & 0.971 \\ 
+  Shift=4 & 0.981 & 1.002 & 1.023 & 1.024 & 1.003 & 0.980 & 0.968 \\ 
+  Shift=5 & 0.979 & 1.001 & 1.023 & 1.028 & 1.012 & 0.983 & 0.951 \\ 
    \hline

### 2025-06-05 R/M_SSA_utility_functions.R

@@ -234,7 +234,6 @@ compute_mssa_BIP_predictors_func<-function(x_mat,lambda_HP,L,date_to_fit,p,q,ht_
     Sigma<-V_obj$Sigma
     Phi<-V_obj$Phi
     Theta<-NULL
-    
   }
     
 #---------------------------------------

### 2025-06-11 Latex/BIP_predictor.tex
up

--- 2025-06-04 Latex/BIP_predictor.tex
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:00:56 2025
+% Wed Jun 11 12:40:38 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:00:56 2025
+% Wed Jun 11 12:40:38 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:00:58 2025
+% Wed Jun 11 12:40:39 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:00:59 2025
+% Wed Jun 11 12:40:50 2025
 \begin{table}[ht]
@@
   \hline
-Target correlation & 0.69 & 0.73 \\ 
-  HT & 8.67 & 4.73 \\ 
+Target correlation & 0.68 & 0.69 \\ 
+  HT & 6.50 & 3.71 \\ 
    \hline
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:01:02 2025
+% Wed Jun 11 12:41:43 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 0.000 & 0.000 & 0.000 & 0.000 & 0.001 & 0.017 & 0.244 \\ 
-  Shift=1 & 0.000 & 0.000 & 0.000 & 0.000 & 0.000 & 0.003 & 0.026 \\ 
-  Shift=2 & 0.001 & 0.000 & 0.000 & 0.000 & 0.000 & 0.001 & 0.007 \\ 
-  Shift=3 & 0.052 & 0.023 & 0.005 & 0.001 & 0.002 & 0.001 & 0.002 \\ 
-  Shift=4 & 0.370 & 0.240 & 0.100 & 0.015 & 0.006 & 0.004 & 0.002 \\ 
-  Shift=5 & 0.792 & 0.673 & 0.450 & 0.157 & 0.016 & 0.013 & 0.007 \\ 
+Shift=0 & 0.000 & 0.000 & 0.001 & 0.002 & 0.005 & 0.020 & 0.112 \\ 
+  Shift=1 & 0.000 & 0.000 & 0.000 & 0.001 & 0.002 & 0.005 & 0.019 \\ 
+  Shift=2 & 0.000 & 0.000 & 0.000 & 0.001 & 0.001 & 0.003 & 0.006 \\ 
+  Shift=3 & 0.018 & 0.007 & 0.002 & 0.000 & 0.002 & 0.003 & 0.004 \\ 
+  Shift=4 & 0.180 & 0.100 & 0.040 & 0.010 & 0.002 & 0.005 & 0.005 \\ 
+  Shift=5 & 0.579 & 0.432 & 0.261 & 0.108 & 0.027 & 0.008 & 0.009 \\ 
    \hline
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:01:02 2025
+% Wed Jun 11 12:41:53 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 0.090 & 0.060 & 0.025 & 0.006 & 0.002 & 0.001 & 0.031 \\ 
-  Shift=1 & 0.271 & 0.202 & 0.126 & 0.053 & 0.007 & 0.001 & 0.006 \\ 
-  Shift=2 & 0.820 & 0.723 & 0.502 & 0.225 & 0.050 & 0.005 & 0.005 \\ 
-  Shift=3 & 0.878 & 0.952 & 0.906 & 0.678 & 0.251 & 0.049 & 0.026 \\ 
-  Shift=4 & 0.641 & 0.916 & 0.986 & 0.938 & 0.549 & 0.147 & 0.020 \\ 
-  Shift=5 & 0.494 & 0.842 & 0.994 & 0.973 & 0.675 & 0.298 & 0.058 \\ 
+Shift=0 & 0.084 & 0.047 & 0.017 & 0.005 & 0.002 & 0.001 & 0.038 \\ 
+  Shift=1 & 0.230 & 0.160 & 0.094 & 0.032 & 0.005 & 0.001 & 0.004 \\ 
+  Shift=2 & 0.801 & 0.666 & 0.414 & 0.173 & 0.026 & 0.005 & 0.006 \\ 
+  Shift=3 & 0.876 & 0.961 & 0.917 & 0.683 & 0.248 & 0.068 & 0.039 \\ 
+  Shift=4 & 0.370 & 0.819 & 0.990 & 0.953 & 0.533 & 0.117 & 0.030 \\ 
+  Shift=5 & 0.353 & 0.802 & 0.983 & 0.943 & 0.571 & 0.239 & 0.043 \\ 
    \hline
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:01:02 2025
+% Wed Jun 11 12:41:53 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 0.997 & 0.966 & 0.911 & 0.856 & 0.847 & 0.890 & 0.953 \\ 
-  Shift=1 & 1.099 & 1.088 & 1.038 & 0.951 & 0.884 & 0.875 & 0.912 \\ 
-  Shift=2 & 1.082 & 1.099 & 1.092 & 1.032 & 0.952 & 0.906 & 0.904 \\ 
-  Shift=3 & 1.033 & 1.068 & 1.097 & 1.076 & 1.009 & 0.954 & 0.934 \\ 
-  Shift=4 & 1.005 & 1.051 & 1.101 & 1.108 & 1.054 & 0.982 & 0.935 \\ 
-  Shift=5 & 1.006 & 1.054 & 1.104 & 1.121 & 1.092 & 1.031 & 0.969 \\ 
+Shift=0 & 0.973 & 0.937 & 0.888 & 0.850 & 0.855 & 0.900 & 0.956 \\ 
+  Shift=1 & 1.073 & 1.051 & 0.997 & 0.918 & 0.870 & 0.873 & 0.906 \\ 
+  Shift=2 & 1.069 & 1.077 & 1.063 & 1.006 & 0.940 & 0.907 & 0.907 \\ 
+  Shift=3 & 1.020 & 1.045 & 1.066 & 1.048 & 0.997 & 0.957 & 0.941 \\ 
+  Shift=4 & 0.983 & 1.016 & 1.056 & 1.064 & 1.023 & 0.974 & 0.942 \\ 
+  Shift=5 & 0.988 & 1.026 & 1.071 & 1.088 & 1.057 & 1.007 & 0.963 \\ 
    \hline
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:01:06 2025
+% Wed Jun 11 12:42:27 2025
 \begin{table}[ht]
@@
   \hline
-Forecast horizon 1 & 0.752 & 0.715 & 0.749 & 0.752 \\ 
-  Forecast horizon 2 & 0.697 & 0.663 & 0.715 & 0.740 \\ 
-  Forecast horizon 3 & 0.554 & 0.622 & 0.669 & 0.706 \\ 
-  Forecast horizon 4 & 0.486 & 0.624 & 0.706 & 0.737 \\ 
-  Forecast horizon 5 & 0.593 & 0.636 & 0.660 & 0.714 \\ 
+Forecast horizon 1 & 0.752 & 0.715 & 0.746 & 0.747 \\ 
+  Forecast horizon 2 & 0.697 & 0.663 & 0.729 & 0.720 \\ 
+  Forecast horizon 3 & 0.554 & 0.622 & 0.684 & 0.709 \\ 
+  Forecast horizon 4 & 0.486 & 0.624 & 0.680 & 0.708 \\ 
+  Forecast horizon 5 & 0.593 & 0.636 & 0.643 & 0.701 \\ 
    \hline
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 12:01:06 2025
+% Wed Jun 11 12:42:27 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 1.157 & 1.120 & 1.056 & 0.992 & 0.982 & 1.032 & 1.105 \\ 
-  Shift=1 & 1.312 & 1.298 & 1.238 & 1.135 & 1.055 & 1.044 & 1.089 \\ 
-  Shift=2 & 1.187 & 1.205 & 1.198 & 1.132 & 1.045 & 0.994 & 0.991 \\ 
-  Shift=3 & 1.025 & 1.060 & 1.089 & 1.067 & 1.001 & 0.946 & 0.927 \\ 
-  Shift=4 & 1.004 & 1.050 & 1.100 & 1.108 & 1.053 & 0.982 & 0.935 \\ 
-  Shift=5 & 0.978 & 1.025 & 1.073 & 1.090 & 1.062 & 1.003 & 0.942 \\ 
+Shift=0 & 1.128 & 1.086 & 1.030 & 0.985 & 0.992 & 1.044 & 1.109 \\ 
+  Shift=1 & 1.280 & 1.254 & 1.189 & 1.096 & 1.037 & 1.041 & 1.081 \\ 
+  Shift=2 & 1.173 & 1.181 & 1.166 & 1.104 & 1.031 & 0.995 & 0.995 \\ 
+  Shift=3 & 1.012 & 1.037 & 1.058 & 1.040 & 0.989 & 0.949 & 0.934 \\ 
+  Shift=4 & 0.982 & 1.016 & 1.056 & 1.063 & 1.022 & 0.974 & 0.941 \\ 
+  Shift=5 & 0.961 & 0.997 & 1.041 & 1.058 & 1.028 & 0.979 & 0.936 \\ 
    \hline
@@
   \hline
-Shift=0 & 0.112 & 0.062 & 0.019 & 0.002 & 0.000 & 0.001 & 0.006 \\ 
-  Shift=1 & 0.451 & 0.327 & 0.206 & 0.108 & 0.042 & 0.031 & 0.073 \\ 
-  Shift=2 & 0.755 & 0.646 & 0.425 & 0.150 & 0.025 & 0.008 & 0.014 \\ 
-  Shift=3 & 0.936 & 0.952 & 0.901 & 0.664 & 0.229 & 0.046 & 0.017 \\ 
-  Shift=4 & 0.502 & 0.838 & 0.956 & 0.918 & 0.613 & 0.230 & 0.092 \\ 
-  Shift=5 & 0.414 & 0.820 & 0.967 & 0.944 & 0.684 & 0.303 & 0.057 \\ 
+Shift=0 & 0.127 & 0.064 & 0.019 & 0.003 & 0.001 & 0.002 & 0.012 \\ 
+  Shift=1 & 0.439 & 0.314 & 0.194 & 0.093 & 0.033 & 0.017 & 0.036 \\ 
+  Shift=2 & 0.718 & 0.565 & 0.326 & 0.098 & 0.020 & 0.008 & 0.010 \\ 
+  Shift=3 & 0.946 & 0.975 & 0.925 & 0.657 & 0.222 & 0.059 & 0.028 \\ 
+  Shift=4 & 0.327 & 0.699 & 0.964 & 0.942 & 0.619 & 0.237 & 0.102 \\ 
+  Shift=5 & 0.383 & 0.755 & 0.946 & 0.910 & 0.581 & 0.241 & 0.055 \\ 
    \hline
@@
   \hline
-Shift=0 & 1.010 & 0.993 & 0.964 & 0.934 & 0.920 & 0.925 & 0.949 \\ 
-  Shift=1 & 1.061 & 1.057 & 1.037 & 1.003 & 0.979 & 0.975 & 0.981 \\ 
-  Shift=2 & 1.034 & 1.040 & 1.035 & 1.005 & 0.969 & 0.952 & 0.958 \\ 
-  Shift=3 & 1.021 & 1.035 & 1.046 & 1.034 & 1.000 & 0.971 & 0.956 \\ 
-  Shift=4 & 0.999 & 1.020 & 1.044 & 1.050 & 1.030 & 1.003 & 0.986 \\ 
-  Shift=5 & 1.005 & 1.027 & 1.052 & 1.063 & 1.050 & 1.021 & 0.989 \\ 
+Shift=0 & 1.005 & 0.986 & 0.960 & 0.935 & 0.927 & 0.939 & 0.964 \\ 
+  Shift=1 & 1.060 & 1.054 & 1.032 & 0.998 & 0.973 & 0.965 & 0.970 \\ 
+  Shift=2 & 1.028 & 1.029 & 1.020 & 0.993 & 0.963 & 0.950 & 0.952 \\ 
+  Shift=3 & 1.018 & 1.027 & 1.034 & 1.022 & 0.994 & 0.972 & 0.961 \\ 
+  Shift=4 & 0.990 & 1.006 & 1.025 & 1.032 & 1.017 & 0.998 & 0.985 \\ 
+  Shift=5 & 1.000 & 1.016 & 1.037 & 1.046 & 1.032 & 1.008 & 0.985 \\ 
    \hline
@@
   \hline
-Shift=0 & 1.244 & 1.223 & 1.187 & 1.150 & 1.133 & 1.140 & 1.169 \\ 
-  Shift=1 & 0.987 & 0.984 & 0.965 & 0.933 & 0.911 & 0.907 & 0.912 \\ 
-  Shift=2 & 1.043 & 1.049 & 1.043 & 1.013 & 0.977 & 0.960 & 0.966 \\ 
-  Shift=3 & 1.044 & 1.059 & 1.071 & 1.058 & 1.024 & 0.993 & 0.978 \\ 
-  Shift=4 & 0.978 & 0.999 & 1.023 & 1.029 & 1.009 & 0.983 & 0.966 \\ 
-  Shift=5 & 0.975 & 0.997 & 1.021 & 1.031 & 1.019 & 0.991 & 0.960 \\ 
+Shift=0 & 1.237 & 1.215 & 1.183 & 1.152 & 1.141 & 1.156 & 1.187 \\ 
+  Shift=1 & 0.986 & 0.980 & 0.960 & 0.929 & 0.905 & 0.898 & 0.903 \\ 
+  Shift=2 & 1.036 & 1.037 & 1.029 & 1.001 & 0.971 & 0.957 & 0.960 \\ 
+  Shift=3 & 1.042 & 1.051 & 1.058 & 1.045 & 1.018 & 0.995 & 0.983 \\ 
+  Shift=4 & 0.970 & 0.985 & 1.004 & 1.010 & 0.996 & 0.978 & 0.965 \\ 
+  Shift=5 & 0.970 & 0.986 & 1.007 & 1.015 & 1.002 & 0.978 & 0.956 \\ 
    \hline

### 2025-06-20 HWI SSA_27_05_2025_with new poster/05.06.2025/paper.tex
up

--- 2025-05-21 archive/overleaf_latex.tex
@@
 % \usepackage[justification=left,hang]{caption}
-\usepackage[justification=raggedright,hang]{caption}
+\usepackage[justification=raggedrig,hang]{caption}
 \usepackage{subcaption}
@@
 \usepackage{appendix}
+\usepackage{float}
 
@@
 Efficient forecasts are often volatile, limiting their practical usefulness.
-Building on the Smooth Sign Accuracy (SSA) framework, we propose a novel method that integrates Multivariate Smooth Sign Accuracy (M-SSA) to extract smoothed components from leading indicators.
+We propose a novel method that employs the framework of Multivariate Smooth Sign Accuracy (M-SSA) to extract smoothed components from leading indicators.
 This approach enhances the signal-to-noise ratio and allows for systematic control over forecast volatility and timing.
-Applied to German GDP, our method yields smoothed forecasts that match or exceed the accuracy of traditional models, particularly over medium-term horizons.
+Applied to German GDP, our method yields smoothed forecasts that improve the accuracy of traditional models, particularly over medium-term horizons.
 While smoother forecasts may lag slightly, this can be mitigated by adjusting the forecast horizon.
-The findings offer practical tools for forecasters and policymakers operating under adjustment constraints.
+Our findings offer practical tools for forecasters and policymakers subject to adjustment costs or frictions in updating forecasts or implementing policy changes.
 
@@
 \begin{quote}
-     \quad As we parse the incoming information, we are focused on separating the signal from the noise as the outlook evolves.\\
+     \quad ``As we parse the incoming information, we are focused on separating the signal from the noise as the outlook evolves.''\\
     \textit{Speech by Federal Reserve Board Chair Jerome H. Powell, March 07, 2025, \\
@@
 leading indicators with smoothed and filtered components derived from the M-SSA procedure, thereby improving the signal-to-noise ratio in the predictor variables used to forecast GDP.
-We apply this framework to well-established German economic indicators to show its ability to improve forecasts of GDP and GDP trends, especially over mid-term horizons ranging from two to four quarters ahead, using a variety of sample periods and performance metrics.
-
-Surprisingly, we find that smoothing indicator-based forecasts often provides useful leading information for GDP.
+We apply this framework to well-established German economic indicators to show its ability to improve forecasts of GDP and GDP trends using a variety of sample periods and performance metrics.
+
+Surprisingly, we find that smoothing indicator-based forecasts often provides useful leading information for GDP, especially over mid-term horizons ranging from two to four quarters ahead.
 We also show that while smoother GDP forecasts lag efficient MSE forecasts slightly, this can be mitigated by adjustments in the forecast horizon.\footnote{The lag reflects the trilemma highlighted by \cite{McElroy2019} in which forecasters choose between forecast accuracy, timeliness and volatility.}
@@
 
-Section \ref{sec:data} describe our data series, the sample period used, and documents their dynamic behaviour.
+Section \ref{sec:data} describes our data series, the sample period used, and documents their dynamic behaviour.
 Section \ref{sec:direct_forecast} provides benchmark forecasts using simple direct projections of GDP based on our indicators.
-It then demonstrates how in-sample forecast performance may improve when we ``de-noise'' the indicators prior to projection with simple smoothing filters.
+We demonstrate how in-sample forecast performance may improve when we ``de-noise'' the indicators prior to projection with simple smoothing filters.
 Section \ref{sec:mSSA} defines the M-SSA method and shows how its forecast performance compares to that of direct projections as well as how it varies with forecast horizon.
@@
 The analysis employs a set of well-established leading indicators to forecast quarterly German GDP \citep{Drechsel2012financial,Heinisch2018bottom}, including industrial production (IP), the ifo Business Climate Index (ifo\_c), the Economic Sentiment Indicator (ESI), and the term spread between 10-year bond yields and the 3-month EURIBOR rate (spre\_10y\_3m) covering the period 1995 to 2024.
-All series are used at a quarterly frequency, with a particular emphasis on forecast horizons of 2 or more quarters, reflecting medium-term forecasting objectives.
-To ensure comparability and mitigate scale effects, all indicators (except the spread, which is differenced only) are log-differenced and standardized, the latter applied solely for ease of visual inspection.
-Additionally, extreme outliers associated with the COVID-19 pandemic are trimmed to prevent distortion of the underlying signal.
-We take into account the publication lag of the indicators and the GDP series.\footnote{Data revisions are not considered, given that previous studies have shown that their overall effect on forecast performance is minor.}
-While GDP is available with a one-quarter delay, and IP is released with a two-month lag, business surveys and financial data are published without any lag.
-All series are therefore re-aligned to reflect the forecasters' available information, with GDP and IP shifted forward by their respective lags.
-Tables \ref{tab:data_stucture} and \ref{tab:data_stucture_aligned} show the data structure, as published by mid-February 2025.
-
-%\begin{table}[ht]
-%\centering
-%\begin{tabular}{rrrrrr}
-%  \hline
-% & GDP & ip & ifo\_c & ESI & spr\_10y\_3m \\ 
-%  \hline
-%2024-07-01 & & 90.700 & 85.800 & 92.500 & -1.300 \\ 
-%  2024-08-01 &  & 93.200 & 84.900 & 90.900 & -1.400 \\ 
-%  2024-09-01 &902.571   & 91.200 & 83.700 & 89.800 & -1.400 \\ 
-%  2024-10-01 &  & 90.800 & 84.100 & 90.700 & -1.200 \\ 
-%  2024-11-01 &  & 92.200 & 83.800 & 89.300 & -0.900 \\ 
-%  2024-12-01 & 900.761 &  & 82.800 & 86.900 & -0.800 \\ 
-%  2025-01-01 &  &  & 82.500 & 88.100 & -0.200 \\ 
-%   \hline
-%\end{tabular}
-%\caption{Ragged end.  } 
-%\label{tab:data_stucture}
-%\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
+We take into account the publication lag of the indicators and the GDP series.\footnote{Data revisions are not considered, given that previous studies \citep{Heinisch2019} have shown that their overall effect on forecast performance is minor.}
+While GDP is available with a one-quarter delay (published in mid of the next quarter), and IP is released with a one-month lag, business surveys and financial data are published without any lag, see Table \ref{tab:data_stucture} which presents the data structure as of the end of January 2025, just before the publication of the Flash estimate for the previous quarter.\footnote{Starting in the second quarter of 2020, the German Federal Statistical Office published for the first time the gross domestic product (GDP) 30 days after the end of the quarter.} All series are consequently re-aligned to correspond with the information available to forecasters, with GDP and IP variables shifted forward according to their respective lags.
+Table \ref{tab:data_stucture_aligned} displays the dataset as of the end of 2024, with GDP and IP variables temporally aligned at the sample endpoint.
+This configuration reflects the matrix of explanatory variables accessible in Q4-2024 for the purposes of nowcasting or forecasting GDP.
+For illustrative purposes, an additional target column is included on the right side of the table, representing the dependent variable for nowcasting applications.
+In the case of an $h$-step ahead forecast, this column is shifted upwards by $h$ quarters.
+
 \begin{table}[ht]
@@
   \hline
-2024-07-01 & 902.571 & 90.700 & 85.800 & 92.500 & -1.300 \\ 
+2024-07-01 &  & 90.700 & 85.800 & 92.500 & -1.300 \\ 
   2024-08-01 &  & 93.200 & 84.900 & 90.900 & -1.400 \\ 
-  2024-09-01 &  & 91.200 & 83.700 & 89.800 & -1.400 \\ 
-  2024-10-01 & 900.761 & 90.800 & 84.100 & 90.700 & -1.200 \\ 
+  2024-09-01 &902.571  & 91.200 & 83.700 & 89.800 & -1.400 \\ 
+  2024-10-01 &  & 90.800 & 84.100 & 90.700 & -1.200 \\ 
   2024-11-01 &  & 92.200 & 83.800 & 89.300 & -0.900 \\ 
@@
 \end{tabular}
-\caption{Ragged end.  } 
+\caption{Ragged end as of January 2025, before publication of the flash estimate.} 
 \label{tab:data_stucture}
 \end{table}
-% Thu May  8 05:39:20 2025
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrr}
-  \hline
- & GDP & ip & ifo\_c & ESI & spr\_10y\_3m \\ 
-  \hline
-2024-05-01 &  & 93.400 & 84.900 & 90.900 & -1.400 \\ 
-  2024-08-01 &  & 90.700 & 83.700 & 89.800 & -1.400 \\ 
-  2024-09-01 & 902.571 & 93.200 & 84.100 & 90.700 & -1.200 \\ 
-  2024-10-01 &  & 91.200 & 83.800 & 89.300 & -0.900 \\ 
-  2024-11-01 &  & 90.800 & 82.800 & 86.900 & -0.800 \\ 
-  2024-12-01 & 900.761 & 92.200 & 82.500 & 88.100 & -0.200 \\ 
-   \hline
-\end{tabular}
-\caption{Data aligned at sample end.} 
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrrrrc}
+  \hline
+ & GDP & ip & ifo\_c & ESI & spr\_10y\_3m &&Target (nowcast)\\ 
+  \hline
+  2024-07-01 & & 93.400 & 85.800 & 92.500 & -1.300&|& \\ 
+  2024-08-01 & & 90.700&  84.900 & 90.900 & -1.400&| &\\ 
+  2024-09-01 & 901.623 &93.200 & 83.700 & 89.800 & -1.400&| &902.571\\ 
+  2024-10-01 &  &91.200 & 84.100 & 90.700 & -1.200&| &\\ 
+  2024-11-01 &  & 90.800& 83.800 & 89.300 & -0.900&| &\\ 
+  2024-12-01 &902.571   & 92.200  & 82.800 & 86.900 & -0.800&| &?\\ 
+    \hline
+\end{tabular}
+\caption{Columns 1 through 5 constitute the matrix of explanatory variables available in Q4-2024 for the establishment of a GDP nowcast or forecast.
+All series are artificially aligned at the sample endpoint.
+Column 6 represents the target variable for the nowcast in Q4-2024.}
 \label{tab:data_stucture_aligned}
@@
 
+To ensure comparability and mitigate scale effects, all indicators (except the spread, which is differenced only) are log-differenced and standardized, the latter applied solely for ease of visual inspection.
+Additionally, extreme outliers associated with the COVID-19 pandemic are trimmed to prevent distortions in subsequent figures.
  Monthly series are transformed to quarterly series and are displayed in Figure \ref{fig:data}, while Figure \ref{fig:data_lags} offers a close-up view of the financial crisis and the pandemic.\footnote{Monthly series are averaged to quarterly values.} The latter figure illustrates that the real-time GDP and industrial production series are synchronized at dips and peaks (coincident), whereas the other indicators tend to lead somewhat.
 The latter can be leveraged in a multivariate approach to enhance forecasting performance compared to univariate benchmarks.
-
-% \begin{figure}[htbp]
-% \begin{center}
-% \includegraphics[height=3in, width=4.5in]{./Figures/Data.pdf}
-% \caption{Data: standardized log-differences, trimmed to $\pm 3$.
-% \label{data}}
-% \end{center}
-% \end{figure}
-
-% \begin{figure}[htbp]
-% \begin{center}
-% \includegraphics[height=3in, width=4.5in]{./Figures/data_lags.pdf}
-% \end{center}
-% \caption{Data: leads and lags during financial crisis (left) and Pandemic (right).
-% \label{data_lags}}
-% \end{figure}
 
@@
 
-%Hence, we are targeting a series that comprises the publication lag and the forecast horizon.
-
-%[Simon and Katja: not sure if we need the following sentences:]
-%Benchmark for analysis are the historical mean and the AR-predictions
-
-%%%%%%%%%%%%%%%%%
-
-%\footnote{Code is provided at \url{ }.}
-%https://github.com/wiaidp/R-package-SSA-Predictor/blob/main/SSA%20Tutorials/tutorial_7.4_MSSA_component_predictor.R}}
-%Among our results to date, we find that industrial production and economic outlook surveys provide much more valuable information than the yield curve.
-We also find that extreme events, such as the Great Financial Crisis or the COVID-19 pandemic, have an important impact on the forecasts.
-This reflects a common tradeoff facing forecasters: the choice between high accuracy in over the whole period and avoiding rare but large and costly forecast errors in extraordinary situations.
-
 %%%%%%%%%%%%%%%%%%% here comes the GIT-file
@@
 
-Figures \ref{CCF} (full sample) and \ref{CCF_wc} (data prior to the pandemic) show the sample cross-correlation functions (CCF's) between current and lagged GDP and the two survey indicators.
-Both show strong correlations that peak at leads relative to GDP.
+Figures \ref{CCF} (full sample) and \ref{CCF_wc} (omitting the Pandemic period) show the sample cross-correlation functions (CCF's) between current and lagged GDP and the two survey indicators.
+Both figures show strong correlations that peak at leads relative to GDP.
 Results for IP showed no lead, while correlations for the term spread were weaker.
-The Figures also show that the precise strengths of the correlations are influenced by the outliers associated with the COVID-19 pandemic.
+The figures also show that the precise strengths of the correlations are influenced by the outliers associated with the COVID-19 pandemic.
 In the interest of model stability, we exclude data from Q4-2019 to Q4-2020 (pandemic) when modelling the dynamics for the multivariate filters developed below.
@@
 
-We then estimated VARMA models to capture these dynamics, selecting a parsimonious VAR(1) specification.\footnote{We used the \texttt{refVARMA()} procedure in the \cite{MTSpackage} \texttt{MTS} package, which sets VARMA coefficients with t-statistics $< 1.5$ to zero.
-See \cite{tsay2013multivariate}.
-We intend to explore richer dynamic structures in subsequent work.} Figure \ref{ma_inv_BIP} shows the implied responses of GDP to shocks to each variable.\footnote{Shown are the MA coefficients for GDP from the vector moving-average (VMA) model obtained by inversion of the VAR(1) representation (i.e.
+We then estimated VARMA models to capture these dynamics, selecting a parsimonious VAR(1) specification.\footnote{We used the \texttt{refVARMA()} procedure in the \cite{MTSpackage} \texttt{MTS} package, which sets VARMA coefficients with t-statistics $< 1.5$ to zero \citep[see][]{tsay2013multivariate}.
+We intend to explore richer dynamic structures in subsequent work.} Figure \ref{ma_inv_BIP} shows the implied responses of GDP to shocks to each indicator variable.\footnote{Shown are the MA coefficients for GDP from the vector moving-average (VMA) model obtained by inversion of the VAR(1) representation (i.e.
 the impluse-response functions, or IRFs, of GDP to the correlated reduced-form shocks.)}
@@
         \includegraphics[height=3in, width=6in]{./Figures/ma_inv_BIP.pdf}
-        \caption{VMA Coefficients for GDP Implied by VAR(1)\\Estimation with pandemic (left) and without pandemic (right).
+        \caption{VMA coefficients for GDP Implied by VAR(1)\\Estimation with pandemic (left) and without pandemic (right).
         \label{ma_inv_BIP}}\end{center}
@@
 There also appears to be a more modest response to IP shocks after a slightly longer lag, but this vanishes once data after the start of the pandemic is included.\footnote{There is also weaker evidence of a response to the term spread (spr\_10y\_3m) at longer lags.
-This should not be confused with the implied response to a \textit{structural} shock monetary policy shock.} The delayed response of GDP to movements in these various indicators suggests that multivariate models may provide better GDP forecasts than univariate models, a question we pursue in the next section.
+} The delayed response of GDP to movements in these various indicators suggests that multivariate models may provide better GDP forecasts than univariate models, a question we pursue in the next section.
 %, where lagging series (GDP and IP, after accounting for publication lags) depend on indicators leading in relative terms.
@@
 \begin{equation}
-    u_{t+h} = \beta \cdot \mathbf{X}_t + \varepsilon_t
+    u_{t+h} = \beta' \cdot \mathbf{X}_t + \varepsilon_t
     \label{eq:direct_forecast}
@@
     \item $\mathbf{X}_t \equiv [1,u_{t-1}, ip_t, ifo\_c_t, ESI_t, spr\_10y\_3m_t]'$ or their transformed counterparts and $u_{t-1}$ accounts for the publication lag of GDP.\footnote{Including lagged GDP enables the nesting of the AR forecast within the broader modeling framework.}
-    \item $\hat{\beta}$ is estimated by least-squares, omitting observations affected by the pandemic.
+    \item $\hat{\beta}$ is estimated by least-squares, omitting observations affected by the pandemic, i.e. Q4-2019 to Q4-2020.
 \end{itemize}
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
     \begin{center}
@@
         %\includegraphics[height=3in, width=4.5in]{./Figures/direct_wc_all.pdf}
-        \caption{Direct forecasts (red lines) based on all indicators and full sample information (without pandemic): GDP (black line) shifted by zero, one, two, and three quarters.
+        \caption{Direct forecasts (red lines) based on all indicators and full sample information (without pandemic): GDP (black line) shifted (i.e. lagged) by zero, one, two, and three quarters.
         \label{direct_wc}}
@@
 Figures \ref{direct_wc} and \ref{direct_wc_financial_crisis} compare actual and fitted values for GDP at forecast horizons $h = 0, \ldots, 3$, using the four indicators and lagged GDP, without applying any filtering.
-Figure \ref{direct_wc} shows results over the sample without the pandemic, while \ref{direct_wc_financial_crisis} provides a closer look at the fit during the financial crisis.
+Figure \ref{direct_wc} shows results over the sample without the pandemic, while Figure \ref{direct_wc_financial_crisis} provides a closer look at the fit during the financial crisis.
 Table \ref{tab:pvaluedhp} shows p-values for tests of the joint null hypothesis that all coefficients on the indicators are zero (HAC-adjusted Wald test).
 
-\begin{figure}[h]
+\begin{figure}[H]
     \begin{center}
@@
 The results indicate that as the forecast horizon $h$ increases, the turning points (i.e., peaks and troughs) of the direct forecasts become increasingly right-shifted (lagged) relative to the target GDP.
-Additionally, the p-values of the Hac-adjusted Wald test precipitously increase from $h=2$ to $h=3$, see Table \ref{tab:pvaluedhp}, implying that the indicators become statistically insignificant for forecast horizons of more than two quarters.\footnote{
-These findings are confirmed by the relative root mean-square errors comparing direct forecasts against the mean benchmark, see Table \ref{rRMSE_mSSA_direct_mean_without_covid8} in the appendix.}
+Additionally, the p-values of the HAC-adjusted Wald test precipitously increase from $h=2$ to $h=3$ (see Table \ref{tab:pvaluedhp}), implying that the indicators become statistically insignificant for forecast horizons of more than two quarters.\footnote{
+These findings are confirmed by the relative root mean-square errors comparing direct forecasts against the mean benchmark, see Table \ref{rRMSE_mSSA_direct_mean_without_covid8} in the Appendix.}
 
@@
 
-Figure \ref{hp_160} presents both classic two-sided and one-sided HP(160) (denoted HP-C for \textit{Concurrent}) filters (top panels) along with their associated amplitude functions in the frequency domain (bottom panels).
+Figure \ref{hp_160} presents both classic two-sided and one-sided HP(160) (denoted HP-C for \textit{concurrent}) filters (top panels) along with their associated amplitude functions in the frequency domain (bottom panels).
 It confirms that the HP-C filter assigns greater weight to yearly components—relevant in a mid-term forecast context— than the classic quarterly HP filter, while also reducing undesirable high-frequency noise.
@@
 
-Figures \ref{direct_hp_forecasts} and \ref{direct_hp_forecasts_financial_crisis} provide results comparable to those of Figures \ref{direct_wc} and \ref{direct_wc_financial_crisis} but now using indicators smoothed with the HP-C filter.
-As before, the first Figure covers the full sample without the pandemic while the latter provides a closer look at performance during the financial crisis.
-We see that the fitted values lag GDP slightly less and track turning points somewhat more quickly.
+Figures \ref{direct_hp_forecasts} and \ref{direct_hp_forecasts_financial_crisis} provide results comparable to those of Figures \ref{direct_wc} and \ref{direct_wc_financial_crisis} but now using indicators smoothed with the HP-C filter, covering the full sample without the pandemic and providing a closer look at performance during the financial crisis.
+The fitted values lag GDP slightly less and track turning points somewhat more quickly.
 Table \ref{tab:pvaluedhp} shows p-values for tests of the joint null hypothesis that all coefficients $\hat{\beta} = 0$.
-We see that the use of filtered indicators improves their statistical significance at almost all forecast horizons, particularly for $h > 2$ quarters.
+The use of filtered indicators improves their statistical significance at forecast horizons $h > 2$ quarters.
 %Figures \ref{direct_hp_forecasts} (all observations with exclusion of the pandemic) and \ref{direct_hp_forecasts_financial_crisis} (financial crisis).
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
     \begin{center}
         \includegraphics[height=5in, width=6in]{./Figures/direct_hp_forecasts_financial_crisis.pdf}
-        \caption{Shifted GDP (black), direct HP forecast (red) and classic direct forecast (blue) over the curse of the financial crisis: regression based on all indicators and full sample information (without pandemic).
+        \caption{Shifted GDP (black), direct HP forecast (red) and classic direct forecast (blue) over the course of the financial crisis: regression based on all indicators and full sample information (without pandemic).
         \label{direct_hp_forecasts_financial_crisis}}
@@
 \centering
-\begin{tabular}{rrrrrrrr}
-  \hline
- & h= 0 & h= 1 & h= 2 & h= 3 & h= 4 & h= 5 & h= 6 \\ 
-  \hline
-  Unfiltered & 0.028 & 0.000 & 0.001 & 0.910 & 0.509 & 0.091 & 0.178 \\ 
-HP-C filtered & 0.000 & 0.001 & 0.036 & 0.046 & 0.049 & 0.012 & 0.022 \\ 
+\begin{tabular}{lrrrrrr}
+  \hline
+  & h= 1 & h= 2 & h= 3 & h= 4 & h= 5 & h= 6 \\ 
+  \hline
+  Unfiltered & 0.000 & 0.001 & 0.910 & 0.509 & 0.091 & 0.178 \\ 
+HP-C filtered & 0.001 & 0.036 & 0.046 & 0.049 & 0.012 & 0.022 \\ 
    \hline
@@
 Since $z_{it}$  depends on future observations $x_{it+1},x_{it+2},...$, the prediction task involves `tracking' $z_{it}$  using the estimate $y^i_{t}=\sum_{j=1}^ny^i_{jt}$, where $y^i_{jt}=\mathbf{b}^i_j~'\mathbf{x}_{jt}=\sum_{k=0}^{L} b^i_{jk}x_{j,t-k}$ are the outputs of a multivariate (causal or one-sided) filter $\mathbf{B}^i$, with columns $\mathbf{b}^i_{j}$ of fixed length $L$, $j=1,...,5$, assigning weights to the last $L$ observations of the $j$-th indicator $\mathbf{x}_{jt}$ (the fixed-length assumption is merely used for ease of exposition).
-For the sake of clarity, we now omit the superscript index $i$ for all variables, assuming $z_t=z^{i_0}_{t}$ for some fixed $i_0$.% Additionally, we assume a fixed filter length $L$, such that $\mathbf{B}^i_t=\mathbf{B}_t=\mathbf{B}$, which has dimension $L\times 5$.
+For the sake of clarity, we now omit the index $i$ for all variables, assuming $z_t=z_{i_0t}$ for some fixed $i_0$.% Additionally, we assume a fixed filter length $L$, such that $\mathbf{B}^i_t=\mathbf{B}_t=\mathbf{B}$, which has dimension $L\times 5$.
 
@@
 \begin{eqnarray}\label{critssa}
-\left.\begin{array}{c}\rho(z,y,h)\to\max\\
+\left.\begin{array}{c}\rho(z,y)\to\max\\
 s.t. \quad \rho(y,1)=\rho_1\end{array}\right\}
@@
 The parameter $\rho_1$ controls the smoothness of the predictor: higher values of $\rho_1$ favour smoother trajectories of $y_t$.
-Assuming that the process $x_t$ has mean zero, \cite{Wildi2024} establishes a connection between $\rho(y,1)$ and the expected duration between consecutive `zero-crossings' (sign changes) of $y_t$, the ``holding time'', denoted by $ht(y)$:
+Assuming that the process $x_t$ has mean zero, \cite{Wildi2024} establishes a connection between $\rho(y,1)$ and the expected duration between consecutive `zero-crossings' (sign changes) of $y_t$, the `holding time', denoted by $HT(y)$\footnote{Formally, the precise relationship between the expected duration between sign changes and the lag-one autocorrelation of the predictor is derived under the assumption of Gaussian time series.
+However, \cite{Wildi2024} demonstrates that this relationship remains robust even when deviations from Gaussianity occur (central limit theorem applied to lowpass-filtered series).}:
 \begin{eqnarray}\label{ht}
-ht(y)=\frac{\pi}{\arccos(\rho(y,1))}.
+HT(y)=\frac{\pi}{\arccos(\rho(y,1))}.
 \label{eq:ht_arccos}
 \end{eqnarray}
-Since \ref{eq:ht_arccos} is strictly monotonic, the SSA criterion can be reformulated as the optimization problem:
+Since eq. \ref{eq:ht_arccos} is strictly monotonic, the SSA criterion can be reformulated as the optimization problem:
 \begin{eqnarray}\label{critssaht}
 \left.\begin{array}{c}\rho(z,y)\to\max\\
-\frac{\arccos(\rho(y,1))}{\pi}=1/ht_1\end{array}\right\},
+\frac{\arccos(\rho(y,1))}{\pi}=1/HT_1\end{array}\right\},
 \end{eqnarray}
-where the smoothness parameter $1/ht_1$ expresses the expected rate of sign changes of the predictor.\footnote{Formally, the exact relationship between the rate of sign changes and the lag-one autocorrelation of the predictor assumes Gaussian time series, but \cite{Wildi2024} demonstrates robustness to deviations from this assumption.} \\
-
+where the smoothness parameter $1/HT_1$ expresses the expected rate of sign changes of the predictor. \\
+
+Eq.
 \eqref{critssaht} has several important implications.
@@
 This ambiguity can be resolved by assuming an arbitrary scale and level for $y_t$ (standardization).
-Alternatively, a mean-square error norm (MSE) may be substituted for the target correlation, as noted in \cite{Wildi2025}.
-In this case, the classic MSE predictor $y_{t,MSE}$ is obtained as a solution to the SSA criterion by insertion of $1/ht_1:=1/ht_{MSE}$ in the constraint, where $ht_{MSE}$ represents the holding time of $y_{t,MSE}$.
-%If the MSE predictor is noisy, which is often the case in applications, one can select $ht_1>ht_{MSE}$  in the constraint to address the rate of noisy crossings or false alarms.
+Alternatively, a mean square error norm (MSE) can be substituted for the target correlation, as noted in \cite{Wildi2025}.
+In this case, the classic MSE predictor $y_{t,MSE}$ is obtained as a solution to the SSA criterion by insertion of $1/HT_1:=1/HT_{MSE}$ in the constraint, where $HT_{MSE}$ represents the holding time of $y_{t,MSE}$.
+%If the MSE predictor is noisy, which is often the case in applications, one can select $HT_1>HT_{MSE}$  in the constraint to address the rate of noisy crossings or false alarms.
 Second, the concept can be extended to a multivariate framework, denoted M-SSA, by generalizing the correlation functions as detailed in \cite{Wildi2025}. 
@@
 Lastly, when $z_t$ is GDP growth, sign changes are indicative of expansions and recessions.
-Unfortunately, classic predictors often generate excessively many false (`noisy') sign changes due to high-frequency leakage as discussed above.
+Unfortunately, classic predictors often generate excessively many false (`noisy') sign changes due to high-frequency leakage, as discussed above.
 %as illustrated by the amplitude function in Figure \eqref{hp_160} (bottom right panel) and further elaborated upon by \cite{Wildi2025}. 
- \cite{Wildi2025} shows that the SSA criterion helps control this phenomenon.
-Using the equivalent dual formulation of criterion \eqref{critssaht}, the author demonstrates that the SSA solution exhibits the smallest rate of zero-crossings of any (linear) predictor with the same target correlation.
-Below, we therefore use the M-SSA framework to construct a `smooth' predictor $y_t$ of $z_t$.
+ \cite{Wildi2025} shows that the SSA criterion helps control this phenomenon,  see Fig.\ref{mssa_msse_zc} for illustration.
+Using the equivalent dual formulation of the criterion \eqref{critssaht}, the author demonstrates that the SSA solution exhibits the lowest zero-crossing rate of any (linear) predictor with the same target correlation.
+In the following, we therefore use the M-SSA framework to construct a `smooth predictor $y_t$ to forecast $z_t$.
 %In a second stage, serving as a generalization and refinement of the univariate HP-C utilized in the direct HP forecast.
 
-\subsection{Example: Out of Sample M-MSE vs M-SSA Nowcasts}
+\subsection{Example: Out of sample M-MSE vs. M-SSA Nowcasts}
 
 For illustration, consider the problem of nowcasting HP-filtered GDP, denoted by HP-GDP, where the target $z_t$ is generated using the two-sided HP(160) filter examined previously.
-We compare the classic multivariate mean squared error (MSE) predictor (denoted by M-MSE) with the M-SSA predictor where we set $ht_1=1.5ht_{MSE}$.
+We compare the classic multivariate mean squared error (MSE) predictor (denoted M-MSE) with the M-SSA predictor where we set $HT_1=1.5\times HT_{MSE}$.
 This constraint ensures that the M-SSA predictor produces roughly 33$\%$ fewer sign changes than M-MSE in large samples.
@@
 Both predictors use the same underlying VAR(1) model introduced in Section \ref{sec:direct_forecast} and estimated on data up to January 2008.
-This ensures a lengthy out-of-sample period that encompasses significant events such as the financial crisis, the sovereign debt crisis, and the COVID-19 pandemic.\footnote{Further details and background information are available in \cite{Wildi2025}.
-All computations are performed utilizing the M-SSA package, as documented in \cite{Wildi2025}.
+This ensures a lengthy out-of-sample period that includes significant events such as the financial crisis, the sovereign debt crisis, and the COVID-19 pandemic\footnote{Further details and background information are available in \cite{Wildi2025}.
+All calculations are performed using the M-SSA package, as documented in \cite{Wildi2025}.
 \url{https://github.com/wiaidp/R-package-SSA-Predictor}}
@@
 
+\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/mssa_msse_zc.pdf}\caption{Sign changes of the M-MSE (green) and M-SSA (blue) nowcasts towards the financial crisis, with sign changes of the M-MSE marked by green vertical lines.
+The sample holding time (HT) of each filter corresponds to the average duration between its consecutive zero-crossings.
+For each filter, its sample HT converges to the respective expected HT asymptotically.
+Here, we impose  $HT_{MSSA}=1.5\times HT_{MSE}$: in the long run, we expect the M-MSE to exhibit approximately $50\%$ more sign changes than the M-SSA.\label{mssa_msse_zc}}\end{center}\end{figure}
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
@@
 The resulting filtered series, presented in Figure \ref{mssa_msse_now}, are standardized to enable direct comparisons.
-The M-SSA nowcast is slightly smoother than that of the M-MSE, exhibiting approximately one-third fewer zero-crossings.
-Table \ref{corhtnow} compares empirical holding times and target correlations for both nowcasts.
-Increased smoothness achieved by the M-SSA comes at the cost of a slight reduction (0.68 vs 0.69) in target correlation.
+The M-SSA nowcast is slightly smoother than that of the M-MSE, exhibiting approximately one-third fewer sign-changes.
+Fig.\ref{mssa_msse_zc} compares both nowcasts in relation to the financial crisis, highlighting that the classic (MSE-based) design tends to produce unwanted 'noisy' zero-crossings at (or in the vicinity of) the transitions between recessions and expansions.
+Further, Table \ref{corhtnow} compares sample target correlations and HTs of the two nowcasts.
+Increased smoothness, i.e.
+holding time increases from 3.7 to 6.5 quarters, achieved by the M-SSA comes at the cost of a slight reduction (0.68 vs 0.69) in target correlation.
 This highlights the inherent trade-off—often referred to as the prediction dilemma—between smoothness (zero-crossing rate) and predictive accuracy (target correlation) within the M-SSA criterion.
-
-The results validate that the smoothing constraint incorporated into the multivariate extension of Criterion \eqref{critssa} effectively manages the zero-crossings of the predictor: sample estimates of the holding times (HTs) align with the expected values derived from Equation \eqref{ht}.\footnote{Effective convergence of sample estimates to their theoretical expected values —assuming the true data-generating process is known— can be validated using the M-SSA package.
+The results validate that the smoothing constraint incorporated into the multivariate extension of criterion \eqref{critssa} effectively manages the sign-changes of the predictor: sample estimates of the holding times align with the expected values derived from Eq.~\eqref{ht}.\footnote{Effective convergence of sample estimates to their theoretical expected values —assuming the true data-generating process is known— can be validated using the M-SSA package.
 This validation is typically performed through analyses of very long samples of simulated (multivariate) data, ensuring the consistency of the estimation procedures under idealized conditions.} An increase of $50\%$ in the HT relative to the M-MSE predictor has only a modest effect on the target correlation, suggesting that this trade-off could be beneficial in our application.
@@
 Next, we extend the previous example to consider forecasts of HP-GDP.
-The only change this requires is to change $h$ to $h=0,...,6$ quarters.
+The only change this requires is to change $h$ to $h=1,...,6$ quarters.
 %The previously described nowcast example can be systematically extended to generate forecasts of the target variable.
@@
 We also see increased weight on the leading indicators relative to that on GDP.\footnote{Weights on IP and ESI are indistinguishable from zero.} Additionally, a phase shift associated with the filter applied to GDP is observed (black line, right panel), reflecting the cyclical characteristics inherent to the HP filter.
-While this phase effect might cause an effective sign change in the GDP filter output as $h$ increases—something to be avoided in this context—the low-pass filters assigned to the additional 'leading' indicators (green and violet lines, right panel) continue to effectively track the low-frequency components of HP-GDP.
+While this phase effect might cause an effective sign change in the GDP filter output as $h$ increases—something to be avoided in this context—the low-pass filters assigned to the additional  `leading' indicators (green and yellow lines, right panel) continue to effectively track the low-frequency components of HP-GDP.
 The combination of the phase effect, applied to GDP, and level-tracking through these additional indicators enables more refined and effective tracking of the target compared to univariate forecasts, which rely solely on the phase-effect of the filter to look ahead.
 
-\begin{figure}[h]
+\begin{figure}[H]
     \begin{center}
         \includegraphics[width=0.85\textwidth]{./Figures/bk_h.pdf}
-        \caption{Multivariate M-SSA: $h=0$ (nowcast, left) and $h=4$ (one year forecast, right).
+        \caption{Multivariate M-SSA filter weights: $h=0$ (nowcast, left) and $h=4$ (one year forecast, right).
         \label{bk_h}}
@@
 All series are standardized for ease of comparison.
-Forecasts using the M-SSA approach consistently shift leftwards $h$ increases; this shift appears to be consistent across all levels, including the peaks and troughs of the series.
+Forecasts using the M-SSA approach consistently shift leftwards as $h$ increases; this shift appears to be consistent across all levels, including the peaks and troughs of the series.
 Conversely, the positions of peaks and troughs in the HP-C series are largely unaffected by increases in $h$.
@@
 
-\subsection{Hit Rate vs. False Alarm Rate}
-
-%As shown in Wildi (2025), the SSA criterion's objective function—the target correlation $\rho(y,z,h)$—can be linked to the sign accuracy of the predictor $SA(y_t):=\text{P}\left(\text{sign}(z_{t+h}) = \text{sign}(y_t)\right)$, i.e., the probability that the signs of the predictor $y_t$ and the target $z_{t+h}$ agree:
-%\begin{eqnarray}\label{sig_acc}
-%SA(y_t)=0.5+\frac{\arcsin(\rho(y,z,h))}{\pi}.
-%\end{eqnarray}
-%Due to the strict monotonicity of the arcsin function, maximizing the target correlation also maximizes the sign accuracy of the predictor.\footnote{Equation \ref{sig_acc} holds exactly for Gaussian processes.
-Due to the central limit theorem, the relationship holds approximately for low-pass filtered non-Gaussian series, see Wildi (2025).}
-\textbf{Remarks:}
-\begin{itemize}
-\item Alternative to MSE, target correlation, holding-time: hit-rate, false alarm. 
-\item We can link objective function (target correlation) with sign accuracy, see Wildi (2025). BUT: risk that reviewer asks for comparison with logit or whatever sign-forecast tool.
-\item Context ROC/AUC  is missing.
-\end{itemize}
-For the ROC curves, we consider predictors of GDP based on both original (unfiltered) and filtered indicators. For the direct forecast, we use the regression
-\[
-GDP_{t+shift}=\beta_0^{shift}+\sum_{i=1}^5\beta_i^{shift}x_{ti}+\epsilon_{t}^{shift,direct}=\hat{GDP}_{t+shift}^{direct}+\epsilon_{t}^{shift,direct},
-\]
-where $x_{ti}$ designates the original indicators. 
-For the filtered indicators, we consider similar regression equations:
-\[GDP_{t+shift}=\beta_0^{shift,h,filt}+\sum_{i=1}^5\beta_i^{shift,h,filt}y_{ti}^{h,filt}+\epsilon_{t}^{h,shift,filt}=\hat{GDP}_{t+shift}^{h,filt}+\epsilon_{t}^{h,shift,filt},\]
-where $y_{ti}^{h,filt}$ represents the $i$-th filtered indicator optimized at forecast horizon $h$, based on either HP-C, M-MSE, or M-SSA. \\
-\textbf{Remarks:}
-\begin{itemize}
-\item I'm computing all combinations of shift and h. 
-\item But we could emphasize the diagonal only (h=shift): simplifies notation/explanation.
-\item The table of AUCs (below) is based on diagonal only
-\end{itemize}
- ROCs are based on the comparison of signs of the target $GDP_{t+shift}$ and of predictors $\hat{GDP}_{t+shift}^{h,filt}$ (remark: we are using standardized GDP but the regressions account automatically for scales or level shifts, i.e., the results apply directly to original GDP).
-For simplicity of exposition, we present results based on the entire sample, excluding the pandemic period.
-Fig.\ref{ROC_GDP_shift_1_4} compares ROCs of the predictors for small and mid-term forecast horizons and Table \ref{auc} presents AUC (area under the curve) for forecast horizons 1-5 quarters ahead.
-\begin{figure}[h]\begin{center}\includegraphics[height=3in, width=5in]{./Figures/ROC_GDP_shift_1_4.pdf}\caption{Hit rate vs.
-false alarm rate of the predictors at small (one quarter ahead:left panel) and yearly forecast horizons (right panel).\label{ROC_GDP_shift_1_4}}\end{center}\end{figure}
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 10:06:12 2025
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrr}
-  \hline
- & Direct forecast & Direct HP forecast & M-MSE & M-SSA \\ 
-  \hline
-Forecast horizon 1 & 0.752 & 0.715 & 0.746 & 0.747 \\ 
-  Forecast horizon 2 & 0.697 & 0.663 & 0.729 & 0.720 \\ 
-  Forecast horizon 3 & 0.554 & 0.622 & 0.684 & 0.709 \\ 
-  Forecast horizon 4 & 0.486 & 0.624 & 0.680 & 0.708 \\ 
-  Forecast horizon 5 & 0.593 & 0.636 & 0.643 & 0.701 \\ 
-   \hline
-\end{tabular}
-\caption{AUCs of predictors against GDP at various forecast horizons.} 
-\label{auc}
-\end{table}At forecast horizons less than or equal to two quarters, the direct forecast challenges the filters.
-However, for horizons $h>2$, the filters outperform the direct approach: the univariate HP-C design is surpassed by the classic multivariate MSE filter, which is in turn dominated by the M-SSA.
-These results confirm that controlling the rate of sign changes of the predictor, while maximizing its target correlation through M-SSA can enhance predictor performance in terms of sign accuracy and false alarm rate.
-
-\newpage
-
 \section{M-SSA Component Predictor}\label{sec:mSSA_component}
@@
 To achieve this, we use the above M-SSA outputs $y_t^i, \: i = 1, \ldots, 5$ (referred to as M-SSA components) as predictors for forward-shifted GDP.
-While we keep the M-SSA fixed, as based on data up to January 2008, we implement an expanding window starting from January 2008 for deriving the component weights, which are up-dated on a quarterly basis.
+While we keep the M-SSA fixed, as based on data up to Q4-2007, we implement an expanding window starting from Q1-2008 for deriving the component weights, which are updated on a quarterly basis. 
 %(this contrasts with Section \ref{sec:direct_forecast}, which was based on the full sample information). 
-All the results that we present here exclude the pandemic; results including the pandemic are in the appendix.\\
-
-\newpage
+All the results that we present here exclude the pandemic; results including the pandemic are in the Appendix.\\
 
@@
 
-Let $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$ denote the M-SSA predictor of HP-GDP at forecast horizon $h$. A predictor of GDP can be obtained from the simple regression equation
+Let $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$ denote the M-SSA predictor of HP-GDP for forecast horizon $h$. A predictor of GDP can be obtained from the simple regression equation
 \begin{equation}
@@
 where the regression parameters $\hat{\beta}_0(shift,h,t),\hat{\beta}_1(shift,h,t)$ are estimated using data up to time point $t-1$ (expanding window).
-Since M-SSA is based on data up to January 2008, $\hat{\textrm{GDP}}_{t+shift}(h)$ constitutes an out-of-sample predictor of $\textrm{GDP}_{t+shift}$ for all positive $shift\geq 0$ and all $t$ larger than January 2008.
+Since M-SSA is based on data up to Q4-2007, $\hat{\textrm{GDP}}_{t+shift}(h)$ constitutes an out-of-sample predictor of $\textrm{GDP}_{t+shift}$ for all positive $shift\geq 0$ and all $t$ larger than Q4-2007.
 Define the out-of-sample forecast error
@@
 \end{equation}
-for observations after January 2008: the corresponding HAC-adjusted p-values are reported in Table \ref{p_val_wc}.\footnote{Statistical significance is influenced by the relatively short out-of-sample span, of length 17 years.} The MSE forecast performance of $\hat{\textrm{GDP}}_{t+shift}(h)$ can be benchmarked against the mean predictor (see Table \ref{rRMSE_mSSA_comp_mean_wc}) as well as against the direct forecast (see Table \ref{rRMSE_mSSA_comp_direct_without_covid7}).
+for observations after Q4-2007: the corresponding HAC-adjusted p-values are reported in Table \ref{p_val_wc}.\footnote{Statistical significance is influenced by the relatively short out-of-sample span, of length 17 years.} The MSE forecast performance of $\hat{\textrm{GDP}}_{t+shift}(h)$ can be benchmarked against the mean predictor (see Table \ref{rRMSE_mSSA_comp_mean_wc}) as well as against the direct forecast (see Table \ref{rRMSE_mSSA_comp_direct_without_covid7}).
 For consistency, both benchmark predictors are also quarterly updated out-of-sample designs.\\
@@
 
+\subsection{Hit Rate vs. False Alarm Rate}
+
+To this point, indicators have been assessed by the extent to which they correlate with or lead movements in the target variable.
+An alternative approach would to be consider their ability to predict the correct ``type'' of outcome, e.g.
+whether growth will be stronger or weaker than average, or whether the economy will be in an expansion or a recession.
+This may be of particular interest to policymakers whose primary concern is to intervene ``in the right direction'' so as to stabilize the economy overall.
+We therefore limit our analysis to binary classifiers, focussing in particular on the ROC curve and the AUC.\footnote{For example, see \cite{YangLahiriPagan2024ROC} for a recent discussion of the ROC, the AUC and their relation to selected other binary classifiers.
+The discussion below follows their notation.}
+
+Consider a binary random variable $Z \in \{0,1\}$ as our target, and a continuous variable $Y$. Our predictor for $Z$ is given by the function
+
+\begin{equation}
+    \mathbf{1}(Y \leq \tau) \equiv I(\tau)
+\end{equation}
+
+where $\mathbf{1}()$ is the indicator function.\footnote{The value of the indicator function is equal to 1 when the condition in its argument is satisfied; otherwise the value of the function is equal to 0.}
+
+The joint distribution of $Z$ and $Y$ allows us to define the Hit Rate $H(\tau)$ and the False Alarm Rate $F(\tau)$ as
+
+\begin{align}
+    H(\tau) &\equiv Pr( Y \leq \tau | Z = 1)\\
+    F(\tau) &\equiv Pr( Y \leq \tau | Z = 0)
+\end{align}
+
+The usefulness of a given variable $Y$ as a predictor will obviously depend on the value of $\tau$ selected, which in turn may be influenced by the relative importance attached to $H(\tau)$ and $F(\tau)$.
+To compare the range of possibilities associated with various choices for $Y$, we may plot on the unit square $F(\tau)$ vs $H(\tau)$ for all values of $\tau$ in the range of $Y$.
+This is called the Receiver Operating Characteristic (ROC) curve.
+
+An ideal predictor would be one where $H(\tau) = 1$ and $F(\tau) = 0 \; \forall \tau $ (i.e.
+$Y > \tau$ whenever $Z=0$ and $Y \leq \tau$ whenever $Z=1$.) In this ideal case, the ROC curve would consist of the extreme left and top edges of the plot.
+Another extreme case would be one where $\{Y,Z\}$ are independent, so $H(\tau) = F(\tau) \; \forall \tau$.
+In this case, the ROC curve would simply be the 45-degree line through the origin.
+When comparing two indicators $Y_a$ and $Y_b$, if the ROC curve of $Y_a$ lies strictly above and to the left of that of $Y_b$, then we would always prefer to use $Y_a$ as our predictor in place of $Y_b$ since for any choice of $\tau$ it will always give us a superior trade-off between $H(\tau)$ and $F(\tau)$.
+When the ROC curves for two indicators cross, however, then our preferred indicator will depend on the relative importance that we attach to the hit and false alarm rates.\footnote{As shown in Wildi (2025), the sign accuracy can be related to the target correlation in the objective function of the (M-)SSA criterion; additionally, the rate of false alarms depends on the holding time (HT), so that the tradeoff underlying the ROC curve is to some extent captured by the M-SSA criterion.}
+
+A popular approximate measure of the overall performance of an indicator $Y$ that abstracts from the choice of $\tau$ is the Area Under the ROC Curve (AUC).
+Because the ROC is defined on the unit square, the maximum value of the AUC = 1.
+Above we saw that when $\{Y,Z\}$ are independent, the ROC curve lies on the 45-degree line, so the area under the curve is then 0.5.
+The range of AUCs between 1 and 0.5 therefore provides some quantitative guidance as to which predictor variables $Y$ have the potential to improve the prediction of $Z$.\footnote{AUCs $< 0.5$ suggest that $-Y$ may give more useful predictions than $Y$.}.
+
+%As shown in Wildi (2025), the SSA criterion's objective function—the target correlation $\rho(y,z,h)$—can be linked to the sign accuracy of the predictor $SA(y_t):=\text{P}\left(\text{sign}(z_{t+h}) = \text{sign}(y_t)\right)$, i.e., the probability that the signs of the predictor $y_t$ and the target $z_{t+h}$ agree:
+%\begin{eqnarray}\label{sig_acc}
+%SA(y_t)=0.5+\frac{\arcsin(\rho(y,z,h))}{\pi}.
+%\end{eqnarray}
+%Due to the strict monotonicity of the arcsin function, maximizing the target correlation also maximizes the sign accuracy of the predictor.\footnote{Equation \ref{sig_acc} holds exactly for Gaussian processes.
+Due to the central limit theorem, the relationship holds approximately for low-pass filtered non-Gaussian series, see Wildi (2025).}
+
+ROC curves are generated by evaluating the relationship between the sign of the target variable $GDP_{t+shift}$  and events where the forecast $\hat{GDP}_{t+shift}(h)$ exceeds a threshold $\tau$.
+The threshold $\tau$ is varied smoothly across the range of the predictor space to construct the ROC curve.\footnote{Although we are using standardized GDP, the regressions automatically account for scales and level shifts, so the results directly apply to the original GDP.} We examined all pairwise combinations of $h$ and $shift$, but for simplicity of exposition, the results below focus on the diagonal case where $h=shift$.
+For clarity, these results are based on the entire sample, excluding the pandemic period.
+Fig.\ref{ROC_GDP_shift_1_4} compares ROCs of the predictors for small and mid-term forecast horizons and Table \ref{auc} presents AUC (area under the curve) for forecast horizons 1-5 quarters ahead.
+\begin{figure}[h]\begin{center}\includegraphics[height=3in, width=5in]{./Figures/ROC_GDP_shift_1_4.pdf}\caption{Hit rate vs.
+false alarm rate of the predictors at small (one quarter ahead:left panel) and yearly forecast horizons (right panel).\label{ROC_GDP_shift_1_4}}\end{center}\end{figure}
+% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Wed May 21 10:06:12 2025
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrr}
+  \hline
+ & Direct forecast & Direct HP forecast & M-MSE & M-SSA \\ 
+  \hline
+Forecast horizon 1 & 0.752 & 0.715 & 0.746 & 0.747 \\ 
+  Forecast horizon 2 & 0.697 & 0.663 & 0.729 & 0.720 \\ 
+  Forecast horizon 3 & 0.554 & 0.622 & 0.684 & 0.709 \\ 
+  Forecast horizon 4 & 0.486 & 0.624 & 0.680 & 0.708 \\ 
+  Forecast horizon 5 & 0.593 & 0.636 & 0.643 & 0.701 \\ 
+   \hline
+\end{tabular}
+\caption{AUCs of predictors against GDP at various forecast horizons.} 
+\label{auc}
+\end{table}At forecast horizons less than or equal to two quarters, the direct forecast challenges the filters.
+However, for horizons $h>2$, the filters outperform the direct approach: the univariate HP-C design is surpassed by the classic multivariate MSE filter, which is in turn dominated by the M-SSA.
+These results confirm that controlling the rate of sign changes of the predictor, while maximizing its target correlation through M-SSA can enhance predictor performance in terms of sign accuracy and false alarm rate.
+
 \section{Revisions}\label{sec:revisions}
@@
 Moreover, the smoothing effect of filters mitigates the effect of revisions when compared to direct forecasts.}, or from updates to predictor parameters.
-The latter can be further separated into changes in VAR parameters and regression weights.
+The latter can be further separated into changes in the regression weights and VAR parameters.
 
@@
 As a confirmation, Figure \ref{revisions2} displays the evolution of the intercept and regression weights over time: with increasing sample size, these estimates seem to converge to fixed points, indicating both the stationarity of the process and the consistency of the estimates.
-(Note, however, the slightly elevated values of the regression weight assigned to the M-SSA component during the financial crisis and the pandemic, suggesting a stronger link between the target and the predictor).
+%(Note, however, the slightly elevated values of the regression weight assigned to the M-SSA component during the financial crisis and the pandemic, suggesting a stronger link between the target and the predictor).
 In contrast, more complex designs based on multiple M-SSA components can exhibit substantial fluctuations in the regression parameters, sometimes changing signs.
@@
 For consistency, we excluded the entire pandemic episode from the data.
-Figure \ref{up_dated_ma_inv_multi_ip} compares the MA-inversions (impulse responses) of the GDP equation derived from the 'old' and the updated VAR models: unlike the former, the latter incorporates ESI as an additional explanatory variable for GDP.
+Figure \ref{up_dated_ma_inv_multi_ip} compares the MA-inversions (impulse responses) of the GDP equation derived from the `old' and the updated VAR models: unlike the former, the latter incorporates ESI as an additional explanatory variable for GDP.
 Figure \ref{bk_2008_all} illustrates the resulting effect on the M-SSA filter targeting HP-GDP at forecast horizon $h=4$.
-The updated filter (right panel) assigns more weight to the additional 'leading' indicators relative to GDP.
+The updated filter (right panel) assigns more weight to the additional `leading' indicators relative to GDP.
 Finally, Figure \ref{revisions3} compares the M-SSA GDP predictors.
 To isolate the effect of the regression revision, the figure displays standardized series.
-The update to the VAR model influences the left-shift of the predictor (blue line), which appears slightly 'faster' than the original predictor (red line).
+The update to the VAR model influences the left-shift of the predictor (blue line), which appears slightly `faster' than the original predictor (red line).
 This suggests that the out-of-sample performance of the M-SSA component predictor may be better than reported in Section \ref{oosp}, which is based on the fixed and increasingly outdated version of M-SSA.
@@
 The resulting predictor tends to outperform the classic direct forecasts in terms of statistical significance, mainly due to a slight left-shift noticeable at sharp recessions such as the Great Financial Crisis or Covid19-Pandemic.
-However, despite the filtering, the predictor remains somewhat noisy, potentially producing excessive noisy sign changes at the transitions between recessions and expansions.
+However, despite the filtering, the predictor remains somewhat noisy, potentially producing excessive noisy sign changes at (or in the vicinity of) the transitions between recessions and expansions.
 This issue is partly due to noise leakage inherent in the one-sided HP filter.
@@
 
-To address these problems, we propose a multivariate extension of SSA based on \cite{Wildi2024}.
+To address these problems, we propose a multivariate extension of SSA framework based on \cite{Wildi2024}.
 Unlike the univariate HP filter, the M-SSA can leverage cross-sectional information from indicators leading GDP in real time while controlling for the rate of zero-crossings of the predictor.
@@
   \hline
-Shift=0 & 1.144 & 1.102 & 1.045 & 0.999 & 1.006 & 1.059 & 1.125 \\ 
-  Shift=1 & 1.238 & 1.213 & 1.151 & 1.060 & 1.004 & 1.007 & 1.047 \\ 
-  Shift=2 & 1.178 & 1.186 & 1.171 & 1.108 & 1.035 & 0.999 & 0.999 \\ 
-  Shift=3 & 1.011 & 1.035 & 1.057 & 1.039 & 0.988 & 0.948 & 0.933 \\ 
-  Shift=4 & 1.010 & 1.045 & 1.086 & 1.094 & 1.051 & 1.002 & 0.968 \\ 
-  Shift=5 & 0.982 & 1.019 & 1.064 & 1.081 & 1.051 & 1.000 & 0.957 \\ 
+Shift=0 & 1.128 & 1.086 & 1.030 & 0.985 & 0.992 & 1.044 & 1.109 \\ 
+  Shift=1 & 1.280 & 1.254 & 1.189 & 1.096 & 1.037 & 1.041 & 1.081 \\ 
+  Shift=2 & 1.173 & 1.181 & 1.166 & 1.104 & 1.031 & 0.995 & 0.995 \\ 
+  Shift=3 & 1.012 & 1.037 & 1.058 & 1.040 & 0.989 & 0.949 & 0.934 \\ 
+  Shift=4 & 0.982 & 1.016 & 1.056 & 1.063 & 1.022 & 0.974 & 0.941 \\ 
+  Shift=5 & 0.961 & 0.997 & 1.041 & 1.058 & 1.028 & 0.979 & 0.936 \\ 
    \hline
@@
   \hline
-RRMSE & 0.855 & 0.874 & 0.919 & 1.023 & 0.985 & 1.010 \\ 
+rRMSE & 0.862 & 0.838 & 0.912 & 1.008 & 1.001 & 1.028 \\  
    \hline
@@
   \hline
-Shift=0 & 1.210 & 1.188 & 1.157 & 1.127 & 1.116 & 1.130 & 1.161 \\ 
-  Shift=1 & 0.982 & 0.977 & 0.957 & 0.925 & 0.902 & 0.895 & 0.899 \\ 
-  Shift=2 & 1.053 & 1.054 & 1.046 & 1.018 & 0.987 & 0.973 & 0.975 \\ 
-  Shift=3 & 1.046 & 1.056 & 1.062 & 1.050 & 1.022 & 0.999 & 0.988 \\ 
-  Shift=4 & 0.969 & 0.984 & 1.003 & 1.010 & 0.996 & 0.977 & 0.964 \\ 
-  Shift=5 & 0.982 & 0.998 & 1.019 & 1.027 & 1.014 & 0.990 & 0.967 \\ 
+Shift=0 & 1.237 & 1.215 & 1.183 & 1.152 & 1.141 & 1.156 & 1.187 \\ 
+  Shift=1 & 0.986 & 0.980 & 0.960 & 0.929 & 0.905 & 0.898 & 0.903 \\ 
+  Shift=2 & 1.036 & 1.037 & 1.029 & 1.001 & 0.971 & 0.957 & 0.960 \\ 
+  Shift=3 & 1.042 & 1.051 & 1.058 & 1.045 & 1.018 & 0.995 & 0.983 \\ 
+  Shift=4 & 0.970 & 0.985 & 1.004 & 1.010 & 0.996 & 0.978 & 0.965 \\ 
+  Shift=5 & 0.970 & 0.986 & 1.007 & 1.015 & 1.002 & 0.978 & 0.956 \\  
    \hline
@@
   \hline
-RRMSE & 0.833 & 1.083 & 0.981 & 0.979 & 1.027 & 1.020 \\ 
+rRMSE & 0.833 & 1.083 & 0.981 & 0.979 & 1.027 & 1.020 \\ 
    \hline

### 2025-06-20 HWI SSA_27_05_2025_with new poster/20.06.2025/paper.tex
up

--- 2025-06-20 HWI SSA_27_05_2025_with new poster/05.06.2025/paper.tex
@@
 
-We then estimated VARMA models to capture these dynamics, selecting a parsimonious VAR(1) specification.\footnote{We used the \texttt{refVARMA()} procedure in the \cite{MTSpackage} \texttt{MTS} package, which sets VARMA coefficients with t-statistics $< 1.5$ to zero \citep[see][]{tsay2013multivariate}.
-We intend to explore richer dynamic structures in subsequent work.} Figure \ref{ma_inv_BIP} shows the implied responses of GDP to shocks to each indicator variable.\footnote{Shown are the MA coefficients for GDP from the vector moving-average (VMA) model obtained by inversion of the VAR(1) representation (i.e.
-the impluse-response functions, or IRFs, of GDP to the correlated reduced-form shocks.)}
+We then estimated a variety of parsimonious VARMA models to capture these dynamics, selecting parsimonious VAR(1) and VAR(3) specifications.\footnote{We used the \texttt{refVARMA()} procedure in the \cite{MTSpackage} \texttt{MTS} package, which sets VARMA coefficients with t-statistics below a critical threshold to zero \citep[see][]{tsay2013multivariate}.
+This led to a VAR(1) specification with a threshold of 1.5.
+The VAR(1) left some residual evidence of higher-order dynamics, but capturing these in our limited sample size was difficult without overfitting.
+We therefore explored VAR($p$) models with coefficients estimated by Elastic Net using the \cite{ElasticNet} \texttt{elasticnet} package.
+This led to a VAR(3) specification with 7 steps.
+For additional details, see \cite{ElasticNet} and the references therein.} The figures \ref{ma_inv_BIP} and \textbf{\textit{[cite new figure here]}} shows the implied responses of GDP to shocks to each indicator variable.\footnote{Shown are the MA coefficients for GDP from the vector moving-average (VMA) model obtained by inversion of the VAR(1) representation (i.e.
+the impulse-response functions, or IRFs, of GDP to the correlated reduced-form shocks.)}
+
+%
+%   SvN: I've added the following very brief description of Elastic-Net
+%
+% Linear regression models often suffer from issues such as overfitting and multicollinearity when the number of predictors exceeds the number of observations.
+Regularization techniques, such as ridge regression and the lasso, offer solutions by imposing penalties on model parameters.
+However, ridge regression maintains all predictors, while the lasso leads to sparse solutions by shrinking coefficients to zero.
+The elastic net estimator, introduced by \cite{ZouHastie2004JRSSB}, addresses limitations inherent in these approaches by blending their penalty functions.
+
+% The elastic net estimator is obtained by solving the following optimization problem:
+
+% \begin{equation}
+% \hat{\beta}^{EN} = \arg\min_{\beta} \left( \sum_{i=1}^{n} (y_i - X_i \beta)^2 + \lambda_1 \sum_{j=1}^{p} |\beta_j| + \lambda_2 \sum_{j=1}^{p} \beta_j^2 \right),
+% \end{equation}
+
+% where $\lambda_1$ and $\lambda_2$ are non-negative tuning parameters, and $X_i$ represents the predictor variables.
+% This formulation encompasses the least-squares estimator as the special case where $\lambda_1 = 0$ and the LASSO as the special case where $\lambda_2 = 0$.
+Unlike the lasso, which may arbitrarily select correlated predictors, the elastic net encourages grouping of strongly correlated variables.
+The combined penalties stabilize estimation in cases with high-dimensional predictors.
+%
+%
+%
 

### 2025-06-20 Sweave/BIP_predictor.Rnw
up

@@
 {
-# Marc's settings for VAR  
+# Marc's settings for EN
+  q<-0
+# 1. Simpler (but possibly too simple)  
   p<-2
-  q<-0
-# Number of non-zero coefficients in VAR for each dimension  
   nsteps<-5
+# 2. Simpler than Simon's original settings (below) but slightly more complex than above ones  (good compromise)
+  p<-3
+  nsteps<-6
+# Simon's original settings (lead to overfitting)  
+#  p<-4
+#  nsteps<-7
 }

### 2025-06-20 HWI SSA_27_05_2025_with new poster/21.06.2025 after revisions/paper.tex
up

--- 2025-06-20 HWI SSA_27_05_2025_with new poster/20.06.2025/paper.tex
@@
 This led to a VAR(3) specification with 7 steps.
-For additional details, see \cite{ElasticNet} and the references therein.} The figures \ref{ma_inv_BIP} and \textbf{\textit{[cite new figure here]}} shows the implied responses of GDP to shocks to each indicator variable.\footnote{Shown are the MA coefficients for GDP from the vector moving-average (VMA) model obtained by inversion of the VAR(1) representation (i.e.
+For additional details, see \cite{ElasticNet} and the references therein.} The figures \ref{ma_inv_BIP} and \ref{ma_inv_BIP_var3_en} show the implied responses of GDP to shocks to each indicator variable.\footnote{Shown are the MA coefficients for GDP from the vector moving-average (VMA) models obtained by inversion of the VAR(1) and VAR(3) representations respectively (i.e.
 the impulse-response functions, or IRFs, of GDP to the correlated reduced-form shocks.)}
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
     \begin{center}
         \includegraphics[height=3in, width=6in]{./Figures/ma_inv_BIP.pdf}
-        \caption{VMA coefficients for GDP Implied by VAR(1)\\Estimation with pandemic (left) and without pandemic (right).
+        \caption{VMA coefficients for GDP implied by VAR(1)\\Estimation with pandemic (left) and without pandemic (right).
         \label{ma_inv_BIP}}\end{center}
@@
 
-The model summarizes key features of the data after we account for publication lags.
-GDP responds most strongly to shocks to the two survey indicators (ifo\_c and ESI) after a modest lag.
-There also appears to be a more modest response to IP shocks after a slightly longer lag, but this vanishes once data after the start of the pandemic is included.\footnote{There is also weaker evidence of a response to the term spread (spr\_10y\_3m) at longer lags.
-} The delayed response of GDP to movements in these various indicators suggests that multivariate models may provide better GDP forecasts than univariate models, a question we pursue in the next section.
-%, where lagging series (GDP and IP, after accounting for publication lags) depend on indicators leading in relative terms.
-In particular, the impulse response of the spread suggests the relevance of a univariate model for this specific leading indicator.
+\begin{figure}[H]
+    \begin{center}
+        \includegraphics[height=3in, width=6in]{./Figures/ma_inv_BIP_var3_en.pdf}
+        \caption{VMA coefficients for GDP implied by VAR(3)\\Estimation with pandemic (left) and without pandemic (right).
+        \label{ma_inv_BIP_var3_en}}\end{center}
+\end{figure}
+
+The models encapsulate key characteristics of the data after adjusting for publication lags.
+Notably, GDP exhibits the strongest response to shocks in the two survey indicators—ifo and ESI\footnote{There is also weaker evidence of a response to the term spread at longer lags.}.
+The impact of the pandemic modifies the lag structure by emphasizing shorter dependencies during that specific episode.
+Removing the COVID-19 breakout from the analysis results in smoother, somewhat longer-tailed dependencies, consistent with the observed empirical autocorrelation functions.
+In particular, the VAR(3) model captures slightly extended lead times and more nuanced lag patterns.
+The delayed response of GDP to movements in these indicators suggests that multivariate models could yield superior forecasting performance compared to univariate approaches, a hypothesis we explore in the next section\footnote{For simplicity, we focus on the VAR(1) model here, while the M-SSA package provides more advanced options, including the discussed VAR(3).}.
 
@@
         %\includegraphics[height=3in, width=4.5in]{./Figures/direct_wc_all.pdf}
-        \caption{Direct forecasts (red lines) based on all indicators and full sample information (without pandemic): GDP (black line) shifted (i.e. lagged) by zero, one, two, and three quarters.
+        \caption{Direct forecasts (red lines) based on all indicators and full sample information (without pandemic): GDP (black line) left-shifted  by zero, one, two, and three quarters.
         \label{direct_wc}}
@@
 
-Figures \ref{direct_wc} and \ref{direct_wc_financial_crisis} compare actual and fitted values for GDP at forecast horizons $h = 0, \ldots, 3$, using the four indicators and lagged GDP, without applying any filtering.
+Figures \ref{direct_wc} and \ref{direct_wc_financial_crisis} compare actual and fitted values for GDP at forecast horizons $h = 0, \ldots, 3$, using the four indicators and the target, i.e.
+GDP shifted forward (leftwards) by $h$ quarters, without applying any filtering.
 Figure \ref{direct_wc} shows results over the sample without the pandemic, while Figure \ref{direct_wc_financial_crisis} provides a closer look at the fit during the financial crisis.
@@
 Using the equivalent dual formulation of the criterion \eqref{critssaht}, the author demonstrates that the SSA solution exhibits the lowest zero-crossing rate of any (linear) predictor with the same target correlation.
-In the following, we therefore use the M-SSA framework to construct a `smooth predictor $y_t$ to forecast $z_t$.
+In the following, we therefore use the M-SSA framework to construct a `smooth' predictor $y_t$ to forecast $z_t$.
 %In a second stage, serving as a generalization and refinement of the univariate HP-C utilized in the direct HP forecast.
@@
 Comparisons with M-MSE allow us to see the cost of this constraint in terms of prediction performance (target correlation).
-Both predictors use the same underlying VAR(1) model introduced in Section \ref{sec:direct_forecast} and estimated on data up to January 2008.
-This ensures a lengthy out-of-sample period that includes significant events such as the financial crisis, the sovereign debt crisis, and the COVID-19 pandemic\footnote{Further details and background information are available in \cite{Wildi2025}.
+Both predictors use the same underlying VAR(1) model introduced in Section \ref{sec:direct_forecast} and estimated on data up to Q4-2007.
+This ensures a lengthy out-of-sample period that includes significant events such as the financial crisis, the sovereign debt crisis, and the COVID-19 pandemic\footnote{The economic sentiment indicator (ESI) is excluded from the parsimonious VAR(1)-model when using the shorter sub-sample, ending in Q4-2007.
+Further details and background information are available in \cite{Wildi2025}.
 All calculations are performed using the M-SSA package, as documented in \cite{Wildi2025}.
@@
 The M-SSA nowcast is slightly smoother than that of the M-MSE, exhibiting approximately one-third fewer sign-changes.
-Fig.\ref{mssa_msse_zc} compares both nowcasts in relation to the financial crisis, highlighting that the classic (MSE-based) design tends to produce unwanted 'noisy' zero-crossings at (or in the vicinity of) the transitions between recessions and expansions.
+Fig.\ref{mssa_msse_zc} compares both nowcasts in relation to the financial crisis, highlighting that the classic (MSE-based) design tends to produce unwanted `noisy' zero-crossings at (or in the vicinity of) the transitions between recessions and expansions.
 Further, Table \ref{corhtnow} compares sample target correlations and HTs of the two nowcasts.
@@
 Next, we extend the previous example to consider forecasts of HP-GDP.
-The only change this requires is to change $h$ to $h=1,...,6$ quarters.
+The only modification this requires is to change $h=0$ (nowcast) to $h=1,...,6$ quarters.
 %The previously described nowcast example can be systematically extended to generate forecasts of the target variable.
@@
 %In comparison, the effect of phase shifts—arising from the cyclical nature of the HP-filter and manifesting in the HP-C forecasts (shown in the left panels)—seems less systematic.
+The systematic left-shift of the M-SSA is a key determinant of its out-of-sample forecast performances, as analyzed in the following section.
 
@@
 \end{equation}
-We now assume that the regression parameters $\hat{\beta}_0(shift,h,t)$ and$,\hat{\beta}_1(shift,h,t)$ are up-dated quarterly.
+We now assume that the regression parameters $\hat{\beta}_0(shift,h,t)$ and$,\hat{\beta}_1(shift,h,t)$ in Equation \ref{oos_pred} are up-dated quarterly.
 Note that Equations \ref{eq:GDP_M-SSA} and \ref{oos_pred} could be extended to incorporate additional M-SSA components, but for illustration purposes, we focus on the simplest (and most robust) design.
@@
 \end{equation}
-for observations after Q4-2007: the corresponding HAC-adjusted p-values are reported in Table \ref{p_val_wc}.\footnote{Statistical significance is influenced by the relatively short out-of-sample span, of length 17 years.} The MSE forecast performance of $\hat{\textrm{GDP}}_{t+shift}(h)$ can be benchmarked against the mean predictor (see Table \ref{rRMSE_mSSA_comp_mean_wc}) as well as against the direct forecast (see Table \ref{rRMSE_mSSA_comp_direct_without_covid7}).
+for observations after Q4-2007: the significance of the corresponding HAC-adjusted statistics is indicated by a `star' symbol in Table \ref{rRMSE_mSSA_comp_mean_wc}\footnote{A single star means significance at the $5\%$ level; a double star means significance at the $0.1\%$ level.
+Note that statistical significance is influenced by the relatively short out-of-sample span, of length 17 years.}.
+The MSE forecast performance of $\hat{\textrm{GDP}}_{t+shift}(h)$ can be benchmarked against the mean predictor (see Table \ref{rRMSE_mSSA_comp_mean_wc}) as well as against the direct forecast (see Table \ref{rRMSE_mSSA_comp_direct_without_covid7}).
 For consistency, both benchmark predictors are also quarterly updated out-of-sample designs.\\
@@
 Our results can be summarized as follows.
-When excluding the pandemic, the direct forecasts outperform the mean forecast up to two quarters ahead, see Table \ref{rRMSE_mSSA_direct_mean_without_covid8}\footnote{However, the pandemic period generally obscures the relationships between predictors and GDP.}.
-However, the GDP component predictors $\hat{\textrm{GDP}}_{t+shift}(h)$, optimized for longer forecast horizons $h\geq 5$, outperform the mean benchmark up to 5Q ahead, see Table \ref{rRMSE_mSSA_comp_mean_wc}.
-Additionally,  $\hat{\textrm{GDP}}_{t+shift}(h)$, optimized for large forecast horizons $h\geq 4$, surpass the direct forecasts for $shift\geq 3$, see Table \ref{rRMSE_mSSA_comp_direct_without_covid7}.
-Statistical significance of the GDP component predictor, optimized for larger forecast horizons, can be established for at least up to one year ahead, see Table \ref{p_val_wc}.
+When excluding the pandemic, the direct forecasts outperform the mean forecast up to two quarters ahead, see Table \ref{rRMSE_mSSA_direct_mean_without_covid8} in the appendix.
+%\footnote{However, the pandemic period generally obscures the relationships between predictors and GDP.}.
+In contrast, the GDP component predictors $\hat{\textrm{GDP}}_{t+shift}(h)$, optimized for longer forecast horizons $h\geq 5$, outperform the mean benchmark up to 5Q ahead, see Table \ref{rRMSE_mSSA_comp_mean_wc}.
+Additionally,  $\hat{\textrm{GDP}}_{t+shift}(h)$, optimized for large forecast horizons $h\geq 4$, surpass the direct forecasts for $shift\geq 3$, see Table \ref{rRMSE_mSSA_comp_direct_without_covid7} in the appendix.
+Statistical significance of the GDP component predictor, optimized for larger forecast horizons, can be established for at least up to one year ahead, see Table \ref{rRMSE_mSSA_comp_mean_wc}.
 Finally, the strong performance of predictors optimized for larger forecast horizons ($h\geq 4$), at smaller shifts ($shifts< 4$) corroborates the importance of the left-shift of the corresponding predictors in relation to forecast performance.
@@
 
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 12:49:50 2025
 \begin{table}[ht]
@@
   \hline
-Shift=0 & 0.084 & 0.047 & 0.017 & 0.005 & 0.002 & 0.001 & 0.038 \\ 
-  Shift=1 & 0.230 & 0.160 & 0.094 & 0.032 & 0.005 & 0.001 & 0.004 \\ 
-  Shift=2 & 0.801 & 0.666 & 0.414 & 0.173 & 0.026 & 0.005 & 0.006 \\ 
-  Shift=3 & 0.876 & 0.961 & 0.917 & 0.683 & 0.248 & 0.068 & 0.039 \\ 
-  Shift=4 & 0.370 & 0.819 & 0.990 & 0.953 & 0.533 & 0.117 & 0.030 \\ 
-  Shift=5 & 0.353 & 0.802 & 0.983 & 0.943 & 0.571 & 0.239 & 0.043 \\ 
-   \hline
-\end{tabular}
-\caption{p-values for $H_0: \hat{\beta_1} = 0$, HAC-adjusted Wald test, Equation \ref{eq_gdp_pred}.\\Sample excludes pandemic.} 
-\label{p_val_wc}
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 12:49:50 2025
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrrrr}
-  \hline
- & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
-  \hline
-Shift=0 & 0.973 & 0.937 & 0.888 & 0.850 & 0.855 & 0.900 & 0.956 \\ 
-  Shift=1 & 1.073 & 1.051 & 0.997 & 0.918 & 0.870 & 0.873 & 0.906 \\ 
-  Shift=2 & 1.069 & 1.077 & 1.063 & 1.006 & 0.940 & 0.907 & 0.907 \\ 
-  Shift=3 & 1.020 & 1.045 & 1.066 & 1.048 & 0.997 & 0.957 & 0.941 \\ 
-  Shift=4 & 0.983 & 1.016 & 1.056 & 1.064 & 1.023 & 0.974 & 0.942 \\ 
-  Shift=5 & 0.988 & 1.026 & 1.071 & 1.088 & 1.057 & 1.007 & 0.963 \\ 
-   \hline
-\end{tabular}
-\caption{rRMSEs: out-of-sample forecast errors Equation \ref{oosfe} vs expanding mean of GDP\\Out-of-sample evaluation starting in Jan-2008, without pandemic.\\Values $< 1$ indicate superior predictions from Equation \ref{oos_pred}.}
+Shift=0 & 0.975 & 0.941^{*} & 0.886^{*} & 0.839^{**} & 0.843^{**} & 0.897^{**} & 0.965 \\ 
+  Shift=1 & 1.085 & 1.068 & 1.011 & 0.927^{*} & 0.871^{**} & 0.875^{**} & 0.919^{**} \\ 
+  Shift=2 & 1.076 & 1.089 & 1.074 & 1.011 & 0.938^{*} & 0.902^{**} & 0.909^{**} \\ 
+  Shift=3 & 1.037 & 1.069 & 1.090 & 1.059 & 0.992 & 0.943^* & 0.930^* \\ 
+  Shift=4 & 1.013 & 1.058 & 1.101 & 1.097 & 1.037 & 0.967 & 0.925^* \\ 
+  Shift=5 & 1.016 & 1.062 & 1.106 & 1.115 & 1.078 & 1.015 & 0.955^* \\  
+   \hline
+\end{tabular}
+\caption{rRMSEs: out-of-sample forecast errors Equation \ref{oosfe} vs expanding mean of GDP\\Out-of-sample evaluation starting in Q1-2008, without pandemic.\\Values $< 1$ indicate superior predictions from Equation \ref{oos_pred}\\
+Row-index: Shift refers to the left-shift of the target $\textrm{GDP}_{t+shift}$\\
+Column-index: $h$ refers to the forecast horizon for which $\textrm{M-SSA}_t^{\textrm{GDP}}(h)$ is optimized\\
+Significance refers to the hypothesis $\alpha_1=0$ in Equation \ref{eq_gdp_pred}.
+} 
 \label{rRMSE_mSSA_comp_mean_wc}
@@
   \hline
-Forecast horizon 1 & 0.752 & 0.715 & 0.746 & 0.747 \\ 
-  Forecast horizon 2 & 0.697 & 0.663 & 0.729 & 0.720 \\ 
-  Forecast horizon 3 & 0.554 & 0.622 & 0.684 & 0.709 \\ 
-  Forecast horizon 4 & 0.486 & 0.624 & 0.680 & 0.708 \\ 
-  Forecast horizon 5 & 0.593 & 0.636 & 0.643 & 0.701 \\ 
+Forecast horizon 1 & 0.752 & 0.715 & 0.750 & 0.756 \\ 
+  Forecast horizon 2 & 0.697 & 0.663 & 0.714 & 0.744 \\ 
+  Forecast horizon 3 & 0.554 & 0.622 & 0.666 & 0.716 \\ 
+  Forecast horizon 4 & 0.486 & 0.624 & 0.707 & 0.732 \\ 
+  Forecast horizon 5 & 0.593 & 0.636 & 0.660 & 0.713 \\ 
    \hline
@@
 These revisions result from quarterly updates to the regression coefficients of equation \ref{oos_pred}.
-The M-SSA filter    itself remains fixed, based on data up to January 2008.
+The M-SSA filter    itself remains fixed, based on data up to Q4-2007.
         \label{revisions1}}
@@
         \includegraphics[height=3in, width=4.5in]{./Figures/up_dated_ma_inv_multi_ip.pdf}
-        \caption{MA-inversion (impulse response) of VAR: data up to Jan-2008 (left) vs. entire data set without pandemic (right).
+        \caption{MA-inversion (impulse response) of VAR: data up to Q4-2007 (left) vs. entire data set without pandemic (right).
         \label{up_dated_ma_inv_multi_ip}}
@@
         \includegraphics[height=3in, width=4.5in]{./Figures/bk_2008_all.pdf}
-        \caption{M-SSA filter tracking HP-GDP at forecast horizon $h=4$: data up to Jan-2008 (left) vs. entire data set without pandemic (right).
+        \caption{M-SSA filter tracking HP-GDP at forecast horizon $h=4$: data up to Q4-2007 (left) vs. entire data set without pandemic (right).
         \label{bk_2008_all}}
@@
 The second source of revisions in the M-SSA component predictor arises from updates to the VAR model within the M-SSA filter.
-We compare the original estimates, based on data up to January 2008, with the final estimates obtained using $h=4$ and $shift=4$ (similar findings apply to other values of forecast horizon and forward-shift).
+We compare the original estimates, based on data up to Q4-2007, with the final estimates obtained using $h=4$ and $shift=4$ (similar findings apply to other values of forecast horizon and forward-shift).
 For consistency, we excluded the entire pandemic episode from the data.
@@
 \end{tabular}
-\caption{Regression of shifted GDP on M-SSA($h$) GDP component.\\rRMSEs benchmarked against the direct forecast.\\Out-of-sample evaluation starting in Jan-2008, without pandemic.} 
+\caption{Regression of shifted GDP on M-SSA($h$) GDP component.\\rRMSEs benchmarked against the direct forecast.\\Out-of-sample evaluation starting in Q1-2008, without pandemic.} 
 \label{rRMSE_mSSA_comp_direct_without_covid7}
@@
 \caption{Direct forecasts.\\
-rRMSEs benchmarked against the expanding mean.\\Out-of-sample evaluation starting in Jan-2008, without pandemic.} 
+rRMSEs benchmarked against the expanding mean.\\Out-of-sample evaluation starting in Q1-2008, without pandemic.} 
 \label{rRMSE_mSSA_direct_mean_without_covid8}
@@
 \end{tabular}
-\caption{rRMSEs: out-of-sample forecast errors Equation \ref{oosfe} vs expanding mean of GDP\\Out-of-sample evaluation starting in Jan-2008, including the pandemic.\\Values $< 1$ indicate superior predictions from Equation \ref{oos_pred}.}
+\caption{rRMSEs: out-of-sample forecast errors Equation \ref{oosfe} vs expanding mean of GDP\\Out-of-sample evaluation starting in Q1-2008, including the pandemic.\\Values $< 1$ indicate superior predictions from Equation \ref{oos_pred}.}
 \label{rRMSE_mSSA_comp_mean2}
@@
 \end{tabular}
-\caption{Regression of shifted GDP on M-SSA($h$) GDP component.\\rRMSEs benchmarked against the direct forecast.\\Out-of-sample evaluation starting in Jan-2008, including the pandemic.} 
+\caption{Regression of shifted GDP on M-SSA($h$) GDP component.\\rRMSEs benchmarked against the direct forecast.\\Out-of-sample evaluation starting in Q1-2008, including the pandemic.} 
 \label{rRMSE_mSSA_comp_direct3}
@@
 \caption{Direct forecasts.\\
-rRMSEs benchmarked against the expanding mean.\\Out-of-sample evaluation starting in Jan-2008, including the pandemic.} 
+rRMSEs benchmarked against the expanding mean.\\Out-of-sample evaluation starting in Q1-2008, including the pandemic.} 
 \label{rRMSE_mSSA_direct_mean4}

### 2025-06-20 Latex/BIP_predictor.tex
up

--- 2025-06-04 Latex/BIP_predictor.tex
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 14:24:40 2025
+% Fri Jun 20 07:05:56 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 14:24:41 2025
+% Fri Jun 20 07:05:56 2025
 \begin{table}[ht]
@@
 \ref{impulse} (full data set) and \eqref{impulse_wc} (prior the pandemic).
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/ma_inv_multi_ip.pdf}\caption{MA inversion of VAR(1): full data set.\label{impulse}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/ma_inv_multi_ip_wc.pdf}\caption{MA inversion of VAR(1): without Pandemic.\label{impulse_wc}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=6in]{./Figures/ma_inv_BIP.pdf}\caption{MA inversion of VAR(1) for GDP: with pandemic (left) and without pandemic (right).\label{ma_inv_BIP}}\end{center}\end{figure}
+\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/ma_inv_multi_ip.pdf}\caption{MA inversion of VAR(1): full data set.\label{impulse}}\end{center}\end{figure}
+
+\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/ma_inv_multi_ip_wc.pdf}\caption{MA inversion of VAR(1): without Pandemic.\label{impulse_wc}}\end{center}\end{figure}
 
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 14:24:43 2025
+% Fri Jun 20 07:05:58 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 14:24:53 2025
+% Fri Jun 20 07:06:08 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 14:25:53 2025
+% Fri Jun 20 07:06:51 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 14:26:07 2025
+% Fri Jun 20 07:07:00 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 14:26:07 2025
+% Fri Jun 20 07:07:00 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 14:26:39 2025
+% Fri Jun 20 07:07:25 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed Jun  4 14:26:39 2025
+% Fri Jun 20 07:07:25 2025
 \begin{table}[ht]

### 2025-06-21 HWI SSA_27_05_2025_with new poster/22.06.2025/paper.tex
up

--- 2025-06-20 HWI SSA_27_05_2025_with new poster/21.06.2025 after revisions/paper.tex
@@
 
-The models encapsulate key characteristics of the data after adjusting for publication lags.
-Notably, GDP exhibits the strongest response to shocks in the two survey indicators—ifo and ESI\footnote{There is also weaker evidence of a response to the term spread at longer lags.}.
-The impact of the pandemic modifies the lag structure by emphasizing shorter dependencies during that specific episode.
+%The models encapsulate key characteristics of the data after adjusting for publication lags. Notably, 
+GDP exhibits the strongest response to shocks in the two survey indicators—ifo and ESI\footnote{There is also weaker evidence of a response to the term spread at longer lags.}.
+The impact of the pandemic modifies the lag structure by emphasizing shorter dependencies triggered by the alternating outliers during that specific episode.
 Removing the COVID-19 breakout from the analysis results in smoother, somewhat longer-tailed dependencies, consistent with the observed empirical autocorrelation functions.
-In particular, the VAR(3) model captures slightly extended lead times and more nuanced lag patterns.
-The delayed response of GDP to movements in these indicators suggests that multivariate models could yield superior forecasting performance compared to univariate approaches, a hypothesis we explore in the next section\footnote{For simplicity, we focus on the VAR(1) model here, while the M-SSA package provides more advanced options, including the discussed VAR(3).}.
+In particular, the VAR(3) model captures slightly extended lead times and more nuanced lag patterns\footnote{The VAR(3) model yields marginal improvements in multivariate filtering performance.
+However, due to its relative simplicity, this study primarily employs the VAR(1) model.
+The M-SSA package (accessible at \url{https://github.com/wiaidp/R-package-SSA-Predictor}) offers more advanced capabilities, including the implementation of the VAR(3) model and various additional (Bayesian) VAR specifications.}.
+The delayed response of GDP to movements in these indicators suggests that multivariate models could yield superior forecasting performance compared to univariate approaches, a hypothesis we explore in the next section
 
@@
 Both predictors use the same underlying VAR(1) model introduced in Section \ref{sec:direct_forecast} and estimated on data up to Q4-2007.
-This ensures a lengthy out-of-sample period that includes significant events such as the financial crisis, the sovereign debt crisis, and the COVID-19 pandemic\footnote{The economic sentiment indicator (ESI) is excluded from the parsimonious VAR(1)-model when using the shorter sub-sample, ending in Q4-2007.
-Further details and background information are available in \cite{Wildi2025}.
-All calculations are performed using the M-SSA package, as documented in \cite{Wildi2025}.
-\url{https://github.com/wiaidp/R-package-SSA-Predictor}}
+This ensures a lengthy out-of-sample period that includes significant events such as the financial crisis, the sovereign debt crisis, and the COVID-19 pandemic\footnote{In the parsimonious VAR(1) model estimated over the shorter sub-sample ending in Q4-2007, the economic sentiment indicator (ESI) does not receive any weight.
+Additional technical details and background information are provided in \cite{Wildi2025}.
+All computations are conducted using the M-SSA package, as documented in \cite{Wildi2025}.
+The package can be accessed at \url{https://github.com/wiaidp/R-package-SSA-Predictor}.}
 
@@
 
-The second source of revisions in the M-SSA component predictor arises from updates to the VAR model within the M-SSA filter.
+The second source of revisions in the M-SSA component predictor arises from updates to the VAR(1) model within the M-SSA filter\footnote{Similar findings apply to more sophisticated alternatives such as the VAR(3) discussed in Section \ref{sec:data}.}.
 We compare the original estimates, based on data up to Q4-2007, with the final estimates obtained using $h=4$ and $shift=4$ (similar findings apply to other values of forecast horizon and forward-shift).
@@
 To isolate the effect of the regression revision, the figure displays standardized series.
-The update to the VAR model influences the left-shift of the predictor (blue line), which appears slightly `faster' than the original predictor (red line).
+The update to the VAR model results in a leftward shift of the predictor (represented by the blue line), which appears marginally more rapid compared to the original predictor (depicted by the red line).
+It is important to note that the smoothness (zero-crossing rate) remains unchanged, though.
 This suggests that the out-of-sample performance of the M-SSA component predictor may be better than reported in Section \ref{oosp}, which is based on the fixed and increasingly outdated version of M-SSA.
@@
 
-To address these problems, we propose a multivariate extension of SSA framework based on \cite{Wildi2024}.
+To address these problems, we propose a multivariate extension of the SSA framework based on \cite{Wildi2024}.
 Unlike the univariate HP filter, the M-SSA can leverage cross-sectional information from indicators leading GDP in real time while controlling for the rate of zero-crossings of the predictor.
@@
 Consequently, a strong link between M-SSA and future HP-GDP also indicates a connection with GDP, even if the statistical significance is obscured by the noise inherent in GDP.
-Out-of-sample performance suggests that this predictor is statistically significant at horizons up to one year.
+Out-of-sample performance suggests that this predictor is statistically significant at horizons exceeding one year.
 Moreover, designs optimized for larger forecast horizons outperform the mean benchmark at all considered forecast horizons, and the predictor exceeds the performance of direct forecasts---based on unfiltered or HP filtered indicators---at horizons longer than two quarters.
-An analysis of revision errors arising from quarterly updates of the proposed M-SSA GDP component predictor indicates that the regression parameters, the VAR model, and the resulting M-SSA filters stabilize after an initial burn-in period.
-This stability enhances the explainability and interpretability of the predictor.\\
+An analysis of hit and false alarm rates within the ROC curve confirms that filtering improves predictive performance for forecast horizons exceeding two quarters.
+Multivariate approaches can further capitalize on leading indicators within the cross-sectional data.
+Additionally, the AUC is optimized through more precise control of the zero-crossing rate within the M-SSA framework.
+Finally, an examination of revision errors associated with quarterly updates of the proposed M-SSA-based GDP component predictor indicates that the regression parameters, VAR model, and resulting M-SSA filters stabilize after an initial burn-in period.
+This stabilization enhances the model’s explainability and interpretability.\\
 

### 2025-06-22 HWI SSA_27_05_2025_with new poster/23.06.2025/poster.tex
up

\documentclass[final]{beamer}
\mode<presentation>
\usetheme{IWH}
% Use the beamerposter package
%\usepackage[orientation=portrait,size=a1,scale=1.15]{beamerposter}

\usepackage[orientation=portrait,scale=1.15]{beamerposter}
\setlength{\paperwidth}{36in}
\setlength{\paperheight}{48in}
\setlength{\textwidth}{0.98\paperwidth}
\setlength{\textheight}{0.98\paperheight}
\usepackage{graphicx} % for including images
\usepackage{amsmath} % for math symbols
\usepackage{amssymb} % for more math symbols
\usepackage{amsfonts}
\usepackage{lmodern} % scalable font package
\usepackage{booktabs}
\usepackage{colortbl}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{subcaption}
\usepackage{setspace}

% Define poster colors and style
\usepackage{color}
\definecolor{darkblue}{rgb}{0.1,0.1,0.5}
\setbeamercolor{title}{fg=darkblue}

%\usepackage[pdftex,hiresbb]{graphicx}
%\usepackage[pageshow]{xtab}
\usepackage[bottom,hang,splitrule]{footmisc}
\usepackage[font={small},labelfont={bf},textfont={bf},format=hang,justification=RaggedRight]{caption}
\usepackage{tabularx}
\usepackage{natbib}
%\usepackage[spaces,hyphens]{url}
\usepackage{breakcites}
\usepackage{fancyhdr}
\usepackage{longtable}
\usepackage{dcolumn}
\usepackage{booktabs}
\usepackage{hyphenat}
%\usepackage[hidelinks]{hyperref}
\usepackage{doi}
\usepackage[UKenglish]{babel}
\usepackage[pdftex]{rotating} % [counterclockwise] %evtl. miteinbinden
\usepackage{color}
\usepackage[section]{placeins}
%\usepackage{subfig}
\usepackage{siunitx}
\sisetup{detect-all}

%%%%%%%%%%%%%%%5
\usepackage{environ}% Required for \NewEnviron, i.e. to read the whole body of the environment
\makeatletter

\newcounter{acolumn}%  Number of current column
\newlength{\acolumnmaxheight}%   Maximum column height

% `column` replacement to measure height
\newenvironment{@acolumn}[1]{%
	\stepcounter{acolumn}%
	\begin{lrbox}{\@tempboxa}%
		\begin{minipage}{#1}%
		}{%
		\end{minipage}
	\end{lrbox}
	\@tempdimc=\dimexpr\ht\@tempboxa+\dp\@tempboxa\relax
	% Save height of this column:
	\expandafter\xdef\csname acolumn@height@\roman{acolumn}\endcsname{\the\@tempdimc}%
	% Save maximum height
	\ifdim\@tempdimc>\acolumnmaxheight
	\global\acolumnmaxheight=\@tempdimc
	\fi
}

% `column` wrapper which sets the height beforehand
\newenvironment{@@acolumn}[1]{%
	\stepcounter{acolumn}%
	% The \autoheight macro contains a \vspace macro with the maximum height minus the natural column height
	\edef\autoheight{\noexpand\vspace*{\dimexpr\acolumnmaxheight-\csname acolumn@height@\roman{acolumn}\endcsname\relax}}%
	% Call original `column`:
	\orig@column{#1}%
}{%
	\endorig@column
}

% Save orignal `column` environment away
\let\orig@column\column
\let\endorig@column\endcolumn

% `columns` variant with automatic height adjustment
\NewEnviron{acolumns}[1][]{%
	% Init vars:
	\setcounter{acolumn}{0}%
	\setlength{\acolumnmaxheight}{0pt}%
	\def\autoheight{\vspace*{0pt}}%
	% Set `column` environment to special measuring environment
	\let\column\@acolumn
	\let\endcolumn\end@acolumn
	\BODY% measure heights
	% Reset counter for second processing round
	\setcounter{acolumn}{0}%
	% Set `column` environment to wrapper
	\let\column\@@acolumn
	\let\endcolumn\end@@acolumn
	% Finally process columns now for real
	\begin{columns}[#1]%
		\BODY
	\end{columns}%
}
\makeatother
%%%%%%%%%%

% Title, author, and date
\title{Smooth and Persistent Forecasts of German GDP}
%\subtitle[]{\hspace{1.9cm}\LARGE{ A quarterly evaluation of EU Commissions' GDP forecasts}}

\author[]{Katja Heinisch$^{\textit{ 1}}$, Simon van Norden$^{\textit{ 2}}$, Marc Wildi$^{\textit{ 3}}$}
\institute[IWH]{\normalsize{$^{\textit{1}}$ Halle Institute for Economic Research (IWH), $^{\textit{2}}$ HEC Montréal and CIREQ, $^{\textit{3}}$ Zurich University of Applied Sciences (ZHAW) }}
%\author[]{Katja Heinisch}
%\institute[IWH]{Halle Institute for Economic Research}
%\vspace{2cm}

\begin{document}

	\begin{frame}
		\begin{columns}[T]
			% Column 1
			\begin{column}{0.48\textwidth}
					\begin{block}{\large Motivation}

					\begin{itemize}
					\item Forecasting dilemma: Efficiency requires full use of information, yet resulting forecasts are often highly volatile.
				%	Trade-off between informational efficiency and forecast smoothness.
					\item Short-term volatility and data revisions challenge reliable forecasting.
				\item \cite{Wildi2024,Wildi2025,McElroy2019,McElroy2020} propose the Smooth Sign Accuracy (SSA) framework, which controls volatility via penalizing forecast sign changes.
				\\
				\vspace{1cm}
					 {\textbf{\textcolor{iwhdarkblue}Research Questions:}}

					\item Can we improve GDP forecast smoothness without sacrificing accuracy?
					\item Do multivariate filtering techniques enhance persistence and coherence?
					\\
					$\rightarrow$ 	We propose methods to produce smooth and persistent forecasts of German GDP, i.e. by replacing raw leading indicators with smoothed and filtered components derived from the M-SSA procedure, thereby improving the signal-to-noise ratio in the predictor variables used to forecast GDP in the mid of a particular quarter.
					\end{itemize}

			\end{block}

		\begin{block}{\large Data}

					  \begin{columns}[T]
						\column{0.495\textwidth}
							\begin{itemize}
							     \item Quarterly German GDP (1995–2024).
							     \item Monthly Indicators: Industrial Production, ifo Business Climate, ESI, Term spread 10y-3m.
							     \item Log-differenced, standardized (except spreads), outliers trimmed ( COVID-19).
							     \item Data aligned to account for publication lags and converted to quarterly frequency 
							     \item Dynamic dependencies are explored via cross-correlation functions (CCFs) and vector moving-average (VMA) representations.
							     \item Survey indicators show strong leading correlations with GDP; IP and term spread exhibit weaker or delayed effects.
							     \item Pandemic-related outliers distort dynamic patterns; hence, data from Q4-2019 to Q4-2020 are excluded for stability.
							\end{itemize}

						\column{0.495\textwidth}
									\begin{center}
										{\small{\textbf{\textcolor{black}{Figure 1: Leads \& lags}}}}

									    \includegraphics[width=1\textwidth]{./Figures/data_lags.pdf}
									\end{center}

								\leftskip2em \singlespacing{ {\footnotesize \emph{Note:} GDP \& Indicators: standardized log-differences, trimmed to $\pm 3$ standard deviations.} }
						\end{columns}

				\end{block}

		\begin{block}{\large{Direct Forecasts}}
			\begin{itemize}
				\item Regression of GDP on contemporaneous indicators, i.e. VARMA dynamics estimated via parsimonious VAR(1).
				\item Impulse responses reveal GDP reacts most strongly to survey shocks, with modest delayed response to IP.
				\item These dynamics motivate the use of multivariate models to improve forecast accuracy beyond univariate benchmarks.
				\item The results indicate that as the forecast horizon $h$ increases, the turning points (i.e., peaks and troughs) of the direct forecasts become increasingly right-shifted (lagged) relative to the target GDP. 
			\end{itemize}

		\begin{columns}[T]
			\column{0.495\textwidth}

				\textcolor{iwhdarkblue}{\textbf{Filtering Indicators}}
					\begin{itemize}
						\item Forecasts are re-estimated using indicators smoothed via the one-sided HP-C filter (HP(160)).
						\item Figures show improved alignment: forecasts lag GDP less and better capture turning points.
						\item Filtering reduces high-frequency noise and enhances signal extraction.
					\end{itemize}
				\begin{footnotesize}
						\begin{table}[h]
						\begin{center}
								{\small{\textbf{\textcolor{black}{Table 1: }}}}
						\begin{tabular}{lrrrrrrr}
							\hline
							& h= 0 & h= 1 & h= 2 & h= 3 & h= 4 & h= 5 & h= 6 \\ 
							\hline
							Unfiltered & 0.028 & 0.000 & 0.001 & 0.910 & 0.509 & 0.091 & 0.178 \\ 
							HP-C filtered & 0.000 & 0.001 & 0.036 & 0.046 & 0.049 & 0.012 & 0.022 \\ 
							\hline
						\end{tabular}
						\leftskip2em \singlespacing{ {\footnotesize \emph{Note:} p-values for $H_0: \hat{\beta} = 0$, HAC-adjusted Wald test.\\Estimation over full sample (without pandemic.)					\label{tab:pvaluedhp}}}
						\end{center}
						\end{table}
				\end{footnotesize}
			\begin{itemize}
				\item HP-C filter improves smoothness but remains univariate and prone to high-frequency leakage.
			\end{itemize}

				\column{0.495\textwidth}

				\begin{center}
				{\small{\textbf{\textcolor{black}{Figure 2: Performance of direct forecasts}}}}
				\includegraphics[width=1\textwidth]{./Figures/direct_hp_forecasts_financial_crisis.pdf}
				\end{center}

				\end{columns}	
			\end{block}

				\begin{block}{\large Empirical Approach -- Multivariate Smooth Sign Accuracy (M-SSA)}
						\addtolength{\itemsep}{10pt}
			%	\textcolor{iwhdarkblue}{\textbf{Filtering Indicators}}
            \begin{columns}[T]
			\column{0.495\textwidth}

				\begin{itemize}

				\item M-SSA extends SSA (\cite{Wildi2025}) to multivariate setting with causal (one-sided) filters.
				\item Target variable $z_t$ (e.g. HP-GDP) is tracked by predictor $y_t$ via multivariate filters on leading indicators.
				\item SSA criterion: maximize forecast-target correlation under constraint on forecast autocorrelation $\rho(y,1)$.
				\item Constraint controls smoothness via expected holding time: $ht(y) = \frac{\pi}{\arccos(\rho(y,1))}$.
				\item Optimization via constraints on sign-changes.
			\end{itemize}
            \column{0.495\textwidth}

				\begin{center}
				{\small{\textbf{\textcolor{black}{Figure 3: M-SSA vs. HP-C}}}}
				\includegraphics[width=1\textwidth]{./Figures/multivar_vs_univar.pdf}
				\end{center}

				\end{columns}

%	\autoheight
			\end{block}

			\end{column}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
			% Column 2
		\begin{column}{0.48\textwidth}

			\begin{block}{\large{Results}}

			\begin{columns}
			\column{0.495\textwidth}
			\textcolor{iwhdarkblue}{\textbf{M-SSA Nowcasts vs M-MSE}}\\
			\begin{itemize}
			\item Comparison of nowcasts for HP-GDP using M-SSA vs M-MSE.
			\item M-SSA reduces sign changes by 33\% with minimal loss of accuracy.
			\item M-SSA achieves longer holding times: a $50\%$ increase compared to M-MSE
			\end{itemize}

		\column{0.495\textwidth}
			  \begin{center}
			\includegraphics[width=0.8\textwidth]{./Figures/mssa_msse_zc.pdf}
				\leftskip2em \singlespacing{ {\footnotesize \emph{Note:}{Two-sided HP(160) applied to GDP (black)\\ Nowcasts: M-SSA (blue) and M-MSE (green). \\The dashed vertical line delimits in- and out-of sample periods.  \\The two-sided filter does not extend to the sample end.
				\label{mssa_msse_now}}}}
		\end{center}
		\end{columns}

					\vspace{0.5cm}
					\textcolor{iwhdarkblue}{\textbf{M-SSA Forecasts for $h > 0$}}\\

						\begin{itemize}	
								\addtolength{\itemsep}{10pt}	
					\item As the forecast horizon hh increases, M-SSA filter weights decrease, indicating greater uncertainty.
					\item M-SSA forecasts shift leftward consistently as $h$ increases, especially around turning points.

						\end{itemize}

			\vspace{0.5cm}

			\textcolor{iwhdarkblue}{\textbf{M-SSA Component Predictor}}\\
				\begin{itemize}
				\item M-SSA components, derived from HP(160)-filtered indicators, are used as predictors for unfiltered GDP via a post-processing regression step.
				\item The M-SSA filters are fixed using data up to January 2008, while component weights are updated quarterly using an expanding window starting from January 2008.
				\item A linear regression maps the M-SSA output at horizon $h$ to forward-shifted GDP at $t + \text{shift}$, correcting for scaling and allowing for $h \neq \text{shift}$.
				\item This setup enables leveraging the forecast-leading behavior (left-shift) of M-SSA predictors, particularly when $\text{shift} < h$.
			\end{itemize}
			\begin{columns}
					\column{0.495\textwidth}

			\begin{itemize}
			\item Out-of-sample forecast performance is evaluated by comparing the M-SSA-based predictor to both a mean forecast and a direct forecast, all updated quarterly.
			\item Results show that for longer horizons ($h \geq 4$), M-SSA predictors outperform both benchmarks, especially at shorter shifts, confirming the predictive value of their anticipatory signal.
			\item Statistical significance of the M-SSA GDP predictor is confirmed for up to one year ahead, excluding the pandemic period.
			\end{itemize}

			\column{0.495\textwidth}
\begin{table}[ht]
	\centering
			{\small{\textbf{\textcolor{black}{Table 2: rRMSEs }}}}\\
\begin{footnotesize}
	\begin{tabular}{rrrrrrrr}
		\hline
		& h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
		\hline
		Shift=0 & 0.978 & 0.942 & 0.893 & 0.855 & 0.860 & 0.906 & 0.962 \\ 
		Shift=1 & 1.082 & 1.060 & 1.005 & 0.926 & 0.877 & 0.880 & 0.914 \\ 
		Shift=2 & 1.083 & 1.090 & 1.077 & 1.019 & 0.952 & 0.918 & 0.919 \\ 
		Shift=3 & 1.034 & 1.059 & 1.081 & 1.063 & 1.010 & 0.970 & 0.954 \\ 
		Shift=4 & 0.995 & 1.029 & 1.069 & 1.077 & 1.035 & 0.986 & 0.953 \\ 
		Shift=5 & 0.993 & 1.030 & 1.075 & 1.093 & 1.062 & 1.011 & 0.967 \\ 
		\hline
	\end{tabular}
\end{footnotesize}
\begin{singlespace}
	{\footnotesize \emph{Notes:} : Out-of-sample forecast errors  vs expanding mean forecast of GDP. Out-of-sample evaluation starting in Q1 2008, without pandemic.\\Values $< 1$ indicate superior predictions.}
\end{singlespace}
\vspace{0.5cm}
\end{table}
\end{columns}

\vspace{0.5cm}
\textcolor{iwhdarkblue}{\textbf{Receiver Operating Characteristic (ROC)}}\\

\begin{columns}
    \column{0.495\textwidth}
      \begin{itemize}
			\item Compute sign accuracy and false positive rate: ROC curves
            \item Compute Area Under Curve (AUC)
			\item All filters outperform direct forecast (larger AUC) for $h>2$ 
			\end{itemize}

\begin{table}[ht]
	\centering
			{\small{\textbf{\textcolor{black}{Table 3: AUCs }}}}\\
\begin{footnotesize}
\begin{tabular}{rrrrr}
  \hline
 & Direct & HP-C & M-MSE & M-SSA \\ 
  \hline
Forecast horizon 1 & 0.752 & 0.715 & 0.746 & 0.747 \\ 
  Forecast horizon 2 & 0.697 & 0.663 & 0.729 & 0.720 \\ 
  Forecast horizon 3 & 0.554 & 0.622 & 0.684 & 0.709 \\ 
  Forecast horizon 4 & 0.486 & 0.624 & 0.680 & 0.708 \\ 
  Forecast horizon 5 & 0.593 & 0.636 & 0.643 & 0.701 \\ 
   \hline
\end{tabular}
\end{footnotesize}
\begin{singlespace}
	{\footnotesize \emph{Notes:} :AUCs of predictors against GDP at various forecast horizons.}
\end{singlespace}
\vspace{0.5cm}
\end{table}

    \column{0.495\textwidth}
        \begin{center}
        \includegraphics[width=0.8\textwidth]{./Figures/ROC_GDP_shift_1_4.pdf}

	     \end{center}

\end{columns}

			\end{block}

			\begin{block}{\large Summary and Conclusion}

			\begin{itemize}
            \item Classic direct forecasts of GDP perform well up to two quarters ahead but suffer from noise and limited horizon; smoothing with an HP(160)-filter improves performance during sharp recessions yet retains some noise-induced sign changes.
            \item To overcome these limitations, a multivariate SSA (M-SSA) framework based on \cite{Wildi2024} is introduced, leveraging cross-sectional indicator data and controlling the predictor’s zero-crossing rate for enhanced signal tracking at turning points.
            \item A simple regression of the M-SSA GDP component on future GDP yields a statistically significant predictor up to one year ahead, outperforming both mean benchmarks and classic direct forecasts at longer horizons.
            \item Stability in regression coefficients and filter design after a burn-in period improves the interpretability of the predictor, offering a robust integration of direct forecasting with multivariate signal extraction targeting quarterly GDP growth.
        \end{itemize}
			\end{block}

	\begin{block}{\large References and Further Information}
	\begin{columns}[T]
		\column{0.3\textwidth}

		\column{0.5\textwidth}
				%\begin{flushright}	
                GIT: https://github.com/wiaidp/R-package-SSA-Predictor
				%\end{flushright}
				\column{0.2\textwidth}

			%\includegraphics[width = 0.55\textwidth]{../Figures/GIT.pdf}
		\end{columns}

	\end{block}
		\end{column}
	\bibliographystyle{apalike}
	\bibliography{references}
\end{columns}

\end{frame}

\end{document}

### 2025-08-08 Achive IWH Simon paper/08.08.2025/paper.tex
up

--- 2025-06-21 HWI SSA_27_05_2025_with new poster/22.06.2025/paper.tex
@@
 % \usepackage[justification=left,hang]{caption}
-\usepackage[justification=raggedrig,hang]{caption}
+\usepackage[justification=raggedright,hang]{caption}
 \usepackage{subcaption}
@@
 
-\title{Smooth and Persistent Forecasts of German GDP}
+\title{Smooth and Persistent Forecasts of German GDP\\ Balancing Accuracy and Stability\thanks{We would like Barbara Rossi and the participants at the Vienna Forecasting Workshop 2025 for their comments and suggestions.}}
 \author{Katja Heinisch\thanks{Halle Institute for Economic Research (IWH).} \hspace{0.2cm}
@@
 Marc Wildi\thanks{Zurich University of Applied Sciences (ZHAW).}}
-\date{May 2025 \\
+\date{July 2025 \\
 \vspace{2cm} \textit{Work in progress, not for circulation}}
@@
 \begin{abstract} %max.
-This paper addresses the trade-off between informational efficiency and forecast smoothness in economic forecasting.
-Efficient forecasts are often volatile, limiting their practical usefulness.
-We propose a novel method that employs the framework of Multivariate Smooth Sign Accuracy (M-SSA) to extract smoothed components from leading indicators.
-This approach enhances the signal-to-noise ratio and allows for systematic control over forecast volatility and timing.
-Applied to German GDP, our method yields smoothed forecasts that improve the accuracy of traditional models, particularly over medium-term horizons.
-While smoother forecasts may lag slightly, this can be mitigated by adjusting the forecast horizon.
-Our findings offer practical tools for forecasters and policymakers subject to adjustment costs or frictions in updating forecasts or implementing policy changes.
+This paper explores the trade-off between informational efficiency and forecast smoothness in economic forecasting.
+Forecasts that minimize mean squared forecast errors often exhibit excessive volatility, limiting their practical applicability.
+We propose the Multivariate Smooth Sign Accuracy (M-SSA) framework as a solution, extracting smoothed components from leading indicators.
+This enhances the signal-to-noise ratio and controls forecast volatility and timing.
+Applied to German GDP, our method yields smoothed forecasts that improve the forecasting accuracy of traditional models, particularly over medium-term horizons.
+While smoother forecasts tend to lag slightly, this drawback can be mitigated by adjusting the forecast horizon.
+These findings highlight the practicality of the M-SSA framework for both forecasters and policymakers, especially in settings where updating forecasts or policy adjustments are costly.
 
@@
 
-Economic forecasters face a dilemma: While efficiency suggests that their forecasts should incorporate all available information, the resulting forecasts are often volatile, which may reflect greater information content or contamination by excessive ``noise''.
+%Economic forecasters face a dilemma: While efficiency suggests that their forecasts should incorporate all available information, the resulting forecasts are often volatile, which may reflect greater information content or contamination by excessive ``noise''.
 Further, forecast users (including public and private decision makers) may be constrained by how rapidly they can adjust to changing forecasts.\footnote{For example, private or public spending decisions may be constrained by annual budgets.
@@
 
-%%% Further recent literature
-% \cite{lehmann2021predicting} show the performance of ifo busines cycles indicator for the nowcast (h=0) and forecast (h=1)
+Economic forecasters face a dilemma
+: while informational efficiency requires that forecasts incorporate all available data, doing so often results in forecasts that are excessively volatile.
+ %Volatile forecasts may be a sign that they efficiently incorporate all available information. 
+ In addition, forecasts may reflect excessive ``noise'' due to overfitting.
+Further, forecast users (including public and private decision makers) may face constraints in their speed of adjustment to changing forecasts.
+For example, fixed annual budgets may constrain spending, or transaction costs may limit changes in investments.
+Others may attempt to smooth their reactions for reasons of signalling.
+Important examples include dividend smoothing by corporate managers and interest rate smoothing by central banks.
 
 \cite{Wildi2024,Wildi2025} and \cite{McElroy2019,McElroy2020} propose novel methods that optimally smooth forecasts by controlling the expected frequency of sign changes.
-The resulting Smooth Sign Accuracy (SSA) framework encompasses efficient mean-square error (MSE) as a special case and produces increasingly smooth forecasts as changes in forecast sign are increasingly penalized.
+The resulting Smooth Sign Accuracy (SSA) framework encompasses efficiently the mean-square error (MSE) as a special case and produces increasingly smooth forecasts as changes in the forecast direction (sign) are increasingly penalized.
 
@@
 
-This study proposes a methodological innovation that extends the classical direct forecasting framework by integrating dimension-reduced information extracted through Multivariate Smooth Sign Accuracy (M-SSA).
+%This study proposes a methodological innovation that extends the classical direct forecasting framework by integrating dimension-reduced information extracted through Multivariate Smooth Sign Accuracy (M-SSA).
 The core idea is to replace raw
 %, unprocessed 
-leading indicators with smoothed and filtered components derived from the M-SSA procedure, thereby improving the signal-to-noise ratio in the predictor variables used to forecast GDP.
+%leading indicators with smoothed and filtered components derived from the M-SSA procedure, thereby improving the signal-to-noise ratio in the predictor variables used to forecast GDP.
 We apply this framework to well-established German economic indicators to show its ability to improve forecasts of GDP and GDP trends using a variety of sample periods and performance metrics.
 
-Surprisingly, we find that smoothing indicator-based forecasts often provides useful leading information for GDP, especially over mid-term horizons ranging from two to four quarters ahead.
-We also show that while smoother GDP forecasts lag efficient MSE forecasts slightly, this can be mitigated by adjustments in the forecast horizon.\footnote{The lag reflects the trilemma highlighted by \cite{McElroy2019} in which forecasters choose between forecast accuracy, timeliness and volatility.}
+This study extends this framework by replacing raw leading indicators with smoothed and filtered components, thereby improving the signal-to-noise ratio in the predictor variables.
+We apply this framework to well-established German economic indicators to show its ability to improve forecasts of GDP \citep{lehmann2021predicting,Heinisch2018bottom}  and GDP trends using a variety of sample periods and performance metrics.
+
+We find that smoothing indicator-based forecasts often provides useful leading information for GDP, especially over medium-term horizons ranging from two to four quarters ahead.
+We also show that while smoother GDP forecasts lag efficient MSE forecasts slightly, this can be mitigated by adjustments in the forecast horizon.\footnote{The lag reflects the trilemma highlighted by \cite{McElroy2019} in which forecasters choose between forecast accuracy, timeliness, and volatility.}
 
 %Previous research often highlights the utility of direct forecasts, particularly for short-term predictions such as nowcasting and forecasts one quarter ahead.
-However, this paper advances the literature by showing that more sophisticated predictors—specifically, filtered components obtained via M-SSA yield superior performance, especially over mid-term horizons ranging from two to four quarters ahead.
-
+However, this paper advances the literature by showing that more sophisticated predictors—specifically, filtered components obtained via M-SSA yield superior performance, especially over medium-term horizons ranging from two to four quarters ahead.
+
+We proceed as follows.
 Section \ref{sec:data} describes our data series, the sample period used, and documents their dynamic behaviour.
-Section \ref{sec:direct_forecast} provides benchmark forecasts using simple direct projections of GDP based on our indicators.
+Section \ref{sec:direct_forecast} provides benchmark forecasts using simple direct projections of GDP based on selected indicators.
 We demonstrate how in-sample forecast performance may improve when we ``de-noise'' the indicators prior to projection with simple smoothing filters.
 Section \ref{sec:mSSA} defines the M-SSA method and shows how its forecast performance compares to that of direct projections as well as how it varies with forecast horizon.
-Section \ref{sec:mSSA_component} considers the ability of our indicators to forecast German GDP growth using the M-SSA framework, and Section \ref{sec:revisions} investigates the temporal stability of the estimated forecasting equations.
-Section \ref{sec:conclusions} summarizes the results and concludes.
-The Appendix provides some additional results.
+Section \ref{sec:mSSA_component} considers the ability of our indicators to forecast German GDP growth out-of-sample using the M-SSA framework, and Section \ref{sec:revisions} investigates the temporal stability of the estimated forecasting equations.
+Section \ref{sec:conclusions} summarizes our results and concludes.
 
 \section{Data and Dependence}\label{sec:data}
-\subsection{Data}
+\subsection{Data and Publication Lags}
 The analysis employs a set of well-established leading indicators to forecast quarterly German GDP \citep{Drechsel2012financial,Heinisch2018bottom}, including industrial production (IP), the ifo Business Climate Index (ifo\_c), the Economic Sentiment Indicator (ESI), and the term spread between 10-year bond yields and the 3-month EURIBOR rate (spre\_10y\_3m) covering the period 1995 to 2024.
-We take into account the publication lag of the indicators and the GDP series.\footnote{Data revisions are not considered, given that previous studies \citep{Heinisch2019} have shown that their overall effect on forecast performance is minor.}
-While GDP is available with a one-quarter delay (published in mid of the next quarter), and IP is released with a one-month lag, business surveys and financial data are published without any lag, see Table \ref{tab:data_stucture} which presents the data structure as of the end of January 2025, just before the publication of the Flash estimate for the previous quarter.\footnote{Starting in the second quarter of 2020, the German Federal Statistical Office published for the first time the gross domestic product (GDP) 30 days after the end of the quarter.} All series are consequently re-aligned to correspond with the information available to forecasters, with GDP and IP variables shifted forward according to their respective lags.
+We take into account the publication lag of the indicators and the GDP series (Table \ref{tab:data_stucture}).\footnote{Data revisions are not considered, given that previous studies \citep{Heinisch2019} have shown that their overall effect on forecast performance is minor.}
+While GDP is available with a one-quarter delay, and IP is released with a two-month lag, business surveys and financial data are published without any lag.
+%All series are consequently re-aligned to correspond with the information available to forecasters, with GDP and IP variables shifted forward according to their respective lags.
 Table \ref{tab:data_stucture_aligned} displays the dataset as of the end of 2024, with GDP and IP variables temporally aligned at the sample endpoint.
@@
   \hline
- & GDP & ip & ifo\_c & ESI & spr\_10y\_3m \\ 
+ & GDP & IP & ifo\_c & ESI & spr\_10y\_3m \\ 
   \hline
@@
 \end{tabular}
-\caption{Ragged end as of January 2025, before publication of the flash estimate.} 
+\caption{Ragged edge structure. As of January 2025, before publication of the GDP flash estimate.} 
 \label{tab:data_stucture}
@@
 
+To allow for missing observations due to publication lags, we re-align series to reflect the information available to forecasters.
+Table \ref{tab:data_stucture_aligned} displays the result as of the end of 2024, with GDP and IP columns shifted to reflect their release dates rather than the periods that they measure.
+This will be the data configuration used for nowcasting or forecasting GDP.
+For illustrative purposes, an additional target column is included on the right side of the table, representing the dependent variable for nowcasting applications.\footnote{In the case of an $h$-step ahead forecast, this column is shifted upwards by $h$ quarters.}
+
 \begin{table}[ht]
@@
   \hline
- & GDP & ip & ifo\_c & ESI & spr\_10y\_3m &&Target (nowcast)\\ 
+ & GDP & IP & ifo\_c & ESI & spr\_10y\_3m &&Target (nowcast)\\ 
   \hline
@@
 \end{tabular}
-\caption{Columns 1 through 5 constitute the matrix of explanatory variables available in Q4-2024 for the establishment of a GDP nowcast or forecast.
-All series are artificially aligned at the sample endpoint.
-Column 6 represents the target variable for the nowcast in Q4-2024.}
+\caption{Aligned dataset.
+At each date, columns 1 through 5 show the most recent values of each predictor available to forecasters.
+Column 6 represents the target variable for a GDP nowcast (i.e.
+the values of column 1 shifted upwards by the publication lag of GDP.}
 \label{tab:data_stucture_aligned}
@@
 
-To ensure comparability and mitigate scale effects, all indicators (except the spread, which is differenced only) are log-differenced and standardized, the latter applied solely for ease of visual inspection.
+To ensure comparability and mitigate scale effects, all indicators (except the term spread, which is differenced only) are log-differenced and standardized.
 Additionally, extreme outliers associated with the COVID-19 pandemic are trimmed to prevent distortions in subsequent figures.
- Monthly series are transformed to quarterly series and are displayed in Figure \ref{fig:data}, while Figure \ref{fig:data_lags} offers a close-up view of the financial crisis and the pandemic.\footnote{Monthly series are averaged to quarterly values.} The latter figure illustrates that the real-time GDP and industrial production series are synchronized at dips and peaks (coincident), whereas the other indicators tend to lead somewhat.
-The latter can be leveraged in a multivariate approach to enhance forecasting performance compared to univariate benchmarks.
+ Monthly series are shifted (Table \ref{tab:data_stucture_aligned}) and averaged to quarterly series (Figure \ref{fig:data}), while Figure \ref{fig:data_lags} offers a close-up view of the financial crisis and the pandemic.
+The latter figure shows that the real-time GDP and industrial production series are synchronized at turning points, whereas the other indicators tend to lead somewhat.
+The challenge for forecasters is to use the latter in a multivariate approach to improve forecasting performance over that of univariate benchmarks.
 
@@
 \subsection{Dependence}
-To examine the dynamic interdependencies among the variables, we estimate their cross-correlations as well as their vector moving-average (VMA) representations.
+To examine the joint dynamics of our variables, we estimate their cross-correlations as well as their vector moving-average (VMA) representations.
 The latter is a key input to the multivariate filters we develop in Sections \ref{sec:mSSA} and \ref{sec:mSSA_component}.
@@
 Figures \ref{CCF} (full sample) and \ref{CCF_wc} (omitting the Pandemic period) show the sample cross-correlation functions (CCF's) between current and lagged GDP and the two survey indicators.
-Both figures show strong correlations that peak at leads relative to GDP.
-Results for IP showed no lead, while correlations for the term spread were weaker.
+Both figures show strong correlations that peak when GDP is slightly lagged.
+Results for IP showed no such lag, while correlations for the term spread were weaker.
 The figures also show that the precise strengths of the correlations are influenced by the outliers associated with the COVID-19 pandemic.
@@
 For additional details, see \cite{ElasticNet} and the references therein.} The figures \ref{ma_inv_BIP} and \ref{ma_inv_BIP_var3_en} show the implied responses of GDP to shocks to each indicator variable.\footnote{Shown are the MA coefficients for GDP from the vector moving-average (VMA) models obtained by inversion of the VAR(1) and VAR(3) representations respectively (i.e.
-the impulse-response functions, or IRFs, of GDP to the correlated reduced-form shocks.)}
+the impulse-response functions (IRFs) of GDP to the correlated reduced-form shocks.)}
 
@@
 %The models encapsulate key characteristics of the data after adjusting for publication lags. Notably, 
-GDP exhibits the strongest response to shocks in the two survey indicators—ifo and ESI\footnote{There is also weaker evidence of a response to the term spread at longer lags.}.
-The impact of the pandemic modifies the lag structure by emphasizing shorter dependencies triggered by the alternating outliers during that specific episode.
-Removing the COVID-19 breakout from the analysis results in smoother, somewhat longer-tailed dependencies, consistent with the observed empirical autocorrelation functions.
-In particular, the VAR(3) model captures slightly extended lead times and more nuanced lag patterns\footnote{The VAR(3) model yields marginal improvements in multivariate filtering performance.
+GDP exhibits the strongest response to shocks in the two survey indicators—ifo and ESI.\footnote{There is also weaker evidence of a response to the term spread at longer lags.} The impact of the pandemic modifies the lag structure by emphasizing shorter dependencies triggered by the alternating outliers during that specific episode.
+Removing the pandemic from the analysis results in smoother, somewhat longer-tailed dependencies, consistent with the observed empirical autocorrelation functions.
+In particular, the VAR(3) model captures slightly longer lead times and more nuanced lag patterns.\footnote{The VAR(3) model yields marginal improvements in multivariate filtering performance.
 However, due to its relative simplicity, this study primarily employs the VAR(1) model.
-The M-SSA package (accessible at \url{https://github.com/wiaidp/R-package-SSA-Predictor}) offers more advanced capabilities, including the implementation of the VAR(3) model and various additional (Bayesian) VAR specifications.}.
-The delayed response of GDP to movements in these indicators suggests that multivariate models could yield superior forecasting performance compared to univariate approaches, a hypothesis we explore in the next section
+The M-SSA package (accessible at \url{https://github.com/wiaidp/R-package-SSA-Predictor}) offers more advanced capabilities, including the implementation of the VAR(3) model and Bayesian VAR specifications.}
+
+The delayed response of GDP to movements in the survey indicators suggests that multivariate models could yield better forecasts than those of univariate model, a hypothesis we explore in the next section.
 
@@
 In this section, we demonstrate how pre-filtering indicator variables may improve the performance of simple forecasts by removing noise.
+We then demonstrate how in-sample forecast performance improves when we ``de-noise'' the indicators prior to projection with simple smoothing filters.\footnote{The smoothing filters are one-sided (concurrent) filters of length 31, therefore, estimation of both forecasting models omits the first 30 observations.}
+
 For simplicity, all the models in this section are direct projections of GDP on our indicators and take the form
@@
 \begin{equation}
-    u_{t+h} = \beta' \cdot \mathbf{X}_t + \varepsilon_t
+    u_{t+h} = \boldsymbol{\beta'} \cdot \mathbf{X}_t + \varepsilon_t
     \label{eq:direct_forecast}
@@
 \begin{itemize}
-    \item $u_{t+h}$ is the growth rate of GDP, $h = \{0, 1, 2, ... \}$
+    \item $u_{t+h}$ is the quarterly growth rate of GDP for forecast horizon $h = \{0, 1, 2, ... \}$
     \item $\mathbf{X}_t \equiv [1,u_{t-1}, ip_t, ifo\_c_t, ESI_t, spr\_10y\_3m_t]'$ or their transformed counterparts and $u_{t-1}$ accounts for the publication lag of GDP.\footnote{Including lagged GDP enables the nesting of the AR forecast within the broader modeling framework.}
-    \item $\hat{\beta}$ is estimated by least-squares, omitting observations affected by the pandemic, i.e. Q4-2019 to Q4-2020.
+    \item $\hat{\boldsymbol{\beta}}$ is estimated by least-squares, omitting observations affected by the pandemic, i.e. Q4-2019 to Q4-2020.
 \end{itemize}
-
-We then demonstrate how in-sample forecast performance improves when we ``de-noise'' the indicators prior to projection with simple smoothing filters.\footnote{The smoothing filters are one-sided (concurrent) filters of length 31, therefore, estimation of both forecasting models omits the first 30 observations.}
 
@@
         %\includegraphics[height=3in, width=4.5in]{./Figures/direct_wc_all.pdf}
-        \caption{Direct forecasts (red lines) based on all indicators and full sample information (without pandemic): GDP (black line) left-shifted  by zero, one, two, and three quarters.
+        \caption{Direct forecasts (red lines) based on all indicators and full sample information (without pandemic.)\\
+        Target (black line): GDP left-shifted  by 0 - 3 quarters.
         \label{direct_wc}}
@@
 
-Figures \ref{direct_wc} and \ref{direct_wc_financial_crisis} compare actual and fitted values for GDP at forecast horizons $h = 0, \ldots, 3$, using the four indicators and the target, i.e.
-GDP shifted forward (leftwards) by $h$ quarters, without applying any filtering.
+Figures \ref{direct_wc} and \ref{direct_wc_financial_crisis} compare actual and fitted values for GDP at forecast horizons $h = 0, \ldots, 3$, using the four indicators and the target, i.e., GDP shifted $h$ quarters ahead.
 Figure \ref{direct_wc} shows results over the sample without the pandemic, while Figure \ref{direct_wc_financial_crisis} provides a closer look at the fit during the financial crisis.
@@
 
-The results indicate that as the forecast horizon $h$ increases, the turning points (i.e., peaks and troughs) of the direct forecasts become increasingly right-shifted (lagged) relative to the target GDP.
+The results indicate that as the forecast horizon $h$ increases, forecast become less volatile and their turning points (i.e., peaks and troughs) become increasingly right-shifted (lagged) relative to the target GDP.
 Additionally, the p-values of the HAC-adjusted Wald test precipitously increase from $h=2$ to $h=3$ (see Table \ref{tab:pvaluedhp}), implying that the indicators become statistically insignificant for forecast horizons of more than two quarters.\footnote{
 These findings are confirmed by the relative root mean-square errors comparing direct forecasts against the mean benchmark, see Table \ref{rRMSE_mSSA_direct_mean_without_covid8} in the Appendix.}
+
+% latex table generated in R 4.2.2 by xtable 1.8-4 package
+% Wed May  7 12:49:46 2025
+
+%\begin{table}[ht]
+%\centering
+%\begin{tabular}{rrrrr}
+%  \hline
+% & h=3 & h=4 & h=5 & h=6 \\ 
+%  \hline
+%rRMSE & 0.962 & 0.967 & 0.990 & 0.967 \\ 
+%   \hline
+%\end{tabular}
+%\caption{Relative root mean-square error of HP-C against classic direct forecast for $h\geq %3$.\\Values $<1$ imply  HP-C has smaller RMSE.  } 
+%\label{tab:rRMSE}
+%\end{table}
+\bigskip
+\begin{table}[h]
+\centering
+\begin{tabular}{lrrrrrr}
+  \hline
+  & h= 1 & h= 2 & h= 3 & h= 4 & h= 5 & h= 6 \\ 
+  \hline
+  Unfiltered & 0.000 & 0.001 & 0.910 & 0.509 & 0.091 & 0.178 \\ 
+HP-C filtered & 0.001 & 0.036 & 0.046 & 0.049 & 0.012 & 0.022 \\ 
+   \hline
+\end{tabular}
+\caption{p-values for $H_0: \boldsymbol{\beta} = 0$ in \ref{eq:direct_forecast} based on HAC-adjusted Wald test.\\Estimation over full sample (without pandemic.)} 
+\label{tab:pvaluedhp}
+\end{table}
 
@@
 
-The Hodrick-Prescott (HP) filter is a classic tool used in business cycle analysis: the filter is specified by a single smoothing parameter $\lambda$ and the value $\lambda=1600$ is specifically recommended for quarterly data.
-However, \cite{Phillips_Jin_2021} noted that the HP(1600) filter removes relevant information due to excessive smoothing.
-In our context, this oversmoothing issue would be further exacerbated when considering forecast horizons shorter than a year---an interval inconsistent with the mean duration of up to several years of business cycles, as highlighted by the HP(1600).
-To address this concern, we suggest a more adaptive HP(160) `target' filter.
+The Hodrick-Prescott (HP) filter is a classic tool used in business cycle analysis: the filter is specified by a single smoothing parameter $\lambda$.
+The value $\lambda=1600$ is the most common value used for business cycle analysis, but \cite{Phillips_Jin_2021} note that this removes relevant information due to excessive smoothing.
+%In our context, this oversmoothing issue would be further exacerbated when considering forecast horizons shorter than a year---an interval inconsistent with the mean duration of up to several years of business cycles, as highlighted by the HP(1600).
+To address this concern, we instead use $\lambda = 160$ to allow for a more adaptive filter.
 
 Figure \ref{hp_160} presents both classic two-sided and one-sided HP(160) (denoted HP-C for \textit{concurrent}) filters (top panels) along with their associated amplitude functions in the frequency domain (bottom panels).
-It confirms that the HP-C filter assigns greater weight to yearly components—relevant in a mid-term forecast context— than the classic quarterly HP filter, while also reducing undesirable high-frequency noise.
+It confirms that the HP-C filter assigns greater weight to yearly components—relevant in a medium-term forecast context— than the classic quarterly HP filter, while also reducing undesirable high-frequency noise.
 Therefore, while the choice of the HP(160) filter in this example
@@
     \begin{center}
+    \textcolor{blue}{\textbf{Notice that the title on top right panel is incorrect!}}
         \includegraphics[height=4in, width=5.5in]{./Figures/hp_160.pdf}
-        \caption{Two-sided HP(160) (top left) and one-sided (concurrent) HP-C filters (top right) with corresponding amplitude functions (bottom).
+        % \caption{Two-sided HP(160) (top left) and one-sided (concurrent) HP-C filters (top right) with corresponding amplitude functions (bottom).
+        \caption{HP$(\lambda=160)$ filters\\
+        Left: HP (Two-sided, symmetric) \qquad Right: HP-C (One-sided, concurrent)\\
+        Top Row: Filter Weights \qquad \qquad \quad Bottom Row: Amplitude Functions.
         \label{hp_160}}
@@
 
-Figures \ref{direct_hp_forecasts} and \ref{direct_hp_forecasts_financial_crisis} provide results comparable to those of Figures \ref{direct_wc} and \ref{direct_wc_financial_crisis} but now using indicators smoothed with the HP-C filter, covering the full sample without the pandemic and providing a closer look at performance during the financial crisis.
+Figures \ref{direct_hp_forecasts} and \ref{direct_hp_forecasts_financial_crisis} provide results comparable to those of Figures \ref{direct_wc} and \ref{direct_wc_financial_crisis} but now using indicators smoothed with the one-sided HP(160), (denoted as the HP-C, for \textit{concurrent}) filter.
 The fitted values lag GDP slightly less and track turning points somewhat more quickly.
-Table \ref{tab:pvaluedhp} shows p-values for tests of the joint null hypothesis that all coefficients $\hat{\beta} = 0$.
+Table \ref{tab:pvaluedhp} shows p-values for tests of the joint null hypothesis that all coefficients ${\boldsymbol{\beta}} = 0$.
 The use of filtered indicators improves their statistical significance at forecast horizons $h > 2$ quarters.
@@
 
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 12:49:46 2025
-
-%\begin{table}[ht]
-%\centering
-%\begin{tabular}{rrrrr}
-%  \hline
-% & h=3 & h=4 & h=5 & h=6 \\ 
-%  \hline
-%rRMSE & 0.962 & 0.967 & 0.990 & 0.967 \\ 
-%   \hline
-%\end{tabular}
-%\caption{Relative root mean-square error of HP-C against classic direct forecast for $h\geq %3$.\\Values $<1$ imply  HP-C has smaller RMSE.  } 
-%\label{tab:rRMSE}
-%\end{table}
-\begin{table}[h]
-\centering
-\begin{tabular}{lrrrrrr}
-  \hline
-  & h= 1 & h= 2 & h= 3 & h= 4 & h= 5 & h= 6 \\ 
-  \hline
-  Unfiltered & 0.000 & 0.001 & 0.910 & 0.509 & 0.091 & 0.178 \\ 
-HP-C filtered & 0.001 & 0.036 & 0.046 & 0.049 & 0.012 & 0.022 \\ 
-   \hline
-\end{tabular}
-\caption{p-values for $H_0: \hat{\beta} = 0$, HAC-adjusted Wald test.\\Estimation over full sample (without pandemic.)} 
-\label{tab:pvaluedhp}
-\end{table}
-
 %We now propose a more refined causal filter design aiming for statistical significance of the resulting predictor at longer forecast horizons, out-of-sample.
 
+The application of the one-sided HP-C filter to the indicators yields some measurable improvements in forecast performance. However, 
+%its univariate nature excludes additional information such as that provided by leading indicators. 
+the noise suppression capability of the one-sided filter is compromised, as explained by \cite{Wildi2025} and illustrated by the high-frequency leakage of the amplitude function in Figure \ref{hp_160}.
+To address these limitations, we turn to the Multivariate Smooth Sign Accuracy (M-SSA) framework proposed by \cite{Wildi2025}.
+
 \section{Multivariate Causal Filter: M-SSA}\label{sec:mSSA}
 
-The application of the one-sided HP-C filter to the indicators yields some measurable improvements in forecast performance.
-However, its univariate nature excludes additional information such as that provided by leading indicators.
-The noise suppression capability of the one-sided filter is compromised, as explained by \cite{Wildi2025} and illustrated by the high-frequency leakage of the amplitude function in Figure \ref{hp_160}.
-To address these limitations, we propose an extension of the Smooth Sign Accuracy (SSA) framework introduced by \cite{Wildi2025}.
-
-\subsection{Optimization criterion}
+To understand the smoothing properties of the M-SSA framework, we begin with the Smooth Sign Accuracy (SSA) case, which considers a univariate smoothing problem, before turning to M-SSA, its multivariate generalization.\footnote{See \cite{Wildi2024} for further details on the SSA framework and its applications.} In the remainder of the section, we demonstrate the properties of M-SSA nowcasts and forecasts for smoothed GDP.
+
+\subsection{SSA Optimization criterion}
 
@@
 The parameter $\rho_1$ controls the smoothness of the predictor: higher values of $\rho_1$ favour smoother trajectories of $y_t$.
-Assuming that the process $x_t$ has mean zero, \cite{Wildi2024} establishes a connection between $\rho(y,1)$ and the expected duration between consecutive `zero-crossings' (sign changes) of $y_t$, the `holding time', denoted by $HT(y)$\footnote{Formally, the precise relationship between the expected duration between sign changes and the lag-one autocorrelation of the predictor is derived under the assumption of Gaussian time series.
+Assuming that the process $x_t$ has mean zero, \cite{Wildi2024} establishes a connection between $\rho(y,1)$ and the expected duration between consecutive `zero-crossings' (sign changes) of $y_t$, the `holding time', denoted by $HT(y)$\footnote{Of course, when $z_t$ is GDP growth, sign changes are indicative of expansions and recessions.
+Formally, the precise relationship between the expected duration between sign changes and the lag-one autocorrelation of the predictor is derived under the assumption of Gaussian time series.
 However, \cite{Wildi2024} demonstrates that this relationship remains robust even when deviations from Gaussianity occur (central limit theorem applied to lowpass-filtered series).}:
@@
 
-Eq.
-\eqref{critssaht} has several important implications.
-First, the objective function is indifferent to an affine transformation of the predictor.
+One implication of Eq.
+\eqref{critssaht} is that the objective function is indifferent to an affine transformation of the predictor.
 This ambiguity can be resolved by assuming an arbitrary scale and level for $y_t$ (standardization).
@@
 %If the MSE predictor is noisy, which is often the case in applications, one can select $HT_1>HT_{MSE}$  in the constraint to address the rate of noisy crossings or false alarms.
-Second, the concept can be extended to a multivariate framework, denoted M-SSA, by generalizing the correlation functions as detailed in \cite{Wildi2025}. 
+
+Unfortunately, classic predictors often generate excessively many false (`noisy') sign changes due to high-frequency leakage, as discussed above (see Figure \ref{hp_160}.)
+%as illustrated by the amplitude function in Figure \eqref{hp_160} (bottom right panel) and further elaborated upon by \cite{Wildi2025}. 
+ \cite{Wildi2025} shows that the SSA criterion helps control this phenomenon (see Fig.\ref{mssa_msse_zc} for illustration.) Using the dual formulation of the criterion \eqref{critssaht}, the author demonstrates that the SSA solution yields the lowest zero-crossing rate among all (linear) predictors with equivalent target correlation.
+
+The concept can be extended to a multivariate framework, denoted M-SSA, by generalizing the correlation functions as detailed in \cite{Wildi2025}.
+
+\subsection{M-SSA Optimization}
+
+\textcolor{blue}{This is where Marc outlines the generalization of the correlation function and other elements of M-SSA.}
+
+ Consequently, the subsequent analysis employs the M-SSA framework to construct a `smooth' predictor $y_t$ for forecasting $z_{t+h}$.
+In particular, it is shown that controlling the zero-crossing rate within the M-SSA framework has a quantifiable effect on the hit rate and false positive rate.
+
 %Lastly, sign changes in trend growth, represented by zero-crossings of the target $z_t$, are indicative of relevant changes in the trajectory of the economy, for example, at transitions between expansions and recessions.
-Lastly, when $z_t$ is GDP growth, sign changes are indicative of expansions and recessions.
-Unfortunately, classic predictors often generate excessively many false (`noisy') sign changes due to high-frequency leakage, as discussed above.
-%as illustrated by the amplitude function in Figure \eqref{hp_160} (bottom right panel) and further elaborated upon by \cite{Wildi2025}. 
- \cite{Wildi2025} shows that the SSA criterion helps control this phenomenon,  see Fig.\ref{mssa_msse_zc} for illustration.
-Using the equivalent dual formulation of the criterion \eqref{critssaht}, the author demonstrates that the SSA solution exhibits the lowest zero-crossing rate of any (linear) predictor with the same target correlation.
-In the following, we therefore use the M-SSA framework to construct a `smooth' predictor $y_t$ to forecast $z_t$.
+
  %In a second stage, serving as a generalization and refinement of the univariate HP-C utilized in the direct HP forecast.
@@
 Both predictors use the same underlying VAR(1) model introduced in Section \ref{sec:direct_forecast} and estimated on data up to Q4-2007.
-This ensures a lengthy out-of-sample period that includes significant events such as the financial crisis, the sovereign debt crisis, and the COVID-19 pandemic\footnote{In the parsimonious VAR(1) model estimated over the shorter sub-sample ending in Q4-2007, the economic sentiment indicator (ESI) does not receive any weight.
-Additional technical details and background information are provided in \cite{Wildi2025}.
+This ensures a lengthy out-of-sample period that includes significant events such as the financial crisis, the sovereign debt crisis, and the COVID-19 pandemic.\footnote{The economic sentiment indicator (ESI) does not receive any weight in the parsimonious VAR(1) model estimated over the shorter sub-sample ending in Q4-2007.
+Technical details and background information on the M-SSA are provided in \cite{Wildi2025}.
 All computations are conducted using the M-SSA package, as documented in \cite{Wildi2025}.
@@
         \includegraphics[height=3in, width=4.5in]{./Figures/bk_gammak.pdf}
-        \caption{Multivariate M-MSE (left) and M-SSA (right) nowcast filter weights.
+        \caption{Multivariate M-MSE (left) and M-SSA (right) nowcast filter weights.\\
+        Note difference in Scales.
         \label{bk_gammak}}
@@
         \includegraphics[height=3in, width=4.5in]{./Figures/mssa_msse_now.pdf}
-        \caption{Two-sided HP(160) applied to GDP (black)\\ Nowcasts: M-SSA (blue) and M-MSE (green).
-\\The dashed vertical line delimits in- and out-of sample periods.
-\\The two-sided filter does not extend to the sample end.
+        \caption{Nowcasts of Smoothed GDP\\
+        \indent Black: Target (Two-sided HP(160) applied to GDP) \\ 
+        \indent Blue: M-SSA Nowcast\\
+        \indent Green: M-MSE Nowcast\\
+        \indent The dashed vertical line delimits in- and out-of sample periods.  \\
+        Note that the two-sided filter does not extend to the sample end.
         \label{mssa_msse_now}}
@@
 
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/mssa_msse_zc.pdf}\caption{Sign changes of the M-MSE (green) and M-SSA (blue) nowcasts towards the financial crisis, with sign changes of the M-MSE marked by green vertical lines.
+\begin{figure}[H]
+    \begin{center}
+    \textcolor{blue}{@Marc: Please change the plot Title to "M-MSE \& M-SSA Nowcasts: Financial Crisis".)}
+        \includegraphics[height=3in, width=4.5in]{./Figures/mssa_msse_zc.pdf}
+        % \caption{Sign changes of the M-MSE (green) and M-SSA (blue) nowcasts towards the financial crisis, with sign changes of the M-MSE marked by green vertical lines.
 The sample holding time (HT) of each filter corresponds to the average duration between its consecutive zero-crossings.
 For each filter, its sample HT converges to the respective expected HT asymptotically.
-Here, we impose  $HT_{MSSA}=1.5\times HT_{MSE}$: in the long run, we expect the M-MSE to exhibit approximately $50\%$ more sign changes than the M-SSA.\label{mssa_msse_zc}}\end{center}\end{figure}
+Here, we impose  $HT_{MSSA}=1.5\times HT_{MSE}$: in the long run, we expect the M-MSE to exhibit approximately $50\%$ more sign changes than the M-SSA.
+        \caption{M-MSE (green) and M-SSA (blue) nowcasts around the financial crisis.\\
+        Sign changes of the M-MSE marked by green vertical lines.\\
+        The holding time (HT) of each filter corresponds to the expected duration between its sign-changes. 
+        % For each filter, its sample HT converges to the respective expected HT asymptotically. 
+        Here, we impose  $HT_{MSSA}=1.5\times HT_{MSE}$. In large samples, we expect the M-SSA to exhibit approximately $33\%$ fewer sign changes than the M-MSE.
+        \label{mssa_msse_zc}}
+    \end{center}
+\end{figure}
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
@@
 \end{tabular}
-\caption{Sample target correlations and holding time (HT) of M-SSA and M-MSE nowcasts.} 
+\caption{Sample target correlations and holding times (HT) of M-SSA and M-MSE nowcasts.} 
 \label{corhtnow}
@@
 Increased smoothness, i.e.
-holding time increases from 3.7 to 6.5 quarters, achieved by the M-SSA comes at the cost of a slight reduction (0.68 vs 0.69) in target correlation.
+sample holding time increases from 3.7 to 6.5 quarters, achieved by the M-SSA comes at the cost of a slight reduction (0.68 vs 0.69) in target correlation.
 This highlights the inherent trade-off—often referred to as the prediction dilemma—between smoothness (zero-crossing rate) and predictive accuracy (target correlation) within the M-SSA criterion.
-The results validate that the smoothing constraint incorporated into the multivariate extension of criterion \eqref{critssa} effectively manages the sign-changes of the predictor: sample estimates of the holding times align with the expected values derived from Eq.~\eqref{ht}.\footnote{Effective convergence of sample estimates to their theoretical expected values —assuming the true data-generating process is known— can be validated using the M-SSA package.
-This validation is typically performed through analyses of very long samples of simulated (multivariate) data, ensuring the consistency of the estimation procedures under idealized conditions.} An increase of $50\%$ in the HT relative to the M-MSE predictor has only a modest effect on the target correlation, suggesting that this trade-off could be beneficial in our application.
-%Going forward, we now omit the M-MSE, as it is a special case of the more general M-SSA framework.
+The results confirm that the smoothing constraint in M-SSA
+% incorporated into the multivariate extension of criterion \eqref{critssa} effectively manages the sign-changes of the predictor: sample estimates of the holding times align with the expected values derived from Eq.~\eqref{ht}.\footnote{Effective convergence of sample estimates to their theoretical expected values —assuming the true data-generating process is known— can be validated using the M-SSA package.
+This validation is typically performed through analyses of very long samples of simulated (multivariate) data, ensuring the consistency of the estimation procedures under idealized conditions.} An
+increases by $50\%$ the HT relative to that of the M-MSE predictor but has only a modest effect on the target correlation, suggesting that this trade-off could be beneficial in our application
+% , as confirmed by a subsequent analysis of hit rate and false alarm rate
+. %Going forward, we now omit the M-MSE, as it is a special case of the more general M-SSA framework.
 
@@
 
-Next, we extend the previous example to consider forecasts of HP-GDP.
-The only modification this requires is to change $h=0$ (nowcast) to $h=1,...,6$ quarters.
+Next, we 
+% extend the previous example to 
+consider forecasts of HP-GDP, changing the previous
+%. The only modification this requires is to change 
+$h=0$ (nowcast) to $h=1,...,6$ quarters. 
 %The previously described nowcast example can be systematically extended to generate forecasts of the target variable.
@@
 %This approach facilitates the assessment of forecast performance across multiple horizons within the established methodological context.
-The influence of the forecast horizon on the M-SSA design is depicted in Figure \ref{bk_h}, which contrasts M-SSA filter weights for the nowcast and the one-year-ahead forecast.
+The influence of the forecast horizon on the M-SSA design is depicted in Figure \ref{bk_h}, which compares M-SSA filter weights for the nowcast and the one-year-ahead forecast.
 Noting the difference in vertical scales, we see that as $h$ increases (right panel), the filter weights diminish, reflecting increased forecast uncertainty.
-We also see increased weight on the leading indicators relative to that on GDP.\footnote{Weights on IP and ESI are indistinguishable from zero.} Additionally, a phase shift associated with the filter applied to GDP is observed (black line, right panel), reflecting the cyclical characteristics inherent to the HP filter.
-While this phase effect might cause an effective sign change in the GDP filter output as $h$ increases—something to be avoided in this context—the low-pass filters assigned to the additional  `leading' indicators (green and yellow lines, right panel) continue to effectively track the low-frequency components of HP-GDP.
+We also see increased weight on the leading indicators relative to that on GDP.\footnote{Weights on IP and ESI are indistinguishable from zero.} Additionally, a phase shift associated with the filter applied to GDP is observed (black line, right panel), reflecting the cyclical characteristics of the HP filter.
+While this phase effect might cause a sign change in the forecast as $h$ increases — something to be avoided in this context — the low-pass filters assigned to the additional  `leading' indicators (green and yellow lines, right panel) continue to effectively track the low-frequency components of HP-GDP.
 The combination of the phase effect, applied to GDP, and level-tracking through these additional indicators enables more refined and effective tracking of the target compared to univariate forecasts, which rely solely on the phase-effect of the filter to look ahead.
@@
         \includegraphics[width=0.85\textwidth]{./Figures/bk_h.pdf}
-        \caption{Multivariate M-SSA filter weights: $h=0$ (nowcast, left) and $h=4$ (one year forecast, right).
+        \caption{Multivariate M-SSA filter weights\\
+        Left: $h=0$ (nowcast) \qquad Right: $h=4$ (one year forecast)\\
+        Note the difference in scales.
         \label{bk_h}}
@@
 \end{figure}
+
+M-SSA predictors are compared to the traditional univariate HP-C in Figures  \ref{multivar_vs_univar} and \ref{mssa_hpc_financial_crisis}.
+In the first figure, we compare both filters across the periods of the financial crisis (top panels) and the COVID-19 pandemic (bottom panels), for forecast horizons $h=0,...,6$.
+In the second figure, we compare both filters to the forward-shifted GDP for shifts larger than two quarters.
+All series are standardized for ease of comparison.
+%The nowcast from the previous section is included for reference.
 
@@
 
-M-SSA predictors are compared to the traditional univariate HP-C in Figures  \ref{multivar_vs_univar} and \ref{mssa_hpc_financial_crisis}.
-In the first figure, we compare both filters across the periods of the financial crisis (top panels) and the COVID-19 pandemic (bottom panels), for forecast horizons $h=0,...,6$.
-%The nowcast from the previous section is included for reference. 
-In the second figure, we compare both filters to the forward-shifted GDP for shifts larger than two quarters.
-All series are standardized for ease of comparison.
 Forecasts using the M-SSA approach consistently shift leftwards as $h$ increases; this shift appears to be consistent across all levels, including the peaks and troughs of the series.
@@
 However, our objective is to develop predictors for unfiltered GDP.
-To achieve this, we use the above M-SSA outputs $y_t^i, \: i = 1, \ldots, 5$ (referred to as M-SSA components) as predictors for forward-shifted GDP.
-While we keep the M-SSA fixed, as based on data up to Q4-2007, we implement an expanding window starting from Q1-2008 for deriving the component weights, which are updated on a quarterly basis. 
+To achieve this, we use the above M-SSA outputs as predictors for forward-shifted GDP.
+
 %(this contrasts with Section \ref{sec:direct_forecast}, which was based on the full sample information). 
@@
 The solid blue line (the nowcast) approaches the zero line near the end of the sample, with a recent upward movement mainly supported by the leading spread component.
-In contrast, the ifo- and ESI-components remain near the zero line, while the ip- and GDP-components appear to be awaiting additional evidence and confirmation before signaling a definitive movement.
+In contrast, the ifo- and ESI-components remain near the zero line, while the IP- and GDP-components appear to be awaiting additional evidence and confirmation before signaling a definitive movement.
 This type of assessment can be useful for evaluating the relevance of recent changes in the forecasted time series dynamics.
@@
 \end{equation}
-The purposes of the regression are twofold: first, the parameters enable affine transformations of the left-hand side variable, so that the standardized target $\textrm{GDP}_{t+shift}$ can be replaced by the original log-returns of GDP; second, the scale adjustment by $\beta_1(shift,h)$ compensates for the deflation (zero-shrinkage) of $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$, reflecting greater uncertainty  as the forecast horizon $h$ extends.
+%The purposes of the regression are twofold: first, the parameters enable affine transformations of the left-hand side variable, so that the standardized target $\textrm{GDP}_{t+shift}$ can be replaced by the original log-returns of GDP; second, the scale adjustment by $\beta_1(shift,h)$ compensates for the deflation (zero-shrinkage) of $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$, reflecting greater uncertainty  as the forecast horizon $h$ extends.
 This simple transformation allows $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$ to serve as a predictor for $\textrm{GDP}_{t+shift}$ even when $h\neq shift$.
 This flexibility enables leveraging the dynamic left-shift of $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$, illustrated in Figure \ref{multivar_vs_univar}, when $shift<h$.\\
-
 Define the (out-of-sample) GDP component predictor $\hat{\textrm{GDP}}_{t+shift}(h)$:  
@@
 where the regression parameters $\hat{\beta}_0(shift,h,t),\hat{\beta}_1(shift,h,t)$ are estimated using data up to time point $t-1$ (expanding window).
-Since M-SSA is based on data up to Q4-2007, $\hat{\textrm{GDP}}_{t+shift}(h)$ constitutes an out-of-sample predictor of $\textrm{GDP}_{t+shift}$ for all positive $shift\geq 0$ and all $t$ larger than Q4-2007.
+Since M-SSA is based on data up to Q4-2007, $\hat{\textrm{GDP}}_{t+shift}(h)$ constitutes an out-of-sample predictor of $\textrm{GDP}_{t+shift}$ for all $shift\geq 0$ and all $t$ larger than Q4-2007.
+\footnote{Equations \ref{eq:GDP_M-SSA} and \ref{oos_pred} could be extended to incorporate additional M-SSA components, but for illustration purposes, we focus on the simplest (and most robust) design.}
 Define the out-of-sample forecast error
@@
 \end{equation}
-We now assume that the regression parameters $\hat{\beta}_0(shift,h,t)$ and$,\hat{\beta}_1(shift,h,t)$ in Equation \ref{oos_pred} are up-dated quarterly.
-Note that Equations \ref{eq:GDP_M-SSA} and \ref{oos_pred} could be extended to incorporate additional M-SSA components, but for illustration purposes, we focus on the simplest (and most robust) design.
-
-\subsection{Out-of-Sample Forecast Performances}\label{oosp}
+%We now assume that the regression parameters $\hat{\beta}_0(shift,h,t)$ and$,\hat{\beta}_1(shift,h,t)$ in Equation \ref{oos_pred} are up-dated quarterly. Note that 
+While we keep the M-SSA outputs fixed, as based on data up to Q4-2007, we implement an expanding window starting from Q1-2008 for deriving the component weights, which are updated on a quarterly basis.
+
+\subsection{Relative MSFEs}\label{oosp}
 
@@
 \end{equation}
-for observations after Q4-2007: the significance of the corresponding HAC-adjusted statistics is indicated by a `star' symbol in Table \ref{rRMSE_mSSA_comp_mean_wc}\footnote{A single star means significance at the $5\%$ level; a double star means significance at the $0.1\%$ level.
+for observations after Q4-2007: the significance of the corresponding HAC-adjusted test statistics for $H_0 : \alpha_1 =0$ is indicated by a `*' symbol in Table \ref{rRMSE_mSSA_comp_mean_wc}\footnote{`*' indicates significance at the $5\%$ level; `**' indicates significance at the $0.1\%$ level.
 Note that statistical significance is influenced by the relatively short out-of-sample span, of length 17 years.}.
-The MSE forecast performance of $\hat{\textrm{GDP}}_{t+shift}(h)$ can be benchmarked against the mean predictor (see Table \ref{rRMSE_mSSA_comp_mean_wc}) as well as against the direct forecast (see Table \ref{rRMSE_mSSA_comp_direct_without_covid7}).
+The MSE forecast performance of $\hat{\textrm{GDP}}_{t+shift}(h)$ can be benchmarked against the mean predictor (see Table \ref{rRMSE_mSSA_comp_mean_wc} which reports relative Root Mean-Square Errors, rRMSE) as well as against the direct forecast (see Table \ref{rRMSE_mSSA_comp_direct_without_covid7}).
 For consistency, both benchmark predictors are also quarterly updated out-of-sample designs.\\
@@
 Statistical significance of the GDP component predictor, optimized for larger forecast horizons, can be established for at least up to one year ahead, see Table \ref{rRMSE_mSSA_comp_mean_wc}.
-Finally, the strong performance of predictors optimized for larger forecast horizons ($h\geq 4$), at smaller shifts ($shifts< 4$) corroborates the importance of the left-shift of the corresponding predictors in relation to forecast performance.
+Finally, the strong performance of predictors optimized for larger forecast horizons ($h\geq 4$), at smaller shifts ($shifts< 4$) corroborates the importance of the left-shift of the corresponding predictors, see Figs.\ref{multivar_vs_univar} and \ref{mssa_hpc_financial_crisis}, in relation to forecast performance.
 
@@
 \end{tabular}
-\caption{rRMSEs: out-of-sample forecast errors Equation \ref{oosfe} vs expanding mean of GDP\\Out-of-sample evaluation starting in Q1-2008, without pandemic.\\Values $< 1$ indicate superior predictions from Equation \ref{oos_pred}\\
+\caption{rRMSEs: out-of-sample forecast errors Equation \ref{oosfe} vs expanding mean of GDP\\Out-of-sample evaluation starting in Q1-2008, without pandemic\\Values $< 1$ indicate superior predictions from Equation \ref{oos_pred}\\
 Row-index: Shift refers to the left-shift of the target $\textrm{GDP}_{t+shift}$\\
@@
 For clarity, these results are based on the entire sample, excluding the pandemic period.
-Fig.\ref{ROC_GDP_shift_1_4} compares ROCs of the predictors for small and mid-term forecast horizons and Table \ref{auc} presents AUC (area under the curve) for forecast horizons 1-5 quarters ahead.
+Fig.\ref{ROC_GDP_shift_1_4} compares ROCs of the predictors for small and medium-term forecast horizons and Table \ref{auc} presents AUC (area under the curve) for forecast horizons 1-5 quarters ahead.
 \begin{figure}[h]\begin{center}\includegraphics[height=3in, width=5in]{./Figures/ROC_GDP_shift_1_4.pdf}\caption{Hit rate vs.
@@
   \hline
-Shift=0 & 1.128 & 1.086 & 1.030 & 0.985 & 0.992 & 1.044 & 1.109 \\ 
-  Shift=1 & 1.280 & 1.254 & 1.189 & 1.096 & 1.037 & 1.041 & 1.081 \\ 
-  Shift=2 & 1.173 & 1.181 & 1.166 & 1.104 & 1.031 & 0.995 & 0.995 \\ 
-  Shift=3 & 1.012 & 1.037 & 1.058 & 1.040 & 0.989 & 0.949 & 0.934 \\ 
-  Shift=4 & 0.982 & 1.016 & 1.056 & 1.063 & 1.022 & 0.974 & 0.941 \\ 
-  Shift=5 & 0.961 & 0.997 & 1.041 & 1.058 & 1.028 & 0.979 & 0.936 \\ 
+Shift=0 & 1.131 & 1.092 & 1.028 & 0.973 & 0.977 & 1.040 & 1.119 \\ 
+  Shift=1 & 1.295 & 1.274 & 1.206 & 1.106 & 1.039 & 1.044 & 1.097 \\ 
+  Shift=2 & 1.180 & 1.194 & 1.178 & 1.109 & 1.029 & 0.990 & 0.997 \\ 
+  Shift=3 & 1.029 & 1.061 & 1.082 & 1.051 & 0.984 & 0.936 & 0.923 \\ 
+  Shift=4 & 1.012 & 1.057 & 1.100 & 1.096 & 1.036 & 0.967 & 0.924 \\ 
+  Shift=5 & 0.988 & 1.033 & 1.076 & 1.084 & 1.048 & 0.987 & 0.929 \\  
    \hline
@@
   \hline
-Shift=0 & 0.127 & 0.064 & 0.019 & 0.003 & 0.001 & 0.002 & 0.012 \\ 
-  Shift=1 & 0.439 & 0.314 & 0.194 & 0.093 & 0.033 & 0.017 & 0.036 \\ 
-  Shift=2 & 0.718 & 0.565 & 0.326 & 0.098 & 0.020 & 0.008 & 0.010 \\ 
-  Shift=3 & 0.946 & 0.975 & 0.925 & 0.657 & 0.222 & 0.059 & 0.028 \\ 
-  Shift=4 & 0.327 & 0.699 & 0.964 & 0.942 & 0.619 & 0.237 & 0.102 \\ 
-  Shift=5 & 0.383 & 0.755 & 0.946 & 0.910 & 0.581 & 0.241 & 0.055 \\ 
-   \hline
-\end{tabular}
-\caption{p-values for $H_0: \hat{\beta_1} = 0$, HAC-adjusted Wald test, Equation \ref{eq_gdp_pred}.\\ including the pandemic.} 
+Shift=0 & 0.080 & 0.042 & 0.011 & 0.001 & 0.000 & 0.002 & 0.011 \\ 
+  Shift=1 & 0.403 & 0.297 & 0.188 & 0.095 & 0.042 & 0.039 & 0.103 \\ 
+  Shift=2 & 0.718 & 0.594 & 0.359 & 0.114 & 0.019 & 0.009 & 0.022 \\ 
+  Shift=3 & 0.955 & 0.950 & 0.871 & 0.569 & 0.158 & 0.031 & 0.013 \\ 
+  Shift=4 & 0.593 & 0.879 & 0.957 & 0.893 & 0.541 & 0.183 & 0.080 \\ 
+  Shift=5 & 0.531 & 0.875 & 0.971 & 0.929 & 0.605 & 0.239 & 0.041 \\ 
+   \hline
+\end{tabular}
+\caption{p-values for $H_0: {\alpha_1} = 0$, HAC-adjusted Wald test, Equation \ref{eq_gdp_pred}.\\ including the pandemic.} 
 \label{p_val1}
@@
   \hline
-Shift=0 & 1.005 & 0.986 & 0.960 & 0.935 & 0.927 & 0.939 & 0.964 \\ 
-  Shift=1 & 1.060 & 1.054 & 1.032 & 0.998 & 0.973 & 0.965 & 0.970 \\ 
-  Shift=2 & 1.028 & 1.029 & 1.020 & 0.993 & 0.963 & 0.950 & 0.952 \\ 
-  Shift=3 & 1.018 & 1.027 & 1.034 & 1.022 & 0.994 & 0.972 & 0.961 \\ 
-  Shift=4 & 0.990 & 1.006 & 1.025 & 1.032 & 1.017 & 0.998 & 0.985 \\ 
-  Shift=5 & 1.000 & 1.016 & 1.037 & 1.046 & 1.032 & 1.008 & 0.985 \\ 
+Shift=0 & 0.999 & 0.980 & 0.951 & 0.925 & 0.915 & 0.926 & 0.955 \\ 
+  Shift=1 & 1.057 & 1.051 & 1.028 & 0.996 & 0.977 & 0.978 & 0.985 \\ 
+  Shift=2 & 1.031 & 1.035 & 1.026 & 0.996 & 0.964 & 0.952 & 0.964 \\ 
+  Shift=3 & 1.022 & 1.035 & 1.042 & 1.026 & 0.992 & 0.965 & 0.953 \\ 
+  Shift=4 & 1.002 & 1.023 & 1.044 & 1.046 & 1.024 & 0.999 & 0.984 \\ 
+  Shift=5 & 1.010 & 1.032 & 1.054 & 1.060 & 1.043 & 1.013 & 0.983 \\  
    \hline
@@
   \hline
-Shift=0 & 1.237 & 1.215 & 1.183 & 1.152 & 1.141 & 1.156 & 1.187 \\ 
-  Shift=1 & 0.986 & 0.980 & 0.960 & 0.929 & 0.905 & 0.898 & 0.903 \\ 
-  Shift=2 & 1.036 & 1.037 & 1.029 & 1.001 & 0.971 & 0.957 & 0.960 \\ 
-  Shift=3 & 1.042 & 1.051 & 1.058 & 1.045 & 1.018 & 0.995 & 0.983 \\ 
-  Shift=4 & 0.970 & 0.985 & 1.004 & 1.010 & 0.996 & 0.978 & 0.965 \\ 
-  Shift=5 & 0.970 & 0.986 & 1.007 & 1.015 & 1.002 & 0.978 & 0.956 \\  
+Shift=0 & 1.230 & 1.207 & 1.172 & 1.139 & 1.127 & 1.140 & 1.177 \\ 
+  Shift=1 & 0.984 & 0.977 & 0.956 & 0.927 & 0.909 & 0.910 & 0.916 \\ 
+  Shift=2 & 1.039 & 1.044 & 1.035 & 1.004 & 0.972 & 0.960 & 0.972 \\ 
+  Shift=3 & 1.046 & 1.059 & 1.067 & 1.050 & 1.015 & 0.987 & 0.975 \\ 
+  Shift=4 & 0.982 & 1.002 & 1.023 & 1.024 & 1.003 & 0.978 & 0.964 \\ 
+  Shift=5 & 0.979 & 1.001 & 1.022 & 1.028 & 1.012 & 0.983 & 0.953 \\   
    \hline
@@
   \hline
-rRMSE & 0.833 & 1.083 & 0.981 & 0.979 & 1.027 & 1.020 \\ 
+rRMSE & 0.812 & 1.075 & 0.992 & 0.977 & 1.021 & 1.031 \\ 
    \hline

### 2025-08-08 Achive IWH Simon paper/08.08.2025/poster_reduced.tex
up

--- 2025-06-22 HWI SSA_27_05_2025_with new poster/23.06.2025/poster.tex
@@
 				\vspace{1cm}
-					 {\textbf{\textcolor{iwhdarkblue}Research Questions:}}
+					 {\textbf{\textcolor{iwhdarkblue}{Research Questions:}}}
 
 					\item Can we improve GDP forecast smoothness without sacrificing accuracy?
-					\item Do multivariate filtering techniques enhance persistence and coherence?
-					\\
-					$\rightarrow$ 	We propose methods to produce smooth and persistent forecasts of German GDP, i.e.
-by replacing raw leading indicators with smoothed and filtered components derived from the M-SSA procedure, thereby improving the signal-to-noise ratio in the predictor variables used to forecast GDP in the mid of a particular quarter.
+					\item Do filtering techniques enhance persistence and coherence?
+                    \begin{itemize}
+                    \begin{normalsize}
+                        \item Replace raw leading indicators with smoothed and filtered components 
+                         \item Improve the signal-to-noise ratio. 
+                    \end{normalsize}
+                 \end{itemize}
                     \end{itemize}
@@
 							     \item Monthly Indicators: Industrial Production, ifo Business Climate, ESI, Term spread 10y-3m.
-							     \item Log-differenced, standardized (except spreads), outliers trimmed ( COVID-19).
+							     \item Log-differenced, standardized (except spreads), outliers trimmed (COVID-19).
 							     \item Data aligned to account for publication lags and converted to quarterly frequency 
@@
 
-						\column{0.495\textwidth}
-									\begin{center}
-										{\small{\textbf{\textcolor{black}{Figure 1: Leads \& lags}}}}
+						\column[T]{0.495\textwidth}
+									\begin{center}{\small{\textbf{\textcolor{black}{Figure 1: Leads \& lags}}}}
 
@@
 									\end{center}
-
-								\leftskip2em \singlespacing{ {\footnotesize \emph{Note:} GDP \& Indicators: standardized log-differences, trimmed to $\pm 3$ standard deviations.} }
+\vspace{-1cm}
+								\leftskip2em \singlespacing{ {\footnotesize \emph{Note:} GDP \& Indicators:  GDP (blue), IP (red), ifo (green), ESI (blue) and term-spread (yellow).
+Data standardized log-differences, trimmed to $\pm 3$ standard deviations.} }
 						\end{columns}
@@
 			\begin{itemize}
-				\item Regression of GDP on contemporaneous indicators, i.e. VARMA dynamics estimated via parsimonious VAR(1).
-				\item Impulse responses reveal GDP reacts most strongly to survey shocks, with modest delayed response to IP.
-				\item These dynamics motivate the use of multivariate models to improve forecast accuracy beyond univariate benchmarks.
-				\item The results indicate that as the forecast horizon $h$ increases, the turning points (i.e., peaks and troughs) of the direct forecasts become increasingly right-shifted (lagged) relative to the target GDP.
+				\item Regression of GDP on contemporaneous indicators.
+				\item Lead/lag structure motivates the use of multivariate models to improve forecast accuracy.
+				\item As the forecast horizon $h$ increases, the turning points (i.e., peaks and troughs) of the direct forecasts become increasingly right-shifted (lagging). 
 			\end{itemize}
@@
 			\column{0.495\textwidth}
-
-				\textcolor{iwhdarkblue}{\textbf{Filtering Indicators}}
+				\\
+				\vspace{1cm}					
+				\textcolor{iwhdarkblue}{\textbf{Filtering Indicators: HP}}
 					\begin{itemize}
@@
 								{\small{\textbf{\textcolor{black}{Table 1: }}}}
+                                \\
+				\vspace{0.3cm}
                             \begin{tabular}{lrrrrrrr}
@@
 						\end{tabular}
-						\leftskip2em \singlespacing{ {\footnotesize \emph{Note:} p-values for $H_0: \hat{\beta} = 0$, HAC-adjusted Wald test.\\Estimation over full sample (without pandemic.)					\label{tab:pvaluedhp}}}
+						\leftskip2em \singlespacing{ {\footnotesize \emph{Note:} p-values for $H_0: \hat{\beta} = 0$, HAC-adjusted Wald test.\\Estimation over full sample (without pandemic).					\label{tab:pvaluedhp}}}
 						\end{center}
@@
 
-				\item M-SSA extends SSA (\cite{Wildi2025}) to multivariate setting with causal (one-sided) filters.
+				\item M-SSA extends SSA \citep{Wildi2025} to multivariate setting with causal (one-sided) filters.
 				\item Target variable $z_t$ (e.g. HP-GDP) is tracked by predictor $y_t$ via multivariate filters on leading indicators.
@@
 				\end{columns}	
+            \begin{itemize}
 
+				\item In contrast to HP (top two panels Figure 3), the multivariate filter (lower two panels) can track recession dips as the forecast horizon increases.
+                \item Main explanation for improved forecast performances at larger horizons ($h>2$).
+			\end{itemize}
 %	\autoheight
@@
 
-			\begin{columns}
+			\begin{columns}[T]
 			\column{0.495\textwidth}
@@
 			\item M-SSA reduces sign changes by 33\% with minimal loss of accuracy.
-			\item M-SSA achieves longer holding times: a $50\%$ increase compared to M-MSE
+			\item M-SSA achieves longer holding times: a $50\%$ increase compared to M-MSE.
 			\end{itemize}
@@
 			\includegraphics[width=0.8\textwidth]{./Figures/mssa_msse_zc.pdf}
-				\leftskip2em \singlespacing{ {\footnotesize \emph{Note:}{Two-sided HP(160) applied to GDP (black)\\ Nowcasts: M-SSA (blue) and M-MSE (green).
-\\The dashed vertical line delimits in- and out-of sample periods.
-\\The two-sided filter does not extend to the sample end.
-				\label{mssa_msse_now}}}}
+				\leftskip2em \singlespacing{ {\footnotesize \emph{Note:} Nowcasts: M-SSA (blue) and M-MSE (green).
+\\The vertical green lines indicate sign changes of M-MSE  \\M-SSA (blue) has less zero-crossings (smoother).
+				\label{mssa_msse_now}}}
 		\end{center}
@@
 
-					\vspace{0.5cm}
-					\textcolor{iwhdarkblue}{\textbf{M-SSA Forecasts for $h > 0$}}\\
+%					\vspace{1.5cm}
+%					\textcolor{iwhdarkblue}{\textbf{M-SSA Forecasts for $h > 0$}}\\
+%					
+%						\begin{itemize}	
+%								\addtolength{\itemsep}{10pt}	
+%					\item M-SSA forecasts shift leftward consistently as $h$ increases, especially around turning points.
 
+%						\end{itemize}
+
+			\vspace{1.5cm}
+
+			\textcolor{iwhdarkblue}{\textbf{M-SSA Forecast}}\\
 				\begin{itemize}
-								\addtolength{\itemsep}{10pt}	
-					\item As the forecast horizon hh increases, M-SSA filter weights decrease, indicating greater uncertainty.
-					\item M-SSA forecasts shift leftward consistently as $h$ increases, especially around turning points.
+				\item Out-of-sample forecast performance is evaluated by comparing the M-SSA-based predictor to both a mean forecast and a direct forecast, all updated quarterly.
 
-						\end{itemize}
-
-			\vspace{0.5cm}
-
-			\textcolor{iwhdarkblue}{\textbf{M-SSA Component Predictor}}\\
-				\begin{itemize}
-				\item M-SSA components, derived from HP(160)-filtered indicators, are used as predictors for unfiltered GDP via a post-processing regression step.
-				\item The M-SSA filters are fixed using data up to January 2008, while component weights are updated quarterly using an expanding window starting from January 2008.
-				\item A linear regression maps the M-SSA output at horizon $h$ to forward-shifted GDP at $t + \text{shift}$, correcting for scaling and allowing for $h \neq \text{shift}$.
-				\item This setup enables leveraging the forecast-leading behavior (left-shift) of M-SSA predictors, particularly when $\text{shift} < h$.
 			\end{itemize}
@@
 			\begin{itemize}
-			\item Out-of-sample forecast performance is evaluated by comparing the M-SSA-based predictor to both a mean forecast and a direct forecast, all updated quarterly.
-			\item Results show that for longer horizons ($h \geq 4$), M-SSA predictors outperform both benchmarks, especially at shorter shifts, confirming the predictive value of their anticipatory signal.
+%			\item Out-of-sample forecast performance is evaluated by comparing the M-SSA-based predictor to both a mean forecast and a direct forecast, all updated quarterly.
+			\item Results show that M-SSA predictors outperform both benchmarks at horizons larger than two quarters.
 			\item Statistical significance of the M-SSA GDP predictor is confirmed for up to one year ahead, excluding the pandemic period.
@@
 			{\small{\textbf{\textcolor{black}{Table 2: rRMSEs }}}}\\
+              				\vspace{0.3cm}
 \begin{footnotesize}
@@
 		\hline
-		Shift=0 & 0.978 & 0.942 & 0.893 & 0.855 & 0.860 & 0.906 & 0.962 \\ 
-		Shift=1 & 1.082 & 1.060 & 1.005 & 0.926 & 0.877 & 0.880 & 0.914 \\ 
-		Shift=2 & 1.083 & 1.090 & 1.077 & 1.019 & 0.952 & 0.918 & 0.919 \\ 
-		Shift=3 & 1.034 & 1.059 & 1.081 & 1.063 & 1.010 & 0.970 & 0.954 \\ 
-		Shift=4 & 0.995 & 1.029 & 1.069 & 1.077 & 1.035 & 0.986 & 0.953 \\ 
-		Shift=5 & 0.993 & 1.030 & 1.075 & 1.093 & 1.062 & 1.011 & 0.967 \\ 
+Shift=0 & 0.973 & 0.937 & 0.888 & 0.850 & 0.855 & 0.900 & 0.956 \\ 
+  Shift=1 & 1.073 & 1.051 & 0.997 & 0.918 & 0.870 & 0.873 & 0.906 \\ 
+  Shift=2 & 1.069 & 1.077 & 1.063 & 1.006 & 0.940 & 0.907 & 0.907 \\ 
+  Shift=3 & 1.020 & 1.045 & 1.066 & 1.048 & 0.997 & 0.957 & 0.941 \\ 
+  Shift=4 & 0.983 & 1.016 & 1.056 & 1.064 & 1.023 & 0.974 & 0.942 \\ 
+  Shift=5 & 0.988 & 1.026 & 1.071 & 1.088 & 1.057 & 1.007 & 0.963 \\
 		\hline
@@
 \begin{singlespace}
-	{\footnotesize \emph{Notes:} : Out-of-sample forecast errors  vs expanding mean forecast of GDP.
-Out-of-sample evaluation starting in Q1 2008, without pandemic.\\Values $< 1$ indicate superior predictions.}
+	\leftskip2em {\footnotesize \emph{Notes:}  Out-of-sample rRMSE of M-SSA vs.
+expanding mean forecast of GDP.
+Out-of-sample evaluation starting in Q1 2008, without pandemic.\\Values $< 1$ indicate superior predictions by M-SSA.}
 \end{singlespace}
@@
 \end{table}
+
+\begin{table}[ht]
+	\centering
+			{\small{\textbf{\textcolor{black}{Table 3: statistical significance}}}}\\
+            \vspace{0.3cm}
+\begin{footnotesize}
+\begin{tabular}{rrrrrrrr}
+  \hline
+ & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
+  \hline
+Shift=0 & 0.084 & 0.047 & 0.017 & 0.005 & 0.002 & 0.001 & 0.038 \\ 
+  Shift=1 & 0.230 & 0.160 & 0.094 & 0.032 & 0.005 & 0.001 & 0.004 \\ 
+  Shift=2 & 0.801 & 0.666 & 0.414 & 0.173 & 0.026 & 0.005 & 0.006 \\ 
+  Shift=3 & 0.876 & 0.961 & 0.917 & 0.683 & 0.248 & 0.068 & 0.039 \\ 
+  Shift=4 & 0.370 & 0.819 & 0.990 & 0.953 & 0.533 & 0.117 & 0.030 \\ 
+  Shift=5 & 0.353 & 0.802 & 0.983 & 0.943 & 0.571 & 0.239 & 0.043 \\ 
+   \hline
+\end{tabular}
+\end{footnotesize}
+\begin{singlespace}
+	\leftskip2em {\footnotesize \emph{Notes:} p-values for $H_0: \hat{\beta_1} = 0$, HAC-adjusted Wald test,\\sample excludes pandemic.}
+\end{singlespace}
+\vspace{0.5cm}
+\end{table}
+
 \end{columns}
 
-\vspace{0.5cm}
+\vspace{1.5cm}
 \textcolor{iwhdarkblue}{\textbf{Receiver Operating Characteristic (ROC)}}\\
-
  \begin{columns}
@@
       \begin{itemize}
-			\item Compute sign accuracy and false positive rate: ROC curves
-            \item Compute Area Under Curve (AUC)
-			\item All filters outperform direct forecast (larger AUC) for $h>2$ 
+			\item Compute hit rate and false alarm rate: ROC curve
+            \item Performance measure: Area Under Curve (AUC)
 			\end{itemize}
@@
 	\centering
-			{\small{\textbf{\textcolor{black}{Table 3: AUCs }}}}\\
+			{\small{\textbf{\textcolor{black}{Table 4: AUCs }}}}\\
+            \vspace{0.3cm}
 \begin{footnotesize}
@@
 \begin{singlespace}
-	{\footnotesize \emph{Notes:} :AUCs of predictors against GDP at various forecast horizons.}
+	{\footnotesize \emph{Notes:} AUCs of predictors against GDP at various forecast horizons.}
 \end{singlespace}
@@
 \end{table}
+            \begin{itemize}
+			\item All filters outperform direct forecast (larger AUC) for $h>2$. 
+            \item For $h>2$, M-SSA maximizes AUC, followed by M-MSE, HP-C and the direct forecast.
+			\end{itemize}
 
@@
 			\begin{itemize}
-            \item Classic direct forecasts of GDP perform well up to two quarters ahead but suffer from noise and limited horizon; smoothing with an HP(160)-filter improves performance during sharp recessions yet retains some noise-induced sign changes.
+            \item Classic direct forecasts of GDP outperform the mean benchmark up to two quarters ahead.  
+            \item Smoothing with an HP(160)-filter improves performances at larger forecast horizons. But the filter cannot leverage leading indicators and remains noisy.
             \item To overcome these limitations, a multivariate SSA (M-SSA) framework based on \cite{Wildi2024} is introduced, leveraging cross-sectional indicator data and controlling the predictor’s zero-crossing rate for enhanced signal tracking at turning points.
@@
 
-	\begin{block}{\large References and Further Information}
-	\begin{columns}[T]
-		\column{0.3\textwidth}
+	 \begin{block}{\large References}
+	\bibliographystyle{apalike}
+    \begin{small}
+        \bibliography{references}
 
-		\column{0.5\textwidth}
-				%\begin{flushright}	
-                GIT: https://github.com/wiaidp/R-package-SSA-Predictor
+        M-SSA package: https://github.com/wiaidp/R-package-SSA-Predictor    
+        \end{small}
+
 				%\end{flushright}
-				\column{0.2\textwidth}
-
-			%\includegraphics[width = 0.55\textwidth]{../Figures/GIT.pdf}
-		\end{columns}
-
     \end{block}
 		\end{column}
-	\bibliographystyle{apalike}
-	\bibliography{references}
+
 \end{columns}

### 2025-08-08 Archive IWH Simon paper/05.06.2025/poster_reduced.tex
up

--- 2025-08-08 Achive IWH Simon paper/08.08.2025/poster_reduced.tex
@@
 				\vspace{1cm}
-					 {\textbf{\textcolor{iwhdarkblue}{Research Questions:}}}
+					 {\textbf{\textcolor{iwhdarkblue}Research Questions:}}
 
@@
                     \begin{itemize}
-                    \begin{normalsize}
                     \item Replace raw leading indicators with smoothed and filtered components 
                     \item Improve the signal-to-noise ratio.
-                    \end{normalsize}
 					\end{itemize}
@@
 							     \item Monthly Indicators: Industrial Production, ifo Business Climate, ESI, Term spread 10y-3m.
-							     \item Log-differenced, standardized (except spreads), outliers trimmed (COVID-19).
+							     \item Log-differenced, standardized (except spreads), outliers trimmed ( COVID-19).
 							     \item Data aligned to account for publication lags and converted to quarterly frequency 
@@
 
-						\column[T]{0.495\textwidth}
-									\begin{center}{\small{\textbf{\textcolor{black}{Figure 1: Leads \& lags}}}}
+						\column{0.495\textwidth}
+									\begin{center}
+										{\small{\textbf{\textcolor{black}{Figure 1: Leads \& lags}}}}
 
@@
 									\end{center}
-\vspace{-1cm}
-								\leftskip2em \singlespacing{ {\footnotesize \emph{Note:} GDP \& Indicators:  GDP (blue), IP (red), ifo (green), ESI (blue) and term-spread (yellow).
-Data standardized log-differences, trimmed to $\pm 3$ standard deviations.} }
+
+								\leftskip2em \singlespacing{ {\footnotesize \emph{Note:} GDP \& Indicators: standardized log-differences, trimmed to $\pm 3$ standard deviations.} }
 						\end{columns}
@@
 			\column{0.495\textwidth}
-				\\
-				\vspace{1cm}					
+
 				\textcolor{iwhdarkblue}{\textbf{Filtering Indicators: HP}}
@@
 								{\small{\textbf{\textcolor{black}{Table 1: }}}}
-                                \\
-				\vspace{0.3cm}
 						\begin{tabular}{lrrrrrrr}
@@
 						\end{tabular}
-						\leftskip2em \singlespacing{ {\footnotesize \emph{Note:} p-values for $H_0: \hat{\beta} = 0$, HAC-adjusted Wald test.\\Estimation over full sample (without pandemic).					\label{tab:pvaluedhp}}}
+						\leftskip2em \singlespacing{ {\footnotesize \emph{Note:} p-values for $H_0: \hat{\beta} = 0$, HAC-adjusted Wald test.\\Estimation over full sample (without pandemic.)					\label{tab:pvaluedhp}}}
 						\end{center}
@@
 
-				\item M-SSA extends SSA \citep{Wildi2025} to multivariate setting with causal (one-sided) filters.
+				\item M-SSA extends SSA (\cite{Wildi2025}) to multivariate setting with causal (one-sided) filters.
 				\item Target variable $z_t$ (e.g. HP-GDP) is tracked by predictor $y_t$ via multivariate filters on leading indicators.
@@
 				\end{columns}
-            \begin{itemize}
 
-				\item In contrast to HP (top two panels Figure 3), the multivariate filter (lower two panels) can track recession dips as the forecast horizon increases.
-                \item Main explanation for improved forecast performances at larger horizons ($h>2$).
-			\end{itemize}
 %	\autoheight
@@
 
-			\begin{columns}[T]
+			\begin{columns}
 			\column{0.495\textwidth}
@@
 			\item M-SSA reduces sign changes by 33\% with minimal loss of accuracy.
-			\item M-SSA achieves longer holding times: a $50\%$ increase compared to M-MSE.
+			\item M-SSA achieves longer holding times: a $50\%$ increase compared to M-MSE
 			\end{itemize}
@@
 			\includegraphics[width=0.8\textwidth]{./Figures/mssa_msse_zc.pdf}
-				\leftskip2em \singlespacing{ {\footnotesize \emph{Note:} Nowcasts: M-SSA (blue) and M-MSE (green).
+				\leftskip2em \singlespacing{ {\footnotesize \emph{Note:}{Financial Crisis\\Nowcasts: M-SSA (blue) and M-MSE (green).
 \\The vertical green lines indicate sign changes of M-MSE  \\M-SSA (blue) has less zero-crossings (smoother).
-				\label{mssa_msse_now}}}
+				\label{mssa_msse_now}}}}
 		\end{center}
@@
 
-			\textcolor{iwhdarkblue}{\textbf{M-SSA Forecast}}\\
+			\textcolor{iwhdarkblue}{\textbf{M-SSA Component Predictor}}\\
 				\begin{itemize}
-				\item Out-of-sample forecast performance is evaluated by comparing the M-SSA-based predictor to both a mean forecast and a direct forecast, all updated quarterly.
+				\item A linear regression maps the M-SSA output at horizon $h$ to forward-shifted GDP at $t + \text{shift}$, correcting for scaling and allowing for $h \neq \text{shift}$.
 
@@
 			\begin{itemize}
-%			\item Out-of-sample forecast performance is evaluated by comparing the M-SSA-based predictor to both a mean forecast and a direct forecast, all updated quarterly.
+			\item Out-of-sample forecast performance is evaluated by comparing the M-SSA-based predictor to both a mean forecast and a direct forecast, all updated quarterly.
 			\item Results show that M-SSA predictors outperform both benchmarks at horizons larger than two quarters.
@@
 			{\small{\textbf{\textcolor{black}{Table 2: rRMSEs }}}}\\
-              				\vspace{0.3cm}
 \begin{footnotesize}
@@
 \begin{singlespace}
-	\leftskip2em {\footnotesize \emph{Notes:}  Out-of-sample rRMSE of M-SSA vs.
+	{\footnotesize \emph{Notes:}  Out-of-sample rRMSE of M-SSA vs.
 expanding mean forecast of GDP.
@@
 			{\small{\textbf{\textcolor{black}{Table 3: statistical significance}}}}\\
-            \vspace{0.3cm}
 \begin{footnotesize}
@@
 \begin{singlespace}
-	\leftskip2em {\footnotesize \emph{Notes:} p-values for $H_0: \hat{\beta_1} = 0$, HAC-adjusted Wald test,\\sample excludes pandemic.}
+	{\footnotesize \emph{Notes:} p-values for $H_0: \hat{\beta_1} = 0$, HAC-adjusted Wald test, Equation \ref{eq_gdp_pred}.\\Sample excludes pandemic.}
 \end{singlespace}
@@
 			{\small{\textbf{\textcolor{black}{Table 4: AUCs }}}}\\
-            \vspace{0.3cm}
 \begin{footnotesize}
@@
             \begin{itemize}
-			\item All filters outperform direct forecast (larger AUC) for $h>2$. 
-            \item For $h>2$, M-SSA maximizes AUC, followed by M-MSE, HP-C and the direct forecast.
+			\item All filters outperform direct forecast (larger AUC) for $h>2$ 
+            \item For $h>2$, M-SSA maximizes AUC, followed by M-MSE, HP-C and the direct forecast
 			\end{itemize}

### 2025-08-08 Archive IWH Simon paper/08.08.2025/paper.tex
up

--- 2025-08-08 Achive IWH Simon paper/08.08.2025/paper.tex
@@
     \begin{center}
-    \textcolor{blue}{\textbf{Notice that the title on top right panel is incorrect!}}
+%    \textcolor{blue}{\textbf{Notice that the title on top right panel is incorrect!}}
         \includegraphics[height=4in, width=5.5in]{./Figures/hp_160.pdf}
@@
 
-\subsection{SSA Optimization criterion}
+\subsection{MSSA Optimization Criterion}
 
@@
 In this framework, $\mathbf{X}_t$ corresponds to a matrix of selected economic indicators with $n=5$, and $z_{it}$, i=1,...,5,   is the output of a two-sided HP(160) filter applied to the four indicators as well as to GDP.
-Since $z_{it}$  depends on future observations $x_{it+1},x_{it+2},...$, the prediction task involves `tracking' $z_{it}$  using the estimate $y^i_{t}=\sum_{j=1}^ny^i_{jt}$, where $y^i_{jt}=\mathbf{b}^i_j~'\mathbf{x}_{jt}=\sum_{k=0}^{L} b^i_{jk}x_{j,t-k}$ are the outputs of a multivariate (causal or one-sided) filter $\mathbf{B}^i$, with columns $\mathbf{b}^i_{j}$ of fixed length $L$, $j=1,...,5$, assigning weights to the last $L$ observations of the $j$-th indicator $\mathbf{x}_{jt}$ (the fixed-length assumption is merely used for ease of exposition).
+Since $z_{it}$  depends on future observations $x_{it+1},x_{it+2},...$, the prediction task involves `tracking' $z_{it}$  using the estimate $y_{it}=\sum_{j=1}^ny_{ijt}$, where $y_{ijt}=\mathbf{b}_{ij}~'\mathbf{x}_{jt}=\sum_{k=0}^{L} b_{ijk}x_{j,t-k}$ are the outputs of a multivariate (causal or one-sided) filter $\mathbf{B}_i^f$, with columns $\mathbf{b}_{ij}$ of fixed length $L$, $j=1,...,5$, assigning weights to the last $L$ observations of the $j$-th indicator $\mathbf{x}_{jt}$ (the fixed-length assumption is merely used for ease of exposition).
 For the sake of clarity, we now omit the index $i$ for all variables, assuming $z_t=z_{i_0t}$ for some fixed $i_0$.% Additionally, we assume a fixed filter length $L$, such that $\mathbf{B}^i_t=\mathbf{B}_t=\mathbf{B}$, which has dimension $L\times 5$.
@@
  \cite{Wildi2025} shows that the SSA criterion helps control this phenomenon (see Fig.\ref{mssa_msse_zc} for illustration.) Using the dual formulation of the criterion \eqref{critssaht}, the author demonstrates that the SSA solution yields the lowest zero-crossing rate among all (linear) predictors with equivalent target correlation.
-
-The concept can be extended to a multivariate framework, denoted M-SSA, by generalizing the correlation functions as detailed in \cite{Wildi2025}.
-
-\subsection{M-SSA Optimization}
-
-\textcolor{blue}{This is where Marc outlines the generalization of the correlation function and other elements of M-SSA.}
+The concept can be extended to a multivariate framework, denoted M-SSA, by generalizing the correlation functions as detailed hereafter.
+
+\subsection{Implementation}
+
+For a better understanding, we here briefly substantiate the optimization criterion \ref{critssa}.
+To start, assume a simple univariate framework with $x_t=\epsilon_t$ white noise and $z_t=\sum_{k=-infty}^\infty \gamma_k x_{t-k}=\sum_{k=-infty}^\infty \gamma_k \epsilon_{t-k}$, where it is assumed that $\gamma_k\neq 0$ for some $k<0$ (acausal filter) and $\sum_{|k|<\infty}\gamma_k^2<\infty$, which ensures stationarity of $z_t$ (extensions to non-stationary integrated processes are discussed in Wildi (2025)).
+Assume we want to predict $z_{t+\delta}$, $\delta\geq 0$ and let $y_{\delta t}=\sum_{k=0}^{L-1} b_{\delta k}x_{t-k}$ designate a causal predictor of length $L$.
+For simplicity of exposition we may assume $\delta=0$ (nowcast) and we may drop $\delta$ from our notation.
+Then, $\rho(z,y)=\frac{\sum_{k=0}^{L-1}b_k\gamma_k}{\sqrt{\sum_{k=0}^{L-1}b_k^2}\sqrt{\sum_{k=0}^{L-1}\gamma_k^2}}$  and $\rho(y,1)=\frac{\mathbf{b}'\mathbf{Mb}}{\mathbf{b}'\mathbf{b}}$, where
+\[
+\mathbf{M}=\left(\begin{array}{ccccccccc}0&0.5&0&0&0&...&0&0&0\\
+0.5&0&0.5&0&0&...&0&0&0\\
+...&&&&&&&&\\
+0&0&0&0&0&...&0.5&0&0.5\\
+0&0&0&0&0&...&0&0.5&0
+\end{array}\right),
+\]
+ofs dimension $L\times L$ is such that $\mathbf{b}'\mathbf{Mb}=\sum_{k=1}^{L-1}b_{k-1}b_k$ corresponds to the order-one autocovariance of $y_t$, up to scaling by the variance $\sigma^2$ of $\epsilon_t$.
+We can now rewrite criterion \ref{critssa} as:
+\begin{eqnarray}\label{crit1}
+\left.\begin{array}{cc}
+&\max_{\mathbf{b}}\mathbf{b}'\boldsymbol{\gamma}_{\delta}\\
+&\mathbf{b}'\mathbf{Mb}=l\rho_1\\
+&\mathbf{b}'\mathbf{b}=l
+\end{array}\right\}.
+\end{eqnarray}
+where the additional length constraint $\mathbf{b}'\mathbf{b}=l$ ensures uniqueness of the SSA solution. 
+By orthogonal projection, the MSE predictor is $y_{MSE,t}=\sum_{k=0}^{L-1}\gamma_kx_{t-k}$, i.e., $b_{MSE,k}=\gamma_k$, $k=0,...,L-1$ and in general we select $\rho_1>\rho_{MSE}$ in criterion l{crit1} to improve smoothness of the resulting SSA predictor $y_t$, where $\rho_{MSE}$ designates the first order ACF of the (benchmark) MSE predictor\footnote{Selecting $\rho_1<\rho_{MSE}$ would result in a SSA predictor with more frequent zero-crossings than the MSE benchmark, which is undesirable in typical applications.}.
+\\
+
+We can extend the SSA criterion \ref{crit1} to a multivariate framework.
+Assume first that $\mathbf{X}_t=\boldsymbol{\epsilon}_t$ is a multivariate white noise sequence with (full rank) variance-covariance matrix $\boldsymbol{\Sigma}$ and let $\mathbf{z}_t$:
+\[
+\mathbf{z}_t=\sum_{|k|<\infty}\boldsymbol{\Gamma}_k\mathbf{x}_{t-k},
+\] 
+where $\boldsymbol{\Gamma}_k$ is a sequence of square-summable $n\times n$ dimensional filter matrices, with entries $(\gamma_{ijk})$ for $i, j = 1, \ldots, n$.
+The condition for square summability is expressed mathematically as:
+\[
+\sum_{|k| < \infty} \text{tr}(\boldsymbol{\Gamma}_k' \boldsymbol{\Gamma}_k) < \infty,
+\]
+where $\text{tr}(\cdot)$ denotes the trace operator. We focus on the estimation of $\mathbf{z}_{t+\delta}$, utilizing the $n-$dimensional predictor defined by
+\[
+\mathbf{y}_t = \sum_{k=0}^{L-1} \mathbf{B}_k \mathbf{x}_{t-k},
+\]
+with $n\times n$-dimensional filter matrices $\mathbf{B}_k$, $k=0,...,L-1$.
+Let $b_{ijk}$ denote the entries of the matrix $\mathbf{B}_k$ for $k = 0, \ldots, L-1$.
+We define several vector notations as follows:
+\begin{eqnarray*}
+\boldsymbol{\epsilon}_{it}=(\epsilon_{it},\epsilon_{it-1},...,\epsilon_{it-(L-1)})'~&,&~\boldsymbol{\epsilon}_{\cdot t}=(\boldsymbol{\epsilon}_{1t}',...,\boldsymbol{\epsilon}_{nt}')'\\
+\boldsymbol{\gamma}_{ij\delta}=(\gamma_{ij\delta},\gamma_{ij\delta+1},...,\gamma_{ij\delta+L-1})'~&,&\boldsymbol{\gamma}_{i\cdot\delta}=(\boldsymbol{\gamma}_{i1\delta}',\boldsymbol{\gamma}_{i2\delta}',...,\boldsymbol{\gamma}_{in\delta}')'\\
+~\mathbf{b}_{ij}=(b_{ij0},b_{ij1},...,b_{ijL-1})'~&,&~\mathbf{b}_{i}=(\mathbf{b}_{i1}',\mathbf{b}_{i2}',...,\mathbf{b}_{in}')'\\
+\boldsymbol{\gamma}_{\delta}=\left(\begin{array}{c}\boldsymbol{\gamma}_{1\cdot\delta}'\\
+\boldsymbol{\gamma}_{2\cdot\delta}'\\
+\cdot\\
+\boldsymbol{\gamma}_{n\cdot\delta}'\end{array}\right)~&,&~\mathbf{b}=\left(\begin{array}{c}\mathbf{b}_{1}'\\\mathbf{b}_{2}'\\\cdot\\\mathbf{b}_{n}'\end{array}\right).
+\end{eqnarray*} 
+We define $y_{ijt} := \mathbf{b}_{ij}'\boldsymbol{\epsilon}_{jt}$, which allows us to express $y_{it} = \mathbf{b}_i'\boldsymbol{\epsilon}_{\cdot t} = \sum_{j=1}^n y_{ijt}$ and $\mathbf{y}_t = \mathbf{b}'\boldsymbol{\epsilon}_{\cdot t}$.
+Through orthogonal projection, the MSE predictor is derived as $\hat{\mathbf{y}}_{t,MSE} = \boldsymbol{\gamma}_{\delta}'\boldsymbol{\epsilon}_{\cdot t}$, with components $\hat{y}_{it,MSE}$, $i=1,...,n$.
+Consider the Kronecker products defined as $\tilde{\mathbf{I}}:=\boldsymbol{\Sigma}\otimes\mathbf{I}_{L\times L}$ and $\tilde{\mathbf{M}}:=\boldsymbol{\Sigma}\otimes\mathbf{M}$, of the $n\times n$ dimensional matrix $\boldsymbol{\Sigma}$, the $L\times L$ dimensional identity  $\mathbf{I}_{L\times L}$ and the autocovariance generating matrix $\mathbf{M}$.
+We can express the following expectations:
+\begin{eqnarray}
+E[z_{it+\delta}y_{it}]&=&E[\hat{y}_{it,MSE}y_{it}]=\boldsymbol{\gamma}_{i\cdot\delta}'\tilde{\mathbf{I}}\mathbf{b}_{i}\nonumber\\%\label{moment1}\\
+E[\hat{z}_{it\delta}^2]&=&\boldsymbol{\gamma}_{i\cdot\delta}'\tilde{\mathbf{I}}\boldsymbol{\gamma}_{i\cdot\delta}\nonumber\\%\label{moment2}\\
+E[y_{it}^2]&=&\mathbf{b}_{i}'\tilde{\mathbf{I}}\mathbf{b}_{i}\nonumber\\%\label{moment3}\\
+E[y_{it-1}y_{it}]&=&\mathbf{b}_{i}'\tilde{\mathbf{M}}\mathbf{b}_{i}.\nonumber
+\end{eqnarray}
+The variance of $z_{it}$ is given by 
+\[
+E[z_{it}^2]=\sum_{|k|<\infty}\boldsymbol{\gamma}_{i k}'\boldsymbol{\Sigma}\boldsymbol{\gamma}_{i k},
+\]
+where $\boldsymbol{\gamma}_{i k}:=(\gamma_{i1k},...,\gamma_{ink})'$. Consequntly, we propose the following multivariate M-SSA criterion for $i=1,...,n$: 
+\begin{eqnarray}\label{mcrit1}
+\textrm{max}_{\mathbf{b}_i}\boldsymbol{\gamma}_{i\cdot\delta}'\tilde{\mathbf{I}}\mathbf{b}_{i}\\
+\mathbf{b}_{i}'\tilde{\mathbf{M}}\mathbf{b}_{i}=\rho_i\nonumber\\
+\mathbf{b}_{i}'\tilde{\mathbf{I}}\mathbf{b}_{i}=1,\nonumber
+\end{eqnarray}
+for $i=1,...,n$. 
+Under the imposed length constraint, the objective function is proportional to the correlation between the predictor and the $i$-th target, or equivalently, the $i$-th MSE predictor.
+Additionally, the HT constraint regulates the first-order ACF or, equivalently, the HT of the predictor.
+The relationships established between sign accuracy and target correlation, as well as between first-order ACF and HT, as discussed in the previous section, remain valid.
+Thus, the M-SSA framework can be reformulated in accordance with any alternative and equivalent expressions of the optimization criterion, as reviewed in the previous section.
+\\
+
+To conclude, we now relax the WN hypothesis and assume a stationary process 
+\begin{equation}\label{x_m_st}
+\mathbf{X}_t = \sum_{k=0}^\infty \boldsymbol{\Xi}_k \boldsymbol{\epsilon}_{t-k},
+\end{equation}
+where $\boldsymbol{\Xi}_0 = \mathbf{I}$. Then, the target and predictor can be expressed formally as:
+\begin{eqnarray*}
+\mathbf{z}_t&=&\sum_{|k|<\infty}(\boldsymbol{\Gamma}\cdot\boldsymbol{\Xi})_k \boldsymbol{\epsilon}_{t-k}\\
+\mathbf{y}_t&=&\sum_{j\geq 0} (\mathbf{B}\cdot\boldsymbol{\Xi})_j\boldsymbol{\epsilon}_{t-j}.
+\end{eqnarray*} 
+Here, $(\boldsymbol{\Gamma}\cdot\boldsymbol{\Xi})_k =\sum_{m\leq k}\boldsymbol{\Gamma}_m\boldsymbol{\Xi}_{k-m}$ and $(\mathbf{B}\cdot\boldsymbol{\Xi})_j=\sum_{m=0}^{\min(L-1,j)}\mathbf{B}_m \boldsymbol{\Xi}_{j-m} $ represent the convolutions of the sequences $\boldsymbol{\Gamma}_k$ and $\mathbf{B}_j$ with the Wold-decomposition $\boldsymbol{\Xi}_m$ of $\mathbf{x}_t$.
+For  given $(\mathbf{B}\cdot\boldsymbol{\Xi})_j$, $j=0,...,L-1$, the original coefficients $\mathbf{B}_{k}$ can be derived through deconvolution:
+\begin{eqnarray}\label{con_inv}
+\mathbf{B}_{k}&=&\left\{\begin{array}{cc}(\mathbf{B}\cdot\boldsymbol{\Xi})_{0}\boldsymbol{\Xi}_0^{-1}&,k=0\\
+\left(  (\mathbf{B}\cdot\boldsymbol{\Xi})_{k}-\sum_{m=0}^{k-1}\mathbf{B}_m \boldsymbol{\Xi}_{j-m}\right)\boldsymbol{\Xi}_0^{-1}&,k=1,...,L-1
+\end{array}\right.
+\end{eqnarray}
+In our parametrization, $\boldsymbol{\Xi}_0^{-1} = \mathbf{I}$ simplifies this expression. Denote the elements of the matrices as follows:
+\[
+(\boldsymbol{\Gamma}\cdot\boldsymbol{\Xi})_{lmk}
+ \quad \text{and} \quad (\mathbf{B}\cdot\boldsymbol{\Xi})_{lmj}.
+ \] 
+We now assume $\delta\geq 0$\footnote{Similar but slightly more complex expressions would arise for $\delta < 0$ (backcasting), which are omitted here.} and $L$ to be sufficiently large to render $(\boldsymbol{\Gamma} \cdot \boldsymbol{\Xi})_{\delta+k} \approx 0$ and $(\mathbf{B} \cdot \boldsymbol{\Xi})_j \approx 0$ for $k,j \geq L$, so that  their contributions to the variances of $\hat{\mathbf{y}}_{t,MSE}$ and $\mathbf{y}_t$ are negligible.
+We define:
+\begin{eqnarray*}
+(\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{ij\delta}&=&\left((\boldsymbol{\Gamma}\cdot\boldsymbol{\Xi})_{ij\delta},(\boldsymbol{\Gamma}\cdot\boldsymbol{\Xi})_{ ij\delta+1}...,(\boldsymbol{\Gamma}\cdot\boldsymbol{\Xi})_{ij\delta+L-1}\right)'\\
+(\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{i\delta}&=&\Big((\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{i1\delta}',(\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{i2\delta}',...,(\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{in\delta}'\Big)'\\
+(\mathbf{b}\cdot\boldsymbol{\xi})_{ij}&=&\left((\mathbf{B}\cdot\boldsymbol{\Xi})_{ ij0},(\mathbf{B}\cdot\boldsymbol{\Xi})_{ij1},...,(\mathbf{B}\cdot\boldsymbol{\Xi})_{ijL-1}\right)'\\
+(\mathbf{b}\cdot\boldsymbol{\xi})_{i}&=&\Big((\mathbf{b}\cdot\boldsymbol{\xi})_{i1}',(\mathbf{b}\cdot\boldsymbol{\xi})_{i2}',...,(\mathbf{b}\cdot\boldsymbol{\xi})_{in}'\Big)'
+\end{eqnarray*}
+It is important to note that if $\boldsymbol{\Xi}_k=\mathbf{0},~k>0$, resulting in $\mathbf{X}_t=\boldsymbol{\epsilon}_t$, then we have $(\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{i\delta}=\boldsymbol{\gamma}_{i\cdot\delta}$ and $(\mathbf{b}\cdot\boldsymbol{\xi})_{i}=\mathbf{b}_i$, as defined in the preceding section.
+The generalized M-SSA criterion can then be expressed as:
+\begin{eqnarray}\label{gen_stat_x}
+%\max_{(\mathbf{b}\cdot\boldsymbol{\xi})_i}\frac{ (\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{i\delta}'\mathbf{\tilde{I}} (\mathbf{b}\cdot\boldsymbol{\xi})_i}{\sqrt{(\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{i\delta}'\mathbf{\tilde{I}}(\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{i\delta}}\sqrt{(\mathbf{b}\cdot\boldsymbol{\xi})_i'\mathbf{\tilde{I}}(\mathbf{b}\cdot\boldsymbol{\xi})_i}}
+\max_{(\mathbf{b}\cdot\boldsymbol{\xi})_i} (\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{i\delta}'\mathbf{\tilde{I}} (\mathbf{b}\cdot\boldsymbol{\xi})_i\\
+(\mathbf{b}\cdot\boldsymbol{\xi})_i'\mathbf{\tilde{M}}(\mathbf{b}\cdot\boldsymbol{\xi})_i=\rho_i\nonumber\\
+(\mathbf{b}\cdot\boldsymbol{\xi})_i'\mathbf{\tilde{I}}(\mathbf{b}\cdot\boldsymbol{\xi})_i=1\nonumber
+\end{eqnarray}
+for $i=1,...,n$.
+This formulation can be solved for $(\mathbf{b}\cdot\boldsymbol{\xi})_i$, as detailed in Theorem \eqref{lambda_mult} in the next section.
+The M-SSA coefficients $\mathbf{b}_i$ can subsequently be determined through deconvolution, as indicated in Equation \eqref{con_inv}.
+Wildi (2025) posits an extension of SSA to non-stationary integrated processes, wherein the HT constraint pertains to stationary differences of the predictor.
+While a corresponding extension to M-SSA is feasible based on findings by McElroy and Wildi (2020), the present discussion is confined to stationary processes, omitting a lengthier treatment of rank-defficient designs (cointegration) in the non-stationary case.\\
+
+%\textcolor{blue}{This is where Marc outlines the generalization of the correlation function and other elements of M-SSA.}
 
@@
   \hline
-Shift=0 & 0.975 & 0.941^{*} & 0.886^{*} & 0.839^{**} & 0.843^{**} & 0.897^{**} & 0.965 \\ 
-  Shift=1 & 1.085 & 1.068 & 1.011 & 0.927^{*} & 0.871^{**} & 0.875^{**} & 0.919^{**} \\ 
-  Shift=2 & 1.076 & 1.089 & 1.074 & 1.011 & 0.938^{*} & 0.902^{**} & 0.909^{**} \\ 
-  Shift=3 & 1.037 & 1.069 & 1.090 & 1.059 & 0.992 & 0.943^* & 0.930^* \\ 
-  Shift=4 & 1.013 & 1.058 & 1.101 & 1.097 & 1.037 & 0.967 & 0.925^* \\ 
-  Shift=5 & 1.016 & 1.062 & 1.106 & 1.115 & 1.078 & 1.015 & 0.955^* \\  
+Shift=0 & 0.975 & $0.941^*$ & $0.886^{*}$ &$ 0.839^{**}$ & $0.843^{**}$ & $0.897^{**}$ & 0.965 \\ 
+  Shift=1 & 1.085 & 1.068 & 1.011 &$ 0.927^{*}$ &$ 0.871^{**}$ & $0.875^{**}$ & $0.919^{**}$ \\ 
+  Shift=2 & 1.076 & 1.089 & 1.074 & 1.011 & $0.938^{*}$ & $0.902^{**}$ & $0.909^{**}$ \\ 
+  Shift=3 & 1.037 & 1.069 & 1.090 & 1.059 & 0.992 & $0.943^*$ & $0.930^*$ \\ 
+  Shift=4 & 1.013 & 1.058 & 1.101 & 1.097 & 1.037 & 0.967 & $0.925^*$ \\ 
+  Shift=5 & 1.016 & 1.062 & 1.106 & 1.115 & 1.078 & 1.015 & $0.955^*$ \\  
    \hline

### 2025-08-08 Archive IWH Simon paper/23.06.2025/Slides_13Feb25/gdp/HWI_gdp.tex
up

\documentclass{beamer}
%\documentclass[notes]{beamer}       % print frame + notes
%\documentclass[notes=only]{beamer}   % only notes
%\documentclass{beamer}              % only frames

\usepackage[latin1]{inputenc}
\usetheme{Warsaw}
\setbeamertemplate{headline}{%
\leavevmode%
  \hbox{%
    \begin{beamercolorbox}[wd=\paperwidth,ht=2.5ex,dp=1.125ex]{palette quaternary}%
    \insertsectionnavigationhorizontal{\paperwidth}{}{\hskip0pt plus1filll}
    \end{beamercolorbox}%
  }
}

\title{A Comparative study of Different Predictor Concepts on a HWI Data Set}
\author{Marc Wildi}
\institute{Institute of data analysis and process design (ZHAW)}
\date{May, 2024}
\date[\today]{}

\usepackage{Sweave}
\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}[allowframebreaks]{Table of Contents}
\tableofcontents
\end{frame}

\section{Introduction}

\begin{frame} {Introduction}
\begin{itemize}
\item See slides on ip (ip-case) for a similar introduction: motivation of trends vs. direct forecasts and theses
\end{itemize}
\end{frame}

\begin{frame} {Zero-crossings BIP: Standardized Series}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{zero_cross_ifo_l_1.pdf}\caption{\label{data}}\end{center}\end{figure}\end{frame}

\begin{frame} {Overview of (and Links to) Performance Metrics in this Study}
\begin{itemize}
\item Relative \textbf{forecast mean-square} error (target: BIP-shifted): slide \eqref{rmse1} (direct AR forecasts full sample), \eqref{rmse2} (same but shorter sample), \eqref{rmse3} (univariate filters, full sample), \eqref{rmse4} (same but before Pandemic), \eqref{slide_multi_rel_mse} (multivariate filters)

\item Relative \textbf{filter mean-square} error (target: BIP-trend): \eqref{rfmse1} (outperformance of multivariate over univariate)

\item Target \textbf{correlations}: \eqref{sample_target_cor1}, \eqref{sample_target_cor2}, \eqref{sample_target_cor3} and \eqref{sample_target_cor4}

\item \textbf{Sign accuracies}: \eqref{signaac1} and \eqref{signaac2}

\item \textbf{t-statistics}: \eqref{tstat1},  (univariate filters and direct AR, full sample), \eqref{tstat2} (same but prior pandemic), \eqref{slide_multi_rel_mse} (multivariate filters)

\item \textbf{Holding times}: \eqref{htzc1} (direct AR and univariate filters), \eqref{htzc2} and \eqref{htzc3} (M-SSA vs. M-MSE)
\end{itemize}
\end{frame}

\section{Data}

\frame{\sectionpage}

\begin{frame} {Data}

\subsection{Ragged End}

\begin{itemize}
\item Selected (small) data set: BIP,ip,ifo-c,ifo-exp,ifo-l,ESI,spr-10y-3m (we did not consider pmi because the series is shorter: for consistency we thus would have to shorten the evaluation span: pmi is OK but not best overall)
\item Transformations: 
\begin{itemize}
\item log returns (except spread which is differenced only) 
\item scaling (we mostly work with standardized series)
\item trim scaled data to $\pm 5$ (to regularize singular pandemic data)
\end{itemize}
\item Target: BIP (shifted by forecast horizon and publication lag, see below)
\item Explanatories: see the above selection 
\item Ragged end: see below 
\end{itemize}
\end{frame}

\begin{frame} {Ragged End}
\begin{itemize}
\item Ragged end: prior to shifting:
% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Sat Feb 15 08:35:27 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & BIP & ip & ifo\_c & ifo\_exp & ifo\_l & ESI & spr\_10y\_3m \\ 
  \hline
2024-09-01 &  & 91.20 & 83.70 & 85.90 & 81.60 & 89.80 & -1.40 \\ 
  2024-10-01 & 900.76 & 90.80 & 84.10 & 87.50 & 80.80 & 90.70 & -1.20 \\ 
  2024-11-01 &  & 92.20 & 83.80 & 86.80 & 81.00 & 89.30 & -0.90 \\ 
  2024-12-01 &  &  & 82.80 & 84.80 & 80.90 & 86.90 & -0.80 \\ 
  2025-01-01 &  &  & 82.50 & 83.20 & 81.70 & 88.10 & -0.20 \\ 
  2025-02-01 &  &  &  &  &  &  & -0.20 \\ 
   \hline
\end{tabular}
\caption{Ragged end } 
\label{perf_var1}
\end{table}\item Findings:
\begin{itemize}
\item Publication lags: lag(BIP)=4, lag(ip) =2, lag(ESI)=1 (same as pmi); the other variables are contemporaneous
\item Real-time data: shift all series to be aligned at sample end
\item Introduce a new `target' column based on BIP. 
\item Target: real-time BIP shifted by sum of publication lag and forecast horizon $h=3$, i.e., shift=$h+lag=$7
\item Ignore data revisions

\end{itemize}
\end{itemize}
\end{frame}

\begin{frame} {Real-Time data: Forecast Horizon $h=3$}

\begin{itemize}

\item Real-time data matrix: shift of target depends on forecast horizon $h=3$:

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Sat Feb 15 08:35:27 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrrr}
  \hline
 & target & BIP & ip & ifo\_c & ifo\_exp & ifo\_l & ESI & spr\_10y\_3m \\ 
  \hline
2024-05-01 & 902.57 & 904.30 & 95.10 & 88.30 & 89.80 & 86.90 & 91.80 & -1.40 \\ 
  2024-06-01 & 902.57 & 904.30 & 94.60 & 89.30 & 90.80 & 87.80 & 92.40 & -1.30 \\ 
  2024-07-01 & 900.76 & 904.30 & 94.70 & 87.60 & 88.20 & 86.90 & 92.20 & -1.30 \\ 
  2024-08-01 & 900.76 & 901.62 & 91.70 & 85.80 & 87.60 & 84.10 & 92.50 & -1.40 \\ 
  2024-09-01 & 900.76 & 901.62 & 93.40 & 84.90 & 86.80 & 83.00 & 90.90 & -1.40 \\ 
  2024-10-01 & 900.76 & 901.62 & 90.70 & 83.70 & 85.90 & 81.60 & 89.80 & -1.20 \\ 
  2024-11-01 & 900.76 & 902.57 & 93.20 & 84.10 & 87.50 & 80.80 & 90.70 & -0.90 \\ 
  2024-12-01 & 900.76 & 902.57 & 91.20 & 83.80 & 86.80 & 81.00 & 89.30 & -0.80 \\ 
  2025-01-01 & 900.76 & 902.57 & 90.80 & 82.80 & 84.80 & 80.90 & 86.90 & -0.20 \\ 
  2025-02-01 & 900.76 & 900.76 & 92.20 & 82.50 & 83.20 & 81.70 & 88.10 & -0.20 \\ 
   \hline
\end{tabular}
\caption{Target column shifted upwards by h+publication lag=7. All other explanatories are shifted to be aligned at sample end.  } 
\label{perf_var1}
\end{table}\end{itemize}

\end{frame}

\subsection{Leads and Lags of `Real-Time' Data}

\begin{frame} {Leads/Lags}
\begin{itemize}
\item Analysis of leads/lags in real-time data matrix
\item Compute cross correlation function (CCF) between target and explanatory variables
\begin{itemize}
\item Peak of CCF indicates lead (left-shift) or lag (right-shift)
\item Note: (shift of) target depends on $h$
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame} {Data and CCF for $h=3$}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4in]{data.pdf}\caption{Selected data: original (top left) and transformed (top right). Cross correlation functions (bottom) \label{cor}}\end{center}\end{figure}\end{frame}

\begin{frame} {CCF for $h=3$}\label{ccf}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4in]{CCF.pdf}\caption{Selected data: original (top left) and transformed (top right). Cross correlation functions (bottom) \label{cor}}\end{center}\end{figure}\end{frame}

\begin{frame} {Leads/Lags}
% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Sat Feb 15 08:35:27 2025
\begin{table}[ht]
\centering
\begin{tabular}{rr}
  \hline
 & Peak correlation \\ 
  \hline
BIP & 7.00 \\ 
  ip & 5.00 \\ 
  ifo\_c & 3.00 \\ 
  ifo\_exp & 3.00 \\ 
  ifo\_l & 4.00 \\ 
  ESI & 3.00 \\ 
  spr\_10y\_3m & -5.00 \\ 
   \hline
\end{tabular}
\caption{Peak correlation: positive numbers signify a lag of the real-time explanatory (shifted backward by its publication lag) relative to the target (shifted forward  by the forecast horizon).} 
\label{perf_var1}
\end{table}
\begin{itemize}
\item Stair-step interpolation of BIP makes an assessment of leads/lags less clear (than for ip)
\end{itemize}
\end{frame}

\subsection{Non-Stationarity: Structural Changes?}

\begin{frame} {Non-Stationarity (Structural Change?)}
\begin{itemize}
\item The \textbf{dependence} structure of some of the series seems to \textbf{change} over time (in particular before/after Pandemic)
\item This can have more or less severe impacts on data \textbf{modelling}, \textbf{filtering}, forecast \textbf{performances}, \textbf{holding times} (to be defined below),...
\item Example: {ACF} of \textbf{ifo-exp} for data prior Pandemic (up to 2019) vs. full data set, see plot on next next 
\end{itemize}
\end{frame}

\begin{frame} {Non-Stationarity/Structural Change in Dependence: ifo-exp}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4in]{ACF_non_stat.pdf}\caption{Selected data: original (top left) and transformed (top right). Cross correlation functions (bottom) \label{cor}}\end{center}\end{figure}\end{frame}

\begin{frame} {Non-Stationarity/Structural Changes}
\begin{itemize}
\item Before Pandemic the data has a stronger and longer dependence (ARMA(model)): the data is \textbf{smoother}
\item On the \textbf{full} data set  the previous dependence structure simplifies to a simple lag-one dependence (MA(1)-model): the data is \textbf{noisier} (higher vola and more frequent zero-crossings)
\end{itemize}
\end{frame}

\section{Benchmarks and Direct AR Forecasts}

\frame{\sectionpage}

\subsection{Design}

\begin{frame} {Benchmarks and Direct AR Predictors}
\begin{itemize}
\item Use data from 1990-02-01 to 2025-02-01 
\begin{itemize}
\item Singular readings during pandemic affect OLS estimation
\item Use trimmed scaled data (threshold $\pm 5$)
\end{itemize}

\item Denote $target_{t}(h)$: first column of real-time data matrix, i.e. real-time $BIP$ shifted forward by lag+h=7
\begin{itemize}
\item Stair-step extrapolation not ideal  
\end{itemize}
\item Benchmark 1: mean of BIP, i.e., $\frac{1}{t}\sum_{k=1}^t BIP_k$
\item Benchmark 2: regression of target on BIP:
\[target_{t}(h)=c+a_1 BIP_{t}+...+a_pBIP_{t-p}+\epsilon_t\]
\item Additional benchmarks: regress any of the real-time indicators on target 
\end{itemize}
\end{frame}

\begin{frame} {ARMA-Models}
\begin{itemize}
\item For each transformed series we can fit an ARMA-model: $p=3$, $q=1$
\item The model is not used for direct forecasts (which are based on OLS regression, see previous slides) 
\item Instead, the ARMA-models are used for computing the real-time filters, see below
\item For each model we can obtain its MA-inversion, see next plot in the case of BIP
\end{itemize}
\end{frame}

\begin{frame} {Benchmark 2: ACF and MA-Inversion of BIP}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{ma_inv_ip.pdf}\caption{\label{data}}\end{center}\end{figure}\end{frame}

\begin{frame} {Findings Benchmark 2}
\begin{itemize}
\item The MA-inversion of BIP as well as the ACF are close to white noise, i.e., \textbf{Benchmark 1$\approx$Benchmark 2}.
\begin{itemize}
\item Stair-step interpolation not ideal 
\item The regressors in the direct AR predictors are insignificant, see t-values further down
\item This result is reliant on the span for estimation (with/without Pandemic), see below for further details
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame} {Forecast horizon}
\begin{itemize}
\item We here emphasize a one quarter ahead forecast horizon $h=3$
\item Alternative horizons such as $h=0$ or $h=12$ lead to similar qualitative outcomes but significance levels (t-tests) are affected
\end{itemize}
\end{frame}

\subsection{Performances}

\begin{frame} {Relative Forecast Performances}
\begin{itemize}

\item Regress the (real-time) indicator $y_{it}$ on the target to obtain the associated $MSE_i=E[\hat{\epsilon}_{it}^2]$, where $\hat{\epsilon}_{it}$ is the residual of the regression

\item Relative forecast MSEs: $MSE_i/MSE(benchmark)$. 
\item We can consider Benchmark 1 or 2 (mean of BIP or AR forecast based on BIP: virtually the same)
\item Numbers smaller one indicate outperformance of the corresponding indicator
\item The following table summarizes performances for the \textbf{entire} sample (pmi would be in the middle)
\end{itemize}

\end{frame}

\begin{frame} {Relative Forecast Performances: Forecast Benchmarks, $h=3$, Full Sample}\label{rmse1}

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Sat Feb 15 08:35:44 2025
\begin{table}[ht]
\centering
\begin{tabular}{rr}
  \hline
 & Relative MSE h=3 \\ 
  \hline
BIP & 1.00000 \\ 
  ip & 0.99472 \\ 
  ifo\_c & 1.00007 \\ 
  ifo\_exp & 1.00054 \\ 
  ifo\_l & 0.99378 \\ 
  ESI & 1.00078 \\ 
  spr\_10y\_3m & 0.99615 \\ 
   \hline
\end{tabular}
\caption{Relative mean-square performances against benchmark 2 (which is virtually identical with benchmark 1, the mean of BIP).} 
\label{perf_var1}
\end{table}\end{frame}

\subsection{Effect of Evaluation Sample}

\begin{frame} {Effect of Sample Selection}
\begin{itemize}
\item The above predictors and the evaluation span rely on the full sample, including Pandemic
\item All results are statistically insignificant, see below for details
\item What happens if we discard the singular (trimmed) Pandemic?
\item The following table reports relative MSEs for data prior to 2019-01-01 
\end{itemize}

\end{frame}

\begin{frame} {Relative Forecast Performances: Evaluation on Data Before Pandemic}\label{rmse2}

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Sat Feb 15 08:35:44 2025
\begin{table}[ht]
\centering
\begin{tabular}{rr}
  \hline
 & Relative MSE h=3 \\ 
  \hline
BIP & 1.00000 \\ 
  ip & 0.99895 \\ 
  ifo\_c & 0.96738 \\ 
  ifo\_exp & 0.97755 \\ 
  ifo\_l & 0.97750 \\ 
  ESI & 0.98354 \\ 
  spr\_10y\_3m & 0.99520 \\ 
   \hline
\end{tabular}
\caption{Relative mean-square performances against benchmark 2 (which is virtually identical with benchmark 1, the mean of BIP).} 
\label{perf_var1}
\end{table}\end{frame}

\begin{frame} {Findings}
\begin{itemize}
\item Indicators perform less well \textbf{after} 2019-01-01 in relative terms
\item Ifo series perform best (before Pandemic)
\item Difference is statistically significant (see below for more detailed results)
\end{itemize}

\end{frame}

\section{Univariate Filters}

\frame{\sectionpage}

\subsection{Alternative Targets}

\begin{frame} {Alternative Targets}
\begin{itemize}
\item Instead of looking at BIP-returns shifted by $lag+h$ (7-steps ahead) we might be interested in looking at the \textbf{trend} of BIP-returns: \emph{trend-growth}
\item Which `trend'?
\begin{itemize}
\item Ideal trend, model-based trend (SEATS, state-space), Henderson, Hodrick Prescott (HP),...
\end{itemize}
\item Due to its wide use in applications we look at \textbf{HP}(14400) (monthly design)
\item On the next slide, we plot and compare the two targets: original shifted BIP (violet line) and HP applied to shifted BIP (brown)
\end{itemize}
\end{frame}

\begin{frame} {Targets 1 and 2: BIP-Shifted (violet), BIP-Trend (brown), ip Real-Time (black): All Series Standardized }
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{alternative_targets_1.pdf}\caption{\label{data}}\end{center}\end{figure}\end{frame}

\begin{frame} {Findings}
\begin{itemize}
\item The filter output (target 2: brown) tracks the (long-term/systematic) trend-growth of target 1 (violet)
\begin{itemize}
\item Business-cycle or \textbf{growth-cycle}: depending on how the filter is implemented, a negative sign can be associated with negative economic growth (growth cycle), under-par growth (below potential output), recessions/crises (see SSA paper)
\item The sign of the filtered data, i.e., its \emph{zero-crossings}, have a potentially meaningful \emph{economic} interpretation
\end{itemize}
\item \textbf{Problem}: two-sided (acausal) filter cannot be computed in real-time 
\end{itemize}
\end{frame}

\subsection{Causal Filter Design}

\begin{frame} {Causal Filter Design}
\begin{itemize}
\item We can apply the two-sided HP to each indicator (ip, ifo, ESI, spread,...)
\item \textbf{Task}: for each indicator compute a \emph{causal} filter which tracks the two-sided HP (applied to that indicator) 
\begin{itemize}
\item \textbf{MSE} and \textbf{SSA} designs: the latter encompasses the former (see SSA paper)
\item We account for $h$ and the publication lag of each indicator (details left aside)
\end{itemize}
\item New \textbf{predictors}: MSE and SSA (applied to indicators)
\item Potentially interesting \textbf{targets}
\begin{itemize}
\item We here continue to look at \emph{BIP-shifted} (violet line in above plot) as our main target of interest 
\item Later, we also consider \emph{BIP-trend}, i.e. the two-sided HP  applied to BIP-shifted (brown line), as an alternative target
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame} {Forecast Performance}\label{overf}
\begin{itemize}
\item For each indicator, we compute the mean-square forecast error $E[(BIP_{t+h}-\hat{T}_{ti})^2]$, where $BIP_{t+h}$ is the target series (shifted BIP) and $\hat{T}_{ti}$ is either the MSE- or the SSA-output when applied to  the $i-$th indicator
\item Note that by design, $\hat{T}_{ti}$ does not target $BIP_{t+h}$. Instead it tracks the two-sided HP applied to the $i$-th indicator (accounting for $h$ and publication lag of that indicator) 
\item \textbf{Idea}: if $\hat{T}_{ti}$ tracks the (future) trend-growth of the $i$-th indicator and if the indicator is informative about the effective target $BIP_{t+h}$, then $\hat{T}_{ti}$ should also track $BIP_{t+h}$ `somehow'
\begin{itemize}
\item Validate MSE- and SSA-trends by computing performances against the `indirect' target $BIP_{t+h}$ 
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame} {MSE and SSA for BIP}
\begin{itemize}
\item The figure on the next slide displays
\begin{itemize}
\item The \textbf{MA inversion} of the ARMA-model fitted to BIP (top left)
\item The \textbf{acausal} two-sided HP (top right)
\item The \textbf{causal} filters: MSE (green), SSA (blue) and classic concurrent HP-C (violet)
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame} {Univariate Filters for BIP}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{filt_coef_INDPRO_uni_x.pdf}\caption{\label{data}}\end{center}\end{figure}\end{frame}

\subsection{Performances: Comparison with AR-Benchmarks}

\begin{frame} {Relative Forecast Performances}
\begin{itemize}
\item In the next table, we compare relative forecast performances of all predictors
\begin{itemize}
\item direct \textbf{AR} forecasts based on indicators: first column
\item trend forecasts \textbf{SSA}: second column
\item trend forecasts \textbf{MSE}: third column
\end{itemize}
\item Relative forecast error: all forecast mean-square errors are normalized by AR-forecast based on BIP (benchmark 2). 
\begin{itemize}
\item Numbers smaller one signify outperformance of the benchmark (significance is analyzed further down)
\end{itemize}
\item We assume \textbf{calibrated} filters, i.e., we regress filters on target to account for (static) level and scale parameters (filters implicitly assume the data to be centered and the scale differs from BIP) 
\end{itemize}
\end{frame}

\begin{frame} {Relative Forecast Performances: $h=3$ Full Sample}\label{rmse3}

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Sat Feb 15 08:35:44 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrr}
  \hline
 & Direct AR forecasts & SSA & MSE \\ 
  \hline
BIP & 1.0000 & 1.0048 & 1.0053 \\ 
  ip & 0.9947 & 1.0066 & 1.0069 \\ 
  ifo\_c & 1.0001 & 0.9897 & 0.9908 \\ 
  ifo\_exp & 1.0005 & 0.9859 & 0.9888 \\ 
  ifo\_l & 0.9938 & 0.9947 & 0.9941 \\ 
  ESI & 1.0008 & 0.9924 & 0.9945 \\ 
  spr\_10y\_3m & 0.9961 & 0.9937 & 0.9931 \\ 
   \hline
\end{tabular}
\caption{Relative mean-square performances against benchmark 2 (which is virtually identical with benchmark 1, the mean of BIP).} 
\label{perf_var1}
\end{table}\end{frame}

%\begin{frame} {Relative Forecast Performances: `Regression'}

%<<label=ats_mba_2,echo=FALSE,results=tex>>=

%mat<-lm_perf_mat
%xtable((mat), dec = 1,digits=rep(2,dim(mat)[2]+1),
%paste("Relative mean-square performances against benchmark 2 (which is virtually identical with benchmark 1, the mean of ip)."),
%label=paste("perf_var1",sep=""),
%center = "centering", file = "", floating = FALSE)
%@
%\end{frame}

\begin{frame} {Discussion}
\begin{itemize}
\item Calibrated trend forecasts (second and third columns) are generally not worse than `direct' AR indicator forecasts (first column)
\item MSE and SSA are similar (the latter is smoother: less zero-crossings, see below for details)

\item \textbf{Significance}: t-values in regressions of predictors on target, see next slide

\end{itemize}

\end{frame}

\begin{frame} {Significance ($h=3$)}\label{tstat1}

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Sat Feb 15 08:35:44 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrr}
  \hline
 & t\_stat direct AR & t-stat SSA & t-stat MSE \\ 
  \hline
BIP & 0.44 & -0.79 & -0.70 \\ 
  ip & 1.40 & -0.31 & -0.08 \\ 
  ifo\_c & 0.40 & 2.30 & 2.22 \\ 
  ifo\_exp & 0.13 & 2.54 & 2.35 \\ 
  ifo\_l & 1.46 & 1.93 & 1.98 \\ 
  ESI & -0.18 & 2.11 & 1.95 \\ 
  spr\_10y\_3m & 1.25 & 2.00 & 2.06 \\ 
  Consensus & 2.11 & 2.51 & 2.48 \\ 
   \hline
\end{tabular}
\caption{Significance of predictors: t-statistics from a regression of the predictors on the target. Consensus forecast is the equally weighted mean of all forecasts.} 
\label{perf_var1}
\end{table}\end{frame}

\begin{frame} {Findings}
\begin{itemize}
\item Direct AR predictors are insignificant (on full sample, including Pandemic)
\item SSA and MSE similar (with respect to t-statistics) and some of them are significant
\item \textbf{Consensus} (equally-weighted) generally perform well
\end{itemize}
\end{frame}

\subsection{Effect of Evaluation Sample}

\begin{frame} {Relative Forecast Performances: Evaluation on Data Before Pandemic}\label{rmse4}

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Sat Feb 15 08:35:44 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrr}
  \hline
 & Direct AR forecasts & SSA & MSE \\ 
  \hline
BIP & 1.00 & 1.00 & 1.00 \\ 
  ip & 1.00 & 1.01 & 1.01 \\ 
  ifo\_c & 0.97 & 0.98 & 0.97 \\ 
  ifo\_exp & 0.98 & 0.96 & 0.96 \\ 
  ifo\_l & 0.98 & 1.00 & 0.99 \\ 
  ESI & 0.98 & 0.98 & 0.98 \\ 
  spr\_10y\_3m & 1.00 & 0.99 & 0.99 \\ 
   \hline
\end{tabular}
\caption{Relative mean-square performances against benchmark 2: data prior to Pandemic.} 
\label{perf_var1}
\end{table}\end{frame}

\begin{frame} {Significance: Data prior to Pandemic}\label{tstat2}

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Sat Feb 15 08:35:44 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrr}
  \hline
 & t\_stat direct AR & t-stat SSA & t-stat MSE \\ 
  \hline
BIP & 0.81 & -1.41 & -1.37 \\ 
  ip & 1.12 & -1.21 & -0.96 \\ 
  ifo\_c & 2.98 & 2.62 & 2.96 \\ 
  ifo\_exp & 2.49 & 3.35 & 3.67 \\ 
  ifo\_l & 2.52 & 1.72 & 1.98 \\ 
  ESI & 2.13 & 2.68 & 2.86 \\ 
  spr\_10y\_3m & 1.50 & 2.24 & 2.32 \\ 
  Consensus & 3.25 & 2.82 & 3.14 \\ 
   \hline
\end{tabular}
\caption{Significance of predictors: t-statistics from a regression of the predictors on the target. Consensus forecast is the equally weighted mean of all forecasts.} 
\label{perf_var1}
\end{table}\end{frame}

\begin{frame} {Findings}
\begin{itemize}
\item \textbf{Main effect}: t-statistics \textbf{markedly larger} (despite shorter sub-span)
\item Direct AR forecasts often strongly significant: performances similar to filters
%\item Ranking (hierarchy) of direct AR-forecasts \textbf{affected}: ifo-exp and ifo-c are more important than ESI on shorter sub-sample 
%\item Hierarchy of trend forecasts less affected

\end{itemize}
\end{frame}

\begin{frame} {Alternative Evaluation Criteria}
\begin{itemize}
\item Until yet we emphasized the mean-square forecast error
\item Question: what about \textbf{smoothness}, i.e., noise suppression and reliability? 
\begin{itemize}
\item Original justification for trend targets: for identical (relative) forecast performances we might prefer a \textbf{more reliable/regular} (smoother/less noisy) forecast rule
\item The above results suggest that MSE forecast performances are \textbf{not} negatively affected by filtering (`trends')
\item What about \textbf{zero-crossings} and shape/dynamics of filter outputs vs. direct AR forecasts
\item And what about \textbf{retardation} (lag/right-shift) of trend forecasts?
\end{itemize}
\end{itemize}
\end{frame}

\subsection{Zero-Crossings and Holding Time}

\begin{frame} {Zero-Crossings and Holding Times}\label{htzc1}
\begin{itemize}
\item \textbf{Holding time} (HT): mean duration between consecutive zero-crossings
\item \textbf{SSA}: impose $50\%$ larger \emph{expected }HT than MSE 
\begin{itemize}
\item Dual formulation of SSA \textbf{optimization} principle: \emph{maximize HT} for given tracking accuracy (target correlation or MSE)
\end{itemize}
\item Compute and compare \emph{empirical} HTs (larger = smoother):
\end{itemize}

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Sat Feb 15 08:35:44 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrr}
  \hline
 & HT AR & HT MSE & HT SSA \\ 
  \hline
BIP & 2.27 & 19.00 & 23.38 \\ 
  ip & 1.68 & 7.07 & 11.26 \\ 
  ifo\_c & 3.01 & 11.26 & 14.48 \\ 
  ifo\_exp & 2.95 & 9.81 & 11.69 \\ 
  ifo\_l & 2.69 & 9.81 & 16.00 \\ 
  ESI & 2.64 & 12.67 & 19.00 \\ 
  spr\_10y\_3m & 2.81 & 9.50 & 12.67 \\ 
   \hline
\end{tabular}
\caption{Empirical (sample) Holding times of predictors: direct AR, MSE and SSA.} 
\label{perf_var1}
\end{table}\end{frame}

\begin{frame} {Zero-Crossings and Holding Times}
\begin{itemize}
\item Filters much smoother (less crossings) than direct AR forecasts
\item Sample HTs of M-SSA approximately 50$\%$ larger than MSE (series are centered and therefore expected and sample HTs match better than in ip-case: see corresponding slides)
\item On the folowing slide we display AR-forecasts (orange) and MSE- (green) and SSA-filters (blue) 
\begin{itemize}
\item Zero-crossings are marked by corresponding colored vertical lines
\item Series are standardized (scaled and centered)
\end{itemize}
\end{itemize}
\end{frame}

%\begin{frame} {Zero-crossings ifo-exp: post Financial Crisis}
%<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
%file = "zero_cross_ifo_exp.pdf"
%cat("\\begin{figure}[H]")
%cat("\\begin{center}")
%cat("\\includegraphics[height=3in, width=4.5in]{", file, "}\n",sep = "")
%cat("\\caption{", sep = "")
%cat("\\label{data}}", sep = "")
%cat("\\end{center}")
%cat("\\end{figure}")
%@
%\end{frame}

\begin{frame} {Zero-crossings BIP: Standardized Series}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{zero_cross_ifo_l_1.pdf}\caption{\label{data}}\end{center}\end{figure}\end{frame}

\begin{frame} {Findings}
\begin{itemize}
\item See similar comments in ip-case (slides ip)
\item Sign of lagged BIP is negative (predictor is not significant): explains `strange' readings of direct AR predictor (orange) during financial crisis
\end{itemize}
\end{frame}

\section{Multivariate Filters}

\frame{\sectionpage}

\subsection{Dependence}

\begin{frame} {Dependence}
\begin{itemize}
\item Multivariate M-MSE and M-SSA: rely on several time series to track the two-sided HP 
\item For the multivariate design we consider BIP, ip, ifo-exp, ESI, spr-10y-3m
\begin{itemize}
\item Criterion for \textbf{selection}: the above series differ in terms of leads/lags (see CCF on slide \eqref{ccf}) \item Therefore a multivariate design can exploit  the mutually differing (as well as the common) features of each series (see below) 
\end{itemize}
\item The next two slides display the \textbf{CCF}s corresponding to the full sample (first slide) and data up to the Pandemic (second slide) 
\begin{itemize}
\item The CCF refers to the \emph{real-time} data: after accounting for publication lags
\end{itemize}
\end{itemize}

\end{frame}

\begin{frame} {CCF: Full Data Set}\label{ccf_pand_multe}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4in]{acf_all.pdf}\caption{MA-inversion of VARMA\label{cor}}\end{center}\end{figure}\end{frame}

\begin{frame} {CCF: Up to Pandemic}\label{ccf_pand_mult}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4in]{acf_pan.pdf}\caption{MA-inversion of VARMA\label{cor}}\end{center}\end{figure}\end{frame}

\begin{frame} {Findings}\label{slide_findings_ccf_mult}
\begin{itemize}
\item Significant ACFs of BIP at multiples of three months: effect of \textbf{stair-step} interpolation
\item \textbf{Non-stationarity}: dependence structure changes after pandemic (structural change?)
\item Dependence: BIP cross-correlates with ip, ifo and ESI (first row in above plots). 
\item Leads/lags correspond to peak-CCF on slide \eqref{ccf} (cross-check)
\item Additional findings: 
\begin{itemize}
\item \textbf{BIP} and \textbf{spread} are nearly \textbf{uncorrelated}! (at observed leads/lags) 
\item \textbf{Spread} seems to correlate with (and to lead) \textbf{ifo}-series (but not ip). Maybe spread enters somehow into construction of ifo?
\end{itemize}
\end{itemize}

\end{frame}

\subsection{VARMA}

\begin{frame} {VARMA}
\begin{itemize}
\item Fit a VARMA model to BIP, ip, ifo-exp, ESI, spr-10y-3m
\begin{itemize}
\item Rely on R-package MTS by Ruey Tsay (we also adopt regularization to avoid overfitting)
\end{itemize}
\item Purpose: VARMA model is used for M-SSA and M-MSE (need to account for leads/lags and CCF)
\item Estimation span: from first observation to $2024-11-01$
%\item Explanatories: BIP, ip, ifo-exp, ESI, spr-10y-3m 
\item Model orders: VARMA(2,0), see below for details 
\item The (V)MA-inversion of the VARMA is displayed on the next slide
\end{itemize}

\end{frame}

\begin{frame} {MA Inversion of VARMA}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4in]{ma_inv_multi_ip.pdf}\caption{MA-inversion of VARMA\label{cor}}\end{center}\end{figure}\end{frame}

\begin{frame} {Findings}
\begin{itemize}
\item Interpretation: see ip-case (slides ip)
\item We can \textbf{interpret} VARMA model: simple and explainable
\end{itemize}

\end{frame}

\subsection{M-SSA and M-MSE Filter Design}

\begin{frame} {M-SSA and M-MSE Filters}
\begin{itemize}
\item For each series $y_{it}$, $i=1,...,5$ of the multivariate filter, the \textbf{target} $T_{it}$ is specified by applying the acausal HP to $y_{it}$ 
\begin{itemize}
\item \textbf{Same trend} targets as in univariate case 
\item But each trend target can rely on \emph{multiple} time series for deriving the \emph{causal} filter
\item \emph{VARMA}-model is used for computing the \emph{causal} filters
\end{itemize}
\item M-SSA: impose the \textbf{same HT} as in univariate case
\item Overall design decisions: compare \textbf{apples with apples} (can compare uni- and multivariate designs in a consistent way)
\item The following two slides display multivariate MSE (M-MSE) and SSA (M-SSA) filters
\end{itemize}

\end{frame}

\begin{frame} {M-MSE}\label{m_mse_p}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4in]{MSE_multi_ip.pdf}\caption{MA-inversion of VARMA\label{cor}}\end{center}\end{figure}\end{frame}

\begin{frame} {M-SSA}\label{m_ssa_p}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4in]{MSSA_multi_ip.pdf}\caption{MA-inversion of VARMA\label{cor}}\end{center}\end{figure}\end{frame}

\begin{frame} {MA-Inversion M-SSA: Filter as Applied to VARMA-Residuals}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4in]{MSSA_multi_eps_ip.pdf}\caption{MA-inversion of VARMA\label{cor}}\end{center}\end{figure}\end{frame}

\begin{frame} {Findings}
\begin{itemize}
\item Interpretation: see ip-case (slides ip)
\item Intuitively \textbf{appealing} (interpretable) outcome
\end{itemize}
\end{frame}

\subsection{M-MSE and M-SSA Outputs as Predictors for BIP-Shifted}

\begin{frame} {Relative (Mean Square) Forecast Error: Target is Shifted BIP}
\begin{itemize}
\item The table on the next slide reports relative mean-square forecast errors 
\item Specifically:
\begin{itemize}
\item Our \textbf{target} for forecast evaluation is the shifted BIP 
\item The $i-$th output of the multivariate filter is designed to track the \emph{acausal HP} applied to the \emph{$i-$th indicator}, $i=1,...,5$
\begin{itemize}
\item By construction, M-MSE and M-SSA do not explicitly `see' or target the shifted BIP. 
\item Nevertheless, we expect that the filter outputs should track BIP `somehow' (see discussion for univariate case above)
\end{itemize}
\end{itemize}
\item Below, we extend the analysis to \emph{acausal HP} targets.

\end{itemize}
\end{frame}

\begin{frame} {Relative (Mean Square) Forecast Error (Full Sample)}\label{slide_multi_rel_mse}

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Sat Feb 15 08:36:40 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrr}
  \hline
 & Relative MSE & t-statistic \\ 
  \hline
M-SSA BIP & 1.01 & -0.64 \\ 
  M-SSA ip & 1.00 & 0.07 \\ 
  M-SSA ifo\_exp & 0.98 & 2.75 \\ 
  M-SSA ESI & 0.98 & 2.35 \\ 
  M-SSA spr\_10y\_3m & 0.99 & 2.01 \\ 
  M-MSE BIP & 1.01 & -0.67 \\ 
  M-MSE ip & 1.00 & 0.02 \\ 
  M-MSE ifo\_exp & 0.98 & 2.67 \\ 
  M-MSE ESI & 0.99 & 2.20 \\ 
  M-MSE spr\_10y\_3m & 0.99 & 2.03 \\ 
   \hline
\end{tabular}
\caption{Relative mean-square forecast performances and t-statistics of multivariate designs (target is shifted BIP)} 
\label{perf_var1}
\end{table}
\end{frame}

\begin{frame} {Relative (Mean Square) Filter Error: Target acausal HP}
\begin{itemize}
\item Findings: the above performances are \textbf{comparable} to the previous univariate designs, see slide \eqref{tstat1}
\item Question: what is the \textbf{added-value} of multivariate designs?
\item For this purpose we need to consider {alternative} \textbf{trend-targets}
\item What is the mean-square \textbf{filter error} with respect to the \textbf{acausal HP} (applied to each indicator)?
\item The table on the next slide computes the relative mean-square \emph{filter errors} of \emph{multivariate} against \emph{univariate} filters
\item \textbf{Added-value} of multivariate designs: numbers smaller 1 signify an outperformance by the  multivariate design (with respect to the \emph{trend} targets)
\end{itemize}
\end{frame}

\begin{frame} {Relative (Mean Square) Filter Error}\label{rfmse1}

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Sat Feb 15 08:36:40 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrr}
  \hline
 & M-MSE & M-SSA \\ 
  \hline
BIP & 0.91 & 0.91 \\ 
  ip & 0.90 & 0.89 \\ 
  ifo\_exp & 0.93 & 0.93 \\ 
  ESI & 1.04 & 1.05 \\ 
  spr\_10y\_3m & 1.06 & 1.08 \\ 
   \hline
\end{tabular}
\caption{Relative mean square filter errors: ratio of multivariate over univariate filters (number smaller one signify outperformance by multivariate designs). Targets of uni and multivariate filter are identical but multivariate designs can rely on additional explanatory series} 
\label{perf_var1}
\end{table}
\end{frame}

\begin{frame} {Findings}
\begin{itemize}
\item Findings similar to ip-case (slides ip forecasting) 
\item However, M-SSA performs worse than SSA for ESI and spread. Explanations:
\begin{itemize}
\item For spread, M-SSA is worse because it is smoother: sample HT of M-SSA is substantially larger (not shown)
\item In the case of ESI, M-SSA is worse probably because of sample variations (and model misspecification: but the latter affects all (M-)SSA filters) 
\item In fact M-SSA trend for ESI `looks better' (not shown): it is slightly smoother and faster (left-shifted) than SSA because the multivariate filter relies mainly on ifo-exp (as additional explanatory variable), which is leading, due to its smaller publication lag
\end{itemize}
\end{itemize}
%<<label=ats_mba_2,echo=FALSE,results=tex>>=

%mat<-cbind(perf_mat_multi_sample[,"HT M-SSA"],perf_mat_uni_sample[select_vec_multi,"HT SSA"])
%colnames(mat)<-c("M-SSA","SSA")

%xtable(mat, dec = 1,digits=rep(2,dim(mat)[2]+1),
%paste("Relative mean square filter errors: ratio of multivariate over univariate filters (number smaller one signify outperformance by multivariate designs). Targets of uni and multivariate filter are identical but multivariate designs can rely on additional explanatory series"),
%label=paste("perf_var1",sep=""),
%center = "centering", file = "", floating = FALSE)
%@

\end{frame}

\subsection{M-MSE and M-SSA Outputs as Predictors for BIP-Trend}

\begin{frame} {BIP-Shifted and BIP-Trend Targets}
\begin{itemize}
\item Up to now, we considered different targets in this study: \textbf{shifted BIP} and two-sided \textbf{HP-trend}  applied to each of the different indicators
\item In the following {evaluation} we restrict this choice by emphasizing \textbf{BIP}. Moreover we consider an alternative performance measure.
\item Evaluation framework:
\begin{enumerate}
\item Two targets: BIP-shifted and BIP-trend, i.e., acausal HP applied to BIP-shifted. In particular, we discard HP applied to the other indicators.
\item Performance measure: instead of mean-square forecast performances we compute the (sample) \textbf{target correlations}, i.e., the (sample) correlations between the outputs of M-MSE (or M-SSA) and each one of the two targets
\end{enumerate}
\end{itemize}
\end{frame}

\begin{frame} {BIP-Shifted and BIP-Trend Targets}
\begin{itemize}
\item The table on the next slide reports corresponding \emph{sample target correlations} for M-MSE and M-SSA against BIP-trend (first two columns) and shifted BIP (last two columns)
\begin{itemize}
\item We consider the 5 filter outputs (the 5 rows in the table) corresponding to BIP, ip, ifo-exp, ESI, spr-10y-3m  in M-MSE and M-SSA as predictors for each one of the two targets
\item Note that M-SSA (or M-MSE) outputs corresponding to ip, ESI, ifo-exp and spread (rows 2-5 in the table) do not target BIP-shift or BIP-trend explicitly (no overfitting) 
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame} {Sample Target Correlations}\label{sample_target_cor1}
% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Sat Feb 15 08:36:40 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & M-MSE HP & M-SSA HP & M-MSE BIP & M-SSA BIP \\ 
  \hline
BIP & 0.21 & 0.17 & -0.04 & -0.04 \\ 
  ip & 0.30 & 0.28 & 0.00 & 0.00 \\ 
  ifo\_exp & 0.46 & 0.47 & 0.15 & 0.16 \\ 
  ESI & 0.66 & 0.66 & 0.13 & 0.13 \\ 
  spread & -0.14 & -0.09 & 0.12 & 0.11 \\ 
   \hline
\end{tabular}
\caption{Target correlations: correlations of M-MSE and M-SSA filter outputs with BIP-trend (first two columns) and with BIP-shifted (last two columns).} 
\label{perf_var1}
\end{table}\end{frame}

\begin{frame} {Findings}\label{slide_target_cor}
\begin{itemize}
%\item M-MSE outperforms M-SSA with respect to trend target (first two columns), by design
\item In general it is \textbf{`easier'} to track the (shifted) BIP-\textbf{trend} target than the (shifted) BIP: correlations in the first two columns (of the above table) tend to be larger
\item Interestingly, the filter outputs corresponding to ip, ESI and ifo-exp correlate significantly  with (shifted) BIP-trend (first two columns, rows 2-4), despite the singular Pandemic
\item Our findings provide \textbf{additional justification} for considering \textbf{trend} targets (instead of shifted data)
\item In particular the correlations in rows 2-4 (first two columns) suggest that the trend outputs corresponding to ip, ESI and ifo in the multivariate filter are \emph{candidates} for predicting the future trend-growth of BIP
\item On the next slide we compare target correlations of M-SSA and SSA with respect to BIP-trend  
\end{itemize}
\end{frame}

\begin{frame} {Sample Target Correlations Against BIP-trend : M-SSA vs. SSA}\label{sample_target_cor2}
% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Sat Feb 15 08:36:40 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrr}
  \hline
 & SSA HP & M-SSA HP \\ 
  \hline
BIP & 0.08 & 0.17 \\ 
  ip & 0.15 & 0.28 \\ 
  ifo\_exp & 0.50 & 0.47 \\ 
  ESI & 0.66 & 0.66 \\ 
  spr\_10y\_3m & -0.11 & -0.09 \\ 
   \hline
\end{tabular}
\caption{Target correlations: SSA vs. M-SSA.} 
\label{perf_var1}
\end{table}\end{frame}

\begin{frame} {Findings}
\begin{itemize}
\item Multivariate filters of lagged series (first two rows) improve over univariate designs: M-SSA can rely on `faster' explanatory series (this information is missing in univariate designs)
\end{itemize}
\end{frame}

\begin{frame} {M-SSA Outputs as Predictors for BIP-Trend}
\begin{itemize}
\item In the following five slides we display the M-SSA filter outputs together with BIP-trend (black line)
\begin{itemize}
\item The first plot displays all M-SSA outputs together with the target (black line)
\item The second plot emphasizes M-SSA for BIP and \textbf{ip} (together with the target)
\item The third plot emphasizes M-SSA for BIP and \textbf{ESI} (together with the target)
\item The fourth plot emphasizes M-SSA for BIP and \textbf{ifo-exp} (together with the target)
\item The last plot emphasizes M-SSA for BIP and \textbf{spread} (together with the target)
\end{itemize}
\item For ease of visual inspection, all series are \textbf{standardized} (standardization is equivalent to filter calibration)

\end{itemize}
\end{frame}

\begin{frame} {BIP-Trend (black) and All M-SSA Outputs}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4in]{M_SSA_output_1.pdf}\caption{MA-inversion of VARMA\label{cor}}\end{center}\end{figure}\end{frame}

\begin{frame} {BIP-Trend (black) and M-SSA: BIP and ip Outputs}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4in]{M_SSA_output_15.pdf}\caption{MA-inversion of VARMA\label{cor}}\end{center}\end{figure}\end{frame}

\begin{frame} {BIP-Trend (black) and M-SSA: BIP and ESI Outputs}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4in]{M_SSA_output_2.pdf}\caption{MA-inversion of VARMA\label{cor}}\end{center}\end{figure}\end{frame}

\begin{frame} {BIP-Trend (black) and M-SSA: BIP and ifo-exp Outputs}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4in]{M_SSA_output_3.pdf}\caption{MA-inversion of VARMA\label{cor}}\end{center}\end{figure}\end{frame}

\begin{frame} {BIP-Trend (black) and M-SSA: BIP and spread Outputs}\label{spr_mssa_o}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4in]{M_SSA_output_4.pdf}\caption{MA-inversion of VARMA\label{cor}}\end{center}\end{figure}\end{frame}

\begin{frame} {Findings}
\begin{itemize}
\item M-SSA BIP (red) is lagging target (black) due to its (largest) publication lag
\item M-SSA ip (yellow-green) is also lagging but less than M-SSA BIP
\item M-SSA ESI (blue) is both `fast' (left-shifted) and smooth: this explains the large target correlation reported in the above table(s)
\item M-SSA ifo-exp (green) is fastest but overlaid with `noisy' cycles
\item Spread (violet) does not seem to correlate with target (black) 
\end{itemize}
\end{frame}

\section{Addressing Retardation (and Smoothness)}

\frame{\sectionpage}

\subsection{Forecast Excess}

\begin{frame} {Addressing Retardation and Smoothness (HT)}
\begin{itemize}
\item Stronger smoothing generally signifies a more pronounced retardation (right shift; lag): dilemma
\item \emph{Retardation} of our filter designs can be addressed by \emph{increasing the forecast horizon} $h$ to $h+\delta$, where $\delta\geq 0$
\begin{itemize}
\item We here illustrate filters with \emph{forecast excesses} $\delta=0,3,6,12,18$ and 24 ($h=3$ is maintained)
\end{itemize}
\item In general, \textbf{increasing} $\delta$ leads ceteris paribus to a \textbf{left-shift} of the predictors (smaller retardation) and to \textbf{noisier} filters with \textbf{smaller HT} (more zero-crossings). 
\item However, in the \textbf{M-SSA} framework we can control and \textbf{fix the HT}: the M-SSA predictor will maintain a fixed HT irrespective of $\delta$
\item Therefore, we can \textbf{maintain smoothness} while \textbf{mitigating retardation} to some extent (by extending the classic forecast dilemma to a more general forecast trilemma, see SSA paper) 
\end{itemize}

\end{frame}

\subsection{M-SSA-BIP and Variable Forecast Excess}

\begin{frame} {Addressing Retardation and Smoothness (HT)}
\begin{itemize}
\item In the previous slides we saw that M-SSA-BIP (the output of M-SSA corresponding to the shifted BIP-target: red lines) was lagging
\item The next two plots illustrate the effect of $\delta=0,3,6,12,18,24$ on M-SSA-BIP  for the whole sample (first plot) and for the financial crisis and the pandemic (second plot)
\begin{itemize}
\item We \textbf{fix} the target, i.e. M-SSA-BIP, and we \textbf{vary} $\delta$
\item Later, we will compare \textbf{all} M-SSA outputs (not only M-SSA-BIP) for  $\delta=12$ \textbf{fixed}
\item Finally, we will analyze \textbf{all combinations}: all selected M-SSA outputs for all selected $\delta$
\item Note: we do not display the M-MSE outputs which are noisier (more `false alarms'), see the ip-case for reference
\end{itemize}
\item For ease of visual inspection we apply \textbf{standardization} (equivalent to filter calibration)
\end{itemize}
\end{frame}

\begin{frame} {Look Ahead: Effect of $\delta$ on Advancement (Left Shift) of M-SSA-BIP }
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4in]{look_ahead_ssa_1.pdf}\caption{Effect of look ahead delta on  left-shift/advancement of M-SSA\label{cor}}\end{center}\end{figure}\end{frame}

\begin{frame} {Financial Crisis (left) and Pandemic (right)}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4in]{look_ahead_ssa_2.pdf}\caption{Effect of look ahead delta on  left-shift/advancement of M-SSA\label{cor}}\end{center}\end{figure}\end{frame}

\begin{frame} {Findings}
\begin{itemize}
\item The main effect of $\delta=3, 6, 12, 18, 24$ on \textbf{M-SSA-BIP} is a \textbf{left-shift} of the filter output 
\begin{itemize}
\item A larger $\delta$  means that zero-crossings at onset and end of crises are detected earlier(advancement, left-shift)
\item Note that M-SSA keeps the (expected) HT fixed for increasing $\delta$ (in contrast to M-MSE which becomes noisier, see slides of ip-case)
\item The filters are also subject to zero-shrinkage for increasing $\delta$ (but this effect is masked by the standardization) 
\end{itemize}
\item We report \textbf{sample HTs} for M-MSE and M-SSA filters as a function of $\delta$: see the table on the next slide

\end{itemize}

\end{frame}

\begin{frame} {Holding-Times}\label{htzc2}

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Sat Feb 15 08:41:28 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
  \hline
 & delta=0 & delta=3 & delta=6 & delta=12 & delta=18 \\ 
  \hline
M-MSE & 9.20 & 9.80 & 8.70 & 7.60 & 6.90 \\ 
  M-SSA & 16.00 & 13.20 & 13.80 & 16.90 & 19.00 \\ 
   \hline
\end{tabular}
\caption{Holding times of M-MSE and M-SSA for various values of look ahead parameter $\delta$} 
\label{perf_var1}
\end{table}\begin{itemize}
\item The above HTs of M-SSA should be compared with the univariate SSA on slide \eqref{htzc1}, first row, last column: by construction expected numbers should match (sample HTs are subject to random deviations)

\end{itemize}

\end{frame}

\begin{frame} {Findings}
\begin{itemize}
\item M-SSA is substantially smoother (less crossings, i.e., less false alarms)
\item The \emph{expected} HT is independent of $\delta$ 
\begin{itemize}
\item The differences in the above \emph{sample} HTs (of M-SSA) reveal mainly \emph{finite sample} variations
\item The \emph{sample} HTs converge to the  \emph{fixed expected} value in sufficiently long samples (see SSA tutorial)
\end{itemize}

%\item We now compare graphically M-MSE and M-SSA for $\delta=12$: zero-crossings are marked by corresponding colored vertical lines

\end{itemize}

\end{frame}

%\begin{frame} {Holding-Times: $\delta=24$}
%<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
%file = "ht_look_ahead.pdf"
%cat("\\begin{figure}[H]")
%cat("\\begin{center}")
%cat("\\includegraphics[height=3in, width=4in]{", file, "}\n",sep = "")
%cat("\\caption{Effect of look ahead delta on  left-shift/advancement of MSE", sep = "")
%cat("\\label{cor}}", sep = "")
%cat("\\end{center}")
%cat("\\end{figure}")
%@
%\end{frame}

\begin{frame} {Effect of $\delta$ on Target Correlation and Sign Accuracy}\label{sample_target_cor3}
\begin{itemize}
\item We analyze the effect of $\delta$ on the target correlation and the sign accuracy, see the following table. 
\end{itemize}

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Sat Feb 15 08:41:28 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrr}
  \hline
 & Target correlation & Sign accuracy \\ 
  \hline
delta=0 & 0.17 & 0.62 \\ 
  delta=3 & 0.22 & 0.63 \\ 
  delta=6 & 0.27 & 0.64 \\ 
  delta=12 & 0.38 & 0.68 \\ 
  delta=18 & 0.56 & 0.77 \\ 
  delta=24 & 0.69 & 0.77 \\ 
   \hline
\end{tabular}
\caption{Target correlations (first column) and sign accuracy (second column) of M-SSA-BIP for various $\delta\geq 0$ values } 
\label{perf_var1}
\end{table}\begin{itemize}
\item Note: the target correlation for $\delta=0$ (first row, first column) corresponds to slide \eqref{sample_target_cor2} (first row, second column)
\end{itemize}

\end{frame}

\begin{frame} {Effect of $\delta$ on Filter Weights}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4in]{look_ahead_ssa_6.pdf}\caption{Effect of look ahead delta on  left-shift/advancement of M-SSA\label{cor}}\end{center}\end{figure}\end{frame}

\begin{frame} {Same as Previous Slide but MA-Inversion of Filter}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4in]{look_ahead_ssa_7.pdf}\caption{Effect of look ahead delta on  left-shift/advancement of M-SSA\label{cor}}\end{center}\end{figure}\end{frame}

\begin{frame} {Findings}
\begin{itemize}
\item The original M-SSA-BIP, based on $\delta=0$, was lagging (due in part to VARMA model misspecification)
\item Increasing $\delta$ leads to substantial improvements of \emph{target correlation} and \emph{sign accuracy}
\item Increasing $\delta$ signifies a \textbf{left-shift} of the filter outputs: this left-shift \emph{explains} mainly the improvement  
\item Increasing $\delta$ induces a zero-shrinkage of the filter weights and a modification of the general decay-pattern, see the plots just above
\begin{itemize}
\item The shrinkage of the filter coefficients does not affect \emph{standardized} filter outputs
\item The different decay pattern affects the left-shift of the (standardized) filter outputs; but it does not affect the (expected) HT
\end{itemize}
\end{itemize}
\end{frame}

\subsection{All M-SSA Outputs and Fixed $\delta=12$}

\begin{frame} {Variable (All) M-SSA Targets and Fixed Forecast Excess}
\begin{itemize}
\item In the above slides we showed the effect of $\delta$ on the M-SSA-BIP output 
\item We now consider \textbf{all} M-SSA outputs for \textbf{fixed} $\delta=12$ (all M-SSA outputs for $\delta=0$ were analyzed previously)
\end{itemize}

\end{frame}

\begin{frame} {All M-SSA Outputs for Fixed $\delta=12$}
\begin{itemize}
\item In contrast to the previous slides, we now \textbf{fix} $\delta:=12$ and we analyze \textbf{all} M-SSA outputs for that $\delta$
\item On the next slide we now display and compare the corresponding look-ahead M-SSA outputs for $\delta=12$  
\item For reference, we also include BIP-trend, i.e., the acausal HP applied to BIP-shifted (black line)
\item For easier visual inspection all series are standardized (equivalent to filter calibration)
\end{itemize}

\end{frame}

\begin{frame} {All M-SSA Outputs for (Fixed) $\delta=12$}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4in]{look_ahead_ssa_3.pdf}\caption{Effect of look ahead delta on  left-shift/advancement of M-SSA\label{cor}}\end{center}\end{figure}\end{frame}

\begin{frame} {Findings}
\begin{itemize}
\item M-SSA outputs for ESI (red), ifo-exp (green), ip (cyan) and BIP (violet) differ mainly during the financial crisis, where especially the former two are left-shifted and faster 
\begin{itemize}
\item The financial crisis cannot be reconciled with the VARMA-model (misspecification) which explains the lesser performance of the BIP M-SSA output in all previous comparisons (details omitted).
\end{itemize}
\item Towards the sample end, M-SSA ifo-exp (and to some extent M-SSA ESI) suggest evidence of a  \textbf{recovery} whereas M-SSA ip and BIP still indicate negative growth.

\end{itemize}

\end{frame}

\begin{frame} {Sign Accuracies}\label{signaac1}
\begin{itemize}
\item In order to complete our evaluation metrics, the table below reports \textbf{sign accuracies} of the M-SSA outputs, i.e., the (sample) probability that a predictor matches the sign of the target (BIP-trend, black line in previous plot)
\end{itemize}

% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Sat Feb 15 08:41:28 2025
\begin{table}[ht]
\centering
\begin{tabular}{rlllll}
  \hline
 & BIP & ip & ifo\_exp & ESI & spr\_10y\_3m \\ 
  \hline
sign acc. & 0.68(0.03) & 0.68(0.03) & 0.65(0.03) & 0.74(0.03) & 0.39(0.03) \\ 
   \hline
\end{tabular}
\caption{Sign accuracies of M-SSA outputs for fixed $\delta=12$ with standard errors in parentheses} 
\label{perf_var1}
\end{table}\begin{itemize}
\item For fixed $\delta=12$,  ESI (red line in previous plot) is best, followed closely by ip, BIP and ifo-exp. The sign accuracy of spread is below $50\%$ 
\end{itemize}

\end{frame}

\subsection{All Combinations  of M-SSA Outputs and of Forecast Excess(es)}

\begin{frame} {All Combinations of Forecast Excess and M-SSA Outputs}
\begin{itemize}
\item Up yet we considered specific combinations of forecast excess and M-SSA outputs
\item The tables in the following three slides report sample \emph{sign-accuracy}, sample \emph{target correlation}  and sample \emph{holding-times} for \textbf{all combinations} of M-SSA outputs and $\delta$
\begin{itemize}
\item Sign-accuracy and sample correlation are evaluated against BIP-trend, i.e., acausal HP applied to BIP-shift
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame} {Sample Sign Accuracy}\label{signaac2}
% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Sat Feb 15 08:41:28 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
  \hline
 & BIP & ip & ifo\_exp & ESI & spr\_10y\_3m \\ 
  \hline
delta=0 & 0.62 & 0.64 & 0.68 & 0.74 & 0.43 \\ 
  delta=3 & 0.63 & 0.65 & 0.67 & 0.73 & 0.42 \\ 
  delta=6 & 0.64 & 0.68 & 0.67 & 0.74 & 0.42 \\ 
  delta=12 & 0.68 & 0.68 & 0.65 & 0.74 & 0.39 \\ 
  delta=18 & 0.77 & 0.74 & 0.61 & 0.67 & 0.36 \\ 
  delta=24 & 0.77 & 0.71 & 0.50 & 0.52 & 0.42 \\ 
   \hline
\end{tabular}
\caption{Sign accuracies of all combinations of M-SSA outputs (columns) and forecast excesses $\delta$ (rows): the sign accuracy is referenced against BIP-trend} 
\label{perf_var1}
\end{table}\end{frame}

\begin{frame} {Sample Target Correlation}\label{sample_target_cor4}
% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Sat Feb 15 08:41:28 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
  \hline
 & BIP & ip & ifo\_exp & ESI & spr\_10y\_3m \\ 
  \hline
delta=0 & 0.17 & 0.28 & 0.47 & 0.66 & -0.09 \\ 
  delta=3 & 0.22 & 0.31 & 0.45 & 0.65 & -0.10 \\ 
  delta=6 & 0.27 & 0.34 & 0.43 & 0.63 & -0.12 \\ 
  delta=12 & 0.38 & 0.41 & 0.38 & 0.58 & -0.18 \\ 
  delta=18 & 0.56 & 0.52 & 0.23 & 0.41 & -0.31 \\ 
  delta=24 & 0.69 & 0.54 & -0.10 & -0.02 & -0.40 \\ 
   \hline
\end{tabular}
\caption{Target correlations of all combinations of M-SSA outputs (columns) and forecast excesses $\delta$ (rows): the target correlation is referenced against BIP-trend} 
\label{perf_var1}
\end{table}\end{frame}

\begin{frame} {Sample HT}\label{htzc3}
% latex table generated in R 4.2.2 by xtable 1.8-4 package
% Sat Feb 15 08:41:29 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
  \hline
 & BIP & ip & ifo\_exp & ESI & spr\_10y\_3m \\ 
  \hline
delta=0 & 16.00 & 11.26 & 11.69 & 19.00 & 25.33 \\ 
  delta=3 & 13.22 & 11.26 & 12.67 & 21.71 & 25.33 \\ 
  delta=6 & 13.82 & 10.48 & 12.67 & 19.00 & 19.00 \\ 
  delta=12 & 16.89 & 12.67 & 11.69 & 19.00 & 21.71 \\ 
  delta=18 & 19.00 & 15.20 & 11.69 & 14.48 & 13.82 \\ 
   \hline
\end{tabular}
\caption{Sample holding times of all combinations of M-SSA outputs (columns) and forecast excesses $\delta$ (rows)} 
\label{perf_var1}
\end{table}\begin{itemize}
\item Remarks:
\begin{itemize}
\item Expected HTs change along columns (from left to right) but are fixed along rows (top to bottom)
\item Sample HTs in the table follow this pattern (though subject to random variation)
\end{itemize}
\end{itemize}

\end{frame}

\begin{frame} {Findings}
\begin{itemize}
\item M-SSA outputs of lagging series (such as BIP and ip: first two columns in above tables) perform better for larger $\delta$
\item M-SSA outputs of coincident series (ESI, ifo-exp, columns 3 and 4) perform best for small values of $\delta$
\begin{itemize}
\item M-SSA ifo-exp is systematically outperformed by M-SSA ESI
\end{itemize}
\item The above findings are intuitively appealing/interpretable
\item M-SSA output of spread is not a good predictor of the target
\item \textbf{Summary}: we can combine M-SSA \textbf{BIP} and \textbf{ip} with \textbf{large} $\delta$ and M-SSA \textbf{ESI} for \textbf{small} $\delta$
\end{itemize}
\end{frame}

\section{(Construction of a) Forward-Looking Consensus BIP-Predictor}

\frame{\sectionpage}

\begin{frame} {Findings}
\begin{itemize}
\item As claimed, we can combine M-SSA \textbf{BIP} and \textbf{ip} with \textbf{large} $\delta$ and M-SSA \textbf{ESI} for \textbf{small} $\delta$ to construct a forward-looking ($h=3$) BIP-predictor 
\begin{itemize}
\item The procedure cyn be extended to arbitrary $h$ (hyperparameter)
\end{itemize}
\item Given similar performances of the selected indicators we rely on an \emph{equally-weighted} `consensus' predictor
\begin{itemize}
\item Equal-weighting of \emph{standardized} predictors (standardization accounts for zero-shrinkage)
\end{itemize}
\item On the following two slides we display the selected (standardized) indicators and their cross-sectional mean,i.e., the \textbf{forward-looking consensus BIP-predictor}
\end{itemize}
\end{frame}

\begin{frame} {BIP-Trend (black) and Selected Predictors}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4in]{bip_predictor1.pdf}\caption{Effect of look ahead delta on  left-shift/advancement of M-SSA\label{cor}}\end{center}\end{figure}\end{frame}

\begin{frame} {BIP-Trend (black) and Forward-Looking Consensus BIP-Predictor}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4in]{bip_predictor2.pdf}\caption{Effect of look ahead delta on  left-shift/advancement of M-SSA\label{cor}}\end{center}\end{figure}\end{frame}

\begin{frame} {Calibrated Forward-Looking Consensus BIP-Predictor}
\begin{itemize}
\item We can also calibrate the predictor on level and scale of BIP-returns by simple linear regression, see next slide
\item We display
\begin{itemize}
\item \textbf{Quarterly BIP} returns shifted forward by $h=3$ (black): interpolated with zero for months 2 and 3
\item \textbf{Two-sided HP} applied to shifted BIP (violet). We multiply HP by 3 to match the \emph{quarterly growth} scale of BIP
\item \textbf{Calibrated consensus} indicator with long-term mean growth (horizontal blue line) and \textbf{mean-crossings} (vertical blue lines)
\end{itemize}
\item Duration between mean-crossings is HT=13.82 ($\approx$mean of HTs of selected indicators)
\end{itemize}
\end{frame}

\begin{frame} {Calibrated Forward-Looking Consensus BIP-Predictor}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4in]{bip_predictor3.pdf}\caption{Effect of look ahead delta on  left-shift/advancement of M-SSA\label{cor}}\end{center}\end{figure}\end{frame}

\begin{frame} {Sensitivity to Estimation Sample }
\begin{itemize}
\item Sources of revisions of the `Calibrated Forward-Looking Consensus BIP-Predictor' are due to calibration (mean and level),  to the VAR-model and to the data (BIP/ip)
\item On the following slide we briefly analyze the effect of the first two source (calibration and estimation sample)
\item Specifically, we recompute the VAR and recalibrate the resulting predictor based on full data (NOv-2024), data prior to 2015 and data prior to 2010
\end{itemize}
\end{frame}

\begin{frame} {Sensitivity to Estimation Sample}
\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4in]{bip_predictor_sensitivity.pdf}\caption{Effect of look ahead delta on  left-shift/advancement of M-SSA\label{cor}}\end{center}\end{figure}\end{frame}

\section{Conclusion}

\frame{\sectionpage}

\begin{frame} {Conclusion}
\begin{itemize}
\item Models and performances are generally reliant on the data sample (for estimation or evaluation): `Pandemic' effect
\item Trend targets can be interpreted meaningfully in terms of economic dynamics (recessions/expansions)
\item Trend targets are easier to forecast (less noisy), see slide \eqref{slide_target_cor}
\item Multivariate designs outperform mainly with respect to trend targets and lagged series (BIP and ip), see slide \eqref{rfmse1}
\item Target correlations, sign accuracies and HT are potentially interesting alternative performance metrics
\item Target correlations and sign accuracies referenced against BIP-filter are strongly significant (not random) even on full sample (including singular Pandemic), see slides  \eqref{sample_target_cor1}, \eqref{sample_target_cor2}, \eqref{sample_target_cor3} and \eqref{sample_target_cor4}  and \eqref{signaac1}  (exception: filtered spread)
\end{itemize}
\end{frame}

\begin{frame} {Conclusion}
\begin{itemize}
\item Filters (uni and multivariate) are similar to direct AR forecasts in terms of forecast mean-square error with respect to BIP-shifted, although they do not fit the target explicitly (no overfitting), see slides \eqref{tstat1} and \eqref{slide_multi_rel_mse}
\item Real-time indicators such as {ESI} and {ifo} (any of the selected ifo series) provide additional information due to \emph{smaller publication lags} (left shifted) 
\item {Spread} is \textbf{ambiguous}:
\begin{itemize}
\item It does not appear to correlate systematically  with BIP, see CCFs on slide \eqref{ccf_pand_multe}
\item The target correlation of the corresponding M-SSA output with BIP-trend is slightly negative, see slide \eqref{sample_target_cor2}: this outcome is confirmed by the plot on slide \eqref{spr_mssa_o}
\item The t-statistics on slide \eqref{tstat1} suggest that the univariate filter based on spread tracks BIP-shifted (not BIP-trend) significantly
\item But the same univariate filter is insignificant when regressed on BIP-trend (value of the t-statistic is 1.3)
\end{itemize}
\item Models and results/outcomes are mostly interpretable (appealing)
\end{itemize}

\end{frame}

\begin{frame} {Conclusion}
\begin{itemize}
\item (M-)SSA can {control} the terms of the forecast trilemma and address user/forecast {priorities} 
\begin{itemize}
\item The {hyperparameters} ($\delta,HT)$ (HT-constraint of M-SSA) allow to {fine-tune} priorities
\item M-SSA {generalizes} classic MSE-paradigm
\end{itemize}
\item Possibility to {combine} various predictors: 
\begin{itemize}
\item {Fast} but noisier: early warning, {look ahead} indicator 
\item A bit slower but more {reliable}: nowcast, {flash} indicator 
\item Very {smooth}/reliable but retarded: {confirmatory} indicator
\end{itemize}
\item Forecast combinations: direct AR with uni- and multivariate filters
\end{itemize}

\end{frame}

\begin{frame} {Summary of (and Links to) Sample Performance Metrics}
\begin{itemize}
\item Relative \textbf{forecast mean-square} error (target: BIP-shifted): slide \eqref{rmse1} (direct AR forecasts full sample), \eqref{rmse2} (same but shorter sample), \eqref{rmse3} (univariate filters, full sample), \eqref{rmse4} (same but before Pandemic), \eqref{slide_multi_rel_mse} (multivariate filters)

\item Relative \textbf{filter mean-square} error (target: BIP-trend): \eqref{rfmse1} (outperformance of multivariate over univariate)

\item Target \textbf{correlations}: \eqref{sample_target_cor1}, \eqref{sample_target_cor2}, \eqref{sample_target_cor3} and \eqref{sample_target_cor4}

\item \textbf{Sign accuracies}: \eqref{signaac1} and \eqref{signaac2}

\item \textbf{t-statistics}: \eqref{tstat1},  (univariate filters and direct AR, full sample), \eqref{tstat2} (same but prior pandemic), \eqref{slide_multi_rel_mse} (multivariate filters)

\item \textbf{Holding times}: \eqref{htzc1} (direct AR and univariate filters), \eqref{htzc2} and \eqref{htzc3} (M-SSA vs. M-MSE)
\end{itemize}
\end{frame}

\begin{frame} {Design Settings (Hyperparameters) in this Study}
\begin{itemize}
\item Data: selected indicators
\item Estimation sample (for models): full data vs. ante Pandemic
\item Evaluation sample (forecast errors): full data vs. ante Pandemic
\item (V)AR(MA) model orders: ARMA(3,1), VAR(2)
\item Trend target: HP(14400)
\item (M-)SSA settings: increase of HT by $50\%$, forecast excess $\delta=0, 3, 6, 12, 18, 24$
\item Forecast horizon: $h=3$
\end{itemize}
\end{frame}

\begin{frame} {Design Settings (Hyperparameters) in this Study}
\begin{itemize}
\item Random-walk or linear interpolation of BIP: default is random walk (na.locf in read function)
\item Use two-sided HP as filter target when evaluating sample mean-square filter errors (=TRUE)
\begin{itemize}
\item One can rely on two-sided HP or on symmetrized HP-C: default is two-sided HP (T)
\end{itemize}
\item Use two-sided HP (or one-sided HP-C) as (M-)SSA/(M-)MSE targets in (M-)SSA optimization (=TRUE)
\begin{itemize}
\item One can rely on two-sided HP or on one-sided HP-C: default is two-sided HP (T)
\item One-sided HP-C is a potentially interesting target (HP-C has some interesting characteristics as a causal filter)
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame} {Sources of Revisions}
\begin{itemize}
\item Data
\item Models: (V)ARMA
\item Calibration: `static' level and scale parameters
\end{itemize}
\end{frame}

\begin{frame} {To Dos}
\begin{itemize}
\item Mixed-frequency approach for BIP
\item Use all series (extension to dimension reduction and/or aggregation techniques)
\item \textbf{Better VAR(MA)}-model
\begin{itemize}
\item The weight assigned to `faster' indicators (small or no publication lag) in multi-step ahead forecasts is too weak
\end{itemize}
\item Additional forecast horizons: $h=0$ (nowcast), $h=12$ one-year forecast. 
\item Alternative holding-time constraints (smoothness)
\item Alternative filter specification(s): stronger suppression of `noisy cycles' of ifo-exp in previous plots?
\item `More' (forecast) combinations
\item Portfolio (of predictors) construction: emphasize fast/smooth tradeoff (instead of classic return/risk approach)
\item R-package, automatic up-dating with new data,...
\end{itemize}
\end{frame}

\begin{frame} {Next Steps}
\begin{itemize}
\item Worth pursuing? 
\begin{itemize}
\item Refresh (latest data)
\end{itemize}
\item Conceptual: \textbf{priorities} 
\begin{itemize}
\item Target, forecast horizon, `consensus' vs. separate predictors, ...
\end{itemize}
\item \textbf{Method}: VAR, mixed-frequency, dimension reduction,...
\item Work with \textbf{R-code} (cross-check, ROC,...): {Github}?
\item \textbf{Paper} 
\item \textbf{Application} (e.g. shiny app, public domain,...)
\end{itemize}
\end{frame}

\end{document}

### 2025-08-08 Archive IWH Simon paper/23.06.2025/paper.tex
up

--- 2025-06-21 HWI SSA_27_05_2025_with new poster/22.06.2025/paper.tex
@@
     \item $\mathbf{X}_t \equiv [1,u_{t-1}, ip_t, ifo\_c_t, ESI_t, spr\_10y\_3m_t]'$ or their transformed counterparts and $u_{t-1}$ accounts for the publication lag of GDP.\footnote{Including lagged GDP enables the nesting of the AR forecast within the broader modeling framework.}
-    \item $\hat{\beta}$ is estimated by least-squares, omitting observations affected by the pandemic, i.e. Q4-2019 to Q4-2020.
+    \item $\hat{\boldsymbol{\beta}}$ is estimated by least-squares, omitting observations affected by the pandemic, i.e. Q4-2019 to Q4-2020.
 \end{itemize}
@@
 
-Figures \ref{direct_wc} and \ref{direct_wc_financial_crisis} compare actual and fitted values for GDP at forecast horizons $h = 0, \ldots, 3$, using the four indicators and the target, i.e.
-GDP shifted forward (leftwards) by $h$ quarters, without applying any filtering.
+Figures \ref{direct_wc} and \ref{direct_wc_financial_crisis} compare actual and fitted values for GDP at forecast horizons $h = 0, \ldots, 3$, using the four indicators and the target, i.e., GDP shifted leftwards by $h$ quarters, without applying any filtering.
 Figure \ref{direct_wc} shows results over the sample without the pandemic, while Figure \ref{direct_wc_financial_crisis} provides a closer look at the fit during the financial crisis.
@@
 The fitted values lag GDP slightly less and track turning points somewhat more quickly.
-Table \ref{tab:pvaluedhp} shows p-values for tests of the joint null hypothesis that all coefficients $\hat{\beta} = 0$.
+Table \ref{tab:pvaluedhp} shows p-values for tests of the joint null hypothesis that all coefficients ${\boldsymbol{\beta}} = 0$.
 The use of filtered indicators improves their statistical significance at forecast horizons $h > 2$ quarters.
@@
 \end{tabular}
-\caption{p-values for $H_0: \hat{\beta} = 0$, HAC-adjusted Wald test.\\Estimation over full sample (without pandemic.)} 
+\caption{p-values for $H_0: \boldsymbol{\beta} = 0$, HAC-adjusted Wald test.\\Estimation over full sample (without pandemic.)} 
 \label{tab:pvaluedhp}
@@
  \cite{Wildi2025} shows that the SSA criterion helps control this phenomenon,  see Fig.\ref{mssa_msse_zc} for illustration.
-Using the equivalent dual formulation of the criterion \eqref{critssaht}, the author demonstrates that the SSA solution exhibits the lowest zero-crossing rate of any (linear) predictor with the same target correlation.
-In the following, we therefore use the M-SSA framework to construct a `smooth' predictor $y_t$ to forecast $z_t$.
+Using the dual formulation of the criterion \eqref{critssaht}, the author demonstrates that the SSA solution yields the lowest zero-crossing rate among all (linear) predictors with equivalent target correlation.
+Consequently, the subsequent analysis employs the M-SSA framework to construct a `smooth' predictor $y_t$ for forecasting $z_{t+h}$.
+In particular, it is shown that controlling the zero-crossing rate within the M-SSA framework has a quantifiable effect on the hit rate and false positive rate.
+
  %In a second stage, serving as a generalization and refinement of the univariate HP-C utilized in the direct HP forecast.
@@
 The results validate that the smoothing constraint incorporated into the multivariate extension of criterion \eqref{critssa} effectively manages the sign-changes of the predictor: sample estimates of the holding times align with the expected values derived from Eq.~\eqref{ht}.\footnote{Effective convergence of sample estimates to their theoretical expected values —assuming the true data-generating process is known— can be validated using the M-SSA package.
-This validation is typically performed through analyses of very long samples of simulated (multivariate) data, ensuring the consistency of the estimation procedures under idealized conditions.} An increase of $50\%$ in the HT relative to the M-MSE predictor has only a modest effect on the target correlation, suggesting that this trade-off could be beneficial in our application.
+This validation is typically performed through analyses of very long samples of simulated (multivariate) data, ensuring the consistency of the estimation procedures under idealized conditions.} An increase of $50\%$ in the HT relative to the M-MSE predictor has only a modest effect on the target correlation, suggesting that this trade-off could be beneficial in our application, as confirmed by a subsequent analysis of hit rate and false alarm rate.
 %Going forward, we now omit the M-MSE, as it is a special case of the more general M-SSA framework.
@@
 Note that statistical significance is influenced by the relatively short out-of-sample span, of length 17 years.}.
-The MSE forecast performance of $\hat{\textrm{GDP}}_{t+shift}(h)$ can be benchmarked against the mean predictor (see Table \ref{rRMSE_mSSA_comp_mean_wc}) as well as against the direct forecast (see Table \ref{rRMSE_mSSA_comp_direct_without_covid7}).
+The MSE forecast performance of $\hat{\textrm{GDP}}_{t+shift}(h)$ can be benchmarked against the mean predictor (see Table \ref{rRMSE_mSSA_comp_mean_wc} which reports relative Root Mean-Square Errors, rRMSE) as well as against the direct forecast (see Table \ref{rRMSE_mSSA_comp_direct_without_covid7}).
 For consistency, both benchmark predictors are also quarterly updated out-of-sample designs.\\

### 2025-08-08 Latex/BIP_predictor.tex
up

@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Fri Jun 20 07:05:56 2025
+% Fri Aug  8 09:40:14 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Fri Jun 20 07:05:56 2025
+% Fri Aug  8 09:40:14 2025
 \begin{table}[ht]
@@
 \ref{impulse} (full data set) and \eqref{impulse_wc} (prior the pandemic).
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/ma_inv_multi_ip.pdf}\caption{MA inversion of VAR(1): full data set.\label{impulse}}\end{center}\end{figure}
-
-\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/ma_inv_multi_ip_wc.pdf}\caption{MA inversion of VAR(1): without Pandemic.\label{impulse_wc}}\end{center}\end{figure}
+\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/ma_inv_multi_ip.pdf}\caption{MA inversion of VAR(1): full data set.\label{impulse}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/ma_inv_multi_ip_wc.pdf}\caption{MA inversion of VAR(1): without Pandemic.\label{impulse_wc}}\end{center}\end{figure}\begin{figure}[H]\begin{center}\includegraphics[height=3in, width=6in]{./Figures/ma_inv_BIP.pdf}\caption{MA inversion of VAR(1) for GDP: with pandemic (left) and without pandemic (right).\label{ma_inv_BIP}}\end{center}\end{figure}
 
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Fri Jun 20 07:05:58 2025
+% Fri Aug  8 09:40:16 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Fri Jun 20 07:06:08 2025
+% Fri Aug  8 09:40:24 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Fri Jun 20 07:06:51 2025
+% Fri Aug  8 09:41:11 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Fri Jun 20 07:07:00 2025
+% Fri Aug  8 09:41:21 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Fri Jun 20 07:07:00 2025
+% Fri Aug  8 09:41:21 2025
 \begin{table}[ht]
@@
 % latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Fri Jun 20 07:07:25 2025
+% Fri Aug  8 09:41:51 2025
 \begin{table}[ht]
@@
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Fri Jun 20 07:07:25 2025
+% Fri Aug  8 09:41:51 2025
 \begin{table}[ht]

### 2025-08-08 Sweave/BIP_predictor.Rnw
up

@@
 hp_c<-HP_obj$hp_trend
-plot(hp_c,main=paste("Two-sided HP(",lambda_HP,")",sep=""),axes=F,type="l",xlab="Lags",ylab="")
+plot(hp_c,main=paste("(One-sided) HP-C",sep=""),axes=F,type="l",xlab="Lags",ylab="")
 abline(h=0)

### 2025-08-08 Archive IWH Simon paper/09.08.2025/Abstract.tex
up

\documentclass{article}
\usepackage{graphicx} % Required for inserting images
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{amsfonts} 
\usepackage{amsthm}
\usepackage{ amssymb }

\title{Smooth and Persistent Forecasts of German GDP}
\author{Katja Heinisch (IWH)\\Simon van Norden (HEC Montréal)\\Marc Wildi (ZHAW)}
\date{February 2025}

\begin{document}

\maketitle

Economic forecasters face a dilemma. Efficiency suggests that their forecasts make use of all available information. The resulting forecasts are often volatile, which may reflect greater information content or contamination by excessive ``noise''. Forecast users (including public and private decision makers) may be constrained by how rapidly they can adjust to changing forecasts.\footnote{For example, private or public spending decisions may be constrained by annual budgets. Similarly, central banks increasingly attempt to smooth interest rate changes to allow them to signal the future direction of policy to markets, while private firms attempt to smooth changes in dividends.}

\cite{Wildi2024,Wildi2025} and \cite{McElroy2019,McElroy2020} propose novel methods that optimally smooth forecasts by controlling the expected frequency of sign changes. The resulting Smooth Sign Accuracy (SSA) framework encompasses efficient mean-square error (MSE) as a special case and produces increasingly smooth forecasts as changes in forecast direction are increasingly penalized.

In this study, we apply the SSA framework to document the ability of key German economic indicators (such as industrial production, ifo business surveys and the yield curve) to provide meaningful and accurate forecasts of GDP and GDP trends.
\cite{Drechsel2012financial,Heinisch2018bottom} have shown that these indicators show a good forecast performance.
We provide results on forecast performance using a variety of sample periods and performance metrics. Surprisingly, we find that smoothing indicator-based forecasts often provides useful leading information for GDP. We then show the degree to which smoother GDP forecasts lag efficient MSE forecasts slightly.\footnote{This reflects the trilemma highlighted by \cite{McElroy2019} in which forecasters choose between forecast accuracy, timeliness and volatility.} Finally, we document the degree to which this can be mitigated by adjustments in the forecast horizon.

Among our results to date, we find that industrial production and economic outlook surveys provide much more valuable information than the yield curve. We also find that extreme events, such as the Great Financial Crisis or the COVID-19 pandemic, have an important impact on the forecasts. This reflects a common tradeoff facing forecasters: the choice between high accuracy in over the whole period and avoiding rare but large and costly forecast errors in extraordinary situations.

\end{document}

### 2025-08-08 Archive IWH Simon paper/09.08.2025/paper.tex
up

--- 2025-08-08 Archive IWH Simon paper/08.08.2025/paper.tex
@@
 &\mathbf{b}'\mathbf{b}=l
-\end{array}\right\}.
+\end{array}\right\},
 \end{eqnarray}
-where the additional length constraint $\mathbf{b}'\mathbf{b}=l$ ensures uniqueness of the SSA solution. 
+where the additional length constraint $\mathbf{b}'\mathbf{b}=l$ serves two purposes, namely to warrant uniqueness of the SSA solution and to ensure that the objective function $\mathbf{b}'\boldsymbol{\gamma}_{\delta}$ is identical to the target correlation, up to an arbitrary scaling term $\sqrt{\mathbf{b}'\mathbf{b}}\sqrt{\sum \gamma_k^2}=\sqrt{l}\sqrt{\sum \gamma_k^2}$.
 By orthogonal projection, the MSE predictor is $y_{MSE,t}=\sum_{k=0}^{L-1}\gamma_kx_{t-k}$, i.e., $b_{MSE,k}=\gamma_k$, $k=0,...,L-1$ and in general we select $\rho_1>\rho_{MSE}$ in criterion l{crit1} to improve smoothness of the resulting SSA predictor $y_t$, where $\rho_{MSE}$ designates the first order ACF of the (benchmark) MSE predictor\footnote{Selecting $\rho_1<\rho_{MSE}$ would result in a SSA predictor with more frequent zero-crossings than the MSE benchmark, which is undesirable in typical applications.}.

### 2025-08-08 Archive IWH Simon paper/09.08.2025/paper.tex
up

@@
 For a better understanding, we here briefly substantiate the optimization criterion \ref{critssa}.
-To start, assume a simple univariate framework with $x_t=\epsilon_t$ white noise and $z_t=\sum_{k=-infty}^\infty \gamma_k x_{t-k}=\sum_{k=-infty}^\infty \gamma_k \epsilon_{t-k}$, where it is assumed that $\gamma_k\neq 0$ for some $k<0$ (acausal filter) and $\sum_{|k|<\infty}\gamma_k^2<\infty$, which ensures stationarity of $z_t$ (extensions to non-stationary integrated processes are discussed in Wildi (2025)).
+To start, assume a simple univariate framework with $x_t=\epsilon_t$ white noise, with variance $\sigma^2$, and $z_t=\sum_{k=-\infty}^\infty \gamma_k x_{t-k}=\sum_{k=-\infty}^\infty \gamma_k \epsilon_{t-k}$, where it is assumed that $\gamma_k\neq 0$ for some $k<0$ (acausal filter) and $\sum_{|k|<\infty}\gamma_k^2<\infty$, ensuring stationarity of $z_t$ (extensions to non-stationary integrated processes are discussed in Wildi (2025)).
 Assume we want to predict $z_{t+\delta}$, $\delta\geq 0$ and let $y_{\delta t}=\sum_{k=0}^{L-1} b_{\delta k}x_{t-k}$ designate a causal predictor of length $L$.
 For simplicity of exposition we may assume $\delta=0$ (nowcast) and we may drop $\delta$ from our notation.
-Then, $\rho(z,y)=\frac{\sum_{k=0}^{L-1}b_k\gamma_k}{\sqrt{\sum_{k=0}^{L-1}b_k^2}\sqrt{\sum_{k=0}^{L-1}\gamma_k^2}}$  and $\rho(y,1)=\frac{\mathbf{b}'\mathbf{Mb}}{\mathbf{b}'\mathbf{b}}$, where
+Then, $\rho(z,y)=\frac{\sum_{k=0}^{L-1}b_k\gamma_k}{\sqrt{\sum_{k=0}^{L-1}b_k^2}\sqrt{\sum_{|k|<\infty}\gamma_k^2}}$  and $\rho(y,1)=\frac{\mathbf{b}'\mathbf{Mb}}{\mathbf{b}'\mathbf{b}}$, where
 \[
@@
 \]
-ofs dimension $L\times L$ is such that $\mathbf{b}'\mathbf{Mb}=\sum_{k=1}^{L-1}b_{k-1}b_k$ corresponds to the order-one autocovariance of $y_t$, up to scaling by the variance $\sigma^2$ of $\epsilon_t$.
+of dimension $L\times L$, is such that $\mathbf{b}'\mathbf{Mb}=\sum_{k=1}^{L-1}b_{k-1}b_k$, the first-order autocovariance of $y_t$ (up to scaling by $\sigma^2$).
 We can now rewrite criterion \ref{critssa} as:
@@
 where the additional length constraint $\mathbf{b}'\mathbf{b}=l$ serves two purposes, namely to warrant uniqueness of the SSA solution and to ensure that the objective function $\mathbf{b}'\boldsymbol{\gamma}_{\delta}$ is identical to the target correlation, up to an arbitrary scaling term $\sqrt{\mathbf{b}'\mathbf{b}}\sqrt{\sum \gamma_k^2}=\sqrt{l}\sqrt{\sum \gamma_k^2}$.
+Criterion \ref{crit1} is easily amenable to optimization, since it involves affine and quadratic functions of $\mathbf{b}$, see Wildi (2025) for details.
 By orthogonal projection, the MSE predictor is $y_{MSE,t}=\sum_{k=0}^{L-1}\gamma_kx_{t-k}$, i.e., $b_{MSE,k}=\gamma_k$, $k=0,...,L-1$ and in general we select $\rho_1>\rho_{MSE}$ in criterion l{crit1} to improve smoothness of the resulting SSA predictor $y_t$, where $\rho_{MSE}$ designates the first order ACF of the (benchmark) MSE predictor\footnote{Selecting $\rho_1<\rho_{MSE}$ would result in a SSA predictor with more frequent zero-crossings than the MSE benchmark, which is undesirable in typical applications.}.

### 2025-08-08 Archive IWH Simon paper/09.08.2025/paper.tex
up

--- 2025-08-08 Archive IWH Simon paper/08.08.2025/paper.tex
@@
 
-\subsection{MSSA Optimization Criterion}
-
-Let $\mathbf{X}_t$ (of dimension $t\times n$) denote a set of $n$ explanatory series $\mathbf{x}_{1t},...,\mathbf{x}_{nt}$, with observations $\mathbf{x}_{it}=(x_{i1},...,x_{it})'$.
-Let $\mathbf{z}_t=(z_{1},...,z_t)$  denote a target series, which typically lies outside the linear space spanned by $\mathbf{X}_t$  (and is unknown at time $t$).
+\subsection{SSA Optimization Criterion}
+
+Let $\mathbf{x}_t$ be a time series representing the data, with observations $(x_{1},...,x_{t})'$.
+Let $\mathbf{z}_t=(z_{1},...,z_t)$  denote a target series, which depends on $\mathbf{x}_t$ as well as on unknown future $x_{t+k}$, $k>0$.
 For simplicity, we assume stationarity of all time series involved.
-In this framework, $\mathbf{X}_t$ corresponds to a matrix of selected economic indicators with $n=5$, and $z_{it}$, i=1,...,5,   is the output of a two-sided HP(160) filter applied to the four indicators as well as to GDP.
-Since $z_{it}$  depends on future observations $x_{it+1},x_{it+2},...$, the prediction task involves `tracking' $z_{it}$  using the estimate $y_{it}=\sum_{j=1}^ny_{ijt}$, where $y_{ijt}=\mathbf{b}_{ij}~'\mathbf{x}_{jt}=\sum_{k=0}^{L} b_{ijk}x_{j,t-k}$ are the outputs of a multivariate (causal or one-sided) filter $\mathbf{B}_i^f$, with columns $\mathbf{b}_{ij}$ of fixed length $L$, $j=1,...,5$, assigning weights to the last $L$ observations of the $j$-th indicator $\mathbf{x}_{jt}$ (the fixed-length assumption is merely used for ease of exposition).
-For the sake of clarity, we now omit the index $i$ for all variables, assuming $z_t=z_{i_0t}$ for some fixed $i_0$.% Additionally, we assume a fixed filter length $L$, such that $\mathbf{B}^i_t=\mathbf{B}_t=\mathbf{B}$, which has dimension $L\times 5$.
+In this framework, we might think of ${z}_t$ as corresponding to the output of a two-sided HP(160) filter applied to the doubly-infinite sequence $\mathbf{x}_{\pm \infty}=(x_k)$, $|k|<\infty$, representing the (log differenced) GDP, for which $\mathbf{x}_t$ is effectively observed.
+Since $z_{t}$  depends on future observations $x_{t+1},x_{t+2},...$\footnote{We here ignore the dependence of $z_t$ on $x_0,x_{-1},...$ since in general the filter weights $\gamma_k$ of the target decay sufficiently fast to zero.}, the prediction task involves `tracking' $z_{t}$  using the (causal) predictor $y_{t}=\sum_{k=0}^{L-1} b_{jk}x_{t-k}$, where $L\leq t$.
 
@@
  \cite{Wildi2025} shows that the SSA criterion helps control this phenomenon (see Fig.\ref{mssa_msse_zc} for illustration.) Using the dual formulation of the criterion \eqref{critssaht}, the author demonstrates that the SSA solution yields the lowest zero-crossing rate among all (linear) predictors with equivalent target correlation.
-The concept can be extended to a multivariate framework, denoted M-SSA, by generalizing the correlation functions as detailed hereafter.
-
-\subsection{Implementation}
-
-For a better understanding, we here briefly substantiate the optimization criterion \ref{critssa}.
-To start, assume a simple univariate framework with $x_t=\epsilon_t$ white noise and $z_t=\sum_{k=-infty}^\infty \gamma_k x_{t-k}=\sum_{k=-infty}^\infty \gamma_k \epsilon_{t-k}$, where it is assumed that $\gamma_k\neq 0$ for some $k<0$ (acausal filter) and $\sum_{|k|<\infty}\gamma_k^2<\infty$, which ensures stationarity of $z_t$ (extensions to non-stationary integrated processes are discussed in Wildi (2025)).
-Assume we want to predict $z_{t+\delta}$, $\delta\geq 0$ and let $y_{\delta t}=\sum_{k=0}^{L-1} b_{\delta k}x_{t-k}$ designate a causal predictor of length $L$.
-For simplicity of exposition we may assume $\delta=0$ (nowcast) and we may drop $\delta$ from our notation.
-Then, $\rho(z,y)=\frac{\sum_{k=0}^{L-1}b_k\gamma_k}{\sqrt{\sum_{k=0}^{L-1}b_k^2}\sqrt{\sum_{k=0}^{L-1}\gamma_k^2}}$  and $\rho(y,1)=\frac{\mathbf{b}'\mathbf{Mb}}{\mathbf{b}'\mathbf{b}}$, where
+The concept can be extended to a multivariate framework, denoted M-SSA, by generalizing the correlation functions as detailed in the next section.\\
+
+We now briefly elucidate the SSA optimization criterion outlined in Equation \ref{critssa}.
+Consider a simplified univariate framework where $x_t=\epsilon_t$ represents white noise with variance $\sigma^2$.
+A generic target specification for $z_t$ is then defined as a linear filter of $x_t$:
+\[z_t=\sum_{k=-\infty}^\infty \gamma_k x_{t-k}=\sum_{k=-\infty}^\infty \gamma_k \epsilon_{t-k},\] 
+where it is assumed that $\gamma_k\neq 0$ for some $k<0$ (implying an acausal filter), and the summation of squared coefficients satisfies $\sum_{|k|<\infty}\gamma_k^2<\infty$, ensuring the stationarity of the target $z_t$.
+With these definitions, all processes are centered at zero and the Gaussian assumption additionally justifies the use of Equation \ref{critssaht}, linking first-order ACF and HT.
+Extensions to non-stationary integrated processes are discussed in Wildi (2025).
+Suppose the goal is to predict  $z_{t+\delta}$, $\delta\geq 0$.
+Let  $y_{\delta t}=\sum_{k=0}^{L-1} b_{\delta k}x_{t-k}$  denote a causal predictor of length $L$.
+To simplify the exposition, we consider the case $\delta=0$ (nowcast), and thus omit $\delta$ from the notation.
+Then, under the postulated white assumption
+\begin{eqnarray*}
+\rho(z,y)&=&\frac{\sum_{k=0}^{L-1}b_k\gamma_k}{\sqrt{\sum_{k=0}^{L-1}b_k^2}\sqrt{\sum_{|k|<\infty}\gamma_k^2}},\\
+\rho(y,1)&=&\frac{\mathbf{b}'\mathbf{Mb}}{\mathbf{b}'\mathbf{b}},
+\end{eqnarray*} 
+where the so-called autocovariance generating matrix $\mathbf{M}$, of dimension $L\times L$, is defined as:
 \[
@@
 0&0&0&0&0&...&0&0.5&0
-\end{array}\right),
+\end{array}\right).
 \]
-ofs dimension $L\times L$ is such that $\mathbf{b}'\mathbf{Mb}=\sum_{k=1}^{L-1}b_{k-1}b_k$ corresponds to the order-one autocovariance of $y_t$, up to scaling by the variance $\sigma^2$ of $\epsilon_t$.
-We can now rewrite criterion \ref{critssa} as:
+The matrix $\mathbf{M}$ is such that $\mathbf{b}'\mathbf{Mb}=\sum_{k=1}^{L-1}b_{k-1}b_k$, representing the first-order autocovariance of $y_t$ (up to the scaling factor $\sigma^2$).
+Using this, the optimization criterion \ref{critssa} can be reformulated as:
 \begin{eqnarray}\label{crit1}
@@
 &\max_{\mathbf{b}}\mathbf{b}'\boldsymbol{\gamma}_{\delta}\\
-&\mathbf{b}'\mathbf{Mb}=l\rho_1\\
+\textrm{subject~to~}&\mathbf{b}'\mathbf{Mb}=l\rho_1\\
 &\mathbf{b}'\mathbf{b}=l
-\end{array}\right\}.
+\end{array}\right\},
 \end{eqnarray}
-where the additional length constraint $\mathbf{b}'\mathbf{b}=l$ ensures uniqueness of the SSA solution. 
-By orthogonal projection, the MSE predictor is $y_{MSE,t}=\sum_{k=0}^{L-1}\gamma_kx_{t-k}$, i.e., $b_{MSE,k}=\gamma_k$, $k=0,...,L-1$ and in general we select $\rho_1>\rho_{MSE}$ in criterion l{crit1} to improve smoothness of the resulting SSA predictor $y_t$, where $\rho_{MSE}$ designates the first order ACF of the (benchmark) MSE predictor\footnote{Selecting $\rho_1<\rho_{MSE}$ would result in a SSA predictor with more frequent zero-crossings than the MSE benchmark, which is undesirable in typical applications.}.
+where the additional length constraint $\mathbf{b}'\mathbf{b}=l$ serves two purposes: firstly, to ensure the uniqueness of the SSA solution; secondly, to normalize the objective function $\mathbf{b}'\boldsymbol{\gamma}_{\delta}$, making it proportional to the target correlation, up to a fixed scaling factor $\sqrt{\mathbf{b}'\mathbf{b}}\sqrt{\sum \gamma_k^2}=\sqrt{l}\sqrt{\sum \gamma_k^2}$.
+Criterion \ref{crit1} is readily amenable to optimization, as it involves affine and quadratic functions of $\mathbf{b}$, see Wildi (2025) for detailed derivations.
 \\
-
-We can extend the SSA criterion \ref{crit1} to a multivariate framework.
-Assume first that $\mathbf{X}_t=\boldsymbol{\epsilon}_t$ is a multivariate white noise sequence with (full rank) variance-covariance matrix $\boldsymbol{\Sigma}$ and let $\mathbf{z}_t$:
+Under the above assumptions, the classical MSE predictor can be expressed via orthogonal projection as:
+\[
+y_{MSE,t}=\sum_{k=0}^{L-1}\gamma_kx_{t-k}
+\]
+which corresponds to setting $b_{MSE,k}=\gamma_k$, $k=0,...,L-1$.
+In general, to enhance the smoothness of the SSA predictor $y_t$, we select $\rho_1>\rho_{MSE}$ in criterion \ref{crit1},  where $\rho_{MSE}$ denotes the first-order ACF of the (benchmark) MSE predictor\footnote{Selecting $\rho_1<\rho_{MSE}$ would produce an SSA predictor with more frequent zero-crossings than the MSE benchmark, which is typically undesirable in practical applications.}.
+
+\subsection{M-SSA Optimization}
+
+We can now extend the SSA criterion \ref{crit1} to a multivariate setting.
+Let $\mathbf{X}_t$ (of dimension $t\times n$) denote a set of $n$ explanatory series $\mathbf{x}_{1t},...,\mathbf{x}_{nt}$, with observations $\mathbf{x}_{it}=(x_{i1},...,x_{it})'$.
+Let $\mathbf{z}_t=(z_{1},...,z_t)$  denote a target series, which typically lies outside the linear space spanned by $\mathbf{X}_t$  (and is unknown at time $t$).
+In this framework, $\mathbf{X}_t$ corresponds to a matrix of selected economic indicators with $n=5$, and $z_{it}$, i=1,...,5,   is the output of a two-sided HP(160) filter applied to the four indicators as well as to GDP.
+%Since $z_{it}$  depends on future observations $x_{it+1},x_{it+2},...$, the prediction task involves `tracking' $z_{it}$  using the estimate $y_{it}=\sum_{j=1}^ny_{ijt}$, where $y_{ijt}=\mathbf{b}_{ij}~'\mathbf{x}_{jt}=\sum_{k=0}^{L} b_{ijk}x_{j,t-k}$ are the outputs of a multivariate (causal or one-sided) filter $\mathbf{B}_i^f$, with columns $\mathbf{b}_{ij}$ of fixed length $L$, $j=1,...,5$, assigning weights to the last $L$ observations of the $j$-th indicator $\mathbf{x}_{jt}$ (the fixed-length assumption is merely used for ease of exposition).
+For the sake of clarity, we now omit the index $i$ for all variables, assuming $z_t=z_{i_0t}$ for some fixed $i_0$.
+
+To introduce the M-SSA, assume that $\mathbf{X}_t=\boldsymbol{\epsilon}_t$ is an $n$-dim white noise sequence with (full rank) variance-covariance matrix $\boldsymbol{\Sigma}$.
+Consider the generalized target $\mathbf{z}_t$:
 \[
@@
 \] 
-where $\boldsymbol{\Gamma}_k$ is a sequence of square-summable $n\times n$ dimensional filter matrices, with entries $(\gamma_{ijk})$ for $i, j = 1, \ldots, n$.
-The condition for square summability is expressed mathematically as:
-\[
-\sum_{|k| < \infty} \text{tr}(\boldsymbol{\Gamma}_k' \boldsymbol{\Gamma}_k) < \infty,
-\]
-where $\text{tr}(\cdot)$ denotes the trace operator. We focus on the estimation of $\mathbf{z}_{t+\delta}$, utilizing the $n-$dimensional predictor defined by
+where $\boldsymbol{\Gamma}_k$  is a sequence of filter matrices of dimension  $n\times n$, satisfying the square-summability condition.
+Each $\boldsymbol{\Gamma}_k$ has entries $(\gamma_{ijk})$ for $i, j = 1, \ldots, n$.
+We focus on the estimation of $\mathbf{z}_{t+\delta}$, utilizing the $n-$dimensional predictor defined by
 \[
@@
 \]
-where $\boldsymbol{\gamma}_{i k}:=(\gamma_{i1k},...,\gamma_{ink})'$. Consequntly, we propose the following multivariate M-SSA criterion for $i=1,...,n$: 
+where $\boldsymbol{\gamma}_{i k}:=(\gamma_{i1k},...,\gamma_{ink})'$. Consequently, we propose the following multivariate M-SSA criterion for $i=1,...,n$: 
 \begin{eqnarray}\label{mcrit1}
 \textrm{max}_{\mathbf{b}_i}\boldsymbol{\gamma}_{i\cdot\delta}'\tilde{\mathbf{I}}\mathbf{b}_{i}\\
-\mathbf{b}_{i}'\tilde{\mathbf{M}}\mathbf{b}_{i}=\rho_i\nonumber\\
+\textrm{subject~to~}\mathbf{b}_{i}'\tilde{\mathbf{M}}\mathbf{b}_{i}=\rho_i\nonumber\\
 \mathbf{b}_{i}'\tilde{\mathbf{I}}\mathbf{b}_{i}=1,\nonumber
@@
 for $i=1,...,n$. 
-Under the imposed length constraint, the objective function is proportional to the correlation between the predictor and the $i$-th target, or equivalently, the $i$-th MSE predictor.
-Additionally, the HT constraint regulates the first-order ACF or, equivalently, the HT of the predictor.
-The relationships established between sign accuracy and target correlation, as well as between first-order ACF and HT, as discussed in the previous section, remain valid.
-Thus, the M-SSA framework can be reformulated in accordance with any alternative and equivalent expressions of the optimization criterion, as reviewed in the previous section.
-\\
-
-To conclude, we now relax the WN hypothesis and assume a stationary process 
+Under the imposed length constraint, the objective function is proportional to the correlation between the predictor and the $i$-th target.
+Furthermore, the criterion relies on affine and quadratic functions of $\mathbf{b}_i$, facilitating the numerical optimization.
+Finally, the HT constraint regulates the first-order ACF or, equivalently, the HT of the predictor.\\
+
+Finally, we can relax the WN hypothesis and assume a stationary process 
 \begin{equation}\label{x_m_st}
@@
 \max_{(\mathbf{b}\cdot\boldsymbol{\xi})_i} (\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{i\delta}'\mathbf{\tilde{I}} (\mathbf{b}\cdot\boldsymbol{\xi})_i\\
-(\mathbf{b}\cdot\boldsymbol{\xi})_i'\mathbf{\tilde{M}}(\mathbf{b}\cdot\boldsymbol{\xi})_i=\rho_i\nonumber\\
+\textrm{subject~to~}(\mathbf{b}\cdot\boldsymbol{\xi})_i'\mathbf{\tilde{M}}(\mathbf{b}\cdot\boldsymbol{\xi})_i=\rho_i\nonumber\\
 (\mathbf{b}\cdot\boldsymbol{\xi})_i'\mathbf{\tilde{I}}(\mathbf{b}\cdot\boldsymbol{\xi})_i=1\nonumber
@@
 for $i=1,...,n$.
-This formulation can be solved for $(\mathbf{b}\cdot\boldsymbol{\xi})_i$, as detailed in Theorem \eqref{lambda_mult} in the next section.
+In analogy to the above white noise case, this formulation can be solved for $(\mathbf{b}\cdot\boldsymbol{\xi})_i$.
 The M-SSA coefficients $\mathbf{b}_i$ can subsequently be determined through deconvolution, as indicated in Equation \eqref{con_inv}.
-Wildi (2025) posits an extension of SSA to non-stationary integrated processes, wherein the HT constraint pertains to stationary differences of the predictor.
-While a corresponding extension to M-SSA is feasible based on findings by McElroy and Wildi (2020), the present discussion is confined to stationary processes, omitting a lengthier treatment of rank-defficient designs (cointegration) in the non-stationary case.\\
-
-%\textcolor{blue}{This is where Marc outlines the generalization of the correlation function and other elements of M-SSA.}
+The present discussion is confined to stationary processes, omitting a lengthier treatment of rank-defficient designs (cointegration) in the non-stationary case.\\
 
@@
 In particular, it is shown that controlling the zero-crossing rate within the M-SSA framework has a quantifiable effect on the hit rate and false positive rate.
-
-%Lastly, sign changes in trend growth, represented by zero-crossings of the target $z_t$, are indicative of relevant changes in the trajectory of the economy, for example, at transitions between expansions and recessions.
-
- %In a second stage, serving as a generalization and refinement of the univariate HP-C utilized in the direct HP forecast.
 
@@
     \begin{center}
-    \textcolor{blue}{@Marc: Please change the plot Title to "M-MSE \& M-SSA Nowcasts: Financial Crisis".)}
+%    \textcolor{blue}{@Marc: Please change the plot Title to "M-MSE \& M-SSA Nowcasts: Financial Crisis".)}
         \includegraphics[height=3in, width=4.5in]{./Figures/mssa_msse_zc.pdf}

### 2025-08-08 Sweave/BIP_predictor.Rnw
up

@@
 # We can select type of VAR model
-VAR_type="BVAR"
+#VAR_type="BVAR"
 VAR_type="VAR"
-VAR_type="elastic_net"
+#VAR_type="elastic_net"
 if (VAR_type=="BVAR")
@@
 colo<-c("blue","green")
-main_title<-"Zero-Crossings of M-SSA and M-MSE towards the financial crisis"
+main_title<-"M-MSE & M-SSA Nowcasts: Financial Crisis"
 plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],lwd=1,ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))

### 2025-08-08 Archive IWH Simon paper/09.08.2025/paper.tex
up

@@
 In this framework, we might think of ${z}_t$ as corresponding to the output of a two-sided HP(160) filter applied to the doubly-infinite sequence $\mathbf{x}_{\pm \infty}=(x_k)$, $|k|<\infty$, representing the (log differenced) GDP, for which $\mathbf{x}_t$ is effectively observed.
-Since $z_{t}$  depends on future observations $x_{t+1},x_{t+2},...$\footnote{We here ignore the dependence of $z_t$ on $x_0,x_{-1},...$ since in general the filter weights $\gamma_k$ of the target decay sufficiently fast to zero.}, the prediction task involves `tracking' $z_{t}$  using the (causal) predictor $y_{t}=\sum_{k=0}^{L-1} b_{jk}x_{t-k}$, where $L\leq t$.
+Since $z_{t}$  depends on future observations $x_{t+1},x_{t+2},...$\footnote{We here ignore the dependence of $z_t$ on $x_0,x_{-1},...$ since in general the filter weights $\gamma_k$ of the target decay sufficiently fast to zero.}, the prediction task involves `tracking' $z_{t}$  using the (causal) predictor $y_{t}=\sum_{k=0}^{L-1} b_{k}x_{t-k}$, where $L\leq t$.
 
@@
 Extensions to non-stationary integrated processes are discussed in Wildi (2025).
-Suppose the goal is to predict  $z_{t+\delta}$, $\delta\geq 0$.
-Let  $y_{\delta t}=\sum_{k=0}^{L-1} b_{\delta k}x_{t-k}$  denote a causal predictor of length $L$.
-To simplify the exposition, we consider the case $\delta=0$ (nowcast), and thus omit $\delta$ from the notation.
-Then, under the postulated white assumption
+Suppose the goal is to predict  $z_{t+h}$, $h\geq 0$.
+Let  $y_{h t}=\sum_{k=0}^{L-1} b_{h k}x_{t-k}$  denote a causal predictor of length $L$.
+To simplify the exposition, we consider the case $h=0$ (nowcast), and thus omit the subscript $h$ from the notation.
+Then, under the postulated white noise assumption
 \begin{eqnarray*}
@@
 \left.\begin{array}{cc}
-&\max_{\mathbf{b}}\mathbf{b}'\boldsymbol{\gamma}_{\delta}\\
+&\max_{\mathbf{b}}\mathbf{b}'\boldsymbol{\gamma}_{h}\\
 \textrm{subject~to~}&\mathbf{b}'\mathbf{Mb}=l\rho_1\\
@@
 \end{eqnarray}
-where the additional length constraint $\mathbf{b}'\mathbf{b}=l$ serves two purposes: firstly, to ensure the uniqueness of the SSA solution; secondly, to normalize the objective function $\mathbf{b}'\boldsymbol{\gamma}_{\delta}$, making it proportional to the target correlation, up to a fixed scaling factor $\sqrt{\mathbf{b}'\mathbf{b}}\sqrt{\sum \gamma_k^2}=\sqrt{l}\sqrt{\sum \gamma_k^2}$.
+where the additional length constraint $\mathbf{b}'\mathbf{b}=l$ serves two purposes: firstly, to ensure the uniqueness of the SSA solution; secondly, to normalize the objective function $\mathbf{b}'\boldsymbol{\gamma}_{h}$, making it proportional to the target correlation, up to a fixed scaling factor $\sqrt{\mathbf{b}'\mathbf{b}}\sqrt{\sum \gamma_k^2}=\sqrt{l}\sqrt{\sum \gamma_k^2}$.
 Criterion \ref{crit1} is readily amenable to optimization, as it involves affine and quadratic functions of $\mathbf{b}$, see Wildi (2025) for detailed derivations.
@@
 Let $\mathbf{z}_t=(z_{1},...,z_t)$  denote a target series, which typically lies outside the linear space spanned by $\mathbf{X}_t$  (and is unknown at time $t$).
-In this framework, $\mathbf{X}_t$ corresponds to a matrix of selected economic indicators with $n=5$, and $z_{it}$, i=1,...,5,   is the output of a two-sided HP(160) filter applied to the four indicators as well as to GDP.
+In this framework, $\mathbf{X}_t$ corresponds to a matrix of selected economic indicators with $n=5$, and $z_{it}$, i=1,...,5,   are the outputs of a two-sided HP(160) filter applied to doubly-infinite extensions $(x_{ik})$, $|k|<\infty$, of the four indicators as well as to GDP.
 %Since $z_{it}$  depends on future observations $x_{it+1},x_{it+2},...$, the prediction task involves `tracking' $z_{it}$  using the estimate $y_{it}=\sum_{j=1}^ny_{ijt}$, where $y_{ijt}=\mathbf{b}_{ij}~'\mathbf{x}_{jt}=\sum_{k=0}^{L} b_{ijk}x_{j,t-k}$ are the outputs of a multivariate (causal or one-sided) filter $\mathbf{B}_i^f$, with columns $\mathbf{b}_{ij}$ of fixed length $L$, $j=1,...,5$, assigning weights to the last $L$ observations of the $j$-th indicator $\mathbf{x}_{jt}$ (the fixed-length assumption is merely used for ease of exposition).
-For the sake of clarity, we now omit the index $i$ for all variables, assuming $z_t=z_{i_0t}$ for some fixed $i_0$.
+%For the sake of clarity, we now omit the index $i$ for all variables, assuming $z_t=z_{i_0t}$ for some fixed $i_0$.
 
@@
 Each $\boldsymbol{\Gamma}_k$ has entries $(\gamma_{ijk})$ for $i, j = 1, \ldots, n$.
-We focus on the estimation of $\mathbf{z}_{t+\delta}$, utilizing the $n-$dimensional predictor defined by
+We focus on the estimation of $\mathbf{z}_{t+h}$, utilizing the $n-$dimensional predictor defined by
 \[
@@
 \boldsymbol{\epsilon}_{it}=(\epsilon_{it},\epsilon_{it-1},...,\epsilon_{it-(L-1)})'~&,&~\boldsymbol{\epsilon}_{\cdot t}=(\boldsymbol{\epsilon}_{1t}',...,\boldsymbol{\epsilon}_{nt}')'\\
-\boldsymbol{\gamma}_{ij\delta}=(\gamma_{ij\delta},\gamma_{ij\delta+1},...,\gamma_{ij\delta+L-1})'~&,&\boldsymbol{\gamma}_{i\cdot\delta}=(\boldsymbol{\gamma}_{i1\delta}',\boldsymbol{\gamma}_{i2\delta}',...,\boldsymbol{\gamma}_{in\delta}')'\\
+\boldsymbol{\gamma}_{ijh}=(\gamma_{ijh},\gamma_{ijh+1},...,\gamma_{ijh+L-1})'~&,&\boldsymbol{\gamma}_{i\cdot h}=(\boldsymbol{\gamma}_{i1h}',\boldsymbol{\gamma}_{i2h}',...,\boldsymbol{\gamma}_{inh}')'\\
 ~\mathbf{b}_{ij}=(b_{ij0},b_{ij1},...,b_{ijL-1})'~&,&~\mathbf{b}_{i}=(\mathbf{b}_{i1}',\mathbf{b}_{i2}',...,\mathbf{b}_{in}')'\\
-\boldsymbol{\gamma}_{\delta}=\left(\begin{array}{c}\boldsymbol{\gamma}_{1\cdot\delta}'\\
-\boldsymbol{\gamma}_{2\cdot\delta}'\\
-\cdot\\
-\boldsymbol{\gamma}_{n\cdot\delta}'\end{array}\right)~&,&~\mathbf{b}=\left(\begin{array}{c}\mathbf{b}_{1}'\\\mathbf{b}_{2}'\\\cdot\\\mathbf{b}_{n}'\end{array}\right).
+\boldsymbol{\gamma}_{h}=(\boldsymbol{\gamma}_{1\cdot h},\boldsymbol{\gamma}_{2\cdot h},...,\boldsymbol{\gamma}_{n\cdot h})
+%\boldsymbol{\gamma}_{h}=\left(\begin{array}{c}\boldsymbol{\gamma}_{1\cdot h}'\\
+%\boldsymbol{\gamma}_{2\cdot h}'\\
+%\cdot\\
+%\boldsymbol{\gamma}_{n\cdot h}'\end{array}\right)
+~&,&~\mathbf{b}=(\mathbf{b}_1,...,\mathbf{b}_n)%\mathbf{b}=\left(\begin{array}{c}\mathbf{b}_{1}'\\\mathbf{b}_{2}'\\\cdot\\\mathbf{b}_{n}'\end{array}\right).
 \end{eqnarray*} 
 We define $y_{ijt} := \mathbf{b}_{ij}'\boldsymbol{\epsilon}_{jt}$, which allows us to express $y_{it} = \mathbf{b}_i'\boldsymbol{\epsilon}_{\cdot t} = \sum_{j=1}^n y_{ijt}$ and $\mathbf{y}_t = \mathbf{b}'\boldsymbol{\epsilon}_{\cdot t}$.
-Through orthogonal projection, the MSE predictor is derived as $\hat{\mathbf{y}}_{t,MSE} = \boldsymbol{\gamma}_{\delta}'\boldsymbol{\epsilon}_{\cdot t}$, with components $\hat{y}_{it,MSE}$, $i=1,...,n$.
+Through orthogonal projection, the MSE predictor is derived as $\hat{\mathbf{y}}_{t,MSE} = \boldsymbol{\gamma}_{h}'\boldsymbol{\epsilon}_{\cdot t}$, with components $\hat{y}_{it,MSE}$, $i=1,...,n$.
 Consider the Kronecker products defined as $\tilde{\mathbf{I}}:=\boldsymbol{\Sigma}\otimes\mathbf{I}_{L\times L}$ and $\tilde{\mathbf{M}}:=\boldsymbol{\Sigma}\otimes\mathbf{M}$, of the $n\times n$ dimensional matrix $\boldsymbol{\Sigma}$, the $L\times L$ dimensional identity  $\mathbf{I}_{L\times L}$ and the autocovariance generating matrix $\mathbf{M}$.
-We can express the following expectations:
+Under the white noise assumption, we can express the following expectations:
 \begin{eqnarray}
-E[z_{it+\delta}y_{it}]&=&E[\hat{y}_{it,MSE}y_{it}]=\boldsymbol{\gamma}_{i\cdot\delta}'\tilde{\mathbf{I}}\mathbf{b}_{i}\nonumber\\%\label{moment1}\\
-E[\hat{z}_{it\delta}^2]&=&\boldsymbol{\gamma}_{i\cdot\delta}'\tilde{\mathbf{I}}\boldsymbol{\gamma}_{i\cdot\delta}\nonumber\\%\label{moment2}\\
+E[z_{it+h}y_{it}]&=&E[\hat{y}_{it,MSE}y_{it}]=\boldsymbol{\gamma}_{i\cdot h}'\tilde{\mathbf{I}}\mathbf{b}_{i}\nonumber\\%\label{moment1}\\
+E[\hat{z}_{ith}^2]&=&\boldsymbol{\gamma}_{i\cdot h}'\tilde{\mathbf{I}}\boldsymbol{\gamma}_{i\cdot h}\nonumber\\%\label{moment2}\\
 E[y_{it}^2]&=&\mathbf{b}_{i}'\tilde{\mathbf{I}}\mathbf{b}_{i}\nonumber\\%\label{moment3}\\
@@
 \begin{eqnarray}\label{mcrit1}
-\textrm{max}_{\mathbf{b}_i}\boldsymbol{\gamma}_{i\cdot\delta}'\tilde{\mathbf{I}}\mathbf{b}_{i}\\
+\textrm{max}_{\mathbf{b}_i}\boldsymbol{\gamma}_{i\cdot h}'\tilde{\mathbf{I}}\mathbf{b}_{i}\\
 \textrm{subject~to~}\mathbf{b}_{i}'\tilde{\mathbf{M}}\mathbf{b}_{i}=\rho_i\nonumber\\
@@
 \end{eqnarray}
-for $i=1,...,n$. 
-Under the imposed length constraint, the objective function is proportional to the correlation between the predictor and the $i$-th target.
+for $i=1,...,n$, where we assume an arbitrary normalization, $l=1$, in the length constraint.
+In analogy to the univariate case, the objective function is then proportional to the target correlation between the predictor and the $i$-th target.
 Furthermore, the criterion relies on affine and quadratic functions of $\mathbf{b}_i$, facilitating the numerical optimization.
@@
  \] 
-We now assume $\delta\geq 0$\footnote{Similar but slightly more complex expressions would arise for $\delta < 0$ (backcasting), which are omitted here.} and $L$ to be sufficiently large to render $(\boldsymbol{\Gamma} \cdot \boldsymbol{\Xi})_{\delta+k} \approx 0$ and $(\mathbf{B} \cdot \boldsymbol{\Xi})_j \approx 0$ for $k,j \geq L$, so that  their contributions to the variances of $\hat{\mathbf{y}}_{t,MSE}$ and $\mathbf{y}_t$ are negligible.
+We now assume $h\geq 0$\footnote{Similar but slightly more complex expressions would arise for $h < 0$ (backcasting), which are omitted here.} and $L$ to be sufficiently large to render $(\boldsymbol{\Gamma} \cdot \boldsymbol{\Xi})_{h+k} \approx 0$ and $(\mathbf{B} \cdot \boldsymbol{\Xi})_j \approx 0$ for $k,j \geq L$, so that  their contributions to the variances of $\hat{\mathbf{y}}_{t,MSE}$ and $\mathbf{y}_t$ are negligible.
 We define:
 \begin{eqnarray*}
-(\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{ij\delta}&=&\left((\boldsymbol{\Gamma}\cdot\boldsymbol{\Xi})_{ij\delta},(\boldsymbol{\Gamma}\cdot\boldsymbol{\Xi})_{ ij\delta+1}...,(\boldsymbol{\Gamma}\cdot\boldsymbol{\Xi})_{ij\delta+L-1}\right)'\\
-(\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{i\delta}&=&\Big((\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{i1\delta}',(\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{i2\delta}',...,(\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{in\delta}'\Big)'\\
+(\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{ijh}&=&\left((\boldsymbol{\Gamma}\cdot\boldsymbol{\Xi})_{ijh},(\boldsymbol{\Gamma}\cdot\boldsymbol{\Xi})_{ ijh+1}...,(\boldsymbol{\Gamma}\cdot\boldsymbol{\Xi})_{ijh+L-1}\right)'\\
+(\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{ih}&=&\Big((\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{i1h}',(\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{i2h}',...,(\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{inh}'\Big)'\\
 (\mathbf{b}\cdot\boldsymbol{\xi})_{ij}&=&\left((\mathbf{B}\cdot\boldsymbol{\Xi})_{ ij0},(\mathbf{B}\cdot\boldsymbol{\Xi})_{ij1},...,(\mathbf{B}\cdot\boldsymbol{\Xi})_{ijL-1}\right)'\\
@@
 \end{eqnarray*}
-It is important to note that if $\boldsymbol{\Xi}_k=\mathbf{0},~k>0$, resulting in $\mathbf{X}_t=\boldsymbol{\epsilon}_t$, then we have $(\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{i\delta}=\boldsymbol{\gamma}_{i\cdot\delta}$ and $(\mathbf{b}\cdot\boldsymbol{\xi})_{i}=\mathbf{b}_i$, as defined in the preceding section.
+It is important to note that if $\boldsymbol{\Xi}_k=\mathbf{0},~k>0$, resulting in $\mathbf{X}_t=\boldsymbol{\epsilon}_t$, then we have $(\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{ih}=\boldsymbol{\gamma}_{i\cdot h}$ and $(\mathbf{b}\cdot\boldsymbol{\xi})_{i}=\mathbf{b}_i$, as defined in the preceding section.
 The generalized M-SSA criterion can then be expressed as:
 \begin{eqnarray}\label{gen_stat_x}
-%\max_{(\mathbf{b}\cdot\boldsymbol{\xi})_i}\frac{ (\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{i\delta}'\mathbf{\tilde{I}} (\mathbf{b}\cdot\boldsymbol{\xi})_i}{\sqrt{(\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{i\delta}'\mathbf{\tilde{I}}(\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{i\delta}}\sqrt{(\mathbf{b}\cdot\boldsymbol{\xi})_i'\mathbf{\tilde{I}}(\mathbf{b}\cdot\boldsymbol{\xi})_i}}
-\max_{(\mathbf{b}\cdot\boldsymbol{\xi})_i} (\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{i\delta}'\mathbf{\tilde{I}} (\mathbf{b}\cdot\boldsymbol{\xi})_i\\
+%\max_{(\mathbf{b}\cdot\boldsymbol{\xi})_i}\frac{ (\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{ih}'\mathbf{\tilde{I}} (\mathbf{b}\cdot\boldsymbol{\xi})_i}{\sqrt{(\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{ih}'\mathbf{\tilde{I}}(\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{ih}}\sqrt{(\mathbf{b}\cdot\boldsymbol{\xi})_i'\mathbf{\tilde{I}}(\mathbf{b}\cdot\boldsymbol{\xi})_i}}
+\max_{(\mathbf{b}\cdot\boldsymbol{\xi})_i} (\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{ih}'\mathbf{\tilde{I}} (\mathbf{b}\cdot\boldsymbol{\xi})_i\\
 \textrm{subject~to~}(\mathbf{b}\cdot\boldsymbol{\xi})_i'\mathbf{\tilde{M}}(\mathbf{b}\cdot\boldsymbol{\xi})_i=\rho_i\nonumber\\
@@
   \hline
-Shift=0 & 0.975 & $0.941^*$ & $0.886^{*}$ &$ 0.839^{**}$ & $0.843^{**}$ & $0.897^{**}$ & 0.965 \\ 
-  Shift=1 & 1.085 & 1.068 & 1.011 &$ 0.927^{*}$ &$ 0.871^{**}$ & $0.875^{**}$ & $0.919^{**}$ \\ 
-  Shift=2 & 1.076 & 1.089 & 1.074 & 1.011 & $0.938^{*}$ & $0.902^{**}$ & $0.909^{**}$ \\ 
-  Shift=3 & 1.037 & 1.069 & 1.090 & 1.059 & 0.992 & $0.943^*$ & $0.930^*$ \\ 
-  Shift=4 & 1.013 & 1.058 & 1.101 & 1.097 & 1.037 & 0.967 & $0.925^*$ \\ 
+Shift=0 & 0.975 & $0.941^{*}$ & $0.886^{*}$ &$ 0.839^{**} $& $0.843^{**}$ &$ 0.897^{**} $& 0.965 \\ 
+  Shift=1 & 1.085 & 1.068 & 1.011 & $0.927^{*} $& $0.871^{**}$ & $0.875^{**}$ & $0.919^{**}$ \\ 
+  Shift=2 & 1.076 & 1.089 & 1.074 & 1.011 & $0.938^{*}$ &$ 0.902^{**} $& $0.909^{**} $\\ 
+  Shift=3 & 1.037 & 1.069 & 1.090 & 1.059 & 0.992 & $0.943^* $&$ 0.930^*$ \\ 
+  Shift=4 & 1.013 & 1.058 & 1.101 & 1.097 & 1.037 & 0.967 &$ 0.925^*$ \\ 
   Shift=5 & 1.016 & 1.062 & 1.106 & 1.115 & 1.078 & 1.015 & $0.955^*$ \\  

### 2025-08-08 Archive IWH Simon paper/09.08.2025/paper.tex
up

@@
 
-Let $\mathbf{x}_t$ be a time series representing the data, with observations $(x_{1},...,x_{t})'$.
-Let $\mathbf{z}_t=(z_{1},...,z_t)$  denote a target series, which depends on $\mathbf{x}_t$ as well as on unknown future $x_{t+k}$, $k>0$.
+Let $\mathbf{x}_t$ be a time series representing the data, with observations $(x_{t},...,x_{1})'$.
+Let $\mathbf{z}_t=(z_{t},...,z_1)'$  denote a target series, which depends on $\mathbf{x}_t$ as well as on unknown future $x_{t+k}$, $k>0$.
 For simplicity, we assume stationarity of all time series involved.
@@
 We can now extend the SSA criterion \ref{crit1} to a multivariate setting.
-Let $\mathbf{X}_t$ (of dimension $t\times n$) denote a set of $n$ explanatory series $\mathbf{x}_{1t},...,\mathbf{x}_{nt}$, with observations $\mathbf{x}_{it}=(x_{i1},...,x_{it})'$.
-Let $\mathbf{z}_t=(z_{1},...,z_t)$  denote a target series, which typically lies outside the linear space spanned by $\mathbf{X}_t$  (and is unknown at time $t$).
-In this framework, $\mathbf{X}_t$ corresponds to a matrix of selected economic indicators with $n=5$, and $z_{it}$, i=1,...,5,   are the outputs of a two-sided HP(160) filter applied to doubly-infinite extensions $(x_{ik})$, $|k|<\infty$, of the four indicators as well as to GDP.
+Let $\mathbf{X}_t$ (of dimension $t\times n$) denote a set of $n$ explanatory series $\mathbf{x}_{1t},...,\mathbf{x}_{nt}$, with observations $\mathbf{x}_{it}=(x_{it},...,x_{i1})'$.
+Let $\mathbf{z}_t=(z_{t},...,z_1)'$  denote a target series, which typically lies outside the linear space spanned by $\mathbf{X}_t$  (and is unknown at time $t$).
+In this framework, $\mathbf{X}_t$ corresponds to a matrix of selected economic indicators with $n=5$, and $z_{it}$, i=1,...,5,   are the outputs of a two-sided HP(160) filter applied to doubly-infinite extensions $(x_{ik})$, $|k|<\infty$, of the four indicators as well as of GDP.
 %Since $z_{it}$  depends on future observations $x_{it+1},x_{it+2},...$, the prediction task involves `tracking' $z_{it}$  using the estimate $y_{it}=\sum_{j=1}^ny_{ijt}$, where $y_{ijt}=\mathbf{b}_{ij}~'\mathbf{x}_{jt}=\sum_{k=0}^{L} b_{ijk}x_{j,t-k}$ are the outputs of a multivariate (causal or one-sided) filter $\mathbf{B}_i^f$, with columns $\mathbf{b}_{ij}$ of fixed length $L$, $j=1,...,5$, assigning weights to the last $L$ observations of the $j$-th indicator $\mathbf{x}_{jt}$ (the fixed-length assumption is merely used for ease of exposition).

### 2025-08-09 Archive IWH Simon paper/09.08.2025/paper.tex
Update paper.tex

@@
 Let $\mathbf{x}_t$ be a time series representing the data, with observations $(x_{t},...,x_{1})'$.
-Let $\mathbf{z}_t=(z_{t},...,z_1)'$  denote a target series, which depends on $\mathbf{x}_t$ as well as on unknown future $x_{t+k}$, $k>0$.
+Let $\mathbf{z}_t=(z_{t},...,z_1)'$  denote a target series, which depends on $\mathbf{x}_t$ as well as on unknown future $x_{t+k}$, $k>0$ (and possibly unknown past $x_k$, $k\leq 0$).
 For simplicity, we assume stationarity of all time series involved.
-In this framework, we might think of ${z}_t$ as corresponding to the output of a two-sided HP(160) filter applied to the doubly-infinite sequence $\mathbf{x}_{\pm \infty}=(x_k)$, $|k|<\infty$, representing the (log differenced) GDP, for which $\mathbf{x}_t$ is effectively observed.
-Since $z_{t}$  depends on future observations $x_{t+1},x_{t+2},...$\footnote{We here ignore the dependence of $z_t$ on $x_0,x_{-1},...$ since in general the filter weights $\gamma_k$ of the target decay sufficiently fast to zero.}, the prediction task involves `tracking' $z_{t}$  using the (causal) predictor $y_{t}=\sum_{k=0}^{L-1} b_{k}x_{t-k}$, where $L\leq t$.
+In this framework, we might think of ${z}_t$ as corresponding to the output of a two-sided HP(160) filter applied to the doubly-infinite sequence $\mathbf{x}_{\pm \infty}=(x_k)$, $|k|<\infty$, for which $\mathbf{x}_t$ is effectively observed, representing the (log differenced) GDP.
+Since $z_{t}$  depends on future observations $x_{t+1},x_{t+2},...$\footnote{We here ignore the dependence of $z_t$ on $x_0,x_{-1},...$ since the HP filter decays sufficiently fast to zero, see Fig.\ref{hp_160}.}, the prediction task involves `tracking' $z_{t}$  using the (causal) predictor $y_{t}=\sum_{k=0}^{L-1} b_{k}x_{t-k}$, where $L\leq t$.
 
@@
 where it is assumed that $\gamma_k\neq 0$ for some $k<0$ (implying an acausal filter), and the summation of squared coefficients satisfies $\sum_{|k|<\infty}\gamma_k^2<\infty$, ensuring the stationarity of the target $z_t$.
-With these definitions, all processes are centered at zero and the Gaussian assumption additionally justifies the use of Equation \ref{critssaht}, linking first-order ACF and HT.
+With these definitions, all processes are centered at zero, justifying the use of Equation \ref{critssaht} which links first-order ACF and HT.
 Extensions to non-stationary integrated processes are discussed in Wildi (2025).
@@
 \end{eqnarray}
-where the additional length constraint $\mathbf{b}'\mathbf{b}=l$ serves two purposes: firstly, to ensure the uniqueness of the SSA solution; secondly, to normalize the objective function $\mathbf{b}'\boldsymbol{\gamma}_{h}$, making it proportional to the target correlation, up to a fixed scaling factor $\sqrt{\mathbf{b}'\mathbf{b}}\sqrt{\sum \gamma_k^2}=\sqrt{l}\sqrt{\sum \gamma_k^2}$.
+where the additional length constraint $\mathbf{b}'\mathbf{b}=l$ serves two purposes: firstly, to ensure the uniqueness of the SSA solution; secondly, to normalize the objective function $\mathbf{b}'\boldsymbol{\gamma}_{h}$, making it proportional to the target correlation $\rho(z,y)$, up to a fixed scaling factor $1/\sqrt{\mathbf{b}'\mathbf{b}}\sqrt{\sum \gamma_k^2}=1/\sqrt{l}\sqrt{\sum \gamma_k^2}$.
 Criterion \ref{crit1} is readily amenable to optimization, as it involves affine and quadratic functions of $\mathbf{b}$, see Wildi (2025) for detailed derivations.
@@
 In general, to enhance the smoothness of the SSA predictor $y_t$, we select $\rho_1>\rho_{MSE}$ in criterion \ref{crit1},  where $\rho_{MSE}$ denotes the first-order ACF of the (benchmark) MSE predictor\footnote{Selecting $\rho_1<\rho_{MSE}$ would produce an SSA predictor with more frequent zero-crossings than the MSE benchmark, which is typically undesirable in practical applications.}.
+Finally,  an extension to autocorrelated $x_t\neq \epsilon_t$ will be discussed below.
 
@@
 Let $\mathbf{z}_t=(z_{t},...,z_1)'$  denote a target series, which typically lies outside the linear space spanned by $\mathbf{X}_t$  (and is unknown at time $t$).
-In this framework, $\mathbf{X}_t$ corresponds to a matrix of selected economic indicators with $n=5$, and $z_{it}$, i=1,...,5,   are the outputs of a two-sided HP(160) filter applied to doubly-infinite extensions $(x_{ik})$, $|k|<\infty$, of the four indicators as well as of GDP.
-%Since $z_{it}$  depends on future observations $x_{it+1},x_{it+2},...$, the prediction task involves `tracking' $z_{it}$  using the estimate $y_{it}=\sum_{j=1}^ny_{ijt}$, where $y_{ijt}=\mathbf{b}_{ij}~'\mathbf{x}_{jt}=\sum_{k=0}^{L} b_{ijk}x_{j,t-k}$ are the outputs of a multivariate (causal or one-sided) filter $\mathbf{B}_i^f$, with columns $\mathbf{b}_{ij}$ of fixed length $L$, $j=1,...,5$, assigning weights to the last $L$ observations of the $j$-th indicator $\mathbf{x}_{jt}$ (the fixed-length assumption is merely used for ease of exposition).
+In this framework, $\mathbf{X}_t$ corresponds to a matrix of selected economic indicators for the German economy, with $n=5$, and $z_{it}$, i=1,...,5,   are the outputs of a two-sided HP(160) filter applied to doubly-infinite extensions $(x_{ik})$, $|k|<\infty$, of the indicators, including GDP.
+In a later step, devised in Section \ref{sec:mSSA_component}, the (filtered) predictors $y_{it}(h)$ of $z_{it+h}$, with $h\geq 0$, will be used to setup up a predictor for GDP.%Since $z_{it}$  depends on future observations $x_{it+1},x_{it+2},...$, the prediction task involves `tracking' $z_{it}$  using the estimate $y_{it}=\sum_{j=1}^ny_{ijt}$, where $y_{ijt}=\mathbf{b}_{ij}~'\mathbf{x}_{jt}=\sum_{k=0}^{L} b_{ijk}x_{j,t-k}$ are the outputs of a multivariate (causal or one-sided) filter $\mathbf{B}_i^f$, with columns $\mathbf{b}_{ij}$ of fixed length $L$, $j=1,...,5$, assigning weights to the last $L$ observations of the $j$-th indicator $\mathbf{x}_{jt}$ (the fixed-length assumption is merely used for ease of exposition).
 %For the sake of clarity, we now omit the index $i$ for all variables, assuming $z_t=z_{i_0t}$ for some fixed $i_0$.
 
-To introduce the M-SSA, assume that $\mathbf{X}_t=\boldsymbol{\epsilon}_t$ is an $n$-dim white noise sequence with (full rank) variance-covariance matrix $\boldsymbol{\Sigma}$.
+To introduce the M-SSA framework, start by assuming that  $\mathbf{X}_t=\boldsymbol{\epsilon}_t$ is an $n$-dimensional white noise sequence with a full-rank variance-covariance matrix $\boldsymbol{\Sigma}$.
 Consider the generalized target $\mathbf{z}_t$:
@@
 %\boldsymbol{\gamma}_{n\cdot h}'\end{array}\right)
-~&,&~\mathbf{b}=(\mathbf{b}_1,...,\mathbf{b}_n)%\mathbf{b}=\left(\begin{array}{c}\mathbf{b}_{1}'\\\mathbf{b}_{2}'\\\cdot\\\mathbf{b}_{n}'\end{array}\right).
+~&,&~\mathbf{b}=(\mathbf{b}_1,...,\mathbf{b}_n).%\mathbf{b}=\left(\begin{array}{c}\mathbf{b}_{1}'\\\mathbf{b}_{2}'\\\cdot\\\mathbf{b}_{n}'\end{array}\right).
 \end{eqnarray*} 
-We define $y_{ijt} := \mathbf{b}_{ij}'\boldsymbol{\epsilon}_{jt}$, which allows us to express $y_{it} = \mathbf{b}_i'\boldsymbol{\epsilon}_{\cdot t} = \sum_{j=1}^n y_{ijt}$ and $\mathbf{y}_t = \mathbf{b}'\boldsymbol{\epsilon}_{\cdot t}$.
-Through orthogonal projection, the MSE predictor is derived as $\hat{\mathbf{y}}_{t,MSE} = \boldsymbol{\gamma}_{h}'\boldsymbol{\epsilon}_{\cdot t}$, with components $\hat{y}_{it,MSE}$, $i=1,...,n$.
+Here, $\boldsymbol{\gamma}_{h}$ and $\mathbf{b}$ are matrices of dimension $(L\cdot n)\times n$: the $i$-th column, $i=1,...,n$, collects the filter weights of the $i$-th target (or the $i$-th predictor) and the corresponding filter weights assigned to the $j=1,...,n$ series are stacked in the (long) column vector.
+The $i$-th column vector of $\boldsymbol{\gamma}_{h}$ (or $\mathbf{b}$) can then be applied to the $L\cdot n$-dimensional $\boldsymbol{\epsilon}_{\cdot t}$, the data.
+Specifically, we define $y_{ijt} := \mathbf{b}_{ij}'\boldsymbol{\epsilon}_{jt}$, which allows us to express $y_{it} = \mathbf{b}_i'\boldsymbol{\epsilon}_{\cdot t} = \sum_{j=1}^n y_{ijt}$ and $\mathbf{y}_t = \mathbf{b}'\boldsymbol{\epsilon}_{\cdot t}$.
+Through orthogonal projection, under the white noise assumption, the MSE predictor is derived as $\hat{\mathbf{y}}_{t,MSE} = \boldsymbol{\gamma}_{h}'\boldsymbol{\epsilon}_{\cdot t}$, with components $\hat{y}_{it,MSE}$, $i=1,...,n$.
 Consider the Kronecker products defined as $\tilde{\mathbf{I}}:=\boldsymbol{\Sigma}\otimes\mathbf{I}_{L\times L}$ and $\tilde{\mathbf{M}}:=\boldsymbol{\Sigma}\otimes\mathbf{M}$, of the $n\times n$ dimensional matrix $\boldsymbol{\Sigma}$, the $L\times L$ dimensional identity  $\mathbf{I}_{L\times L}$ and the autocovariance generating matrix $\mathbf{M}$.

### 2025-08-09 Archive IWH Simon paper/10.08.2025/paper.tex
up

--- 2025-08-09 Archive IWH Simon paper/09.08.2025/paper.tex
@@
 Let $\mathbf{X}_t$ (of dimension $t\times n$) denote a set of $n$ explanatory series $\mathbf{x}_{1t},...,\mathbf{x}_{nt}$, with observations $\mathbf{x}_{it}=(x_{it},...,x_{i1})'$.
-Let $\mathbf{z}_t=(z_{t},...,z_1)'$  denote a target series, which typically lies outside the linear space spanned by $\mathbf{X}_t$  (and is unknown at time $t$).
-In this framework, $\mathbf{X}_t$ corresponds to a matrix of selected economic indicators for the German economy, with $n=5$, and $z_{it}$, i=1,...,5,   are the outputs of a two-sided HP(160) filter applied to doubly-infinite extensions $(x_{ik})$, $|k|<\infty$, of the indicators, including GDP.
-In a later step, devised in Section \ref{sec:mSSA_component}, the (filtered) predictors $y_{it}(h)$ of $z_{it+h}$, with $h\geq 0$, will be used to setup up a predictor for GDP.%Since $z_{it}$  depends on future observations $x_{it+1},x_{it+2},...$, the prediction task involves `tracking' $z_{it}$  using the estimate $y_{it}=\sum_{j=1}^ny_{ijt}$, where $y_{ijt}=\mathbf{b}_{ij}~'\mathbf{x}_{jt}=\sum_{k=0}^{L} b_{ijk}x_{j,t-k}$ are the outputs of a multivariate (causal or one-sided) filter $\mathbf{B}_i^f$, with columns $\mathbf{b}_{ij}$ of fixed length $L$, $j=1,...,5$, assigning weights to the last $L$ observations of the $j$-th indicator $\mathbf{x}_{jt}$ (the fixed-length assumption is merely used for ease of exposition).
+Let $\mathbf{Z}_t$, of dimension $t\times m$ (in our case $m=n$) a set of $m$ target series $\mathbf{z}_{it}=(z_{it},...,z_{i1})'$, $i=1,...m$, which typically lie outside the linear space spanned by $\mathbf{X}_t$  (and are unknown at time $t$).
+In this framework, $\mathbf{X}_t$ represents a matrix of selected economic indicators for the German economy, with $n=5$.
+The targets $z_{it}$, i=1,...,5,   are obtained by applying a two-sided HP(160) filter to the doubly-infinite extensions $(x_{ik})$, $|k|<\infty$, of these indicators, including GDP.
+In a subsequent step, as delineated in Section \ref{sec:mSSA_component}, the (filtered) predictors $y_{hit}$ of $z_{it+h}$, with $h\geq 0$,  will be utilized to construct predictors for GDP.
+Unlike the approach in the previous section, these predictors will be based on multivariate M-SSA filters rather than univariate HP-C designs.
+%Since $z_{it}$  depends on future observations $x_{it+1},x_{it+2},...$, the prediction task involves `tracking' $z_{it}$  using the estimate $y_{it}=\sum_{j=1}^ny_{ijt}$, where $y_{ijt}=\mathbf{b}_{ij}~'\mathbf{x}_{jt}=\sum_{k=0}^{L} b_{ijk}x_{j,t-k}$ are the outputs of a multivariate (causal or one-sided) filter $\mathbf{B}_i^f$, with columns $\mathbf{b}_{ij}$ of fixed length $L$, $j=1,...,5$, assigning weights to the last $L$ observations of the $j$-th indicator $\mathbf{x}_{jt}$ (the fixed-length assumption is merely used for ease of exposition).
 %For the sake of clarity, we now omit the index $i$ for all variables, assuming $z_t=z_{i_0t}$ for some fixed $i_0$.
 
-To introduce the M-SSA framework, start by assuming that  $\mathbf{X}_t=\boldsymbol{\epsilon}_t$ is an $n$-dimensional white noise sequence with a full-rank variance-covariance matrix $\boldsymbol{\Sigma}$.
+To introduce the M-SSA framework, we begin by assuming that  $\mathbf{X}_t=\boldsymbol{\epsilon}_t$ is an $n$-dimensional white noise sequence with a full-rank variance-covariance matrix $\boldsymbol{\Sigma}$.
 Consider the generalized target $\mathbf{z}_t$:
@@
 \end{eqnarray*} 
-Here, $\boldsymbol{\gamma}_{h}$ and $\mathbf{b}$ are matrices of dimension $(L\cdot n)\times n$: the $i$-th column, $i=1,...,n$, collects the filter weights of the $i$-th target (or the $i$-th predictor) and the corresponding filter weights assigned to the $j=1,...,n$ series are stacked in the (long) column vector.
-The $i$-th column vector of $\boldsymbol{\gamma}_{h}$ (or $\mathbf{b}$) can then be applied to the $L\cdot n$-dimensional $\boldsymbol{\epsilon}_{\cdot t}$, the data.
-Specifically, we define $y_{ijt} := \mathbf{b}_{ij}'\boldsymbol{\epsilon}_{jt}$, which allows us to express $y_{it} = \mathbf{b}_i'\boldsymbol{\epsilon}_{\cdot t} = \sum_{j=1}^n y_{ijt}$ and $\mathbf{y}_t = \mathbf{b}'\boldsymbol{\epsilon}_{\cdot t}$.
-Through orthogonal projection, under the white noise assumption, the MSE predictor is derived as $\hat{\mathbf{y}}_{t,MSE} = \boldsymbol{\gamma}_{h}'\boldsymbol{\epsilon}_{\cdot t}$, with components $\hat{y}_{it,MSE}$, $i=1,...,n$.
-Consider the Kronecker products defined as $\tilde{\mathbf{I}}:=\boldsymbol{\Sigma}\otimes\mathbf{I}_{L\times L}$ and $\tilde{\mathbf{M}}:=\boldsymbol{\Sigma}\otimes\mathbf{M}$, of the $n\times n$ dimensional matrix $\boldsymbol{\Sigma}$, the $L\times L$ dimensional identity  $\mathbf{I}_{L\times L}$ and the autocovariance generating matrix $\mathbf{M}$.
+Here, $\boldsymbol{\gamma}_{h}$ (or $\mathbf{b}$) is a matrix of dimension $(L\cdot n)\times n$.
+The $i$-th column, $i=1,...,n$, contains the filter weights associated with the $i$-th target (or $i$-th predictor), while the filter weights corresponding to the $j=1,...,n$ series (indicators) are stacked into column vectors of length $L\cdot  n$.
+The $i$-th column vector of $\boldsymbol{\gamma}_{h}$ (or $\mathbf{b}$) can be applied to the $L\cdot n$-dimensional $\boldsymbol{\epsilon}_{\cdot t}$, which stacks all $n$ indicators into a single long data vector.
+Specifically, we define
+\[
+y_{ijt} := \mathbf{b}_{ij}'\boldsymbol{\epsilon}_{jt},
+\]
+which allows us to express the $i$-th predictor as:
+\[
+y_{it} = \mathbf{b}_i'\boldsymbol{\epsilon}_{\cdot t} = \sum_{j=1}^n y_{ijt}.
+\] 
+Similarly, the multivariate predictor vector can be written as:
+\[
+\mathbf{y}_t = \mathbf{b}'\boldsymbol{\epsilon}_{\cdot t}
+\]
+and the same applies to the target variables.
+Note that, for setting up the optimization criterion, it is not necessary to utilize the doubly infinite expressions of the target(s), as illustrated below.
+Under the white noise assumption, the orthogonal projection leads to the mean-squared error (MSE) predictor expressed as:
+\[
+\hat{\mathbf{y}}_{t,MSE} = \boldsymbol{\gamma}_{h}'\boldsymbol{\epsilon}_{\cdot t},
+\]
+whose components are $\hat{y}_{it,MSE}$, $i=1,...,n$. We define the Kronecker products: 
+\[
+\tilde{\mathbf{I}}:=\boldsymbol{\Sigma}\otimes\mathbf{I}_{L\times L},\textrm{ ~and~} \tilde{\mathbf{M}}:=\boldsymbol{\Sigma}\otimes\mathbf{M},
+\]
+where $\boldsymbol{\Sigma}$ is the variance-covariance matrix of the (white noise) data,  $\mathbf{I}_{L\times L}$ is the  $L\times L$ identity matrix  and  $\mathbf{M}$ is the autocovariance generating matrix introduced earlier.
 Under the white noise assumption, we can express the following expectations:
 \begin{eqnarray}
-E[z_{it+h}y_{it}]&=&E[\hat{y}_{it,MSE}y_{it}]=\boldsymbol{\gamma}_{i\cdot h}'\tilde{\mathbf{I}}\mathbf{b}_{i}\nonumber\\%\label{moment1}\\
-E[\hat{z}_{ith}^2]&=&\boldsymbol{\gamma}_{i\cdot h}'\tilde{\mathbf{I}}\boldsymbol{\gamma}_{i\cdot h}\nonumber\\%\label{moment2}\\
+E[z_{it+h}y_{it}]&=&\boldsymbol{\gamma}_{i\cdot h}'\tilde{\mathbf{I}}\mathbf{b}_{i}\nonumber\\%\label{moment1}\\
+%\label{moment2}\\
 E[y_{it}^2]&=&\mathbf{b}_{i}'\tilde{\mathbf{I}}\mathbf{b}_{i}\nonumber\\%\label{moment3}\\
-E[y_{it-1}y_{it}]&=&\mathbf{b}_{i}'\tilde{\mathbf{M}}\mathbf{b}_{i}.\nonumber
+E[y_{it-1}y_{it}]&=&\mathbf{b}_{i}'\tilde{\mathbf{M}}\mathbf{b}_{i},\nonumber
 \end{eqnarray}
-The variance of $z_{it}$ is given by 
-\[
-E[z_{it}^2]=\sum_{|k|<\infty}\boldsymbol{\gamma}_{i k}'\boldsymbol{\Sigma}\boldsymbol{\gamma}_{i k},
-\]
-where $\boldsymbol{\gamma}_{i k}:=(\gamma_{i1k},...,\gamma_{ink})'$. Consequently, we propose the following multivariate M-SSA criterion for $i=1,...,n$: 
+where, for notational simplicity, we assume $\sigma^2=1$. 
+%The variance of $z_{it}$ is given by 
+%\[
+%E[z_{it}^2]=\sum_{|k|<\infty}\boldsymbol{\gamma}_{i %k}'\boldsymbol{\Sigma}\boldsymbol{\gamma}_{i k},
+%\]
+%where $\boldsymbol{\gamma}_{i k}:=(\gamma_{i1k},...,\gamma_{ink})'$ is the weight assigned to the $n$ indicators at lag $k$. 
+Consequently, we propose the following multivariate M-SSA criterion for $i=1,...,n$: 
 \begin{eqnarray}\label{mcrit1}
@@
 \end{eqnarray}
-for $i=1,...,n$, where we assume an arbitrary normalization, $l=1$, in the length constraint.
-In analogy to the univariate case, the objective function is then proportional to the target correlation between the predictor and the $i$-th target.
-Furthermore, the criterion relies on affine and quadratic functions of $\mathbf{b}_i$, facilitating the numerical optimization.
-Finally, the HT constraint regulates the first-order ACF or, equivalently, the HT of the predictor.\\
+for $i=1,...,n$, where we assume an arbitrary normalization, $l=1$, in the length constraint $\mathbf{b}_{i}'\tilde{\mathbf{I}}\mathbf{b}_{i}=1$.
+If explicitly necessary, the arbitrary scaling $l=1$ can be revised and optimized separately at a later stage, after an optimal `unity-length' M-SSA predictor has been determined.
+In analogy to the univariate case, the objective function is proportional to the target correlation $\rho(z_i,y_i)$ between the $i$-th predictor and the $i$-th target.
+Moreover, the criterion involves affine and quadratic forms of $\mathbf{b}_i$, which simplifies the process of numerical optimization.
+Finally, the HT constraint $\mathbf{b}_{i}'\tilde{\mathbf{M}}\mathbf{b}_{i}=\rho_i$, where we used $l=1$, serves to control the first-order ACF or, equivalently, the HT of the predictor.
+\\
 
@@
 
- Consequently, the subsequent analysis employs the M-SSA framework to construct a `smooth' predictor $y_t$ for forecasting $z_{t+h}$.
+The subsequent analysis employs the M-SSA framework to construct `smooth' predictors for forecasting $z_{it+h}$ as well as for forecasting the original (unfiltered) GDP.
 In particular, it is shown that controlling the zero-crossing rate within the M-SSA framework has a quantifiable effect on the hit rate and false positive rate.
@@
 This validation is typically performed through analyses of very long samples of simulated (multivariate) data, ensuring the consistency of the estimation procedures under idealized conditions.} An
-increases by $50\%$ the HT relative to that of the M-MSE predictor but has only a modest effect on the target correlation, suggesting that this trade-off could be beneficial in our application
+increases by $50\%$ the HT relative to that of the M-MSE predictor\footnote{The convergence of the sample HT to the effectively imposed $HT_1$  can be verified in a simulation study, utilizing the M-SSA package.} but has only a modest effect on the target correlation, suggesting that this trade-off could be beneficial in our application
 % , as confirmed by a subsequent analysis of hit rate and false alarm rate

### 2025-08-09 Archive IWH Simon paper/10.08.2025/paper.tex
up

@@
 For simplicity, assume a fixed horizon, say $h=h_0$, so that we may drop the reference to $h$ in our notation.
-In the univariate case ($n=1$), \cite{Wildi2025} proposes the Smooth Sign Accuracy (SSA) as an optimization criterion:
+In the univariate case, \cite{Wildi2025} proposes the Smooth Sign Accuracy (SSA) as an optimization criterion:
 \begin{eqnarray}\label{critssa}
@@
 We now briefly elucidate the SSA optimization criterion outlined in Equation \ref{critssa}.
-Consider a simplified univariate framework where $x_t=\epsilon_t$ represents white noise with variance $\sigma^2$.
+Consider a simplified univariate framework where $x_t=\epsilon_t$ represents white noise with variance $\sigma^2$ and assume, for notational simplicity, that $\sigma^2=1$.
 A generic target specification for $z_t$ is then defined as a linear filter of $x_t$:
@@
 \]
-The matrix $\mathbf{M}$ is such that $\mathbf{b}'\mathbf{Mb}=\sum_{k=1}^{L-1}b_{k-1}b_k$, representing the first-order autocovariance of $y_t$ (up to the scaling factor $\sigma^2$).
+The matrix $\mathbf{M}$ is such that $\mathbf{b}'\mathbf{Mb}=\sum_{k=1}^{L-1}b_{k-1}b_k$, representing the first-order autocovariance of $y_t$.
 Using this, the optimization criterion \ref{critssa} can be reformulated as:
@@
 Let $\mathbf{X}_t$ (of dimension $t\times n$) denote a set of $n$ explanatory series $\mathbf{x}_{1t},...,\mathbf{x}_{nt}$, with observations $\mathbf{x}_{it}=(x_{it},...,x_{i1})'$.
-Let $\mathbf{Z}_t$, of dimension $t\times m$ (in our case $m=n$) a set of $m$ target series $\mathbf{z}_{it}=(z_{it},...,z_{i1})'$, $i=1,...m$, which typically lie outside the linear space spanned by $\mathbf{X}_t$  (and are unknown at time $t$).
+Let $\mathbf{Z}_t$, of dimension $t\times m$ (in our case $m=n$) a set of $m$ target series $\mathbf{z}_{it}=(z_{it},...,z_{i1})'$, $i=1,...m$, which typically lie outside the linear space spanned by   $\mathbf{X}_t$ (and are unknown at time $t$ but not independent of $\mathbf{X}_t$).
 In this framework, $\mathbf{X}_t$ represents a matrix of selected economic indicators for the German economy, with $n=5$.
@@
 In a subsequent step, as delineated in Section \ref{sec:mSSA_component}, the (filtered) predictors $y_{hit}$ of $z_{it+h}$, with $h\geq 0$,  will be utilized to construct predictors for GDP.
-Unlike the approach in the previous section, these predictors will be based on multivariate M-SSA filters rather than univariate HP-C designs.
+However, unlike the approach in the previous section, these predictors will be based on multivariate M-SSA filters rather than univariate HP-C designs.
 %Since $z_{it}$  depends on future observations $x_{it+1},x_{it+2},...$, the prediction task involves `tracking' $z_{it}$  using the estimate $y_{it}=\sum_{j=1}^ny_{ijt}$, where $y_{ijt}=\mathbf{b}_{ij}~'\mathbf{x}_{jt}=\sum_{k=0}^{L} b_{ijk}x_{j,t-k}$ are the outputs of a multivariate (causal or one-sided) filter $\mathbf{B}_i^f$, with columns $\mathbf{b}_{ij}$ of fixed length $L$, $j=1,...,5$, assigning weights to the last $L$ observations of the $j$-th indicator $\mathbf{x}_{jt}$ (the fixed-length assumption is merely used for ease of exposition).
@@
 where $\boldsymbol{\Sigma}$ is the variance-covariance matrix of the (white noise) data,  $\mathbf{I}_{L\times L}$ is the  $L\times L$ identity matrix  and  $\mathbf{M}$ is the autocovariance generating matrix introduced earlier.
-Under the white noise assumption, we can express the following expectations:
+Under the simplifying standardized white noise assumption, we can express the following expectations:
 \begin{eqnarray}
@@
 E[y_{it}^2]&=&\mathbf{b}_{i}'\tilde{\mathbf{I}}\mathbf{b}_{i}\nonumber\\%\label{moment3}\\
-E[y_{it-1}y_{it}]&=&\mathbf{b}_{i}'\tilde{\mathbf{M}}\mathbf{b}_{i},\nonumber
+E[y_{it-1}y_{it}]&=&\mathbf{b}_{i}'\tilde{\mathbf{M}}\mathbf{b}_{i}.\nonumber
 \end{eqnarray}
-where, for notational simplicity, we assume $\sigma^2=1$. 
 %The variance of $z_{it}$ is given by 
@@
  \] 
-We now assume $h\geq 0$\footnote{Similar but slightly more complex expressions would arise for $h < 0$ (backcasting), which are omitted here.} and $L$ to be sufficiently large to render $(\boldsymbol{\Gamma} \cdot \boldsymbol{\Xi})_{h+k} \approx 0$ and $(\mathbf{B} \cdot \boldsymbol{\Xi})_j \approx 0$ for $k,j \geq L$, so that  their contributions to the variances of $\hat{\mathbf{y}}_{t,MSE}$ and $\mathbf{y}_t$ are negligible.
+We now assume $h\geq 0$\footnote{Similar but slightly more complex expressions would arise for $h < 0$ (backcasting), which are omitted here.} and $L$ to be sufficiently large to render $(\boldsymbol{\Gamma} \cdot \boldsymbol{\Xi})_{h+k} \approx 0$ and $(\mathbf{B} \cdot \boldsymbol{\Xi})_j \approx 0$ for $k,j \geq L$.
 We define:
@@
 This validation is typically performed through analyses of very long samples of simulated (multivariate) data, ensuring the consistency of the estimation procedures under idealized conditions.} An
-increases by $50\%$ the HT relative to that of the M-MSE predictor\footnote{The convergence of the sample HT to the effectively imposed $HT_1$  can be verified in a simulation study, utilizing the M-SSA package.} but has only a modest effect on the target correlation, suggesting that this trade-off could be beneficial in our application
+increases by $50\%$ the HT relative to that of the M-MSE predictor\footnote{The convergence of the sample HT to the effectively imposed $HT_1$  can be verified in a simulation study, based on  very long samples of the time series, utilizing the M-SSA package.} but has only a modest effect on the target correlation, suggesting that this trade-off could be beneficial in our application
 % , as confirmed by a subsequent analysis of hit rate and false alarm rate

### 2025-08-10 Archive IWH Simon paper/10.08.2025/paper.tex
Update paper.tex

@@
 We proceed as follows.
-Section \ref{sec:data} describes our data series, the sample period used, and documents their dynamic behaviour.
-Section \ref{sec:direct_forecast} provides benchmark forecasts using simple direct projections of GDP based on selected indicators.
+Section \eqref{sec:data} describes our data series, the sample period used, and documents their dynamic behaviour.
+Section \eqref{sec:direct_forecast} provides benchmark forecasts using simple direct projections of GDP based on selected indicators.
 We demonstrate how in-sample forecast performance may improve when we ``de-noise'' the indicators prior to projection with simple smoothing filters.
-Section \ref{sec:mSSA} defines the M-SSA method and shows how its forecast performance compares to that of direct projections as well as how it varies with forecast horizon.
-Section \ref{sec:mSSA_component} considers the ability of our indicators to forecast German GDP growth out-of-sample using the M-SSA framework, and Section \ref{sec:revisions} investigates the temporal stability of the estimated forecasting equations.
-Section \ref{sec:conclusions} summarizes our results and concludes.
+Section \eqref{sec:mSSA} defines the M-SSA method and shows how its forecast performance compares to that of direct projections as well as how it varies with forecast horizon.
+Section \eqref{sec:mSSA_component} considers the ability of our indicators to forecast German GDP growth out-of-sample using the M-SSA framework, and Section \eqref{sec:revisions} investigates the temporal stability of the estimated forecasting equations.
+Section \eqref{sec:conclusions} summarizes our results and concludes.
 
@@
 The analysis employs a set of well-established leading indicators to forecast quarterly German GDP \citep{Drechsel2012financial,Heinisch2018bottom}, including industrial production (IP), the ifo Business Climate Index (ifo\_c), the Economic Sentiment Indicator (ESI), and the term spread between 10-year bond yields and the 3-month EURIBOR rate (spre\_10y\_3m) covering the period 1995 to 2024.
-We take into account the publication lag of the indicators and the GDP series (Table \ref{tab:data_stucture}).\footnote{Data revisions are not considered, given that previous studies \citep{Heinisch2019} have shown that their overall effect on forecast performance is minor.}
+We take into account the publication lag of the indicators and the GDP series (Table \eqref{tab:data_stucture}).\footnote{Data revisions are not considered, given that previous studies \citep{Heinisch2019} have shown that their overall effect on forecast performance is minor.}
 While GDP is available with a one-quarter delay, and IP is released with a two-month lag, business surveys and financial data are published without any lag.
 %All series are consequently re-aligned to correspond with the information available to forecasters, with GDP and IP variables shifted forward according to their respective lags.
-Table \ref{tab:data_stucture_aligned} displays the dataset as of the end of 2024, with GDP and IP variables temporally aligned at the sample endpoint.
+Table \eqref{tab:data_stucture_aligned} displays the dataset as of the end of 2024, with GDP and IP variables temporally aligned at the sample endpoint.
 This configuration reflects the matrix of explanatory variables accessible in Q4-2024 for the purposes of nowcasting or forecasting GDP.
@@
 To allow for missing observations due to publication lags, we re-align series to reflect the information available to forecasters.
-Table \ref{tab:data_stucture_aligned} displays the result as of the end of 2024, with GDP and IP columns shifted to reflect their release dates rather than the periods that they measure.
+Table \eqref{tab:data_stucture_aligned} displays the result as of the end of 2024, with GDP and IP columns shifted to reflect their release dates rather than the periods that they measure.
 This will be the data configuration used for nowcasting or forecasting GDP.
@@
 Additionally, extreme outliers associated with the COVID-19 pandemic are trimmed to prevent distortions in subsequent figures.
- Monthly series are shifted (Table \ref{tab:data_stucture_aligned}) and averaged to quarterly series (Figure \ref{fig:data}), while Figure \ref{fig:data_lags} offers a close-up view of the financial crisis and the pandemic.
+ Monthly series are shifted (Table \eqref{tab:data_stucture_aligned}) and averaged to quarterly series (Figure \eqref{fig:data}), while Figure \eqref{fig:data_lags} offers a close-up view of the financial crisis and the pandemic.
 The latter figure shows that the real-time GDP and industrial production series are synchronized at turning points, whereas the other indicators tend to lead somewhat.
@@
 To examine the joint dynamics of our variables, we estimate their cross-correlations as well as their vector moving-average (VMA) representations.
-The latter is a key input to the multivariate filters we develop in Sections \ref{sec:mSSA} and \ref{sec:mSSA_component}.
-
-Figures \ref{CCF} (full sample) and \ref{CCF_wc} (omitting the Pandemic period) show the sample cross-correlation functions (CCF's) between current and lagged GDP and the two survey indicators.
+The latter is a key input to the multivariate filters we develop in Sections \eqref{sec:mSSA} and \eqref{sec:mSSA_component}.
+
+Figures \eqref{CCF} (full sample) and \eqref{CCF_wc} (omitting the Pandemic period) show the sample cross-correlation functions (CCF's) between current and lagged GDP and the two survey indicators.
 Both figures show strong correlations that peak when GDP is slightly lagged.
@@
 This led to a VAR(3) specification with 7 steps.
-For additional details, see \cite{ElasticNet} and the references therein.} The figures \ref{ma_inv_BIP} and \ref{ma_inv_BIP_var3_en} show the implied responses of GDP to shocks to each indicator variable.\footnote{Shown are the MA coefficients for GDP from the vector moving-average (VMA) models obtained by inversion of the VAR(1) and VAR(3) representations respectively (i.e.
+For additional details, see \cite{ElasticNet} and the references therein.} The figures \eqref{ma_inv_BIP} and \eqref{ma_inv_BIP_var3_en} show the implied responses of GDP to shocks to each indicator variable.\footnote{Shown are the MA coefficients for GDP from the vector moving-average (VMA) models obtained by inversion of the VAR(1) and VAR(3) representations respectively (i.e.
 the impulse-response functions (IRFs) of GDP to the correlated reduced-form shocks.)}
@@
 To illustrate the concept, we rely on all indicators and full sample information, excluding observations affected by the pandemic.
-In-sample forecasts and the forward-shifted GDP, which serves as the target, are presented in Figures \ref{direct_wc} (full dataset without the pandemic) and \ref{direct_wc_financial_crisis} (financial crisis) for forecast horizons up to three quarters ahead.
+In-sample forecasts and the forward-shifted GDP, which serves as the target, are presented in Figures \eqref{direct_wc} (full dataset without the pandemic) and \eqref{direct_wc_financial_crisis} (financial crisis) for forecast horizons up to three quarters ahead.
 
@@
 
-Figures \ref{direct_wc} and \ref{direct_wc_financial_crisis} compare actual and fitted values for GDP at forecast horizons $h = 0, \ldots, 3$, using the four indicators and the target, i.e., GDP shifted $h$ quarters ahead.
-Figure \ref{direct_wc} shows results over the sample without the pandemic, while Figure \ref{direct_wc_financial_crisis} provides a closer look at the fit during the financial crisis.
-Table \ref{tab:pvaluedhp} shows p-values for tests of the joint null hypothesis that all coefficients on the indicators are zero (HAC-adjusted Wald test).
+Figures \eqref{direct_wc} and \eqref{direct_wc_financial_crisis} compare actual and fitted values for GDP at forecast horizons $h = 0, \ldots, 3$, using the four indicators and the target, i.e., GDP shifted $h$ quarters ahead.
+Figure \eqref{direct_wc} shows results over the sample without the pandemic, while Figure \eqref{direct_wc_financial_crisis} provides a closer look at the fit during the financial crisis.
+Table \eqref{tab:pvaluedhp} shows p-values for tests of the joint null hypothesis that all coefficients on the indicators are zero (HAC-adjusted Wald test).
 
@@
 The results indicate that as the forecast horizon $h$ increases, forecast become less volatile and their turning points (i.e., peaks and troughs) become increasingly right-shifted (lagged) relative to the target GDP.
-Additionally, the p-values of the HAC-adjusted Wald test precipitously increase from $h=2$ to $h=3$ (see Table \ref{tab:pvaluedhp}), implying that the indicators become statistically insignificant for forecast horizons of more than two quarters.\footnote{
-These findings are confirmed by the relative root mean-square errors comparing direct forecasts against the mean benchmark, see Table \ref{rRMSE_mSSA_direct_mean_without_covid8} in the Appendix.}
+Additionally, the p-values of the HAC-adjusted Wald test precipitously increase from $h=2$ to $h=3$ (see Table \eqref{tab:pvaluedhp}), implying that the indicators become statistically insignificant for forecast horizons of more than two quarters.\footnote{
+These findings are confirmed by the relative root mean-square errors comparing direct forecasts against the mean benchmark, see Table \eqref{rRMSE_mSSA_direct_mean_without_covid8} in the Appendix.}
 
@@
 \end{tabular}
-\caption{p-values for $H_0: \boldsymbol{\beta} = 0$ in \ref{eq:direct_forecast} based on HAC-adjusted Wald test.\\Estimation over full sample (without pandemic.)} 
+\caption{p-values for $H_0: \boldsymbol{\beta} = 0$ in \eqref{eq:direct_forecast} based on HAC-adjusted Wald test.\\Estimation over full sample (without pandemic.)} 
 \label{tab:pvaluedhp}
@@
 
-Figure \ref{hp_160} presents both classic two-sided and one-sided HP(160) (denoted HP-C for \textit{concurrent}) filters (top panels) along with their associated amplitude functions in the frequency domain (bottom panels).
+Figure \eqref{hp_160} presents both classic two-sided and one-sided HP(160) (denoted HP-C for \textit{concurrent}) filters (top panels) along with their associated amplitude functions in the frequency domain (bottom panels).
 It confirms that the HP-C filter assigns greater weight to yearly components—relevant in a medium-term forecast context— than the classic quarterly HP filter, while also reducing undesirable high-frequency noise.
@@
 
-Figures \ref{direct_hp_forecasts} and \ref{direct_hp_forecasts_financial_crisis} provide results comparable to those of Figures \ref{direct_wc} and \ref{direct_wc_financial_crisis} but now using indicators smoothed with the one-sided HP(160), (denoted as the HP-C, for \textit{concurrent}) filter.
+Figures \eqref{direct_hp_forecasts} and \eqref{direct_hp_forecasts_financial_crisis} provide results comparable to those of Figures \eqref{direct_wc} and \eqref{direct_wc_financial_crisis} but now using indicators smoothed with the one-sided HP(160), (denoted as the HP-C, for \textit{concurrent}) filter.
 The fitted values lag GDP slightly less and track turning points somewhat more quickly.
-Table \ref{tab:pvaluedhp} shows p-values for tests of the joint null hypothesis that all coefficients ${\boldsymbol{\beta}} = 0$.
+Table \eqref{tab:pvaluedhp} shows p-values for tests of the joint null hypothesis that all coefficients ${\boldsymbol{\beta}} = 0$.
 The use of filtered indicators improves their statistical significance at forecast horizons $h > 2$ quarters.
-%Figures \ref{direct_hp_forecasts} (all observations with exclusion of the pandemic) and \ref{direct_hp_forecasts_financial_crisis} (financial crisis).
+%Figures \eqref{direct_hp_forecasts} (all observations with exclusion of the pandemic) and \eqref{direct_hp_forecasts_financial_crisis} (financial crisis).
 A comparison with the classic (unfiltered) direct forecasts in the latter figure %in Fig.\eqref{direct_wc_financial_crisis}
 % indicates that the new predictors are slightly less retarded and track the turning points more promptly (`faster'). 
-% The filtered approach also exhibits superior statistical properties. As shown in Table \ref{tab:f_stats}, 
+% The filtered approach also exhibits superior statistical properties. As shown in Table \eqref{tab:f_stats}, 
 %The p-values of the F-statistics for the HP-based forecasts are smaller than for the direct forecasts specifically at $h=3$, but the predictor remains statistically insignificant at forecast horizons exceeding two quarters (even though the evaluation is conducted in-sample and based on all available indicators).
@@
 %its univariate nature excludes additional information such as that provided by leading indicators. 
-the noise suppression capability of the one-sided filter is compromised, as explained by \cite{Wildi2025} and illustrated by the high-frequency leakage of the amplitude function in Figure \ref{hp_160}.
+the noise suppression capability of the one-sided filter is compromised, as explained by \cite{Wildi2025} and illustrated by the high-frequency leakage of the amplitude function in Figure \eqref{hp_160}.
 To address these limitations, we turn to the Multivariate Smooth Sign Accuracy (M-SSA) framework proposed by \cite{Wildi2025}.
@@
 For simplicity, we assume stationarity of all time series involved.
-In this framework, we might think of ${z}_t$ as corresponding to the output of a two-sided HP(160) filter applied to the doubly-infinite sequence $\mathbf{x}_{\pm \infty}=(x_k)$, $|k|<\infty$, for which $\mathbf{x}_t$ is effectively observed, representing the (log differenced) GDP.
-Since $z_{t}$  depends on future observations $x_{t+1},x_{t+2},...$\footnote{We here ignore the dependence of $z_t$ on $x_0,x_{-1},...$ since the HP filter decays sufficiently fast to zero, see Fig.\ref{hp_160}.}, the prediction task involves `tracking' $z_{t}$  using the (causal) predictor $y_{t}=\sum_{k=0}^{L-1} b_{k}x_{t-k}$, where $L\leq t$.
-
-The task of `tracking' a target can be formalized in different ways; here, we focus on the correlation $\rho(z,y,h)$ between $y_t$ and target $z_{t+h}$, where $h\geq 0$ denotes the forecast horizon (backcasting with $h<0$ is ignored in this context).
+In our current framwork, $z_t$ corresponds to the output of a two-sided Hodrick–Prescott filter with smoothing parameter $\lambda = 160$ applied to the doubly infinite sequence $x_k$, $k\in\mathbb{Z}$.
+The finite sample $\mathbf{x}_t$ is the portion effectively observed and represents the log-differenced GDP (the dependence on $x_0, x_{-1},...$ is neglected because the HP filter weights decay rapidly to zero; see Fig.
+\eqref{hp_160}), the prediction task comprises tracking $z_{t}$  using the (causal) predictor $y_{t}=\sum_{k=0}^{L-1} b_{k}x_{t-k}=\mathbf{b}'\mathbf{x}_t$, where the filter $\mathbf{b}$ with weights $b_0,...,b_{L-1}'$ (of length $L=t$) is available for optimization purposes.
+
+The task of tracking a target can be formalized in different ways; here, we focus on the correlation $\rho(z,y,h)$ between $y_t$ and target $z_{t+h}$, where $h\geq 0$ denotes the forecast horizon (backcasting with $h<0$ is ignored in this context).
 For simplicity, assume a fixed horizon, say $h=h_0$, so that we may drop the reference to $h$ in our notation.
-In the univariate case, \cite{Wildi2025} proposes the Smooth Sign Accuracy (SSA) as an optimization criterion:
+In the univariate case, \cite{Wildi2025} proposes the Smooth Sign Accuracy (SSA) as an optimization criterion for determining $\mathbf{b}$:
 \begin{eqnarray}\label{critssa}
-\left.\begin{array}{c}\rho(z,y)\to\max\\
+\left.\begin{array}{c}\max_{\mathbf{b}}~\rho(z,y)\\
 s.t. \quad \rho(y,1)=\rho_1\end{array}\right\}
@@
 \end{eqnarray}
-Since eq. \ref{eq:ht_arccos} is strictly monotonic, the SSA criterion can be reformulated as the optimization problem:
+Since eq. \eqref{eq:ht_arccos} is strictly monotonic, the SSA criterion can be reformulated as the optimization problem:
 \begin{eqnarray}\label{critssaht}
-\left.\begin{array}{c}\rho(z,y)\to\max\\
+\left.\begin{array}{c}\textrm{max}_{\mathbf{b}}~\rho(z,y)\\
 \frac{\arccos(\rho(y,1))}{\pi}=1/HT_1\end{array}\right\},
@@
 
-One implication of Eq.
-\eqref{critssaht} is that the objective function is indifferent to an affine transformation of the predictor.
+One implication of criterion \eqref{critssaht} is that the objective function as well as the constraint are indifferent to an affine transformation of the predictor.
 This ambiguity can be resolved by assuming an arbitrary scale and level for $y_t$ (standardization).
@@
 
-Unfortunately, classic predictors often generate excessively many false (`noisy') sign changes due to high-frequency leakage, as discussed above (see Figure \ref{hp_160}.)
+Unfortunately, classic predictors often generate excessively many false (`noisy') sign changes due to high-frequency leakage, as discussed above (see Figure \eqref{hp_160}.)
 %as illustrated by the amplitude function in Figure \eqref{hp_160} (bottom right panel) and further elaborated upon by \cite{Wildi2025}. 
- \cite{Wildi2025} shows that the SSA criterion helps control this phenomenon (see Fig.\ref{mssa_msse_zc} for illustration.) Using the dual formulation of the criterion \eqref{critssaht}, the author demonstrates that the SSA solution yields the lowest zero-crossing rate among all (linear) predictors with equivalent target correlation.
+ \cite{Wildi2025} shows that the SSA criterion helps control this phenomenon (see Fig.\eqref{mssa_msse_zc} for illustration.) Using the dual formulation of the criterion \eqref{critssaht}, the author demonstrates that the SSA solution yields the lowest zero-crossing rate among all (linear) predictors with equivalent target correlation.
 The concept can be extended to a multivariate framework, denoted M-SSA, by generalizing the correlation functions as detailed in the next section.\\
 
-We now briefly elucidate the SSA optimization criterion outlined in Equation \ref{critssa}.
+We now briefly elucidate the SSA optimization criterion outlined in Eq.
+\eqref{critssa}.
 Consider a simplified univariate framework where $x_t=\epsilon_t$ represents white noise with variance $\sigma^2$ and assume, for notational simplicity, that $\sigma^2=1$.
@@
 \[z_t=\sum_{k=-\infty}^\infty \gamma_k x_{t-k}=\sum_{k=-\infty}^\infty \gamma_k \epsilon_{t-k},\] 
-where it is assumed that $\gamma_k\neq 0$ for some $k<0$ (implying an acausal filter), and the summation of squared coefficients satisfies $\sum_{|k|<\infty}\gamma_k^2<\infty$, ensuring the stationarity of the target $z_t$.
-With these definitions, all processes are centered at zero, justifying the use of Equation \ref{critssaht} which links first-order ACF and HT.
+where it is assumed that $\gamma_k\neq 0$ for some $k<0$ (implying an acausal filter), and $\sum_{|k|<\infty}\gamma_k^2<\infty$ (squared summability), ensuring the stationarity of the target $z_t$.
+With these definitions, all processes are centered at zero, justifying the use of Equation \eqref{critssaht} which links first-order ACF and HT.
 Extensions to non-stationary integrated processes are discussed in Wildi (2025).
@@
 Let  $y_{h t}=\sum_{k=0}^{L-1} b_{h k}x_{t-k}$  denote a causal predictor of length $L$.
-To simplify the exposition, we consider the case $h=0$ (nowcast), and thus omit the subscript $h$ from the notation.
+For notational simplicity we now omit the superscript $h$ from the predictor and we denote $\boldsymbol{\gamma}_h=(\gamma_{h},\gamma_{h+1},...,\gamma_{h+L-1})'$.
+% To simplify the exposition, we consider the case $h=0$ (nowcast), and thus omit the subscript $h$ from the notation.
 Then, under the postulated white noise assumption
 \begin{eqnarray*}
-\rho(z,y)&=&\frac{\sum_{k=0}^{L-1}b_k\gamma_k}{\sqrt{\sum_{k=0}^{L-1}b_k^2}\sqrt{\sum_{|k|<\infty}\gamma_k^2}},\\
+\rho(z,y)&=&\frac{\mathbf{b}'\boldsymbol{\gamma}_h}{\sqrt{\mathbf{b}'\mathbf{b}}\sqrt{\boldsymbol{\gamma}_{\infty}'\boldsymbol{\gamma}_{\infty}}},\\
 \rho(y,1)&=&\frac{\mathbf{b}'\mathbf{Mb}}{\mathbf{b}'\mathbf{b}},
 \end{eqnarray*} 
-where the so-called autocovariance generating matrix $\mathbf{M}$, of dimension $L\times L$, is defined as:
+where $\boldsymbol{\gamma}_{\infty}'\boldsymbol{\gamma}_{\infty}:=\sum_{|k|<\infty}\gamma_k^2<\infty$ and where the so-called autocovariance generating matrix $\mathbf{M}$, of dimension $L\times L$, is defined as:
 \[
@@
 The matrix $\mathbf{M}$ is such that $\mathbf{b}'\mathbf{Mb}=\sum_{k=1}^{L-1}b_{k-1}b_k$, representing the first-order autocovariance of $y_t$.
-Using this, the optimization criterion \ref{critssa} can be reformulated as:
+Using this, the optimization criterion \eqref{critssa} can be reformulated as:
 \begin{eqnarray}\label{crit1}
@@
 \end{eqnarray}
-where the additional length constraint $\mathbf{b}'\mathbf{b}=l$ serves two purposes: firstly, to ensure the uniqueness of the SSA solution; secondly, to normalize the objective function $\mathbf{b}'\boldsymbol{\gamma}_{h}$, making it proportional to the target correlation $\rho(z,y)$, up to a fixed scaling factor $1/\sqrt{\mathbf{b}'\mathbf{b}}\sqrt{\sum \gamma_k^2}=1/\sqrt{l}\sqrt{\sum \gamma_k^2}$.
-Criterion \ref{crit1} is readily amenable to optimization, as it involves affine and quadratic functions of $\mathbf{b}$, see Wildi (2025) for detailed derivations.
+where the additional length constraint $\mathbf{b}'\mathbf{b}=l$ serves two purposes: firstly, to ensure the uniqueness of the SSA solution; secondly, to normalize the objective function $\mathbf{b}'\boldsymbol{\gamma}_{h}$, making it proportional to the target correlation $\rho(z,y)$, up to a fixed scaling factor $\sqrt{\mathbf{b}'\mathbf{b}}\sqrt{\boldsymbol{\gamma}_{\infty}'\boldsymbol{\gamma}_{\infty}}=\sqrt{l}\sqrt{\boldsymbol{\gamma}_{\infty}'\boldsymbol{\gamma}_{\infty}}$ (in the denominator of the target correlation).
+Criterion \eqref{crit1} is readily amenable to optimization, as it involves affine and quadratic functions of $\mathbf{b}$, see Wildi (2025) for detailed derivations.
 \\
@@
 \[
-y_{MSE,t}=\sum_{k=0}^{L-1}\gamma_kx_{t-k}
+y_{MSE,t}=\sum_{k=0}^{L-1}\gamma_{k+h}x_{t-k}
 \]
-which corresponds to setting $b_{MSE,k}=\gamma_k$, $k=0,...,L-1$.
-In general, to enhance the smoothness of the SSA predictor $y_t$, we select $\rho_1>\rho_{MSE}$ in criterion \ref{crit1},  where $\rho_{MSE}$ denotes the first-order ACF of the (benchmark) MSE predictor\footnote{Selecting $\rho_1<\rho_{MSE}$ would produce an SSA predictor with more frequent zero-crossings than the MSE benchmark, which is typically undesirable in practical applications.}.
+which corresponds to setting $b_{MSE,k}=\gamma_{k+h}$, $k=0,...,L-1$.
+In general, to enhance the smoothness of the SSA predictor $y_t$, we select $\rho_1>\rho_{MSE}$ in criterion \eqref{crit1},  where $\rho_{MSE}$ denotes the first-order ACF of the (benchmark) MSE predictor\footnote{Selecting $\rho_1<\rho_{MSE}$ would produce an SSA predictor with more frequent zero-crossings than the MSE benchmark, which is typically undesirable in practical applications.}.
 Finally,  an extension to autocorrelated $x_t\neq \epsilon_t$ will be discussed below.
@@
 
-We can now extend the SSA criterion \ref{crit1} to a multivariate setting.
+We can now extend the SSA criterion \eqref{crit1} to a multivariate setting.
 Let $\mathbf{X}_t$ (of dimension $t\times n$) denote a set of $n$ explanatory series $\mathbf{x}_{1t},...,\mathbf{x}_{nt}$, with observations $\mathbf{x}_{it}=(x_{it},...,x_{i1})'$.
@@
 In this framework, $\mathbf{X}_t$ represents a matrix of selected economic indicators for the German economy, with $n=5$.
-The targets $z_{it}$, i=1,...,5,   are obtained by applying a two-sided HP(160) filter to the doubly-infinite extensions $(x_{ik})$, $|k|<\infty$, of these indicators, including GDP.
-In a subsequent step, as delineated in Section \ref{sec:mSSA_component}, the (filtered) predictors $y_{hit}$ of $z_{it+h}$, with $h\geq 0$,  will be utilized to construct predictors for GDP.
-However, unlike the approach in the previous section, these predictors will be based on multivariate M-SSA filters rather than univariate HP-C designs.
+The targets $z_{it}$, i=1,...,5,   are obtained by applying a two-sided HP(160) filter to the doubly-infinite extensions $x_{ik}$, $k\in \mathbf{Z}$, of these indicators.
+In a subsequent step, as delineated in Section \eqref{sec:mSSA_component}, the (filtered) predictors $y_{hit}$ of $z_{it+h}$, with $h\geq 0$,  will be utilized to construct predictors for GDP.
+However, unlike the approach in Section \eqref{sec:direct_forecast}, these predictors will be based on multivariate M-SSA filters rather than univariate HP-C designs.
 %Since $z_{it}$  depends on future observations $x_{it+1},x_{it+2},...$, the prediction task involves `tracking' $z_{it}$  using the estimate $y_{it}=\sum_{j=1}^ny_{ijt}$, where $y_{ijt}=\mathbf{b}_{ij}~'\mathbf{x}_{jt}=\sum_{k=0}^{L} b_{ijk}x_{j,t-k}$ are the outputs of a multivariate (causal or one-sided) filter $\mathbf{B}_i^f$, with columns $\mathbf{b}_{ij}$ of fixed length $L$, $j=1,...,5$, assigning weights to the last $L$ observations of the $j$-th indicator $\mathbf{x}_{jt}$ (the fixed-length assumption is merely used for ease of exposition).
@@
 
-To introduce the M-SSA framework, we begin by assuming that  $\mathbf{X}_t=\boldsymbol{\epsilon}_t$ is an $n$-dimensional white noise sequence with a full-rank variance-covariance matrix $\boldsymbol{\Sigma}$.
+For this purpose, to introduce the M-SSA framework, we begin by assuming that  $\mathbf{X}_t=\boldsymbol{\epsilon}_t$ is an $n$-dimensional white noise sequence with a full-rank variance-covariance matrix $\boldsymbol{\Sigma}$.
 Consider the generalized target $\mathbf{z}_t$:
@@
 \]
-with $n\times n$-dimensional filter matrices $\mathbf{B}_k$, $k=0,...,L-1$.
+with $n\times n$-dimensional filter matrices $\mathbf{B}_k$, $k=0,...,L-1$ (for notational simplicity we drop the subscript $h$ referring to the forecast horizon in these expressions).
 Let $b_{ijk}$ denote the entries of the matrix $\mathbf{B}_k$ for $k = 0, \ldots, L-1$.
@@
 Here, $\boldsymbol{\gamma}_{h}$ (or $\mathbf{b}$) is a matrix of dimension $(L\cdot n)\times n$.
-The $i$-th column, $i=1,...,n$, contains the filter weights associated with the $i$-th target (or $i$-th predictor), while the filter weights corresponding to the $j=1,...,n$ series (indicators) are stacked into column vectors of length $L\cdot  n$.
-The $i$-th column vector of $\boldsymbol{\gamma}_{h}$ (or $\mathbf{b}$) can be applied to the $L\cdot n$-dimensional $\boldsymbol{\epsilon}_{\cdot t}$, which stacks all $n$ indicators into a single long data vector.
+The $i$-th column, $i=1,...,n$, contains the filter weights associated with the $i$-th target (or $i$-th predictor), and the filter weights corresponding to the $j=1,...,n$ series (indicators) are stacked into this column vector, of total length $n\cdot L$.
+The $i$-th column vector of $\boldsymbol{\gamma}_{h}$ (or $\mathbf{b}$) can be applied to the $L\cdot n$-dimensional $\boldsymbol{\epsilon}_{\cdot t}$, which stacks all $n$ (artificial) `indicators' into a single long data vector.
 Specifically, we define
@@
 and the same applies to the target variables.
-Note that, for setting up the optimization criterion, it is not necessary to utilize the doubly infinite expressions of the target(s), as illustrated below.
+Recall from the SSA criterion \eqref{crit1}  that it is not necessary to rely on the doubly infinite representations of the targets when formulating the optimization criterion, a point that is subsequently confirmed.
 Under the white noise assumption, the orthogonal projection leads to the mean-squared error (MSE) predictor expressed as:
@@
 \]
-whose components are $\hat{y}_{it,MSE}$, $i=1,...,n$. We define the Kronecker products: 
+%whose components are $\hat{y}_{it,MSE}$, $i=1,...,n$. 
+We define the Kronecker products: 
 \[
@@
 Comparisons with M-MSE allow us to see the cost of this constraint in terms of prediction performance (target correlation).
-Both predictors use the same underlying VAR(1) model introduced in Section \ref{sec:direct_forecast} and estimated on data up to Q4-2007.
+Both predictors use the same underlying VAR(1) model introduced in Section \eqref{sec:direct_forecast} and estimated on data up to Q4-2007.
 This ensures a lengthy out-of-sample period that includes significant events such as the financial crisis, the sovereign debt crisis, and the COVID-19 pandemic.\footnote{The economic sentiment indicator (ESI) does not receive any weight in the parsimonious VAR(1) model estimated over the shorter sub-sample ending in Q4-2007.
@@
 
-Figure \ref{bk_gammak} displays the coefficients of the multivariate nowcasts, where M-SSA's increased smoothing reduces the maximum amplitudes of the filter weights and slows their rate of decay at longer lags.
-The resulting filtered series, presented in Figure \ref{mssa_msse_now}, are standardized to enable direct comparisons.
+Figure \eqref{bk_gammak} displays the coefficients of the multivariate nowcasts, where M-SSA's increased smoothing reduces the maximum amplitudes of the filter weights and slows their rate of decay at longer lags.
+The resulting filtered series, presented in Figure \eqref{mssa_msse_now}, are standardized to enable direct comparisons.
 The M-SSA nowcast is slightly smoother than that of the M-MSE, exhibiting approximately one-third fewer sign-changes.
-Fig.\ref{mssa_msse_zc} compares both nowcasts in relation to the financial crisis, highlighting that the classic (MSE-based) design tends to produce unwanted `noisy' zero-crossings at (or in the vicinity of) the transitions between recessions and expansions.
-Further, Table \ref{corhtnow} compares sample target correlations and HTs of the two nowcasts.
+Fig.\eqref{mssa_msse_zc} compares both nowcasts in relation to the financial crisis, highlighting that the classic (MSE-based) design tends to produce unwanted `noisy' zero-crossings at (or in the vicinity of) the transitions between recessions and expansions.
+Further, Table \eqref{corhtnow} compares sample target correlations and HTs of the two nowcasts.
 Increased smoothness, i.e.
@@
 %This approach facilitates the assessment of forecast performance across multiple horizons within the established methodological context.
-The influence of the forecast horizon on the M-SSA design is depicted in Figure \ref{bk_h}, which compares M-SSA filter weights for the nowcast and the one-year-ahead forecast.
+The influence of the forecast horizon on the M-SSA design is depicted in Figure \eqref{bk_h}, which compares M-SSA filter weights for the nowcast and the one-year-ahead forecast.
 Noting the difference in vertical scales, we see that as $h$ increases (right panel), the filter weights diminish, reflecting increased forecast uncertainty.
@@
 
-M-SSA predictors are compared to the traditional univariate HP-C in Figures  \ref{multivar_vs_univar} and \ref{mssa_hpc_financial_crisis}.
+M-SSA predictors are compared to the traditional univariate HP-C in Figures  \eqref{multivar_vs_univar} and \eqref{mssa_hpc_financial_crisis}.
 In the first figure, we compare both filters across the periods of the financial crisis (top panels) and the COVID-19 pandemic (bottom panels), for forecast horizons $h=0,...,6$.
@@
 
-%(this contrasts with Section \ref{sec:direct_forecast}, which was based on the full sample information). 
+%(this contrasts with Section \eqref{sec:direct_forecast}, which was based on the full sample information). 
 All the results that we present here exclude the pandemic; results including the pandemic are in the Appendix.\\
@@
 % To enhance interpretability and assessment of the M-SSA predictor, we propose analyzing its individual components.
-For illustration, Figure \ref{mssa_predictor_interpretability} shows the nowcast at $h=0$.
+For illustration, Figure \eqref{mssa_predictor_interpretability} shows the nowcast at $h=0$.
 The solid blue line (the nowcast) approaches the zero line near the end of the sample, with a recent upward movement mainly supported by the leading spread component.
@@
 This simple transformation allows $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$ to serve as a predictor for $\textrm{GDP}_{t+shift}$ even when $h\neq shift$.
-This flexibility enables leveraging the dynamic left-shift of $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$, illustrated in Figure \ref{multivar_vs_univar}, when $shift<h$.\\
+This flexibility enables leveraging the dynamic left-shift of $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$, illustrated in Figure \eqref{multivar_vs_univar}, when $shift<h$.\\
 Define the (out-of-sample) GDP component predictor $\hat{\textrm{GDP}}_{t+shift}(h)$:  
@@
 Since M-SSA is based on data up to Q4-2007, $\hat{\textrm{GDP}}_{t+shift}(h)$ constitutes an out-of-sample predictor of $\textrm{GDP}_{t+shift}$ for all $shift\geq 0$ and all $t$ larger than Q4-2007.
-\footnote{Equations \ref{eq:GDP_M-SSA} and \ref{oos_pred} could be extended to incorporate additional M-SSA components, but for illustration purposes, we focus on the simplest (and most robust) design.}
+\footnote{Equations \eqref{eq:GDP_M-SSA} and \eqref{oos_pred} could be extended to incorporate additional M-SSA components, but for illustration purposes, we focus on the simplest (and most robust) design.}
 Define the out-of-sample forecast error
@@
 \end{equation}
-%We now assume that the regression parameters $\hat{\beta}_0(shift,h,t)$ and$,\hat{\beta}_1(shift,h,t)$ in Equation \ref{oos_pred} are up-dated quarterly. Note that 
+%We now assume that the regression parameters $\hat{\beta}_0(shift,h,t)$ and$,\hat{\beta}_1(shift,h,t)$ in Equation \eqref{oos_pred} are up-dated quarterly. Note that 
 While we keep the M-SSA outputs fixed, as based on data up to Q4-2007, we implement an expanding window starting from Q1-2008 for deriving the component weights, which are updated on a quarterly basis.
@@
 
-The statistical significance of the (out of sample) GDP component predictor in Equation \ref{oos_pred}  can be evaluated by regressing 
+The statistical significance of the (out of sample) GDP component predictor in Equation \eqref{oos_pred}  can be evaluated by regressing 
 $\textrm{GDP}_{t+shift}$ on $\hat{\textrm{GDP}}_{t+shift}(h)$:
@@
 \end{equation}
-for observations after Q4-2007: the significance of the corresponding HAC-adjusted test statistics for $H_0 : \alpha_1 =0$ is indicated by a `*' symbol in Table \ref{rRMSE_mSSA_comp_mean_wc}\footnote{`*' indicates significance at the $5\%$ level; `**' indicates significance at the $0.1\%$ level.
+for observations after Q4-2007: the significance of the corresponding HAC-adjusted test statistics for $H_0 : \alpha_1 =0$ is indicated by a `*' symbol in Table \eqref{rRMSE_mSSA_comp_mean_wc}\footnote{`*' indicates significance at the $5\%$ level; `**' indicates significance at the $0.1\%$ level.
 Note that statistical significance is influenced by the relatively short out-of-sample span, of length 17 years.}.
-The MSE forecast performance of $\hat{\textrm{GDP}}_{t+shift}(h)$ can be benchmarked against the mean predictor (see Table \ref{rRMSE_mSSA_comp_mean_wc} which reports relative Root Mean-Square Errors, rRMSE) as well as against the direct forecast (see Table \ref{rRMSE_mSSA_comp_direct_without_covid7}).
+The MSE forecast performance of $\hat{\textrm{GDP}}_{t+shift}(h)$ can be benchmarked against the mean predictor (see Table \eqref{rRMSE_mSSA_comp_mean_wc} which reports relative Root Mean-Square Errors, rRMSE) as well as against the direct forecast (see Table \eqref{rRMSE_mSSA_comp_direct_without_covid7}).
 For consistency, both benchmark predictors are also quarterly updated out-of-sample designs.\\
 
-% Obviously, predicting HP-GDP (see Table \ref{p_val_hpbip_wc}) is less challenging than predicting GDP (see Table \eqref{p_val_wc}), due to smoothness (autocorrelation) of the target when compared to GDP.
+% Obviously, predicting HP-GDP (see Table \eqref{p_val_hpbip_wc}) is less challenging than predicting GDP (see Table \eqref{p_val_wc}), due to smoothness (autocorrelation) of the target when compared to GDP.
 
 Our results can be summarized as follows.
-When excluding the pandemic, the direct forecasts outperform the mean forecast up to two quarters ahead, see Table \ref{rRMSE_mSSA_direct_mean_without_covid8} in the appendix.
+When excluding the pandemic, the direct forecasts outperform the mean forecast up to two quarters ahead, see Table \eqref{rRMSE_mSSA_direct_mean_without_covid8} in the appendix.
 %\footnote{However, the pandemic period generally obscures the relationships between predictors and GDP.}.
-In contrast, the GDP component predictors $\hat{\textrm{GDP}}_{t+shift}(h)$, optimized for longer forecast horizons $h\geq 5$, outperform the mean benchmark up to 5Q ahead, see Table \ref{rRMSE_mSSA_comp_mean_wc}.
-Additionally,  $\hat{\textrm{GDP}}_{t+shift}(h)$, optimized for large forecast horizons $h\geq 4$, surpass the direct forecasts for $shift\geq 3$, see Table \ref{rRMSE_mSSA_comp_direct_without_covid7} in the appendix.
-Statistical significance of the GDP component predictor, optimized for larger forecast horizons, can be established for at least up to one year ahead, see Table \ref{rRMSE_mSSA_comp_mean_wc}.
-Finally, the strong performance of predictors optimized for larger forecast horizons ($h\geq 4$), at smaller shifts ($shifts< 4$) corroborates the importance of the left-shift of the corresponding predictors, see Figs.\ref{multivar_vs_univar} and \ref{mssa_hpc_financial_crisis}, in relation to forecast performance.
+In contrast, the GDP component predictors $\hat{\textrm{GDP}}_{t+shift}(h)$, optimized for longer forecast horizons $h\geq 5$, outperform the mean benchmark up to 5Q ahead, see Table \eqref{rRMSE_mSSA_comp_mean_wc}.
+Additionally,  $\hat{\textrm{GDP}}_{t+shift}(h)$, optimized for large forecast horizons $h\geq 4$, surpass the direct forecasts for $shift\geq 3$, see Table \eqref{rRMSE_mSSA_comp_direct_without_covid7} in the appendix.
+Statistical significance of the GDP component predictor, optimized for larger forecast horizons, can be established for at least up to one year ahead, see Table \eqref{rRMSE_mSSA_comp_mean_wc}.
+Finally, the strong performance of predictors optimized for larger forecast horizons ($h\geq 4$), at smaller shifts ($shifts< 4$) corroborates the importance of the left-shift of the corresponding predictors, see Figs.\eqref{multivar_vs_univar} and \eqref{mssa_hpc_financial_crisis}, in relation to forecast performance.
 
@@
 \end{tabular}
-\caption{rRMSEs: out-of-sample forecast errors Equation \ref{oosfe} vs expanding mean of GDP\\Out-of-sample evaluation starting in Q1-2008, without pandemic\\Values $< 1$ indicate superior predictions from Equation \ref{oos_pred}\\
+\caption{rRMSEs: out-of-sample forecast errors Equation \eqref{oosfe} vs expanding mean of GDP\\Out-of-sample evaluation starting in Q1-2008, without pandemic\\Values $< 1$ indicate superior predictions from Equation \eqref{oos_pred}\\
 Row-index: Shift refers to the left-shift of the target $\textrm{GDP}_{t+shift}$\\
 Column-index: $h$ refers to the forecast horizon for which $\textrm{M-SSA}_t^{\textrm{GDP}}(h)$ is optimized\\
-Significance refers to the hypothesis $\alpha_1=0$ in Equation \ref{eq_gdp_pred}.
+Significance refers to the hypothesis $\alpha_1=0$ in Equation \eqref{eq_gdp_pred}.
 } 
@@
 %\end{eqnarray}
-%Due to the strict monotonicity of the arcsin function, maximizing the target correlation also maximizes the sign accuracy of the predictor.\footnote{Equation \ref{sig_acc} holds exactly for Gaussian processes.
+%Due to the strict monotonicity of the arcsin function, maximizing the target correlation also maximizes the sign accuracy of the predictor.\footnote{Equation \eqref{sig_acc} holds exactly for Gaussian processes.
 Due to the central limit theorem, the relationship holds approximately for low-pass filtered non-Gaussian series, see Wildi (2025).}
@@
 For clarity, these results are based on the entire sample, excluding the pandemic period.
-Fig.\ref{ROC_GDP_shift_1_4} compares ROCs of the predictors for small and medium-term forecast horizons and Table \ref{auc} presents AUC (area under the curve) for forecast horizons 1-5 quarters ahead.
+Fig.\eqref{ROC_GDP_shift_1_4} compares ROCs of the predictors for small and medium-term forecast horizons and Table \eqref{auc} presents AUC (area under the curve) for forecast horizons 1-5 quarters ahead.
 \begin{figure}[h]\begin{center}\includegraphics[height=3in, width=5in]{./Figures/ROC_GDP_shift_1_4.pdf}\caption{Hit rate vs.
@@
         \caption{Revisions of the M-SSA GDP component predictor are shown, with the final predictor in blue and the real-time design in red.
-These revisions result from quarterly updates to the regression coefficients of equation \ref{oos_pred}.
+These revisions result from quarterly updates to the regression coefficients of equation \eqref{oos_pred}.
 The M-SSA filter    itself remains fixed, based on data up to Q4-2007.
@@
         \includegraphics[height=3in, width=4.5in]{./Figures/revisions2.pdf}
-        \caption{Revisions of intercept and regression weight of Equation \ref{oos_pred}.
+        \caption{Revisions of intercept and regression weight of Equation \eqref{oos_pred}.
         \label{revisions2}}
@@
 %??? Justification HP(160): replace HP(1600) by AR(2) with similar persistency but shorter cycle lengthi, i.e. $a_1=-2"persistency*acos(\theta, a_2=persistency$.
-Figure \ref{revisions1} compares the final and real-time predictors: as the sample size increases, the real-time predictor converges steadily to the final predictor.\footnote{The rate of convergence is partially determined by the simplicity of our design, which avoids incorporating additional M-SSA components.}
-As a confirmation, Figure \ref{revisions2} displays the evolution of the intercept and regression weights over time: with increasing sample size, these estimates seem to converge to fixed points, indicating both the stationarity of the process and the consistency of the estimates.
+Figure \eqref{revisions1} compares the final and real-time predictors: as the sample size increases, the real-time predictor converges steadily to the final predictor.\footnote{The rate of convergence is partially determined by the simplicity of our design, which avoids incorporating additional M-SSA components.}
+As a confirmation, Figure \eqref{revisions2} displays the evolution of the intercept and regression weights over time: with increasing sample size, these estimates seem to converge to fixed points, indicating both the stationarity of the process and the consistency of the estimates.
 %(Note, however, the slightly elevated values of the regression weight assigned to the M-SSA component during the financial crisis and the pandemic, suggesting a stronger link between the target and the predictor).
@@
 
-The second source of revisions in the M-SSA component predictor arises from updates to the VAR(1) model within the M-SSA filter\footnote{Similar findings apply to more sophisticated alternatives such as the VAR(3) discussed in Section \ref{sec:data}.}.
+The second source of revisions in the M-SSA component predictor arises from updates to the VAR(1) model within the M-SSA filter\footnote{Similar findings apply to more sophisticated alternatives such as the VAR(3) discussed in Section \eqref{sec:data}.}.
 We compare the original estimates, based on data up to Q4-2007, with the final estimates obtained using $h=4$ and $shift=4$ (similar findings apply to other values of forecast horizon and forward-shift).
 For consistency, we excluded the entire pandemic episode from the data.
-Figure \ref{up_dated_ma_inv_multi_ip} compares the MA-inversions (impulse responses) of the GDP equation derived from the `old' and the updated VAR models: unlike the former, the latter incorporates ESI as an additional explanatory variable for GDP.
-Figure \ref{bk_2008_all} illustrates the resulting effect on the M-SSA filter targeting HP-GDP at forecast horizon $h=4$.
+Figure \eqref{up_dated_ma_inv_multi_ip} compares the MA-inversions (impulse responses) of the GDP equation derived from the `old' and the updated VAR models: unlike the former, the latter incorporates ESI as an additional explanatory variable for GDP.
+Figure \eqref{bk_2008_all} illustrates the resulting effect on the M-SSA filter targeting HP-GDP at forecast horizon $h=4$.
 The updated filter (right panel) assigns more weight to the additional `leading' indicators relative to GDP.
-Finally, Figure \ref{revisions3} compares the M-SSA GDP predictors.
+Finally, Figure \eqref{revisions3} compares the M-SSA GDP predictors.
 To isolate the effect of the regression revision, the figure displays standardized series.
@@
 It is important to note that the smoothness (zero-crossing rate) remains unchanged, though.
-This suggests that the out-of-sample performance of the M-SSA component predictor may be better than reported in Section \ref{oosp}, which is based on the fixed and increasingly outdated version of M-SSA.
+This suggests that the out-of-sample performance of the M-SSA component predictor may be better than reported in Section \eqref{oosp}, which is based on the fixed and increasingly outdated version of M-SSA.
 
@@
 \end{tabular}
-\caption{p-values for $H_0: {\alpha_1} = 0$, HAC-adjusted Wald test, Equation \ref{eq_gdp_pred}.\\ including the pandemic.} 
+\caption{p-values for $H_0: {\alpha_1} = 0$, HAC-adjusted Wald test, Equation \eqref{eq_gdp_pred}.\\ including the pandemic.} 
 \label{p_val1}
@@
 \end{tabular}
-\caption{rRMSEs: out-of-sample forecast errors Equation \ref{oosfe} vs expanding mean of GDP\\Out-of-sample evaluation starting in Q1-2008, including the pandemic.\\Values $< 1$ indicate superior predictions from Equation \ref{oos_pred}.}
+\caption{rRMSEs: out-of-sample forecast errors Equation \eqref{oosfe} vs expanding mean of GDP\\Out-of-sample evaluation starting in Q1-2008, including the pandemic.\\Values $< 1$ indicate superior predictions from Equation \eqref{oos_pred}.}
 \label{rRMSE_mSSA_comp_mean2}

### 2025-08-10 Archive IWH Simon paper/10.08.2025/paper.tex
up

@@
 \usepackage{amsmath}
+\usepackage{amsfonts}
 \usepackage[a4paper,includehead]{geometry}
@@
 The finite sample $\mathbf{x}_t$ is the portion effectively observed and represents the log-differenced GDP (the dependence on $x_0, x_{-1},...$ is neglected because the HP filter weights decay rapidly to zero; see Fig.
-\eqref{hp_160}), the prediction task comprises tracking $z_{t}$  using the (causal) predictor $y_{t}=\sum_{k=0}^{L-1} b_{k}x_{t-k}=\mathbf{b}'\mathbf{x}_t$, where the filter $\mathbf{b}$ with weights $b_0,...,b_{L-1}'$ (of length $L=t$) is available for optimization purposes.
-
-The task of tracking a target can be formalized in different ways; here, we focus on the correlation $\rho(z,y,h)$ between $y_t$ and target $z_{t+h}$, where $h\geq 0$ denotes the forecast horizon (backcasting with $h<0$ is ignored in this context).
+\eqref{hp_160}).
+The prediction task comprises tracking $z_{t}$  using the (causal) predictor $y_{t}=\sum_{k=0}^{L-1} b_{k}x_{t-k}=\mathbf{b}'\mathbf{x}_t$, where the filter $\mathbf{b}$ with weights $b_0,...,b_{L-1}'$ (actually of length $L=t$) is available for optimization purposes.
+
+The task of tracking a target `optimally' can be formalized in different ways; here, we focus on the correlation $\rho(z,y,h)$ between $y_t$ and target $z_{t+h}$, where $h\geq 0$ denotes the forecast horizon (backcasting with $h<0$ is ignored in this context).
 For simplicity, assume a fixed horizon, say $h=h_0$, so that we may drop the reference to $h$ in our notation.
@@
 
-Unfortunately, classic predictors often generate excessively many false (`noisy') sign changes due to high-frequency leakage, as discussed above (see Figure \eqref{hp_160}.)
+Traditional predictors often yield an excessive number of false sign changes due to leakage of high-frequency components (high-frequency leakage) in the predicted series, as discussed above (see Figure \eqref{hp_160}).
+This noisiness is particularly problematic near the onset or termination of economic crises (recessions), where the target fluctuates around zero; reliable tracking of sign changes would facilitate real-time assessment of the occurrence and timing of such events.
 %as illustrated by the amplitude function in Figure \eqref{hp_160} (bottom right panel) and further elaborated upon by \cite{Wildi2025}. 
- \cite{Wildi2025} shows that the SSA criterion helps control this phenomenon (see Fig.\eqref{mssa_msse_zc} for illustration.) Using the dual formulation of the criterion \eqref{critssaht}, the author demonstrates that the SSA solution yields the lowest zero-crossing rate among all (linear) predictors with equivalent target correlation.
+ \cite{Wildi2025} shows that the SSA criterion helps control this phenomenon (see Fig.\eqref{mssa_msse_zc} for illustration).
+Using the dual formulation of the criterion \eqref{critssaht}, the author demonstrates that the SSA solution yields the \emph{lowest} zero-crossing rate among all (linear) predictors with equivalent target correlation.
 The concept can be extended to a multivariate framework, denoted M-SSA, by generalizing the correlation functions as detailed in the next section.\\
@@
 Let  $y_{h t}=\sum_{k=0}^{L-1} b_{h k}x_{t-k}$  denote a causal predictor of length $L$.
-For notational simplicity we now omit the superscript $h$ from the predictor and we denote $\boldsymbol{\gamma}_h=(\gamma_{h},\gamma_{h+1},...,\gamma_{h+L-1})'$.
+For notational convenience we now omit the subscript $h$ from the predictor and we denote $\boldsymbol{\gamma}_h:==(\gamma_{h},\gamma_{h+1},...,\gamma_{h+L-1})'$.
 % To simplify the exposition, we consider the case $h=0$ (nowcast), and thus omit the subscript $h$ from the notation.
-Then, under the postulated white noise assumption
+Then, under the postulated white noise assumption we obtain:
 \begin{eqnarray*}
-\rho(z,y)&=&\frac{\mathbf{b}'\boldsymbol{\gamma}_h}{\sqrt{\mathbf{b}'\mathbf{b}}\sqrt{\boldsymbol{\gamma}_{\infty}'\boldsymbol{\gamma}_{\infty}}},\\
-\rho(y,1)&=&\frac{\mathbf{b}'\mathbf{Mb}}{\mathbf{b}'\mathbf{b}},
+\rho(z,y)=\frac{\mathbf{b}'\boldsymbol{\gamma}_h}{\sqrt{\mathbf{b}'\mathbf{b}}\sqrt{\boldsymbol{\gamma}_{\infty}'\boldsymbol{\gamma}_{\infty}}} \textrm{~and~}\rho(y,1)=\frac{\mathbf{b}'\mathbf{Mb}}{\mathbf{b}'\mathbf{b}},
 \end{eqnarray*} 
@@
 &\max_{\mathbf{b}}\mathbf{b}'\boldsymbol{\gamma}_{h}\\
-\textrm{subject~to~}&\mathbf{b}'\mathbf{Mb}=l\rho_1\\
+\textrm{s.t.~}&\mathbf{b}'\mathbf{Mb}=l\rho_1\\
 &\mathbf{b}'\mathbf{b}=l
@@
 \end{eqnarray}
-where the additional length constraint $\mathbf{b}'\mathbf{b}=l$ serves two purposes: firstly, to ensure the uniqueness of the SSA solution; secondly, to normalize the objective function $\mathbf{b}'\boldsymbol{\gamma}_{h}$, making it proportional to the target correlation $\rho(z,y)$, up to a fixed scaling factor $\sqrt{\mathbf{b}'\mathbf{b}}\sqrt{\boldsymbol{\gamma}_{\infty}'\boldsymbol{\gamma}_{\infty}}=\sqrt{l}\sqrt{\boldsymbol{\gamma}_{\infty}'\boldsymbol{\gamma}_{\infty}}$ (in the denominator of the target correlation).
-Criterion \eqref{crit1} is readily amenable to optimization, as it involves affine and quadratic functions of $\mathbf{b}$, see Wildi (2025) for detailed derivations.
+where the additional length constraint $\mathbf{b}'\mathbf{b}=l$ serves two purposes: firstly, to ensure the uniqueness of the SSA solution; secondly, to normalize the objective function $\mathbf{b}'\boldsymbol{\gamma}_{h}$, making it proportional to the target correlation $\rho(z,y)$, up to a fixed scaling factor $\sqrt{\mathbf{b}'\mathbf{b}}\sqrt{\boldsymbol{\gamma}_{\infty}'\boldsymbol{\gamma}_{\infty}}=\sqrt{l}\sqrt{\boldsymbol{\gamma}_{\infty}'\boldsymbol{\gamma}_{\infty}}$ (for convenience we often select $l=1$, leading to a standardized $y_t$).
+Criterion \eqref{crit1} is readily amenable to optimization, as it involves affine and quadratic functions of $\mathbf{b}$, see Wildi (2024) for background (detailed derivations are obtained in Wildi (2025)).
 \\
-Under the above assumptions, the classical MSE predictor can be expressed via orthogonal projection as:
+Under the above assumptions, the classical MSE predictor of length $L$ can be expressed via orthogonal projection as:
 \[
@@
 In general, to enhance the smoothness of the SSA predictor $y_t$, we select $\rho_1>\rho_{MSE}$ in criterion \eqref{crit1},  where $\rho_{MSE}$ denotes the first-order ACF of the (benchmark) MSE predictor\footnote{Selecting $\rho_1<\rho_{MSE}$ would produce an SSA predictor with more frequent zero-crossings than the MSE benchmark, which is typically undesirable in practical applications.}.
-Finally,  an extension to autocorrelated $x_t\neq \epsilon_t$ will be discussed below.
+Finally,  an extension to autocorrelated $x_t\neq \epsilon_t$ is discussed below.
 
@@
 Let $\mathbf{X}_t$ (of dimension $t\times n$) denote a set of $n$ explanatory series $\mathbf{x}_{1t},...,\mathbf{x}_{nt}$, with observations $\mathbf{x}_{it}=(x_{it},...,x_{i1})'$.
-Let $\mathbf{Z}_t$, of dimension $t\times m$ (in our case $m=n$) a set of $m$ target series $\mathbf{z}_{it}=(z_{it},...,z_{i1})'$, $i=1,...m$, which typically lie outside the linear space spanned by   $\mathbf{X}_t$ (and are unknown at time $t$ but not independent of $\mathbf{X}_t$).
+Let $\mathbf{Z}_t$, of dimension $t\times m$  a set of $m$ target series $\mathbf{z}_{it}=(z_{it},...,z_{i1})'$, $i=1,...m$, which typically lie outside the linear space spanned by   $\mathbf{X}_t$ (and are unknown at time $t$ but not independent of $\mathbf{X}_t$).
 In this framework, $\mathbf{X}_t$ represents a matrix of selected economic indicators for the German economy, with $n=5$.
 The targets $z_{it}$, i=1,...,5,   are obtained by applying a two-sided HP(160) filter to the doubly-infinite extensions $x_{ik}$, $k\in \mathbf{Z}$, of these indicators.
-In a subsequent step, as delineated in Section \eqref{sec:mSSA_component}, the (filtered) predictors $y_{hit}$ of $z_{it+h}$, with $h\geq 0$,  will be utilized to construct predictors for GDP.
+In a subsequent step, as delineated in Section \eqref{sec:mSSA_component}, the (filtered) predictors $y_{hit}$ of $z_{it+h}$, with $h\geq 0$,  will be utilized to construct predictors for the GDP.
 However, unlike the approach in Section \eqref{sec:direct_forecast}, these predictors will be based on multivariate M-SSA filters rather than univariate HP-C designs.
@@
 \] 
-where $\boldsymbol{\Gamma}_k$  is a sequence of filter matrices of dimension  $n\times n$, satisfying the square-summability condition.
+where $\boldsymbol{\Gamma}_k$  is a sequence of filter matrices of dimension  $n\times n$ (hence we assume $m=n$), satisfying the square-summability condition.
 Each $\boldsymbol{\Gamma}_k$ has entries $(\gamma_{ijk})$ for $i, j = 1, \ldots, n$.
@@
 \[
-\mathbf{y}_t = \mathbf{b}'\boldsymbol{\epsilon}_{\cdot t}
+\mathbf{y}_t = \mathbf{b}'\boldsymbol{\epsilon}_{\cdot t}.
 \]
-and the same applies to the target variables.
-Recall from the SSA criterion \eqref{crit1}  that it is not necessary to rely on the doubly infinite representations of the targets when formulating the optimization criterion, a point that is subsequently confirmed.
-Under the white noise assumption, the orthogonal projection leads to the mean-squared error (MSE) predictor expressed as:
+Also, under the white noise assumption, the orthogonal projection leads to the MSE predictor expressed as:
 \[
-\hat{\mathbf{y}}_{t,MSE} = \boldsymbol{\gamma}_{h}'\boldsymbol{\epsilon}_{\cdot t},
+\hat{\mathbf{y}}_{t,MSE} = \boldsymbol{\gamma}_{h}'\boldsymbol{\epsilon}_{\cdot t}.
 \]
+Note that we here assume the MSE predictor to rely on sub-filters of length $L$ for  each subseries $\mathbf{x}_{it}$.  
+Recall from the SSA criterion \eqref{crit1}  that it is not necessary to rely on the doubly infinite representations of the target when formulating the optimization criterion, a point that is subsequently confirmed.
 %whose components are $\hat{y}_{it,MSE}$, $i=1,...,n$. 
@@
 \textrm{max}_{\mathbf{b}_i}\boldsymbol{\gamma}_{i\cdot h}'\tilde{\mathbf{I}}\mathbf{b}_{i}\\
-\textrm{subject~to~}\mathbf{b}_{i}'\tilde{\mathbf{M}}\mathbf{b}_{i}=\rho_i\nonumber\\
+\textrm{s.t.~}\mathbf{b}_{i}'\tilde{\mathbf{M}}\mathbf{b}_{i}=\rho_i\nonumber\\
 \mathbf{b}_{i}'\tilde{\mathbf{I}}\mathbf{b}_{i}=1,\nonumber
@@
 \max_{(\mathbf{b}\cdot\boldsymbol{\xi})_i} (\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{ih}'\mathbf{\tilde{I}} (\mathbf{b}\cdot\boldsymbol{\xi})_i\\
-\textrm{subject~to~}(\mathbf{b}\cdot\boldsymbol{\xi})_i'\mathbf{\tilde{M}}(\mathbf{b}\cdot\boldsymbol{\xi})_i=\rho_i\nonumber\\
+\textrm{s.t.~}(\mathbf{b}\cdot\boldsymbol{\xi})_i'\mathbf{\tilde{M}}(\mathbf{b}\cdot\boldsymbol{\xi})_i=\rho_i\nonumber\\
 (\mathbf{b}\cdot\boldsymbol{\xi})_i'\mathbf{\tilde{I}}(\mathbf{b}\cdot\boldsymbol{\xi})_i=1\nonumber
@@
 for $i=1,...,n$.
-In analogy to the above white noise case, this formulation can be solved for $(\mathbf{b}\cdot\boldsymbol{\xi})_i$.
-The M-SSA coefficients $\mathbf{b}_i$ can subsequently be determined through deconvolution, as indicated in Equation \eqref{con_inv}.
+In analogy to the above white noise case, the criterion involves affine and quadratic functions of $(\mathbf{b}\cdot\boldsymbol{\xi})_i$ and thus can be solved for the convolution weights.
+The requested M-SSA coefficients $\mathbf{b}_i$ can subsequently be determined through deconvolution, as indicated in Equation \eqref{con_inv}.
 The present discussion is confined to stationary processes, omitting a lengthier treatment of rank-defficient designs (cointegration) in the non-stationary case.\\

### 2025-08-11 Archive IWH Simon paper/10.08.2025/paper.tex
up

@@
 
-\title{Smooth and Persistent Forecasts of German GDP\\ Balancing Accuracy and Stability\thanks{We would like Barbara Rossi and the participants at the Vienna Forecasting Workshop 2025 for their comments and suggestions.}}
+\title{Smooth and Persistent Forecasts of German GDP:\\ Balancing Accuracy and Stability\thanks{We would like Barbara Rossi and the participants at the Vienna Forecasting Workshop 2025 for their comments and suggestions.}}
 \author{Katja Heinisch\thanks{Halle Institute for Economic Research (IWH).} \hspace{0.2cm}
@@
 Marc Wildi\thanks{Zurich University of Applied Sciences (ZHAW).}}
-\date{July 2025 \\
+\date{August 2025 \\
 \vspace{2cm} \textit{Work in progress, not for circulation}}
@@
 This enhances the signal-to-noise ratio and controls forecast volatility and timing.
-Applied to German GDP, our method yields smoothed forecasts that improve the forecasting accuracy of traditional models, particularly over medium-term horizons.
+Applied to German GDP growth, our method yields smoothed forecasts that improve the forecasting accuracy of traditional models, particularly over medium-term horizons.
 While smoother forecasts tend to lag slightly, this drawback can be mitigated by adjusting the forecast horizon.
@@
 \bigskip \vspace{10pt}
-\textbf{Keywords:} Forecast Smoothing, Smooth Sign Accuracy 
+\textbf{Keywords:} Forecast Smoothing, Smooth Sign Accuracy, Time Series Filtering 
 \newline
@@
 
-Economic forecasters face a dilemma
-: while informational efficiency requires that forecasts incorporate all available data, doing so often results in forecasts that are excessively volatile.
+Economic forecasters face a dilemma: while informational efficiency requires that forecasts incorporate all available data, doing so often results in forecasts that are excessively volatile.
  %Volatile forecasts may be a sign that they efficiently incorporate all available information. 
@@
 This study extends this framework by replacing raw leading indicators with smoothed and filtered components, thereby improving the signal-to-noise ratio in the predictor variables.
-We apply this framework to well-established German economic indicators to show its ability to improve forecasts of GDP \citep{lehmann2021predicting,Heinisch2018bottom}  and GDP trends using a variety of sample periods and performance metrics.
+We apply this framework to well-established German economic indicators to show its ability to improve forecasts of GDP growth \citep{lehmann2021predicting,Heinisch2018bottom}  and GDP trends using a variety of sample periods and performance metrics.
 
@@
 We proceed as follows.
-Section \eqref{sec:data} describes our data series, the sample period used, and documents their dynamic behaviour.
-Section \eqref{sec:direct_forecast} provides benchmark forecasts using simple direct projections of GDP based on selected indicators.
+Section \ref{sec:data} describes our data series, the sample period used, and documents their dynamic behaviour.
+Section \ref{sec:direct_forecast} provides benchmark forecasts using simple direct projections of GDP based on selected indicators.
 We demonstrate how in-sample forecast performance may improve when we ``de-noise'' the indicators prior to projection with simple smoothing filters.
-Section \eqref{sec:mSSA} defines the M-SSA method and shows how its forecast performance compares to that of direct projections as well as how it varies with forecast horizon.
-Section \eqref{sec:mSSA_component} considers the ability of our indicators to forecast German GDP growth out-of-sample using the M-SSA framework, and Section \eqref{sec:revisions} investigates the temporal stability of the estimated forecasting equations.
-Section \eqref{sec:conclusions} summarizes our results and concludes.
+Section \ref{sec:mSSA} defines the M-SSA method and shows how its forecast performance compares to that of direct projections as well as how it varies with forecast horizon.
+Section \ref{sec:mSSA_component} considers the ability of our indicators to forecast German GDP growth out-of-sample using the M-SSA framework, and Section \ref{sec:revisions} investigates the temporal stability of the estimated forecasting equations.
+Section \ref{sec:conclusions} summarizes our results and concludes.
 
@@
 The analysis employs a set of well-established leading indicators to forecast quarterly German GDP \citep{Drechsel2012financial,Heinisch2018bottom}, including industrial production (IP), the ifo Business Climate Index (ifo\_c), the Economic Sentiment Indicator (ESI), and the term spread between 10-year bond yields and the 3-month EURIBOR rate (spre\_10y\_3m) covering the period 1995 to 2024.
-We take into account the publication lag of the indicators and the GDP series (Table \eqref{tab:data_stucture}).\footnote{Data revisions are not considered, given that previous studies \citep{Heinisch2019} have shown that their overall effect on forecast performance is minor.}
+We take into account the publication lag of the indicators and the GDP series (Table \ref{tab:data_stucture}).\footnote{Data revisions are not considered, given that previous studies \citep{Heinisch2019} have shown that their overall effect on forecast performance is minor.}
 While GDP is available with a one-quarter delay, and IP is released with a two-month lag, business surveys and financial data are published without any lag.
@@
 \begin{table}[ht]
+\caption{Ragged edge structure. As of January 2025, before publication of the GDP flash estimate.} 
+\label{tab:data_stucture}
 \centering
@@
 \end{tabular}
-\caption{Ragged edge structure. As of January 2025, before publication of the GDP flash estimate.} 
-\label{tab:data_stucture}
+
 \end{table}
@@
 To allow for missing observations due to publication lags, we re-align series to reflect the information available to forecasters.
-Table \eqref{tab:data_stucture_aligned} displays the result as of the end of 2024, with GDP and IP columns shifted to reflect their release dates rather than the periods that they measure.
-This will be the data configuration used for nowcasting or forecasting GDP.
+Table \ref{tab:data_stucture_aligned} displays the artificial result as of the end of 2024, with GDP and IP columns shifted to reflect their release dates rather than the periods that they measure.
+This data configuration will be the used for nowcasting or forecasting GDP.
 For illustrative purposes, an additional target column is included on the right side of the table, representing the dependent variable for nowcasting applications.\footnote{In the case of an $h$-step ahead forecast, this column is shifted upwards by $h$ quarters.}
@@
 \begin{table}[ht]
+\caption{Aligned dataset.
+At each date, various columns show the most recent values of each predictor available to forecasters.
+Column 6 represents the target variable for a GDP nowcast (i.e.
+the values of column 1 shifted upwards by the publication lag of GDP.}
+\label{tab:data_stucture_aligned}
 \centering
@@
 \end{tabular}
-\caption{Aligned dataset.
-At each date, columns 1 through 5 show the most recent values of each predictor available to forecasters.
-Column 6 represents the target variable for a GDP nowcast (i.e.
-the values of column 1 shifted upwards by the publication lag of GDP.}
-\label{tab:data_stucture_aligned}
+
 \end{table}
@@
 Additionally, extreme outliers associated with the COVID-19 pandemic are trimmed to prevent distortions in subsequent figures.
- Monthly series are shifted (Table \eqref{tab:data_stucture_aligned}) and averaged to quarterly series (Figure \eqref{fig:data}), while Figure \eqref{fig:data_lags} offers a close-up view of the financial crisis and the pandemic.
+ Monthly series are shifted (Table \ref{tab:data_stucture_aligned}) and averaged to quarterly series (Figure \ref{fig:data}), while Figure \ref{fig:data_lags} offers a close-up view of the financial crisis and the pandemic.
 The latter figure shows that the real-time GDP and industrial production series are synchronized at turning points, whereas the other indicators tend to lead somewhat.
@@
 To examine the joint dynamics of our variables, we estimate their cross-correlations as well as their vector moving-average (VMA) representations.
-The latter is a key input to the multivariate filters we develop in Sections \eqref{sec:mSSA} and \eqref{sec:mSSA_component}.
-
-Figures \eqref{CCF} (full sample) and \eqref{CCF_wc} (omitting the Pandemic period) show the sample cross-correlation functions (CCF's) between current and lagged GDP and the two survey indicators.
+The latter is a key input to the multivariate filters we develop in Sections \ref{sec:mSSA} and \ref{sec:mSSA_component}.
+
+Figures \ref{CCF} (full sample) and \ref{CCF_wc} (omitting the Pandemic period) show the sample cross-correlation functions (CCF's) between current and lagged GDP and the two survey indicators.
 Both figures show strong correlations that peak when GDP is slightly lagged.
@@
 This led to a VAR(3) specification with 7 steps.
-For additional details, see \cite{ElasticNet} and the references therein.} The figures \eqref{ma_inv_BIP} and \eqref{ma_inv_BIP_var3_en} show the implied responses of GDP to shocks to each indicator variable.\footnote{Shown are the MA coefficients for GDP from the vector moving-average (VMA) models obtained by inversion of the VAR(1) and VAR(3) representations respectively (i.e.
+For additional details, see \cite{ElasticNet} and the references therein.} The figures \ref{ma_inv_BIP} and \ref{ma_inv_BIP_var3_en} show the implied responses of GDP to shocks to each indicator variable.\footnote{Shown are the MA coefficients for GDP from the vector moving-average (VMA) models obtained by inversion of the VAR(1) and VAR(3) representations respectively (i.e.
 the impulse-response functions (IRFs) of GDP to the correlated reduced-form shocks.)}
@@
 
-Figures \eqref{direct_wc} and \eqref{direct_wc_financial_crisis} compare actual and fitted values for GDP at forecast horizons $h = 0, \ldots, 3$, using the four indicators and the target, i.e., GDP shifted $h$ quarters ahead.
-Figure \eqref{direct_wc} shows results over the sample without the pandemic, while Figure \eqref{direct_wc_financial_crisis} provides a closer look at the fit during the financial crisis.
-Table \eqref{tab:pvaluedhp} shows p-values for tests of the joint null hypothesis that all coefficients on the indicators are zero (HAC-adjusted Wald test).
+Figures \ref{direct_wc} and \ref{direct_wc_financial_crisis} compare actual and fitted values for GDP at forecast horizons $h = 0, \ldots, 3$, using the four indicators and the target, i.e., GDP shifted $h$ quarters ahead.
+Figure \ref{direct_wc} shows results over the sample without the pandemic, while Figure \ref{direct_wc_financial_crisis} provides a closer look at the fit during the financial crisis.
+Table \ref{tab:pvaluedhp} shows p-values for tests of the joint null hypothesis that all coefficients on the indicators are zero (HAC-adjusted Wald test).
 
@@
 Additionally, the p-values of the HAC-adjusted Wald test precipitously increase from $h=2$ to $h=3$ (see Table \eqref{tab:pvaluedhp}), implying that the indicators become statistically insignificant for forecast horizons of more than two quarters.\footnote{
-These findings are confirmed by the relative root mean-square errors comparing direct forecasts against the mean benchmark, see Table \eqref{rRMSE_mSSA_direct_mean_without_covid8} in the Appendix.}
+These findings are confirmed by the relative root mean-square errors comparing direct forecasts against the mean benchmark, see Table \ref{rRMSE_mSSA_direct_mean_without_covid8} in the Appendix.}
 
@@
 \begin{table}[h]
+\caption{p-values for $H_0: \boldsymbol{\beta} = 0$ in \eqref{eq:direct_forecast} based on HAC-adjusted Wald test.\\Estimation over full sample (without pandemic.)} 
+\label{tab:pvaluedhp}
 \centering
@@
 \end{tabular}
-\caption{p-values for $H_0: \boldsymbol{\beta} = 0$ in \eqref{eq:direct_forecast} based on HAC-adjusted Wald test.\\Estimation over full sample (without pandemic.)} 
-\label{tab:pvaluedhp}
+
 \end{table}
@@
 
-Figure \eqref{hp_160} presents both classic two-sided and one-sided HP(160) (denoted HP-C for \textit{concurrent}) filters (top panels) along with their associated amplitude functions in the frequency domain (bottom panels).
+Figure \ref{hp_160} presents both classic two-sided and one-sided HP(160) (denoted HP-C for \textit{concurrent}) filters (top panels) along with their associated amplitude functions in the frequency domain (bottom panels).
 It confirms that the HP-C filter assigns greater weight to yearly components—relevant in a medium-term forecast context— than the classic quarterly HP filter, while also reducing undesirable high-frequency noise.
@@
 
-Figures \eqref{direct_hp_forecasts} and \eqref{direct_hp_forecasts_financial_crisis} provide results comparable to those of Figures \eqref{direct_wc} and \eqref{direct_wc_financial_crisis} but now using indicators smoothed with the one-sided HP(160), (denoted as the HP-C, for \textit{concurrent}) filter.
+Figures \ref{direct_hp_forecasts} and \ref{direct_hp_forecasts_financial_crisis} provide results comparable to those of Figures \ref{direct_wc} and \ref{direct_wc_financial_crisis} but now using indicators smoothed with the one-sided HP(160), (denoted as the HP-C, for \textit{concurrent}) filter.
 The fitted values lag GDP slightly less and track turning points somewhat more quickly.
-Table \eqref{tab:pvaluedhp} shows p-values for tests of the joint null hypothesis that all coefficients ${\boldsymbol{\beta}} = 0$.
+Table \ref{tab:pvaluedhp} shows p-values for tests of the joint null hypothesis that all coefficients ${\boldsymbol{\beta}} = 0$.
 The use of filtered indicators improves their statistical significance at forecast horizons $h > 2$ quarters.
-%Figures \eqref{direct_hp_forecasts} (all observations with exclusion of the pandemic) and \eqref{direct_hp_forecasts_financial_crisis} (financial crisis).
+%Figures \ref{direct_hp_forecasts} (all observations with exclusion of the pandemic) and \ref{direct_hp_forecasts_financial_crisis} (financial crisis).
 A comparison with the classic (unfiltered) direct forecasts in the latter figure %in Fig.\eqref{direct_wc_financial_crisis}
 % indicates that the new predictors are slightly less retarded and track the turning points more promptly (`faster'). 
-% The filtered approach also exhibits superior statistical properties. As shown in Table \eqref{tab:f_stats}, 
+% The filtered approach also exhibits superior statistical properties. As shown in Table \ref{tab:f_stats}, 
 %The p-values of the F-statistics for the HP-based forecasts are smaller than for the direct forecasts specifically at $h=3$, but the predictor remains statistically insignificant at forecast horizons exceeding two quarters (even though the evaluation is conducted in-sample and based on all available indicators).
@@
 The finite sample $\mathbf{x}_t$ is the portion effectively observed and represents the log-differenced GDP (the dependence on $x_0, x_{-1},...$ is neglected because the HP filter weights decay rapidly to zero; see Fig.
-\eqref{hp_160}).
+\ref{hp_160}).
 The prediction task comprises tracking $z_{t}$  using the (causal) predictor $y_{t}=\sum_{k=0}^{L-1} b_{k}x_{t-k}=\mathbf{b}'\mathbf{x}_t$, where the filter $\mathbf{b}$ with weights $b_0,...,b_{L-1}'$ (actually of length $L=t$) is available for optimization purposes.
@@
 
-Traditional predictors often yield an excessive number of false sign changes due to leakage of high-frequency components (high-frequency leakage) in the predicted series, as discussed above (see Figure \eqref{hp_160}).
+Traditional predictors often yield an excessive number of false sign changes due to leakage of high-frequency components (high-frequency leakage) in the predicted series, as discussed above (see Figure \ref{hp_160}).
 This noisiness is particularly problematic near the onset or termination of economic crises (recessions), where the target fluctuates around zero; reliable tracking of sign changes would facilitate real-time assessment of the occurrence and timing of such events.
-%as illustrated by the amplitude function in Figure \eqref{hp_160} (bottom right panel) and further elaborated upon by \cite{Wildi2025}. 
- \cite{Wildi2025} shows that the SSA criterion helps control this phenomenon (see Fig.\eqref{mssa_msse_zc} for illustration).
-Using the dual formulation of the criterion \eqref{critssaht}, the author demonstrates that the SSA solution yields the \emph{lowest} zero-crossing rate among all (linear) predictors with equivalent target correlation.
-The concept can be extended to a multivariate framework, denoted M-SSA, by generalizing the correlation functions as detailed in the next section.\\
+%as illustrated by the amplitude function in Figure \ref{hp_160} (bottom right panel) and further elaborated upon by \cite{Wildi2025}. 
+ \cite{Wildi2025} shows that the SSA criterion helps control this phenomenon (see Fig.\ref{mssa_msse_zc} for illustration).\\ Using the dual formulation of the criterion \eqref{critssaht}, the author demonstrates that the SSA solution yields the \emph{lowest} zero-crossing rate among all (linear) predictors with equivalent target correlation.
+The concept can be extended to a multivariate framework, denoted M-SSA, by generalizing the correlation functions as detailed in the next section.
+As we shall see, controlling the zero-crossing rate within the M-SSA framework will have a quantifiable effect on the hit rate and false positive rate, thereby explicitly addressing the detection of present (nowcasting) and future (forecasting) economic contractions and expansions.\\
 
@@
 Let $\mathbf{X}_t$ (of dimension $t\times n$) denote a set of $n$ explanatory series $\mathbf{x}_{1t},...,\mathbf{x}_{nt}$, with observations $\mathbf{x}_{it}=(x_{it},...,x_{i1})'$.
-Let $\mathbf{Z}_t$, of dimension $t\times m$  a set of $m$ target series $\mathbf{z}_{it}=(z_{it},...,z_{i1})'$, $i=1,...m$, which typically lie outside the linear space spanned by   $\mathbf{X}_t$ (and are unknown at time $t$ but not independent of $\mathbf{X}_t$).
+Let $\mathbf{Z}_t$, of dimension $t\times m$, represent  a set of $m$ target series $\mathbf{z}_{it}=(z_{it},...,z_{i1})'$, $i=1,...m$, which typically lie outside the linear space spanned by   $\mathbf{X}_t$ (and are unknown at time $t$ but not independent of $\mathbf{X}_t$).
 In this framework, $\mathbf{X}_t$ represents a matrix of selected economic indicators for the German economy, with $n=5$.
@@
 \begin{table}[ht]
+\caption{Sample target correlations and holding times (HT) of M-SSA and M-MSE nowcasts.} 
+\label{corhtnow}
 \centering
@@
 \end{tabular}
-\caption{Sample target correlations and holding times (HT) of M-SSA and M-MSE nowcasts.} 
-\label{corhtnow}
+
 \end{table}
@@
 \begin{table}[ht]
+
+\caption{rRMSEs: out-of-sample forecast errors Equation \eqref{oosfe} vs expanding mean of GDP\\Out-of-sample evaluation starting in Q1-2008, without pandemic\\Values $< 1$ indicate superior predictions from Equation \eqref{oos_pred}\\
+Row-index: Shift refers to the left-shift of the target $\textrm{GDP}_{t+shift}$\\
+Column-index: $h$ refers to the forecast horizon for which $\textrm{M-SSA}_t^{\textrm{GDP}}(h)$ is optimized\\
+Significance refers to the hypothesis $\alpha_1=0$ in Equation \eqref{eq_gdp_pred}.
+} 
+\label{rRMSE_mSSA_comp_mean_wc}
 \centering
@@
 \end{tabular}
-\caption{rRMSEs: out-of-sample forecast errors Equation \eqref{oosfe} vs expanding mean of GDP\\Out-of-sample evaluation starting in Q1-2008, without pandemic\\Values $< 1$ indicate superior predictions from Equation \eqref{oos_pred}\\
-Row-index: Shift refers to the left-shift of the target $\textrm{GDP}_{t+shift}$\\
-Column-index: $h$ refers to the forecast horizon for which $\textrm{M-SSA}_t^{\textrm{GDP}}(h)$ is optimized\\
-Significance refers to the hypothesis $\alpha_1=0$ in Equation \eqref{eq_gdp_pred}.
-} 
-\label{rRMSE_mSSA_comp_mean_wc}
 \end{table}
@@
 % Wed May 21 10:06:12 2025
+
 \begin{table}[ht]
+\caption{AUCs of predictors against GDP at various forecast horizons.} 
+\label{auc}
 \centering
@@
 \end{tabular}
-\caption{AUCs of predictors against GDP at various forecast horizons.} 
-\label{auc}
-\end{table}At forecast horizons less than or equal to two quarters, the direct forecast challenges the filters.
+
+\end{table}
+
+At forecast horizons less than or equal to two quarters, the direct forecast challenges the filters.
 However, for horizons $h>2$, the filters outperform the direct approach: the univariate HP-C design is surpassed by the classic multivariate MSE filter, which is in turn dominated by the M-SSA.
@@
 \begin{table}[ht]
+\caption{Regression of shifted GDP on M-SSA($h$) GDP component.\\rRMSEs benchmarked against the direct forecast.\\Out-of-sample evaluation starting in Q1-2008, without pandemic.} 
+\label{rRMSE_mSSA_comp_direct_without_covid7}
 \centering
@@
 \end{tabular}
-\caption{Regression of shifted GDP on M-SSA($h$) GDP component.\\rRMSEs benchmarked against the direct forecast.\\Out-of-sample evaluation starting in Q1-2008, without pandemic.} 
-\label{rRMSE_mSSA_comp_direct_without_covid7}
+
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
@@
 \begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrrr}
-  \hline
- & Shift=0 & Shift=1 & Shift=2 & Shift=3 & Shift=4 & Shift=5 \\ 
-  \hline
-rRMSE & 0.862 & 0.838 & 0.912 & 1.008 & 1.001 & 1.028 \\  
-   \hline
-\end{tabular}
 \caption{Direct forecasts.\\
@@
 \label{rRMSE_mSSA_direct_mean_without_covid8}
+\centering
+\begin{tabular}{rrrrrrr}
+  \hline
+ & Shift=0 & Shift=1 & Shift=2 & Shift=3 & Shift=4 & Shift=5 \\ 
+  \hline
+rRMSE & 0.862 & 0.838 & 0.912 & 1.008 & 1.001 & 1.028 \\  
+   \hline
+\end{tabular}
+
 \end{table}
@@
 \begin{table}[ht]
+\caption{p-values for $H_0: {\alpha_1} = 0$, HAC-adjusted Wald test, Equation \eqref{eq_gdp_pred}.\\ including the pandemic.} 
+\label{p_val1}
 \centering
@@
 \end{tabular}
-\caption{p-values for $H_0: {\alpha_1} = 0$, HAC-adjusted Wald test, Equation \eqref{eq_gdp_pred}.\\ including the pandemic.} 
-\label{p_val1}
+
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
@@
 \begin{table}[ht]
+\caption{rRMSEs: out-of-sample forecast errors Equation \eqref{oosfe} vs expanding mean of GDP\\Out-of-sample evaluation starting in Q1-2008, including the pandemic.\\Values $< 1$ indicate superior predictions from Equation \eqref{oos_pred}.}
+\label{rRMSE_mSSA_comp_mean2}
 \centering
@@
 \end{tabular}
-\caption{rRMSEs: out-of-sample forecast errors Equation \eqref{oosfe} vs expanding mean of GDP\\Out-of-sample evaluation starting in Q1-2008, including the pandemic.\\Values $< 1$ indicate superior predictions from Equation \eqref{oos_pred}.}
-\label{rRMSE_mSSA_comp_mean2}
+
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
@@
 \begin{table}[ht]
+\caption{Regression of shifted GDP on M-SSA($h$) GDP component.\\rRMSEs benchmarked against the direct forecast.\\Out-of-sample evaluation starting in Q1-2008, including the pandemic.} 
+\label{rRMSE_mSSA_comp_direct3}
 \centering
@@
 \end{tabular}
-\caption{Regression of shifted GDP on M-SSA($h$) GDP component.\\rRMSEs benchmarked against the direct forecast.\\Out-of-sample evaluation starting in Q1-2008, including the pandemic.} 
-\label{rRMSE_mSSA_comp_direct3}
+
 \end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
@@
 \begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrrr}
-  \hline
- & Shift=0 & Shift=1 & Shift=2 & Shift=3 & Shift=4 & Shift=5 \\ 
-  \hline
-rRMSE & 0.812 & 1.075 & 0.992 & 0.977 & 1.021 & 1.031 \\ 
-   \hline
-\end{tabular}
 \caption{Direct forecasts.\\
@@
 \label{rRMSE_mSSA_direct_mean4}
+\centering
+\begin{tabular}{rrrrrrr}
+  \hline
+ & Shift=0 & Shift=1 & Shift=2 & Shift=3 & Shift=4 & Shift=5 \\ 
+  \hline
+rRMSE & 0.812 & 1.075 & 0.992 & 0.977 & 1.021 & 1.031 \\ 
+   \hline
+\end{tabular}
+
 \end{table}
 
+\begin{figure}[htpb]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/CCF.pdf}\caption{Sample cross-correlation function (CCF): full data set.\label{CCF}}\end{center}\end{figure}
+\begin{figure}[htpb]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/CCF_wc.pdf}\caption{Sample cross-correlation function (CCF): without Pandemic.\label{CCF_wc}}\end{center}\end{figure}
+
 \end{document}

### 2025-08-15 Archive IWH Simon paper/12.08.2025/paper.tex
up

--- 2025-08-11 Archive IWH Simon paper/10.08.2025/paper.tex
@@
 \begin{abstract} %max.
-This paper explores the trade-off between informational efficiency and forecast smoothness in economic forecasting.
-Forecasts that minimize mean squared forecast errors often exhibit excessive volatility, limiting their practical applicability.
-We propose the Multivariate Smooth Sign Accuracy (M-SSA) framework as a solution, extracting smoothed components from leading indicators.
-This enhances the signal-to-noise ratio and controls forecast volatility and timing.
-Applied to German GDP growth, our method yields smoothed forecasts that improve the forecasting accuracy of traditional models, particularly over medium-term horizons.
-While smoother forecasts tend to lag slightly, this drawback can be mitigated by adjusting the forecast horizon.
-These findings highlight the practicality of the M-SSA framework for both forecasters and policymakers, especially in settings where updating forecasts or policy adjustments are costly.
+Forecasts that minimize mean squared forecast error (MSE) often exhibit excessive volatility, limiting their practical applicability.
+We address this accuracy–smoothness trade-off by introducing a Multivariate Smooth Sign Accuracy (M-SSA) framework, which extracts smoothed components from leading indicators to enhance the signal-to-noise ratio and control the forecast volatility and timing.
+Applied to quarterly German GDP growth, our method yields smoothed forecasts that improve the forecasting accuracy of traditional models, particularly over medium-term horizons.
+While smoother forecasts tend to lag slightly in turning points, adjusting the forecast horizon offsets this effect.
+These findings highlight the practicality of the M-SSA framework for both forecasters and policymakers, especially in settings where forecast revisions or policy adjustments are costly.
 
@@
 \bigskip \vspace{10pt}
-\textbf{Keywords:} Forecast Smoothing, Smooth Sign Accuracy, Time Series Filtering 
+\textbf{Keywords:} Forecast Smoothing, Smooth Sign Accuracy, Time-Series Filtering 
 \newline
@@
  In addition, forecasts may reflect excessive ``noise'' due to overfitting.
-Further, forecast users (including public and private decision makers) may face constraints in their speed of adjustment to changing forecasts.
+At the same time, forecast users (including public and private decision makers) may face constraints in their speed of adjustment to changing forecasts.
 For example, fixed annual budgets may constrain spending, or transaction costs may limit changes in investments.
@@
 
-\cite{Wildi2024,Wildi2025} and \cite{McElroy2019,McElroy2020} propose novel methods that optimally smooth forecasts by controlling the expected frequency of sign changes.
+Recent work by \cite{Wildi2024,Wildi2025} and \cite{McElroy2019,McElroy2020} formalipropose novel methods that optimally smooth forecasts by controlling the expected frequency of sign changes.
 The resulting Smooth Sign Accuracy (SSA) framework encompasses efficiently the mean-square error (MSE) as a special case and produces increasingly smooth forecasts as changes in the forecast direction (sign) are increasingly penalized.
@@
 \subsection{Dependence}
-To examine the joint dynamics of our variables, we estimate their cross-correlations as well as their vector moving-average (VMA) representations.
+To examine the joint dynamics of our selected variables, we estimate their cross-correlations as well as their vector moving-average (VMA) representations.
 The latter is a key input to the multivariate filters we develop in Sections \ref{sec:mSSA} and \ref{sec:mSSA_component}.
@@
 This led to a VAR(3) specification with 7 steps.
-For additional details, see \cite{ElasticNet} and the references therein.} The figures \ref{ma_inv_BIP} and \ref{ma_inv_BIP_var3_en} show the implied responses of GDP to shocks to each indicator variable.\footnote{Shown are the MA coefficients for GDP from the vector moving-average (VMA) models obtained by inversion of the VAR(1) and VAR(3) representations respectively (i.e.
-the impulse-response functions (IRFs) of GDP to the correlated reduced-form shocks.)}
+For additional details, see \cite{ElasticNet} and the references therein.} The Figures \ref{ma_inv_BIP} and \ref{ma_inv_BIP_var3_en} show the implied responses of GDP to shocks to each indicator variable.\footnote{Shown are the MA coefficients for GDP from the vector moving-average (VMA) models obtained by inversion of the VAR(1) and VAR(3) representations respectively (i.e.
+the impulse-response functions (IRFs) of GDP to the correlated reduced-form shocks).}
 
@@
 %
-%
-
-\begin{figure}[H]
+
+\begin{figure}[htpb]
+    \begin{subfigure}[htpb]{\textwidth}
         \begin{center}
             \includegraphics[height=3in, width=6in]{./Figures/ma_inv_BIP.pdf}
-        \caption{VMA coefficients for GDP implied by VAR(1)\\Estimation with pandemic (left) and without pandemic (right).
-        \label{ma_inv_BIP}}\end{center}
-\end{figure}
-
-\begin{figure}[H]
+            \caption{implied by VAR(1) \label{ma_inv_BIP}}
+        \end{center}
+    \end{subfigure}
+    \\
+    \begin{subfigure}[htpb]{\textwidth}
         \begin{center}
             \includegraphics[height=3in, width=6in]{./Figures/ma_inv_BIP_var3_en.pdf}
-        \caption{VMA coefficients for GDP implied by VAR(3)\\Estimation with pandemic (left) and without pandemic (right).
-        \label{ma_inv_BIP_var3_en}}\end{center}
+            \caption{implied by VAR(3)\label{ma_inv_BIP_var3_en}}
+        \end{center}
+    \end{subfigure}
+    \caption{VMA coefficients for GDP.\\Estimation with pandemic (left) and without pandemic (right).}
 \end{figure}
@@
 In this section, we demonstrate how pre-filtering indicator variables may improve the performance of simple forecasts by removing noise.
-We then demonstrate how in-sample forecast performance improves when we ``de-noise'' the indicators prior to projection with simple smoothing filters.\footnote{The smoothing filters are one-sided (concurrent) filters of length 31, therefore, estimation of both forecasting models omits the first 30 observations.}
+We demonstrate how in-sample forecast performance improves when we ``de-noise'' the indicators prior to projection with simple smoothing filters.\footnote{The smoothing filters are one-sided (concurrent) filters of length 31, therefore, estimation of both forecasting models omits the first 30 observations.}
 
@@
 \begin{equation}
-    u_{t+h} = \boldsymbol{\beta'} \cdot \mathbf{X}_t + \varepsilon_t
+    y_{t+h} = \boldsymbol{\beta'} \cdot \mathbf{X}_t + \varepsilon_t
     \label{eq:direct_forecast}
@@
 \begin{itemize}
-    \item $u_{t+h}$ is the quarterly growth rate of GDP for forecast horizon $h = \{0, 1, 2, ... \}$
-    \item $\mathbf{X}_t \equiv [1,u_{t-1}, ip_t, ifo\_c_t, ESI_t, spr\_10y\_3m_t]'$ or their transformed counterparts and $u_{t-1}$ accounts for the publication lag of GDP.\footnote{Including lagged GDP enables the nesting of the AR forecast within the broader modeling framework.}
+    \item $y_{t+h}$ is the quarterly growth rate of GDP for forecast horizon $h = \{0, 1, 2, ... \}$ (labeled shifted GDP in the following )
+    \item $\mathbf{X}_t \equiv [1,y_{t-1}, ip_t, ifo\_c_t, ESI_t, spr\_10y\_3m_t]'$ or their transformed counterparts and $y_{t-1}$ accounts for the publication lag of GDP.\footnote{Including lagged GDP enables the nesting of the AR forecast within the broader modeling framework.}
     \item $\hat{\boldsymbol{\beta}}$ is estimated by least-squares, omitting observations affected by the pandemic, i.e. Q4-2019 to Q4-2020.
@@
 
-\begin{figure}[H]
+\begin{figure}[htbp]
+
+ \begin{subfigure}[htpb]{\textwidth}
         \begin{center}
              \includegraphics[width=0.9\textwidth]{./Figures/direct_wc_all.pdf}
-        %\includegraphics[height=3in, width=4.5in]{./Figures/direct_wc_all.pdf}
-        \caption{Direct forecasts (red lines) based on all indicators and full sample information (without pandemic.)\\
-        Target (black line): GDP left-shifted  by 0 - 3 quarters.
-        \label{direct_wc}}
-    \end{center}
+            \caption{Full sample information (without pandemic)\label{direct_wc}}
+        \end{center}
+    \end{subfigure}
+
+ \begin{subfigure}[htpb]{\textwidth}
+        \begin{center}
+            \includegraphics[width=0.9\textwidth]{./Figures/direct_wc_financial_crisis.pdf}
+            \caption{Financial crisis (without pandemic)\label{direct_wc_financial_crisis}}
+        \end{center}
+    \end{subfigure}
+    \caption{Direct forecasts (red lines) based on all indicators and target (black line): GDP left-shifted  by 0 - 3 quarters.}
 \end{figure}
@@
 
-\begin{figure}[H]
-    \begin{center}
-        \includegraphics[width=0.9\textwidth]{./Figures/direct_wc_financial_crisis.pdf}
-        \caption{Direct forecast across the financial crisis:  regression based on all indicators and full sample information, without pandemic.
-        \label{direct_wc_financial_crisis}}
-    \end{center}
-\end{figure}
-
-The results indicate that as the forecast horizon $h$ increases, forecast become less volatile and their turning points (i.e., peaks and troughs) become increasingly right-shifted (lagged) relative to the target GDP.
-Additionally, the p-values of the HAC-adjusted Wald test precipitously increase from $h=2$ to $h=3$ (see Table \eqref{tab:pvaluedhp}), implying that the indicators become statistically insignificant for forecast horizons of more than two quarters.\footnote{
+The results indicate that as the forecast horizon $h$ increases, forecasts become less volatile and their turning points (i.e., peaks and troughs) become increasingly right-shifted (lagged) relative to the target GDP.
+Additionally, the p-values of the HAC-adjusted Wald test precipitously increase from $h=2$ to $h=3$ (see Table \ref{tab:pvaluedhp}), implying that the indicators become statistically insignificant for forecast horizons of more than two quarters.\footnote{
 These findings are confirmed by the relative root mean-square errors comparing direct forecasts against the mean benchmark, see Table \ref{rRMSE_mSSA_direct_mean_without_covid8} in the Appendix.}
@@
 \begin{table}[h]
-\caption{p-values for $H_0: \boldsymbol{\beta} = 0$ in \eqref{eq:direct_forecast} based on HAC-adjusted Wald test.\\Estimation over full sample (without pandemic.)} 
+\caption{p-values for $H_0: \boldsymbol{\beta} = 0$ in eq.~\eqref{eq:direct_forecast} based on HAC-adjusted Wald test.\\Estimation over full sample (without pandemic.)} 
 \label{tab:pvaluedhp}
@@
         % \caption{Two-sided HP(160) (top left) and one-sided (concurrent) HP-C filters (top right) with corresponding amplitude functions (bottom).
-        \caption{HP$(\lambda=160)$ filters\\
+        \caption{HP$(\lambda=160)$ filters.\\
         Left: HP (Two-sided, symmetric) \qquad Right: HP-C (One-sided, concurrent)\\
@@
 
-Figures \ref{direct_hp_forecasts} and \ref{direct_hp_forecasts_financial_crisis} provide results comparable to those of Figures \ref{direct_wc} and \ref{direct_wc_financial_crisis} but now using indicators smoothed with the one-sided HP(160), (denoted as the HP-C, for \textit{concurrent}) filter.
+Figures \ref{direct_hp_forecasts} and \ref{direct_hp_forecasts_financial_crisis} provide results comparable to those of Figures \ref{direct_wc} and \ref{direct_wc_financial_crisis} but now using indicators smoothed with the one-sided HP(160), (denoted as the HP-C, for \textit{concurrent} filter).
 The fitted values lag GDP slightly less and track turning points somewhat more quickly.
@@
         \includegraphics[width=0.85\textwidth]{./Figures/direct_hp_forecasts.pdf}
-        \caption{Direct HP forecasts: entire data set with exclusion of the pandemic.
+        \caption{Direct HP-C forecasts: entire data set (without pandemic).
         \label{direct_hp_forecasts}}
@@
 %its univariate nature excludes additional information such as that provided by leading indicators. 
-the noise suppression capability of the one-sided filter is compromised, as explained by \cite{Wildi2025} and illustrated by the high-frequency leakage of the amplitude function in Figure \eqref{hp_160}.
+the noise suppression capability of the one-sided filter is compromised, as explained by \cite{Wildi2025} and illustrated by the high-frequency leakage of the amplitude function in Figure \ref{hp_160}.
 To address these limitations, we turn to the Multivariate Smooth Sign Accuracy (M-SSA) framework proposed by \cite{Wildi2025}.
@@
 For simplicity, we assume stationarity of all time series involved.
-In our current framwork, $z_t$ corresponds to the output of a two-sided Hodrick–Prescott filter with smoothing parameter $\lambda = 160$ applied to the doubly infinite sequence $x_k$, $k\in\mathbb{Z}$.
-The finite sample $\mathbf{x}_t$ is the portion effectively observed and represents the log-differenced GDP (the dependence on $x_0, x_{-1},...$ is neglected because the HP filter weights decay rapidly to zero; see Fig.
-\ref{hp_160}).
+In our current framework, $z_t$ corresponds to the output of a two-sided Hodrick–Prescott filter with smoothing parameter $\lambda = 160$ applied to the doubly infinite sequence $x_k$, $k\in\mathbb{Z}$.
+The finite sample $\mathbf{x}_t$ is the portion effectively observed and represents the log-differenced GDP (the dependence on $x_0, x_{-1},...$ is neglected because the HP filter weights decay rapidly to zero; see Figure \ref{hp_160}).
 The prediction task comprises tracking $z_{t}$  using the (causal) predictor $y_{t}=\sum_{k=0}^{L-1} b_{k}x_{t-k}=\mathbf{b}'\mathbf{x}_t$, where the filter $\mathbf{b}$ with weights $b_0,...,b_{L-1}'$ (actually of length $L=t$) is available for optimization purposes.
@@
 %as illustrated by the amplitude function in Figure \ref{hp_160} (bottom right panel) and further elaborated upon by \cite{Wildi2025}. 
- \cite{Wildi2025} shows that the SSA criterion helps control this phenomenon (see Fig.\ref{mssa_msse_zc} for illustration).\\ Using the dual formulation of the criterion \eqref{critssaht}, the author demonstrates that the SSA solution yields the \emph{lowest} zero-crossing rate among all (linear) predictors with equivalent target correlation.
+ \cite{Wildi2025} shows that the SSA criterion helps control this phenomenon (Figure \ref{mssa_msse_zc}).
+
+ Using the dual formulation of the criterion \eqref{critssaht}, the author demonstrates that the SSA solution yields the \emph{lowest} zero-crossing rate among all (linear) predictors with equivalent target correlation.
 The concept can be extended to a multivariate framework, denoted M-SSA, by generalizing the correlation functions as detailed in the next section.
-As we shall see, controlling the zero-crossing rate within the M-SSA framework will have a quantifiable effect on the hit rate and false positive rate, thereby explicitly addressing the detection of present (nowcasting) and future (forecasting) economic contractions and expansions.\\
-
-We now briefly elucidate the SSA optimization criterion outlined in Eq.
-\eqref{critssa}.
+As we shall see, controlling the zero-crossing rate within the M-SSA framework will have a quantifiable effect on the hit rate and false positive rate, thereby explicitly addressing the detection of present (nowcasting) and future (forecasting) economic contractions and expansions.
+
+We now briefly elucidate the SSA optimization criterion outlined in Eq.\eqref{critssa}.
 Consider a simplified univariate framework where $x_t=\epsilon_t$ represents white noise with variance $\sigma^2$ and assume, for notational simplicity, that $\sigma^2=1$.
@@
 Let  $y_{h t}=\sum_{k=0}^{L-1} b_{h k}x_{t-k}$  denote a causal predictor of length $L$.
-For notational convenience we now omit the subscript $h$ from the predictor and we denote $\boldsymbol{\gamma}_h:==(\gamma_{h},\gamma_{h+1},...,\gamma_{h+L-1})'$.
+For notational convenience, we now omit the subscript $h$ from the predictor $\mathbf{b}=\mathbf{b}_h$ and we denote $\boldsymbol{\gamma}_h:=(\gamma_{h},\gamma_{h+1},...,\gamma_{h+L-1})'$.
 % To simplify the exposition, we consider the case $h=0$ (nowcast), and thus omit the subscript $h$ from the notation.
-Then, under the postulated white noise assumption we obtain:
+Then, under the postulated white noise assumption, we obtain:
 \begin{eqnarray*}
@@
 \end{eqnarray}
-where the additional length constraint $\mathbf{b}'\mathbf{b}=l$ serves two purposes: firstly, to ensure the uniqueness of the SSA solution; secondly, to normalize the objective function $\mathbf{b}'\boldsymbol{\gamma}_{h}$, making it proportional to the target correlation $\rho(z,y)$, up to a fixed scaling factor $\sqrt{\mathbf{b}'\mathbf{b}}\sqrt{\boldsymbol{\gamma}_{\infty}'\boldsymbol{\gamma}_{\infty}}=\sqrt{l}\sqrt{\boldsymbol{\gamma}_{\infty}'\boldsymbol{\gamma}_{\infty}}$ (for convenience we often select $l=1$, leading to a standardized $y_t$).
-Criterion \eqref{crit1} is readily amenable to optimization, as it involves affine and quadratic functions of $\mathbf{b}$, see Wildi (2024) for background (detailed derivations are obtained in Wildi (2025)).
-\\
+where the additional length constraint $\mathbf{b}'\mathbf{b}=l$ serves two purposes: firstly, to ensure the uniqueness of the SSA solution; secondly, to normalize the objective function $\mathbf{b}'\boldsymbol{\gamma}_{h}$, making it proportional to the target correlation $\rho(z,y)$, up to a fixed scaling factor $\sqrt{\mathbf{b}'\mathbf{b}}\sqrt{\boldsymbol{\gamma}_{\infty}'\boldsymbol{\gamma}_{\infty}}=\sqrt{l}\sqrt{\boldsymbol{\gamma}_{\infty}'\boldsymbol{\gamma}_{\infty}}$.
+For convenience, $l$ is typically set to 1, yielding a standardized $y_t$.
+Consequently, the solution to the criterion is invariant to the choice between these two objective formulations.
+However,
+Criterion \eqref{crit1} is readily amenable to optimization, as it involves affine and quadratic functions of $\mathbf{b}$ \citep{Wildi2024}.\footnote{Detailed derivations are obtained in \cite{Wildi2025}.} \\
 Under the above assumptions, the classical MSE predictor of length $L$ can be expressed via orthogonal projection as:
@@
 which corresponds to setting $b_{MSE,k}=\gamma_{k+h}$, $k=0,...,L-1$.
-In general, to enhance the smoothness of the SSA predictor $y_t$, we select $\rho_1>\rho_{MSE}$ in criterion \eqref{crit1},  where $\rho_{MSE}$ denotes the first-order ACF of the (benchmark) MSE predictor\footnote{Selecting $\rho_1<\rho_{MSE}$ would produce an SSA predictor with more frequent zero-crossings than the MSE benchmark, which is typically undesirable in practical applications.}.
-Finally,  an extension to autocorrelated $x_t\neq \epsilon_t$ is discussed below.
+In general, to enhance the smoothness of the SSA predictor $y_t$, we select $\rho_1>\rho_{MSE}$ in criterion \eqref{crit1},  where $\rho_{MSE}$ denotes the first-order ACF of the (benchmark) MSE predictor.\footnote{Selecting $\rho_1<\rho_{MSE}$ would produce an SSA predictor with more frequent zero-crossings than the MSE benchmark, which is typically undesirable in practical applications.} Finally,  an extension to autocorrelated $x_t\neq \epsilon_t$ is discussed below.
 
@@
 
-We can now extend the SSA criterion \eqref{crit1} to a multivariate setting.
+The SSA criterion \eqref{crit1} is now extended to a multivariate setting.
 Let $\mathbf{X}_t$ (of dimension $t\times n$) denote a set of $n$ explanatory series $\mathbf{x}_{1t},...,\mathbf{x}_{nt}$, with observations $\mathbf{x}_{it}=(x_{it},...,x_{i1})'$.
@@
 \] 
-where $\boldsymbol{\Gamma}_k$  is a sequence of filter matrices of dimension  $n\times n$ (hence we assume $m=n$), satisfying the square-summability condition.
+where $\boldsymbol{\Gamma}_k$  is a sequence of filter matrices of dimension  $n\times n$, satisfying the square-summability condition.
 Each $\boldsymbol{\Gamma}_k$ has entries $(\gamma_{ijk})$ for $i, j = 1, \ldots, n$.
@@
 
-Finally, we can relax the WN hypothesis and assume a stationary process 
+Finally, we can relax the white noise hypothesis and assume a stationary process 
 \begin{equation}\label{x_m_st}
@@
  \] 
-We now assume $h\geq 0$\footnote{Similar but slightly more complex expressions would arise for $h < 0$ (backcasting), which are omitted here.} and $L$ to be sufficiently large to render $(\boldsymbol{\Gamma} \cdot \boldsymbol{\Xi})_{h+k} \approx 0$ and $(\mathbf{B} \cdot \boldsymbol{\Xi})_j \approx 0$ for $k,j \geq L$.
+We now assume $h\geq 0$ (note: analogous but slightly more complex expressions arise for $h < 0$, i.e., backcasting, which are omitted here).
+In addition, $L$ should be chosen sufficiently large to ensure  $(\boldsymbol{\Gamma} \cdot \boldsymbol{\Xi})_{h+k} \approx 0$ and $(\mathbf{B} \cdot \boldsymbol{\Xi})_j \approx 0$ for $k,j \geq L$.
+This restriction is typically satisfied in practical applications; otherwise, the criterion can be generalized to accommodate atypical long-memory prediction problems (see Wildi 2025).
 We define:
@@
 In analogy to the above white noise case, the criterion involves affine and quadratic functions of $(\mathbf{b}\cdot\boldsymbol{\xi})_i$ and thus can be solved for the convolution weights.
-The requested M-SSA coefficients $\mathbf{b}_i$ can subsequently be determined through deconvolution, as indicated in Equation \eqref{con_inv}.
-The present discussion is confined to stationary processes, omitting a lengthier treatment of rank-defficient designs (cointegration) in the non-stationary case.\\
+The requested M-SSA coefficients $\mathbf{b}_i$ can subsequently be determined through deconvolution, as indicated in Eq.~\eqref{con_inv}.
+The present discussion is confined to stationary processes, omitting a lengthier treatment of rank-defficient designs (cointegration) in the non-stationary case.
 
@@
 
-\subsection{Example: Out of sample M-MSE vs. M-SSA Nowcasts}
+\subsection{Out of sample M-MSE vs. M-SSA Nowcasts}
 
 For illustration, consider the problem of nowcasting HP-filtered GDP, denoted by HP-GDP, where the target $z_t$ is generated using the two-sided HP(160) filter examined previously.
-We compare the classic multivariate mean squared error (MSE) predictor (denoted M-MSE) with the M-SSA predictor where we set $HT_1=1.5\times HT_{MSE}$.
+We compare the classic multivariate mean squared error (MSE) predictor (denoted M-MSE) with the M-SSA predictor, where we set $HT_1=1.5\times HT_{MSE}$.
 This constraint ensures that the M-SSA predictor produces roughly 33$\%$ fewer sign changes than M-MSE in large samples.
 Comparisons with M-MSE allow us to see the cost of this constraint in terms of prediction performance (target correlation).
-Both predictors use the same underlying VAR(1) model introduced in Section \eqref{sec:direct_forecast} and estimated on data up to Q4-2007.
+Both predictors use the same underlying VAR(1) model introduced in Section \ref{sec:direct_forecast} and estimated on data up to Q4-2007.
 This ensures a lengthy out-of-sample period that includes significant events such as the financial crisis, the sovereign debt crisis, and the COVID-19 pandemic.\footnote{The economic sentiment indicator (ESI) does not receive any weight in the parsimonious VAR(1) model estimated over the shorter sub-sample ending in Q4-2007.
@@
         \caption{Multivariate M-MSE (left) and M-SSA (right) nowcast filter weights.\\
-        Note difference in Scales.
+        Note the difference in scales.
         \label{bk_gammak}}
@@
 \begin{table}[ht]
-\caption{Sample target correlations and holding times (HT) of M-SSA and M-MSE nowcasts.} 
+\caption{Sample target correlations and holding times (HT) of M-SSA and M-MSE nowcasts. Full Sample (excl. pandemic).} 
 \label{corhtnow}
@@
 Figure \eqref{bk_gammak} displays the coefficients of the multivariate nowcasts, where M-SSA's increased smoothing reduces the maximum amplitudes of the filter weights and slows their rate of decay at longer lags.
-The resulting filtered series, presented in Figure \eqref{mssa_msse_now}, are standardized to enable direct comparisons.
+The resulting filtered series, presented in Figure \ref{mssa_msse_now}, are standardized to enable direct comparisons.
 The M-SSA nowcast is slightly smoother than that of the M-MSE, exhibiting approximately one-third fewer sign-changes.
-Fig.\eqref{mssa_msse_zc} compares both nowcasts in relation to the financial crisis, highlighting that the classic (MSE-based) design tends to produce unwanted `noisy' zero-crossings at (or in the vicinity of) the transitions between recessions and expansions.
-Further, Table \eqref{corhtnow} compares sample target correlations and HTs of the two nowcasts.
+Figure \ref{mssa_msse_zc} compares both nowcasts in relation to the financial crisis, highlighting that the classic (MSE-based) design tends to produce unwanted `noisy' zero-crossings at (or in the vicinity of) the transitions between recessions and expansions.
+Further, Table \ref{corhtnow} compares sample target correlations and HTs of the two nowcasts.
 Increased smoothness, i.e.
@@
 This validation is typically performed through analyses of very long samples of simulated (multivariate) data, ensuring the consistency of the estimation procedures under idealized conditions.} An
-increases by $50\%$ the HT relative to that of the M-MSE predictor\footnote{The convergence of the sample HT to the effectively imposed $HT_1$  can be verified in a simulation study, based on  very long samples of the time series, utilizing the M-SSA package.} but has only a modest effect on the target correlation, suggesting that this trade-off could be beneficial in our application
+increases by $50\%$ the HT relative to that of the M-MSE predictor\footnote{The convergence of the sample HT to the effectively imposed $HT_1$  can be verified in a simulation study, based on  very long samples of the time series, utilizing the M-SSA package.} but has only a modest effect on the target correlation, suggesting that this trade-off could be beneficial in our application.
 % , as confirmed by a subsequent analysis of hit rate and false alarm rate
@@
 
-\subsection{Example: M-SSA Forecasts}
+\subsection{M-SSA Forecasts}
 
@@
 %This approach facilitates the assessment of forecast performance across multiple horizons within the established methodological context.
-The influence of the forecast horizon on the M-SSA design is depicted in Figure \eqref{bk_h}, which compares M-SSA filter weights for the nowcast and the one-year-ahead forecast.
+The influence of the forecast horizon on the M-SSA design is depicted in Figure \eqref{bk_h}, which compares M-SSA filter weights for the nowcast and the one-year-ahead forecast($h=4$).
 Noting the difference in vertical scales, we see that as $h$ increases (right panel), the filter weights diminish, reflecting increased forecast uncertainty.
@@
         \includegraphics[width=0.85\textwidth]{./Figures/bk_h.pdf}
-        \caption{Multivariate M-SSA filter weights\\
+        \caption{Multivariate M-SSA filter weights.\\
         Left: $h=0$ (nowcast) \qquad Right: $h=4$ (one year forecast)\\
@@
 
-M-SSA predictors are compared to the traditional univariate HP-C in Figures  \eqref{multivar_vs_univar} and \eqref{mssa_hpc_financial_crisis}.
+M-SSA predictors are compared to the traditional univariate HP-C in Figures  \ref{multivar_vs_univar} and \ref{mssa_hpc_financial_crisis}.
 In the first figure, we compare both filters across the periods of the financial crisis (top panels) and the COVID-19 pandemic (bottom panels), for forecast horizons $h=0,...,6$.
@@
 %(this contrasts with Section \eqref{sec:direct_forecast}, which was based on the full sample information). 
-All the results that we present here exclude the pandemic; results including the pandemic are in the Appendix.\\
+All the results that we present here exclude the pandemic; results including the pandemic are available in the Appendix.\\
 
@@
 
-Let $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$ denote the M-SSA predictor of HP-GDP for forecast horizon $h$. A predictor of GDP can be obtained from the simple regression equation
+Let $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$ denote the M-SSA predictor of HP-GDP at time $t$ for forecast horizon $h$. A predictor of GDP can be obtained from the following regression specification
 \begin{equation}
@@
 This flexibility enables leveraging the dynamic left-shift of $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$, illustrated in Figure \eqref{multivar_vs_univar}, when $shift<h$.\\
-Define the (out-of-sample) GDP component predictor $\hat{\textrm{GDP}}_{t+shift}(h)$:  
+The corresponding out-of-sample predictor $\hat{\textrm{GDP}}_{t+shift}(h)$ is given by :  
 \begin{equation}\label{oos_pred}
@@
 \end{equation}
-where the regression parameters $\hat{\beta}_0(shift,h,t),\hat{\beta}_1(shift,h,t)$ are estimated using data up to time point $t-1$ (expanding window).
+where the regression parameters $\hat{\beta}_0(shift,h,t),\hat{\beta}_1(shift,h,t)$ are estimated recursively  using data up to time point $t-1$ (expanding window).
 Since M-SSA is based on data up to Q4-2007, $\hat{\textrm{GDP}}_{t+shift}(h)$ constitutes an out-of-sample predictor of $\textrm{GDP}_{t+shift}$ for all $shift\geq 0$ and all $t$ larger than Q4-2007.
 \footnote{Equations \eqref{eq:GDP_M-SSA} and \eqref{oos_pred} could be extended to incorporate additional M-SSA components, but for illustration purposes, we focus on the simplest (and most robust) design.}
-Define the out-of-sample forecast error
+The associated out-of-sample forecast error is defined as
 \begin{equation}\label{oosfe}
@@
 %We now assume that the regression parameters $\hat{\beta}_0(shift,h,t)$ and$,\hat{\beta}_1(shift,h,t)$ in Equation \eqref{oos_pred} are up-dated quarterly. Note that 
-While we keep the M-SSA outputs fixed, as based on data up to Q4-2007, we implement an expanding window starting from Q1-2008 for deriving the component weights, which are updated on a quarterly basis.
+In the empirical implementation, the M-SSA forecasts are held fixed, while the regression parameters  $\hat{\beta}_0(\cdot)$ and $\hat{\beta}_1(\cdot)$ are updated each quarter using an expanding window beginning in 2008Q1.
+This procedure yields a sequence of optimally weighted M-SSA forecasts that adapt over time to evolving GDP dynamics.
 
@@
 
-The statistical significance of the (out of sample) GDP component predictor in Equation \eqref{oos_pred}  can be evaluated by regressing 
-$\textrm{GDP}_{t+shift}$ on $\hat{\textrm{GDP}}_{t+shift}(h)$:
+The statistical significance of the (out of sample) GDP component predictor in Equation \eqref{oos_pred}  can be evaluated by:
 \begin{equation}\label{eq_gdp_pred}
@@
 \end{equation}
-for observations after Q4-2007: the significance of the corresponding HAC-adjusted test statistics for $H_0 : \alpha_1 =0$ is indicated by a `*' symbol in Table \eqref{rRMSE_mSSA_comp_mean_wc}\footnote{`*' indicates significance at the $5\%$ level; `**' indicates significance at the $0.1\%$ level.
-Note that statistical significance is influenced by the relatively short out-of-sample span, of length 17 years.}.
-The MSE forecast performance of $\hat{\textrm{GDP}}_{t+shift}(h)$ can be benchmarked against the mean predictor (see Table \eqref{rRMSE_mSSA_comp_mean_wc} which reports relative Root Mean-Square Errors, rRMSE) as well as against the direct forecast (see Table \eqref{rRMSE_mSSA_comp_direct_without_covid7}).
-For consistency, both benchmark predictors are also quarterly updated out-of-sample designs.\\
+for observations after Q4-2007.
+Statistical inference is based on heteroskedasticity- and autocorrelation-consistent (HAC) standard errors, with the null hypothesis $H_0 : \alpha_1 =0$.
+The significance levels of the corresponding test statistics are indicated by $^{*}$ in Table \ref{rRMSE_mSSA_comp_mean_wc}.\footnote{$^{*}$ denotes significance at the $5\%$ level; $^{**}$ indicates significance at the $0.1\%$ level.
+Note that statistical significance is influenced by the relatively short out-of-sample period of 17 years.} Forecast accuracy is evaluated in terms of relative root mean squared forecast errors (rRMSEs), where the MSE forecast performance of $\hat{\textrm{GDP}}_{t+shift}(h)$ can be benchmarked against the mean predictor (Table \ref{rRMSE_mSSA_comp_mean_wc}) as well as against the direct forecast (see Table \ref{rRMSE_mSSA_comp_direct_without_covid7}).
+For consistency, both benchmark predictors are quarterly updated out-of-sample designs.\\
 
@@
 Our results can be summarized as follows.
-When excluding the pandemic, the direct forecasts outperform the mean forecast up to two quarters ahead, see Table \eqref{rRMSE_mSSA_direct_mean_without_covid8} in the appendix.
+When excluding the pandemic, the direct forecasts outperform the mean forecast up to two quarters ahead, see Table \ref{rRMSE_mSSA_direct_mean_without_covid8} in the appendix.
 %\footnote{However, the pandemic period generally obscures the relationships between predictors and GDP.}.
-In contrast, the GDP component predictors $\hat{\textrm{GDP}}_{t+shift}(h)$, optimized for longer forecast horizons $h\geq 5$, outperform the mean benchmark up to 5Q ahead, see Table \eqref{rRMSE_mSSA_comp_mean_wc}.
-Additionally,  $\hat{\textrm{GDP}}_{t+shift}(h)$, optimized for large forecast horizons $h\geq 4$, surpass the direct forecasts for $shift\geq 3$, see Table \eqref{rRMSE_mSSA_comp_direct_without_covid7} in the appendix.
-Statistical significance of the GDP component predictor, optimized for larger forecast horizons, can be established for at least up to one year ahead, see Table \eqref{rRMSE_mSSA_comp_mean_wc}.
-Finally, the strong performance of predictors optimized for larger forecast horizons ($h\geq 4$), at smaller shifts ($shifts< 4$) corroborates the importance of the left-shift of the corresponding predictors, see Figs.\eqref{multivar_vs_univar} and \eqref{mssa_hpc_financial_crisis}, in relation to forecast performance.
+In contrast, the GDP predictors $\hat{\textrm{GDP}}_{t+shift}(h)$, optimized for longer forecast horizons $h\geq 5$, outperform the mean benchmark up to 5 quarters ahead, see Table \ref{rRMSE_mSSA_comp_mean_wc}.
+Additionally,  $\hat{\textrm{GDP}}_{t+shift}(h)$, optimized for large forecast horizons $h\geq 4$, surpass the direct forecasts for $shift\geq 3$, see Table \ref{rRMSE_mSSA_comp_direct_without_covid7} in the appendix.
+Statistical significance of the GDP predictor, optimized for larger forecast horizons, can be established for at least up to one year ahead, see Table \eqref{rRMSE_mSSA_comp_mean_wc}.
+Finally, the strong performance of predictors optimized for larger forecast horizons ($h\geq 4$), at smaller shifts ($shifts< 4$) corroborates the importance of the left-shift of the corresponding predictors, see Figures \ref{multivar_vs_univar} and \ref{mssa_hpc_financial_crisis}, in relation to forecast performance.
 
@@
 
-\caption{rRMSEs: out-of-sample forecast errors Equation \eqref{oosfe} vs expanding mean of GDP\\Out-of-sample evaluation starting in Q1-2008, without pandemic\\Values $< 1$ indicate superior predictions from Equation \eqref{oos_pred}\\
+\caption{rRMSEs: Out-of-sample forecast errors Equation \eqref{oosfe} vs expanding mean of GDP.\\Out-of-sample evaluation starting in Q1-2008, without pandemic\\Values $< 1$ indicate superior predictions from Equation \eqref{oos_pred}\\
 Row-index: Shift refers to the left-shift of the target $\textrm{GDP}_{t+shift}$\\
@@
 \centering
-\begin{tabular}{rrrrrrrr}
+\begin{tabular}{llllllll}
   \hline
@@
 To this point, indicators have been assessed by the extent to which they correlate with or lead movements in the target variable.
-An alternative approach would to be consider their ability to predict the correct ``type'' of outcome, e.g.
+An alternative approach would be to consider their ability to predict the correct ``type'' of outcome, e.g.
 whether growth will be stronger or weaker than average, or whether the economy will be in an expansion or a recession.
 This may be of particular interest to policymakers whose primary concern is to intervene ``in the right direction'' so as to stabilize the economy overall.
-We therefore limit our analysis to binary classifiers, focussing in particular on the ROC curve and the AUC.\footnote{For example, see \cite{YangLahiriPagan2024ROC} for a recent discussion of the ROC, the AUC and their relation to selected other binary classifiers.
+We therefore limit our analysis to binary classifiers, focusing in particular on the ROC (Receiver Operating Characteristic Curve) curve and the AUC (Area Under the Curve).\footnote{For example, see \cite{YangLahiriPagan2024ROC} for a recent discussion of the ROC, the AUC and their relation to selected other binary classifiers.
 The discussion below follows their notation.}
@@
 
-where $\mathbf{1}()$ is the indicator function.\footnote{The value of the indicator function is equal to 1 when the condition in its argument is satisfied; otherwise the value of the function is equal to 0.}
+where $\mathbf{1}()$ is the indicator function.\footnote{The value of the indicator function is equal to 1 when the condition in its argument is satisfied; otherwise, the value of the function is equal to 0.}
 
@@
 
-The usefulness of a given variable $Y$ as a predictor will obviously depend on the value of $\tau$ selected, which in turn may be influenced by the relative importance attached to $H(\tau)$ and $F(\tau)$.
-To compare the range of possibilities associated with various choices for $Y$, we may plot on the unit square $F(\tau)$ vs $H(\tau)$ for all values of $\tau$ in the range of $Y$.
-This is called the Receiver Operating Characteristic (ROC) curve.
+The usefulness of a given variable $Y$ as a predictor will depend on the value of $\tau$ selected, which in turn may be influenced by the relative importance attached to $H(\tau)$ and $F(\tau)$.
+To compare the range of possibilities associated with various choices for $Y$, we plot on the unit square $F(\tau)$ vs $H(\tau)$ for all values of $\tau$ in the range of $Y$.
+This is called ROC.
 
@@
 
-A popular approximate measure of the overall performance of an indicator $Y$ that abstracts from the choice of $\tau$ is the Area Under the ROC Curve (AUC).
+A popular approximate measure of the overall performance of an indicator $Y$ that abstracts from the choice of $\tau$ is the AUC.
 Because the ROC is defined on the unit square, the maximum value of the AUC = 1.
-Above we saw that when $\{Y,Z\}$ are independent, the ROC curve lies on the 45-degree line, so the area under the curve is then 0.5.
-The range of AUCs between 1 and 0.5 therefore provides some quantitative guidance as to which predictor variables $Y$ have the potential to improve the prediction of $Z$.\footnote{AUCs $< 0.5$ suggest that $-Y$ may give more useful predictions than $Y$.}.
+Above, we saw that when $\{Y,Z\}$ are independent, the ROC curve lies on the 45-degree line, so the area under the curve is then 0.5.
+The range of AUCs between 1 and 0.5 therefore provides some quantitative guidance as to which predictor variables $Y$ have the potential to improve the prediction of $Z$.\footnote{AUCs $< 0.5$ suggest that $-Y$ may give more useful predictions than $Y$.}
 
@@
 For clarity, these results are based on the entire sample, excluding the pandemic period.
-Fig.\eqref{ROC_GDP_shift_1_4} compares ROCs of the predictors for small and medium-term forecast horizons and Table \eqref{auc} presents AUC (area under the curve) for forecast horizons 1-5 quarters ahead.
+Figure \ref{ROC_GDP_shift_1_4} compares ROCs of the predictors for small and medium-term forecast horizons and Table \ref{auc} presents AUC (area under the curve) for forecast horizons 1-5 quarters ahead.
+
 \begin{figure}[h]\begin{center}\includegraphics[height=3in, width=5in]{./Figures/ROC_GDP_shift_1_4.pdf}\caption{Hit rate vs.
@@
 
-\subsection{Revisions: Regression Equation Only}
+\subsection{Revisions: Regression Equation}
 
@@
 %??? Justification HP(160): replace HP(1600) by AR(2) with similar persistency but shorter cycle lengthi, i.e. $a_1=-2"persistency*acos(\theta, a_2=persistency$.
-Figure \eqref{revisions1} compares the final and real-time predictors: as the sample size increases, the real-time predictor converges steadily to the final predictor.\footnote{The rate of convergence is partially determined by the simplicity of our design, which avoids incorporating additional M-SSA components.}
-As a confirmation, Figure \eqref{revisions2} displays the evolution of the intercept and regression weights over time: with increasing sample size, these estimates seem to converge to fixed points, indicating both the stationarity of the process and the consistency of the estimates.
+Figure \ref{revisions1} compares the final and real-time predictors: as the sample size increases, the real-time predictor converges steadily to the final predictor.\footnote{The rate of convergence is partially determined by the simplicity of our design, which avoids incorporating additional M-SSA components.}
+As a confirmation, Figure \ref{revisions2} displays the evolution of the intercept and regression weights over time: with increasing sample size, these estimates seem to converge to fixed points, indicating both the stationarity of the process and the consistency of the estimates.
 %(Note, however, the slightly elevated values of the regression weight assigned to the M-SSA component during the financial crisis and the pandemic, suggesting a stronger link between the target and the predictor).
@@
 
-\subsection{Revisions: M-SSA Only}
+\subsection{Revisions: M-SSA}
 
@@
 To isolate the effect of the regression revision, all series are standardized.
-The up-dated M-SSA (blue) is left-shifted at the zero-crossings and appears slightly faster.
-For consistency, the entire pandemic episode has been removed.
+The updated M-SSA (blue) is left-shifted at the zero-crossings and appears slightly faster.
+Full sample excluding the pandemic.
         \label{revisions3}}
@@
 
-The second source of revisions in the M-SSA component predictor arises from updates to the VAR(1) model within the M-SSA filter\footnote{Similar findings apply to more sophisticated alternatives such as the VAR(3) discussed in Section \eqref{sec:data}.}.
-We compare the original estimates, based on data up to Q4-2007, with the final estimates obtained using $h=4$ and $shift=4$ (similar findings apply to other values of forecast horizon and forward-shift).
+The second source of revisions in the M-SSA component predictor arises from updates to the VAR(1) model within the M-SSA filter.\footnote{Similar findings apply to more sophisticated alternatives such as the VAR(3) discussed in Section \ref{sec:data}.} We compare the original estimates, based on data up to Q4-2007, with the final estimates obtained using $h=4$ and $shift=4$ (similar findings apply to other values of forecast horizon and forward-shift).
 For consistency, we excluded the entire pandemic episode from the data.
-Figure \eqref{up_dated_ma_inv_multi_ip} compares the MA-inversions (impulse responses) of the GDP equation derived from the `old' and the updated VAR models: unlike the former, the latter incorporates ESI as an additional explanatory variable for GDP.
-Figure \eqref{bk_2008_all} illustrates the resulting effect on the M-SSA filter targeting HP-GDP at forecast horizon $h=4$.
+Figure \ref{up_dated_ma_inv_multi_ip} compares the MA-inversions (impulse responses) of the GDP equation derived from the `old' and the updated VAR models: unlike the former, the latter incorporates ESI as an additional explanatory variable for GDP.
+Figure \ref{bk_2008_all} illustrates the resulting effect on the M-SSA filter targeting HP-GDP at forecast horizon $h=4$.
 The updated filter (right panel) assigns more weight to the additional `leading' indicators relative to GDP.
-Finally, Figure \eqref{revisions3} compares the M-SSA GDP predictors.
-To isolate the effect of the regression revision, the figure displays standardized series.
-The update to the VAR model results in a leftward shift of the predictor (represented by the blue line), which appears marginally more rapid compared to the original predictor (depicted by the red line).
+Finally, Figure \ref{revisions3} compares the M-SSA GDP predictors.
+To isolate the effect of the regression revision, the figure displays standardized series.The update to the VAR model results in a leftward shift of the predictor (represented by the blue line), which appears marginally more rapid compared to the original predictor (depicted by the red line).
 It is important to note that the smoothness (zero-crossing rate) remains unchanged, though.
-This suggests that the out-of-sample performance of the M-SSA component predictor may be better than reported in Section \eqref{oosp}, which is based on the fixed and increasingly outdated version of M-SSA.
+This suggests that the out-of-sample performance of the M-SSA component predictor may be better than reported in Section \ref{oosp}, which is based on the fixed and increasingly outdated version of M-SSA.
 
@@
 This issue is partly due to noise leakage inherent in the one-sided HP filter.
-\\
 
@@
 For illustration, we consider a simple approach based on regressing the single M-SSA GDP output on future GDP, thereby ignoring the other filter outputs.
-This predictor is intuitively appealing because the future HP-GDP, i.e., the target of M-SSA—is the low-frequency component of the future GDP target.
+This predictor is intuitively appealing because the future HP-GDP, i.e., the target of M-SSA, is the low-frequency component of the future GDP target.
 Consequently, a strong link between M-SSA and future HP-GDP also indicates a connection with GDP, even if the statistical significance is obscured by the noise inherent in GDP.
@@
 Finally, an examination of revision errors associated with quarterly updates of the proposed M-SSA-based GDP component predictor indicates that the regression parameters, VAR model, and resulting M-SSA filters stabilize after an initial burn-in period.
-This stabilization enhances the model’s explainability and interpretability.\\
+This stabilization enhances the model’s explainability and interpretability.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%
+\appendix
 \newpage
-\section{Appendix}\label{sec:Appendix}
+%\pagestyle{fancy}
+%\fancyhf{} % Setzt alle Kopf- und Fußzeilen zurück
+%\fancyhead[L]{M-SSA}
+%\fancyfoot[C]{\thepage} % Seitenzahlen in der Fußzeile wieder hinzufügen
+
+	\section{Appendix}\label{sec:appendix}
+	\setcounter{table}{0}
+	\renewcommand{\thetable}{A\arabic{table}}
+	\setcounter{figure}{0}
+	\renewcommand{\thefigure}{A\arabic{figure}}
+	\renewcommand{\thesubsection}{\Alph{subsection}}
+		\setcounter{page}{1}
+	\renewcommand{\thepage}{A-\arabic{page}}
 
@@
 
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 12:49:51 2025
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 12:49:51 2025
 \begin{table}[ht]
@@
 \end{tabular}
-
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 12:49:51 2025
+\end{table}
+
 \begin{table}[ht]
@@
 
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 12:49:51 2025
 \begin{table}[ht]
@@
 
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 12:49:51 2025
+\end{table}
+
 \begin{table}[ht]
-\caption{rRMSEs: out-of-sample forecast errors Equation \eqref{oosfe} vs expanding mean of GDP\\Out-of-sample evaluation starting in Q1-2008, including the pandemic.\\Values $< 1$ indicate superior predictions from Equation \eqref{oos_pred}.}
+\caption{rRMSEs: Out-of-sample forecast errors Equation \eqref{oosfe} vs expanding mean of GDP.\\Out-of-sample evaluation starting in Q1-2008, including the pandemic.\\Values $< 1$ indicate superior predictions from Equation \eqref{oos_pred}.}
 \label{rRMSE_mSSA_comp_mean2}
@@
 
-\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 12:49:51 2025
+\end{table}
+
 \begin{table}[ht]
@@
 
-\begin{figure}[htpb]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/CCF.pdf}\caption{Sample cross-correlation function (CCF): full data set.\label{CCF}}\end{center}\end{figure}
-\begin{figure}[htpb]\begin{center}\includegraphics[height=3in, width=4.5in]{./Figures/CCF_wc.pdf}\caption{Sample cross-correlation function (CCF): without Pandemic.\label{CCF_wc}}\end{center}\end{figure}
-
 \end{document}

### 2025-08-15 Archive IWH Simon paper/15.08.2025 afetr Katja's feedback/paper.tex
up

--- 2025-08-15 Archive IWH Simon paper/12.08.2025/paper.tex
@@
 \usepackage{color}
+\usepackage{fancyhdr}
 %\usepackage{subfig}
@@
 
-\geometry{left=2.25cm,right=2.25cm,top=1.5cm,bottom=2cm}
-%\hyphenation{EMU} \hyphenation{CEECs}
-%\setlength{\parindent}{0pt} \setlength{\parskip}{10pt} %\setlength{\abovecaptionskip}{-5pt}
-%\setlength{\belowcaptionskip}{-5pt} \setlength{\floatsep}{10pt}
-\clubpenalty = 10000 
+\geometry{left=2cm,right=2cm,top=1.5cm,bottom=1.5cm}
+
+\setlength{\parindent}{0pt} %%no space at the beginning of paragraph
+\setlength{\parskip}{6pt} %% space between paragraphs
+
+\setlength{\abovecaptionskip}{0pt} %%Abstand unter der Überschrift
+\setlength{\belowcaptionskip}{0pt} %%Abstand ÜBER der Überschrift
+\setlength{\floatsep}{6pt}
+\setlength{\textfloatsep}{6pt}
+\setlength{\intextsep}{1\baselineskip}
+
+\clubpenalty = 10000 %keine einzelnen Seiten eines Absatzes oben oder unten auf einer Seite
 \widowpenalty = 10000
 \displaywidowpenalty = 10000
-
-% \makeatletter
-% \setlength{\@fptop}{5pt} %setzt den abstand der floats vom oberen Rand auf Null- 5 entspricht der gleichen Höhe wie die wo der Text anfängt
-% 		%\setlength\@fptop{0\p@ \@plus 1fil}
-% \setlength\@fpsep{10pt plus 2pt minus 2pt} %Abstand zwischen den Floats
-% \makeatother
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{document}
+%
+\makeatletter
+\setlength{\@fptop}{5pt} %setzt den abstand der floats vom oberen Rand auf Null- 5 entspricht der gleichen H?he wie die wo der Text anf?ngt
+%\setlength\@fptop{0\p@ \@plus 1fil}
+\setlength\@fpsep{30pt plus 2pt minus 2pt} %Abstand zwischen den Floats
+\makeatother
+
 \setlength{\abovedisplayskip}{2cm}
@@
 \setlength{\belowdisplayskip}{0cm}
-\setlength{\belowdisplayshortskip}{2cm} %Abstand zwischen Text und Formel
+\setlength{\belowdisplayshortskip}{2cm} %abstand zwischen text und formel
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{document}
 
@@
 Marc Wildi\thanks{Zurich University of Applied Sciences (ZHAW).}}
-\date{August 2025 \\
-\vspace{2cm} \textit{Work in progress, not for circulation}}
+\date{August 2025} 
+%\vspace{2cm} \textit{Work in progress, not for circulation}}
 
@@
 
-%Economic forecasters face a dilemma: While efficiency suggests that their forecasts should incorporate all available information, the resulting forecasts are often volatile, which may reflect greater information content or contamination by excessive ``noise''.
-Further, forecast users (including public and private decision makers) may be constrained by how rapidly they can adjust to changing forecasts.\footnote{For example, private or public spending decisions may be constrained by annual budgets.
-Similarly, central banks increasingly attempt to smooth interest rate changes to allow them to signal the future direction of policy to markets, while private firms attempt to smooth changes in dividends.}
-
 Economic forecasters face a dilemma: while informational efficiency requires that forecasts incorporate all available data, doing so often results in forecasts that are excessively volatile. 
- %Volatile forecasts may be a sign that they efficiently incorporate all available information. 
 In addition, forecasts may reflect excessive ``noise'' due to overfitting.
@@
 
-Recent work by \cite{Wildi2024,Wildi2025} and \cite{McElroy2019,McElroy2020} formalipropose novel methods that optimally smooth forecasts by controlling the expected frequency of sign changes.
-The resulting Smooth Sign Accuracy (SSA) framework encompasses efficiently the mean-square error (MSE) as a special case and produces increasingly smooth forecasts as changes in the forecast direction (sign) are increasingly penalized.
+Recent work by \cite{Wildi2024,Wildi2025} and \cite{McElroy2019,McElroy2020} formalizes this problem and propose a novel method that optimally smooths forecasts by controlling the expected frequency of sign changes.
+This Smooth Sign Accuracy (SSA) framework encompasses efficiently the mean-squared error (MSE) as a special case and produces increasingly smooth forecasts as changes in the forecast direction (sign) are increasingly penalized.
 
@@
 
-%This study proposes a methodological innovation that extends the classical direct forecasting framework by integrating dimension-reduced information extracted through Multivariate Smooth Sign Accuracy (M-SSA).
-The core idea is to replace raw
-%, unprocessed 
-%leading indicators with smoothed and filtered components derived from the M-SSA procedure, thereby improving the signal-to-noise ratio in the predictor variables used to forecast GDP.
-We apply this framework to well-established German economic indicators to show its ability to improve forecasts of GDP and GDP trends using a variety of sample periods and performance metrics.
-
-This study extends this framework by replacing raw leading indicators with smoothed and filtered components, thereby improving the signal-to-noise ratio in the predictor variables.
+This study extends the SSA-framework in two ways: First, replacing raw leading indicators with smoothed and filtered components, thereby improving the signal-to-noise ratio in the predictor variables.
+Second, we generalize SSA to a multivariate context (M-SSA), exploiting cross-sectional information among leading indicators to enhance forecast accuracy and stability.
 We apply this framework to well-established German economic indicators to show its ability to improve forecasts of GDP growth \citep{lehmann2021predicting,Heinisch2018bottom} and GDP trends using a variety of sample periods and performance metrics.
+Previous research often highlights the utility of direct forecasts, particularly for short-term predictions such as nowcasting and forecasts one quarter ahead.
+However, this paper advances the literature by showing that more sophisticated predictors—specifically, filtered components obtained via M-SSA yield superior performance.
 
@@
 
-%Previous research often highlights the utility of direct forecasts, particularly for short-term predictions such as nowcasting and forecasts one quarter ahead.
-However, this paper advances the literature by showing that more sophisticated predictors—specifically, filtered components obtained via M-SSA yield superior performance, especially over medium-term horizons ranging from two to four quarters ahead.
-
 We proceed as follows.
-Section \ref{sec:data} describes our data series, the sample period used, and documents their dynamic behaviour.
+Section \ref{sec:data} describes our data, the sample period used, and documents their dynamic behaviour.
 Section \ref{sec:direct_forecast} provides benchmark forecasts using simple direct projections of GDP based on selected indicators.
 We demonstrate how in-sample forecast performance may improve when we ``de-noise'' the indicators prior to projection with simple smoothing filters.
-Section \ref{sec:mSSA} defines the M-SSA method and shows how its forecast performance compares to that of direct projections as well as how it varies with forecast horizon.
-Section \ref{sec:mSSA_component} considers the ability of our indicators to forecast German GDP growth out-of-sample using the M-SSA framework, and Section \ref{sec:revisions} investigates the temporal stability of the estimated forecasting equations.
+Section \ref{sec:mSSA} defines the M-SSA method.
+Section \ref{sec:performance_mssa} reports empirical results and shows the forecast performance compared to that of direct projections as well as how it varies with forecast horizon. 
+%%%Section \ref{sec:mSSA_component} considers the ability of our indicators to forecast German GDP growth out-of-sample using the M-SSA framework.
+Section \ref{sec:revisions} examines robustness and stability.
 Section \ref{sec:conclusions} summarizes our results and concludes.
+Additional Figures and Tables are presented in an Online Appendix.\footnote{\url{www.iwh-halle.de}}
 
@@
 \subsection{Data and Publication Lags}
-The analysis employs a set of well-established leading indicators to forecast quarterly German GDP \citep{Drechsel2012financial,Heinisch2018bottom}, including industrial production (IP), the ifo Business Climate Index (ifo\_c), the Economic Sentiment Indicator (ESI), and the term spread between 10-year bond yields and the 3-month EURIBOR rate (spre\_10y\_3m) covering the period 1995 to 2024.
+Our analysis employs a set of well-established leading indicators to forecast quarterly German GDP \citep{Drechsel2012financial,Heinisch2018bottom}, including industrial production (IP), the ifo Business Climate Index (ifo\_c), the Economic Sentiment Indicator (ESI), and the term spread between 10-year bond yields and the 3-month EURIBOR rate (spre\_10y\_3m) covering the period 1991 to 2024.
 We take into account the publication lag of the indicators and the GDP series (Table \ref{tab:data_stucture}).\footnote{Data revisions are not considered, given that previous studies \citep{Heinisch2019} have shown that their overall effect on forecast performance is minor.}
-While GDP is available with a one-quarter delay, and IP is released with a two-month lag, business surveys and financial data are published without any lag.
-%All series are consequently re-aligned to correspond with the information available to forecasters, with GDP and IP variables shifted forward according to their respective lags.
-Table \eqref{tab:data_stucture_aligned} displays the dataset as of the end of 2024, with GDP and IP variables temporally aligned at the sample endpoint.
-This configuration reflects the matrix of explanatory variables accessible in Q4-2024 for the purposes of nowcasting or forecasting GDP.
-For illustrative purposes, an additional target column is included on the right side of the table, representing the dependent variable for nowcasting applications.
-In the case of an $h$-step ahead forecast, this column is shifted upwards by $h$ quarters.
+While GDP is available with a one-quarter delay, and IP is released with a two-month lag, business surveys and financial data are published contemporaneously.
 
 \begin{table}[ht]
-\caption{Ragged edge structure. As of January 2025, before publication of the GDP flash estimate.} 
+    \caption{Ragged edge structure.\\ Information set available to forecasters as of January 2025, before publication of the GDP flash estimate.} 
     \label{tab:data_stucture}
@@
     \end{tabular}
-
 \end{table}
@@
 \begin{table}[ht]
-\caption{Aligned dataset.
-At each date, various columns show the most recent values of each predictor available to forecasters.
+    \caption{Aligned dataset.\\
+    Indicators realigned to reflect real-time availability, with the GDP column shifted by its publication lag.
 Column 6 represents the target variable for a GDP nowcast (i.e.
@@
 Additionally, extreme outliers associated with the COVID-19 pandemic are trimmed to prevent distortions in subsequent figures.
- Monthly series are shifted (Table \ref{tab:data_stucture_aligned}) and averaged to quarterly series (Figure \ref{fig:data}), while Figure \ref{fig:data_lags} offers a close-up view of the financial crisis and the pandemic.
+ Monthly series are aligned (Table \ref{tab:data_stucture_aligned}) and averaged to quarterly series (Figure \ref{fig:data}), while Figure \ref{fig:data_lags} offers a close-up view of the financial crisis and the pandemic.
 The latter figure shows that the real-time GDP and industrial production series are synchronized at turning points, whereas the other indicators tend to lead somewhat.
@@
     \end{subfigure}
-\caption{GDP \& Indicators: standardized log-differences, trimmed to $\pm 3$ standard deviations.}
+    \caption{Quarterly German GDP and Indicators.\\ standardized log-differences, trimmed to $\pm 3$ standard deviations to facilitate visual inspection.}
 \end{figure}
@@
 The latter is a key input to the multivariate filters we develop in Sections \ref{sec:mSSA} and \ref{sec:mSSA_component}.
-
-Figures \ref{CCF} (full sample) and \ref{CCF_wc} (omitting the Pandemic period) show the sample cross-correlation functions (CCF's) between current and lagged GDP and the two survey indicators.
-Both figures show strong correlations that peak when GDP is slightly lagged.
-Results for IP showed no such lag, while correlations for the term spread were weaker.
-The figures also show that the precise strengths of the correlations are influenced by the outliers associated with the COVID-19 pandemic.
+The sample cross-correlation functions (CCF's) between current and lagged GDP and the two survey indicators show strong correlations that peak when GDP is slightly lagged.\footnote{See Figures \ref{fig:CCF} (full sample) and \ref{fig:CCF_wc} (omitting the Pandemic period) in the Online Appendix.} Results for IP showed no such lag, while correlations for the term spread were weaker.
+However, the precise strengths of the correlations are influenced by the outliers associated with the COVID-19 pandemic.
 In the interest of model stability, we exclude data from Q4-2019 to Q4-2020 (pandemic) when modelling the dynamics for the multivariate filters developed below.
 
-\begin{figure}[ht]
-    \begin{subfigure}[htpb]{\textwidth}
-        \begin{center}
-            \includegraphics[height=3in, width=4.5in]{./Figures/CCF.pdf}
-            \caption{Full data set from Q2-1991 to Q4-2024.\label{CCF}}
-        \end{center}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[htpb]{\textwidth}
-        \begin{center}
-            \includegraphics[height=3in, width=4.5in]{./Figures/CCF_wc.pdf}
-            \caption{Without Pandemic (omitting Q4-2019 to Q4-2020).\label{CCF_wc}}
-        \end{center}
-    \end{subfigure}
-    \caption{Sample cross-correlation functions (CCF) showing correlations with lags of GDP growth.}
-\end{figure}
-
-We then estimated a variety of parsimonious VARMA models to capture these dynamics, selecting parsimonious VAR(1) and VAR(3) specifications.\footnote{We used the \texttt{refVARMA()} procedure in the \cite{MTSpackage} \texttt{MTS} package, which sets VARMA coefficients with t-statistics below a critical threshold to zero \citep[see][]{tsay2013multivariate}.
+We then estimate a variety of parsimonious VARMA models to capture these dynamics, selecting parsimonious VAR(1) and VAR(3) specifications.\footnote{We used the \texttt{refVARMA()} procedure in the \cite{MTSpackage} \texttt{MTS} package, which sets VARMA coefficients with t-statistics below a critical threshold to zero \citep[see][]{tsay2013multivariate}.
 This led to a VAR(1) specification with a threshold of 1.5.
@@
 This led to a VAR(3) specification with 7 steps.
-For additional details, see \cite{ElasticNet} and the references therein.} The Figures \ref{ma_inv_BIP} and \ref{ma_inv_BIP_var3_en} show the implied responses of GDP to shocks to each indicator variable.\footnote{Shown are the MA coefficients for GDP from the vector moving-average (VMA) models obtained by inversion of the VAR(1) and VAR(3) representations respectively (i.e.
-the impulse-response functions (IRFs) of GDP to the correlated reduced-form shocks).}
-
-%
+For additional details, see \cite{ElasticNet} and the references therein.} The implied responses of GDP to shocks to each indicator variable are indicating that survey indicators have the strongest and most persistent impact on GDP, with pandemic data shortening the apparent lead time.\footnote{See Figures \ref{fig:ma_inv_BIP} and \ref{fig:ma_inv_BIP_var3_en} in the Online Appendix.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %   SvN: I've added the following very brief description of Elastic-Net
@@
 %
-%
-
-\begin{figure}[htpb]
-    \begin{subfigure}[htpb]{\textwidth}
-        \begin{center}
-            \includegraphics[height=3in, width=6in]{./Figures/ma_inv_BIP.pdf}
-            \caption{implied by VAR(1) \label{ma_inv_BIP}}
-        \end{center}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[htpb]{\textwidth}
-        \begin{center}
-            \includegraphics[height=3in, width=6in]{./Figures/ma_inv_BIP_var3_en.pdf}
-            \caption{implied by VAR(3)\label{ma_inv_BIP_var3_en}}
-        \end{center}
-    \end{subfigure}
-    \caption{VMA coefficients for GDP.\\Estimation with pandemic (left) and without pandemic (right).}
-\end{figure}
-
-%The models encapsulate key characteristics of the data after adjusting for publication lags. Notably, 
+
 GDP exhibits the strongest response to shocks in the two survey indicators—ifo and ESI.\footnote{There is also weaker evidence of a response to the term spread at longer lags.} The impact of the pandemic modifies the lag structure by emphasizing shorter dependencies triggered by the alternating outliers during that specific episode.
@@
 
-\section{Forecasts with Pre-Filtering}\label{sec:direct_forecast}
-
-In this section, we demonstrate how pre-filtering indicator variables may improve the performance of simple forecasts by removing noise.
-We demonstrate how in-sample forecast performance improves when we ``de-noise'' the indicators prior to projection with simple smoothing filters.\footnote{The smoothing filters are one-sided (concurrent) filters of length 31, therefore, estimation of both forecasting models omits the first 30 observations.}
-
-For simplicity, all the models in this section are direct projections of GDP on our indicators and take the form
+\section{Baseline Forecasting Approaches}
+\label{sec:direct_forecast}
+
+In this section, we evaluate the performance of simple indicator-based forecasts of German GDP growth and examine how pre-filtering can improve their accuracy by removing noise. 
+Our starting point is a direct in-sample forecasting framework in which GDP growth at horizon $h$ is regressed on a set of leading indicators and lagged GDP, estimated over the pre-pandemic period.
 
 \begin{equation}
-    y_{t+h} = \boldsymbol{\beta'} \cdot \mathbf{X}_t + \varepsilon_t
+    u_{t+h} = \boldsymbol{\beta'} \cdot \mathbf{X}_t + \varepsilon_t
     \label{eq:direct_forecast}
@@
 \begin{itemize}
-    \item $y_{t+h}$ is the quarterly growth rate of GDP for forecast horizon $h = \{0, 1, 2, ... \}$ (labeled shifted GDP in the following )
-    \item $\mathbf{X}_t \equiv [1,y_{t-1}, ip_t, ifo\_c_t, ESI_t, spr\_10y\_3m_t]'$ or their transformed counterparts and $y_{t-1}$ accounts for the publication lag of GDP.\footnote{Including lagged GDP enables the nesting of the AR forecast within the broader modeling framework.}
+    \item $u_{t+h}$ is the quarterly growth rate of GDP for forecast horizon $h = \{0, 1, 2, ... \}$ (labeled shifted GDP in the following )
+    \item $\mathbf{X}_t \equiv [1,u_{t-1}, ip_t, ifo\_c_t, ESI_t, spr\_10y\_3m_t]'$ or their transformed counterparts and $u_{t-1}$ accounts for the publication lag of GDP.\footnote{Including lagged GDP enables the nesting of the AR forecast within the broader modeling framework.}
     \item $\hat{\boldsymbol{\beta}}$ is estimated by least-squares, omitting observations affected by the pandemic, i.e. Q4-2019 to Q4-2020.
@@
 
-\subsection{Unfiltered Indicators}\label{cdf}
-
-%So-called direct forecasts are obtained by regressing the relevant indicators on forward-shifted GDP, accounting for the additional publication lag.
-To illustrate the concept, we rely on all indicators and full sample information, excluding observations affected by the pandemic.
-In-sample forecasts and the forward-shifted GDP, which serves as the target, are presented in Figures \eqref{direct_wc} (full dataset without the pandemic) and \eqref{direct_wc_financial_crisis} (financial crisis) for forecast horizons up to three quarters ahead.
-
-\begin{figure}[htbp]
-
- \begin{subfigure}[htpb]{\textwidth}
-        \begin{center}
-             \includegraphics[width=0.9\textwidth]{./Figures/direct_wc_all.pdf}
-            \caption{Full sample information (without pandemic)\label{direct_wc}}
-        \end{center}
-    \end{subfigure}
-
- \begin{subfigure}[htpb]{\textwidth}
-        \begin{center}
-            \includegraphics[width=0.9\textwidth]{./Figures/direct_wc_financial_crisis.pdf}
-            \caption{Financial crisis (without pandemic)\label{direct_wc_financial_crisis}}
-        \end{center}
-    \end{subfigure}
-    \caption{Direct forecasts (red lines) based on all indicators and target (black line): GDP left-shifted  by 0 - 3 quarters.}
-\end{figure}
-
-Figures \ref{direct_wc} and \ref{direct_wc_financial_crisis} compare actual and fitted values for GDP at forecast horizons $h = 0, \ldots, 3$, using the four indicators and the target, i.e., GDP shifted $h$ quarters ahead.
-Figure \ref{direct_wc} shows results over the sample without the pandemic, while Figure \ref{direct_wc_financial_crisis} provides a closer look at the fit during the financial crisis.
-Table \ref{tab:pvaluedhp} shows p-values for tests of the joint null hypothesis that all coefficients on the indicators are zero (HAC-adjusted Wald test).
-
-The results indicate that as the forecast horizon $h$ increases, forecasts become less volatile and their turning points (i.e., peaks and troughs) become increasingly right-shifted (lagged) relative to the target GDP.
-Additionally, the p-values of the HAC-adjusted Wald test precipitously increase from $h=2$ to $h=3$ (see Table \ref{tab:pvaluedhp}), implying that the indicators become statistically insignificant for forecast horizons of more than two quarters.\footnote{
-These findings are confirmed by the relative root mean-square errors comparing direct forecasts against the mean benchmark, see Table \ref{rRMSE_mSSA_direct_mean_without_covid8} in the Appendix.}
-
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 12:49:46 2025
-
-%\begin{table}[ht]
-%\centering
-%\begin{tabular}{rrrrr}
-%  \hline
-% & h=3 & h=4 & h=5 & h=6 \\ 
-%  \hline
-%rRMSE & 0.962 & 0.967 & 0.990 & 0.967 \\ 
-%   \hline
-%\end{tabular}
-%\caption{Relative root mean-square error of HP-C against classic direct forecast for $h\geq %3$.\\Values $<1$ imply  HP-C has smaller RMSE.  } 
-%\label{tab:rRMSE}
-%\end{table}
+We first estimate forecasts using unfiltered indicators, which reflect the raw information available to forecasters in real time.
+We then apply a one-sided concurrent smoothing filter to the same indicators and demonstrate how in-sample forecast performance improves when we ``de-noise'' the indicators before projection with simple smoothing filters.\footnote{The smoothing filters are one-sided (concurrent) filters of length 31, therefore, estimation of both forecasting models omits the first 30 observations.}
+Comparing these two designs allows us to isolate the effect of pre-filtering in a straightforward, transparent way.
+
+\subsection{Direct Forecasts with Unfiltered Indicators}\label{sec:cdf}
+
+As a benchmark, we estimate direct projections of GDP growth at horizons $h=0$ to $h=6$ quarters ahead using all indicators without filtering.   
+Figures \ref{fig:direct_wc} and \ref{fig:direct_wc_financial_crisis} compare actual and fitted values for GDP at forecast horizons $h = 0, \ldots, 3$, using the four indicators plus lagged GDP. 
+%At short horizons, forecasts track turning points relatively well but exhibit considerable volatility.
+With increasing horizon, the forecasts become smoother; however, this is achieved at the expense of timeliness---turning points, i.e., peaks and troughs, are increasingly delayed relative to the target series GDP.
+At short horizons, forecasts track turning points relatively well.
+As the forecast horizon increases, the forecasts become increasingly damped (zero-shrinkage) and delayed, with peaks and troughs progressively shifted to the right relative to the target series GDP.
+
 \bigskip
 \begin{table}[h]
-\caption{p-values for $H_0: \boldsymbol{\beta} = 0$ in eq.~\eqref{eq:direct_forecast} based on HAC-adjusted Wald test.\\Estimation over full sample (without pandemic.)} 
+\caption{Statistical Significance of Direct Forecasts.\\
+p-values for $H_0: \boldsymbol{\beta} = 0$ in eq.~\eqref{eq:direct_forecast} based on HAC-adjusted Wald test.\\Estimation over full sample (without pandemic.)} 
 \label{tab:pvaluedhp}
@@
 
-%\begin{table}[ht]
-%\centering
-%\begin{tabular}{rrrrrrrr}
-%  \hline
-% & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
-%  \hline
-%P-values & 0.000 & 0.000 & 0.007 & 0.946 & 0.603 & 0.096 & 0.181 \\ 
-%   \hline
-%\end{tabular}
-%\caption{Statistical significance of direct forecasts.  } 
-%\label{tab:f_stats}
-%\end{table}
-
-\subsection{Direct Forecasts Based on Filtered Indicators}\label{hpdf}
-%\subsection{Filter: HP(160)}
-
-%As shown, the performance of direct forecasts sharply declines for longer forecast horizons. 
+Table \ref{tab:pvaluedhp} reports p-values for tests of the joint null hypothesis that all coefficients on the indicators are zero (HAC-adjusted Wald test). 
+Additionally, these results confirm that the predictive power of the unfiltered indicators declines sharply beyond $h=2$, with significance levels falling to insignificance for horizons beyond two quarters.%\footnote{These findings are confirmed by the relative root mean-square errors comparing direct forecasts against the mean benchmark, see Table \ref{tab:rRMSE_mSSA_direct_mean_without_covid8} in the Appendix.}
+ This finding motivates the use of pre-filtering to extract a clearer signal from the indicator set.
+
+\begin{figure}[htbp]
+
+ \begin{subfigure}[htpb]{\textwidth}
+        \begin{center}
+             \includegraphics[width=0.9\textwidth]{./Figures/direct_wc_all.pdf}
+            \caption{Full sample information (without pandemic)\label{fig:direct_wc}}
+        \end{center}
+    \end{subfigure}
+
+ \begin{subfigure}[htpb]{\textwidth}
+        \begin{center}
+            \includegraphics[width=0.9\textwidth]{./Figures/direct_wc_financial_crisis.pdf}
+            \caption{Financial crisis (without pandemic)\label{fig:direct_wc_financial_crisis}}
+        \end{center}
+    \end{subfigure}
+    \caption{Direct forecasts of GDP (red lines) using unfiltered indicators and target (black line): GDP left-shifted  by 0 - 3 quarters.}\label{fig:direct_forecast}
+\end{figure}
+
+\subsection{Direct Forecasts with HP-C-Filtered Indicators}\label{sec:hpdf}
+
 We conjecture that economic indicators may be contaminated by unpredictable high-frequency noise, thereby obscuring the effective `signal' and making a direct regression more susceptible to overfitting.
@@
 The value $\lambda=1600$ is the most common value used for business cycle analysis, but \cite{Phillips_Jin_2021} note that this removes relevant information due to excessive smoothing.
-%In our context, this oversmoothing issue would be further exacerbated when considering forecast horizons shorter than a year---an interval inconsistent with the mean duration of up to several years of business cycles, as highlighted by the HP(1600).
+%% In our context, this oversmoothing issue would be further exacerbated when considering forecast horizons shorter than a year---an interval inconsistent with the mean duration of up to several years of business cycles, as highlighted by the HP(1600).
 To address this concern, we instead use $\lambda = 160$ to allow for a more adaptive filter.
-
-Figure \ref{hp_160} presents both classic two-sided and one-sided HP(160) (denoted HP-C for \textit{concurrent}) filters (top panels) along with their associated amplitude functions in the frequency domain (bottom panels).
-It confirms that the HP-C filter assigns greater weight to yearly components—relevant in a medium-term forecast context— than the classic quarterly HP filter, while also reducing undesirable high-frequency noise.
-Therefore, while the choice of the HP(160) filter in this example
-is somewhat arbitrary, it should serve the purpose of reducing high-frequency noise while preserving longer-duration movements.\footnote{A comprehensive technical analysis of the effect of $\lambda$ on the resulting GDP predictor can be conducted using the M-SSA package (\url{https://github.com/wiaidp/R-package-SSA-Predictor}).}
-
-\begin{figure}[h]
-    \begin{center}
-%    \textcolor{blue}{\textbf{Notice that the title on top right panel is incorrect!}}
-        \includegraphics[height=4in, width=5.5in]{./Figures/hp_160.pdf}
-        % \caption{Two-sided HP(160) (top left) and one-sided (concurrent) HP-C filters (top right) with corresponding amplitude functions (bottom).
-        \caption{HP$(\lambda=160)$ filters.\\
-        Left: HP (Two-sided, symmetric) \qquad Right: HP-C (One-sided, concurrent)\\
-        Top Row: Filter Weights \qquad \qquad \quad Bottom Row: Amplitude Functions.
-        \label{hp_160}}
-    \end{center}
-\end{figure}
-
-%\subsection{Direct HP Forecast}\label{hpdf}
-
-%Extending the `classic' direct forecast design, we consider filtered indicators, based on the one-sided HP(160), denoted as the HP-C filter: the filtered indicators are explanatory variables for the regressions on forward-shifted GDP.
-As in the previous section, we rely on all indicators as well as the full sample estimate for illustration,
-
-Figures \ref{direct_hp_forecasts} and \ref{direct_hp_forecasts_financial_crisis} provide results comparable to those of Figures \ref{direct_wc} and \ref{direct_wc_financial_crisis} but now using indicators smoothed with the one-sided HP(160), (denoted as the HP-C, for \textit{concurrent} filter).
+Comparing classic two-sided and one-sided HP(160) (denoted HP-C for \textit{concurrent}) filters, we find that the HP-C filter assigns greater weight to yearly components---relevant in a medium-term forecast context---than the classic quarterly HP filter, while also reducing undesirable high-frequency noise.\footnote{See Figure \ref{fig:hp_160} in the Online Appendix.
+A comprehensive technical analysis of the effect of $\lambda$ on the resulting GDP predictor can be conducted using the M-SSA package (\url{https://github.com/wiaidp/R-package-SSA-Predictor}).}
+
+Figure \ref{fig:direct_hp_forecasts_financial_crisis} provides results comparable to those of \ref{fig:direct_forecast} but now using indicators smoothed with HP-C.
 The fitted values lag GDP slightly less and track turning points somewhat more quickly.
-Table \ref{tab:pvaluedhp} shows p-values for tests of the joint null hypothesis that all coefficients ${\boldsymbol{\beta}} = 0$.
-The use of filtered indicators improves their statistical significance at forecast horizons $h > 2$ quarters.
-%Figures \ref{direct_hp_forecasts} (all observations with exclusion of the pandemic) and \ref{direct_hp_forecasts_financial_crisis} (financial crisis).
-A comparison with the classic (unfiltered) direct forecasts in the latter figure %in Fig.\eqref{direct_wc_financial_crisis}
-% indicates that the new predictors are slightly less retarded and track the turning points more promptly (`faster'). 
-% The filtered approach also exhibits superior statistical properties. As shown in Table \ref{tab:f_stats}, 
-%The p-values of the F-statistics for the HP-based forecasts are smaller than for the direct forecasts specifically at $h=3$, but the predictor remains statistically insignificant at forecast horizons exceeding two quarters (even though the evaluation is conducted in-sample and based on all available indicators).
+Table \ref{tab:pvaluedhp},reporting the p-values for tests of the joint null hypothesis that all coefficients ${\boldsymbol{\beta}} = 0$, also shows that filtering improves the statistical significance of the indicators for horizons beyond two quarters, suggesting that much of the predictive content of the raw series is concentrated in their lower-frequency components.
 
@@
     \begin{center}
-        \includegraphics[width=0.85\textwidth]{./Figures/direct_hp_forecasts.pdf}
-        \caption{Direct HP-C forecasts: entire data set (without pandemic).
-        \label{direct_hp_forecasts}}
-    \end{center}
-\end{figure}
-
-\begin{figure}[H]
-    \begin{center}
         \includegraphics[height=5in, width=6in]{./Figures/direct_hp_forecasts_financial_crisis.pdf}
-        \caption{Shifted GDP (black), direct HP forecast (red) and classic direct forecast (blue) over the course of the financial crisis: regression based on all indicators and full sample information (without pandemic).
-        \label{direct_hp_forecasts_financial_crisis}}
-    \end{center}
-\end{figure}
-
-%We now propose a more refined causal filter design aiming for statistical significance of the resulting predictor at longer forecast horizons, out-of-sample.
-
-The application of the one-sided HP-C filter to the indicators yields some measurable improvements in forecast performance. However, 
-%its univariate nature excludes additional information such as that provided by leading indicators. 
-the noise suppression capability of the one-sided filter is compromised, as explained by \cite{Wildi2025} and illustrated by the high-frequency leakage of the amplitude function in Figure \ref{hp_160}.
-To address these limitations, we turn to the Multivariate Smooth Sign Accuracy (M-SSA) framework proposed by \cite{Wildi2025}.
-
-\section{Multivariate Causal Filter: M-SSA}\label{sec:mSSA}
-
-To understand the smoothing properties of the M-SSA framework, we begin with the Smooth Sign Accuracy (SSA) case, which considers a univariate smoothing problem, before turning to M-SSA, its multivariate generalization.\footnote{See \cite{Wildi2024} for further details on the SSA framework and its applications.} In the remainder of the section, we demonstrate the properties of M-SSA nowcasts and forecasts for smoothed GDP.
-
-\subsection{SSA Optimization Criterion}
+        \caption{Forecasts During the Financial Crisis: HP-C vs. Unfiltered Indicators.\\
+        Shifted GDP (black), forecasts based on HP-C filtered indicators (blue) and unfiltered indicators (red) during the financial crisis: regression based on full sample information (without pandemic).
+        \label{fig:direct_hp_forecasts_financial_crisis}}
+    \end{center}
+\end{figure}
+
+The application of the one-sided HP-C filter to the indicators yields some measurable improvements in forecast performance.
+However, the noise suppression capability of the one-sided filter is compromised, as explained by \cite{Wildi2025} and illustrated by the high-frequency leakage of the amplitude function.\footnote{Figure \ref{fig:hp_160} in the Online Appendix.} To address these limitations, we turn to the Multivariate Smooth Sign Accuracy (M-SSA) framework proposed by \cite{Wildi2025}.
+
+\section{Methodology: The M-SSA Framework
+%Multivariate Causal Filter: M-SSA}
+\label{sec:mSSA}}
+
+While pre-filtering improves the performance of direct forecasts by removing high-frequency noise, it does not directly control the smoothness of the resulting forecast series.
+Excessive sign changes in the forecast growth rate can occur near turning points, leading to spurious recession or recovery signals.
+
+The Smooth Sign Accuracy (SSA) framework proposed by \cite{Wildi2024,Wildi2025} addresses this by explicitly controlling the forecast’s holding time (HT), i.e..
+the expected duration between sign changes, while preserving a high correlation with the target series.\footnote{See \cite{Wildi2024} for further details on the SSA framework and its applications.}
+
+The Multivariate Smooth Sign Accuracy (M-SSA) framework extends SSA to multiple predictors.
+
+\subsection{
+ Smooth Sign Accuracy (SSA)}
 
@@
 In our current framework, $z_t$ corresponds to the output of a two-sided Hodrick–Prescott filter with smoothing parameter $\lambda = 160$ applied to the doubly infinite sequence $x_k$, $k\in\mathbb{Z}$.
-The finite sample $\mathbf{x}_t$ is the portion effectively observed and represents the log-differenced GDP (the dependence on $x_0, x_{-1},...$ is neglected because the HP filter weights decay rapidly to zero; see Figure \ref{hp_160}).
+The finite sample $\mathbf{x}_t$ is the portion effectively observed and represents the log-differenced GDP (the dependence on $x_0, x_{-1},...$ is neglected because the HP filter weights decay rapidly to zero; see Figure \ref{fig:hp_160}).
 The prediction task comprises tracking $z_{t}$ using the (causal) predictor $y_{t}=\sum_{k=0}^{L-1} b_{k}x_{t-k}=\mathbf{b}'\mathbf{x}_t$, where the filter $\mathbf{b}$ with weights $b_0,...,b_{L-1}'$ (actually of length $L=t$) is available for optimization purposes.
@@
 In the univariate case, \cite{Wildi2025} proposes the Smooth Sign Accuracy (SSA) as an optimization criterion for determining $\mathbf{b}$:
-\begin{eqnarray}\label{critssa}
+\begin{eqnarray}\label{eq:critssa}
 \left.\begin{array}{c}\max_{\mathbf{b}}~\rho(z,y)\\
@@
 The parameter $\rho_1$ controls the smoothness of the predictor: higher values of $\rho_1$ favour smoother trajectories of $y_t$.
-Assuming that the process $x_t$ has mean zero, \cite{Wildi2024} establishes a connection between $\rho(y,1)$ and the expected duration between consecutive `zero-crossings' (sign changes) of $y_t$, the `holding time', denoted by $HT(y)$\footnote{Of course, when $z_t$ is GDP growth, sign changes are indicative of expansions and recessions.
+Assuming that the process $x_t$ has mean zero, \cite{Wildi2024} establishes a connection between $\rho(y,1)$ and the expected duration between consecutive sign changes (or zero-crossings) of $y_t$, the `holding time', denoted by $HT(y)$\footnote{Of course, when $z_t$ is GDP growth, sign changes are indicative of expansions and recessions.
 Formally, the precise relationship between the expected duration between sign changes and the lag-one autocorrelation of the predictor is derived under the assumption of Gaussian time series.
 However, \cite{Wildi2024} demonstrates that this relationship remains robust even when deviations from Gaussianity occur (central limit theorem applied to lowpass-filtered series).}:
-\begin{eqnarray}\label{ht}
+\begin{eqnarray}\label{eq:ht}
 HT(y)=\frac{\pi}{\arccos(\rho(y,1))}.
@@
 Since eq. \eqref{eq:ht_arccos} is strictly monotonic, the SSA criterion can be reformulated as the optimization problem:
-\begin{eqnarray}\label{critssaht}
+\begin{eqnarray}\label{eq:critssaht}
 \left.\begin{array}{c}\textrm{max}_{\mathbf{b}}~\rho(z,y)\\
@@
 
-One implication of criterion \eqref{critssaht} is that the objective function as well as the constraint are indifferent to an affine transformation of the predictor.
+One implication of criterion \eqref{eq:critssaht} is that the objective function as well as the constraint are indifferent to an affine transformation of the predictor.
 This ambiguity can be resolved by assuming an arbitrary scale and level for $y_t$ (standardization).
@@
 
-Traditional predictors often yield an excessive number of false sign changes due to leakage of high-frequency components (high-frequency leakage) in the predicted series, as discussed above (see Figure \ref{hp_160}).
+Traditional predictors often yield an excessive number of false sign changes due to leakage of high-frequency components (high-frequency leakage) in the predicted series, as discussed above (see Figure \ref{fig:hp_160}).
 This noisiness is particularly problematic near the onset or termination of economic crises (recessions), where the target fluctuates around zero; reliable tracking of sign changes would facilitate real-time assessment of the occurrence and timing of such events.
-%as illustrated by the amplitude function in Figure \ref{hp_160} (bottom right panel) and further elaborated upon by \cite{Wildi2025}. 
- \cite{Wildi2025} shows that the SSA criterion helps control this phenomenon (Figure \ref{mssa_msse_zc}).
-
- Using the dual formulation of the criterion \eqref{critssaht}, the author demonstrates that the SSA solution yields the \emph{lowest} zero-crossing rate among all (linear) predictors with equivalent target correlation.
+\cite{Wildi2025} shows that the SSA criterion helps control this phenomenon (Figure \ref{fig:mssa_msse_zc}).
+
+Using the dual formulation of the criterion \eqref{eq:critssaht}, the author demonstrates that the SSA solution yields the \emph{lowest} sign change rate among all (linear) predictors with equivalent target correlation.
 The concept can be extended to a multivariate framework, denoted M-SSA, by generalizing the correlation functions as detailed in the next section.
-As we shall see, controlling the zero-crossing rate within the M-SSA framework will have a quantifiable effect on the hit rate and false positive rate, thereby explicitly addressing the detection of present (nowcasting) and future (forecasting) economic contractions and expansions.
-
-We now briefly elucidate the SSA optimization criterion outlined in Eq.\eqref{critssa}.
+As we shall see, controlling the sign change rate within the M-SSA framework will have a quantifiable effect on the hit rate and false positive rate, thereby explicitly addressing the detection of present (nowcasting) and future (forecasting) economic contractions and expansions.
+
+We now briefly elucidate the SSA optimization criterion outlined in Eq.\eqref{eq:critssa}.
 Consider a simplified univariate framework where $x_t=\epsilon_t$ represents white noise with variance $\sigma^2$ and assume, for notational simplicity, that $\sigma^2=1$.
@@
 where it is assumed that $\gamma_k\neq 0$ for some $k<0$ (implying an acausal filter), and $\sum_{|k|<\infty}\gamma_k^2<\infty$ (squared summability), ensuring the stationarity of the target $z_t$.
-With these definitions, all processes are centered at zero, justifying the use of Equation \eqref{critssaht} which links first-order ACF and HT.
+With these definitions, all processes are centered at zero, justifying the use of Equation \eqref{eq:critssaht} which links first-order ACF and HT.
 Extensions to non-stationary integrated processes are discussed in Wildi (2025).
@@
 The matrix $\mathbf{M}$ is such that $\mathbf{b}'\mathbf{Mb}=\sum_{k=1}^{L-1}b_{k-1}b_k$, representing the first-order autocovariance of $y_t$.
-Using this, the optimization criterion \eqref{critssa} can be reformulated as:
-\begin{eqnarray}\label{crit1}
+Using this, the optimization criterion \eqref{eq:critssa} can be reformulated as:
+\begin{eqnarray}\label{eq:crit1}
 \left.\begin{array}{cc}
@@
 However,
-Criterion \eqref{crit1} is readily amenable to optimization, as it involves affine and quadratic functions of $\mathbf{b}$ \citep{Wildi2024}.\footnote{Detailed derivations are obtained in \cite{Wildi2025}.} \\
+Criterion \eqref{eq:crit1} is readily amenable to optimization, as it involves affine and quadratic functions of $\mathbf{b}$ \citep{Wildi2024}.\footnote{Detailed derivations are obtained in \cite{Wildi2025}.} \\
 Under the above assumptions, the classical MSE predictor of length $L$ can be expressed via orthogonal projection as:
@@
 which corresponds to setting $b_{MSE,k}=\gamma_{k+h}$, $k=0,...,L-1$.
-In general, to enhance the smoothness of the SSA predictor $y_t$, we select $\rho_1>\rho_{MSE}$ in criterion \eqref{crit1},  where $\rho_{MSE}$ denotes the first-order ACF of the (benchmark) MSE predictor.\footnote{Selecting $\rho_1<\rho_{MSE}$ would produce an SSA predictor with more frequent zero-crossings than the MSE benchmark, which is typically undesirable in practical applications.} Finally,  an extension to autocorrelated $x_t\neq \epsilon_t$ is discussed below.
-
-\subsection{M-SSA Optimization}
-
-The SSA criterion \eqref{crit1} is now extended to a multivariate setting.
+In general, to enhance the smoothness of the SSA predictor $y_t$, we select $\rho_1>\rho_{MSE}$ in criterion \eqref{eq:crit1}, where $\rho_{MSE}$ denotes the first-order ACF of the (benchmark) MSE predictor.\footnote{Selecting $\rho_1<\rho_{MSE}$ would produce an SSA predictor with more frequent sign changes than the MSE benchmark, which is typically undesirable in practical applications.} Finally, an extension to autocorrelated $x_t\neq \epsilon_t$ is discussed below.
+
+\subsection{Multivariate Extension of SSA (M-SSA)}
+
+The SSA criterion \eqref{eq:crit1} is now extended to a multivariate setting.
 Let $\mathbf{X}_t$ (of dimension $t\times n$) denote a set of $n$ explanatory series $\mathbf{x}_{1t},...,\mathbf{x}_{nt}$, with observations $\mathbf{x}_{it}=(x_{it},...,x_{i1})'$.
@@
 The targets $z_{it}$, i=1,...,5, are obtained by applying a two-sided HP(160) filter to the doubly-infinite extensions $x_{ik}$, $k\in \mathbf{Z}$, of these indicators.
-In a subsequent step, as delineated in Section \eqref{sec:mSSA_component}, the (filtered) predictors $y_{hit}$ of $z_{it+h}$, with $h\geq 0$,  will be utilized to construct predictors for the GDP.
-However, unlike the approach in Section \eqref{sec:direct_forecast}, these predictors will be based on multivariate M-SSA filters rather than univariate HP-C designs.
+In a subsequent step, as delineated in Section \ref{sec:mSSA_component}, the (filtered) predictors $y_{hit}$ of $z_{it+h}$, with $h\geq 0$, will be utilized to construct predictors for the GDP.
+However, unlike the approach in Section \ref{sec:direct_forecast}, these predictors will be based on multivariate M-SSA filters rather than univariate HP-C designs.
 %Since $z_{it}$ depends on future observations $x_{it+1},x_{it+2},...$, the prediction task involves `tracking' $z_{it}$ using the estimate $y_{it}=\sum_{j=1}^ny_{ijt}$, where $y_{ijt}=\mathbf{b}_{ij}~'\mathbf{x}_{jt}=\sum_{k=0}^{L} b_{ijk}x_{j,t-k}$ are the outputs of a multivariate (causal or one-sided) filter $\mathbf{B}_i^f$, with columns $\mathbf{b}_{ij}$ of fixed length $L$, $j=1,...,5$, assigning weights to the last $L$ observations of the $j$-th indicator $\mathbf{x}_{jt}$ (the fixed-length assumption is merely used for ease of exposition).
@@
 Note that we here assume the MSE predictor to rely on sub-filters of length $L$ for each subseries $\mathbf{x}_{it}$. 
-Recall from the SSA criterion \eqref{crit1}  that it is not necessary to rely on the doubly infinite representations of the target when formulating the optimization criterion, a point that is subsequently confirmed.
+Recall from the SSA criterion \eqref{eq:crit1} that it is not necessary to rely on the doubly infinite representations of the target when formulating the optimization criterion, a point that is subsequently confirmed.
 %whose components are $\hat{y}_{it,MSE}$, $i=1,...,n$. 
@@
 Consequently, we propose the following multivariate M-SSA criterion for $i=1,...,n$: 
-\begin{eqnarray}\label{mcrit1}
+\begin{eqnarray}\label{eq:mcrit1}
 \textrm{max}_{\mathbf{b}_i}\boldsymbol{\gamma}_{i\cdot h}'\tilde{\mathbf{I}}\mathbf{b}_{i}\\
@@
 Finally, the HT constraint $\mathbf{b}_{i}'\tilde{\mathbf{M}}\mathbf{b}_{i}=\rho_i$, where we used $l=1$, serves to control the first-order ACF or, equivalently, the HT of the predictor.
-\\
 
 Finally, we can relax the white noise hypothesis and assume a stationary process 
-\begin{equation}\label{x_m_st}
+\begin{equation}\label{eq:x_m_st}
 \mathbf{X}_t = \sum_{k=0}^\infty \boldsymbol{\Xi}_k \boldsymbol{\epsilon}_{t-k},
@@
 For given $(\mathbf{B}\cdot\boldsymbol{\Xi})_j$, $j=0,...,L-1$, the original coefficients $\mathbf{B}_{k}$ can be derived through deconvolution:
-\begin{eqnarray}\label{con_inv}
+\begin{eqnarray}\label{eq:con_inv}
 \mathbf{B}_{k}&=&\left\{\begin{array}{cc}(\mathbf{B}\cdot\boldsymbol{\Xi})_{0}\boldsymbol{\Xi}_0^{-1}&,k=0\\
-\left(  (\mathbf{B}\cdot\boldsymbol{\Xi})_{k}-\sum_{m=0}^{k-1}\mathbf{B}_m \boldsymbol{\Xi}_{j-m}\right)\boldsymbol{\Xi}_0^{-1}&,k=1,...,L-1
+\left((\mathbf{B}\cdot\boldsymbol{\Xi})_{k}-\sum_{m=0}^{k-1}\mathbf{B}_m \boldsymbol{\Xi}_{j-m}\right)\boldsymbol{\Xi}_0^{-1}&,k=1,...,L-1
 \end{array}\right.
@@
 The generalized M-SSA criterion can then be expressed as:
-\begin{eqnarray}\label{gen_stat_x}
+\begin{eqnarray}\label{eq:gen_stat_x}
 %\max_{(\mathbf{b}\cdot\boldsymbol{\xi})_i}\frac{ (\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{ih}'\mathbf{\tilde{I}} (\mathbf{b}\cdot\boldsymbol{\xi})_i}{\sqrt{(\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{ih}'\mathbf{\tilde{I}}(\boldsymbol{\gamma}\cdot\boldsymbol{\xi})_{ih}}\sqrt{(\mathbf{b}\cdot\boldsymbol{\xi})_i'\mathbf{\tilde{I}}(\mathbf{b}\cdot\boldsymbol{\xi})_i}}
@@
 In analogy to the above white noise case, the criterion involves affine and quadratic functions of $(\mathbf{b}\cdot\boldsymbol{\xi})_i$ and thus can be solved for the convolution weights.
-The requested M-SSA coefficients $\mathbf{b}_i$ can subsequently be determined through deconvolution, as indicated in Eq.~\eqref{con_inv}.
+The requested M-SSA coefficients $\mathbf{b}_i$ can subsequently be determined through deconvolution, as indicated in Eq.~\eqref{eq:con_inv}.
 The present discussion is confined to stationary processes, omitting a lengthier treatment of rank-defficient designs (cointegration) in the non-stationary case.
@@
 The subsequent analysis employs the M-SSA framework to construct `smooth' predictors for forecasting $z_{it+h}$ as well as for forecasting the original (unfiltered) GDP.
-In particular, it is shown that controlling the zero-crossing rate within the M-SSA framework has a quantifiable effect on the hit rate and false positive rate.
-
-\subsection{Out of sample M-MSE vs. M-SSA Nowcasts}
-
+In particular, it is shown that controlling the sign change rate within the M-SSA framework has a quantifiable effect on the hit rate and false positive rate.
+
+\section{Empirical Results: M-SSA Performance\label{sec:performance_mssa}}
+
+\subsection{M-MSE vs. M-SSA Nowcasts}
+%Out of sample?
 For illustration, consider the problem of nowcasting HP-filtered GDP, denoted by HP-GDP, where the target $z_t$ is generated using the two-sided HP(160) filter examined previously.
@@
         \includegraphics[height=3in, width=4.5in]{./Figures/bk_gammak.pdf}
-        \caption{Multivariate M-MSE (left) and M-SSA (right) nowcast filter weights.\\
+        \caption{M-MSE (left) and M-SSA (right) nowcast filter weights.\\
         Note the difference in scales.
-        \label{bk_gammak}}
+        \label{fig:bk_gammak}}
     \end{center}
@@
         \includegraphics[height=3in, width=4.5in]{./Figures/mssa_msse_now.pdf}
-        \caption{Nowcasts of Smoothed GDP\\
-        \indent Black: Target (Two-sided HP(160) applied to GDP) \\ 
-        \indent Blue: M-SSA Nowcast\\
-        \indent Green: M-MSE Nowcast\\
-        \indent The dashed vertical line delimits in- and out-of sample periods.  \\
+        \caption{Nowcasts of Smoothed GDP.\\
+        M-SSA (blue) and M-MSE (green) nowcasts are compared with the two-sided HP(160) GDP target (black). The dashed vertical line delimits in- and out-of sample periods. 
         Note that the two-sided filter does not extend to the sample end.
-        \label{mssa_msse_now}}
+        \label{fig:mssa_msse_now}}
     \end{center}
@@
     \begin{center}
-%    \textcolor{blue}{@Marc: Please change the plot Title to "M-MSE \& M-SSA Nowcasts: Financial Crisis".)}
         \includegraphics[height=3in, width=4.5in]{./Figures/mssa_msse_zc.pdf}
-        % \caption{Sign changes of the M-MSE (green) and M-SSA (blue) nowcasts towards the financial crisis, with sign changes of the M-MSE marked by green vertical lines.
-The sample holding time (HT) of each filter corresponds to the average duration between its consecutive zero-crossings.
-For each filter, its sample HT converges to the respective expected HT asymptotically.
-Here, we impose  $HT_{MSSA}=1.5\times HT_{MSE}$: in the long run, we expect the M-MSE to exhibit approximately $50\%$ more sign changes than the M-SSA.
-        \caption{M-MSE (green) and M-SSA (blue) nowcasts around the financial crisis.\\
-        Sign changes of the M-MSE marked by green vertical lines.\\
+        \caption{Nowcasts of Smoothed GDP during the financial crisis.\\
+        M-SSA (blue) and M-MSE (green) nowcasts. 
+        Sign changes of the M-MSE and M-SSA are marked by green and blue vertical lines, respectively; overlap of these lines at start and end of the recession indicates no loss of timeliness despite a stronger smoothing.
         The holding time (HT) of each filter corresponds to the expected duration between its sign-changes. 
@@
         Here, we impose  $HT_{MSSA}=1.5\times HT_{MSE}$. In large samples, we expect the M-SSA to exhibit approximately $33\%$ fewer sign changes than the M-MSE.
-        \label{mssa_msse_zc}}
-    \end{center}
-\end{figure}
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May  7 12:49:47 2025
-\begin{table}[ht]
-\caption{Sample target correlations and holding times (HT) of M-SSA and M-MSE nowcasts. Full Sample (excl. pandemic).} 
-\label{corhtnow}
-\centering
-\begin{tabular}{lrr}
-  \hline
- & M-SSA & M-MSE \\ 
-  \hline
-Target correlation & 0.68 & 0.69 \\ 
-  HT & 6.50 & 3.71 \\ 
-   \hline
-\end{tabular}
-
-\end{table}
-
-Figure \eqref{bk_gammak} displays the coefficients of the multivariate nowcasts, where M-SSA's increased smoothing reduces the maximum amplitudes of the filter weights and slows their rate of decay at longer lags.
-The resulting filtered series, presented in Figure \ref{mssa_msse_now}, are standardized to enable direct comparisons.
+        \label{fig:mssa_msse_zc}}
+    \end{center}
+\end{figure}
+
+%\begin{table}[ht]
+%\caption{Target Correlations and Holding Times of M-SSA and M-%MSE Nowcasts.\\Full Sample (excl. pandemic).} 
+%\label{tab:corhtnow}
+%\centering
+%\begin{tabular}{lrr}
+%  \hline
+% & M-SSA & M-MSE \\ 
+%  \hline
+%Target correlation & 0.68 & 0.69 \\ 
+% HT & 6.50 & 3.71 \\ 
+%   \hline
+%\end{tabular}
+
+%\end{table}
+
+Figure \ref{fig:bk_gammak} displays the coefficients of the multivariate nowcasts, where M-SSA's increased smoothing reduces the maximum amplitudes of the filter weights and slows their rate of decay at longer lags.
+The resulting filtered series, presented in Figure \ref{fig:mssa_msse_now}, are standardized to enable direct comparisons.
 The M-SSA nowcast is slightly smoother than that of the M-MSE, exhibiting approximately one-third fewer sign-changes.
-Figure \ref{mssa_msse_zc} compares both nowcasts in relation to the financial crisis, highlighting that the classic (MSE-based) design tends to produce unwanted `noisy' zero-crossings at (or in the vicinity of) the transitions between recessions and expansions.
-Further, Table \ref{corhtnow} compares sample target correlations and HTs of the two nowcasts.
-Increased smoothness, i.e.
-sample holding time increases from 3.7 to 6.5 quarters, achieved by the M-SSA comes at the cost of a slight reduction (0.68 vs 0.69) in target correlation.
-This highlights the inherent trade-off—often referred to as the prediction dilemma—between smoothness (zero-crossing rate) and predictive accuracy (target correlation) within the M-SSA criterion.
+Figure \ref{fig:mssa_msse_zc} compares both nowcasts in relation to the financial crisis, highlighting that the classic (MSE-based) design tends to produce unwanted `noisy' sign changes at (or in the vicinity of) the transitions between recessions and expansions.
+
+%Table \ref{tab:corhtnow}
+We find that increased smoothness by the M-SSA, i.e., sample holding time increases from 3.7 to 6.5 quarters, comes at the cost of a slight reduction (0.68 vs 0.69) in target correlation.
+This highlights the inherent trade-off—often referred to as the prediction dilemma—between smoothness (sign change rate) and predictive accuracy (target correlation) within the M-SSA criterion.
 The results confirm that the smoothing constraint in M-SSA 
-% incorporated into the multivariate extension of criterion \eqref{critssa} effectively manages the sign-changes of the predictor: sample estimates of the holding times align with the expected values derived from Eq.~\eqref{ht}.\footnote{Effective convergence of sample estimates to their theoretical expected values —assuming the true data-generating process is known— can be validated using the M-SSA package.
-This validation is typically performed through analyses of very long samples of simulated (multivariate) data, ensuring the consistency of the estimation procedures under idealized conditions.} An
 increases by $50\%$ the HT relative to that of the M-MSE predictor\footnote{The convergence of the sample HT to the effectively imposed $HT_1$ can be verified in a simulation study, based on very long samples of the time series, utilizing the M-SSA package.} but has only a modest effect on the target correlation, suggesting that this trade-off could be beneficial in our application.
-% , as confirmed by a subsequent analysis of hit rate and false alarm rate
-. %Going forward, we now omit the M-MSE, as it is a special case of the more general M-SSA framework.
 
@@
 
-Next, we 
-% extend the previous example to 
-consider forecasts of HP-GDP, changing the previous
-%. The only modification this requires is to change 
+Next, we consider forecasts of HP-GDP, changing the previous 
 $h=0$ (nowcast) to $h=1,...,6$ quarters. 
-%The previously described nowcast example can be systematically extended to generate forecasts of the target variable.
-This extension utilizes the same empirical framework, incorporating an assumed $50\%$ increase in the HT over the M-MSE, and considers forward shifts of   (note that M-SSA does not target GDP explicitly).
-%This approach facilitates the assessment of forecast performance across multiple horizons within the established methodological context.
-The influence of the forecast horizon on the M-SSA design is depicted in Figure \eqref{bk_h}, which compares M-SSA filter weights for the nowcast and the one-year-ahead forecast($h=4$).
+The influence of the forecast horizon on the M-SSA design is depicted in Figure \ref{fig:bk_h}, which compares M-SSA filter weights for the nowcast and the one-year-ahead forecast($h=4$).
 Noting the difference in vertical scales, we see that as $h$ increases (right panel), the filter weights diminish, reflecting increased forecast uncertainty.
@@
 While this phase effect might cause a sign change in the forecast as $h$ increases — something to be avoided in this context — the low-pass filters assigned to the additional `leading' indicators (green and yellow lines, right panel) continue to effectively track the low-frequency components of HP-GDP.
-The combination of the phase effect, applied to GDP, and level-tracking through these additional indicators enables more refined and effective tracking of the target compared to univariate forecasts, which rely solely on the phase-effect of the filter to look ahead.
+The combination of the phase effect, applied to GDP, and level-tracking through these additional indicators enables more refined and effective tracking of the target compared to univariate forecasts, which rely solely on the phase effect of the filter to look ahead.
 
@@
         \includegraphics[width=0.85\textwidth]{./Figures/bk_h.pdf}
-        \caption{Multivariate M-SSA filter weights.\\
+        \caption{M-SSA filter weights by forecast horizon.\\
         Left: $h=0$ (nowcast) \qquad Right: $h=4$ (one year forecast)\\
         Note the difference in scales.
-        \label{bk_h}}
-    \end{center}
-\end{figure}
-
-M-SSA predictors are compared to the traditional univariate HP-C in Figures  \ref{multivar_vs_univar} and \ref{mssa_hpc_financial_crisis}.
-In the first figure, we compare both filters across the periods of the financial crisis (top panels) and the COVID-19 pandemic (bottom panels), for forecast horizons $h=0,...,6$.
+        \label{fig:bk_h}}
+    \end{center}
+\end{figure}
+
+M-SSA predictors are compared to the traditional univariate HP-C in Figures \ref{fig:multivar_vs_univar} and \ref{fig:mssa_hpc_financial_crisis}.
+In the first figure, we compare both filters across the periods of the financial crisis (left panels) and the COVID-19 pandemic (right panels), for forecast horizons $h=0,...,6$.
 In the second figure, we compare both filters to the forward-shifted GDP for shifts larger than two quarters.
@@
         \includegraphics[width=0.85\textwidth]{./Figures/multivar_vs_univar.pdf}
-        \caption{Now- and forecasts of HP-GDP for $h=0,...,6$: HP-C (top) vs. M-SSA (bottom) across the financial crisis (left) and the pandemic (right).
-        \label{multivar_vs_univar}}
+        \caption{Now- and forecasts across horizons.\\ Forecasts with HP-C (top) vs.
+M-SSA (bottom) for horizons $h=0,...,6$ are shown for the financial crisis (left panels) and the pandemic (right panels).
+        \label{fig:multivar_vs_univar}}
     \end{center}
@@
         \includegraphics[width=0.85\textwidth]{./Figures/mssa_hpc_financial_crisis.pdf}
-        \caption{Forward-shifted GDP (black), M-SSA (blue) and HP-C (red) for $h=3,4,5,6$ across the financial crisis. All series are standardized for ease of comparison.
-        \label{mssa_hpc_financial_crisis}}
+        \caption{Forward-shifted GDP vs. forecasts.\\
+         M-SSA (blue) and HP-C (red) forecasts to forward-shifted GDP (black) for horizons h=3 – 6 during the financial crisis. All series are standardized for ease of comparison.
+        \label{fig:mssa_hpc_financial_crisis}}
     \end{center}
@@
 
-\section{M-SSA Component Predictor}\label{sec:mSSA_component}
-
-The targets of the multivariate filters above were the smoothed outputs obtained from the acausal two-sided HP(160) filter applied to the five indicators.
-However, our objective is to develop predictors for unfiltered GDP.
-To achieve this, we use the above M-SSA outputs as predictors for forward-shifted GDP.
+\subsection{M-SSA GDP Predictor}\label{sec:mSSA_component}
+
+%The final step in our framework uses the M-SSA results, obtained from the acausal two-sided HP(160) filter applied to the five indicators, as a predictor for actual GDP growth.
+This is achieved through a simple regression of the target series on the so-called M-SSA GDP predictor, estimated separately for each forecast horizon.
+
+The final step in our framework uses the M-SSA filtered series as a predictor for actual GDP growth.
+This is achieved through a simple regression of forward-shifted GDP on the M-SSA predictor, where M-SSA is optimized for tracking (the acausal) HP-GDP at various forecast horizons.
 
 %(this contrasts with Section \eqref{sec:direct_forecast}, which was based on the full sample information). 
-All the results that we present here exclude the pandemic; results including the pandemic are available in the Appendix.\\
+All the results that we present exclude the pandemic; further results including the pandemic are available in the Online Appendix.\\
 
@@
 
-\subsection{Tracking GDP: Optimal Weighting}
-
-Let $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$ denote the M-SSA predictor of HP-GDP at time $t$ for forecast horizon $h$. A predictor of GDP can be obtained from the following regression specification
+\subsection*{Tracking GDP: Optimal Weighting}
+
+Let $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$ denote the M-SSA predictor of HP-GDP at time $t$ for forecast horizon $h$.
+A GDP forecast can be obtained from the following regression specification\footnote{By default, the M-SSA criterion \eqref{eq:gen_stat_x} generates a standardized predictor and the regression coefficients are calibrated to match the level and scale of GDP.}
 \begin{equation}
-\textrm{GDP}_{t+shift}=\beta_0(shift,h)+\beta_1(shift,h) \cdot \textrm{M-SSA}^{\textrm{GDP}}_t(h) + \epsilon_t^{shift,h}
+\textrm{GDP}_{t+shift}=\beta_0(shift,h)+\beta_1(shift,h) \cdot \textrm{M-SSA}^{\textrm{GDP}}_t(h) + \epsilon_t^{shift,h}.
 \label{eq:GDP_M-SSA}
@@
 This flexibility enables leveraging the dynamic left-shift of $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$, illustrated in Figure \eqref{multivar_vs_univar}, when $shift<h$.\\
-The corresponding out-of-sample predictor $\hat{\textrm{GDP}}_{t+shift}(h)$ is given by :  
-\begin{equation}\label{oos_pred}
-\hat{\textrm{GDP}}_{t+shift}(h):=\hat{\beta}_0(shift,h,t)+\hat{\beta}_1(shift,h,t) \cdot \textrm{M-SSA}^{\textrm{GDP}}_t(h),
+The corresponding out-of-sample predictor $\widehat{\textrm{GDP}}_{t+shift}(h)$ is given by: 
+\begin{equation}\label{eq:oos_pred}
+\widehat{\textrm{GDP}}_{t+shift}(h):=\hat{\beta}_0(shift,h,t)+\hat{\beta}_1(shift,h,t) \cdot \textrm{M-SSA}^{\textrm{GDP}}_t(h),
 \end{equation}
 where the regression parameters $\hat{\beta}_0(shift,h,t),\hat{\beta}_1(shift,h,t)$ are estimated recursively using data up to time point $t-1$ (expanding window).
-Since M-SSA is based on data up to Q4-2007, $\hat{\textrm{GDP}}_{t+shift}(h)$ constitutes an out-of-sample predictor of $\textrm{GDP}_{t+shift}$ for all $shift\geq 0$ and all $t$ larger than Q4-2007.
-\footnote{Equations \eqref{eq:GDP_M-SSA} and \eqref{oos_pred} could be extended to incorporate additional M-SSA components, but for illustration purposes, we focus on the simplest (and most robust) design.}
+Since the M-SSA predictor is based on data up to Q4-2007, $\widehat{\textrm{GDP}}_{t+shift}(h)$ constitutes an out-of-sample predictor of $\textrm{GDP}_{t+shift}$ for all $shift\geq 0$ and all $t$ larger than Q4-2007.\footnote{Equations \eqref{eq:GDP_M-SSA} and \eqref{eq:oos_pred} could be extended to incorporate additional M-SSA components, but for illustration purposes, we focus on the simplest (and most robust) design.}
 The associated out-of-sample forecast error is defined as
-\begin{equation}\label{oosfe}
-\hat{\epsilon}_t^{shift,h}:={\textrm{GDP}}_{t+shift}-\hat{\textrm{GDP}}_{t+shift}(h).
+\begin{equation}\label{eq:oosfe}
+\hat{\epsilon}_t^{shift,h}:={\textrm{GDP}}_{t+shift}-\widehat{\textrm{GDP}}_{t+shift}(h).
 \end{equation}
@@
 
-\subsection{Relative MSFEs}\label{oosp}
-
-The statistical significance of the (out of sample) GDP component predictor in Equation \eqref{oos_pred}  can be evaluated by:
-\begin{equation}\label{eq_gdp_pred}
-\textrm{GDP}_{t+shift}=\alpha_0+\alpha_1\hat{\textrm{GDP}}_{t+shift}(h)+\nu_{t}^{shift,h},
+\subsection*{Relative MSFEs}\label{sec:oosp}
+
+The statistical significance of the (out of sample) M-SSA GDP predictor in Eq.~\eqref{eq:oos_pred} can be evaluated by:
+\begin{equation}\label{eq:eq_gdp_pred}
+\textrm{GDP}_{t+shift}=\alpha_0+\alpha_1\widehat{\textrm{GDP}}_{t+shift}(h)+\nu_{t}^{shift,h},
 \end{equation}
@@
 Statistical inference is based on heteroskedasticity- and autocorrelation-consistent (HAC) standard errors, with the null hypothesis $H_0 : \alpha_1 =0$.
-The significance levels of the corresponding test statistics are indicated by $^{*}$ in Table \ref{rRMSE_mSSA_comp_mean_wc}.\footnote{$^{*}$ denotes significance at the $5\%$ level; $^{**}$ indicates significance at the $0.1\%$ level.
-Note that statistical significance is influenced by the relatively short out-of-sample period of 17 years.} Forecast accuracy is evaluated in terms of relative root mean squared forecast errors (rRMSEs), where the MSE forecast performance of $\hat{\textrm{GDP}}_{t+shift}(h)$ can be benchmarked against the mean predictor (Table \ref{rRMSE_mSSA_comp_mean_wc}) as well as against the direct forecast (see Table \ref{rRMSE_mSSA_comp_direct_without_covid7}).
-For consistency, both benchmark predictors are quarterly updated out-of-sample designs.\\
+The significance levels of the corresponding test statistics are indicated by $^{*}$ in Table \ref{tab:rRMSE_mSSA_comp_mean_wc}.\footnote{$^{*}$ denotes significance at the $5\%$ level; $^{**}$ indicates significance at the $1\%$ level.
+Note that statistical significance is influenced by the relatively short out-of-sample period of 17 years.} Forecast accuracy is evaluated in terms of relative root mean squared forecast errors (rRMSEs), where the MSE forecast performance of $\widehat{\textrm{GDP}}_{t+shift}(h)$ can be benchmarked against the mean predictor (Table \ref{tab:rRMSE_mSSA_comp_mean_wc}) as well as against the direct forecast (see Table \ref{tab:rRMSE_mSSA_comp_direct_without_covid7}).
+For consistency, both benchmark predictors are quarterly updated out-of-sample designs.
 
 % Obviously, predicting HP-GDP (see Table \eqref{p_val_hpbip_wc}) is less challenging than predicting GDP (see Table \eqref{p_val_wc}), due to smoothness (autocorrelation) of the target when compared to GDP.
-
 Our results can be summarized as follows.
-When excluding the pandemic, the direct forecasts outperform the mean forecast up to two quarters ahead, see Table \ref{rRMSE_mSSA_direct_mean_without_covid8} in the appendix.
+When excluding the pandemic, the direct forecasts outperform the mean forecast up to two quarters ahead (Table \ref{tab:rRMSE_mSSA_direct_mean_without_covid8} in the appendix).
 %\footnote{However, the pandemic period generally obscures the relationships between predictors and GDP.}.
-In contrast, the GDP predictors $\hat{\textrm{GDP}}_{t+shift}(h)$, optimized for longer forecast horizons $h\geq 5$, outperform the mean benchmark up to 5 quarters ahead, see Table \ref{rRMSE_mSSA_comp_mean_wc}.
-Additionally,  $\hat{\textrm{GDP}}_{t+shift}(h)$, optimized for large forecast horizons $h\geq 4$, surpass the direct forecasts for $shift\geq 3$, see Table \ref{rRMSE_mSSA_comp_direct_without_covid7} in the appendix.
-Statistical significance of the GDP predictor, optimized for larger forecast horizons, can be established for at least up to one year ahead, see Table \eqref{rRMSE_mSSA_comp_mean_wc}.
-Finally, the strong performance of predictors optimized for larger forecast horizons ($h\geq 4$), at smaller shifts ($shifts< 4$) corroborates the importance of the left-shift of the corresponding predictors, see Figures \ref{multivar_vs_univar} and \ref{mssa_hpc_financial_crisis}, in relation to forecast performance.
+In contrast, the GDP predictors $\widehat{\textrm{GDP}}_{t+shift}(h)$, optimized for longer forecast horizons $h\geq 5$, outperform the mean benchmark up to 5 quarters ahead, see Table \ref{tab:rRMSE_mSSA_comp_mean_wc}.
+Additionally, $\widehat{\textrm{GDP}}_{t+shift}(h)$, optimized for large forecast horizons $h\geq 4$, surpass the direct forecasts for $shift\geq 3$ (Table \ref{tab:rRMSE_mSSA_comp_direct_without_covid7}).
+Statistical significance of the GDP predictor, optimized for larger forecast horizons, can be established for at least up to one year ahead.
+Finally, the strong performance of predictors optimized for larger forecast horizons ($h\geq 4$), at smaller shifts ($shifts< 4$) corroborates the importance of the left-shift of the corresponding predictors, see Figures \ref{fig:multivar_vs_univar} and \ref{fig:mssa_hpc_financial_crisis}, in relation to forecast performance.
 
@@
 
-\caption{rRMSEs: Out-of-sample forecast errors Equation \eqref{oosfe} vs expanding mean of GDP.\\Out-of-sample evaluation starting in Q1-2008, without pandemic\\Values $< 1$ indicate superior predictions from Equation \eqref{oos_pred}\\
-Row-index: Shift refers to the left-shift of the target $\textrm{GDP}_{t+shift}$\\
-Column-index: $h$ refers to the forecast horizon for which $\textrm{M-SSA}_t^{\textrm{GDP}}(h)$ is optimized\\
-Significance refers to the hypothesis $\alpha_1=0$ in Equation \eqref{eq_gdp_pred}.
+\caption{Relative RMSE of M-SSA GDP Predictors vs. Mean Benchmark.\\
+Out-of-sample rRMSEs for various shifts and horizons, excluding the pandemic.\\
+Shift refers to the left-shift of the target $\textrm{GDP}_{t+shift}$; $h$ refers to the forecast horizon for which $\textrm{M-SSA}_t^{\textrm{GDP}}(h)$ is optimized. 
+Values $< 1$ indicate superior predictions from Eq.~ \eqref{eq:oos_pred}. Significance refers to the hypothesis $\alpha_1=0$ in Eq.~\eqref{eq:eq_gdp_pred}.
 } 
-\label{rRMSE_mSSA_comp_mean_wc}
+\label{tab:rRMSE_mSSA_comp_mean_wc}
 \centering
@@
 
+\begin{table}[ht]
+\caption{Relative RMSEs of M-SSA GDP Predictors vs. Direct Forecasts.\\
+Out-of-sample evaluation starting in Q1-2008, without pandemic. Values below 1 indicate improvement over direct forecasting.
+\label{tab:rRMSE_mSSA_comp_direct_without_covid7}}
+\centering
+\begin{tabular}{rrrrrrrr}
+  \hline
+ & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
+  \hline
+Shift=0 & 1.131 & 1.092 & 1.028 & 0.973 & 0.977 & 1.040 & 1.119 \\ 
+  Shift=1 & 1.295 & 1.274 & 1.206 & 1.106 & 1.039 & 1.044 & 1.097 \\ 
+  Shift=2 & 1.180 & 1.194 & 1.178 & 1.109 & 1.029 & 0.990 & 0.997 \\ 
+  Shift=3 & 1.029 & 1.061 & 1.082 & 1.051 & 0.984 & 0.936 & 0.923 \\ 
+  Shift=4 & 1.012 & 1.057 & 1.100 & 1.096 & 1.036 & 0.967 & 0.924 \\ 
+  Shift=5 & 0.988 & 1.033 & 1.076 & 1.084 & 1.048 & 0.987 & 0.929 \\  
+   \hline
+\end{tabular}
+\end{table}
+
 \subsection{Hit Rate vs. False Alarm Rate}
@@
 
-ROC curves are generated by evaluating the relationship between the sign of the target variable $GDP_{t+shift}$  and events where the forecast $\hat{GDP}_{t+shift}(h)$ exceeds a threshold $\tau$.
+ROC curves are generated by evaluating the relationship between the sign of the target variable $GDP_{t+shift}$ and events where the forecast $\widehat{GDP}_{t+shift}(h)$ exceeds a threshold $\tau$.
 The threshold $\tau$ is varied smoothly across the range of the predictor space to construct the ROC curve.\footnote{Although we are using standardized GDP, the regressions automatically account for scales and level shifts, so the results directly apply to the original GDP.} We examined all pairwise combinations of $h$ and $shift$, but for simplicity of exposition, the results below focus on the diagonal case where $h=shift$.
 For clarity, these results are based on the entire sample, excluding the pandemic period.
-Figure \ref{ROC_GDP_shift_1_4} compares ROCs of the predictors for small and medium-term forecast horizons and Table \ref{auc} presents AUC (area under the curve) for forecast horizons 1-5 quarters ahead.
-
-\begin{figure}[h]\begin{center}\includegraphics[height=3in, width=5in]{./Figures/ROC_GDP_shift_1_4.pdf}\caption{Hit rate vs.
-false alarm rate of the predictors at small (one quarter ahead:left panel) and yearly forecast horizons (right panel).\label{ROC_GDP_shift_1_4}}\end{center}\end{figure}
-% latex table generated in R 4.2.2 by xtable 1.8-4 package
-% Wed May 21 10:06:12 2025
+Figure \ref{fig:ROC_GDP_shift_1_4} compares ROCs of the predictors for small and medium-term forecast horizons and Table \ref{tab:auc} presents AUC (area under the curve) for forecast horizons 1-5 quarters ahead.
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[height=3in, width=5in]{./Figures/ROC_GDP_shift_1_4.pdf}
+\caption{Hit rate vs. false alarm rate.\\ 
+ROC curves compare the classification performance of direct forecasts, HP-C forecasts, M-MSE, and M-SSA at one-quarter (left panel) and one-year (right panel) horizons.
+\label{fig:ROC_GDP_shift_1_4}}
+\end{center}
+\end{figure}
 
@@
 \caption{AUCs of predictors against GDP at various forecast horizons.} 
-\label{auc}
+\label{tab:auc}
 \centering
@@
 
-\section{Revisions}\label{sec:revisions}
-
-Revisions refer to changes in the historical values of a time series resulting from updates with new information.
-Revisions of the M-SSA component predictor can arise from data updates, which we do not analyze further here\footnote{Except for GDP, the other indicators are not or only slightly revised.
+\section{Robustness and Stability}\label{sec:revisions}
+Model parameters evolve as new observations become available, and forecasts are revised as incoming data update the information set.
+Similarly, revisions of the M-SSA GDP predictor can arise from data updates, which we do not analyze further here\footnote{Except for GDP, the other indicators are not or only slightly revised.
 Moreover, the smoothing effect of filters mitigates the effect of revisions when compared to direct forecasts.}, or from updates to predictor parameters.
@@
 
-\subsection{Revisions: Regression Equation}
-
-\begin{figure}[h]
+\subsection{Parameter Stability and Revisions}
+
+\begin{figure}[htpb]
     \begin{center}
         \includegraphics[height=3in, width=4.5in]{./Figures/revisions1.pdf}
-        \caption{Revisions of the M-SSA GDP component predictor are shown, with the final predictor in blue and the real-time design in red.
-These revisions result from quarterly updates to the regression coefficients of equation \eqref{oos_pred}.
+        \caption{Revisions of the M-SSA GDP predictor.\\
+        Comparison of final (blue) and real-time (red) M-SSA GDP  predictors for h=4, shift=4.
+These revisions result from quarterly updates to the regression coefficients of equation \eqref{eq:oos_pred}.
 The M-SSA filter    itself remains fixed, based on data up to Q4-2007.
-        \label{revisions1}}
+        \label{fig:revisions1}}
     \end{center}
@@
         \includegraphics[height=3in, width=4.5in]{./Figures/revisions2.pdf}
-        \caption{Revisions of intercept and regression weight of Equation \eqref{oos_pred}.
-        \label{revisions2}}
-    \end{center}
-\end{figure}
-
-For illustration, we focus on a forecast horizon $h=4$ for the GDP component predictor and a forward-shift $shift=4$ for GDP, noting that similar results would be obtained with other combinations of $h$ and $shift$.
-%???Check left-shift of MA-inversion of M-SSA and M-MSE
-%??? Check forecast excess for HP-C
-%??? Justification HP(160): replace HP(1600) by AR(2) with similar persistency but shorter cycle lengthi, i.e. $a_1=-2"persistency*acos(\theta, a_2=persistency$.
-Figure \ref{revisions1} compares the final and real-time predictors: as the sample size increases, the real-time predictor converges steadily to the final predictor.\footnote{The rate of convergence is partially determined by the simplicity of our design, which avoids incorporating additional M-SSA components.}
-As a confirmation, Figure \ref{revisions2} displays the evolution of the intercept and regression weights over time: with increasing sample size, these estimates seem to converge to fixed points, indicating both the stationarity of the process and the consistency of the estimates.
+        \caption{Revisions of intercept and regression weights.\\
+        Quarterly estimates of the intercept and GDP coefficient in the M-SSA regression for h=4, shift=4 according to Equation \eqref{eq:oos_pred}.
+        \label{fig:revisions2}}
+    \end{center}
+\end{figure}
+
+A key concern for real-time applications is the stability of forecast parameters and the magnitude of forecast revisions as new data arrive.
+For illustration, we focus on a forecast horizon $h=4$ for the M-SSA GDP predictor and a forward-shift $shift=4$ for GDP, noting that similar results would be obtained with other combinations of $h$ and $shift$.
+
+Figure \ref{fig:revisions1} compares the final and real-time predictors: as the sample size increases, the real-time predictor converges steadily to the final predictor.\footnote{The rate of convergence is partially determined by the simplicity of our design, which avoids incorporating additional M-SSA components.}
+In addition, Figure \ref{fig:revisions2} displays the evolution of the intercept and regression weights over time: with increasing sample size, these estimates seem to converge to fixed points, indicating both the stationarity of the process and the consistency of the estimates.
 %(Note, however, the slightly elevated values of the regression weight assigned to the M-SSA component during the financial crisis and the pandemic, suggesting a stronger link between the target and the predictor).
@@
 
-\subsection{Revisions: M-SSA}
-
-\begin{figure}[h!]
+\subsection{VAR Model Updates and Filter Revisions}
+%{Revisions: M-SSA}
+The second source of revisions in the M-SSA GDP predictor arises from updates to the VAR(1) model within the M-SSA filter.\footnote{Similar findings apply to more sophisticated alternatives such as the VAR(3) discussed in Section \ref{sec:data}.}
+We compare the original estimates, based on data up to Q4-2007, with the final estimates obtained using $h=4$ and $shift=4$ (similar findings apply to other values of forecast horizon and forward-shift).
+For consistency, we excluded the entire pandemic episode from the data.
+Figure \ref{fig:up_dated_ma_inv_multi_ip} compares the MA-inversions (impulse responses) of the GDP equation derived from the `old' and the updated VAR models.
+Distinctly, the updated model incorporates the ESI as an additional explanatory variable for GDP, reflecting evidence from the longer data sample.
+Figure \ref{fig:bk_2008_all} illustrates the resulting effect on the M-SSA filter targeting HP-GDP at forecast horizon $h=4$.
+The updated filter (right panel) assigns more weight to the additional `leading' indicators relative to GDP.
+Finally, Figure \ref{fig:revisions3} compares the M-SSA GDP predictors from the initial and updated VAR models.
+To isolate the effect of the regression revision, the figure displays standardized series.
+The update to the VAR model results in a leftward shift of the predictor (represented by the blue line), which appears marginally more rapid compared to the original predictor (depicted by the red line).
+It is important to note that the smoothness (sign change rate) remains unchanged, though.
+This suggests that the out-of-sample performance of the M-SSA GDP predictor may be better than reported in Section \ref{sec:oosp}, which is based on the fixed and increasingly outdated version of M-SSA.
+
+\begin{figure}[htbp]
     \begin{center}
         \includegraphics[height=3in, width=4.5in]{./Figures/up_dated_ma_inv_multi_ip.pdf}
-        \caption{MA-inversion (impulse response) of VAR: data up to Q4-2007 (left) vs. entire data set without pandemic (right).
-        \label{up_dated_ma_inv_multi_ip}}
-    \end{center}
-\end{figure}
-
-\begin{figure}[h!]
+        \caption{Impulse Responses from VAR Models: Initial vs. Full-Sample Estimates.\\
+        MA-inverted impulse responses for GDP from VAR models estimated with data up to 2007Q4 (left) and the full sample excluding the pandemic (right). 
+        \label{fig:up_dated_ma_inv_multi_ip}}
+    \end{center}
+\end{figure}
+
+\begin{figure}[htbp]
     \begin{center}
         \includegraphics[height=3in, width=4.5in]{./Figures/bk_2008_all.pdf}
-        \caption{M-SSA filter tracking HP-GDP at forecast horizon $h=4$: data up to Q4-2007 (left) vs. entire data set without pandemic (right).
-        \label{bk_2008_all}}
-    \end{center}
-\end{figure}
-
-\begin{figure}[h!]
+        \caption{M-SSA Filter Weights: Initial vs. Full-Sample Estimates.\\
+        M-SSA filters for HP-GDP at $h=4$ using 2007Q4 (left panel) and full-sample (right panel) VAR models (without pandemic).
+        \label{fig:bk_2008_all}}
+    \end{center}
+\end{figure}
+
+\begin{figure}[htpb]
     \begin{center}
         \includegraphics[height=3in, width=4.5in]{./Figures/revisions3.pdf}
-        \caption{Revisions resulting from updates to the VAR model in M-SSA optimized for $h=shift=4$.
+        \caption{M-SSA GDP Predictors: Initial vs. Full-Sample Estimates.\\
+Standardized M-SSA GDP predictors from initial (red) and updated (blue) VAR models for $h=shift=4$ (excluding pandemic). 
+                %Revisions resulting from updates to the VAR model in M-SSA optimized for $h=shift=4$.
 To isolate the effect of the regression revision, all series are standardized.
@@
 Full sample excluding the pandemic.
-        \label{revisions3}}
-    \end{center}
-\end{figure}
-
-The second source of revisions in the M-SSA component predictor arises from updates to the VAR(1) model within the M-SSA filter.\footnote{Similar findings apply to more sophisticated alternatives such as the VAR(3) discussed in Section \ref{sec:data}.} We compare the original estimates, based on data up to Q4-2007, with the final estimates obtained using $h=4$ and $shift=4$ (similar findings apply to other values of forecast horizon and forward-shift).
-For consistency, we excluded the entire pandemic episode from the data.
-Figure \ref{up_dated_ma_inv_multi_ip} compares the MA-inversions (impulse responses) of the GDP equation derived from the `old' and the updated VAR models: unlike the former, the latter incorporates ESI as an additional explanatory variable for GDP.
-Figure \ref{bk_2008_all} illustrates the resulting effect on the M-SSA filter targeting HP-GDP at forecast horizon $h=4$.
-The updated filter (right panel) assigns more weight to the additional `leading' indicators relative to GDP.
-Finally, Figure \ref{revisions3} compares the M-SSA GDP predictors.
-To isolate the effect of the regression revision, the figure displays standardized series.The update to the VAR model results in a leftward shift of the predictor (represented by the blue line), which appears marginally more rapid compared to the original predictor (depicted by the red line).
-It is important to note that the smoothness (zero-crossing rate) remains unchanged, though.
-This suggests that the out-of-sample performance of the M-SSA component predictor may be better than reported in Section \ref{oosp}, which is based on the fixed and increasingly outdated version of M-SSA.
-
-\section{Summary and Conclusion}\label{sec:conclusions}
+        \label{fig:revisions3}}
+    \end{center}
+\end{figure}
+
+\clearpage
+\section{Conclusion}\label{sec:conclusions}
 
@@
 To address these problems, we propose a multivariate extension of the SSA framework based on \cite{Wildi2024}.
-Unlike the univariate HP filter, the M-SSA can leverage cross-sectional information from indicators leading GDP in real time while controlling for the rate of zero-crossings of the predictor.
-Consequently, the multivariate filter outputs become increasingly left-shifted as the forecast horizon lengthens, allowing dips and peaks of the target series to be tracked more systematically, while reducing the occurrence of noisy zero-crossings.
-However, since M-SSA does not explicitly target GDP, an additional step is necessary to derive the final GDP predictor.\\% We propose two approaches: a simple equally-weighted aggregate, assuming all M-SSA components are equally important for predicting GDP, and an optimally weighted predictor based on a regression of the M-SSA components on future GDP.\\
-
-For illustration, we consider a simple approach based on regressing the single M-SSA GDP output on future GDP, thereby ignoring the other filter outputs.
+Unlike the univariate HP filter, the M-SSA can leverage cross-sectional information from indicators leading GDP in real time while controlling for the rate of sign changes of the predictor.
+Consequently, the multivariate filter outputs become increasingly left-shifted as the forecast horizon lengthens, allowing dips and peaks of the target series to be tracked more systematically, while reducing the occurrence of noisy sign changes.
+However, since M-SSA does not explicitly target GDP, an additional step is necessary to derive the final GDP predictor.% We propose two approaches: a simple equally-weighted aggregate, assuming all M-SSA components are equally important for predicting GDP, and an optimally weighted predictor based on a regression of the M-SSA components on future GDP.\\
+
+For illustration, we consider a simple approach based on regressing the M-SSA predictor (of HP-GDP) on future GDP.
 This predictor is intuitively appealing because the future HP-GDP, i.e., the target of M-SSA, is the low-frequency component of the future GDP target.
@@
 Multivariate approaches can further capitalize on leading indicators within the cross-sectional data.
-Additionally, the AUC is optimized through more precise control of the zero-crossing rate within the M-SSA framework.
-Finally, an examination of revision errors associated with quarterly updates of the proposed M-SSA-based GDP component predictor indicates that the regression parameters, VAR model, and resulting M-SSA filters stabilize after an initial burn-in period.
+Additionally, the AUC is optimized through more precise control of the sign change rate within the M-SSA framework.
+Finally, an examination of revision errors associated with quarterly updates of the proposed M-SSA-based GDP predictor indicates that the regression parameters, VAR model, and resulting M-SSA filters stabilize after an initial burn-in period.
 This stabilization enhances the model’s explainability and interpretability.
@@
 In conclusion, the new predictor design integrates the traditional direct forecast approach with a novel multivariate filter to target quarterly GDP growth up to one year ahead.
-This multivariate approach leverages a set of economic indicators leading GDP in real time while controlling the smoothness of the predictor through the zero-crossing rate.
-
+This multivariate approach leverages a set of economic indicators leading GDP in real time while controlling the smoothness of the predictor through the sign change rate.
+
+%%%%% Main findings:
+
+%Our analysis yields four main findings.
+%First, pre-filtering leading indicators with a one-sided HP-C filter improves the performance of simple direct forecasts by attenuating high-frequency noise, particularly at medium-term horizons.
+%Second, M-SSA further enhances forecast quality by jointly filtering multiple indicators, systematically shifting forecast peaks and troughs forward in time and improving turning-point detection.
+%Third, the trade-off between smoothness and accuracy can be explicitly managed via the holding-time constraint, enabling forecasters to suppress spurious zero-crossings without materially reducing predictive correlations.
+%Fourth, M-SSA is robust in real-time conditions: parameter estimates stabilize quickly, forecast revisions converge within a short window, and turning points in the M-SSA GDP Signal remain consistent under alternative VAR specifications.
 %%%%%%%%%%%%%%%%%%%
@@
 \newpage
-%\pagestyle{fancy}
-%\fancyhf{} % Setzt alle Kopf- und Fußzeilen zurück
-%\fancyhead[L]{M-SSA}
-%\fancyfoot[C]{\thepage} % Seitenzahlen in der Fußzeile wieder hinzufügen
-
-	\section{Appendix}\label{sec:appendix}
+\pagestyle{fancy}
+\fancyhf{} % Setzt alle Kopf- und Fußzeilen zurück
+\fancyhead[L]{Heinisch, van Norden, Wildi: Smooth and Persistent Forecasts of German GDP}
+\fancyfoot[C]{\thepage} % Seitenzahlen in der Fußzeile wieder hinzufügen
+
+	\section{Online Appendix}\label{sec:appendix}
 	\setcounter{table}{0}
@@
 
-\subsection{Additional Performances Results: Without Pandemic}
+\subsection{Additional Figures and Tables}
+
+\begin{figure}[ht]
+  \begin{subfigure}[htpb]{\textwidth}
+        \begin{center}
+            \includegraphics[height=3in, width=4.5in]{./Figures/CCF.pdf}
+            \caption{Full data set from Q2-1991 to Q4-2024.\label{fig:CCF}}
+        \end{center}
+    \end{subfigure}
+    \\
+    \begin{subfigure}[htpb]{\textwidth}
+        \begin{center}
+            \includegraphics[height=3in, width=4.5in]{./Figures/CCF_wc.pdf}
+            \caption{Without Pandemic (omitting Q4-2019 to Q4-2020).\label{fig:CCF_wc}}
+        \end{center}
+    \end{subfigure}
+    \caption{Sample cross-correlation functions (CCF) between GDP growth and Indicators.}
+\end{figure}
+
+\begin{figure}[htpb]
+    \begin{subfigure}[htpb]{\textwidth}
+        \begin{center}
+            \includegraphics[height=3in, width=6in]{./Figures/ma_inv_BIP.pdf}
+            \caption{implied by VAR(1) \label{fig:ma_inv_BIP}}
+        \end{center}
+    \end{subfigure}
+    \\
+    \begin{subfigure}[htpb]{\textwidth}
+        \begin{center}
+            \includegraphics[height=3in, width=6in]{./Figures/ma_inv_BIP_var3_en.pdf}
+            \caption{implied by VAR(3)\label{fig:ma_inv_BIP_var3_en}}
+        \end{center}
+    \end{subfigure}
+    \caption{Impulse Responses of GDP to Indicator Shocks from VAR Models.\\  VMA coefficients for GDP implied by VAR models.
+    \\Estimation with pandemic (left) and without pandemic (right).}
+\end{figure}
+
+\begin{figure}[H]
+    \begin{center}
+        \includegraphics[width=0.85\textwidth]{./Figures/direct_hp_forecasts.pdf}
+        \caption{Direct GDP forecasts using HP-C filtered Indicators (blue) vs. GDP shifted forward (to the left) by 0-3 quarters (black line).\\ Full data set (without pandemic).
+        \label{fig:direct_hp_forecasts}}
+    \end{center}
+\end{figure}
+
+\begin{figure}[h]
+    \begin{center}
+%    \textcolor{blue}{\textbf{Notice that the title on top right panel is incorrect!}}
+        \includegraphics[height=4in, width=5.5in]{./Figures/hp_160.pdf}
+        % \caption{Two-sided HP(160) (top left) and one-sided (concurrent) HP-C filters (top right) with corresponding amplitude functions (bottom).
+        \caption{HP$(\lambda=160)$ filters and Frequency Response.\\
+        Left: HP (Two-sided, symmetric) \qquad Right: HP-C (One-sided, concurrent)\\
+        Top Row: Filter Weights \qquad \qquad \quad Bottom Row: Amplitude Functions.
+        \label{fig:hp_160}}
+    \end{center}
+\end{figure}
 
 \begin{table}[ht]
-\caption{Regression of shifted GDP on M-SSA($h$) GDP component.\\rRMSEs benchmarked against the direct forecast.\\Out-of-sample evaluation starting in Q1-2008, without pandemic.} 
-\label{rRMSE_mSSA_comp_direct_without_covid7}
+\caption{Relative RMSEs of Direct Forecasts vs. Mean Benchmark.\\
+Out-of-sample rRMSEs for direct forecasts benchmarked against the expanding-mean, excluding the pandemic. Values below 1 indicate improvement over the mean.
+\label{tab:rRMSE_mSSA_direct_mean_without_covid8}}
 \centering
-\begin{tabular}{rrrrrrrr}
-  \hline
- & h=0 & h=1 & h=2 & h=3 & h=4 & h=5 & h=6 \\ 
-  \hline
-Shift=0 & 1.131 & 1.092 & 1.028 & 0.973 & 0.977 & 1.040 & 1.119 \\ 
-  Shift=1 & 1.295 & 1.274 & 1.206 & 1.106 & 1.039 & 1.044 & 1.097 \\ 
-  Shift=2 & 1.180 & 1.194 & 1.178 & 1.109 & 1.029 & 0.990 & 0.997 \\ 
-  Shift=3 & 1.029 & 1.061 & 1.082 & 1.051 & 0.984 & 0.936 & 0.923 \\ 
-  Shift=4 & 1.012 & 1.057 & 1.100 & 1.096 & 1.036 & 0.967 & 0.924 \\ 
-  Shift=5 & 0.988 & 1.033 & 1.076 & 1.084 & 1.048 & 0.987 & 0.929 \\  
+\begin{tabular}{rrrrrrr}
+  \hline
+ & Shift=0 & Shift=1 & Shift=2 & Shift=3 & Shift=4 & Shift=5 \\ 
+  \hline
+rRMSE & 0.862 & 0.838 & 0.912 & 1.008 & 1.001 & 1.028 \\  
    \hline
@@
 
+\newpage
+\subsection{Additional Tables and Performance Metrics (including Pandemic)}
+
 \begin{table}[ht]
-\caption{Direct forecasts.\\
-rRMSEs benchmarked against the expanding mean.\\Out-of-sample evaluation starting in Q1-2008, without pandemic.} 
-\label{rRMSE_mSSA_direct_mean_without_covid8}
-\centering
-\begin{tabular}{rrrrrrr}
-  \hline
- & Shift=0 & Shift=1 & Shift=2 & Shift=3 & Shift=4 & Shift=5 \\ 
-  \hline
-rRMSE & 0.862 & 0.838 & 0.912 & 1.008 & 1.001 & 1.028 \\  
-   \hline
-\end{tabular}
-
-\end{table}
-
-\newpage
-\subsection{Performances: Including Singular Pandemic Data}
-
-\begin{table}[ht]
-\caption{p-values for $H_0: {\alpha_1} = 0$, HAC-adjusted Wald test, Equation \eqref{eq_gdp_pred}.\\ including the pandemic.} 
-\label{p_val1}
+\caption{Significance of M-SSA GDP Component Predictor (Including Pandemic).\\
+p-values for $H_0: {\alpha_1} = 0$, HAC-adjusted Wald test, Equation \eqref{eq:eq_gdp_pred}.\\ including the pandemic.} 
+\label{tab:p_val1}
 \centering
@@
 \end{tabular}
-
 \end{table}
@@
 \begin{table}[ht]
-\caption{rRMSEs: Out-of-sample forecast errors Equation \eqref{oosfe} vs expanding mean of GDP.\\Out-of-sample evaluation starting in Q1-2008, including the pandemic.\\Values $< 1$ indicate superior predictions from Equation \eqref{oos_pred}.}
-\label{rRMSE_mSSA_comp_mean2}
+\caption{Relative RMSEs of M-SSA GDP Component Predictors vs. Mean Benchmark (Including Pandemic).
+%rRMSEs: Out-of-sample forecast errors Equation \eqref{oosfe} vs expanding mean of GDP.\\Out-of-sample evaluation starting in Q1-2008, including the pandemic.\\Values $< 1$ indicate superior predictions from Equation \eqref{oos_pred}.}
+\label{tab:rRMSE_mSSA_comp_mean2}}
 \centering
@@
 \end{tabular}
-
 \end{table}
 
-\begin{table}[ht]
-\caption{Regression of shifted GDP on M-SSA($h$) GDP component.\\rRMSEs benchmarked against the direct forecast.\\Out-of-sample evaluation starting in Q1-2008, including the pandemic.} 
-\label{rRMSE_mSSA_comp_direct3}
+\begin{table}[htpb]
+\caption{Relative RMSEs of Direct Forecasts vs. Mean Benchmark (Including Pandemic).
+%Relative RMSEs of M-SSA GDP Component Predictors vs. Direct Forecasts (Including Pandemic).
+%Regression of shifted GDP on M-SSA($h$) GDP component.\\rRMSEs benchmarked against the direct forecast.\\Out-of-sample evaluation starting in Q1-2008, including the pandemic.} 
+\label{tab:rRMSE_mSSA_comp_direct3}}
 \centering
@@
 rRMSEs benchmarked against the expanding mean.\\Out-of-sample evaluation starting in Q1-2008, including the pandemic.} 
-\label{rRMSE_mSSA_direct_mean4}
+\label{tab:rRMSE_mSSA_direct_mean4}
 \centering

### 2025-08-15 Sweave/BIP_predictor.Rnw
up

@@
 # We can select type of VAR model
+# This is the setting used in the paper
+VAR_type="VAR"
 #VAR_type="BVAR"
-VAR_type="VAR"
 #VAR_type="elastic_net"

### 2025-08-16 Archive IWH Simon paper/16.08.2025/paper.tex
up

--- 2025-08-15 Archive IWH Simon paper/15.08.2025 afetr Katja's feedback/paper.tex
@@
 Let $\textrm{M-SSA}^{\textrm{GDP}}_t(h)$ denote the M-SSA predictor of HP-GDP at time $t$ for forecast horizon $h$.
-A GDP forecast can be obtained from the following regression specification\footnote{By default, the M-SSA criterion \eqref{eq:gen_stat_x} generates a standardized predictor and the regression coefficients are calibrated to match the level and scale of GDP.}
+A GDP forecast can be obtained from the following regression specification\footnote{By default, the M-SSA criterion \eqref{eq:gen_stat_x} generates a standardized predictor and the regression coefficients are calibrated to match the level and scale of the forecast target optimally.}
 \begin{equation}
@@
 Finally, the strong performance of predictors optimized for larger forecast horizons ($h\geq 4$), at smaller shifts ($shifts< 4$) corroborates the importance of the left-shift of the corresponding predictors, see Figures \ref{fig:multivar_vs_univar} and \ref{fig:mssa_hpc_financial_crisis}, in relation to forecast performance.
-
-%???check whether rMSE is decreasing towards sample end (because initial estimate towards 2007 is unreliable)???
+Remarkably, these findings hold even though M-SSA was restricted to a relatively short in-sample window ending in Q4-2007, i.e., prior to the onset of the financial crisis.
 
 \begin{table}[ht]
-
 \caption{Relative RMSE of M-SSA GDP Predictors vs. Mean Benchmark.\\ 
-Out-of-sample rRMSEs for various shifts and horizons, excluding the pandemic.\\
+M-SSA is fixed, based on data up to Q4-2007. Out-of-sample evaluation starting in Q1-2008, without pandemic.\\
 Shift refers to the left-shift of the target $\textrm{GDP}_{t+shift}$; $h$ refers to the forecast horizon for which $\textrm{M-SSA}_t^{\textrm{GDP}}(h)$ is optimized. 
@@
   \hline
-Shift=0 & 0.975 & $0.941^{*}$ & $0.886^{*}$ &$ 0.839^{**} $& $0.843^{**}$ &$ 0.897^{**} $& 0.965 \\ 
-  Shift=1 & 1.085 & 1.068 & 1.011 & $0.927^{*} $& $0.871^{**}$ & $0.875^{**}$ & $0.919^{**}$ \\ 
-  Shift=2 & 1.076 & 1.089 & 1.074 & 1.011 & $0.938^{*}$ &$ 0.902^{**} $& $0.909^{**} $\\ 
-  Shift=3 & 1.037 & 1.069 & 1.090 & 1.059 & 0.992 & $0.943^* $&$ 0.930^*$ \\ 
-  Shift=4 & 1.013 & 1.058 & 1.101 & 1.097 & 1.037 & 0.967 &$ 0.925^*$ \\ 
-  Shift=5 & 1.016 & 1.062 & 1.106 & 1.115 & 1.078 & 1.015 & $0.955^*$ \\  
+Shift=0 & 0.973 & $0.937^*$ & $0.888^*$ & $0.850^{**}$ & $0.855^{**}$ & $0.900^{**}$ & $0.956^*$ \\ 
+  Shift=1 & 1.073 & 1.051 & 0.997 & $0.918^*$ & $0.870^{**}$ & $0.873^{**}$ & $0.906^{**}$ \\ 
+  Shift=2 & 1.069 & 1.077 & 1.063 & 1.006 & $0.940^*$ & $0.907^{**}$ & $0.907^{**}$ \\ 
+  Shift=3 & 1.020 & 1.045 & 1.066 & 1.048 & 0.997 & 0.957 & $0.941^*$ \\ 
+  Shift=4 & 0.983 & 1.016 & 1.056 & 1.064 & 1.023 & 0.974 & $0.942^*$ \\ 
+  Shift=5 & 0.988 & 1.026 & 1.071 & 1.088 & 1.057 & 1.007 & $0.963^*$ \\ 
    \hline
@@
 \caption{Relative RMSEs of M-SSA GDP Predictors vs. Direct Forecasts.\\ 
-Out-of-sample evaluation starting in Q1-2008, without pandemic. Values below 1 indicate improvement over direct forecasting.
+M-SSA is fixed, based on data up to Q4-2007. Out-of-sample evaluation starting in Q1-2008, without pandemic. 
+Values below 1 indicate improvement over direct forecasting.
 \label{tab:rRMSE_mSSA_comp_direct_without_covid7}}
@@
   \hline
-Shift=0 & 1.131 & 1.092 & 1.028 & 0.973 & 0.977 & 1.040 & 1.119 \\ 
-  Shift=1 & 1.295 & 1.274 & 1.206 & 1.106 & 1.039 & 1.044 & 1.097 \\ 
-  Shift=2 & 1.180 & 1.194 & 1.178 & 1.109 & 1.029 & 0.990 & 0.997 \\ 
-  Shift=3 & 1.029 & 1.061 & 1.082 & 1.051 & 0.984 & 0.936 & 0.923 \\ 
-  Shift=4 & 1.012 & 1.057 & 1.100 & 1.096 & 1.036 & 0.967 & 0.924 \\ 
-  Shift=5 & 0.988 & 1.033 & 1.076 & 1.084 & 1.048 & 0.987 & 0.929 \\  
+Shift=0 & 1.128 & 1.086 & 1.030 & 0.985 & 0.992 & 1.044 & 1.109 \\ 
+  Shift=1 & 1.280 & 1.254 & 1.189 & 1.096 & 1.037 & 1.041 & 1.081 \\ 
+  Shift=2 & 1.173 & 1.181 & 1.166 & 1.104 & 1.031 & 0.995 & 0.995 \\ 
+  Shift=3 & 1.012 & 1.037 & 1.058 & 1.040 & 0.989 & $0.949^*$ & $0.934^*$ \\ 
+  Shift=4 & 0.982 & 1.016 & 1.056 & 1.063 & 1.022 & 0.974 & $0.941^*$ \\ 
+  Shift=5 & 0.961 & 0.997 & 1.041 & 1.058 & 1.028 & 0.979 & $0.936^{*}$ \\ 
   \hline
@@
   \hline
-Forecast horizon 1 & 0.752 & 0.715 & 0.750 & 0.756 \\ 
-  Forecast horizon 2 & 0.697 & 0.663 & 0.714 & 0.744 \\ 
-  Forecast horizon 3 & 0.554 & 0.622 & 0.666 & 0.716 \\ 
-  Forecast horizon 4 & 0.486 & 0.624 & 0.707 & 0.732 \\ 
-  Forecast horizon 5 & 0.593 & 0.636 & 0.660 & 0.713 \\ 
+Forecast horizon 1 & 0.752 & 0.715 & 0.746 & 0.747 \\ 
+  Forecast horizon 2 & 0.697 & 0.663 & 0.729 & 0.720 \\ 
+  Forecast horizon 3 & 0.554 & 0.622 & 0.684 & 0.709 \\ 
+  Forecast horizon 4 & 0.486 & 0.624 & 0.680 & 0.708 \\ 
+  Forecast horizon 5 & 0.593 & 0.636 & 0.643 & 0.701 \\ 
    \hline
@@
 Additionally, the AUC is optimized through more precise control of the sign change rate within the M-SSA framework.
+Remarkably, these results hold even though the M-SSA filter was optimized on a relatively short in-sample window ending prior to the onset of the financial crisis.
 Finally, an examination of revision errors associated with quarterly updates of the proposed M-SSA-based GDP predictor indicates that the regression parameters, VAR model, and resulting M-SSA filters stabilize after an initial burn-in period.
@@
   \hline
-Shift=0 & 0.080 & 0.042 & 0.011 & 0.001 & 0.000 & 0.002 & 0.011 \\ 
-  Shift=1 & 0.403 & 0.297 & 0.188 & 0.095 & 0.042 & 0.039 & 0.103 \\ 
-  Shift=2 & 0.718 & 0.594 & 0.359 & 0.114 & 0.019 & 0.009 & 0.022 \\ 
-  Shift=3 & 0.955 & 0.950 & 0.871 & 0.569 & 0.158 & 0.031 & 0.013 \\ 
-  Shift=4 & 0.593 & 0.879 & 0.957 & 0.893 & 0.541 & 0.183 & 0.080 \\ 
-  Shift=5 & 0.531 & 0.875 & 0.971 & 0.929 & 0.605 & 0.239 & 0.041 \\ 
+Shift=0 & 0.127 & 0.064 & 0.019 & 0.003 & 0.001 & 0.002 & 0.012 \\ 
+  Shift=1 & 0.439 & 0.314 & 0.194 & 0.093 & 0.033 & 0.017 & 0.036 \\ 
+  Shift=2 & 0.718 & 0.565 & 0.326 & 0.098 & 0.020 & 0.008 & 0.010 \\ 
+  Shift=3 & 0.946 & 0.975 & 0.925 & 0.657 & 0.222 & 0.059 & 0.028 \\ 
+  Shift=4 & 0.327 & 0.699 & 0.964 & 0.942 & 0.619 & 0.237 & 0.102 \\ 
+  Shift=5 & 0.383 & 0.755 & 0.946 & 0.910 & 0.581 & 0.241 & 0.055 \\ 
    \hline
@@
   \hline
-Shift=0 & 0.999 & 0.980 & 0.951 & 0.925 & 0.915 & 0.926 & 0.955 \\ 
-  Shift=1 & 1.057 & 1.051 & 1.028 & 0.996 & 0.977 & 0.978 & 0.985 \\ 
-  Shift=2 & 1.031 & 1.035 & 1.026 & 0.996 & 0.964 & 0.952 & 0.964 \\ 
-  Shift=3 & 1.022 & 1.035 & 1.042 & 1.026 & 0.992 & 0.965 & 0.953 \\ 
-  Shift=4 & 1.002 & 1.023 & 1.044 & 1.046 & 1.024 & 0.999 & 0.984 \\ 
-  Shift=5 & 1.010 & 1.032 & 1.054 & 1.060 & 1.043 & 1.013 & 0.983 \\  
+Shift=0 & 1.005 & 0.986 & 0.960 & 0.935 & 0.927 & 0.939 & 0.964 \\ 
+  Shift=1 & 1.060 & 1.054 & 1.032 & 0.998 & 0.973 & 0.965 & 0.970 \\ 
+  Shift=2 & 1.028 & 1.029 & 1.020 & 0.993 & 0.963 & 0.950 & 0.952 \\ 
+  Shift=3 & 1.018 & 1.027 & 1.034 & 1.022 & 0.994 & 0.972 & 0.961 \\ 
+  Shift=4 & 0.990 & 1.006 & 1.025 & 1.032 & 1.017 & 0.998 & 0.985 \\ 
+  Shift=5 & 1.000 & 1.016 & 1.037 & 1.046 & 1.032 & 1.008 & 0.985 \\  
    \hline
@@
 \begin{table}[htpb]
-\caption{Relative RMSEs of Direct Forecasts vs. Mean Benchmark (Including Pandemic).
+\caption{Relative RMSEs of of M-SSA GDP Component Predictors  vs. Direct Forecasts (Including Pandemic).
 %Relative RMSEs of M-SSA GDP Component Predictors vs. Direct Forecasts (Including Pandemic).
@@
   \hline
-Shift=0 & 1.230 & 1.207 & 1.172 & 1.139 & 1.127 & 1.140 & 1.177 \\ 
-  Shift=1 & 0.984 & 0.977 & 0.956 & 0.927 & 0.909 & 0.910 & 0.916 \\ 
-  Shift=2 & 1.039 & 1.044 & 1.035 & 1.004 & 0.972 & 0.960 & 0.972 \\ 
-  Shift=3 & 1.046 & 1.059 & 1.067 & 1.050 & 1.015 & 0.987 & 0.975 \\ 
-  Shift=4 & 0.982 & 1.002 & 1.023 & 1.024 & 1.003 & 0.978 & 0.964 \\ 
-  Shift=5 & 0.979 & 1.001 & 1.022 & 1.028 & 1.012 & 0.983 & 0.953 \\   
+Shift=0 & 1.237 & 1.215 & 1.183 & 1.152 & 1.141 & 1.156 & 1.187 \\ 
+  Shift=1 & 0.986 & 0.980 & 0.960 & 0.929 & 0.905 & 0.898 & 0.903 \\ 
+  Shift=2 & 1.036 & 1.037 & 1.029 & 1.001 & 0.971 & 0.957 & 0.960 \\ 
+  Shift=3 & 1.042 & 1.051 & 1.058 & 1.045 & 1.018 & 0.995 & 0.983 \\ 
+  Shift=4 & 0.970 & 0.985 & 1.004 & 1.010 & 0.996 & 0.978 & 0.965 \\ 
+  Shift=5 & 0.970 & 0.986 & 1.007 & 1.015 & 1.002 & 0.978 & 0.956 \\   
    \hline

### 2025-08-16 Sweave/BIP_predictor.Rnw
up

@@
 x_mat<-data[,select_vec_multi]
+head(x_mat)
 tail(x_mat)
@@
 colnames(data_table)[1]<-"GDP"
+head(data_table,20)
 tail(data_table,20)
